### 2015-08-11 doc/IQS.tex
adding IQS note

%% Based on a TeXnicCenter-Template, which was
%% created by Christoph Börensen
%% and slightly modified by Tino Weinkauf.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{scrartcl} %This is a special class provided by the KOMA script, which does a lot of adjustments to adapt the standard LaTeX classes to european habits, change to [a4paper,12pt,twoside] for doublesided layout

%########################### Preferences #################################

% ******** vmargin settings *********
\usepackage{vmargin} %This give you full control over the used page area, it maybe not the idea od Latex to do so, but I wanted to reduce to amount of white space on the page
%\setpapersize{A4}
%\setmargins{3.5cm}%			%linker Rand, left edge
					 %{1.5cm}%     %oberer Rand, top edge
           %{14.7cm}%		%Textbreite, text width
           %{23.42cm}%   %Texthoehe, text hight
           %{14pt}%			%Kopfzeilenhöhe, header hight
           %{1cm}%   	  %Kopfzeilenabstand, header distance
           %{0pt}%				%Fußzeilenhoehe footer hight
           %{2cm}%    	  %Fusszeilenabstand, footer distance

% ********* Font definiton ************
%\usepackage{t1enc} % as usual
%\usepackage[latin1]{inputenc} % as usual
%\usepackage{times}		
%\usepackage{mathptmx}  	%mathematical fonts for use with times, I encountered some problems using this package togather with pdftex, which I was not able to resolve

% ********* Graphics definition *******
%\usepackage[pdftex]{graphicx} % required to import graphic files
\usepackage{color} %allows to mark some entries in the tables with color
%\usepackage{eso-pic} % these two are required to add the little picture on top of every page
%\usepackage{everyshi} % these two are required to add the little picture on top of every page
%\renewcommand{\floatpagefraction}{0.7} %default:0.5 allows two big pictures on one page

\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amstext}
\usepackage{amsbsy}
\usepackage{mathbbol} 
%\usepackage{pmatrix}

%********** Enybeling Hyperlinks *******
%\usepackage[pdfborder=000,pdftex=true]{hyperref}% this enables jumping from a reference and table of content in the pdf file to its target

% ********* Table layout **************
\usepackage{booktabs}	  	%design of table, has an excellent documentation
%\usepackage{lscape}			%use this if you want to rotate the table together with the lines around the table

%  new definitions
\renewcommand{\div}{\bs{\nabla}\! \cdot \!}
\newcommand{\grad}{\bs{\nabla}}
% extra space
\newcommand{\qq}{\quad\quad}
% common reference commands
\newcommand{\eqt}[1]{Eq.~(\ref{#1})}                     % equation
\newcommand{\fig}[1]{Fig.~\ref{#1}}                      % figure
\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table
\newcommand{\sct}[1]{Section~\ref{#1}}                   % section
\newcommand{\app}[1]{Appendix~\ref{#1}}                   % appendix

\newcommand{\bs}[1]{\mathbf{#1}}
\newcommand{\dd}{\mathrm{d}}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\vn}{\vec{n}}
\newcommand{\vel}{\vec{\mathrm{v}}}
\newcommand{\adj}{\Phi^\dagger_0}

% ********* Caption Layout ************
\usepackage{ccaption} % allows special formating of the captions
\captionnamefont{\bf\footnotesize\sffamily} % defines the font of the caption name (e.g. Figure: or Table:)
\captiontitlefont{\footnotesize\sffamily} % defines the font of the caption text (same as above, but not bold)
\setlength{\abovecaptionskip}{0mm} %lowers the distace of captions to the figure

% ********* Header and Footer **********
% This is something to play with forever. I use here the advanced settings of the KOMA script

\usepackage{scrpage2} %header and footer using the options for the KOMA script
\renewcommand{\headfont}{\footnotesize\sffamily} % font for the header
\renewcommand{\pnumfont}{\footnotesize\sffamily} % font for the pagenumbers

%the following lines define the pagestyle for the main document
\defpagestyle{cb}{%
(\textwidth,0pt)% sets the border line above the header
{\pagemark\hfill\headmark\hfill}% doublesided, left page
{\hfill\headmark\hfill\pagemark}% doublesided, right page
{\hfill\headmark\hfill\pagemark}%  onesided
(\textwidth,1pt)}% sets the border line below the header
%
{(\textwidth,1pt)% sets the border line above the footer
{{\it Jean Ragusa}\hfill TAMU}% doublesided, left page
{Jean Ragusa\hfill{\it TAMU}}% doublesided, right page
{Jean Ragusa\hfill{\it TAMU}} % one sided printing
(\textwidth,0pt)% sets the border line below the footer
}

%this defines the page style for the first pages: all empty
\renewpagestyle{plain}%
	{(\textwidth,0pt)%
		{\hfill}{\hfill}{\hfill}%
	(\textwidth,0pt)}%
	{(\textwidth,0pt)%	
		{\hfill}{\hfill}{\hfill}%
	(\textwidth,0pt)}

%********** Footnotes **********
\renewcommand{\footnoterule}{\rule{5cm}{0.2mm} \vspace{0.3cm}} %increases the distance of footnotes from the text
\deffootnote[1em]{1em}{1em}{\textsuperscript{\normalfont\thefootnotemark}} %some moe formattion on footnotes

\renewcommand{\labelitemii}{$\diamond$}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% boxes
\usepackage{color}
\definecolor{myblue}{rgb}{.8, .8, 1}
\usepackage{empheq}

\newlength\mytemplen
\newsavebox\mytempbox

\makeatletter
\newcommand\mybluebox{%
    \@ifnextchar[%]
       {\@mybluebox}%
       {\@mybluebox[0pt]}}

\def\@mybluebox[#1]{%
    \@ifnextchar[%]
       {\@@mybluebox[#1]}%
       {\@@mybluebox[#1][0pt]}}

\def\@@mybluebox[#1][#2]#3{
    \sbox\mytempbox{#3}%
    \mytemplen\ht\mytempbox
    \advance\mytemplen #1\relax
    \ht\mytempbox\mytemplen
    \mytemplen\dp\mytempbox
    \advance\mytemplen #2\relax
    \dp\mytempbox\mytemplen
    \colorbox{myblue}{\hspace{1em}\usebox{\mytempbox}\hspace{1em}}}

\makeatother

%################ End Preferences, Begin Document #####################

\pagestyle{plain} % on headers or footers on the first page

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}

\begin{figure}[th]
    \centering
		%\includegraphics[width=10cm]{logo.jpg}
	\label{fig:logo}
\end{figure}

\vspace{2cm}
\vspace{2cm}

% There might be better solutions for the title page, giving all distances and sizes manually was simply the easiest solution

{\Huge\bf\sf IQS Notes}

\vspace{.5cm}

{\Huge\bf\sf TAMU}

\vspace{.5cm}

%{\Huge\bf\sf to use with TeXnicCenter}

\vspace{2cm}

{\Large\bf\sf Jean C. Ragusa}
\vspace{2cm}

{\Large\bf\sf \today} %adds the current date

\vspace{\fill}

\tt{jean.ragusa@tamu.edu}

\end{center}
\newpage

%%The following loads the picture on top of every page, the numbers in \put() define the position on the page:
%\AddToShipoutPicture{\setlength\unitlength{0.1mm}\put(604,2522){\includegraphics[width=1.5cm]{logo.jpg}}}

\pagestyle{cb} % now we want to have headers and footers

\tableofcontents

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Time-dependent diffusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\be
\frac{1}{v} \frac{\partial \Phi }{\partial t} = \nu_p \sigma_f \Phi - \left( -\div D \grad \Phi + \sigma_a \Phi \right) + \lambda C
\ee
\be
\frac{dC}{dt} = \nu_d \sigma_f \Phi - \lambda C
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{IQS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Factorization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Factorization into amplitude and shape functions:
\be
\Phi(\vec{r},t) = p(t) \varphi(\vec{r},t)
\ee

Factorization is not unique:
\[
\Phi(\vec{r},t) = \frac{p(t)}{f(t)} \left( f(t) \varphi(\vec{r},t) \right)
\]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Equations for the shape $\varphi$}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\be
\frac{1}{v} \frac{\partial \varphi }{\partial t} = \nu_p \sigma_f \varphi - \left( -\div D \grad \varphi + (\sigma_a+\alpha(t)) \varphi \right) + \frac{1}{p}\lambda C
\ee
where
\[
\alpha(t) = \frac{1}{v} \frac{1}{p(t)}\frac{dp}{dt}
\]
Very little change in the precursor equation:
\be
\frac{dC}{dt} = \nu_d \sigma_f p \varphi - \lambda C
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Equations for the amplitude $p$}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
These are simply obtained by integration of the shape equations over the entire space. We use the following short-cut notation:
\[
\langle f, g \rangle = \int_{\text{domain}} f(\vec{r})g(\vec{r}) d^3r
\]
The starting point is the shape equation (re-written below before division by $p$ occurred).
\be
\frac{1}{v} p \frac{\partial \varphi }{\partial t} + \frac{1}{v} \varphi \frac{dp}{dt} = \nu_p \sigma_f p \varphi - p \left( -\div D \grad \varphi + \sigma_a \varphi \right) + \lambda C
\ee
Before carrying out the integration over space, we multiply the equation by a time-independent weighting function (a common choice for this function is the initial adjoint flux, $\adj$).

After a bit of algebra, we obtain
\be
\frac{dp}{dt} = \left(\frac{\rho}{\Lambda}-\frac{\beta}{\Lambda} - \theta \right) p + \lambda \xi
\ee
\be
\frac{d\xi}{dt} = \frac{\beta}{\Lambda} p - \left( \lambda + \theta \right) \xi
\ee 
with the following definitions
\begin{align}
\frac{\rho}{\Lambda} &= \frac{\langle \adj,  \left[ \nu \sigma_f - \left( -\div D \grad  + \sigma_a  \right)\right] \varphi \rangle}{\langle \adj, v^{-1} \,\varphi \rangle}\\
\frac{\beta}{\Lambda}&= \frac{\langle \adj,  \nu_d \sigma_f \varphi \rangle}{\langle \adj, v^{-1} \,\varphi \rangle}\\
\theta               &= \frac{\langle \adj,  v^{-1} \frac{\partial \varphi}{\partial t} \rangle}{\langle \adj, v^{-1} \,\varphi \rangle}
                      = \frac{\frac{d}{dt}\langle \adj,  v^{-1} \varphi \rangle}{\langle \adj, v^{-1} \,\varphi \rangle}
\end{align}
The total time derivative expression for $\theta$ is due to the fact that $v$ and the initial adjoint function are both independent of time.

We have also defined
\[
\xi = \frac{ \langle \adj, C \rangle}{\langle \adj, v^{-1} \,\varphi \rangle}
\]
which gives
\[
\langle \adj, \frac{dC}{dt} \rangle = \langle \adj, v^{-1} \,\varphi \rangle \frac{d \xi}{dt} + \xi \frac{d}{dt}\langle \adj, v^{-1} \,\varphi \rangle
\]
and explains the presence of the $\theta$ term in the PRKE equation for the precursor concentrations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Normalization choice}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We choose to have the following quantity constant in time
\be
\langle \adj, v^{-1} \,\varphi \rangle = K_0
\ee
even though the shape itself may still be allowed to change in time.

This makes the factorization {\bf unique} and lets $\theta=0$, yielding the standard form of the PRKE equations (note that
in the PRKE approximation, i.e., when the shape is assume independent of time, we automatically have $\theta=0$).

\be
\frac{dp}{dt} = \left(\frac{\rho}{\Lambda}-\frac{\beta}{\Lambda} \right) p + \lambda \xi
\ee
\be
\frac{d\xi}{dt} = \frac{\beta}{\Lambda} p - \lambda \xi
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IQS solution procedure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the IQS method, the shape is solved for on macro time steps $[t^n,t^{n+1}]$ (one solve for the shape over that time step) while the 
PRKE equations are solve on micro time steps: $t^n, t^{n+1/M}, \ldots, t^{n+m/M}, \ldots, t^{n+1}$ (for a total of $M$ solves 
during one macro time step).

Because of the factorization approximation, we are actually solving a nonlinear problem over the macro time step. Convergence is said to be reached 
when the new shape $\varphi(\vec{r},t^{n+1})$ satisfies the normalize condition $\langle \adj, v^{-1} \,\varphi \rangle = K_0$.

The process over $[t^n,t^{n+1}]$ is as follows. Let the nonlinear iteration index be $\ell$.
\begin{enumerate}
\item Set $\ell=0$
\item Estimate the shape at the end time $\varphi^{n+1,\ell}$ (could be extrapolation, or just assume equal to begin time shape function)
\item Using the begin time and end time shape to compute $\rho/\Lambda$ and $\beta/\Lambda$ at the micro time steps 
(this is a bit costly, so in reality we only do this for a certain multiple of the micro time steps, assuming a linear variation in time of the shape and using the exact values of the XS at the evaluation instants). 
\item Solve the PRKE equations using data from 3.
\item Compute $\alpha(t^{n+1}) = \frac{1}{v} \frac{1}{p(t^{n+1})}\left.\frac{dp}{dt}\right|_{t^{n+1}}$
\item Compute the shape at the end time $\varphi^{n+1,\ell+1}$
\item Check for convergence in $K_0$. If so, EXIT
\item Otherwise $\ell \leftarrow \ell+1$ and GO BACK to 3.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Homogeneous perturbation in an homogenous domain}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this Section I consider a ``simple'' test case which I was hoping to test the 
implementation of IQS.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Problem description}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Consider an initially critical homogeneous slab reactor solved with 1-g diffusion theory.
The initial flux/shape is therefore a cosine. We can make the reactor critical simply by dividing
the $\nu$'s with k$_\text{eff}$.

At time $t=0$, the $\sigma_a$ is subjected to a variation that is linear in time (and this, at all times,
the domain remains homogenous, i.e., no spatially varying cross sections).

Because the initial flux only contains the steady-state fundamental mode and because the modification is homogeneous,
no other modes are excited and the flux function only changes in magnitude. That is, the shape function remains constant
and the flux amplitude is actually given by the solution of the PRKE.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Brute force temporal discretization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A backward Euler time discretization yields, for the flux,

\be
\label{eq:ex1}
\frac{1}{v} \frac{\Phi^{n+1}-\Phi^n }{\Delta t} = (\nu_p \sigma_f)^{n+1} \Phi^{n+1} - \left( -\div D \grad + \sigma_a \right)^{n+1} \Phi^{n+1} + \lambda C^{n+1}
\ee
For a constant shape, this becomes:
\be
\frac{\varphi_0}{v} \frac{p^{n+1}-p^n }{\Delta t} = (\nu_p \sigma_f)^{n+1} \varphi_0 p^{n+1} - \left( -\div D \grad + \sigma_a \right)^{n+1} \varphi_0 p^{n+1} + \lambda \varphi_0 \xi^{n+1}
\ee
What equations should be satisfied by $p^{n+1}$ and $\xi^{n+1}$? Simply the PRKE discretized with backward Euler over $[t^n,t^{n+1}]$ (one step!). This
can be easily verified if you take the above equation, multiply it by $\adj$ and integrate over the entire domain.

We numerically observed that the brute-force solve yielded the same results as the PRKE solve when using the initial shape in the PRKE parameters.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IQS approach with NO subcycling for the PRKE}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
If the shape and the amplitude functions are solved using the same time step, then we have

\be
\label{eq:ex2}
\frac{1}{v} \frac{\varphi^{n+1} -\varphi^n}{\Delta t} = (\nu_p \sigma_f)^{n+1} \varphi^{n+1} - \left( -\div D \grad + (\sigma_a+\alpha) \right)^{n+1} \varphi^{n+1} + \frac{1}{p^{n+1}}\lambda C^{n+1}
\ee
where 
\[
\alpha^{n+1} = \frac{1}{v} \frac{1}{p^{n+1}}\frac{p^{n+1}-p^{n}}{\Delta t}
\]
This form of $\alpha$ is due to the fact that we assumed here to employ only one time interval for the PRKE (no subcycling).
Note that by letting $\varphi^{n+1}=\varphi^n=\varphi_0$,  we can re-arrange \eqt{eq:ex2} as \eqt{eq:ex1}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IQS approach with subcycling for the PRKE}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Now, \eqt{eq:ex2} tells you also something quite important. If you solve \eqt{eq:ex2} for $\varphi^{n+1}$, you obtain a solution (we verified if 
the PRKE are done with the same time step as the shape, that the shape solution remains constant). Now, 
\eqt{eq:ex2} is a linear system ($A \varphi^{n+1} = b$), so if I change slightly the entry of matrix $A$, I should get another solution for 
$\varphi^{n+1}$. Well, subcycling the PRKE gives us a better value for $\alpha$ (it is more accurate since we solve the PRKE on a finer time grid),
so we automatically see that $\varphi^{n+1}$ cannot remain constant equal to its initial value. Where's the flaw ????

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-08-11 matlab/kinetics_driver.m
playing with theta

@@ -6,7 +6,7 @@ close all; clc;
 global dat npar
 
 % verbose/output parameters
-testing = false;
+testing = true;
 console_print = true;
 plot_transient_figure = true;
 plot_power_figure = true;
@@ -16,7 +16,7 @@ make_movie = false;
 npar.set_bc_last=true;
 
 % select problem
-pbID=10; refinements=3;
+pbID=2; refinements=8;
 problem_init(pbID,refinements);
 
 % compute fundamental eigenmode
@@ -39,11 +39,12 @@ u0=u;
 
 phi_adjoint = npar.phi_adj;
 IV   = assemble_mass(     dat.inv_vel ,curr_time);
+npar.IV=IV;
 
 % time steping data
-dt=0.001;
-ntimes=1000; % 150*2;
-iqs_factor=100;
+dt=0.01;
+ntimes=100; % 150*2;
+iqs_factor=10;
 
 
 if make_movie
@@ -118,7 +119,7 @@ Pnorm_prkeIQS(1)=X(1);
 dt=dt*iqs_factor; ntimes=ntimes/iqs_factor;
 
 n_micro=10;
-freq_react=2;
+freq_react=1;
 
 %%% loop on time steps %%%
 for it=1:ntimes
@@ -155,6 +156,9 @@ dt=dt/iqs_factor; ntimes=ntimes*iqs_factor;
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%% standard PRKE with initial shape
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% if testing
+%     return
+% end
 
 shape0=phi0;
 [~,beff_MGT]=compute_prke_parameters(0.,shape0);

### 2015-08-11 matlab/solve_IQS_diffusion.m

@@ -2,8 +2,9 @@ function [u_shape, X] = solve_IQS_diffusion(u_shape,X,dt_macro,time_end,n_micro,
 
 global dat npar
 
-max_iter_iqs = 2;
-tol_iqs=1e-7;
+max_iter_iqs = 3;
+tol_iqs=1e-6;
+npar.theta_old=[];
 
 IV = assemble_mass(dat.inv_vel,time_end);
 
@@ -41,8 +42,8 @@ for iter = 1: max_iter_iqs
     if err<tol_iqs
         break
     else
-        u_shape = u_shape / ((npar.phi_adj)'*IV*shape_end/npar.K0);
-        shape_end=u_shape(1:npar.n);
+%         u_shape = u_shape / ((npar.phi_adj)'*IV*shape_end/npar.K0);
+%         shape_end=u_shape(1:npar.n);
     end
 end
 

### 2015-08-11 matlab/solve_prke_iqs.m

@@ -1,7 +1,7 @@
 function [X,dpdt] =  solve_prke_iqs(X,dt_macro,time_end,shape_beg,shape_end,n_micro,freq_react)
 
 % make the problem-data a global variable
-global dat 
+global dat npar
 
 % time step for prke solve
 dt = dt_macro/n_micro;
@@ -15,6 +15,20 @@ end
 % time step
 time_beg = time_end - dt_macro;
 times_react_update = time_beg + linspace(0,n_react,n_react+1) * dt_macro/n_react;
+
+theta_beg = npar.phi_adj' * npar.IV(1:npar.n,1:npar.n) * shape_beg;
+theta_end = npar.phi_adj' * npar.IV(1:npar.n,1:npar.n) * shape_end;
+theta = -(theta_end-theta_beg)/dt_macro/npar.K0;
+
+if ~isempty(npar.theta_old)
+    % damping
+    a=1;
+    theta = a*theta+(1-a)*npar.theta_old;
+    theta=theta;
+else
+    npar.theta_old=theta;
+end
+
 
 % compute prke parameters at the above times 
 for i=1:n_react+1 % same as length(times_react_update)
@@ -41,11 +55,13 @@ end
 % loop over micro time steps
 for it=1:n_micro
     % build prke matrix
-    A=[(rho_MGT_iqs(it)-beff_MGT_iqs(it)) dat.lambda ; ...
-        beff_MGT_iqs(it)                 -dat.lambda];
+    A=[(rho_MGT_iqs(it)-beff_MGT_iqs(it)-theta) dat.lambda ; ...
+        beff_MGT_iqs(it)                 -(dat.lambda+theta)];
     % solve
     X=(eye(2)-dt*A)\X;
 end
+
+fprintf('theta = %g, power= %g \n',theta,X(1));
 
 % save value for dpdt at the end of the macro time step
 dXdt = A*X;

### 2015-10-06 matlab/kinetics_driver.m
adding back the possibility of using homemade prke solver

@@ -69,7 +69,7 @@ i=0;
 % i=i+1; list_runs{i}= 'brute_force';
 % i=i+1; list_runs{i}= 'brute_force_an_prec';
 i=i+1; list_runs{i}= 'iqs_an_prec';
-% i=i+1; list_runs{i}= 'iqs';
+i=i+1; list_runs{i}= 'iqs';
 % i=i+1; list_runs{i}= 'prke_initial_shape';
 % i=i+1; list_runs{i}= 'prke_exact_shape';
 % i=i+1; list_runs{i}= 'prke_qs_shape';

### 2015-10-06 matlab/problem_init.m

@@ -197,124 +197,13 @@ npar.add_ones_on_diagonal=~npar.add_zero_on_diagonal;
 
 % IQS options
 npar.solve_prke_compute_rho_each_time = false;
+npar.prke_solve = 'matlab' ;
 
 return
 end
 
 
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-function [u]=solve_ss_fem()
-
-% perform the Newton solve
-u = newton();
-
-
-return
-end
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-function  T=newton()
-
-global dat npar prt
-
-% initial guess
-T=ones(sum(npar.ndofs),1)*400;
-T=linspace(1000,400,sum(npar.ndofs))';
-T=T/npar.scale;
-
-curr_time=0;
-% compute the residual
-resi = dat.myfunc(curr_time,T);
-% compute its norm
-norm_resi=norm(resi);
-% compute stopping criterion
-tol_newton = npar.atol_newton + npar.rtol_newton*norm_resi;
-
-eigstudy=false;
-if(npar.newton_solve_option>3),eigstudy=false;end
-
-inewton=0;
-while (inewton<npar.max_newton && norm_resi>tol_newton)
-    inewton=inewton+1;
-    if(prt.lev>=0), fprintf('Newton iteration %i, ',inewton); end
-    % compute the jacobian matrix
-    J = dat.myjac(curr_time,T,resi);
-    P = dat.myprec(curr_time,T);
-    if(eigstudy)
-%         P = compute_precond(T,npar,dat,1);
-        figure(2)
-        subplot(2,1,1);
-        eigplot(full(J),2);
-        subplot(2,1,2);
-        JiP=J*inv(P);
-        eigplot(full(JiP),2);
-        fprintf('\n\tEst. of cond number for J: %g, \t for JiP: %g\n',condest(J),condest(JiP));
-        figure(3);
-        subplot(2,1,1);   spy(J);
-        subplot(2,1,2);   spy(JiP);
-    end
-    % newton solve
-    switch npar.newton_solve_option
-        case {1}
-            dT = -J\resi;
-        case{2,3}
-            % gmres(A,b,restart,tol,maxit,M1,M2,x0)
-            restart=[];
-            tol_newton_lin=npar.tol_newton_lin;
[…]

### 2015-10-06 matlab/solve_IQS_diffusion.m

@@ -18,8 +18,11 @@ shape_end=shape_beg;
 for iter = 1: max_iter_iqs
     
     % solve for amplitude function
-%     [X,dpdt] =  solve_prke_iqs(X_beg,dt_macro,time_end,shape_beg,shape_end,n_micro,freq_react);
-    [X,dpdt] =  solve_prke_ode(X_beg,dt_macro,time_end,shape_beg,shape_end);
+    if strcmpi(npar.prke_solve,'matlab')
+        [X,dpdt,~,~] =  solve_prke_ode(X_beg,dt_macro,time_end,shape_beg,shape_end);
+    else
+        [X,dpdt,~,~] =  solve_prke_iqs(X_beg,dt_macro,time_end,shape_beg,shape_end,n_micro,freq_react);
+    end
     % assemble IQS    
     TR = assemble_transient_operator_iqs(time_end,X(1),dpdt);
     M  = assemble_time_dependent_operator(time_end);

### 2015-10-06 matlab/solve_IQS_diffusion_an_prec.m

@@ -21,18 +21,24 @@ shape_end=shape_beg;
 for iter = 1: max_iter_iqs
     
     % solve for amplitude function
-    %     [X,dpdt] =  solve_prke_iqs(X_beg,dt_macro,time_end,shape_beg,shape_end,n_micro,freq_react);
-    [X,dpdt,t,y] =  solve_prke_ode(X_beg,dt_macro,time_end,shape_beg,shape_end);
-    
+    if strcmpi(npar.prke_solve,'matlab')
+        [X,dpdt,t,y] =  solve_prke_ode(X_beg,dt_macro,time_end,shape_beg,shape_end);
+    else
+        [X,dpdt,t,y] =  solve_prke_iqs(X_beg,dt_macro,time_end,shape_beg,shape_end,n_micro,freq_react);
+    end
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     % assemble IQS
     
     % shortcut
     p=X(1);
     % interpolating polynomial
-    pp = interp1(t,y,'pchip','pp');
+    if strcmpi(npar.prke_solve,'matlab')
+        pp = interp1(t,y,'pchip','pp');
+    else
+        pp = interp1(t,y,'linear','pp');
+    end
     f = @(t) ppval(pp,t);
-
+    
     TR  = sparse(npar.n,npar.n,npar.nnz);
     
     D    = assemble_stiffness(dat.cdiff   ,time_end);
@@ -69,7 +75,7 @@ for iter = 1: max_iter_iqs
     % build system matrix
     A = IV-dt*TR;
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-         
+    
     if npar.set_bc_last
         [A,rhs]=apply_BC(A,rhs,npar.add_ones_on_diagonal);
     else
@@ -77,7 +83,7 @@ for iter = 1: max_iter_iqs
     end
     % solve for new shape_end: M(unew-uold)/dt=TR.unew
     shape_end = A\rhs;
-
+    
     % update precursors
     C_new =  C_old*exp(-lambda*dt) + ( a1*NFId_old + a2*NFId_new )*shape_beg + ( a2*NFId_old + a3*NFId_new )*shape_end ;
     % re-package as single solution vector

### 2015-10-06 matlab/solve_prke_iqs.m

@@ -1,4 +1,4 @@
-function [X,dpdt] =  solve_prke_iqs(X,dt_macro,time_end,shape_beg,shape_end,n_micro,freq_react)
+function [X,dpdt,t,y] =  solve_prke_iqs(X,dt_macro,time_end,shape_beg,shape_end,n_micro,freq_react)
 
 % make the problem-data a global variable
 global dat npar
@@ -31,7 +31,7 @@ end
 
 theta = 0;
 
-% compute prke parameters at the above times 
+% compute prke parameters at the above times
 for i=1:n_react+1 % same as length(times_react_update)
     % weights for linear interpolation of the shape
     w1 = (time_end-times_react_update(i))/dt_macro;
@@ -53,6 +53,10 @@ for i=1:n_react
     beff_MGT_iqs = [ beff_MGT_iqs aux(2:end)];
 end
 
+% storage for t and y (for post-processing)
+t=zeros(nmicro+1,1); t(1) = time_beg;
+y=zeros(nmicro+1,1); y(1) = X(1);
+
 % loop over micro time steps
 for it=1:n_micro
     % build prke matrix
@@ -60,6 +64,9 @@ for it=1:n_micro
         beff_MGT_iqs(it)                 -(dat.lambda+theta)];
     % solve
     X=(eye(2)-dt*A)\X;
+    % storage
+    t(it+1)=t(it)+dt;
+    y(it+1)=X(1);
 end
 
 fprintf('theta = %g, power= %g \n',theta,X(1));

### 2015-10-07 matlab/kinetics_driver.m
Including theta discretization option for precursors

@@ -43,7 +43,7 @@ npar.IV=IV;
 
 % time steping data
 dt=0.005;
-ntimes=10/dt; % 150*2;
+ntimes=2/dt; % 150*2;
 iqs_factor=1;
 
 
@@ -69,6 +69,7 @@ i=0;
 % i=i+1; list_runs{i}= 'brute_force';
 % i=i+1; list_runs{i}= 'brute_force_an_prec';
 i=i+1; list_runs{i}= 'iqs_an_prec';
+i=i+1; list_runs{i}= 'iqs_theta_prec';
 i=i+1; list_runs{i}= 'iqs';
 % i=i+1; list_runs{i}= 'prke_initial_shape';
 % i=i+1; list_runs{i}= 'prke_exact_shape';
@@ -196,7 +197,7 @@ for it=1:ntimes
     
     time_end=it*dt;
     if console_print, fprintf('time end = %g \n',time_end); end
-    
+
     % solve time-dependent diffusion for flux
     [u_shape,X,t,y] = solve_IQS_diffusion_an_prec(u_shape,X,dt,time_end);
     
@@ -216,6 +217,59 @@ for it=1:ntimes
     time_prke_iqs=[time_prke_iqs ; t];
     power_prke_iqs=[power_prke_iqs; y];
 
+    
+end
+
+% make movie
+if plot_transient_figure && make_movie
+    close(gcf)
+    % save as AVI file
+    movie2avi(mov, 'PbID10_v2_iqs.avi', 'compression','None', 'fps',1);
+end
+% undo time step increase
+dt=dt/iqs_factor; ntimes=ntimes*iqs_factor;
+
+end
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% IQS IQS IQS with Theta Discretized precursors
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+if should_I_run_this(list_runs,'iqs_theta_prec')
+    
+amplitude_norm_iqs3=1;
+
+dt=dt*iqs_factor; ntimes=ntimes/iqs_factor;
+
+% initial solution vector
+u_shape=[phi0;C0];
+[~,beff_MGT]=compute_prke_parameters(0.,phi0);
+X=[1;beff_MGT/dat.lambda];
+Pnorm_prkeIQS3(1)=X(1);
+
[…]

### 2015-10-07 matlab/solve_IQS_diffusion_td_prec.m

@@ -1,0 +1,90 @@
+function [u_shape, X, err] = solve_IQS_diffusion_td_prec(u_shape,X,dt_macro,time_end,theta)
+
+global dat npar
+
+% shortcuts
+lambda = dat.lambda;
+dt = dt_macro;
+C_old = u_shape(npar.n+1:end);
+
+max_iter_iqs = 1;
+tol_iqs=1e-6;
+npar.theta_old=[];
+
+% save values at beginning of macro time step: they are needed in the IQS iteration
+X_beg=X;
+u_shape_beg=u_shape;
+% for clarity, I also single out the shape function at the beginning/end of the macro time step
+shape_beg=u_shape(1:npar.n);
+shape_end=shape_beg;
+
+for iter = 1: max_iter_iqs
+    
+    % solve for amplitude function
+    %     [X,dpdt] =  solve_prke_iqs(X_beg,dt_macro,time_end,shape_beg,shape_end,n_micro,freq_react);
+    [X,dpdt,t,y] =  solve_prke_ode(X_beg,dt_macro,time_end,shape_beg,shape_end);
+    
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    % assemble IQS
+    
+    % shortcut
+    p_old=X_beg(1);
+    p=X(1);
+
+    TR  = sparse(npar.n,npar.n,npar.nnz);
+    
+    D    = assemble_stiffness(dat.cdiff   ,time_end);
+    A    = assemble_mass(     dat.siga    ,time_end);
+    NFIp = assemble_mass(     dat.nusigf_p,time_end) / npar.keff;
+    IV   = assemble_mass(     dat.inv_vel ,time_end);
+    Aiqs = assemble_mass(     dat.inv_vel ,time_end) * dpdt/p;
+    
+    NFId_old = assemble_mass(dat.nusigf_d,time_end-dt) / npar.keff ;
+    NFId_new = assemble_mass(dat.nusigf_d,time_end)    / npar.keff ;
+    
+    % flux-flux matrix
+    TR=NFIp-(D+A+Aiqs);
+    
+    a1= (1-(1-theta)*dt*lambda)/(1+theta*dt*lambda);
+    a2= (1-theta)*dt/(1+theta*dt*lambda)*p_old;
+    a3= theta*dt/(1+theta*dt*lambda)*p;
+    
+    
+    % build transient matrix (divide lambda by p)
+    TR = TR + lambda/p * a3*NFId_new;
+    
[…]

### 2015-10-08 doc/note-to-yaqi/precusor-integration.tex
Precursor integration note for Yaqi

\documentclass[11pt]{tamurmemo} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{graphics}
\usepackage{float}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsbsy}
\usepackage{xspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% operators
\renewcommand{\div}{\vec{\nabla}\! \cdot \!}
\newcommand{\grad}{\vec{\nabla}}
% latex shortcuts
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bal}{\begin{align}}
\newcommand{\eali}{\end{align}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\ben}{\begin{enumerate}}
\newcommand{\een}{\end{enumerate}}
% DGFEM commands
\newcommand{\jmp}[1]{[\![#1]\!]}                     % jump
\newcommand{\mvl}[1]{\{\!\!\{#1\}\!\!\}}             % mean value
\newcommand{\kef}{\ensuremath{k_{\textit{eff}}}}
%\newcommand{\keff}{{\text{k}$_\textit{eff}$}\xspace}
\newcommand{\keff}{\kef\xspace}
% shortcut for domain notation
%\newcommand{\D}{\mathcal{D}}
% vector shortcuts
\newcommand{\vo}{\vec{\Omega}}
\newcommand{\vr}{\vec{r}}
\newcommand{\vn}{\vec{n}}
\newcommand{\vnk}{\vec{\mathbf{n}}}
\newcommand{\vj}{\vec{J}}
% extra space
\newcommand{\qq}{\quad\quad}
% common reference commands
\newcommand{\eqt}[1]{Eq.~(\ref{#1})}                     % equation
\newcommand{\fig}[1]{Fig.~\ref{#1}}                      % figure
\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table

\newcommand{\ud}{\,\mathrm{d}}
\newcommand{\mt}[1]{\marginpar{\tiny #1}}
\newcommand{\D}{\ensuremath{\mathcal{D}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%   BEGIN DOCUMENT
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%---------------------------------------------------------------------------%%
%% OPTIONS FOR NOTE
%%---------------------------------------------------------------------------%%

\toms{Distribution}
\subject{Precursor Evaluation}

%-------NO CHANGES
\collegename{Dwight Look College of Engineering}
\deptname{Department of Nuclear Engineering}
\fromms{Zachary\ M.\ Prince}
\originator{zmp}
\typist{zmp}
\date{\today}
%-------NO CHANGES

%-------OPTIONS
%\reference{NPB Star Reimbursable Project}
%\thru{}
%\enc{list}      
%\attachments{list}
\cy{File}
%\encas
%\attachmentas
%\attachmentsas 
%-------OPTIONS

%%---------------------------------------------------------------------------%%
%% DISTRIBUTION LIST
%%---------------------------------------------------------------------------%%

\distribution {
Yaqi Wang, INL, {\em yaqi.wang@inl.gov},\\
Jean Ragusa, TAMU, {\em jean.ragusa@tamu.edu}
}

%%---------------------------------------------------------------------------%%
%% BEGIN NOTE
%%---------------------------------------------------------------------------%%

\opening

%%---------------------------------------------------------------------------%%
\section{Precursor Solve for Shape Evaluation}
%%---------------------------------------------------------------------------%%

This document shows two different methods to solve precursor equations or IQS shape evaluation.  To begin, below is the shape equation coupled with its precursors; for simplicity, the problem is single group with a single precursor group.

\be
\frac{1}{v}\frac{\partial\varphi}{\partial t}=\nu\Sigma_f(1-\beta)\varphi-\left(-\div D \grad + \Sigma_a + \frac{1}{v}\frac{1}{p}\frac{dp}{dt}\right)\varphi+\frac{1}{p}\lambda C 
\ee
\be
\frac{dC}{dt} = \beta\nu \Sigma_f \varphi p - \lambda C
\ee

It is possible to keep these equations separate and discretize them together; however, this is unnecessary and a memory expensive endeavor.  So solving the precursor equation using some discretization or integration scheme to solve the precursors with respect to the shape first, then insert that into the shape equation, solving only one equation at a time.  This document will discuss two techniques to solving the precursor equation.  First is a time discretization method that is currently being implemented in YAK.  The second is a analytical integration of the precursors, this method has proven to be more beneficial for IQS.

\bigskip

%%---------------------------------------------------------------------------%%
\section{Time Discretization using Theta Method}
%%---------------------------------------------------------------------------%%

The simplest way to evaluate the precursor equation is to perform a implicit/explicit finite difference approximation known as the theta method.  The method entails discretizing the the precursor in time and including a weight $(\theta)$ to the implicit and explicit components.  Most generally, if there is a function $u$ defined as:

\be
\frac{du}{dt}=f(u,t)
\ee

Then the discretization would look like:

\be
\frac{u^{n+1}-u^n}{\Delta t}=(1-\theta)f(u^n,t) + \theta f(u^{n+1},t)
\ee

Applying this to the precursor equation:

\be
\frac{C^{n+1}-C^n}{\Delta t}=(1-\theta)\beta(\nu\Sigma_f)^n\varphi^n p^n-(1-\theta)\lambda C^n + \theta\beta(\nu\Sigma_f)^{n+1}\varphi^{n+1}p^{n+1}-\theta\lambda C^{n+1}
\ee

Rearranging to solve for the precursor at the new time step:

\be
C^{n+1}=\frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta (\nu\Sigma_f)^n}{1+\theta\Delta t\lambda}\varphi^n p^n +  \frac{(1-\theta)\Delta t \beta (\nu\Sigma_f)^{n+1}}{1+\theta\Delta t\lambda}\varphi^{n+1} p^{n+1}
\ee

This equation can then be plugged into the shape equation and solving for shape alone.  $\theta$ is a somewhat arbitrary constant $0\le\theta\le1$.  However, the purely explicit method shows when $\theta=0$, implicit when $\theta=1$, and Crank-Nicholson when $\theta=1/2$.  Due to instability, the explicit method is not a preferable option.  YAK currently implements both implicit and Crank-Nicholson as options for precursor evaluation.

\bigskip

%%---------------------------------------------------------------------------%%
\section{Analytical Integration}
%%---------------------------------------------------------------------------%%

Through prototyping, it has been found that neither implicit nor Crank-Nicholson time discretization of precursors are preferable methods for solving the shape equation in IQS.  It has been found that these discretizations result in a lack of convergence of the shape over the IQS iteration process.  In order to remedy the error, a analytical representation of the precursors was implemented in the prototype and the shape solution was able to converge.  The following section shows how this method was implemented in the prototype and the desired implementation in YAK.

Using an exponential operator, the precursor equation can be analytically solved for:

\be
Ce^{\lambda t}=\int\beta\nu\Sigma_f\varphi pe^{\lambda t'}dt'
\ee

The right hand side can be integrated over a time step in order to acquire the precursor at the next time step:

\be
C^{n+1}e^{\lambda t_{n+1}}=\beta\int_{t_n}^{t_{n+1}}\nu\Sigma_f(t')\varphi(t') p(t')e^{\lambda t'}dt' + \lambda C^n e^{\lambda t_n}
\ee

Dividing by $e^{\lambda t_{n+1}}$:

\be
C^{n+1} = \beta\int_{t_n}^{t_{n+1}}\nu\Sigma_f(t')\varphi(t') p(t')e^{-\lambda (t_{n+1}-t')}dt' + \lambda C^n e^{-\lambda \Delta t}
\ee

Because $\nu\Sigma_f$ and $\varphi$ being integrated are not known continuously over the time step, they can be interpolated linearly over the macro step.  Such that:

\be
f(t) \cong \frac{t-t_n}{t_{n+1}-t_n}(f^{n+1}-f^n) \quad t_n \le t \le t_{n+1}
\ee

So:

\be
C^{n+1} = \beta\left(a_3(\nu\Sigma_f)^{n+1}\varphi^{n+1}+a_2(\nu\Sigma_f)^{n+1}\varphi^n + a_2(\nu\Sigma_f)^n\varphi^{n+1} + a_1(\nu\Sigma_f)^n\varphi^n\right) + \lambda C^n e^{-\lambda \Delta t}
\ee

Where the integration coefficients are defined as:

\begin{align}
&a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)^2p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
&a_2= \int_{t_n}^{t_{n+1}}\frac{(t'-t_n)(t_{n+1}-t')}{(\Delta t)^2}p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
&a_3 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)^2p(t')e^{-\lambda(t_{n+1}-t')}dt'
\end{align}

The amplitude $(p)$ is included in the integration coefficient because it has been highly accurately calculated in the micro step scheme, so a piecewise interpolation between those points can be done to maximize accuracy.  However, this term in the integration makes it impossible to do analytically.  The prototype uses Matlab software to interpolate the amplitude between micro steps and a quadrature integration for the coefficients.  So the challenge for YAK is to replicate this procedure: passing the amplitude vector to the DNP auxkernel, interpolating it, and integrating the coefficients.  Applying these coefficients to the "fission source" term seems trivial after the procedure is completed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\closing
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-10-12 doc/note-to-yaqi/precusor-integration.tex
update to prec analytical note

@@
 \deptname{Department of Nuclear Engineering}
-\fromms{Zachary\ M.\ Prince}
-\originator{zmp}
-\typist{zmp}
+\fromms{Zachary\ M.\ Prince, Jean C. Ragusa}
+\originator{zmp,jcr}
+\typist{zmp,jcr}
 \date{\today}
@@
 \distribution {
-Yaqi Wang, INL, {\em yaqi.wang@inl.gov},\\
-Jean Ragusa, TAMU, {\em jean.ragusa@tamu.edu}
+Yaqi Wang, INL, {\em yaqi.wang@inl.gov}
+%,\\
+%Jean Ragusa, TAMU, {\em jean.ragusa@tamu.edu}
 }
@@
 
-This document shows two different methods to solve precursor equations or IQS shape evaluation.
-To begin, below is the shape equation coupled with its precursors; for simplicity, the problem is single group with a single precursor group.
+This document presents two different time-integration methods to solve coupled IQS shape + precursor equations, recalled below
+using, for simplicity, a single neutron group and a single precursor group.
 
@@
 
-It is possible to keep these equations separate and discretize them together; however, this is unnecessary and a memory expensive endeavor.
-So solving the precursor equation using some discretization or integration scheme to solve the precursors with respect to the shape first, then insert that into the shape equation, solving only one equation at a time.
-This document will discuss two techniques to solving the precursor equation.
+First, we note that we could keep this system of two time-dependent equations and solve it as a coupled system. 
+However, this is unnecessary and a memory expensive endeavor because the precursor equation is only an ODE and not a PDE. 
+Instead, one may discretize in time the shape equation, which typically requires the knowledge of the
+precursor concentrations at the end of the time step. This precursor value is taken from the solution, numerical or anlaytical,
+of the precursors ODE.
+This document will discuss two techniques for solving the precursor equation.
 First is a time discretization method that is currently being implemented in YAK.
-The second is a analytical integration of the precursors, this method has proven to be more beneficial for IQS.
-
-\bigskip
+The second is a analytical integration of the precursors, the latter method has proven to be more beneficial for IQS convergence.
 
 %%---------------------------------------------------------------------------%%
-\section{Time Discretization using Theta Method}
+\section{Time Discretization using the Theta Method}
 %%---------------------------------------------------------------------------%%
 
-The simplest way to evaluate the precursor equation is to perform a implicit/explicit finite difference approximation known as the theta method.
-The method entails discretizing the the precursor in time and including a weight $(\theta)$ to the implicit and explicit components.
-Most generally, if there is a function $u$ defined as:
+A fairly simple way to evaluate the precursor equation is to employ the $\theta$-scheme ($0\le\theta\le1$,
+explicit when $\theta=0$, implicit when $\theta=1$, and Crank-Nicholson when $\theta=1/2$).
+Generally, if there is a function $u$ whose governing equation is $\frac{du}{dt}=f(u,t)$,
+then the $\theta$-discretization is
 
 \be
-\frac{du}{dt}=f(u,t)
-\ee
-
-Then the discretization would look like:
-
-\be
-\frac{u^{n+1}-u^n}{\Delta t}=(1-\theta)f(u^n,t) + \theta f(u^{n+1},t)
+\frac{u^{n+1}-u^n}{\Delta t}=(1-\theta)f(u^n,t) + \theta f(u^{n+1},t) \,.
 \ee
@@
 
-Rearranging to solve for the precursor at the new time step:
+Rearranging to solve for the precursor at the end of the time step yields
 
 \be
-C^{n+1}=\frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta (\nu\Sigma_f)^n}{1+\theta\Delta t\lambda}\varphi^n p^n +  \frac{(1-\theta)\Delta t \beta (\nu\Sigma_f)^{n+1}}{1+\theta\Delta t\lambda}\varphi^{n+1} p^{n+1}
+C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta (\nu\Sigma_f)^n}{1+\theta\Delta t\lambda}\varphi^n p^n +  \frac{(1-\theta)\Delta t \beta (\nu\Sigma_f)^{n+1}}{1+\theta\Delta t\lambda}\varphi^{n+1} p^{n+1}
 \ee
-
-This equation can then be plugged into the shape equation and solving for shape alone.
-$\theta$ is a somewhat arbitrary constant $0\le\theta\le1$.
-However, the purely explicit method shows when $\theta=0$, implicit when $\theta=1$, and Crank-Nicholson when $\theta=1/2$.
-Due to instability, the explicit method is not a preferable option.
-YAK currently implements both implicit and Crank-Nicholson as options for precursor evaluation.
+Reporting this value of $C^{n+1}$, one can solve for the shape $\varphi^{n+1}$ as a function of $\varphi^n$ and $C^n$
+(and $p^n$, $p^{n+1}$, $dp/dt|_n$ and  $dp/dt|_{n+1}$).
+Once $\varphi^{n+1}$ has been determined, $C^{n+1}$ is updated. YAK currently implements both implicit and Crank-Nicholson as options for precursor evaluation.
 
@@
 It has been found that these discretizations result in a lack of convergence of the shape over the IQS iteration process.
-In order to remedy the error, a analytical representation of the precursors was implemented in the prototype and the shape solution was able to converge.
-The following section shows how this method was implemented in the prototype and the desired implementation in YAK.
+In order to remedy the error, a analytical representation of the precursors was implemented in the prototype and the shape solution was able to converge (the normalization constant of the IQS method can be preserved to $10^{-10}$ while the theta-scheme only allowed convergence in the normalization factor to about $10^{-3}$).
+The following section shows how this method was implemented in the prototype and the desired implementation for YAK.
 
@@
 \be
-Ce^{\lambda t}=\int\beta\nu\Sigma_f\varphi pe^{\lambda t'}dt'
+\int_{t_n}^{t_{n+1}} C(t')e^{\lambda t'} dt' = \int_{t_n}^{t_{n+1}} \beta\nu\Sigma_f(t')\varphi(t') p(t')e^{\lambda t'}dt'
 \ee
 
-The right hand side can be integrated over a time step in order to acquire the precursor at the next time step:
+yielding
 
 \be
-C^{n+1}e^{\lambda t_{n+1}}=\beta\int_{t_n}^{t_{n+1}}\nu\Sigma_f(t')\varphi(t') p(t')e^{\lambda t'}dt' + \lambda C^n e^{\lambda t_n}
+C^{n+1} =  C^n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta\nu\Sigma_f(t')\varphi(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
 \ee
 
-Dividing by $e^{\lambda t_{n+1}}$:
+Because $\beta\nu\Sigma_f$ and $\varphi$ being integrated are not known continuously over the time step, they can be interpolated linearly over the macro step.  Such that:
 
 \be
-C^{n+1} = \beta\int_{t_n}^{t_{n+1}}\nu\Sigma_f(t')\varphi(t') p(t')e^{-\lambda (t_{n+1}-t')}dt' + \lambda C^n e^{-\lambda \Delta t}
+h(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}h^n  + \frac{t-t_n}{t_{n+1}-t_n}h^{n+1}  \quad t_n \le t \le t_{n+1}
 \ee
+% We let $\ell_n(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}$ and  $\ell_{n+1}(t) = \frac{t-t_n}{t_{n+1}-t_n}$ for simplicity.
 
-Because $\nu\Sigma_f$ and $\varphi$ being integrated are not known continuously over the time step, they can be interpolated linearly over the macro step.  Such that:
+However, for the PRKE solve, we do have a very accurate representation of $p(t')$ over the time interval $[t_n,t_{n+1}]$.
+
+Finally, we have the final expression for the analytical value for $C^{n+1}$:
 
 \be
-f(t) \cong \frac{t-t_n}{t_{n+1}-t_n}(f^{n+1}-f^n) \quad t_n \le t \le t_{n+1}
-\ee
-
-So:
-
-\be
-C^{n+1} = \beta\left(a_3(\nu\Sigma_f)^{n+1}\varphi^{n+1}+a_2(\nu\Sigma_f)^{n+1}\varphi^n + a_2(\nu\Sigma_f)^n\varphi^{n+1} + a_1(\nu\Sigma_f)^n\varphi^n\right) + \lambda C^n e^{-\lambda \Delta t}
+C^{n+1} = C^n e^{-\lambda \Delta t} 
++ \left(a_3(\beta\nu\Sigma_f)^{n+1}+a_2(\beta\nu\Sigma_f)^n\right)\varphi^{n+1}
++ \left(a_2(\beta\nu\Sigma_f)^{n+1}+a_1(\beta\nu\Sigma_f)^n\right)\varphi^n 
 \ee
@@
 The amplitude $(p)$ is included in the integration coefficient because it has been highly accurately calculated in the micro step scheme, so a piecewise interpolation between those points can be done to maximize accuracy.
-However, this term in the integration makes it impossible to do analytically.
-The prototype uses Matlab software to interpolate the amplitude between micro steps and a quadrature integration for the coefficients.
+
+Our prototype code uses Matlab software to interpolate the amplitude between micro steps and a quadrature integration for the coefficients.
 So the challenge for YAK is to replicate this procedure: passing the amplitude vector to the DNP auxkernel, interpolating it, and integrating the coefficients.
-Applying these coefficients to the "fission source" term seems trivial after the procedure is completed.
+
+\underline{``Note-1'':} Applying this procedure to the fission source (assume a linear in time variation for 
+$\beta\nu\Sigma_f(t')\varphi(t')$) is simple. We just need to agree what to do with beta.
+
+\underline{``Note-2'':} We can apply this to the precursor spatial representations as done in YAK, even though
+we think the proper way to handle their spatial discretization is the one we proposed in our previous note.
 

### 2015-10-26 doc/note-to-yaqi/precusor-integration.tex
Fission source term change to precursor-integration note

@@
 \be
-\frac{C^{n+1}-C^n}{\Delta t}=(1-\theta)\beta(\nu\Sigma_f)^n\varphi^n p^n-(1-\theta)\lambda C^n + \theta\beta(\nu\Sigma_f)^{n+1}\varphi^{n+1}p^{n+1}-\theta\lambda C^{n+1}
+\frac{C^{n+1}-C^n}{\Delta t}=(1-\theta)\beta S_f^np^n-(1-\theta)\lambda C^n + \theta\beta S_f^{n+1}p^{n+1}-\theta\lambda C^{n+1}
+\ee
+
+Where $S_f$ is the fission source equivalent for shape:
+
+\be
+S_f^n=(\nu\Sigma_f)^n\varphi^n
 \ee
@@
 \be
-C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta (\nu\Sigma_f)^n}{1+\theta\Delta t\lambda}\varphi^n p^n +  \frac{(1-\theta)\Delta t \beta (\nu\Sigma_f)^{n+1}}{1+\theta\Delta t\lambda}\varphi^{n+1} p^{n+1}
+C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta}{1+\theta\Delta t\lambda}S_f^n p^n +  \frac{\theta\Delta t \beta}{1+\theta\Delta t\lambda}S_f^{n+1} p^{n+1}
 \ee
@@
 \be
-\int_{t_n}^{t_{n+1}} C(t')e^{\lambda t'} dt' = \int_{t_n}^{t_{n+1}} \beta\nu\Sigma_f(t')\varphi(t') p(t')e^{\lambda t'}dt'
+\int_{t_n}^{t_{n+1}} C(t')e^{\lambda t'} dt' = \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{\lambda t'}dt'
 \ee
@@
 \be
-C^{n+1} =  C^n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta\nu\Sigma_f(t')\varphi(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
+C^{n+1} =  C^n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
 \ee
 
-Because $\beta\nu\Sigma_f$ and $\varphi$ being integrated are not known continuously over the time step, they can be interpolated linearly over the macro step.  Such that:
+Because $\beta$ and $\nu\Sigma_f\varphi$ being integrated are not known continuously over the time step, they can be interpolated linearly over the macro step.  Such that:
 
@@
 C^{n+1} = C^n e^{-\lambda \Delta t} 
-+ \left(a_3(\beta\nu\Sigma_f)^{n+1}+a_2(\beta\nu\Sigma_f)^n\right)\varphi^{n+1}
-+ \left(a_2(\beta\nu\Sigma_f)^{n+1}+a_1(\beta\nu\Sigma_f)^n\right)\varphi^n 
++ \left(a_3\beta^{n+1}+a_2\beta^n\right)S_f^{n+1}
++ \left(a_2\beta^{n+1}+a_1\beta^n\right)S_f^n 
 \ee
@@
 \underline{``Note-1'':} Applying this procedure to the fission source (assume a linear in time variation for 
-$\beta\nu\Sigma_f(t')\varphi(t')$) is simple. We just need to agree what to do with beta.
+$\beta\nu\Sigma_f(t')\varphi(t')$) is simple. We just need to agree what to do with beta. If beta is independent of time, the equation reduces to:
+
+\be
+C^{n+1} = C^n e^{-\lambda \Delta t} + \left(\hat{a}_2 S_f^{n+1}+\hat{a}_1 S_f^n\right)\beta
+\ee
+
+\begin{align}
+&\hat{a}_1= \int_{t_n}^{t_{n+1}}\frac{t_{n+1}-t'}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
+&\hat{a}_2 = \int_{t_n}^{t_{n+1}}\frac{t'-t_n}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt'
+\end{align}
 

### 2016-03-08 doc/note-to-yaqi/predictor-corrector_note.tex
IQS-PC note for Yaqi

\documentclass[11pt]{tamurmemo} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{graphics}
\usepackage{float}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsbsy}
\usepackage{xspace}
\usepackage{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% operators
\renewcommand{\div}{\vec{\nabla}\! \cdot \!}
\newcommand{\grad}{\vec{\nabla}}
% latex shortcuts
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bal}{\begin{align}}
\newcommand{\eali}{\end{align}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\ben}{\begin{enumerate}}
\newcommand{\een}{\end{enumerate}}
% DGFEM commands
\newcommand{\jmp}[1]{[\![#1]\!]}                     % jump
\newcommand{\mvl}[1]{\{\!\!\{#1\}\!\!\}}             % mean value
\newcommand{\kef}{\ensuremath{k_{\textit{eff}}}}
%\newcommand{\keff}{{\text{k}$_\textit{eff}$}\xspace}
\newcommand{\keff}{\kef\xspace}
% shortcut for domain notation
%\newcommand{\D}{\mathcal{D}}
% vector shortcuts
\newcommand{\vo}{\vec{\Omega}}
\newcommand{\vr}{\vec{r}}
\newcommand{\vn}{\vec{n}}
\newcommand{\vnk}{\vec{\mathbf{n}}}
\newcommand{\vj}{\vec{J}}
% extra space
\newcommand{\qq}{\quad\quad}
% common reference commands
\newcommand{\eqt}[1]{Eq.~(\ref{#1})}                     % equation
\newcommand{\fig}[1]{Fig.~\ref{#1}}                      % figure
\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table

\newcommand{\ud}{\,\mathrm{d}}
\newcommand{\mt}[1]{\marginpar{\tiny #1}}
\newcommand{\D}{\ensuremath{\mathcal{D}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%   BEGIN DOCUMENT
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%---------------------------------------------------------------------------%%
%% OPTIONS FOR NOTE
%%---------------------------------------------------------------------------%%

\toms{Distribution}
\subject{Predictor Corrector IQS}

%-------NO CHANGES
\collegename{Dwight Look College of Engineering}
\deptname{Department of Nuclear Engineering}
\fromms{Zachary\ M.\ Prince, Jean C. Ragusa}
\originator{zmp,jcr}
\typist{zmp,jcr}
\date{\today}
%-------NO CHANGES

%-------OPTIONS
%\reference{NPB Star Reimbursable Project}
%\thru{}
%\enc{list}      
%\attachments{list}
\cy{File}
%\encas
%\attachmentas
%\attachmentsas 
%-------OPTIONS

%%---------------------------------------------------------------------------%%
%% DISTRIBUTION LIST
%%---------------------------------------------------------------------------%%

\distribution {
Yaqi Wang, INL, {\em yaqi.wang@inl.gov}
%,\\
%Jean Ragusa, TAMU, {\em jean.ragusa@tamu.edu}
}

%%---------------------------------------------------------------------------%%
%% BEGIN NOTE
%%---------------------------------------------------------------------------%%

\opening

%%---------------------------------------------------------------------------%%
\section{Standard IQS Review}
%%---------------------------------------------------------------------------%%

Multigroup diffusion equation with delayed neutron precrusors:
\begin{align}
\frac{1}{v^g} \frac{\partial \phi^g }{\partial t} =& \frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \phi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g  \nonumber \\
&  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \phi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
\label{eq:diffusion}
\end{align}
\be
\frac{dC_i}{dt} = \frac{\beta_i}{k_{eff}}\sum_{g=1}^G\nu^{g} \Sigma_f^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:prec}
\ee

IQS method involves factorizing flux ($\phi^g(\vec{r},t)$) into space-time dependent shape ($\varphi^g(\vec{r},t)$) and time-dependent amplitude ($p(t)$).  This results coupled sytem of shape diffusion and point reactor kinetics for amplitude:
\begin{align}
\frac{1}{v^g} \frac{\partial \varphi^g }{\partial t} =& \frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g  \nonumber \\
&  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}  + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
\label{eq:shape}
\end{align}
\be
\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
\ee
\be
\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
\ee

Where the functional coefficients are calculated using the space-/time-dependent shape function as follows:
\be
\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{k_{eff}}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:rmb}
\ee
\be
\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{k_{eff}}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:b}
\ee
\be
\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
\label{eq:l}
\ee

The shape diffusion is solved at the beginning time step with a already known amplitude, then the PRKE is evaulted at small (micro) time steps until the next large (macro) time step.  The final amplitude from the PRKE is used to solve the shape diffusion equation.  Since the PRKE parameters would be more accurate with this new shape, they are recomputed and the PRKE and shape evaluations are iterated.  The system is iterated until the IQS error is converged to 0:
\be
\text{Error}_{\text{IQS}}=\left|\frac{\left(\phi^*_g,\frac{1}{v_g}\varphi_g^n\right)}{\left(\phi_g^*,\frac{1}{v_g}\varphi_g^0\right)}-1\right| \rightarrow 0
\ee

%%---------------------------------------------------------------------------%%
\section{Predictor-Corrector IQS}
%%---------------------------------------------------------------------------%%

The Predictor-Corrector version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the evaluation of the coupled system is different.  This version first solves the flux diffusion (represented by Equations \ref{eq:diffusion} and \ref{eq:prec}) to get a predicted flux.  The predicted flux at this step is then converted to shape by rescaling as follows:
\be
\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
\ee
Where:
\be
K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
\ee
\be
K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
\ee

The PRKE parameters are then computed with this shape using Equations \ref{eq:rmb} - \ref{eq:l} and interpolated over the macro step, then the PRKE is evaluated.  With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1}
\ee

The precursors need to then be updated, using any of a variety of methods previously discussed in the "Precursor Integration" memo.  Theta method evaluation looks like:
\be
C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta}{1+\theta\Delta t\lambda}S_f^n  +  \frac{\theta\Delta t \beta}{1+\theta\Delta t\lambda}S_f^{n+1} 
\ee
Where:
\be
S_f^n=\sum_{g'=1}^G  \left(\nu^{g'} \Sigma_f^{g'}\right)_n \underbrace{\phi^g_{n}}_{\text{corrected}}
\ee

There is no iteration necessary for this method and in turn is much simpler and faster than the standard IQS.

%%---------------------------------------------------------------------------%%
\section{Rattlesnake Implementation}
%%---------------------------------------------------------------------------%%

In order to preserve the already implemented standard version of IQS, another executioner called IQStmp was created that can be specified in the input deck.  This executioner references the original IQS executioner to execute the PRKE.  Because the diffusion solve is flux instead of shape, when IQStmp is specied, the IQS removal kernel ($\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$) is bypassed, while all the postprocessors are still executed.  However, it is difficult to rescale the flux to shape before the PRKE parameter postprocessor are executed.  So the parameters are computed using the full flux, but amplitude is space independent and comes out of the integrals.  As seen in parameter definitions, when shape is replaced with flux, the amplitude comes out of the integral and cancels out.  So instead of rescaling shape, the predicted flux is rescaled as follows:
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}  p_{n+1} 
\ee

How it is implemented in Rattlesnake is as follows:
\begin{verbatim}
IQStmp::endStep(Real /*input_time*/)
{
  IQS::endStep();
  Real z = (*_Lambda)/_Lambda_orig;
  _nl.update();
  _nl.solution().close();
  _nl.solution().scale(_X.get_values()[0]/z);
  _nl.solution().close();
  _nl.update();
}
\end{verbatim}

However, how to update the precursors auxilary variable is still not understood.  The predicted easiest way is to simply re-evaluate the auxkernel, instead of doing something complicated in the executioner.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\closing
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-06-05 doc/ANS_winter_2016/ANS_summary.tex
ANS Winter 2016 summary draft

\documentclass{anstrans}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Improved Quasi-Static Method with Step Doubling Time Adaptation}
\author{Zachary M. Prince$^{*}$, Dr. Jean C. Ragusa$^{*}$, Yaqi Wang$^{\dagger}$}

\institute{
$^{*}$Department of Nuclear Engineering, Texas A\&M University, College Station, TX
\and
$^{\dagger}$Idaho National Laboratory, Idaho Falls, ID
}

\email{zachmprince@tamu.edu \and jean.ragusa@tamu.edu \and yaqi.wang@inl.gov}

% Optional disclaimer: remove this command to hide
%\disclaimer{Notice: this manuscript is a work of fiction. Any resemblance to actual articles, living or dead, is purely coincidental.}

%%%% packages and definitions (optional)
\usepackage{graphicx} % allows inclusion of graphics
\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{microtype} % improves typography for PDF
\usepackage{subcaption}

\newcommand{\SN}{S$_N$}
\renewcommand{\vec}[1]{\bm{#1}} %vector is bold italic
\newcommand{\vd}{\bm{\cdot}} % slightly bold vector dot
\newcommand{\ud}{\mathop{}\!\mathrm{d}} % upright derivative symbol
%  new definitions
\newcommand{\bs}[1]{\mathbf{#1}}
\renewcommand{\div}{\bs{\nabla}\! \cdot \!}
\newcommand{\grad}{\bs{\nabla}}
% extra space
\newcommand{\qq}{\quad\quad}
% common reference commands
\newcommand{\eqt}[1]{Eq.~(\ref{#1})}                     % equation
\newcommand{\fig}[1]{Fig.~\ref{#1}}                      % figure
\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table
\newcommand{\sct}[1]{Section~\ref{#1}}                   % section
\newcommand{\app}[1]{Appendix~\ref{#1}}                   % appendix

\newcommand{\keff}{k_\textit{eff}}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\vn}{\vec{n}}
\newcommand{\vel}{\vec{\mathrm{v}}}
\newcommand{\adj}{\Phi^\dagger_0}
\newcommand{\tcr}[1]{\textcolor{red}{#1}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert_{L^2}}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
The imminent restart of the Transient Reactor Testing Facility (TREAT) at Idaho National Laboratory (INL) has brought significant attention to the development of transient reactor modeling.  TREAT was designed to subject fuels and other reactor components to various degrees of neutron pulses, from minor transients to accident scenario.  Inherently, this task requires predicting pulse responses due to various control rod movements; which is a complicated endeavor.  Originally, this was done using an iterative process of rod movements and measuring the resulting response.  With the advances in computer modeling, the current goal, preceding the restart, is to use simulation to characterize responses before experiments take place.  However, neutron transient modeling is extremely computationally expensive due to the required implicit time-stepping.  Therefore, methods that mitigate computation time, with minimal detriment to accuracy, are highly desired.  This paper will describe the improved quasi-static method (IQS), which hopes to decrease computation time significantly from traditional "brute force" transient methods.

IQS is a spatial kinetics method that involves factorizing the flux solution into time-dependent amplitude and space- and time-dependent shape \cite{Ott_1966,Dulla2008}.  The impetus of the method is the assumption that shape is weakly time-dependent, so expensive space-dependent evaluations are only required on large time steps.  While inexpensive amplitude evaluations are performed on much smaller time steps to retain accuracy.  IQS has been tested extensively on kinetics benchmarks with constant time steps.  However, to test it's viability for TREAT simulations, evaluating IQS's performance with time adaptation is required.  The rest of thispaper will brefly desribe the derivation of IQS, the time adaptation technique used, and results from two kinetic benchmarks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
In this Section, we recall the equation for the IQS method, starting from the standard multi-group diffusion equations with delayed neutron precursors in operator form:

\begin{align}
\frac{1}{v^g} \frac{\partial \phi^g }{\partial t} =& \frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \phi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g  \nonumber \\
&  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \phi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
\label{eq:flux}
\end{align}
\be
\frac{dC_i}{dt} = \frac{\beta_i}{k_{eff}}\sum_{g=1}^G\nu^{g} \Sigma_f^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:precursor}
\ee

Factorization is an important step in the derivation of the IQS method. The factorization approach leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent multigroup shape ($\varphi$):
\be
\phi^g(\vec{r},t)=p(t)\varphi^g(\vec{r},t)
\ee
To obtain the amplitude equations, the multigroup equations are multiplied by a weighting function, typically the initial adjoint flux ($\phi^*$), and then integrated over phase-space.  For brevity, the inner product over space will be represented with parenthetical notation:
\be
\int_D\phi^{*g}(\vec{r})f^g(\vec{r})d^3r=\left(\phi^{*g},f^g\right)
\ee
In order to impose uniqueness of the factorization, one requires the following:
\be
\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}
\ee
After some manipulation, the standard point reactor kinetics equations (PRKE) for the amplitude solution are obtained:
\be
\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
\ee
\be
\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
\ee
Where the functional coefficients are calculated using the space-/time-dependent shape function as follows:
\begin{align}
\frac{\rho-\bar{\beta}}{\Lambda}=&\Bigg[\sum_{g=1}^G \bigg(\phi^{*g},\frac{\chi_p^g}{k_{eff}}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\
& -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g \bigg)\Bigg] \times \Bigg[\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)\Bigg]^{-1}
\label{eq:rmb}
\end{align}
\be
\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{k_{eff}}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:b}
\ee
\be
\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
\label{eq:l}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IQS Predictor-Corrector (IQS P-C)}

This version of IQS first solves the flux diffusion (represented by Equations \ref{eq:flux} and \ref{eq:precursor}) to get a predicted flux.  The predicted flux at this step is then converted to shape by rescaling as follows:
\be
\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
\label{eq:rescale}
\ee
Where:
\be
K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
\ee
\be
K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
\ee

The PRKE parameters are then computed with this shape using Equations \ref{eq:rmb} - \ref{eq:l} and interpolated over the macro step, then the PRKE is evaluated.  With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Adaptation}

The time adaptation used for quantifying IQS's ability is step doubling.  The step doubling technique involves estimating the local truncation error for a certain time step by taking the difference between a solution with one full step and a solution with two half steps.  If the step is small enough, the error will be smaller than a user driven tolerance and the magnitude of the next step will be calculated based on the error.  If the step is too large, the step will be repeated with a smaller step calculated with the resulting error.  The error of the time step is approximated by Equation \ref{eq:edt2}, where $\varphi^g_1$ and $\varphi^g_2$ are the solutions with the full step and half step, respectively.  $\varphi$ is changed to $\phi$ for regular flux evaluation and IQS P-C.  If $e_n > e_{max}$ the time step is repeated; if $e_n < e_{max}$ the system moves to the next time step.  The next $\Delta t$ is calculated using Equation \ref{eq:dt2}, where $\mu$ is the convergence order of the time integration scheme being used.  $e_{max}$ and $e_{tol}$ are user defined parameters; $e_{max}$ is usually less than $e_{tol}$ to better guarantee that the calculated $\Delta t_{new}$ will pass the error criteria so time steps won't be repeated.

\be
e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_2 - \sum_{g=1}^G\varphi^g_1}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_2},\norm{\sum_{g=1}^G\varphi^g_1}\right)}
\label{eq:edt2}
\ee
\be
\Delta t_{new} = \Delta t_{old} \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{\mu+1}}
\label{eq:dt2}
\ee

To be clear, each solution is re-scaled by $p$ when calculating $e_n$.  Additionally, if the PJFNK iteration does not converge, the entire step doubling process is repeated with half the time original time step.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Analysis}
This section describes results of an examples that tests the IQS implementation and shows its effectiveness on computation speed and accuracy.  Two examples were selected for this purpose.  The first is the two-dimensional TWIGL ramp transient benchmark, described further.  The second is the LRA benchmark, which a two-dimensional, temperature feedback example.  Each example was solved with regular diffusion (brute force) and IQS P-C.  Each method was tested with backward-Euler for the TWIGL benchmark and Crank-Nicholson was used for the LRA benchmark.  The performance for each method are represented by the number of macro time steps and solves with step doubling time adaptation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TWIGL Benchmark}

This benchmark problem originates from the Argonne National Lab Benchmark Problem Book \cite{BPB}.  It is a 2D, 2-group reactor core model with no reflector region shown in Figure \ref{fig:TWIGL_reg}.  Table \ref{tab:TWIGL_mat} shows the material properties of each fuel region and the ramp perturbation of Material 1.

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{TWIGL_regions.jpg}
\caption{TWIGL benchmark problem description}
\label{fig:TWIGL_reg}
\end{center}
\end{figure}

\begin{table}[htp]
\begin{center}
\resizebox{0.48\textwidth}{!}{
\begin{tabular}{lllllll}
\hline
  &  &  &  &  &  &  $\underline{\Sigma_s (cm^{-1})} $ \\
Material & Group & $D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu\Sigma_f (cm^{-1})$ & $\chi$ & $g \rightarrow 2$ \\
\hline
1 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.01 \\
  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.00  \\
2 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.01  \\
  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.00  \\
3 & 1 & 1.3 & 0.008 & 0.003 & 1.0 & 0.01  \\
  & 2 & 0.5 & 0.050 & 0.060 & 0.0 & 0.00  \\
\hline
  & $\nu$ & $v_1 (cm/s)$ & $v_2 (cm/s)$ & $\beta$ & $\lambda (1/s)$ &  \\
\hline
  & 2.43 & 1.0E7 & 2.0E5 & 0.0075 & 0.08 & \\
\hline
 \multicolumn{7}{l}{\footnotesize Material 1 ramp perturbation:} \\
\multicolumn{7}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (1-0.11667t) \quad t \leq 0.2 s$} \\
\multicolumn{7}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (0.97666t) \quad t > 0.2 s$} \normalsize
\end{tabular}}
\end{center}
\caption{TWIGL material properties and ramp perturbation}
\label{tab:TWIGL_mat}
\end{table}

Figure \ref{fig:TWIGL_power} shows the IQS  solution as compared with the Brute Force solution with minimal time steps.  Table \ref{tab:TWIGLdt2} shows the results for time adaptation.  The results show that IQS performs exceptionally well compared to brute force for this highly transient example.

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.45\textwidth]{TWIGL_power_plot.png}
\caption{Power level comarison of TWIGL benchmark between IQS and Brute Force using $e_{max} = 0.001$}
\label{fig:TWIGL_power}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\resizebox{0.45\textwidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|l|l|}
\hline
  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Test & $e_{max}$ & Error & Steps & Solves & Error & Steps & Solves \\
\hline
1 &	0.05    &	0.00012677 &	9   &	29   &	0.03380433 &	4  &	9   \\
2 &	0.01    &	3.5555e-05 &	5   &	40   &	0.00263068 &	5  &	12  \\
3 &	0.005   &	4.0364e-05 &	5   &	40   &	0.00160486 &	6  &	21  \\
4 &	0.001   &	0.00294822 &	5   &	36   &	1.7527e-05 &	10 &	35  \\
5 &	0.0005  &	0.00099778 &	6   &	55   &	1.4185e-05 &	16 &	74  \\
6 &	0.0001  &	0.00019510 &	7   &	65   &	6.2903e-06 &	19 &	78  \\
7 &	5.0e-05 &	0.00018372 &	12  &	163  &	1.5247e-06 &	24 &	92  \\
8 &	1.0e-05 &	8.0564e-05 &	379 &	5729 &	9.8321e-07 &	48 &	210 \\
\hline

\end{tabular}}
\end{center}
\caption{TWIGL step doubling results}
\label{tab:TWIGLdt2}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{LRA Benchmark}

The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{BPB}.  It is a super prompt-critical transient.  To have better understanding on the cross sections given later, we present the equations here:
\begin{align}
-\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\grad D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 \nonumber \\
&- \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
-\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\grad D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
S_f &= \sum_{g=1}^2 \Sigma_{f,g} \phi_g, \\
\frac{\partial C_i}{\partial t} &= \nu\beta_i f - \lambda_i C_i, \quad i=1,2, \\
\frac{\partial T}{\partial t} &= \alpha f, \label{eq:lra-temp} \\
\Sigma_{a,1} &= \Sigma_{a,1}(\vec{r}, t=0) \left[1+\gamma\left(\sqrt{T} - \sqrt{T_0}\right)\right], \\
P &= \kappa S_f,
\end{align}
where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_6f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0=T(\vec{r}, t=0)$.
The two-group diffusion equation are solved with zero flux boundary conditions on external surfaces, reflecting conditions at symmetry boundaries and steady state initial conditions which are obtained by solving
\begin{align}
-\grad D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
-\grad D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1
\end{align}
The eigenvalue $k$ is used to modify the fission cross section for the transient simulations with $\frac{1}{k}\Sigma_{f,g}, g=1,2$.  The initial flux distribution shall be normalized such that the averaged power density
\begin{align}
\bar{P} \equiv \frac{\int_{V_{core}} P(\vec{r}, t=0) d\vec{r}}{\int_{V_{core}} d\vec{r}},
\end{align}
where $V_{core}$ is the core region with fuels, is equal to $10^{-6} W\cdot cm^{-3}$.
The initial precursor concentrations are in equilibrium with the initial critical flux distribution.\\

The geometry is illustrated in~\fig{fig:lra-geometry}.\\
\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\linewidth]{lra.png}
\caption{LRA benchmark geometry with region assignment.}
\label{fig:lra-geometry}
\end{figure}

Initial two-group constants are presented in~\tbl{tab:lra-xs}.
$\nu$ is equal to 2.43.
Axial bulking $B^2 = 10^{-4}$ is applied for both energy groups.
Delayed neutron data are presented in~\tbl{tab:lra-dnp}.
All fuel materials have the same delayed neutron data.
Some scalar data are listed in~\tbl{tab:lra-scalar}.\\
\begin{table}[!htbp]
    \centering
    \resizebox{0.5\textwidth}{!}{
      \begin{tabular}{|c|c|c|c|l|l|l|l|l|}
      \hline
             &                 & Group   & $D_g$      & $\Sigma_{a,g}$ & $\nu\Sigma_{f,g}$ & $\Sigma_{s,1\rightarrow 2}$ & $\chi_g$ & $v_g$              \\
      Region & Material        &       g & (cm)       &  ($cm^{-1}$)   &  ($cm^{-1}$)      &  ($cm^{-1}$)                &          & ($cm\cdot s^{-1}$) \\
      \hline
      1      & Fuel 1 with rod & 1       & 1.255      & 0.008252       & 0.004602          &                             & 1        & $3.0\times10^7$    \\
             &                 & 2       & 0.211      & 0.1003         & 0.1091            & 0.02533                     & 0        & $3.0\times10^5$    \\
      \hline
      2      & Fuel 1 without rod & 1    & 1.268      & 0.007181       & 0.004609          &                             & 1        & $3.0\times10^7$    \\
             &                    & 2    & 0.1902     & 0.07047        & 0.08675           & 0.02767                     & 0        & $3.0\times10^5$    \\
      \hline
      3      & Fuel 2 with rod & 1       & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
             &                 & 2       & 0.2091     & 0.08344        & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
      \hline
      4      & Fuel 2 without rod & 1    & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
             &                    & 2    & 0.2091     & 0.073324       & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
      \hline
      5      & Reflector        & 1      & 1.257      & 0.0006034      & -                 &                             & -        & $3.0\times10^7$    \\
             &                  & 2      & 0.1592     & 0.01911        & -                 & 0.04754                     & -        & $3.0\times10^5$    \\
      \hline
      \end{tabular}}
    \caption{LRA benchmark initial two-group constants.\label{tab:lra-xs}}
\end{table}

\begin{table}[!htbp]
    \centering
      \begin{tabular}{|c|l|l|l|l|}
      \hline
      Group i & $\beta_i$ & $\lambda_i$ ($s^{-1}$) & $\chi_{d,i,1}$ & $\chi_{d,i,2}$ \\
      \hline
      1       & 0.0054    & 0.0654  & 1 & 0 \\
      2       & 0.001087  & 1.35    & 1 & 0 \\
      \hline
      \end{tabular}
    \caption{LRA benchmark delayed neutron data.\label{tab:lra-dnp}}
\end{table}

\begin{table}[!htbp]
    \centering
    \resizebox{0.5\textwidth}{!}{
      \begin{tabular}{|l|c|l|}
      \hline
      Meaning & Notation & value \\
      \hline
      Axial buckling for both energy groups & $B_g^2$  & $10^{-4}$ ($cm^{-2}$)\\
      Mean number of neutrons per fission   & $\nu$    & 2.43 \\
      Conversion factor                     & $\alpha$ & $3.83\times 10^{-11}$ ($K\cdot cm^{3}$) \\
      Feedback constant                     & $\gamma$ & $3.034\times 10^{-3}$ ($K^{1/2}$) \\
      Energy released per fission           & $\kappa$ & $3.204\times 10^{-11} $ ($J/fission$) \\
      Initial and reference temperature     & $T_0$    & 300 (K) \\
      Active core volume                    & $V_{core}$ & 17550 ($cm^2$)\\
      \hline
      \end{tabular}}
    \caption{LRA benchmark scalar values.\label{tab:lra-scalar}}
\end{table}

The transient is initiated by changing the thermal absorption cross section as the following:
\begin{align}
\Sigma_{a,2}(t) = \Sigma_{a,2}(t=0) \left\{\begin{array}{lr} 1-0.0606184t, & t\leq 2 \\
                                                             0.8787631, & t>2
                                           \end{array}\right.
\end{align}
where $t$ is time in seconds.\\

Figure \ref{fig:LRA_plots} shows the IQS P-C power profile as compared with the Brute Force solution.    These plots show that IQS is consistent with Doppler feedback problems.  Table  \ref{tab:LRAdt2} shows the time adaptation performance of both methods.  These results show that IQS, again, performs much better than the brute force method by taking significantly less time steps while retaining similar accuracy.

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{lra_power_profile.jpg}
\caption{Power level comarison of LRA benchmark between IQS and Brute Force using $e_{max} = 0.01$}
\label{fig:LRA_plots}
\end{center}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\resizebox{0.48\textwidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|l|l|}
\hline
  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Test & $e_{max}$ & Error & Steps & Solves & Error & Steps & Solves \\
\hline
1 &	0.01    &	3.5555e-05 &	145  &	735  &	0.00263068 &	105 &	616  \\
2 &	0.005   &	4.0364e-05 &	266  &	948  &	0.00160486 &	157 &	814  \\
3 &	0.001   &	0.00294822 &	585  &	2046 &	1.7527e-05 &	218 &	1062  \\
4 &	0.0005  &	0.00099778 &	833  &	2908 &	1.4185e-05 &	480 &	2038  \\
5 &	0.0001  &	0.00019510 &	1793 &	6155 &	6.2903e-06 &	434 &	1458  \\
6 &	5.0e-05 &	0.00018372 &	2485 &	8511 &	1.5247e-06 &	542 &	1719  \\
\hline

\end{tabular}}
\end{center}
\caption{LRA step doubling results}
\label{tab:LRAdt2}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

The purpose of this paper was to show IQS's performance with step doubling time adaptation.  Two examples were used to quantify the method's performance, including the TWIGL and LRA benchmarks.  The results from the TWIGL example shows significant improvement from the traditional brute force method.  This improvement was anticipated because the shape for TWIGL changes very little through the transient, so the PRKE in IQS does most of the work.  The results for the LRA benchmark showed similar improvement with less time steps for the same accuracy.  In conclusion, IQS shows impressive performance for complex multi-physics problems with time adaptation.  However, to rigorously test IQS, more complex problems including TREAT models and transport problems need to be applied.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\appendix
%\section{Appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}

This work was supported by the Department of Energy and Idaho National Laboratory.  Special thanks to Mark Dehart, NEAMS, and the Moose/Rattlesnake team at INL.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{ans}
\bibliography{references_IQS}
\end{document}

### 2016-06-07 doc/NEAMS Report/NEAMS_Report_V2.tex
moved figures to a subfolder

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{scrartcl} 
%########################### Preferences #################################

\usepackage{vmargin}
\usepackage{color}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amstext}
\usepackage{amsbsy}
%\usepackage{mathbbol}
\usepackage{graphicx} 
\usepackage{verbatim}
\usepackage{csvsimple} 
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{multirow}

%  new definitions
\renewcommand{\div}{\bs{\nabla}\! \cdot \!}
\newcommand{\grad}{\bs{\nabla}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert_{L^2}}
% extra space
\newcommand{\qq}{\quad\quad}
% common reference commands
\newcommand{\eqt}[1]{Equation~(\ref{#1})}                     % equation
\newcommand{\fig}[1]{Figure~\ref{#1}}                      % figure
\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table
\newcommand{\sct}[1]{Section~\ref{#1}}                   % section
\newcommand{\app}[1]{Appendix~\ref{#1}}                   % appendix

\newcommand{\bs}[1]{\mathbf{#1}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\keff}{k_\textit{eff}}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\vn}{\vec{n}}
\newcommand{\vel}{\vec{\mathrm{v}}}
\newcommand{\adj}{\Phi^\dagger_0}
\newcommand{\tcr}[1]{\textcolor{red}{#1}}
\newcommand{\tcb}[1]{\textcolor{blue}{#1}}

% ********* Caption Layout ************
%\usepackage{ccaption} % allows special formating of the captions
%\captionnamefont{\bf\footnotesize\sffamily} % defines the font of the caption name (e.g. Figure: or Table:)
%\captiontitlefont{\footnotesize\sffamily} % defines the font of the caption text (same as above, but not bold)
%\setlength{\abovecaptionskip}{0mm} %lowers the distance of captions to the figure
\setlength\parindent{0pt}
\setlength{\oddsidemargin}{1in}
\setlength{\evensidemargin}{1in}
\setlength{\textwidth}{6.5in}
% ********* Header and Footer **********
% This is something to play with forever. I use here the advanced settings of the KOMA script

\title{Improved Quasi-Static Methods in RATTLESNAKE}
\subtitle{Development and Implementation of Improved Quasi-Static (IQS) methods for time-dependent neutron diffusion and neutron transport solvers in RATTLESNAKE}
\author{ \normalsize
  \textbf{Zachary M. Prince$^\dagger$, Jean C. Ragusa$^\dagger$, Yaqi Wang$^\star$} \\
\normalsize $^\dagger$Department of Nuclear Engineering \\
 \normalsize Texas A\&M University, College Station, TX, USA\\
 \normalsize $^\star$Idaho National Laboratory\\
 \normalsize \href{mailto:zachmprince@tamu.edu}{zachmprince@tamu.edu}, \href{jean.ragusa@tamu.edu}{jean.ragusa@tamu.edu}, \href{yaqi.wang@inl.gov}{yaqi.wang@inl.gov} 
}

\hypersetup{
  colorlinks=true,
  urlcolor=blue,
}

\pagestyle{fancy}
\fancyhf{}
\lhead{INL/EXT-16-38059}
\rfoot{DRAFT}
\cfoot{\thepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% boxes
\usepackage{color}
\definecolor{myblue}{rgb}{.8, .8, 1}
\usepackage{empheq}

\newlength\mytemplen
\newsavebox\mytempbox

\makeatletter
\newcommand\mybluebox{%
    \@ifnextchar[%]
       {\@mybluebox}%
       {\@mybluebox[0pt]}}

\def\@mybluebox[#1]{%
    \@ifnextchar[%]
       {\@@mybluebox[#1]}%
       {\@@mybluebox[#1][0pt]}}

\def\@@mybluebox[#1][#2]#3{
    \sbox\mytempbox{#3}%
    \mytemplen\ht\mytempbox
    \advance\mytemplen #1\relax
    \ht\mytempbox\mytemplen
    \mytemplen\dp\mytempbox
    \advance\mytemplen #2\relax
    \dp\mytempbox\mytemplen
    \colorbox{myblue}{\hspace{1em}\usebox{\mytempbox}\hspace{1em}}}

\makeatother

%################ End Preferences, Begin Document #####################

\pagestyle{plain} % on headers or footers on the first page

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% set relative directory for figures. can be changed at any time later with \renewcommand
\newcommand{\FiguresDir}{./figs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
\pagenumbering{arabic}

\tcb{I will fix the cover page and page footers later}

\tcb{I removed CFEM from the title}

\tcr{you need Yaqi's concurrence since you list him as co-author}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{Abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tcr{we need a short abstract}

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\par
The improved quasi-static (IQS) method \tcr{you need to give 3-4 references} is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (the shape) and a time-dependent component (the amplitude). The technique relies on the shape being less rapidly varying in time compared to the flux, hence
requiring fewer shape computations or updates. 
The IQS method has primarily been developped in the context of the neutron diffusion approximation \tcr{give refs}, but \tcr{cite sedat} provides an extension of the technique to transport, implemented in TDTORT. 
The method has mostly being applied to neutron kinetics, without thermal-hydraulic feedback. In this report, we extend the IQS
method in the context of transient multiphysics simulations.
%The method These components include the flux's power and shape. Power is time-dependent, while the shape is both space- and time-dependent.  However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the shape may not require computation at every time step, invoking the “quasi-static” nature.
\\

In this Section, we recall the equations for the IQS method, using the standard multigroup diffusion equations written below:
\begin{align}
\frac{1}{v^g} \frac{\partial \phi^g }{\partial t} =& \frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \phi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g  \nonumber \\
&  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \phi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
\label{eq:flux}
\end{align}
\be
\frac{dC_i}{dt} = \frac{\beta_i}{\keff}\sum_{g=1}^G\nu^{g} \Sigma_f^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:precursor}
\ee
with
\be
\beta = \sum_{i=1}^I \beta_{i} 
\ee
(extension to the transport equations will be provided subsequently, once a concise operator notation has been introduced). 
In the above equations, \tcr{give a long text naming each variables, for example: $\phi^g$ is the neutron flux in group $g$, ....}. Note that the prompt fission terms have been normalized by $\keff$, the eigenvalue of the initial state, as transient simulations are often carried out for initially critical reactors. The expressions for subcritical and source-driven problems are straightforward to derive but omitted in this report for brevity.

The factorization approach in IQS leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent multigroup shape ($\varphi^g$):
\be
\phi^g(\vec{r},t)=p(t)\varphi^g(\vec{r},t)
\ee
The factorization is not an approximation. Reporting it in the flux and precursor equations yields the following equations
\begin{enumerate}
\item for the shape:
\begin{align}
\frac{1}{v^g}\left(\frac{dp}{dt}\varphi^g + p\frac{\partial \varphi^g}{\partial t}\right) =&\frac{\chi_p^g}{k_{eff}}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} p \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right)p \varphi^g \nonumber \\
&  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} p \varphi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
\label{eq:shape}
\end{align}
\be
\frac{dC_i}{dt} = \frac{\beta_i}{k_{eff}}\sum_{g=1}^G\nu^{g} \Sigma_f^g p \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:prec}
\ee

PRKE formulation (amplitude equations):

\begin{align}
\sum_{g=1}^G\int_D &\left[\frac{1}{v^g}\left(\frac{dp}{dt}\varphi^g + p\frac{\partial \varphi^g}{\partial t}\right)\right]w^gd^3r =  \nonumber \\
& \sum_{g=1}^G\int_D\left[\frac{\chi_p^g}{k_{eff}} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} p \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right)p \varphi^g\right]w^gd^3r   \\ 
 +&\sum_{g=1}^G\int_D \left[\sum_{g'\neq g}^G\Sigma_s^{g'\to g} p \varphi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i\right]w^gd^3r \ , \quad 1 \le g \le G  \nonumber
\end{align}
\be 
\int_D\frac{dC_i}{dt}\sum_{g=1}^G\chi_{d,i}^gw^gd^3r = \int_D \frac{\beta_i}{k_{eff}}\left[\sum_{g=1}^G\left(\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'} - \lambda_i C_i \right)\chi_{d,i}^g w^g\right]d^3r \ , \quad 1 \le i \le I 
\ee

It can be shown that the most appropriate weighting function $(w^g)$ is the initial adjoint flux $(\phi^{*g})$.  For brevity, the following definition will be applied: $\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r=\left(\phi^{*g},f\right)$
\begin{align}
& \quad \frac{dp}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)+ p\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)= \nonumber \\
& p\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{k_{eff}}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)  \\
& + \sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i) \nonumber
\end{align}
\be
\frac{d}{dt}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)=\frac{1}{k_{eff}}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})p - \sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \quad 1 \le i \le I 
\ee

In order to impose uniqueness of the factorization, one imposes:
\be
\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=constant
\ee

Therefore the PRKE formulation reduces to:
\begin{align}
\frac{dp}{dt}= &\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{k_{eff}}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g}  \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p \nonumber \\ 
& +\sum_{i=1}^I\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i
\end{align}
\be
\frac{d\xi_i}{dt}=\frac{1}{k_{eff}}\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p - \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i \quad 1 \le i \le I 
\ee
Where:
\be
\xi_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g C_i)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\ee

It is convenient to define the effective reactivity, delay-neutron fraction, and delayed-neutron precursor decay constant:
\be
\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{k_{eff}}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:rmb}
\ee
\be
\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{k_{eff}}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:b}
\ee
\be
\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
\label{eq:l}
\ee

So:
\be
\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
\label{eq:p}
\ee
\be
\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
\label{eq:c}
\ee

Equations (\ref{eq:p}) and (\ref{eq:c}) are a formulation of the point reactor kinetics equation (PRKE), but the parameters (Equations (\ref{eq:rmb})-(\ref{eq:l})) are dependent on the shape.  If the assumption is made that the shape is time-independent, the shape is computed once at the first time step and used for the PRKE parameter evaluation at all other steps.  However, if the shape is dependent on time, the shape needs to be computed in transient using equation (\ref{eq:shape}) and (\ref{eq:prec}) in order retain accuracy.  Equations (\ref{eq:shape2}) and (\ref{eq:preq2}) shows the usual form of the shape and precursor equations with amplitude put on the right hand side.  Equation (\ref{eq:shape2}) is very similar to the multigroup flux equation (\ref{eq:flux}), except the removal cross-section term is augmented by a $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term and the precursor contribution has a $\frac{1}{p}$ mulitiplier.  Equation (\ref{eq:preq2}) is very similar to the normal precursor equation (\ref{eq:precursor}), except the fission source term is multiplied by $p$.  These differences are crucial for IQS implementation in Rattlesnake.

\begin{align}
\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\ 
& -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
\label{eq:shape2}
\end{align}
\be
\frac{dC_i}{dt} = \frac{\beta_i}{k_{eff}}p \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:preq2}
\ee

%\begin{align}
%& \quad \sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}\right)= \nonumber \\ 
%&\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g\right) \\
%&+\frac{1}{p}\sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \nonumber
%\end{align}

Computing this shape can become expensive, especially in two or three dimensions.  Subsequently, it is attractive to make the assumption that the shape is weakly time-dependent so the shape can be computed after a multitude of PRKE calculations which is the root of IQS.  To visualize:
\begin{figure}[!htbp]
\includegraphics[width=\linewidth]{\FiguresDir/IQS_visualization.jpg}
\caption{IQS method visualization}
\end{figure}

Additionally, to improve consistency and accuracy, each macro time step can be iterated so the best shape is used to compute power at the micro time steps.  This iteration process must converge the shape such that the uniqueness condition $(\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=0)$ is preserved.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IQS Predictor-Corrector (IQS P-C)}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Predictor-Corrector version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the evaluation of the coupled system is different.  This version first solves the flux diffusion (represented by Equations \ref{eq:flux} and \ref{eq:precursor}) to get a predicted flux.  The predicted flux at this step is then converted to shape by rescaling as follows:
\be
\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
\label{eq:rescale}
\ee
Where:
\be
K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
\ee
\be
K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
\ee

The PRKE parameters are then computed with this shape using Equations \ref{eq:rmb} - \ref{eq:l} and interpolated over the macro step, then the PRKE is evaluated.  With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1}
\ee

The advantage to the predictor-corrector method is there is no iteration necessary for this method and in turn is much simpler and faster than the standard IQS.  The disadvantage is that the method assumes $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)$ is inherently constant.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rattlesnake Implementation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section will explain how, thus far, IQS has been implemented in Rattlesnake.

\subsection{Executioner}
The IQS executioner derives from the Transient executioner in MOOSE.  The IQS executioner contains a loop over micro time steps that computes the PRKE and then passes $p$ and $\frac{dp}{dt}$ for the Transient executioner to evaluate the shape equation at each macro step.  The PRKE is computed with a user specified option of backward-Euler, Crank-Nicolson, or SDIRK33.   The IQS executioner also supplements Transient’s Picard iteration process by adding its own error criteria:
\be
Error_{IQS}=\left|\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,n}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,0}\right)}-1\right|
\ee

\subsection{Action System}
IQS defines its uniqueness from its executioner type; however, many changes needed to be made in the Rattlesnake action system in order to support IQS execution.   First, changes needed to be made in order to evaluate the shape equation.  The shape equation, after some manipulation, is very similar to the time-dependent, which Rattlesnake is already set up to solve:

\begin{align}
\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
& - \underbrace{\frac{1}{v^g} \overbrace{\frac{1}{p}\frac{dp}{dt}}^{From Executioner}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
\end{align}

To enable Rattlesnake to solve this equation, another kernel was created that evaluated $\sum_{g=1}^G\frac{1}{v^gp}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.  Second, four postprocessors were created in order to calculate the PRKE parameters.  The parameter calculations were separated by $\frac{\bar{\beta}_i}{\Lambda}$ numerator, $\bar{\lambda}_i$ numerator/denominator, $\frac{\rho}{\Lambda}/\frac{\bar{\beta}}{\Lambda}$ denominator, and $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator.  The first three are relatively simple, only relying on material properties and solution quantities.  The $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator requires the use of the MOOSE save in feature, which saves the residual from a calculated kernel or boundary contribution in the shape evaluation to an auxiliary variable.  Finally, a user object was created to pull together all the postprocessor values and carryout the numerator/denominator divisions that were then passed to the executioner.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Precursor Integration}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
This section presents two different time-integration methods to solve coupled IQS shape + precursor equations, recalled below
using, for simplicity, a single neutron group and a single precursor group.

\be
\frac{1}{v}\frac{\partial\varphi}{\partial t}=\nu\Sigma_f(1-\beta)\varphi-\left(-\div D \grad + \Sigma_a + \frac{1}{v}\frac{1}{p}\frac{dp}{dt}\right)\varphi+\frac{1}{p}\lambda C 
\ee
\be
\frac{dC}{dt} = \beta\nu \Sigma_f \varphi p - \lambda C
\ee

First, we note that we could keep this system of two time-dependent equations and solve it as a coupled system. 
However, this is unnecessary and a memory expensive endeavor because the precursor equation is only an ODE and not a PDE. 
Instead, one may discretize in time the shape equation, which typically requires the knowledge of the
precursor concentrations at the end of the time step. This precursor value is taken from the solution, numerical or anlaytical,
of the precursors ODE. This document will discuss two techniques for solving the precursor equation.  First is a time discretization method that is currently being implemented in Rattlesnake.  The second is a analytical integration of the precursors, the latter method has proven to be more beneficial for IQS convergence.

\subsubsection{Time Discretization using the Theta Method}

A fairly simple way to evaluate the precursor equation is to employ the $\theta$-scheme ($0\le\theta\le1$),
explicit when $\theta=0$, implicit when $\theta=1$, and Crank-Nicholson when $\theta=1/2$).
Generally, if there is a function $u$ whose governing equation is $\frac{du}{dt}=f(u,t)$,
then the $\theta$-discretization is

\be
\frac{u^{n+1}-u^n}{\Delta t}=(1-\theta)f(u^n,t) + \theta f(u^{n+1},t) \,.
\ee

Applying this to the precursor equation:

\be
\frac{C^{n+1}-C^n}{\Delta t}=(1-\theta)\beta S_f^np^n-(1-\theta)\lambda C^n + \theta\beta S_f^{n+1}p^{n+1}-\theta\lambda C^{n+1}
\ee

Where $S_f$ is the fission source equivalent for shape:

\be
S_f^n=(\nu\Sigma_f)^n\varphi^n
\ee

Rearranging to solve for the precursor at the end of the time step yields

\be
C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta}{1+\theta\Delta t\lambda}S_f^n p^n +  \frac{\theta\Delta t \beta}{1+\theta\Delta t\lambda}S_f^{n+1} p^{n+1}
\ee
Reporting this value of $C^{n+1}$, one can solve for the shape $\varphi^{n+1}$ as a function of $\varphi^n$ and $C^n$
(and $p^n$, $p^{n+1}$, $dp/dt|_n$ and  $dp/dt|_{n+1}$).
Once $\varphi^{n+1}$ has been determined, $C^{n+1}$ is updated. Rattlesnake currently implements both implicit and Crank-Nicholson as options for precursor evaluation.

\subsubsection{Analytical Integration}

Through prototyping, it has been found that neither implicit nor Crank-Nicholson time discretization of precursors are preferable methods for solving the shape equation in IQS.  It has been found that these discretizations result in a lack of convergence of the shape over the IQS iteration process.  In order to remedy the error, a analytical representation of the precursors was implemented in the prototype and the shape solution was able to converge (the normalization constant of the IQS method can be preserved to $10^{-10}$ while the theta-scheme only allowed convergence in the normalization factor to about $10^{-3}$).  The following section shows how this method was implemented in the prototype and the desired implementation for Rattlesnake.

Using an exponential operator, the precursor equation can be analytically solved for:

\be
\int_{t_n}^{t_{n+1}} C(t')e^{\lambda t'} dt' = \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{\lambda t'}dt'
\ee

yielding

\be
C^{n+1} =  C^n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
\ee

Because $\beta$ and $S_f$ being integrated are not known continuously over the time step, they can be interpolated linearly over the macro step.  Such that:

\be
h(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}h^n  + \frac{t-t_n}{t_{n+1}-t_n}h^{n+1}  \quad t_n \le t \le t_{n+1}
\ee
% We let $\ell_n(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}$ and  $\ell_{n+1}(t) = \frac{t-t_n}{t_{n+1}-t_n}$ for simplicity.

However, for the PRKE solve, we do have a very accurate representation of $p(t')$ over the time interval $[t_n,t_{n+1}]$.

Finally, we have the final expression for the analytical value for $C^{n+1}$:

\be
C^{n+1} = C^n e^{-\lambda \Delta t} 
+ \left(a_3\beta^{n+1}+a_2\beta^n\right)S_f^{n+1}
+ \left(a_2\beta^{n+1}+a_1\beta^n\right)S_f^n 
\ee

Where the integration coefficients are defined as:

\begin{align}
&a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)^2p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
&a_2= \int_{t_n}^{t_{n+1}}\frac{(t'-t_n)(t_{n+1}-t')}{(\Delta t)^2}p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
&a_3 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)^2p(t')e^{-\lambda(t_{n+1}-t')}dt'
\end{align}

The amplitude $(p)$ is included in the integration coefficient because it has been highly accurately calculated in the micro step scheme, so a piecewise interpolation between those points can be done to maximize accuracy.

The prototype code uses Matlab software to interpolate the amplitude between micro steps and a quadrature integration for the coefficients.  So the challenge for Rattlesnake is to replicate this procedure: passing the amplitude vector to the DNP auxkernel, interpolating it, and integrating the coefficients.

\subsection{Predictor-Corrector Modification}
In order to preserve the already implemented standard version of IQS, an option in the IQS executioner was created to specify which method is desired.  Because the diffusion solve is flux instead of shape, when predictor-corrector option is specied, the IQS removal kernel ($\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$) and the modified precursor kernel are bypassed, while all the postprocessors are still executed.  However, it is difficult to rescale the flux to shape before the PRKE parameter postprocessor are executed.  So the parameters are computed using the full flux, but amplitude is space independent and comes out of the integrals.  As seen in parameter definitions, when shape is replaced with flux, the amplitude comes out of the integral and cancels out.  So the conversion of the predicted flux in Equation \ref{eq:rescale} to shape is unnecessary if the corrected flux is solved with Equation \ref{eq:phicorr2}.  After obtaining the corrected flux, the precursors are re-evaluated using a \texttt{EXEC\_LINEAR}.
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}} p_{n+1}
\label{eq:phicorr2}
\ee

\subsection{Input}
The input deck for IQS is very similar to the current transient diffusion input file.  The IQS input has a different executioner type and parameters.  The executioner type is simply “IQS” and input parameters include number of micro time steps, IQS error tolerance, and initial power.  The Rattlesnake transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $k_{eff}$, which is present in the transient input deck.  However, IQS also requires an initial evaluation of the adjoint flux, for the weighting function.  So another input file was made for the adjoint calculation, as well as including another multi-app and transfer in the IQS input deck.

\subsection{Unintended Contributions}
The implementation of IQS in Rattlesnake put pressure on many features of Rattlesnake and MOOSE that reveled bugs and possible improvement.  Two significant issues in Rattlesnake that were found involved the adjoint solve and the diffusion fission kernel.  When testing IQS, it was found that the adjoint flux solution was not the same as the forward flux solution in a single group test, which is obviously invalid.  Also, when investigating which action to include the IQS kernel in, it turned out that the fission diffusion kernel was placed in the neutron transport action, so this kernel was demoted to the neutron diffusion actions for clarity.  Additionally, the pressure on the save in feature in MOOSE propagated its application to boundary conditions and initial solves.  MOOSE also updated its ability to restart dense vector data and to set MooseApp executioner right after executioner is created.  Merge requests: \#5474, \#5489, \#5495, and \#5497

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IQS in Other Action Systems}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For simplicity and intuitiveness the previous section only described IQS implementation in CFEM diffusion.  Rattlesnake has other action systems capable of transient simulation, where IQS can be implemented and be effective.  One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equation \ref{eq:flux} and \ref{eq:shape2}.  However, in the derivation for IQS is term is unaffected between shape and flux evaluation.  So saving the residual for this diffusion kernel in the \texttt{save\_in} variable is the only alteration to this action for IQS to function. \\

The derivation of IQS was done using multi-group diffusion, but the derivation with transport is very similar.  For simplicity, the transport equation is written in operator form in Equation \ref{eq:transport}.  Factorization is defined as $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ and the resulting shape equation is defined by Equation \ref{eq:tshape}.  Finally, the PRKE parameters are defined by Equations \ref{eq:trmb} and \ref{eq:tb}, where $\left(\Psi^{*g},f^g\right) = \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d\Omega$.
\be
\frac{1}{v^g}\frac{\partial \Psi^g}{\partial d} = \left(H^g + P_p^g - L^g\right)\Psi^g + S_d^g
\label{eq:transport}
\ee
\be
\frac{1}{v^g}\frac{\partial \psi^g}{\partial d} = \left(H^g + P_p^g - L^g - \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right)\psi^g + \frac{1}{p} S_d^g
\label{eq:tshape}
\ee
\be
\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},(H^g+P_p^g-L^g)\psi^g\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
\label{eq:trmb}
\ee
\be
\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{\sum_{g=1}^G(\Psi^{*g}, P_{d,i}^g \psi^g)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
\label{eq:tb}
\ee

The derivation of IQS in transport is simple and straight-forward, and its implementation into Rattlesnake is even more so.  The main differences between diffusion implementation in transport is outlined below:
\begin{enumerate}
\item The form of the operators in the shape equations is different, but Rattlesnake has already implemented all the kernels necessary to represent these operators.  So no change is necessary to these kernels is necessary for IQS.  Additionally, the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ is the same, so both actions can use the same kernel.
\item The PRKE parameters also change because of the operators.  For the $(\rho-\bar{\beta})$ parameter, the same post-processor can be used with the \texttt{save\_in} functionality in MOOSE.  The post-processor for $\bar{\beta}_i$ must be re-written for transport, but takes a very similar form as diffusion.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Time Adaptation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A very important aspect of IQS is how it performs with time adaptation because this will most likely be used with IQS.  The time adaptation used for quantifying IQS's ability is step doubling.  The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step and a solution with two half steps.  If the step is small enough, the error will be smaller than a user driven tolerance and the magnitude of the next step will be calculated based on the error.  If the step is too large, the step will be repeated with a smaller step calculated with the resulting error.  The error of the time step is approximated by Equation \ref{eq:edt2}, where $\varphi^g_1$ and $\varphi^g_2$ are the solutions with the full step and half step, respectively.  $\varphi$ is changed to $\phi$ for regular flux evaluation and IQS P-C.  If $e_n > e_{max}$ the time step is repeated; if $e_n < e_{max}$ the system moves to the next time step.  The next $\Delta t$ is calculated using Equation \ref{eq:dt2}, where $p$ is the convergence order of the time integration scheme being used.  $e_{max}$ and $e_{tol}$ are user defined parameters; $e_{max}$ is usually less than $e_{tol}$ to better guarantee that the calculated $\Delta t_{new}$ will pass the error criteria so time steps won't be repeated.

\be
e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_2 - \sum_{g=1}^G\varphi^g_1}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_2},\norm{\sum_{g=1}^G\varphi^g_1}\right)}
\label{eq:edt2}
\ee
\be
\Delta t_{new} = \Delta t_{old} \left(\frac{e_{tol}}{e_n}\right)
\label{eq:dt2}
\ee

To be clear, each solution undergoes Picard iterations until $Error_{IQS}$ converges before the error is calculated; for IQS P-C, each solution is re-scaled by $p$.  Additionally, if the PJFNK iteration does not converge, the entire step doubling process is repeated with half the time original time step.

\newpage
%------------------------------------------------------------------------------
%
%------------------------------------------------------------------------------
\section{RESULTS} 
\label{sect::results}

This section describes results of an examples that tests the IQS implementation and shows its effectiveness on computation speed and accuracy.  Three examples were selected for this purpose.  The first is a homogeneous one-group problem, subjected to a heterogeneous material change (absorption cross-section change as a ramp in time for a subset of the geometry).  The second is the two-dimensional TWIGL ramp transient benchmark, described further.  The final is the LRA benchmark, which a two-dimensional, temperature feedback example.  Each example was solved with regular diffusion (brute force), IQS, and IQS P-C.  Each method was tested with backward-Euler and BDF2 for the first two examples; just Crank-Nicholson was used for third example.  The performance for each method are represented by convergence plots with constant $\Delta t$ and by the number of time steps with step doubling time adaptation.

%------------------------------------------------------------------------------
\subsection{One-Dimensional Custom Example}

The example is very simple and computes quickly, it entails a one dimensional, heterogeneous 400 cm slab with a varying absorption cross section.   Figure \ref{fig:slab} how the regions of the slab are divided and Table \ref{tab:1Dmat} shows the initial material properties.  Table \ref{tab:1Dslope} shows the ramp of the absorption cross-section of each region.

\begin{figure}[!htbp]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l |}
\hline \hline \hline
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 4 & 4 & 1 & 1 & 1 & 1 \\
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
\hline \hline \hline
\end{tabular}
\caption{1-D heterogeneous slab region identification}
\label{fig:slab}
\end{center}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{1-D heterogeneous slab material properties and problem parameters}
\label{tab:1Dmat}
\begin{tabular}{lllllll}
\hline
Region & $D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu \Sigma_f (cm^{-1})$ & $v (cm/s)$ & $\beta$ & $\lambda (s^{-1})$ \\
\hline
1 & 1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\
2 & 1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\
3 & 1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\
4 & 1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[!htbp]
\begin{center}
\caption{1-D heterogeneous slab absorption cross-section slope perturbation}
\label{tab:1Dslope}
\begin{tabular}{lllllll}
\hline
Material Property & 0.0 s & 0.1 s & 0.6 s & 1.0 s & 1.7 s \\
\hline
$\Sigma_{a,2} (cm^{-1})$ & 1.1 & 1.1 & 1.095 & 1.095 & 1.095 \\
$\Sigma_{a,3} (cm^{-1})$ & 1.1 & 1.1 & 1.09 & 1.09 & 1.1 \\
$\Sigma_{a,4} (cm^{-1})$ & 1.1 & 1.1 & 1.105 & 1.105 & 1.105 \\
\hline
\end{tabular}
\end{center}
\end{table}

Figure \ref{fig:power} shows the power at each macro time step as compared to the traditional brute force (full flux time discretization) method.  The strong correlation between the two curves shows that IQS is consistent with a proven method for a highly transient example.  Figure \ref{fig:conv} shows that IQS is not only consistent for this example, but also has a better error constant in the convergence study.  Figures \ref{fig:initial} - \ref{fig:final} plots shape changes in the IQS method, showing where the shape solution is necessary and a simple PRKE evaluation is inadequate.  Figure \ref{fig:conv} shows the error convergence for brute force and IQS methods.  This plot shows that IQS has consistent convergence slope for first and second order time schemes and that IQS has a better error constant for the convergence.  Table \ref{tab:1Ddt2} shows the results for time adaptation.  These results show that traditional IQS does not perform well with step doubling time adaptation and IQS P-C performs moderately better than brute force.

\begin{figure}[!htbp]
\includegraphics[width=\linewidth]{\FiguresDir/1D_het_power_plot.png}
\caption{Power level of 1D heterogeneous example using IQS and Brute Force with $\Delta t = 0.025$}
\label{fig:power}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=5in,height=2in]{\FiguresDir/initial_flux.jpg}
\caption{Initial Flux Plot}
\label{fig:initial}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=5in,height=2in]{\FiguresDir/max_flux.jpg}
\caption{Flux Plot when Absorption Cross Section is at Minimum}
\label{fig:max}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=5in,height=2in]{\FiguresDir/final_flux.jpg}
\caption{Final Flux Computation (not steady-state)}
\label{fig:final}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=5in]{\FiguresDir/1D_het_convergence.jpg}
\caption{Error convergence comparison of 1D hetergenous example}
\label{fig:conv}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{1-D heterogeneous slab step doubling results}
\label{tab:1Ddt2}
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
\hline
  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Test & $e_{max}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
\hline
1 &	0.05 	&	0.021596 &	10  &	32	 &	0.10084 &	12	  &	175		&	0.028019 &	10	&	33 \\
2 &	0.01 	&	0.032864 &	21  &	80	 &	0.08030 &	41	  &	850	 	&	0.044721 &	20	&	74 \\
3 &	0.005 	&	0.053159 &	27  &	96	 &	0.01215 &	115	  &	2220 	&	0.052095 &	25	&	96 \\
4 & 0.001 	&	0.056546 &	56  &	188	 &	0.03242 &	848	  &	12511 	&	0.061274 &	50	&	162 \\
5 &	0.0005 	&	0.062882 &	77  &	245	 &	0.03491 &	1718  &	25841 	&	0.061240 &	69	&	225 \\
6 &	0.0001 	&	0.060089 &	177 &	537	 &	0.03554 &	8702  &	129985 	&	0.060824 &	159	&	480 \\
7 &	5.0e-05	&	0.059513 &	252 &	767	 &	0.03622 &	17282 &	256463 	&	0.061078 &	224	&	680 \\
8 &	1.0e-05	&	0.061063 &	561 &	1691 &	0.03142 &	79988 &	1104227 &	0.061901 &	501	&	1509 \\
\hline
\end{tabular}
\end{center}
\end{table}

%------------------------------------------------------------------------------
\subsection{TWIGL Benchmark}

This benchmark problem originates from the Argonne National Lab Benchmark Problem Book.  It is a 2D, 2-group reactor core model with no reflector region shown in Figure \ref{fig:TWIGL_reg}.  Table \ref{tab:TWIGL_mat} shows the material properties of each fuel region and the ramp perturbation of Material 1.

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.75\textwidth]{\FiguresDir/TWIGL_regions.jpg}
\caption{TWIGL benchmark problem description}
\label{fig:TWIGL_reg}
\end{center}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{1-D heterogeneous slab absorption cross-section slope perturbation}
\label{tab:TWIGL_mat}
\begin{tabular}{llllllll}
\hline
  &  &  &  &  &  &  \multicolumn{2}{c}{$\underline{\Sigma_s (cm^{-1})} $} \\
Material & Group & $D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu\Sigma_f (cm^{-1})$ & $\chi$ & $g \rightarrow 1$ & $g \rightarrow 2$ \\
\hline
1 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01 \\
  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
2 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01  \\
  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
3 & 1 & 1.3 & 0.008 & 0.003 & 1.0 & 0.0 & 0.01  \\
  & 2 & 0.5 & 0.050 & 0.060 & 0.0 & 0.0 & 0.00  \\
\hline
  & $\nu$ & $v_1 (cm/s)$ & $v_2 (cm/s)$ & $\beta$ & $\lambda (1/s)$ &   &   \\
\hline
  & 2.43 & 1.0E7 & 2.0E5 & 0.0075 & 0.08 &   &   \\
\hline
 \multicolumn{8}{l}{\footnotesize Material 1 ramp perturbation:} \\
\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (1-0.11667t) \quad t \leq 0.2 s$} \\
\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (0.97666t) \quad t > 0.2 s$} \normalsize
\end{tabular}
\end{center}
\end{table}

Figures \ref{fig:TWIGL_power} and \ref{fig:TWIGL_plots} show the IQS  solution as compared with the Brute Force solution.  It is important to note the IQS shape plot is scaled differently than the Brute Force flux plot because the amplitude term is not included, but the gradients of colors is comparable.  These plots show that IQS is consistent in more complex, higher dimensional problems in RATTLESNAKE.  Finally, Figure \ref{fig:TWIGL_conv} plots the error convergence of IQS and the Brute Force methods.  The curves show the impressive convergence of IQS for the highly transient TWIGL example.  Table \ref{tab:TWIGLdt2} shows the results for time adaptation.  The results show that both IQS methods perform exceptionally well compared to brute force.  It also shows that traditional IQS performed better with large $e_{max}$, while IQS P-C was better with smaller $e_{max}$.

\begin{figure}[!htbp]
\centering
\includegraphics[height=3.0in]{\FiguresDir/TWIGL_power_plot.png}
\caption{Power level comarison of 1D heterogeneous example between IQS and Brute Force using $\Delta t = 0.004$}
\label{fig:TWIGL_power}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\begin{subfigure}[!htbp]{0.4\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/ndiff_ramp_flux.jpg}
\caption{Brute force flux}
\end{subfigure}
\quad
\begin{subfigure}[!htbp]{0.4\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/iqs_ramp_shape.jpg}
\caption{IQS Shape}
\end{subfigure}
\caption{TWIGL Benchmark flux/shape comparison at $t=0.2$}
\label{fig:TWIGL_plots}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=5in]{\FiguresDir/TWIGL_convergence.jpg}
\caption{Error convergence comparison of TWIGL Benchmark}
\label{fig:TWIGL_conv}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{TWIGL step doubling results}
\label{tab:TWIGLdt2}
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
\hline
  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Test & $e_{max}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
\hline
1 &	0.05    &	0.00012677 &	9   &	29  &	0.03380433 &	4   &	20   &	0.00323100 &	4  &	9   \\
2 &	0.01    &	3.5555e-05 &	11  &	35  &	0.00166991 &	5   &	40   &	0.00263068 &	5  &	12  \\
3 &	0.005   &	4.0364e-05 &	11  &	31  &	0.00886584 &	5   &	40   &	0.00160486 &	6  &	21  \\
4 &	0.001   &	0.00294822 &	33  &	122 &	0.02976305 &	5   &	36   &	1.7527e-05 &	10 &	35  \\
5 &	0.0005  &	0.00099778 &	39  &	131 &	0.00143781 &	6   &	55   &	1.4185e-05 &	16 &	74  \\
6 &	0.0001  &	0.00019510 &	78  &	236 &	0.00016175 &	7   &	65   &	6.2903e-06 &	19 &	78  \\
7 &	5.0e-05 &	0.00018372 &	112 &	342 &	6.0328e-05 &	12  &	163  &	1.5247e-06 &	24 &	92  \\
8 &	1.0e-05 &	8.0564e-05 &	263 &	794 &	7.7103e-05 &	379 &	5729 &	9.8321e-07 &	48 &	210 \\
\hline

\end{tabular}
\end{center}
\end{table}

\newpage

%------------------------------------------------------------------------------
\subsection{LRA Benchmark}

The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor.  It is a super prompt-critical transient.  To have better understanding on the cross sections given later, we present the equations here:
\begin{align}
-\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\grad D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 - \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
-\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\grad D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
S_f &= \sum_{g=1}^2 \Sigma_{f,g} \phi_g, \\
\frac{\partial C_i}{\partial t} &= \nu\beta_i f - \lambda_i C_i, \quad i=1,2, \\
\frac{\partial T}{\partial t} &= \alpha f, \label{eq:lra-temp} \\
\Sigma_{a,1} &= \Sigma_{a,1}(\vec{r}, t=0) \left[1+\gamma\left(\sqrt{T} - \sqrt{T_0}\right)\right], \\
P &= \kappa S_f,
\end{align}
where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_6f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0=T(\vec{r}, t=0)$.
The two-group diffusion equation are solved with zero flux boundary conditions on external surfaces, reflecting conditions at symmetry boundaries and steady state initial conditions which are obtained by solving
\begin{align}
-\grad D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
-\grad D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1. \\
\end{align}
The eigenvalue $k$ is used to modify the fission cross section for the transient simulations with $\frac{1}{k}\Sigma_{f,g}, g=1,2$.  The initial flux distribution shall be normalized such that the averaged power density
\begin{align}
\bar{P} \equiv \frac{\int_{V_{core}} P(\vec{r}, t=0) d\vec{r}}{\int_{V_{core}} d\vec{r}},
\end{align}
where $V_{core}$ is the core region with fuels, is equal to $10^{-6} W\cdot cm^{-3}$.
The initial precursor concentrations are in equilibrium with the initial critical flux distribution.\\

The geometry is illustrated in~\fig{fig:lra-geometry}.\\
\begin{figure}[H]
\centering
\includegraphics[width=0.8\linewidth]{\FiguresDir/lra.png}
\caption{LRA benchmark geometry with region assignment.}
\label{fig:lra-geometry}
\end{figure}

Initial two-group constants are presented in~\tbl{tab:lra-xs}.
$\nu$ is equal to 2.43.
Axial bulking $B^2 = 10^{-4}$ is applied for both energy groups.
Delayed neutron data are presented in~\tbl{tab:lra-dnp}.
All fuel materials have the same delayed neutron data.
Some scalar data are listed in~\tbl{tab:lra-scalar}.\\
\begin{table}[H]
    \centering
    \caption{LRA benchmark initial two-group constants.\label{tab:lra-xs}}
      \begin{tabular}{|c|c|c|c|l|l|l|l|l|}
      \hline
             &                 & Group   & $D_g$      & $\Sigma_{a,g}$ & $\nu\Sigma_{f,g}$ & $\Sigma_{s,1\rightarrow 2}$ & $\chi_g$ & $v_g$              \\
      Region & Material        &       g & (cm)       &  ($cm^{-1}$)   &  ($cm^{-1}$)      &  ($cm^{-1}$)                &          & ($cm\cdot s^{-1}$) \\
      \hline
      1      & Fuel 1 with rod & 1       & 1.255      & 0.008252       & 0.004602          &                             & 1        & $3.0\times10^7$    \\
             &                 & 2       & 0.211      & 0.1003         & 0.1091            & 0.02533                     & 0        & $3.0\times10^5$    \\
      \hline
      2      & Fuel 1 without rod & 1    & 1.268      & 0.007181       & 0.004609          &                             & 1        & $3.0\times10^7$    \\
             &                    & 2    & 0.1902     & 0.07047        & 0.08675           & 0.02767                     & 0        & $3.0\times10^5$    \\
      \hline
      3      & Fuel 2 with rod & 1       & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
             &                 & 2       & 0.2091     & 0.08344        & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
      \hline
      4      & Fuel 2 without rod & 1    & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
             &                    & 2    & 0.2091     & 0.073324       & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
      \hline
      5      & Reflector        & 1      & 1.257      & 0.0006034      & -                 &                             & -        & $3.0\times10^7$    \\
             &                  & 2      & 0.1592     & 0.01911        & -                 & 0.04754                     & -        & $3.0\times10^5$    \\
      \hline
      \end{tabular}
\end{table}

\begin{table}[H]
    \centering
    \caption{LRA benchmark delayed neutron data.\label{tab:lra-dnp}}
      \begin{tabular}{|c|l|l|l|l|}
      \hline
      Group i & $\beta_i$ & $\lambda_i$ ($s^{-1}$) & $\chi_{d,i,1}$ & $\chi_{d,i,2}$ \\
      \hline
      1       & 0.0054    & 0.0654  & 1 & 0 \\
      2       & 0.001087  & 1.35    & 1 & 0 \\
      \hline
      \end{tabular}
\end{table}

\begin{table}[H]
    \centering
    \caption{LRA benchmark scalar values.\label{tab:lra-scalar}}
      \begin{tabular}{|l|c|l|}
      \hline
      Meaning & Notation & value \\
      \hline
      Axial buckling for both energy groups & $B_g^2$  & $10^{-4}$ ($cm^{-2}$)\\
      Mean number of neutrons per fission   & $\nu$    & 2.43 \\
      Conversion factor                     & $\alpha$ & $3.83\times 10^{-11}$ ($K\cdot cm^{3}$) \\
      Feedback constant                     & $\gamma$ & $3.034\times 10^{-3}$ ($K^{1/2}$) \\
      Energy released per fission           & $\kappa$ & $3.204\times 10^{-11} $ ($J/fission$) \\
      Initial and reference temperature     & $T_0$    & 300 (K) \\
      Active core volume                    & $V_{core}$ & 17550 ($cm^2$)\\
      \hline
      \end{tabular}
\end{table}

The transient is initiated by changing the thermal absorption cross section as the following:
\begin{align}
\Sigma_{a,2}(t) = \Sigma_{a,2}(t=0) \left\{\begin{array}{lr} 1-0.0606184t, & t\leq 2 \\
                                                             0.8787631, & t>2
                                           \end{array}\right.
\end{align}
where $t$ is time in seconds.\\

Figure \ref{fig:LRA_plots} show the IQS  solution as compared with the Brute Force solution.    These plots show that IQS is consistent with Doppler feedback problems in RATTLESNAKE.  Finally, Figure \ref{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods.

\begin{figure}[!htbp]
\begin{center}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/lra_power_profile.jpg}
\caption{Power Profile}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/lra_temp_profile.jpg}
\caption{Temperature Profile}
\end{subfigure}
\caption{LRA Benchmark power and temperature comparison}
\label{fig:LRA_plots}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=5in]{\FiguresDir/lra_convergence.jpg}
\caption{Error convergence comparison of LRA Benchmark}
\label{fig:lra_conv}
\end{figure}

\end{document}

### 2016-06-07 doc/NEAMS_Report/NEAMS_Report.tex
removed blank space from folder name

--- 2016-06-07 doc/NEAMS Report/NEAMS_Report_V2.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[12pt]{scrartcl} 
-%########################### Preferences #################################
+\documentclass[10pt]{scrartcl}
 
@@
 \newcommand{\adj}{\Phi^\dagger_0}
-\newcommand{\tcr}[1]{\textcolor{red}{#1}}
-\newcommand{\tcb}[1]{\textcolor{blue}{#1}}
 
@@
 \setlength{\textwidth}{6.5in}
-% ********* Header and Footer **********
-% This is something to play with forever. I use here the advanced settings of the KOMA script
 
-\title{Improved Quasi-Static Methods in RATTLESNAKE}
-\subtitle{Development and Implementation of Improved Quasi-Static (IQS) methods for time-dependent neutron diffusion and neutron transport solvers in RATTLESNAKE}
+\title{Improved Quasi-Static Method}
+\subtitle{IQS Method Implemention for CFEM Diffusion in Rattlesnake}
 \author{ \normalsize
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% boxes
-\usepackage{color}
-\definecolor{myblue}{rgb}{.8, .8, 1}
-\usepackage{empheq}
-
-\newlength\mytemplen
-\newsavebox\mytempbox
-
-\makeatletter
-\newcommand\mybluebox{%
-    \@ifnextchar[%]
-       {\@mybluebox}%
-       {\@mybluebox[0pt]}}
-
-\def\@mybluebox[#1]{%
-    \@ifnextchar[%]
-       {\@@mybluebox[#1]}%
-       {\@@mybluebox[#1][0pt]}}
-
-\def\@@mybluebox[#1][#2]#3{
-    \sbox\mytempbox{#3}%
-    \mytemplen\ht\mytempbox
-    \advance\mytemplen #1\relax
-    \ht\mytempbox\mytemplen
-    \mytemplen\dp\mytempbox
-    \advance\mytemplen #2\relax
-    \dp\mytempbox\mytemplen
-    \colorbox{myblue}{\hspace{1em}\usebox{\mytempbox}\hspace{1em}}}
-
-\makeatother
-
-%################ End Preferences, Begin Document #####################
-
-\pagestyle{plain} % on headers or footers on the first page
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% set relative directory for figures. can be changed at any time later with \renewcommand
-\newcommand{\FiguresDir}{./figs}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \maketitle
 \pagenumbering{arabic}
-
-\tcb{I will fix the cover page and page footers later}
-
-\tcb{I removed CFEM from the title}
-
-\tcr{you need Yaqi's concurrence since you list him as co-author}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{Abstract}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\tcr{we need a short abstract}
-
-\pagebreak
 
@@
 \par
-The improved quasi-static (IQS) method \tcr{you need to give 3-4 references} is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (the shape) and a time-dependent component (the amplitude).
-The technique relies on the shape being less rapidly varying in time compared to the flux, hence
-requiring fewer shape computations or updates. 
-The IQS method has primarily been developped in the context of the neutron diffusion approximation \tcr{give refs}, but \tcr{cite sedat} provides an extension of the technique to transport, implemented in TDTORT.
-The method has mostly being applied to neutron kinetics, without thermal-hydraulic feedback. In this report, we extend the IQS
-method in the context of transient multiphysics simulations.
-%The method These components include the flux's power and shape.
+The improved quasi-static (IQS) method is a transient spatial kinetics method that involves factorizing flux into space- and time-dependent components.
+These components include the flux's power and shape.
 Power is time-dependent, while the shape is both space- and time-dependent.
@@
 
-In this Section, we recall the equations for the IQS method, using the standard multigroup diffusion equations written below:
+In this Section, we recall the equations for the IQS method, starting from the standard multigroup diffusion equations written below:
 \begin{align}
@@
 \be
-\frac{dC_i}{dt} = \frac{\beta_i}{\keff}\sum_{g=1}^G\nu^{g} \Sigma_f^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
+\frac{dC_i}{dt} = \frac{\beta_i}{k_{eff}}\sum_{g=1}^G\nu^{g} \Sigma_f^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
 \label{eq:precursor}
@@
 \ee
-(extension to the transport equations will be provided subsequently, once a concise operator notation has been introduced). 
-In the above equations, \tcr{give a long text naming each variables, for example: $\phi^g$ is the neutron flux in group $g$, ....}.
-Note that the prompt fission terms have been normalized by $\keff$, the eigenvalue of the initial state, as transient simulations are often carried out for initially critical reactors.
-The expressions for subcritical and source-driven problems are straightforward to derive but omitted in this report for brevity.
 
-The factorization approach in IQS leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent multigroup shape ($\varphi^g$):
+Factorization is the most important step in the derivation of the IQS method.
+The factorization approach leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent multigroup shape ($\varphi^g$):
 \be
@@
 \ee
-The factorization is not an approximation. Reporting it in the flux and precursor equations yields the following equations
-\begin{enumerate}
-\item for the shape:
+Then the flux and precursor equations become:
 \begin{align}
@@
 \begin{figure}[!htbp]
-\includegraphics[width=\linewidth]{\FiguresDir/IQS_visualization.jpg}
+\includegraphics[width=\linewidth]{IQS_visualization.jpg}
 \caption{IQS method visualization}
@@
 \begin{figure}[!htbp]
-\includegraphics[width=\linewidth]{\FiguresDir/1D_het_power_plot.png}
+\includegraphics[width=\linewidth]{1D_het_power_plot.png}
 \caption{Power level of 1D heterogeneous example using IQS and Brute Force with $\Delta t = 0.025$}
@@
 \begin{center}
-\includegraphics[width=5in,height=2in]{\FiguresDir/initial_flux.jpg}
+\includegraphics[width=5in,height=2in]{initial_flux.jpg}
 \caption{Initial Flux Plot}
@@
 \begin{center}
-\includegraphics[width=5in,height=2in]{\FiguresDir/max_flux.jpg}
+\includegraphics[width=5in,height=2in]{max_flux.jpg}
 \caption{Flux Plot when Absorption Cross Section is at Minimum}
@@
 \begin{center}
-\includegraphics[width=5in,height=2in]{\FiguresDir/final_flux.jpg}
+\includegraphics[width=5in,height=2in]{final_flux.jpg}
 \caption{Final Flux Computation (not steady-state)}
@@
 \centering
-\includegraphics[width=5in]{\FiguresDir/1D_het_convergence.jpg}
+\includegraphics[width=5in]{1D_het_convergence.jpg}
 \caption{Error convergence comparison of 1D hetergenous example}
@@
 \begin{center}
-\includegraphics[width=0.75\textwidth]{\FiguresDir/TWIGL_regions.jpg}
+\includegraphics[width=0.75\textwidth]{TWIGL_regions.jpg}
 \caption{TWIGL benchmark problem description}
@@
 \centering
-\includegraphics[height=3.0in]{\FiguresDir/TWIGL_power_plot.png}
+\includegraphics[height=3.0in]{TWIGL_power_plot.png}
 \caption{Power level comarison of 1D heterogeneous example between IQS and Brute Force using $\Delta t = 0.004$}
@@
 \begin{subfigure}[!htbp]{0.4\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/ndiff_ramp_flux.jpg}
+\includegraphics[width=\textwidth]{ndiff_ramp_flux.jpg}
 \caption{Brute force flux}
@@
 \begin{subfigure}[!htbp]{0.4\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/iqs_ramp_shape.jpg}
+\includegraphics[width=\textwidth]{iqs_ramp_shape.jpg}
 \caption{IQS Shape}
@@
 \centering
-\includegraphics[width=5in]{\FiguresDir/TWIGL_convergence.jpg}
+\includegraphics[width=5in]{TWIGL_convergence.jpg}
 \caption{Error convergence comparison of TWIGL Benchmark}
@@
 \centering
-\includegraphics[width=0.8\linewidth]{\FiguresDir/lra.png}
+\includegraphics[width=0.8\linewidth]{lra.png}
 \caption{LRA benchmark geometry with region assignment.}
@@
 \begin{subfigure}[!htbp]{0.49\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/lra_power_profile.jpg}
+\includegraphics[width=\textwidth]{lra_power_profile.jpg}
 \caption{Power Profile}
@@
 \begin{subfigure}[!htbp]{0.49\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/lra_temp_profile.jpg}
+\includegraphics[width=\textwidth]{lra_temp_profile.jpg}
 \caption{Temperature Profile}
@@
 \centering
-\includegraphics[width=5in]{\FiguresDir/lra_convergence.jpg}
+\includegraphics[width=5in]{lra_convergence.jpg}
 \caption{Error convergence comparison of LRA Benchmark}

### 2016-06-07 doc/NEAMS_Report/NEAMS_Report.tex
working on theory part of report

--- 2016-06-07 doc/NEAMS Report/NEAMS_Report_V2.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Overview}
+\section{Theory}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Power is time-dependent, while the shape is both space- and time-dependent.
-However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the shape may not require computation at every time step, invoking the “quasi-static” nature.
-\\
+However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the shape may not require computation at every time step, invoking the Աuasi-staticԠnature.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{IQS for Multigroup Diffusion Equations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 In this Section, we recall the equations for the IQS method, using the standard multigroup diffusion equations written below:
+\begin{subequations}
 \begin{align}
@@
 \ee
-with
-\be
+\end{subequations}
+%
+Extension to the transport equations will be provided subsequently, once a concise operator notation has been introduced. 
+In the above equations, \tcr{give a long text naming each variables, for example: $\phi^g$ is the neutron flux in group $g$, the removal, fission, scattering cross sections are denoted by ...; the total delayed neutron fraction is
+\[
 \beta = \sum_{i=1}^I \beta_{i} 
-\ee
-(extension to the transport equations will be provided subsequently, once a concise operator notation has been introduced). 
-In the above equations, \tcr{give a long text naming each variables, for example: $\phi^g$ is the neutron flux in group $g$, ....}.
+\]
+}
 Note that the prompt fission terms have been normalized by $\keff$, the eigenvalue of the initial state, as transient simulations are often carried out for initially critical reactors.
@@
 \ee
-The factorization is not an approximation. Reporting it in the flux and precursor equations yields the following equations
-\begin{enumerate}
-\item for the shape:
+The factorization is not an approximation. Reporting it in the flux and precursor equations yields, after some simple algebra, the following equations for the shape and the precursors:
+\begin{subequations}
 \begin{align}
-\frac{1}{v^g}\left(\frac{dp}{dt}\varphi^g + p\frac{\partial \varphi^g}{\partial t}\right) =&\frac{\chi_p^g}{k_{eff}}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} p \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right)p \varphi^g \nonumber \\
-&  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} p \varphi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
-\label{eq:shape}
+\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\ 
+& -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
+\label{eq:shape2}
 \end{align}
 \be
-\frac{dC_i}{dt} = \frac{\beta_i}{k_{eff}}\sum_{g=1}^G\nu^{g} \Sigma_f^g p \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
-\label{eq:prec}
+\frac{dC_i}{dt} = \frac{\beta_i}{\keff}p \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
+\label{eq:preq2}
 \ee
+\end{subequations}
 
-PRKE formulation (amplitude equations):
+We note that the time-dependent shape equation is similar to the time-dependent flux, with the following modifications:
+\begin{enumerate}
+\item the shape equation contains an additional term equivalent to a removal cross section,  $ \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$,
+\item the delayed neutron source term is divided by $p$,
+\item the system of equations is now nonlinear due to the factorization,
+\item an equation is needed to obtain the amplitude $p$.
+\end{enumerate}
 
-\begin{align}
-\sum_{g=1}^G\int_D &\left[\frac{1}{v^g}\left(\frac{dp}{dt}\varphi^g + p\frac{\partial \varphi^g}{\partial t}\right)\right]w^gd^3r =  \nonumber \\
-& \sum_{g=1}^G\int_D\left[\frac{\chi_p^g}{k_{eff}} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} p \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right)p \varphi^g\right]w^gd^3r   \\ 
- +&\sum_{g=1}^G\int_D \left[\sum_{g'\neq g}^G\Sigma_s^{g'\to g} p \varphi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i\right]w^gd^3r \ , \quad 1 \le g \le G  \nonumber
-\end{align}
-\be 
-\int_D\frac{dC_i}{dt}\sum_{g=1}^G\chi_{d,i}^gw^gd^3r = \int_D \frac{\beta_i}{k_{eff}}\left[\sum_{g=1}^G\left(\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'} - \lambda_i C_i \right)\chi_{d,i}^g w^g\right]d^3r \ , \quad 1 \le i \le I
-\ee
-
-It can be shown that the most appropriate weighting function $(w^g)$ is the initial adjoint flux $(\phi^{*g})$.
-For brevity, the following definition will be applied: $\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r=\left(\phi^{*g},f\right)$
-\begin{align}
-& \quad \frac{dp}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)+ p\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)= \nonumber \\
-& p\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{k_{eff}}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)  \\
-& + \sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i) \nonumber
-\end{align}
-\be
-\frac{d}{dt}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)=\frac{1}{k_{eff}}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})p - \sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \quad 1 \le i \le I
-\ee
-
-In order to impose uniqueness of the factorization, one imposes:
-\be
-\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=constant
-\ee
-
-Therefore the PRKE formulation reduces to:
-\begin{align}
-\frac{dp}{dt}= &\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{k_{eff}}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g}  \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p \nonumber \\
-& +\sum_{i=1}^I\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i
-\end{align}
-\be
-\frac{d\xi_i}{dt}=\frac{1}{k_{eff}}\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p - \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i \quad 1 \le i \le I
-\ee
-Where:
-\be
-\xi_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g C_i)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-\ee
-
-It is convenient to define the effective reactivity, delay-neutron fraction, and delayed-neutron precursor decay constant:
-\be
-\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{k_{eff}}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-\label{eq:rmb}
-\ee
-\be
-\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{k_{eff}}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-\label{eq:b}
-\ee
-\be
-\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
-\label{eq:l}
-\ee
-
-So:
+To derive the amplitude equation (also known as the PRKE equations), the shape/precursors equations are weighed by a space-dependent function (typically the initial adjoint flux, $\phi^{*g}$) and integrated over the phase-space.
+The final expressions are given below (the derivations are given in appendix for the interested reader).
+\begin{subequations}
 \be
@@
 \ee
-
-Equations (\ref{eq:p}) and (\ref{eq:c}) are a formulation of the point reactor kinetics equation (PRKE), but the parameters (Equations (\ref{eq:rmb})-(\ref{eq:l})) are dependent on the shape.
-If the assumption is made that the shape is time-independent, the shape is computed once at the first time step and used for the PRKE parameter evaluation at all other steps.
-However, if the shape is dependent on time, the shape needs to be computed in transient using equation (\ref{eq:shape}) and (\ref{eq:prec}) in order retain accuracy.
-Equations (\ref{eq:shape2}) and (\ref{eq:preq2}) shows the usual form of the shape and precursor equations with amplitude put on the right hand side.
-Equation (\ref{eq:shape2}) is very similar to the multigroup flux equation (\ref{eq:flux}), except the removal cross-section term is augmented by a $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term and the precursor contribution has a $\frac{1}{p}$ mulitiplier.
-Equation (\ref{eq:preq2}) is very similar to the normal precursor equation (\ref{eq:precursor}), except the fission source term is multiplied by $p$.
-These differences are crucial for IQS implementation in Rattlesnake.
-
-\begin{align}
-\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\ 
-& -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
-\label{eq:shape2}
-\end{align}
+\end{subequations}
+where we haved define the reactivity, effective delayed-neutron fraction, and delayed-neutron precursor decay constant as follows:
+\begin{subequations}
 \be
-\frac{dC_i}{dt} = \frac{\beta_i}{k_{eff}}p \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
-\label{eq:preq2}
+\frac{\rho-\bar{\beta}}{\Lambda} = 
+\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}
+{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
+\label{eq:rmb}
 \ee
-
-%\begin{align}
-%& \quad \sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}\right)= \nonumber \\ 
-%&\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g\right) \\
-%&+\frac{1}{p}\sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \nonumber
-%\end{align}
+\be
+\frac{\bar{\beta}}{\Lambda} = \sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda} = 
+\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}
+{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
+\label{eq:b}
+\ee
+\be
+\bar{\lambda}_i = \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
+\label{eq:l}
+\ee 
+\end{subequations}
+The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.
 
@@
 The PRKE is computed with a user specified option of backward-Euler, Crank-Nicolson, or SDIRK33.
-The IQS executioner also supplements Transient’s Picard iteration process by adding its own error criteria:
+The IQS executioner also supplements Transient Picard iteration process by adding its own error criteria:
 \be
@@
 Instead, one may discretize in time the shape equation, which typically requires the knowledge of the
-precursor concentrations at the end of the time step. This precursor value is taken from the solution, numerical or anlaytical,
+precursor concentrations at the end of the time step. This precursor value is taken from the solution, numerical or analytical,
 of the precursors ODE.
@@
 The IQS input has a different executioner type and parameters.
-The executioner type is simply “IQS” and input parameters include number of micro time steps, IQS error tolerance, and initial power.
-The Rattlesnake transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $k_{eff}$, which is present in the transient input deck.
+The executioner type is simply ԉQSԠand input parameters include number of micro time steps, IQS error tolerance, and initial power.
+The Rattlesnake transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $\keff$, which is present in the transient input deck.
 However, IQS also requires an initial evaluation of the adjoint flux, for the weighting function.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\appendix
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\tcr{putting stuff here for now}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+It can be shown that the most appropriate weighting function $(w^g)$ is the initial adjoint flux $(\phi^{*g})$.
+For brevity, the following definition will be applied: $\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r=\left(\phi^{*g},f\right)$
+\begin{align}
+& \quad \frac{dp}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)+ p\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)= \nonumber \\
+& p\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)  \\
+& + \sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i) \nonumber
+\end{align}
+\be
+\frac{d}{dt}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)=\frac{1}{\keff}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})p - \sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \quad 1 \le i \le I
+\ee
+
+In order to impose uniqueness of the factorization, one imposes:
+\be
+\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=constant
+\ee
+
+Therefore the PRKE formulation reduces to:
+\begin{align}
+\frac{dp}{dt}= &\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g}  \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p \nonumber \\
+& +\sum_{i=1}^I\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i
+\end{align}
+\be
+\frac{d\xi_i}{dt}=\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p - \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i \quad 1 \le i \le I
+\ee
+Where:
+\be
+\xi_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g C_i)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
+\ee
+
+It is convenient to define the effective reactivity, delay-neutron fraction, and delayed-neutron precursor decay constant:
+\be
+\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
+\label{eq:rmb}
+\ee
+\be
+\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
+\label{eq:b}
+\ee
+\be
+\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
+\label{eq:l}
+\ee
+
+So:
+\be
+\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
+\label{eq:p}
+\ee
+\be
+\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
+\label{eq:c}
+\ee
+
+Equations (\ref{eq:p}) and (\ref{eq:c}) are a formulation of the point reactor kinetics equation (PRKE), but the parameters (Equations (\ref{eq:rmb})-(\ref{eq:l})) are dependent on the shape.
+If the assumption is made that the shape is time-independent, the shape is computed once at the first time step and used for the PRKE parameter evaluation at all other steps.
+However, if the shape is dependent on time, the shape needs to be computed in transient using equation (\ref{eq:shape}) and (\ref{eq:prec}) in order retain accuracy.
+Equations (\ref{eq:shape2}) and (\ref{eq:preq2}) shows the usual form of the shape and precursor equations with amplitude put on the right hand side.
+Equation (\ref{eq:shape2}) is very similar to the multigroup flux equation (\ref{eq:flux}), except the removal cross-section term is augmented by a $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term and the precursor contribution has a $\frac{1}{p}$ multiplier.
+Equation (\ref{eq:preq2}) is very similar to the normal precursor equation (\ref{eq:precursor}), except the fission source term is multiplied by $p$.
+These differences are crucial for IQS implementation in Rattlesnake.
+
+\begin{align}
+\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\ 
+& -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
+\label{eq:shape2}
+\end{align}
+\be
+\frac{dC_i}{dt} = \frac{\beta_i}{\keff}p \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
+\label{eq:preq2}
+\ee
+
+%\begin{align}
+%& \quad \sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}\right)= \nonumber \\ 
+%&\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g\right) \\
+%&+\frac{1}{p}\sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \nonumber
+%\end{align}
+
 \end{document}

### 2016-06-07 doc/NEAMS_Report/NEAMS_Report.tex
lots of changes

@@
 
-Computing this shape can become expensive, especially in two or three dimensions.
-Subsequently, it is attractive to make the assumption that the shape is weakly time-dependent so the shape can be computed after a multitude of PRKE calculations which is the root of IQS.
-To visualize:
-\begin{figure}[!htbp]
-\includegraphics[width=\linewidth]{\FiguresDir/IQS_visualization.jpg}
-\caption{IQS method visualization}
-\end{figure}
-
-Additionally, to improve consistency and accuracy, each macro time step can be iterated so the best shape is used to compute power at the micro time steps.
-This iteration process must converge the shape such that the uniqueness condition $(\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=0)$ is preserved.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{IQS Predictor-Corrector (IQS P-C)}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The Predictor-Corrector version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the evaluation of the coupled system is different.
-This version first solves the flux diffusion (represented by Equations \ref{eq:flux} and \ref{eq:precursor}) to get a predicted flux.
-The predicted flux at this step is then converted to shape by rescaling as follows:
-\be
-\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
-\label{eq:rescale}
-\ee
-Where:
-\be
-K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
-\ee
-\be
-K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
-\ee
-
-The PRKE parameters are then computed with this shape using Equations \ref{eq:rmb} - \ref{eq:l} and interpolated over the macro step, then the PRKE is evaluated.
-With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
-\be
-\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1}
-\ee
-
-The advantage to the predictor-corrector method is there is no iteration necessary for this method and in turn is much simpler and faster than the standard IQS.
-The disadvantage is that the method assumes $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)$ is inherently constant.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Rattlesnake Implementation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This section will explain how, thus far, IQS has been implemented in Rattlesnake.
-
-\subsection{Executioner}
-The IQS executioner derives from the Transient executioner in MOOSE.
-The IQS executioner contains a loop over micro time steps that computes the PRKE and then passes $p$ and $\frac{dp}{dt}$ for the Transient executioner to evaluate the shape equation at each macro step.
-The PRKE is computed with a user specified option of backward-Euler, Crank-Nicolson, or SDIRK33.
-The IQS executioner also supplements Transient Picard iteration process by adding its own error criteria:
-\be
-Error_{IQS}=\left|\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,n}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,0}\right)}-1\right|
-\ee
-
-\subsection{Action System}
-IQS defines its uniqueness from its executioner type; however, many changes needed to be made in the Rattlesnake action system in order to support IQS execution.
-First, changes needed to be made in order to evaluate the shape equation.
-The shape equation, after some manipulation, is very similar to the time-dependent, which Rattlesnake is already set up to solve:
-
-\begin{align}
-\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
-& - \underbrace{\frac{1}{v^g} \overbrace{\frac{1}{p}\frac{dp}{dt}}^{From Executioner}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
-\end{align}
-
-To enable Rattlesnake to solve this equation, another kernel was created that evaluated $\sum_{g=1}^G\frac{1}{v^gp}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
-Second, four postprocessors were created in order to calculate the PRKE parameters.
-The parameter calculations were separated by $\frac{\bar{\beta}_i}{\Lambda}$ numerator, $\bar{\lambda}_i$ numerator/denominator, $\frac{\rho}{\Lambda}/\frac{\bar{\beta}}{\Lambda}$ denominator, and $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator.
-The first three are relatively simple, only relying on material properties and solution quantities.
-The $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator requires the use of the MOOSE save in feature, which saves the residual from a calculated kernel or boundary contribution in the shape evaluation to an auxiliary variable.
-Finally, a user object was created to pull together all the postprocessor values and carryout the numerator/denominator divisions that were then passed to the executioner.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Precursor Integration}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-This section presents two different time-integration methods to solve coupled IQS shape + precursor equations, recalled below
-using, for simplicity, a single neutron group and a single precursor group.
-
-\be
-\frac{1}{v}\frac{\partial\varphi}{\partial t}=\nu\Sigma_f(1-\beta)\varphi-\left(-\div D \grad + \Sigma_a + \frac{1}{v}\frac{1}{p}\frac{dp}{dt}\right)\varphi+\frac{1}{p}\lambda C 
-\ee
-\be
-\frac{dC}{dt} = \beta\nu \Sigma_f \varphi p - \lambda C
-\ee
-
-First, we note that we could keep this system of two time-dependent equations and solve it as a coupled system. 
-However, this is unnecessary and a memory expensive endeavor because the precursor equation is only an ODE and not a PDE. 
-Instead, one may discretize in time the shape equation, which typically requires the knowledge of the
-precursor concentrations at the end of the time step. This precursor value is taken from the solution, numerical or analytical,
-of the precursors ODE.
-This document will discuss two techniques for solving the precursor equation.
-First is a time discretization method that is currently being implemented in Rattlesnake.
-The second is a analytical integration of the precursors, the latter method has proven to be more beneficial for IQS convergence.
-
-\subsubsection{Time Discretization using the Theta Method}
-
-A fairly simple way to evaluate the precursor equation is to employ the $\theta$-scheme ($0\le\theta\le1$),
-explicit when $\theta=0$, implicit when $\theta=1$, and Crank-Nicholson when $\theta=1/2$).
-Generally, if there is a function $u$ whose governing equation is $\frac{du}{dt}=f(u,t)$,
-then the $\theta$-discretization is
-
-\be
-\frac{u^{n+1}-u^n}{\Delta t}=(1-\theta)f(u^n,t) + \theta f(u^{n+1},t) \,.
-\ee
-
-Applying this to the precursor equation:
-
-\be
-\frac{C^{n+1}-C^n}{\Delta t}=(1-\theta)\beta S_f^np^n-(1-\theta)\lambda C^n + \theta\beta S_f^{n+1}p^{n+1}-\theta\lambda C^{n+1}
-\ee
-
-Where $S_f$ is the fission source equivalent for shape:
-
-\be
-S_f^n=(\nu\Sigma_f)^n\varphi^n
-\ee
-
-Rearranging to solve for the precursor at the end of the time step yields
-
-\be
-C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta}{1+\theta\Delta t\lambda}S_f^n p^n +  \frac{\theta\Delta t \beta}{1+\theta\Delta t\lambda}S_f^{n+1} p^{n+1}
-\ee
-Reporting this value of $C^{n+1}$, one can solve for the shape $\varphi^{n+1}$ as a function of $\varphi^n$ and $C^n$
-(and $p^n$, $p^{n+1}$, $dp/dt|_n$ and  $dp/dt|_{n+1}$).
-Once $\varphi^{n+1}$ has been determined, $C^{n+1}$ is updated. Rattlesnake currently implements both implicit and Crank-Nicholson as options for precursor evaluation.
-
-\subsubsection{Analytical Integration}
-
-Through prototyping, it has been found that neither implicit nor Crank-Nicholson time discretization of precursors are preferable methods for solving the shape equation in IQS.
-It has been found that these discretizations result in a lack of convergence of the shape over the IQS iteration process.
-In order to remedy the error, a analytical representation of the precursors was implemented in the prototype and the shape solution was able to converge (the normalization constant of the IQS method can be preserved to $10^{-10}$ while the theta-scheme only allowed convergence in the normalization factor to about $10^{-3}$).
-The following section shows how this method was implemented in the prototype and the desired implementation for Rattlesnake.
-
-Using an exponential operator, the precursor equation can be analytically solved for:
-
-\be
-\int_{t_n}^{t_{n+1}} C(t')e^{\lambda t'} dt' = \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{\lambda t'}dt'
-\ee
-
-yielding
-
-\be
-C^{n+1} =  C^n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
-\ee
-
-Because $\beta$ and $S_f$ being integrated are not known continuously over the time step, they can be interpolated linearly over the macro step.  Such that:
-
-\be
-h(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}h^n  + \frac{t-t_n}{t_{n+1}-t_n}h^{n+1}  \quad t_n \le t \le t_{n+1}
-\ee
-% We let $\ell_n(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}$ and  $\ell_{n+1}(t) = \frac{t-t_n}{t_{n+1}-t_n}$ for simplicity.
-
-However, for the PRKE solve, we do have a very accurate representation of $p(t')$ over the time interval $[t_n,t_{n+1}]$.
-
-Finally, we have the final expression for the analytical value for $C^{n+1}$:
-
-\be
-C^{n+1} = C^n e^{-\lambda \Delta t} 
-+ \left(a_3\beta^{n+1}+a_2\beta^n\right)S_f^{n+1}
-+ \left(a_2\beta^{n+1}+a_1\beta^n\right)S_f^n 
-\ee
-
-Where the integration coefficients are defined as:
-
-\begin{align}
-&a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)^2p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
-&a_2= \int_{t_n}^{t_{n+1}}\frac{(t'-t_n)(t_{n+1}-t')}{(\Delta t)^2}p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
-&a_3 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)^2p(t')e^{-\lambda(t_{n+1}-t')}dt'
-\end{align}
-
-The amplitude $(p)$ is included in the integration coefficient because it has been highly accurately calculated in the micro step scheme, so a piecewise interpolation between those points can be done to maximize accuracy.
-
-The prototype code uses Matlab software to interpolate the amplitude between micro steps and a quadrature integration for the coefficients.
-So the challenge for Rattlesnake is to replicate this procedure: passing the amplitude vector to the DNP auxkernel, interpolating it, and integrating the coefficients.
-
-\subsection{Predictor-Corrector Modification}
-In order to preserve the already implemented standard version of IQS, an option in the IQS executioner was created to specify which method is desired.
-Because the diffusion solve is flux instead of shape, when predictor-corrector option is specied, the IQS removal kernel ($\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$) and the modified precursor kernel are bypassed, while all the postprocessors are still executed.
-However, it is difficult to rescale the flux to shape before the PRKE parameter postprocessor are executed.
-So the parameters are computed using the full flux, but amplitude is space independent and comes out of the integrals.
-As seen in parameter definitions, when shape is replaced with flux, the amplitude comes out of the integral and cancels out.
-So the conversion of the predicted flux in Equation \ref{eq:rescale} to shape is unnecessary if the corrected flux is solved with Equation \ref{eq:phicorr2}.
-After obtaining the corrected flux, the precursors are re-evaluated using a \texttt{EXEC\_LINEAR}.
-\be
-\underbrace{\phi^g_{n+1}}_{\text{corrected}} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}} p_{n+1}
-\label{eq:phicorr2}
-\ee
-
-\subsection{Input}
-The input deck for IQS is very similar to the current transient diffusion input file.
-The IQS input has a different executioner type and parameters.
-The executioner type is simply ԉQSԠand input parameters include number of micro time steps, IQS error tolerance, and initial power.
-The Rattlesnake transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $\keff$, which is present in the transient input deck.
-However, IQS also requires an initial evaluation of the adjoint flux, for the weighting function.
-So another input file was made for the adjoint calculation, as well as including another multi-app and transfer in the IQS input deck.
-
-\subsection{Unintended Contributions}
-The implementation of IQS in Rattlesnake put pressure on many features of Rattlesnake and MOOSE that reveled bugs and possible improvement.
-Two significant issues in Rattlesnake that were found involved the adjoint solve and the diffusion fission kernel.
-When testing IQS, it was found that the adjoint flux solution was not the same as the forward flux solution in a single group test, which is obviously invalid.
-Also, when investigating which action to include the IQS kernel in, it turned out that the fission diffusion kernel was placed in the neutron transport action, so this kernel was demoted to the neutron diffusion actions for clarity.
-Additionally, the pressure on the save in feature in MOOSE propagated its application to boundary conditions and initial solves.
-MOOSE also updated its ability to restart dense vector data and to set MooseApp executioner right after executioner is created.
-Merge requests: \#5474, \#5489, \#5495, and \#5497
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{IQS in Other Action Systems}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\tcr{add the normalization condition I forgot to mention thus far}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Operator Notation and Extension of IQS to Multigroup Transport Equations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\tcr{introduce operator notation, then switch to transport}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Time Adaptation}
+\subsection{Solving the IQS System of Equations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Solving for the shape can become expensive, especially in two or three dimensions.  Using IQS, one \tcr{hopes} that the 
+shape's time dependency is weaker than that of the flux itself,  thus allowing larger time step sizes in computing the shape.
+The PRKE equations form a small ODE system and can be solved using a much smaller time step size.
+In transients where the
+shape varies much less than the flux, IQS can thus be very computationally effective. The two time scale solution process,
+a micro scale for the PRKE and a macro scale for the shape, is illustrated in \tcr{add label for fig}.
+
+\begin{figure}[!htbp]
+\includegraphics[width=\linewidth]{\FiguresDir/IQS_visualization.jpg}
+\caption{IQS method visualization}
+\end{figure}
+
+\tcr{add a hint that additional befefits can be relaized using time adaptation and refer to that upcoming Section}
+
+As we noted earlier, \tcr{add blurb about IQS is a nonlinear system, Picard, ...}
+
+\tcr{add another blurb + schematics about how IQS can be used in multiphysics simulation, where neutronics is nonlinear coupled to, say thermal-hydraulics}
+
+Additionally, to improve consistency and accuracy, each macro time step can be iterated so the best shape is used to compute power at the micro time steps.
+This iteration process must converge the shape such that the uniqueness condition $(\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=0)$ is preserved.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{A Predictor-Corrector version of IQS (IQS P-C)}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The Predictor-Corrector version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the evaluation of the coupled system is different.
+This version first solves the flux diffusion (represented by Equations \ref{eq:flux} and \ref{eq:precursor}) to get a predicted flux.
+The predicted flux at this step is then converted to shape by rescaling as follows:
+\be
+\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
+\label{eq:rescale}
+\ee
+Where:
+\be
+K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
+\ee
+\be
+K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
+\ee
+
+The PRKE parameters are then computed with this shape using Equations \ref{eq:rmb} - \ref{eq:l} and interpolated over the macro step, then the PRKE is evaluated.
+With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
+\be
+\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1}
+\ee
+
+The advantage to the predictor-corrector method is there is no iteration necessary for this method and in turn is much simpler and faster than the standard IQS.
+The disadvantage is that the method assumes $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)$ is inherently constant.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Precursor Integration}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\tcr{ this subsection and its 2 subsubsections on precursors is way to long. make it short and sweet}
+
+This section presents two different time-integration methods to solve coupled IQS shape + precursor equations, recalled below
+using, for simplicity, a single neutron group and a single precursor group.
+
+\be
+\frac{1}{v}\frac{\partial\varphi}{\partial t}=\nu\Sigma_f(1-\beta)\varphi-\left(-\div D \grad + \Sigma_a + \frac{1}{v}\frac{1}{p}\frac{dp}{dt}\right)\varphi+\frac{1}{p}\lambda C 
+\ee
+\be
+\frac{dC}{dt} = \beta\nu \Sigma_f \varphi p - \lambda C
+\ee
+
+First, we note that we could keep this system of two time-dependent equations and solve it as a coupled system. 
+However, this is unnecessary and a memory expensive endeavor because the precursor equation is only an ODE and not a PDE. 
+Instead, one may discretize in time the shape equation, which typically requires the knowledge of the
+precursor concentrations at the end of the time step. This precursor value is taken from the solution, numerical or analytical,
+of the precursors ODE.
+This document will discuss two techniques for solving the precursor equation.
+First is a time discretization method that is currently being implemented in Rattlesnake.
+The second is a analytical integration of the precursors, the latter method has proven to be more beneficial for IQS convergence.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Time Discretization using the Theta Method}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A fairly simple way to evaluate the precursor equation is to employ the $\theta$-scheme ($0\le\theta\le1$),
+explicit when $\theta=0$, implicit when $\theta=1$, and Crank-Nicholson when $\theta=1/2$).
+Generally, if there is a function $u$ whose governing equation is $\frac{du}{dt}=f(u,t)$,
+then the $\theta$-discretization is
+
+\be
+\frac{u^{n+1}-u^n}{\Delta t}=(1-\theta)f(u^n,t) + \theta f(u^{n+1},t) \,.
+\ee
+
+Applying this to the precursor equation:
+
+\be
+\frac{C^{n+1}-C^n}{\Delta t}=(1-\theta)\beta S_f^np^n-(1-\theta)\lambda C^n + \theta\beta S_f^{n+1}p^{n+1}-\theta\lambda C^{n+1}
+\ee
+
+Where $S_f$ is the fission source equivalent for shape:
+
+\be
+S_f^n=(\nu\Sigma_f)^n\varphi^n
+\ee
+
+Rearranging to solve for the precursor at the end of the time step yields
+
+\be
+C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta}{1+\theta\Delta t\lambda}S_f^n p^n +  \frac{\theta\Delta t \beta}{1+\theta\Delta t\lambda}S_f^{n+1} p^{n+1}
+\ee
+Reporting this value of $C^{n+1}$, one can solve for the shape $\varphi^{n+1}$ as a function of $\varphi^n$ and $C^n$
+(and $p^n$, $p^{n+1}$, $dp/dt|_n$ and  $dp/dt|_{n+1}$).
+Once $\varphi^{n+1}$ has been determined, $C^{n+1}$ is updated. Rattlesnake currently implements both implicit and Crank-Nicholson as options for precursor evaluation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Analytical Integration}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Through prototyping, it has been found that neither implicit nor Crank-Nicholson time discretization of precursors are preferable methods for solving the shape equation in IQS.
+It has been found that these discretizations result in a lack of convergence of the shape over the IQS iteration process.
+In order to remedy the error, a analytical representation of the precursors was implemented in the prototype and the shape solution was able to converge (the normalization constant of the IQS method can be preserved to $10^{-10}$ while the theta-scheme only allowed convergence in the normalization factor to about $10^{-3}$).
+The following section shows how this method was implemented in the prototype and the desired implementation for Rattlesnake.
+
+Using an exponential operator, the precursor equation can be analytically solved for:
+
+\be
+\int_{t_n}^{t_{n+1}} C(t')e^{\lambda t'} dt' = \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{\lambda t'}dt'
+\ee
+
+yielding
+
+\be
+C^{n+1} =  C^n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
+\ee
+
+Because $\beta$ and $S_f$ being integrated are not known continuously over the time step, they can be interpolated linearly over the macro step.  Such that:
+
+\be
+h(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}h^n  + \frac{t-t_n}{t_{n+1}-t_n}h^{n+1}  \quad t_n \le t \le t_{n+1}
+\ee
+% We let $\ell_n(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}$ and  $\ell_{n+1}(t) = \frac{t-t_n}{t_{n+1}-t_n}$ for simplicity.
+
+However, for the PRKE solve, we do have a very accurate representation of $p(t')$ over the time interval $[t_n,t_{n+1}]$.
+
+Finally, we have the final expression for the analytical value for $C^{n+1}$:
+
+\be
+C^{n+1} = C^n e^{-\lambda \Delta t} 
++ \left(a_3\beta^{n+1}+a_2\beta^n\right)S_f^{n+1}
++ \left(a_2\beta^{n+1}+a_1\beta^n\right)S_f^n 
+\ee
+
+Where the integration coefficients are defined as:
+
+\begin{align}
+&a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)^2p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
+&a_2= \int_{t_n}^{t_{n+1}}\frac{(t'-t_n)(t_{n+1}-t')}{(\Delta t)^2}p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
+&a_3 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)^2p(t')e^{-\lambda(t_{n+1}-t')}dt'
+\end{align}
+
+The amplitude $(p)$ is included in the integration coefficient because it has been highly accurately calculated in the micro step scheme, so a piecewise interpolation between those points can be done to maximize accuracy.
+
+The prototype code uses Matlab software to interpolate the amplitude between micro steps and a quadrature integration for the coefficients.
+So the challenge for Rattlesnake is to replicate this procedure: passing the amplitude vector to the DNP auxkernel, interpolating it, and integrating the coefficients.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Time Adaptation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Rattlesnake Implementation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\tcr{I think we need an introduction on how things work in a MOOSE-based application, before going into the details of each part}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Executioner}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The IQS executioner derives from the Transient executioner in MOOSE.
+The IQS executioner contains a loop over micro time steps that computes the PRKE and then passes $p$ and $\frac{dp}{dt}$ for the Transient executioner to evaluate the shape equation at each macro step.
+The PRKE is computed with a user specified option of backward-Euler, Crank-Nicolson, or SDIRK33.
+The IQS executioner also supplements Transient Picard iteration process by adding its own error criteria:
+\be
+Error_{IQS}=\left|\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,n}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,0}\right)}-1\right|
+\ee
+
+\tcr{need to make sure this error is also explained int he theory section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Action System}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+IQS defines its uniqueness from its executioner type; however, many changes needed to be made in the Rattlesnake action system in order to support IQS execution.
+First, changes needed to be made in order to evaluate the shape equation.
+The shape equation, after some manipulation, is very similar to the time-dependent, which Rattlesnake is already set up to solve:
+
+\begin{align}
+\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
+& - \underbrace{\frac{1}{v^g} \overbrace{\frac{1}{p}\frac{dp}{dt}}^{From Executioner}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
+\end{align}
+
+To enable Rattlesnake to solve this equation, another kernel was created that evaluated $\sum_{g=1}^G\frac{1}{v^gp}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
+Second, four postprocessors were created in order to calculate the PRKE parameters.
+The parameter calculations were separated by $\frac{\bar{\beta}_i}{\Lambda}$ numerator, $\bar{\lambda}_i$ numerator/denominator, $\frac{\rho}{\Lambda}/\frac{\bar{\beta}}{\Lambda}$ denominator, and $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator.
+The first three are relatively simple, only relying on material properties and solution quantities.
+The $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator requires the use of the MOOSE save in feature, which saves the residual from a calculated kernel or boundary contribution in the shape evaluation to an auxiliary variable.
+Finally, a user object was created to pull together all the postprocessor values and carryout the numerator/denominator divisions that were then passed to the executioner.
+
+\tcr{need to make it clear what action system will have IQS in the end. a list would be good}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{PRKE coefficients}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\tcr{here you should talk about save in and postprocessors. Remove the relevant info, neatly, from the theory section and bring it here}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Predictor-Corrector Modification}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In order to preserve the already implemented standard version of IQS, an option in the IQS executioner was created to specify which method is desired.
+Because the diffusion solve is flux instead of shape, when predictor-corrector option is specied, the IQS removal kernel ($\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$) and the modified precursor kernel are bypassed, while all the postprocessors are still executed.
+However, it is difficult to rescale the flux to shape before the PRKE parameter postprocessor are executed.
+So the parameters are computed using the full flux, but amplitude is space independent and comes out of the integrals.
+As seen in parameter definitions, when shape is replaced with flux, the amplitude comes out of the integral and cancels out.
+So the conversion of the predicted flux in Equation \ref{eq:rescale} to shape is unnecessary if the corrected flux is solved with Equation \ref{eq:phicorr2}.
+After obtaining the corrected flux, the precursors are re-evaluated using a \texttt{EXEC\_LINEAR}.
+\be
+\underbrace{\phi^g_{n+1}}_{\text{corrected}} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}} p_{n+1}
+\label{eq:phicorr2}
+\ee
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Input}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The input deck for IQS is very similar to the current transient diffusion input file.
+The IQS input has a different executioner type and parameters.
+The executioner type is simply IQS and input parameters include number of micro time steps, IQS error tolerance, and initial power.
+The Rattlesnake transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $\keff$, which is present in the transient input deck.
+However, IQS also requires an initial evaluation of the adjoint flux, for the weighting function.
+So another input file was made for the adjoint calculation, as well as including another multi-app and transfer in the IQS input deck.
+
+\tcr{in the yak manual you may find how to incorporate nicely a piece of the input file as a table or figure here.
+the list of parameters and available options for each must be given.
+Maybe an additional appendix should be a smaple input file, with comments}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Additional Contributions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The implementation of IQS in Rattlesnake put pressure on many features of Rattlesnake and MOOSE that reveled bugs and possible improvement.
+Two significant issues in Rattlesnake that were found involved the adjoint solve and the diffusion fission kernel.
+When testing IQS, it was found that the adjoint flux solution was not the same as the forward flux solution in a single group test, which is obviously invalid.
+Also, when investigating which action to include the IQS kernel in, it turned out that the fission diffusion kernel was placed in the neutron transport action, so this kernel was demoted to the neutron diffusion actions for clarity.
+Additionally, the pressure on the save in feature in MOOSE propagated its application to boundary conditions and initial solves.
+MOOSE also updated its ability to restart dense vector data and to set MooseApp executioner right after executioner is created.
+Merge requests: \#5474, \#5489, \#5495, and \#5497
+
+\tcr{is this list exhaustive?}
+
 \newpage
@@
 %------------------------------------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{RESULTS} 
 \label{sect::results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%------------------------------------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{One-Dimensional Custom Example}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%------------------------------------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{TWIGL Benchmark}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%------------------------------------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{LRA Benchmark}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+
+\tcr{needed}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\clearpage
+\pagebreak
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Appendix: aaaa}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-06-09 doc/NEAMS_Report/NEAMS_Report.tex
Addressing Section 1  comments

@@
 \usepackage{multirow}
+\usepackage{tikz}
 
@@
 \par
-The improved quasi-static (IQS) method \tcr{you need to give 3-4 references} is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (the shape) and a time-dependent component (the amplitude).
-The technique relies on the shape being less rapidly varying in time compared to the flux, hence
-requiring fewer shape computations or updates. 
-The IQS method has primarily been developped in the context of the neutron diffusion approximation \tcr{give refs}, but \tcr{cite sedat} provides an extension of the technique to transport, implemented in TDTORT.
-The method has mostly being applied to neutron kinetics, without thermal-hydraulic feedback. In this report, we extend the IQS
-method in the context of transient multiphysics simulations.
-%The method These components include the flux's power and shape.
-Power is time-dependent, while the shape is both space- and time-dependent.
-However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the shape may not require computation at every time step, invoking the Աuasi-staticԠnature.
+The improved quasi-static (IQS) method is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (the shape) and a time-dependent component (the amplitude) \cite{Ott_1966,Dulla2008}.
+The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates.
+The IQS method has primarily been developed in the context of the neutron diffusion approximation \cite{Ban_2012,Ferguson_1973}, but \cite{Devooght_1980} provides an extension of the technique to transport, implemented in TDTORT.
+The method has mostly being applied to neutron kinetics, without thermal-hydraulic feedback.
+In this report, we extend the IQS method in the context of transient multiphysics simulations.
 
@@
 Extension to the transport equations will be provided subsequently, once a concise operator notation has been introduced. 
-In the above equations, \tcr{give a long text naming each variables, for example: $\phi^g$ is the neutron flux in group $g$, the removal, fission, scattering cross sections are denoted by ...; the total delayed neutron fraction is
-\[
-\beta = \sum_{i=1}^I \beta_{i} 
-\]
-}
+In the above equations,  \\
+
+\begin{tabular}{lll}
+$\phi^g$				&	$=$	&	Scalar flux in energy group $g$ \\
+$C_i$					&	$=$	&	Concentration of delayed neutron precursor $i$ \\
+$\keff$					&	$=$	&	Effective multiplication factor \\
+$\Sigma_f^{g}$			&	$=$	&	Fission cross section in energy group $g$ \\
+$\Sigma_r^{g}$			&	$=$	&	Removal cross section in energy group $g$ \\
+$\Sigma_s^{g' \to g}$	&	$=$	&	Scattering cross section from energy group $g'$ to $g$ \\
+$v^g$					&	$=$	&	Neutron velocity in energy group $g$ \\
+$\chi_p^g$				&	$=$	&	Fission spectrum of prompt neutrons \\
+$\chi_{d,i}^g$			&	$=$	&	Fission spectrum of delayed neutrons from precursor $i$ \\
+$\nu^g$					&	$=$	&	Total number of neutrons per fission \\
+$D^g$					&	$=$	&	Diffusion coefficient \\
+$\lambda_i$				&	$=$	&	Decay constant of precursor $i$ \\
+$\beta_i$				&	$=$	&	Delayed neutron fraction  from precursor $i$ \\
+$\beta$			 	&	$=$	&	Total delayed neutron fraction ($\beta = \sum_{i=1}^I \beta_{i}$) \\
+  & & 
+\end{tabular}
+
 Note that the prompt fission terms have been normalized by $\keff$, the eigenvalue of the initial state, as transient simulations are often carried out for initially critical reactors.
 The expressions for subcritical and source-driven problems are straightforward to derive but omitted in this report for brevity.
+\\
 
@@
 \begin{enumerate}
-\item the shape equation contains an additional term equivalent to a removal cross section,  $ \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$,
-\item the delayed neutron source term is divided by $p$,
-\item the system of equations is now nonlinear due to the factorization,
-\item an equation is needed to obtain the amplitude $p$.
+\item The shape equation contains an additional term equivalent to a removal cross section,  $ \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$.
+\item The delayed neutron source term is divided by $p$.
+\item The system of equations is now nonlinear due to the factorization.
+\item An equation is needed to obtain the amplitude $p$.
 \end{enumerate}
@@
 The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.
-
-\tcr{add the normalization condition I forgot to mention thus far}
+Additionally, in order to impose uniqueness on the factorization and to derive the PRKE the following normalization condition is imposed: $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}$.
+This condition is later used  as the convergence criteria when solving the nonlinear system of equations.
 
@@
 
-Solving for the shape can become expensive, especially in two or three dimensions.  Using IQS, one \tcr{hopes} that the 
-shape's time dependency is weaker than that of the flux itself,  thus allowing larger time step sizes in computing the shape.
+Solving for the shape can become expensive, especially in two or three dimensions.
+Using IQS, one expects that the shape's time dependency is weaker than that of the flux itself,  thus allowing larger time step sizes in computing the shape.
 The PRKE equations form a small ODE system and can be solved using a much smaller time step size.
-In transients where the
-shape varies much less than the flux, IQS can thus be very computationally effective. The two time scale solution process,
-a micro scale for the PRKE and a macro scale for the shape, is illustrated in \tcr{add label for fig}.
+In transients where the shape varies much less than the flux, IQS can thus be very computationally effective.
+The two time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in Figure \ref{fig:iqsviz}.
 
@@
 \caption{IQS method visualization}
-\end{figure}
-
-\tcr{add a hint that additional befefits can be relaized using time adaptation and refer to that upcoming Section}
-
-As we noted earlier, \tcr{add blurb about IQS is a nonlinear system, Picard, ...}
-
-\tcr{add another blurb + schematics about how IQS can be used in multiphysics simulation, where neutronics is nonlinear coupled to, say thermal-hydraulics}
-
-Additionally, to improve consistency and accuracy, each macro time step can be iterated so the best shape is used to compute power at the micro time steps.
+\label{fig:iqsviz}
+\end{figure}
+
+By this design, the benefits of IQS is most realized when using time adaptive techniques, one of these is described in Section \ref{sect:dt2}.
+As we noted earlier, shape-PRKE equations are a nonlinear system and thus need to be solved in a iterative manner.
+Each macro time step can be iterated so the best shape is used to compute power at the micro time steps.
 This iteration process must converge the shape such that the uniqueness condition $(\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=0)$ is preserved.
+The most convenient method to implement in Rattlesnake are Picard iterations.
+Where initially, the shape is evaluated with a guessed amplitude, if the normalization condition changes significantly, then the PRKE is evaluated with parameters computed from the last solution.
+This process is repeated until convergence.
+Picard iteration is predominantly utilized for multiphysics simulation, where neutronics is nonlinearly coupled to thermal-hydraulics.
+So the shape-PRKE couple is easily fit into this process, described by Figure \ref{fig:picard}.
+
+\begin{figure}[!htpb]
+\centering
+\begin{tikzpicture}[
+redbox/.style = {rectangle, rounded corners, draw, align=center, top color=white, bottom color=red!20, ultra thick, minimum size=20mm}, 
+bluebox/.style = {rectangle, rounded corners, draw, align=center, top color=white, bottom color=blue!20, ultra thick, minimum size=20mm},
+greenbox/.style = {rectangle, rounded corners, draw, align=center, top color=white, bottom color=green!20, ultra thick, minimum size=20mm},
+every node/.style       = {font=\large}
+]
+\node[redbox](iqs) at (0,0) {IQS};
+\node[bluebox](shape) at (6,2) {Shape};
+\node[greenbox](prke) at (11,2) {PRKE};
+\node[bluebox](mp) at (6,-2) {Multi-Physics};
+
+\draw[<->,ultra thick,](iqs.north east) -- node[above,sloped] {Picard}(shape.west);
+%\draw[<->,ultra thick](iqs.east) -- node[above] {Picard}(prke.west);
+\draw[<->,ultra thick](iqs.south east) -- node[below,sloped] {Picard} (mp.west);
+
+\draw[->,ultra thick](prke.south) to [bend left=60] node[above,sloped,midway] {$p,\frac{dp}{dt}$}(shape.south);
+\draw[->,ultra thick](shape.east) -- (prke.west) node[above,sloped,midway] {$\rho,\bar{\beta},\bar{\lambda},\Lambda$};
+
+\end{tikzpicture}
+\caption{Visualization of Picard iteration process}
+\label{fig:picard}
+\end{figure}
 
@@
 
-\tcr{ this subsection and its 2 subsubsections on precursors is way to long. make it short and sweet}
-
-This section presents two different time-integration methods to solve coupled IQS shape + precursor equations, recalled below
-using, for simplicity, a single neutron group and a single precursor group.
+This section presents two different time-integration methods to solve coupled IQS shape + precursor equations, recalled below using, for simplicity, a single neutron group and a single precursor group.
 
@@
 However, this is unnecessary and a memory expensive endeavor because the precursor equation is only an ODE and not a PDE.
-Instead, one may discretize in time the shape equation, which typically requires the knowledge of the
-precursor concentrations at the end of the time step. This precursor value is taken from the solution, numerical or analytical,
-of the precursors ODE.
-This document will discuss two techniques for solving the precursor equation.
-First is a time discretization method that is currently being implemented in Rattlesnake.
-The second is a analytical integration of the precursors, the latter method has proven to be more beneficial for IQS convergence.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Time Discretization using the Theta Method}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A fairly simple way to evaluate the precursor equation is to employ the $\theta$-scheme ($0\le\theta\le1$),
-explicit when $\theta=0$, implicit when $\theta=1$, and Crank-Nicholson when $\theta=1/2$).
-Generally, if there is a function $u$ whose governing equation is $\frac{du}{dt}=f(u,t)$,
-then the $\theta$-discretization is
-
-\be
-\frac{u^{n+1}-u^n}{\Delta t}=(1-\theta)f(u^n,t) + \theta f(u^{n+1},t) \,.
-\ee
-
-Applying this to the precursor equation:
-
-\be
-\frac{C^{n+1}-C^n}{\Delta t}=(1-\theta)\beta S_f^np^n-(1-\theta)\lambda C^n + \theta\beta S_f^{n+1}p^{n+1}-\theta\lambda C^{n+1}
-\ee
-
-Where $S_f$ is the fission source equivalent for shape:
-
-\be
-S_f^n=(\nu\Sigma_f)^n\varphi^n
-\ee
-
-Rearranging to solve for the precursor at the end of the time step yields
-
-\be
-C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta}{1+\theta\Delta t\lambda}S_f^n p^n +  \frac{\theta\Delta t \beta}{1+\theta\Delta t\lambda}S_f^{n+1} p^{n+1}
-\ee
-Reporting this value of $C^{n+1}$, one can solve for the shape $\varphi^{n+1}$ as a function of $\varphi^n$ and $C^n$
-(and $p^n$, $p^{n+1}$, $dp/dt|_n$ and  $dp/dt|_{n+1}$).
-Once $\varphi^{n+1}$ has been determined, $C^{n+1}$ is updated. Rattlesnake currently implements both implicit and Crank-Nicholson as options for precursor evaluation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Analytical Integration}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Through prototyping, it has been found that neither implicit nor Crank-Nicholson time discretization of precursors are preferable methods for solving the shape equation in IQS.
+Instead, one may discretize in time the shape equation, which typically requires the knowledge of the precursor concentrations at the end of the time step.
+This precursor value is taken from the solution, numerical or analytical, of the precursors ODE.
+\\
+
+Through prototyping, it has been found that typical implicit nor Crank-Nicholson time discretizations of precursors are not preferable methods for solving the shape equation in IQS.
 It has been found that these discretizations result in a lack of convergence of the shape over the IQS iteration process.
-In order to remedy the error, a analytical representation of the precursors was implemented in the prototype and the shape solution was able to converge (the normalization constant of the IQS method can be preserved to $10^{-10}$ while the theta-scheme only allowed convergence in the normalization factor to about $10^{-3}$).
-The following section shows how this method was implemented in the prototype and the desired implementation for Rattlesnake.
-
-Using an exponential operator, the precursor equation can be analytically solved for:
-
-\be
-\int_{t_n}^{t_{n+1}} C(t')e^{\lambda t'} dt' = \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{\lambda t'}dt'
-\ee
-
-yielding
-
-\be
-C^{n+1} =  C^n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
-\ee
-
-Because $\beta$ and $S_f$ being integrated are not known continuously over the time step, they can be interpolated linearly over the macro step.  Such that:
-
-\be
-h(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}h^n  + \frac{t-t_n}{t_{n+1}-t_n}h^{n+1}  \quad t_n \le t \le t_{n+1}
-\ee
-% We let $\ell_n(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}$ and  $\ell_{n+1}(t) = \frac{t-t_n}{t_{n+1}-t_n}$ for simplicity.
-
-However, for the PRKE solve, we do have a very accurate representation of $p(t')$ over the time interval $[t_n,t_{n+1}]$.
-
-Finally, we have the final expression for the analytical value for $C^{n+1}$:
+In order to remedy the error, a analytical representation of the precursors was implemented in the prototype and the shape solution was able to converge.
+The method entails linearly interpolating the fission source ($S_f = \nu \Sigma_f \varphi$) and integrating the resulting continuous function analytically.
+Equation \ref{eq:c_dis} is the equation to solve for the precursors and Equations \ref{eq:a1} and \ref{eq:a2} are integrating factors.
 
@@
 C^{n+1} = C^n e^{-\lambda \Delta t} 
-+ \left(a_3\beta^{n+1}+a_2\beta^n\right)S_f^{n+1}
-+ \left(a_2\beta^{n+1}+a_1\beta^n\right)S_f^n 
-\ee
-
-Where the integration coefficients are defined as:
-
-\begin{align}
-&a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)^2p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
-&a_2= \int_{t_n}^{t_{n+1}}\frac{(t'-t_n)(t_{n+1}-t')}{(\Delta t)^2}p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
-&a_3 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)^2p(t')e^{-\lambda(t_{n+1}-t')}dt'
-\end{align}
++ \beta_i \left(a_2 S_f^{n+1} + a_1 S_f^{n}\right)
+\label{eq:c_dis}
+\ee
+
+\be
+a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)p(t')e^{-\lambda(t_{n+1}-t')}dt'
+\label{eq:a1}
+\ee
+\be
+a_2 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)p(t')e^{-\lambda(t_{n+1}-t')}dt'
+\label{eq:a2}
+\ee
 
@@
 
-The prototype code uses Matlab software to interpolate the amplitude between micro steps and a quadrature integration for the coefficients.
-So the challenge for Rattlesnake is to replicate this procedure: passing the amplitude vector to the DNP auxkernel, interpolating it, and integrating the coefficients.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Time Adaptation}
+\label{sect:dt2}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 If $e_n > e_{max}$ the time step is repeated; if $e_n < e_{max}$ the system moves to the next time step.
-The next $\Delta t$ is calculated using Equation \ref{eq:dt2}, where $p$ is the convergence order of the time integration scheme being used.
+The next $\Delta t$ is calculated using Equation \ref{eq:dt2}, where $\mu$ is the convergence order of the time integration scheme being used.
 $e_{max}$ and $e_{tol}$ are user defined parameters; $e_{max}$ is usually less than $e_{tol}$ to better guarantee that the calculated $\Delta t_{new}$ will pass the error criteria so time steps won't be repeated.
@@
 \be
-\Delta t_{new} = \Delta t_{old} \left(\frac{e_{tol}}{e_n}\right)
+\Delta t_{new} = \Delta t_{old} \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+\mu}}
 \label{eq:dt2}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\bibliographystyle{ans}
+\bibliography{references_IQS}
 \end{document}

### 2016-06-09 doc/NEAMS_Report/NEAMS_Report.tex
Fixing things in Section 2 and 3

@@
 \subsection{Operator Notation and Extension of IQS to Multigroup Transport Equations}
+\label{sect:transport}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-For simplicity and intuitiveness the previous section only described IQS implementation in CFEM diffusion.
-Rattlesnake has other action systems capable of transient simulation, where IQS can be implemented and be effective.
-One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equation \ref{eq:flux} and \ref{eq:shape2}.
-However, in the derivation for IQS is term is unaffected between shape and flux evaluation.
-So saving the residual for this diffusion kernel in the \texttt{save\_in} variable is the only alteration to this action for IQS to function.
-\\
+For simplicity and intuitiveness the previous section only described IQS implementation in CFEM diffusion.   \\
 
@@
 \ee
-
-The derivation of IQS in transport is simple and straight-forward, and its implementation into Rattlesnake is even more so.
-The main differences between diffusion implementation in transport is outlined below:
-\begin{enumerate}
-\item The form of the operators in the shape equations is different, but Rattlesnake has already implemented all the kernels necessary to represent these operators.
-So no change is necessary to these kernels is necessary for IQS.
-Additionally, the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ is the same, so both actions can use the same kernel.
-\item The PRKE parameters also change because of the operators.
-For the $(\rho-\bar{\beta})$ parameter, the same post-processor can be used with the \texttt{save\_in} functionality in MOOSE.
-The post-processor for $\bar{\beta}_i$ must be re-written for transport, but takes a very similar form as diffusion.
-\end{enumerate}
 
@@
 \subsection{Precursor Integration}
+\label{sect:prec}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\tcr{I think we need an introduction on how things work in a MOOSE-based application, before going into the details of each part}
+Rattlesnake is a MOOSE-based application developed INL specific to solving radiation transport problems with multiphysics capabilities.
+MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like Rattlesnake.
+Because radiation transport is such a complex problem, Rattlesnake is a relatively large application with a lot of functionality.
+At the heart of Rattlesnake is the action system.
+The Rattlesnake action system gives the capability of consolidating MOOSE input syntax, so that a user does not have to define every kernel, variable, etc.
+in the problem, which could be on the order of thousands with multi-group problems.
+The user, instead, is able to input an equation description (Diffusion, S$_n$, P$_n$, etc.) and solve method (SAAF,LS,CFEM,DFEM,etc.), and the action system will incorporate all the necessary physics involved (kernels, boundary conditions, postprocessor, etc.).
 
@@
 
-\tcr{need to make sure this error is also explained int he theory section}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-To enable Rattlesnake to solve this equation, another kernel was created that evaluated $\sum_{g=1}^G\frac{1}{v^gp}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
-Second, four postprocessors were created in order to calculate the PRKE parameters.
-The parameter calculations were separated by $\frac{\bar{\beta}_i}{\Lambda}$ numerator, $\bar{\lambda}_i$ numerator/denominator, $\frac{\rho}{\Lambda}/\frac{\bar{\beta}}{\Lambda}$ denominator, and $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator.
-The first three are relatively simple, only relying on material properties and solution quantities.
-The $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator requires the use of the MOOSE save in feature, which saves the residual from a calculated kernel or boundary contribution in the shape evaluation to an auxiliary variable.
-Finally, a user object was created to pull together all the postprocessor values and carryout the numerator/denominator divisions that were then passed to the executioner.
-
-\tcr{need to make it clear what action system will have IQS in the end. a list would be good}
+To enable Rattlesnake to solve this equation, another kernel was created that evaluates $\sum_{g=1}^G\frac{1}{v^gp}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
+Also, the precursor kernel was modified to include the $\frac{1}{p}$ term.
+Finally, the precursor auxkernel that evaluates Equation \ref{eq:preq2} using the analytical integration method described in Section \ref{sect:prec}.
 
@@
 
-\tcr{here you should talk about save in and postprocessors. Remove the relevant info, neatly, from the theory section and bring it here}
+In order to evaluate the PRKE coefficients, defined by Equations \ref{eq:rmb} - \ref{eq:l}, four postprocessors were created.
+The parameter calculations were separated by $\frac{\bar{\beta}_i}{\Lambda}$ numerator, $\bar{\lambda}_i$ numerator/denominator, $\frac{\rho-\bar{\beta}}{\Lambda}/\frac{\bar{\beta}}{\Lambda}$ denominator, and $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator.
+The first three are relatively simple, only relying on material properties and solution quantities, then computing the elemental integral.
+The $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator requires the use of the MOOSE \texttt{save\_in} feature.
+This feature saves the residual from a calculated kernel or boundary contribution in the shape evaluation to an auxiliary variable.
+The postprocessor then computes the inner product of this variable and the initial adjoint solution.
+After each of these postprocessors are evaluated, a user object pulls together all the values and performs the numerator/denominator divisions.
+The resulting values are then passed to the executioner for the PRKE evaluation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Other Action Systems}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For simplicity, IQS implementation has only been described for CFEM diffusion.
+However, Rattlesnake has other action systems capable of transient simulation, where IQS can be implemented and be effective.
+One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equation \ref{eq:flux} and \ref{eq:shape2}.
+However, in the derivation for IQS is term is unaffected between shape and flux evaluation.
+So saving the residual for this diffusion kernel in the \texttt{save\_in} variable is the only alteration to this action for IQS to function.
+
+Additional action systems involve transport, but it is evident from Section \ref{sect:transport} that IQS implementation in these is straightforward.
+The main differences between diffusion implementation in transport is outlined below:
+\begin{enumerate}
+\item The form of the operators in the shape equations is different, but Rattlesnake has already implemented all the kernels necessary to represent these operators.
+So no change is necessary to these kernels is necessary for IQS.
+Additionally, the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ is the same, so both actions can use the same kernel.
+\item The PRKE parameters also change because of the operators.
+For the $(\rho-\bar{\beta})$ parameter, the same post-processor can be used with the \texttt{save\_in} functionality in MOOSE.
+The post-processor for $\bar{\beta}_i$ must be re-written for transport, but takes a very similar form as diffusion.
+\end{enumerate}
+
+IQS plans to be available to any action system capable of transient simulation, which include:
+\begin{itemize}
+\item CFEM-Diffusion
+\item DFEM-Diffusion
+\item SAAF-CFEM-SN
+\item SAAF-CFEM-PN
+\item LS-CFEM-SN
+\item DFEM-SN
+\end{itemize}
 
@@
 However, IQS also requires an initial evaluation of the adjoint flux, for the weighting function.
-So another input file was made for the adjoint calculation, as well as including another multi-app and transfer in the IQS input deck.
+So another input file and multi-app transfer was made for the adjoint calculation.
+An example input file with detailed commentary can be found in the Appendix.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\tcr{I dont think this section is really necessary}
 
@@
 \begin{figure}[!htbp]
-\includegraphics[width=\linewidth]{\FiguresDir/1D_het_power_plot.png}
-\caption{Power level of 1D heterogeneous example using IQS and Brute Force with $\Delta t = 0.025$}
+\centering
+\includegraphics[width=0.8\linewidth]{\FiguresDir/1D_het_power_plot.png}
+\caption{Power level of 1D heterogeneous example with $\Delta t = 0.01$}
 \label{fig:power}
@@
 \label{tab:1Ddt2}
+\resizebox{\textwidth}{!}{
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
@@
 \hline
-\end{tabular}
+\end{tabular}}
 \end{center}
@@
 \includegraphics[height=3.0in]{\FiguresDir/TWIGL_power_plot.png}
-\caption{Power level comarison of 1D heterogeneous example between IQS and Brute Force using $\Delta t = 0.004$}
+\caption{Power level comparison of TWIGL Benchmark with $\Delta t = 0.025$}
 \label{fig:TWIGL_power}
@@
 \label{tab:TWIGLdt2}
+\resizebox{\textwidth}{!}{
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
@@
 
-\end{tabular}
+\end{tabular}}
 \end{center}
@@
 The geometry is illustrated in~\fig{fig:lra-geometry}.\\
-\begin{figure}[H]
+\begin{figure}[!htbp]
 \centering
@@
 Some scalar data are listed in~\tbl{tab:lra-scalar}.\\
-\begin{table}[H]
+\begin{table}[!htbp]
     \centering
     \caption{LRA benchmark initial two-group constants.\label{tab:lra-xs}}
+    \resizebox{\textwidth}{!}{
       \begin{tabular}{|c|c|c|c|l|l|l|l|l|}
@@
       \hline
-      \end{tabular}
+      \end{tabular}}
 \end{table}
 
-\begin{table}[H]
+\begin{table}[!htbp]
     \centering
@@
 
-\begin{table}[H]
+\begin{table}[!htbp]
     \centering
@@
 These plots show that IQS is consistent with Doppler feedback problems in RATTLESNAKE.
-Finally, Figure \ref{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods.
+Figure \ref{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods.
+Table  \ref{tab:LRAdt2} shows the time adaptation performance of both methods.
+These results show that IQS, again, performs much better than the brute force method by taking significantly less diffusion solves while retaining similar accuracy.
 
@@
 \begin{subfigure}[!htbp]{0.49\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/lra_power_profile.jpg}
+\includegraphics[width=\textwidth]{\FiguresDir/lra_power_profile.png}
 \caption{Power Profile}
@@
 \begin{subfigure}[!htbp]{0.49\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/lra_temp_profile.jpg}
+\includegraphics[width=\textwidth]{\FiguresDir/lra_temp_profile.png}
 \caption{Temperature Profile}
@@
 \end{figure}
+
+\begin{table}[!htbp]
+\caption{LRA step doubling results}
+\label{tab:LRAdt2}
+\begin{center}
+\begin{tabular}{|l|l|l|l|l|l|l|l|}
+\hline
+  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS P-C} \\
+\hline
+Test & $e_{max}$ & Error & Steps & Solves & Error & Steps & Solves \\
+\hline
+1 &	 0.01 &	 0.0437566 &	 145 &	 735 &	 0.143855 &	 105 &	 616 \\ 
+2 &	 0.005 &	 0.0309732 &	 266 &	 948 &	 0.104782 &	 157 &	 814 \\ 
+3 &	 0.001 &	 0.158242 &	 585 &	 2046 &	 0.155676 &	 218 &	 1062 \\ 
+4 &	 0.0005 &	 0.270949 &	 833 &	 2908 &	 0.211994 &	 480 &	 2038 \\ 
+5 &	 0.0001 &	 0.280468 &	 1793 &	 6155 &	 0.0437761 &	 434 &	 1458 \\ 
+6 &	 5e-05 &	 0.212274 &	 2485 &	 8511 &	 0.0385312 &	 542 &	 1719 \\ 
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 \frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-\label{eq:rmb}
+%\label{eq:rmb}
 \ee
@@
 \frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-\label{eq:b}
+%\label{eq:b}
 \ee
@@
 \bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
-\label{eq:l}
+%\label{eq:l}
 \ee
@@
 \frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
-\label{eq:p}
+%\label{eq:p}
 \ee
@@
 \frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
-\label{eq:c}
+%\label{eq:c}
 \ee
@@
 & -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
-\label{eq:shape2}
+%\label{eq:shape2}
 \end{align}
@@
 \frac{dC_i}{dt} = \frac{\beta_i}{\keff}p \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
-\label{eq:preq2}
+%\label{eq:preq2}
 \ee

### 2016-06-10 doc/NEAMS_Report/NEAMS_Report.tex
corrections. did operator stuff. added reminders

@@
 $\nu^g$					&	$=$	&	Total number of neutrons per fission \\
-$D^g$					&	$=$	&	Diffusion coefficient \\
+$D^g$					&	$=$	&	Diffusion coefficient in energy group $g$\\
 $\lambda_i$				&	$=$	&	Decay constant of precursor $i$ \\
@@
 \end{subequations}
-where we haved define the reactivity, effective delayed-neutron fraction, and delayed-neutron precursor decay constant as follows:
+where we have defined the reactivity, effective delayed-neutron fraction, and delayed-neutron precursor decay constant as follows:
 \begin{subequations}
@@
 \end{subequations}
-The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.
-Additionally, in order to impose uniqueness on the factorization and to derive the PRKE the following normalization condition is imposed: $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}$.
-This condition is later used  as the convergence criteria when solving the nonlinear system of equations.
+The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.\\  
+Additionally, in order to impose uniqueness on the factorization and to derive the PRKE, the following normalization condition is imposed: $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}$.
+This condition is later used  as the convergence criteria when solving the nonlinear system of equations resulting from the IQS method.
 
@@
 
-\tcr{introduce operator notation, then switch to transport}
-
-For simplicity and intuitiveness the previous section only described IQS implementation in CFEM diffusion.   \\
-
-The derivation of IQS was done using multi-group diffusion, but the derivation with transport is very similar.
-For simplicity, the transport equation is written in operator form in Equation \ref{eq:transport}.
+For simplicity, the previous section only described the IQS implementation using the neutron diffusion equation.
+However, deriving the IQS form for other neutron balance equations (e.g., transport, simplified transport, etc.) is very similar.
+To this end, we now re-write the neutron conservation equations in operator form in Equation \ref{eq:transport}.
 Factorization is defined as $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ and the resulting shape equation is defined by Equation \ref{eq:tshape}.
@@
 \be
-\frac{1}{v^g}\frac{\partial \Psi^g}{\partial d} = \left(H^g + P_p^g - L^g\right)\Psi^g + S_d^g
+\frac{1}{v^g}\frac{\partial \Psi^g}{\partial t} = \sum_{g'} \left(H^{g'\to g} + P_p^{g' \to g} \right) \Psi^{g'} - L^g\Psi^g + S_d^g
 \label{eq:transport}
 \ee
-\be
-\frac{1}{v^g}\frac{\partial \psi^g}{\partial d} = \left(H^g + P_p^g - L^g - \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right)\psi^g + \frac{1}{p} S_d^g
+where $\Psi^g$ is the multigroup neutron flux (angular flux in the case of transport), $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt neutron production operator, $L^g$ is the loss operator, and $S_d^g$ is the delayed neutron source operator.
+Using \eqt{eq:flux}, the reader may easily obtain the functional form for these operators in the case of a diffusion approximation.
+Next, the factorization $\Psi^g = p \, \psi^g$ is introduced, where the shape is denoted by $\psi^g$, leading to the following shape equations:
+\be
+\frac{1}{v^g}\frac{\partial \psi^g}{\partial t} = \sum_{g'} \left(H^{g' \to g} + P_p^{g' \to g}\right) \psi^{g'} - \left( L^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \psi^g + \frac{1}{p} S_d^g
 \label{eq:tshape}
@@
 \be
-\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},(H^g+P_p^g-L^g)\psi^g\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
+\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},\sum_{g'}(H^{' \to g}g+P_p^{g' \to g}-L^{g'}\delta_{g'g})\psi^{g'}\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
 \label{eq:trmb}
@@
 \ee
+where $P_{d,i}^g$ is the delayed-neutron operator for precursor group $i$.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Additional Contributions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\tcr{I dont think this section is really necessary}
-
-The implementation of IQS in Rattlesnake put pressure on many features of Rattlesnake and MOOSE that reveled bugs and possible improvement.
-Two significant issues in Rattlesnake that were found involved the adjoint solve and the diffusion fission kernel.
-When testing IQS, it was found that the adjoint flux solution was not the same as the forward flux solution in a single group test, which is obviously invalid.
-Also, when investigating which action to include the IQS kernel in, it turned out that the fission diffusion kernel was placed in the neutron transport action, so this kernel was demoted to the neutron diffusion actions for clarity.
-Additionally, the pressure on the save in feature in MOOSE propagated its application to boundary conditions and initial solves.
-MOOSE also updated its ability to restart dense vector data and to set MooseApp executioner right after executioner is created.
-Merge requests: \#5474, \#5489, \#5495, and \#5497
-
-\tcr{is this list exhaustive?}
+\tcr{I am sending you the pretty latex example in an email}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\tcr{I think the time-adaptation results for TWIGL was be a great addition}
+
+\tcr{re-do the figures 4 and 10 as discussed}
 

### 2016-06-10 doc/ANS_winter_2016/ANS_summary.tex
minor comments/suggestions for ans

@@
 \title{Improved Quasi-Static Method with Step Doubling Time Adaptation}
-\author{Zachary M. Prince$^{*}$, Dr. Jean C. Ragusa$^{*}$, Yaqi Wang$^{\dagger}$}
+\author{Zachary M. Prince$^{*}$, Jean C. Ragusa$^{*}$, Yaqi Wang$^{\dagger}$}
 
@@
 %%%% packages and definitions (optional)
+\usepackage{color} % allows inclusion of graphics
 \usepackage{graphicx} % allows inclusion of graphics
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\tcr{you may have to remove yaqi due to short time for approval}
+
 \section{Introduction}
@@
 However, to test it's viability for TREAT simulations, evaluating IQS's performance with time adaptation is required.
-The rest of thispaper will brefly desribe the derivation of IQS, the time adaptation technique used, and results from two kinetic benchmarks.
+The rest of this paper will briefly describe the derivation of IQS, the time adaptation technique used, and results from two kinetic benchmarks.
+\tcr{likely only TWIGL for now. amend as needed}
 
@@
 \section{Theory}
+\tcr{this is a bit long for a summary. How about using my operator section instead. it could be shorter and sweeter. in a summary, there is no way you can give that much details}
 In this Section, we recall the equation for the IQS method, starting from the standard multi-group diffusion equations with delayed neutron precursors in operator form:

### 2016-06-10 doc/ANS_winter_2016/ANS_summary.tex
ANS summary: changed to operator format and deleted LRA

--- 2016-06-05 doc/ANS_winter_2016/ANS_summary.tex
@@
 \title{Improved Quasi-Static Method with Step Doubling Time Adaptation}
-\author{Zachary M. Prince$^{*}$, Dr. Jean C. Ragusa$^{*}$, Yaqi Wang$^{\dagger}$}
+\author{Zachary M. Prince, Jean C. Ragusa}
 
 \institute{
-$^{*}$Department of Nuclear Engineering, Texas A\&M University, College Station, TX
-\and
-$^{\dagger}$Idaho National Laboratory, Idaho Falls, ID
+Department of Nuclear Engineering, Texas A\&M University, College Station, TX
 }
 
-\email{zachmprince@tamu.edu \and jean.ragusa@tamu.edu \and yaqi.wang@inl.gov}
+\email{zachmprince@tamu.edu \and jean.ragusa@tamu.edu}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Introduction}
@@
 However, to test it's viability for TREAT simulations, evaluating IQS's performance with time adaptation is required.
-The rest of thispaper will brefly desribe the derivation of IQS, the time adaptation technique used, and results from two kinetic benchmarks.
+The rest of this paper will briefly describe the derivation of IQS, the time adaptation technique used, and results from a kinetic benchmarks.
 
@@
 \section{Theory}
-In this Section, we recall the equation for the IQS method, starting from the standard multi-group diffusion equations with delayed neutron precursors in operator form:
+In this Section, we recall the equation for the IQS method, starting from the standard multi-group transport equations with delayed neutron precursors in operator form:
 
-\begin{align}
-\frac{1}{v^g} \frac{\partial \phi^g }{\partial t} =& \frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \phi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g  \nonumber \\
-&  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \phi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
+\be
+\frac{1}{v^g}\frac{\partial \Psi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \Psi^{g'} - L^g\Psi^g + S_{d}^g
+\ee 
 \label{eq:flux}
-\end{align}
 \be
-\frac{dC_i}{dt} = \frac{\beta_i}{k_{eff}}\sum_{g=1}^G\nu^{g} \Sigma_f^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
+\frac{dC_i}{dt} = \sum_{g=1}^G P_{d,i}^g \Psi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
 \label{eq:precursor}
@@
 Factorization is an important step in the derivation of the IQS method.
-The factorization approach leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent multigroup shape ($\varphi$):
+The factorization approach leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent multigroup shape ($\psi$):
 \be
-\phi^g(\vec{r},t)=p(t)\varphi^g(\vec{r},t)
+\Psi^g(\vec{r},\Omega,t)=p(t)\psi^g(\vec{r},\Omega,t)
 \ee
@@
 \be
-\int_D\phi^{*g}(\vec{r})f^g(\vec{r})d^3r=\left(\phi^{*g},f^g\right)
+\left(\Psi^{*g},f^g\right) = \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d^2\Omega
 \ee
@@
 \be
-\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}
+\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right) = \textit{constant}
 \ee
@@
 Where the functional coefficients are calculated using the space-/time-dependent shape function as follows:
-\begin{align}
-\frac{\rho-\bar{\beta}}{\Lambda}=&\Bigg[\sum_{g=1}^G \bigg(\phi^{*g},\frac{\chi_p^g}{k_{eff}}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\
-& -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g \bigg)\Bigg] \times \Bigg[\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)\Bigg]^{-1}
+%\begin{align}
+%\frac{\rho-\bar{\beta}}{\Lambda}=&\Bigg[\sum_{g=1}^G \bigg(\phi^{*g},\frac{\chi_p^g}{k_{eff}}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\
+%& -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g \bigg)\Bigg] \times \Bigg[\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)\Bigg]^{-1}
+%\label{eq:rmb}
+%\end{align}
+%\be
+%\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{k_{eff}}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
+%\label{eq:b}
+%\ee
+%\be
+%\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
+%\label{eq:l}
+%\ee
+\be
+\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},\sum_{g'}(H^{' \to g}g+P_p^{g' \to g}-L^{g'}\delta_{g'g})\psi^{g'}\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
 \label{eq:rmb}
-\end{align}
-\be
-\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{k_{eff}}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-\label{eq:b}
 \ee
 \be
-\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
+\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{\sum_{g=1}^G(\Psi^{*g}, P_{d,i}^g \psi^g)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
+\ee
+\be
+\bar{\lambda}_i=\frac{\sum_{g=1}^G(\Psi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
 \label{eq:l}
@@
 \be
-\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
+\psi^g_{n+1} = \underbrace{\Psi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
 \label{eq:rescale}
@@
 \be
-K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
+K_{n+1} =\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\Psi^g_{n+1}\right)
 \ee
 \be
-K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
+K_{0} =\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g_{n+1}\right)=\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\Psi^g_{0}\right)
 \ee
@@
 \be
-\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1}
+\underbrace{\Psi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \psi^g_{n+1}
 \ee
@@
 \be
-e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_2 - \sum_{g=1}^G\varphi^g_1}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_2},\norm{\sum_{g=1}^G\varphi^g_1}\right)}
+e_n = \frac{\norm{\sum_{g=1}^G\int_{4\pi}\Psi^g_2 d^2\Omega - \sum_{g=1}^G\int_{4\pi}\Psi^g_1 d^2\Omega}}{\text{max}\left(\norm{\sum_{g=1}^G\int_{4\pi}\Psi^g_2 d^2\Omega},\norm{\sum_{g=1}^G \int_{4\pi}\Psi^g_1 d^2\Omega}\right)}
 \label{eq:edt2}
@@
 \section{Results and Analysis}
-This section describes results of an examples that tests the IQS implementation and shows its effectiveness on computation speed and accuracy.
-Two examples were selected for this purpose.
-The first is the two-dimensional TWIGL ramp transient benchmark, described further.
-The second is the LRA benchmark, which a two-dimensional, temperature feedback example.
-Each example was solved with regular diffusion (brute force) and IQS P-C.
-Each method was tested with backward-Euler for the TWIGL benchmark and Crank-Nicholson was used for the LRA benchmark.
+This section describes results of an example that tests the IQS implementation and shows its effectiveness on computation speed and accuracy.
+A benchmark example was selected for this purpose.
+The example is the two-dimensional TWIGL ramp transient benchmark, described further.
+The example was solved with regular diffusion (brute force) and IQS P-C with backward-Euler time discretization.
 The performance for each method are represented by the number of macro time steps and solves with step doubling time adaptation.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{LRA Benchmark}
-
-The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{BPB}.
+%\subsection{LRA Benchmark}
+%
+%The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{BPB}.
 It is a super prompt-critical transient.
 To have better understanding on the cross sections given later, we present the equations here:
-\begin{align}
--\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\grad D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 \nonumber \\
-&- \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
--\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\grad D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
-S_f &= \sum_{g=1}^2 \Sigma_{f,g} \phi_g, \\
-\frac{\partial C_i}{\partial t} &= \nu\beta_i f - \lambda_i C_i, \quad i=1,2, \\
-\frac{\partial T}{\partial t} &= \alpha f, \label{eq:lra-temp} \\
-\Sigma_{a,1} &= \Sigma_{a,1}(\vec{r}, t=0) \left[1+\gamma\left(\sqrt{T} - \sqrt{T_0}\right)\right], \\
-P &= \kappa S_f,
-\end{align}
-where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_6f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0=T(\vec{r}, t=0)$.
-The two-group diffusion equation are solved with zero flux boundary conditions on external surfaces, reflecting conditions at symmetry boundaries and steady state initial conditions which are obtained by solving
-\begin{align}
--\grad D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
--\grad D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1
-\end{align}
-The eigenvalue $k$ is used to modify the fission cross section for the transient simulations with $\frac{1}{k}\Sigma_{f,g}, g=1,2$.
+%\begin{align}
+%-\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\grad D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 \nonumber \\
+%&- \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
+%-\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\grad D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
+%S_f &= \sum_{g=1}^2 \Sigma_{f,g} \phi_g, \\
+%\frac{\partial C_i}{\partial t} &= \nu\beta_i f - \lambda_i C_i, \quad i=1,2, \\
+%\frac{\partial T}{\partial t} &= \alpha f, \label{eq:lra-temp} \\
+%\Sigma_{a,1} &= \Sigma_{a,1}(\vec{r}, t=0) \left[1+\gamma\left(\sqrt{T} - \sqrt{T_0}\right)\right], \\
+%P &= \kappa S_f,
+%\end{align}
+%where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_6f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0=T(\vec{r}, t=0)$.
+%The two-group diffusion equation are solved with zero flux boundary conditions on external surfaces, reflecting conditions at symmetry boundaries and steady state initial conditions which are obtained by solving
+%\begin{align}
+%-\grad D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
+%-\grad D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1
+%\end{align}
+%The eigenvalue $k$ is used to modify the fission cross section for the transient simulations with $\frac{1}{k}\Sigma_{f,g}, g=1,2$.
 The initial flux distribution shall be normalized such that the averaged power density
-\begin{align}
-\bar{P} \equiv \frac{\int_{V_{core}} P(\vec{r}, t=0) d\vec{r}}{\int_{V_{core}} d\vec{r}},
-\end{align}
-where $V_{core}$ is the core region with fuels, is equal to $10^{-6} W\cdot cm^{-3}$.
-The initial precursor concentrations are in equilibrium with the initial critical flux distribution.\\
-
-The geometry is illustrated in~\fig{fig:lra-geometry}.\\
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.8\linewidth]{lra.png}
-\caption{LRA benchmark geometry with region assignment.}
-\label{fig:lra-geometry}
-\end{figure}
-
-Initial two-group constants are presented in~\tbl{tab:lra-xs}.
-$\nu$ is equal to 2.43.
-Axial bulking $B^2 = 10^{-4}$ is applied for both energy groups.
-Delayed neutron data are presented in~\tbl{tab:lra-dnp}.
-All fuel materials have the same delayed neutron data.
-Some scalar data are listed in~\tbl{tab:lra-scalar}.\\
-\begin{table}[!htbp]
-    \centering
-    \resizebox{0.5\textwidth}{!}{
-      \begin{tabular}{|c|c|c|c|l|l|l|l|l|}
-      \hline
-             &                 & Group   & $D_g$      & $\Sigma_{a,g}$ & $\nu\Sigma_{f,g}$ & $\Sigma_{s,1\rightarrow 2}$ & $\chi_g$ & $v_g$              \\
-      Region & Material        &       g & (cm)       &  ($cm^{-1}$)   &  ($cm^{-1}$)      &  ($cm^{-1}$)                &          & ($cm\cdot s^{-1}$) \\
-      \hline
-      1      & Fuel 1 with rod & 1       & 1.255      & 0.008252       & 0.004602          &                             & 1        & $3.0\times10^7$    \\
-             &                 & 2       & 0.211      & 0.1003         & 0.1091            & 0.02533                     & 0        & $3.0\times10^5$    \\
-      \hline
-      2      & Fuel 1 without rod & 1    & 1.268      & 0.007181       & 0.004609          &                             & 1        & $3.0\times10^7$    \\
-             &                    & 2    & 0.1902     & 0.07047        & 0.08675           & 0.02767                     & 0        & $3.0\times10^5$    \\
-      \hline
-      3      & Fuel 2 with rod & 1       & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
-             &                 & 2       & 0.2091     & 0.08344        & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
-      \hline
-      4      & Fuel 2 without rod & 1    & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
-             &                    & 2    & 0.2091     & 0.073324       & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
-      \hline
-      5      & Reflector        & 1      & 1.257      & 0.0006034      & -                 &                             & -        & $3.0\times10^7$    \\
-             &                  & 2      & 0.1592     & 0.01911        & -                 & 0.04754                     & -        & $3.0\times10^5$    \\
-      \hline
-      \end{tabular}}
-    \caption{LRA benchmark initial two-group constants.\label{tab:lra-xs}}
-\end{table}
-
-\begin{table}[!htbp]
-    \centering
-      \begin{tabular}{|c|l|l|l|l|}
-      \hline
-      Group i & $\beta_i$ & $\lambda_i$ ($s^{-1}$) & $\chi_{d,i,1}$ & $\chi_{d,i,2}$ \\
-      \hline
-      1       & 0.0054    & 0.0654  & 1 & 0 \\
-      2       & 0.001087  & 1.35    & 1 & 0 \\
-      \hline
-      \end{tabular}
-    \caption{LRA benchmark delayed neutron data.\label{tab:lra-dnp}}
-\end{table}
-
-\begin{table}[!htbp]
-    \centering
-    \resizebox{0.5\textwidth}{!}{
-      \begin{tabular}{|l|c|l|}
-      \hline
-      Meaning & Notation & value \\
-      \hline
-      Axial buckling for both energy groups & $B_g^2$  & $10^{-4}$ ($cm^{-2}$)\\
-      Mean number of neutrons per fission   & $\nu$    & 2.43 \\
-      Conversion factor                     & $\alpha$ & $3.83\times 10^{-11}$ ($K\cdot cm^{3}$) \\
-      Feedback constant                     & $\gamma$ & $3.034\times 10^{-3}$ ($K^{1/2}$) \\
-      Energy released per fission           & $\kappa$ & $3.204\times 10^{-11} $ ($J/fission$) \\
-      Initial and reference temperature     & $T_0$    & 300 (K) \\
-      Active core volume                    & $V_{core}$ & 17550 ($cm^2$)\\
-      \hline
-      \end{tabular}}
-    \caption{LRA benchmark scalar values.\label{tab:lra-scalar}}
-\end{table}
-
-The transient is initiated by changing the thermal absorption cross section as the following:
-\begin{align}
-\Sigma_{a,2}(t) = \Sigma_{a,2}(t=0) \left\{\begin{array}{lr} 1-0.0606184t, & t\leq 2 \\
-                                                             0.8787631, & t>2
-                                           \end{array}\right.
-\end{align}
-where $t$ is time in seconds.\\
-
-Figure \ref{fig:LRA_plots} shows the IQS P-C power profile as compared with the Brute Force solution.
+%\begin{align}
+%\bar{P} \equiv \frac{\int_{V_{core}} P(\vec{r}, t=0) d\vec{r}}{\int_{V_{core}} d\vec{r}},
+%\end{align}
+%where $V_{core}$ is the core region with fuels, is equal to $10^{-6} W\cdot cm^{-3}$.
+%The initial precursor concentrations are in equilibrium with the initial critical flux distribution.\\
+%
+%The geometry is illustrated in~\fig{fig:lra-geometry}.\\
+%\begin{figure}[!htbp]
+%\centering
+%\includegraphics[width=0.8\linewidth]{lra.png}
+%\caption{LRA benchmark geometry with region assignment.}
+%\label{fig:lra-geometry}
+%\end{figure}
+%
+%Initial two-group constants are presented in~\tbl{tab:lra-xs}.
+%$\nu$ is equal to 2.43.
+%Axial bulking $B^2 = 10^{-4}$ is applied for both energy groups.
+%Delayed neutron data are presented in~\tbl{tab:lra-dnp}.
+%All fuel materials have the same delayed neutron data.
+%Some scalar data are listed in~\tbl{tab:lra-scalar}.\\
+%\begin{table}[!htbp]
+%    \centering
+%    \resizebox{0.5\textwidth}{!}{
+%      \begin{tabular}{|c|c|c|c|l|l|l|l|l|}
+%      \hline
+%             &                 & Group   & $D_g$      & $\Sigma_{a,g}$ & $\nu\Sigma_{f,g}$ & $\Sigma_{s,1\rightarrow 2}$ & $\chi_g$ & $v_g$              \\
+%      Region & Material        &       g & (cm)       &  ($cm^{-1}$)   &  ($cm^{-1}$)      &  ($cm^{-1}$)                &          & ($cm\cdot s^{-1}$) \\
+%      \hline
+%      1      & Fuel 1 with rod & 1       & 1.255      & 0.008252       & 0.004602          &                             & 1        & $3.0\times10^7$    \\
+%             &                 & 2       & 0.211      & 0.1003         & 0.1091            & 0.02533                     & 0        & $3.0\times10^5$    \\
+%      \hline
+%      2      & Fuel 1 without rod & 1    & 1.268      & 0.007181       & 0.004609          &                             & 1        & $3.0\times10^7$    \\
+%             &                    & 2    & 0.1902     & 0.07047        & 0.08675           & 0.02767                     & 0        & $3.0\times10^5$    \\
+%      \hline
+%      3      & Fuel 2 with rod & 1       & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
+%             &                 & 2       & 0.2091     & 0.08344        & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
+%      \hline
+%      4      & Fuel 2 without rod & 1    & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
+%             &                    & 2    & 0.2091     & 0.073324       & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
+%      \hline
+%      5      & Reflector        & 1      & 1.257      & 0.0006034      & -                 &                             & -        & $3.0\times10^7$    \\
+%             &                  & 2      & 0.1592     & 0.01911        & -                 & 0.04754                     & -        & $3.0\times10^5$    \\
+%      \hline
+%      \end{tabular}}
+%    \caption{LRA benchmark initial two-group constants.\label{tab:lra-xs}}
+%\end{table}
+%
+%\begin{table}[!htbp]
+%    \centering
+%      \begin{tabular}{|c|l|l|l|l|}
+%      \hline
+%      Group i & $\beta_i$ & $\lambda_i$ ($s^{-1}$) & $\chi_{d,i,1}$ & $\chi_{d,i,2}$ \\
+%      \hline
+%      1       & 0.0054    & 0.0654  & 1 & 0 \\
+%      2       & 0.001087  & 1.35    & 1 & 0 \\
+%      \hline
+%      \end{tabular}
+%    \caption{LRA benchmark delayed neutron data.\label{tab:lra-dnp}}
+%\end{table}
+%
+%\begin{table}[!htbp]
+%    \centering
+%    \resizebox{0.5\textwidth}{!}{
+%      \begin{tabular}{|l|c|l|}
+%      \hline
+%      Meaning & Notation & value \\
+%      \hline
+%      Axial buckling for both energy groups & $B_g^2$  & $10^{-4}$ ($cm^{-2}$)\\
+%      Mean number of neutrons per fission   & $\nu$    & 2.43 \\
+%      Conversion factor                     & $\alpha$ & $3.83\times 10^{-11}$ ($K\cdot cm^{3}$) \\
+%      Feedback constant                     & $\gamma$ & $3.034\times 10^{-3}$ ($K^{1/2}$) \\
+%      Energy released per fission           & $\kappa$ & $3.204\times 10^{-11} $ ($J/fission$) \\
+%      Initial and reference temperature     & $T_0$    & 300 (K) \\
+%      Active core volume                    & $V_{core}$ & 17550 ($cm^2$)\\
+%      \hline
+%      \end{tabular}}
+%    \caption{LRA benchmark scalar values.\label{tab:lra-scalar}}
+%\end{table}
+%
+%
+%The transient is initiated by changing the thermal absorption cross section as the following:
+%\begin{align}
+%\Sigma_{a,2}(t) = \Sigma_{a,2}(t=0) \left\{\begin{array}{lr} 1-0.0606184t, & t\leq 2 \\
+%                                                             0.8787631, & t>2
+%                                           \end{array}\right.
+%\end{align}
+%where $t$ is time in seconds.\\
+%
+%Figure \ref{fig:LRA_plots} shows the IQS P-C power profile as compared with the Brute Force solution.
 These plots show that IQS is consistent with Doppler feedback problems.
@@
 These results show that IQS, again, performs much better than the brute force method by taking significantly less time steps while retaining similar accuracy.
-
-\begin{figure}[!htbp]
-\begin{center}
-\includegraphics[width=0.5\textwidth]{lra_power_profile.jpg}
-\caption{Power level comarison of LRA benchmark between IQS and Brute Force using $e_{max} = 0.01$}
-\label{fig:LRA_plots}
-\end{center}
-\end{figure}
-
-\begin{table}[!htbp]
-\begin{center}
-\resizebox{0.48\textwidth}{!}{
-\begin{tabular}{|l|l|l|l|l|l|l|l|}
-\hline
-  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS P-C} \\
-\hline
-Test & $e_{max}$ & Error & Steps & Solves & Error & Steps & Solves \\
-\hline
-1 &	0.01    &	3.5555e-05 &	145  &	735  &	0.00263068 &	105 &	616  \\
-2 &	0.005   &	4.0364e-05 &	266  &	948  &	0.00160486 &	157 &	814  \\
-3 &	0.001   &	0.00294822 &	585  &	2046 &	1.7527e-05 &	218 &	1062  \\
-4 &	0.0005  &	0.00099778 &	833  &	2908 &	1.4185e-05 &	480 &	2038  \\
-5 &	0.0001  &	0.00019510 &	1793 &	6155 &	6.2903e-06 &	434 &	1458  \\
-6 &	5.0e-05 &	0.00018372 &	2485 &	8511 &	1.5247e-06 &	542 &	1719  \\
-\hline
-
-\end{tabular}}
-\end{center}
-\caption{LRA step doubling results}
-\label{tab:LRAdt2}
-\end{table}
+%
+%\begin{figure}[!htbp]
+%\begin{center}
+%\includegraphics[width=0.5\textwidth]{lra_power_profile.jpg}
+%\caption{Power level comarison of LRA benchmark between IQS and Brute Force using $e_{max} = 0.01$}
+%\label{fig:LRA_plots}
+%\end{center}
+%\end{figure}
+%
+%\begin{table}[!htbp]
+%\begin{center}
+%\resizebox{0.48\textwidth}{!}{
+%\begin{tabular}{|l|l|l|l|l|l|l|l|}
+%\hline
+%  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS P-C} \\
+%\hline
+%Test & $e_{max}$ & Error & Steps & Solves & Error & Steps & Solves \\
+%\hline
+%1 &	0.01    &	3.5555e-05 &	145  &	735  &	0.00263068 &	105 &	616  \\
+%2 &	0.005   &	4.0364e-05 &	266  &	948  &	0.00160486 &	157 &	814  \\
+%3 &	0.001   &	0.00294822 &	585  &	2046 &	1.7527e-05 &	218 &	1062  \\
+%4 &	0.0005  &	0.00099778 &	833  &	2908 &	1.4185e-05 &	480 &	2038  \\
+%5 &	0.0001  &	0.00019510 &	1793 &	6155 &	6.2903e-06 &	434 &	1458  \\
+%6 &	5.0e-05 &	0.00018372 &	2485 &	8511 &	1.5247e-06 &	542 &	1719  \\
+%\hline
+%
+%\end{tabular}}
+%\end{center}
+%\caption{LRA step doubling results}
+%\label{tab:LRAdt2}
+%\end{table}
 
@@
 The purpose of this paper was to show IQS's performance with step doubling time adaptation.
-Two examples were used to quantify the method's performance, including the TWIGL and LRA benchmarks.
+The TWIGL benchmark was used to quantify the method's performance.
 The results from the TWIGL example shows significant improvement from the traditional brute force method.
 This improvement was anticipated because the shape for TWIGL changes very little through the transient, so the PRKE in IQS does most of the work.
-The results for the LRA benchmark showed similar improvement with less time steps for the same accuracy.
-In conclusion, IQS shows impressive performance for complex multi-physics problems with time adaptation.
+In conclusion, IQS shows impressive performance for complex neutronics problems with time adaptation.
 However, to rigorously test IQS, more complex problems including TREAT models and transport problems need to be applied.

### 2016-06-11 doc/ANS_winter_2016/ANS_summary.tex
ANS Summary: Final Changes

@@
 To obtain the amplitude equations, the multigroup equations are multiplied by a weighting function, typically the initial adjoint flux ($\phi^*$), and then integrated over phase-space.
-For brevity, the inner product over space will be represented with parenthetical notation:
-\be
-\left(\Psi^{*g},f^g\right) = \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d^2\Omega
-\ee
-In order to impose uniqueness of the factorization, one requires the following:
-\be
-\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right) = \textit{constant}
-\ee
-After some manipulation, the standard point reactor kinetics equations (PRKE) for the amplitude solution are obtained:
+For brevity, the inner product over space will be represented with parenthetical notation ($\left(\Psi^{*g},f^g\right) = \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d^2\Omega
+$).
+In order to impose uniqueness of the factorization, one requires $\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)$ to be constant.
+And after some manipulation, the standard point reactor kinetics equations (PRKE) for the amplitude solution are obtained:
 \be
@@
 
-To be clear, each solution is re-scaled by $p$ when calculating $e_n$.
-Additionally, if the PJFNK iteration does not converge, the entire step doubling process is repeated with half the time original time step.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results and Analysis}
-This section describes results of an example that tests the IQS implementation and shows its effectiveness on computation speed and accuracy.
-A benchmark example was selected for this purpose.
-The example is the two-dimensional TWIGL ramp transient benchmark, described further.
-The example was solved with regular diffusion (brute force) and IQS P-C with backward-Euler time discretization.
+This section describes results of the TWIGL benchmark that tests the IQS implementation and shows its effectiveness on computation speed and accuracy.
+The benchmark was solved with regular diffusion (brute force) and IQS P-C with backward-Euler time discretization.
 The performance for each method are represented by the number of macro time steps and solves with step doubling time adaptation.
@@
 
-This benchmark problem originates from the Argonne National Lab Benchmark Problem Book \cite{BPB}.
-It is a 2D, 2-group reactor core model with no reflector region shown in Figure \ref{fig:TWIGL_reg}.
-Table \ref{tab:TWIGL_mat} shows the material properties of each fuel region and the ramp perturbation of Material 1.
-
-\begin{figure}[!htbp]
-\begin{center}
-\includegraphics[width=0.5\textwidth]{TWIGL_regions.jpg}
-\caption{TWIGL benchmark problem description}
-\label{fig:TWIGL_reg}
-\end{center}
-\end{figure}
-
-\begin{table}[htp]
-\begin{center}
-\resizebox{0.48\textwidth}{!}{
-\begin{tabular}{lllllll}
-\hline
-  &  &  &  &  &  &  $\underline{\Sigma_s (cm^{-1})} $ \\
-Material & Group & $D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu\Sigma_f (cm^{-1})$ & $\chi$ & $g \rightarrow 2$ \\
-\hline
-1 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.01 \\
-  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.00  \\
-2 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.01  \\
-  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.00  \\
-3 & 1 & 1.3 & 0.008 & 0.003 & 1.0 & 0.01  \\
-  & 2 & 0.5 & 0.050 & 0.060 & 0.0 & 0.00  \\
-\hline
-  & $\nu$ & $v_1 (cm/s)$ & $v_2 (cm/s)$ & $\beta$ & $\lambda (1/s)$ &  \\
-\hline
-  & 2.43 & 1.0E7 & 2.0E5 & 0.0075 & 0.08 & \\
-\hline
- \multicolumn{7}{l}{\footnotesize Material 1 ramp perturbation:} \\
-\multicolumn{7}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (1-0.11667t) \quad t \leq 0.2 s$} \\
-\multicolumn{7}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (0.97666t) \quad t > 0.2 s$} \normalsize
-\end{tabular}}
-\end{center}
-\caption{TWIGL material properties and ramp perturbation}
-\label{tab:TWIGL_mat}
-\end{table}
-
+This benchmark problem originates from the Argonne National Lab Benchmark Problem Book.
+It is a 2D, 2-group reactor core model with no reflector region, the geometry, material properties, and transient perturbation can be seen in \cite{BPB}.
 Figure \ref{fig:TWIGL_power} shows the IQS  solution as compared with the Brute Force solution with minimal time steps.
@@
 
-This work was supported by the Department of Energy and Idaho National Laboratory.  Special thanks to Mark Dehart, NEAMS, and the Moose/Rattlesnake team at INL.
+This work was supported by the Department of Energy and Idaho National Laboratory.  Special thanks to Mark Dehart, Yaqi Wang, NEAMS, and the Moose/Rattlesnake team at INL.
 

### 2016-06-12 doc/NEAMS_Report/NEAMS_Report.tex
NEAMS Report: Lots of changes, only a conclusion is needed

@@
 \usepackage{tikz}
+\usepackage{listings}
+\usepackage[numbers,sort&compress]{natbib}[2010/09/13]
 
@@
 
+% tikz stuff
+\usetikzlibrary{shapes,arrows,positioning}
+\pgfdeclarelayer{background}
+\pgfdeclarelayer{foreground}
+\pgfsetlayers{background,main,foreground}
+\tikzstyle{greenblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=green!20, ultra thick, minimum width=60mm, minimum height=15mm,]
+\tikzstyle{blueblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=blue!20, ultra thick, minimum width=60mm, minimum height=15mm]
+\tikzstyle{reddiamond}=[diamond, draw, align=center, top color=white, bottom color=red!20, ultra thick, minimum width=60mm, aspect=2]
+\newcommand{\tikzback}[5]{
+ \begin{pgfonlayer}{background}
+  \path (#1.west |- #2.north)+(-0.5,0.75) node (a1) {};
+  \path (#3.east |- #4.south)+(+0.5,-0.25) node (a2) {};
+  \path[fill=yellow!10,rounded corners, draw=black!100, dashed] (a1) rectangle (a2);
+   \path (#3.east |- #2.north)+(0,1.25)--(#1.west |- #2.north) node[midway] (#5-n) {};
+   \path (#3.east |- #2.south)+(0,-0.35)--(#1.west |- #2.south) node[midway] (#5-s) {};
+   \path (#1.west |- #2.north)+(-0.75,0)--(#1.west |- #4.south) node[midway] (#5-w) {};
+ \end{pgfonlayer}}
+
+% input stuff
+\lstset{%
+%  language=C++,
+  showstringspaces=false,
+% basicstyle=\scriptsize\ttfamily,
+  basicstyle=\small\ttfamily,
+  commentstyle=\color{inl@green},
+%  keywordstyle=\bfseries,
+%  escapeinside=$$
+%  framesep=0pt,
+%  rulesep=0pt,
+  frame=single
+}
+% Make nice big tildes inside lstlistings instead of
+% tiny ones.
+\lstset{
+    literate={~} {$\sim$}{1}
+}
+
 % ********* Caption Layout ************
@@
 
-\title{Improved Quasi-Static Methods in RATTLESNAKE}
-\subtitle{Development and Implementation of Improved Quasi-Static (IQS) methods for time-dependent neutron diffusion and neutron transport solvers in RATTLESNAKE}
+\title{Improved Quasi-Static Methods in Rattlesnake}
+\subtitle{Development and Implementation of Improved Quasi-Static (IQS) methods for time-dependent neutron diffusion and neutron transport solvers in Rattlesnake}
 \author{ \normalsize
@@
 
-\tcb{I will fix the cover page and page footers later}
-
-\tcb{I removed CFEM from the title}
-
-\tcr{you need Yaqi's concurrence since you list him as co-author}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\tcr{we need a short abstract}
+Because of the recent interest in reactor transient modeling and the restart of the Transient Reactor (TREAT) Facility, there has been a need for more efficient, robust methods in computation frameworks.
+This is the impetus of implementing the Improved Quasi-Static method (IQS) in the Rattlesnake framework.
+IQS has implemented with CFEM diffusion by factorizing flux into time-dependent amplitude and spacial- and weakly time-dependent shape.
+The shape evaluation is very similar to a flux diffusion solve and is computed at large (macro) time steps.
+While the amplitude evaluation is a PRKE solve where the parameters are dependent on the shape and is computed at small (micro) time steps.
+IQS has been tested with a custom one-dimensional example, the TWIGL ramp benchmark, and LRA benchmark.
+The results from each example include error convergence plots and performance with time adaptation.
+These examples prove it to be a viable and effective method for highly transient cases.
+More complex cases are intended to be applied to further test the method and its implementation.
 
@@
 \centering
-\begin{tikzpicture}[
-redbox/.style = {rectangle, rounded corners, draw, align=center, top color=white, bottom color=red!20, ultra thick, minimum size=20mm}, 
-bluebox/.style = {rectangle, rounded corners, draw, align=center, top color=white, bottom color=blue!20, ultra thick, minimum size=20mm},
-greenbox/.style = {rectangle, rounded corners, draw, align=center, top color=white, bottom color=green!20, ultra thick, minimum size=20mm},
-every node/.style       = {font=\large}
-]
-\node[redbox](iqs) at (0,0) {IQS};
-\node[bluebox](shape) at (6,2) {Shape};
-\node[greenbox](prke) at (11,2) {PRKE};
-\node[bluebox](mp) at (6,-2) {Multi-Physics};
-
-\draw[<->,ultra thick,](iqs.north east) -- node[above,sloped] {Picard}(shape.west);
-%\draw[<->,ultra thick](iqs.east) -- node[above] {Picard}(prke.west);
-\draw[<->,ultra thick](iqs.south east) -- node[below,sloped] {Picard} (mp.west);
-
-\draw[->,ultra thick](prke.south) to [bend left=60] node[above,sloped,midway] {$p,\frac{dp}{dt}$}(shape.south);
-\draw[->,ultra thick](shape.east) -- (prke.west) node[above,sloped,midway] {$\rho,\bar{\beta},\bar{\lambda},\Lambda$};
+\begin{tikzpicture}[every node/.style = {font=\normalsize}]
+
+\node[greenblock](p1) at (0,0) {Solve shape equations \\ over macro time step};
+\node[greenblock](p2) at (0,-2.5) {Compute PRKE parameters \\ over macro time step};
+\node[greenblock](p3) at (0,-5) {Solve PRKE using micro \\ steps};
+\node[greenblock](p4) at (0,-7.5) {Update $p$ and $\frac{dp}{dt}$};
+\node[blueblock] (p5) at (0,-11) {Solve other physics \\ components};
+\node[reddiamond] (check) at (0,-15) {Check for \\ convergence};
+\node [above =0.1mm of p1] {\large{IQS Solve:}};
+\node [above =0.1mm of p5] {\large{Multi-physics:}};
+
+\tikzback{p1}{p1}{p4}{p4}{bk1}
+\tikzback{p5}{p5}{p5}{p5}{bk2}
+
+\draw[->,ultra thick](p1.south) -- (p2.north);
+\draw[->,ultra thick](p2.south) -- (p3.north);
+\draw[->,ultra thick](p3.south) -- (p4.north);
+\draw [->,ultra thick] (p4.south)+(0,-0.25) -- node [above] {} (bk2-n);
+\draw[->,ultra thick](p5.south)+(0,-0.25) -- (check.north);
+%\draw[->,ultra thick](check.west) -- (-5,-15) -- node[above,sloped] {\large{no}}(-5,0) -- node[above] {}(bk1-w);
+\draw[->,ultra thick](check.west) -- node[above,sloped] {\large{no}} (-5,-15) |-  node[above] {}(bk1-w);
+\draw[->,ultra thick](check.south) -- node[right] {\large{yes}} ++(0,-2);
 
@@
 Error_{IQS}=\left|\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,n}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,0}\right)}-1\right|
+\label{eq:eiqs}
 \ee
@@
 So another input file and multi-app transfer was made for the adjoint calculation.
-An example input file with detailed commentary can be found in the Appendix.
-
-\tcr{in the yak manual you may find how to incorporate nicely a piece of the input file as a table or figure here.
-the list of parameters and available options for each must be given.
-Maybe an additional appendix should be a smaple input file, with comments}
-
-\tcr{I am sending you the pretty latex example in an email}
+\\
+
+Below is the syntax for the executioner block for an IQS input file.
+The \texttt{predictor\_corrector} logical determines whether to do IQS P-C or regular IQS.
+The \texttt{IQS\_error\_tol} is the tolerance for the IQS error represented by Equation \ref{eq:eiqs} and will at most \texttt{picard\_max\_its} iterations until convergence.
+The \texttt{prke\_scheme} defines the time discretization used for the PRKE evaluation.
+Where \texttt{RK} is SDIRK33, \texttt{CN} is Crank-Nicholson, and \texttt{IE} is implicit-Euler.
+\begin{lstlisting}
+[Executioner]
+  type = IQS
+  predictor_corrector = true/false
+  picard_max_its = 5
+  ...
+  n_micro = 10000
+  IQS_error_tol = 1e-7
+  prke_scheme = 'RK'
+[]
+\end{lstlisting}
+
+Since IQS needs to use the adjoint solution for the PRKE parameter evaluation, auxiliary variables need to be created for each group in the input file.  Below is an example of their definition:
+\begin{lstlisting}
+[AuxVariables]
+  [./adjoint_flux_g0]
+    family = LAGRANGE
+    order = FIRST
+  [../]
+  [./adjoint_flux_g1]
+    family = LAGRANGE
+    order = FIRST
+  [../]
+  ...
+[]
+\end{lstlisting}
+
+Below is the syntax for the multi-app block to perform forward and adjoint steady-state evaluations:
+\begin{lstlisting}
+[MultiApps]
+ [./initial_solve]
+   type = FullSolveMultiApp
+   execute_on = initial
+   input_files = initial.i
+ [../]
+ [./adjoint_solve]
+   type = FullSolveMultiApp
+   execute_on = initial
+   input_files = adjoint.i
+ [../]
+[]
+\end{lstlisting}
+
+Below is the syntax for the Transfer block to copy the initial and adjoint solutions from the multi-apps.
+\begin{lstlisting}
+[Transfers]
+ [./copy_solution]
+   type = TransportSystemVariableTransfer
+   direction = from_multiapp
+   multi_app = initial_solve
+   execute_on = initial
+   from_transport_system = diff
+   to_transport_system = diff
+ [../]
+ [./copy_adjoint]
+   type = MultiAppVariableTransfer
+   execute_on = initial
+   direction = from_multiapp
+   multi_app = adjoint_solve
+   from_variables = 'sflux_g0 sflux_g1 ...'
+   to_variables = 'adjoint_flux_g0 adjoint_flux_g1 ...'
+ [../]
+ [./copy_eigenvalue]
+   type = EigenvalueTransfer
+   execute_on = initial
+   direction = from_multiapp
+   multi_app = initial_solve
+ [../]
+[]
+\end{lstlisting}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\tcr{I think the time-adaptation results for TWIGL was be a great addition}
-
-\tcr{re-do the figures 4 and 10 as discussed}
 
@@
 Figure \ref{fig:slab} how the regions of the slab are divided and Table \ref{tab:1Dmat} shows the initial material properties.
-Table \ref{tab:1Dslope} shows the ramp of the absorption cross-section of each region.
+Regions 2, 3, and 4 have slope perturbations at different points in time, Table \ref{tab:1Dslope} shows the values of the absorption cross-section in each region at the times of interest.
+The values of $\Sigma_a$ between these times of interest are linear interpolations between the given values.
 
@@
   &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
-  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
 1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 4 & 4 & 1 & 1 & 1 & 1 \\
-  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
   &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
@@
 \label{tab:1Dmat}
-\begin{tabular}{lllllll}
-\hline
-Region & $D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu \Sigma_f (cm^{-1})$ & $v (cm/s)$ & $\beta$ & $\lambda (s^{-1})$ \\
-\hline
-1 & 1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\
-2 & 1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\
-3 & 1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\
-4 & 1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\
+\begin{tabular}{llllll}
+\hline
+$D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu \Sigma_f (cm^{-1})$ & $v (cm/s)$ & $\beta$ & $\lambda (s^{-1})$ \\
+\hline
+1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\
+
 \hline
@@
 \begin{center}
-\caption{1-D heterogeneous slab absorption cross-section slope perturbation}
+\caption{1-D heterogeneous slab absorption cross-section at times of interest}
 \label{tab:1Dslope}
-\begin{tabular}{lllllll}
-\hline
-Material Property & 0.0 s & 0.1 s & 0.6 s & 1.0 s & 1.7 s \\
-\hline
-$\Sigma_{a,2} (cm^{-1})$ & 1.1 & 1.1 & 1.095 & 1.095 & 1.095 \\
-$\Sigma_{a,3} (cm^{-1})$ & 1.1 & 1.1 & 1.09 & 1.09 & 1.1 \\
-$\Sigma_{a,4} (cm^{-1})$ & 1.1 & 1.1 & 1.105 & 1.105 & 1.105 \\
+\begin{tabular}{llllllll}
+\hline
+Region & Material Property & 0.0 s & 0.1 s & 0.6 s & 1.0 s & 1.7 s \\
+\hline
+2 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.095 & 1.095 & 1.095 \\
+3 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.09 & 1.09 & 1.1 \\
+4 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.105 & 1.105 & 1.105 \\
 \hline
@@
 Figure \ref{fig:conv} shows that IQS is not only consistent for this example, but also has a better error constant in the convergence study.
-Figures \ref{fig:initial} - \ref{fig:final} plots shape changes in the IQS method, showing where the shape solution is necessary and a simple PRKE evaluation is inadequate.
 Figure \ref{fig:conv} shows the error convergence for brute force and IQS methods.
@@
 
+%Figures \ref{fig:initial} - \ref{fig:final} plots shape changes in the IQS method, showing where the shape solution is necessary and a simple PRKE evaluation is inadequate.
+
 \begin{figure}[!htbp]
 \centering
-\includegraphics[width=0.8\linewidth]{\FiguresDir/1D_het_power_plot.png}
-\caption{Power level of 1D heterogeneous example with $\Delta t = 0.01$}
+\includegraphics[width=0.49\textwidth]{\FiguresDir/1D_het_power_plot.png}
+\includegraphics[width=0.5\textwidth]{\FiguresDir/1D_het_power_plot2.png}
+\caption{Power level of 1D heterogeneous example}
 \label{fig:power}
@@
 
-\begin{figure}[!htbp]
-\begin{center}
-\includegraphics[width=5in,height=2in]{\FiguresDir/initial_flux.jpg}
-\caption{Initial Flux Plot}
-\label{fig:initial}
-\end{center}
-\end{figure}
-
-\begin{figure}[!htbp]
-\begin{center}
-\includegraphics[width=5in,height=2in]{\FiguresDir/max_flux.jpg}
-\caption{Flux Plot when Absorption Cross Section is at Minimum}
-\label{fig:max}
-\end{center}
-\end{figure}
-
-\begin{figure}[!htbp]
-\begin{center}
-\includegraphics[width=5in,height=2in]{\FiguresDir/final_flux.jpg}
-\caption{Final Flux Computation (not steady-state)}
-\label{fig:final}
-\end{center}
-\end{figure}
+%\begin{figure}[!htbp]
+%\begin{center}
+%\includegraphics[width=5in,height=2in]{\FiguresDir/initial_flux.jpg}
+%\caption{Initial Flux Plot}
+%\label{fig:initial}
+%\end{center}
+%\end{figure}
+%
+%\begin{figure}[!htbp]
+%\begin{center}
+%\includegraphics[width=5in,height=2in]{\FiguresDir/max_flux.jpg}
+%\caption{Flux Plot when Absorption Cross Section is at Minimum}
+%\label{fig:max}
+%\end{center}
+%\end{figure}
+%
+%\begin{figure}[!htbp]
+%\begin{center}
+%\includegraphics[width=5in,height=2in]{\FiguresDir/final_flux.jpg}
+%\caption{Final Flux Computation (not steady-state)}
+%\label{fig:final}
+%\end{center}
+%\end{figure}
 
@@
 
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-This benchmark problem originates from the Argonne National Lab Benchmark Problem Book.
+This benchmark problem originates from the Argonne National Lab Benchmark Problem Book \cite{ANL_BPB}.
 It is a 2D, 2-group reactor core model with no reflector region shown in Figure \ref{fig:TWIGL_reg}.
@@
 
-Figures \ref{fig:TWIGL_power} and \ref{fig:TWIGL_plots} show the IQS  solution as compared with the Brute Force solution.
-It is important to note the IQS shape plot is scaled differently than the Brute Force flux plot because the amplitude term is not included, but the gradients of colors is comparable.
-These plots show that IQS is consistent in more complex, higher dimensional problems in RATTLESNAKE.
+Figures \ref{fig:TWIGL_power} show the IQS  solution as compared with the Brute Force solution.
+These plots show that IQS is consistent in more complex, higher dimensional problems in Rattlesnake.
 Finally, Figure \ref{fig:TWIGL_conv} plots the error convergence of IQS and the Brute Force methods.
 The curves show the impressive convergence of IQS for the highly transient TWIGL example.
-Table \ref{tab:TWIGLdt2} shows the results for time adaptation.
+Table \ref{tab:TWIGLdt2} and Figure \ref{fig:TWIGL_power_dt2} show the results for time adaptation.
 The results show that both IQS methods perform exceptionally well compared to brute force.
@@
 
+%It is important to note the IQS shape plot is scaled differently than the Brute Force flux plot because the amplitude term is not included, but the gradients of colors is comparable.
+
 \begin{figure}[!htbp]
 \centering
-\includegraphics[height=3.0in]{\FiguresDir/TWIGL_power_plot.png}
-\caption{Power level comparison of TWIGL Benchmark with $\Delta t = 0.025$}
+\includegraphics[width=0.49\linewidth]{\FiguresDir/TWIGL_power_plot.png}
+\includegraphics[width=0.5\linewidth]{\FiguresDir/TWIGL_power_plot2.png}
+\caption{Power level comparison of TWIGL Benchmark}
 \label{fig:TWIGL_power}
@@
 
-\begin{figure}[!htbp]
-\begin{center}
-\begin{subfigure}[!htbp]{0.4\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/ndiff_ramp_flux.jpg}
-\caption{Brute force flux}
-\end{subfigure}
-\quad
-\begin{subfigure}[!htbp]{0.4\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/iqs_ramp_shape.jpg}
-\caption{IQS Shape}
-\end{subfigure}
-\caption{TWIGL Benchmark flux/shape comparison at $t=0.2$}
-\label{fig:TWIGL_plots}
-\end{center}
-\end{figure}
+%\begin{figure}[!htbp]
+%\begin{center}
+%\begin{subfigure}[!htbp]{0.4\textwidth}
+%\includegraphics[width=\textwidth]{\FiguresDir/ndiff_ramp_flux.jpg}
+%\caption{Brute force flux}
+%\end{subfigure}
+%\quad
+%\begin{subfigure}[!htbp]{0.4\textwidth}
+%\includegraphics[width=\textwidth]{\FiguresDir/iqs_ramp_shape.jpg}
+%\caption{IQS Shape}
+%\end{subfigure}
+%\caption{TWIGL Benchmark flux/shape comparison at $t=0.2$}
+%\label{fig:TWIGL_plots}
+%\end{center}
+%\end{figure}
 
@@
 
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.8\linewidth]{\FiguresDir/TWIGL_power_plot_dt2.png}
+\caption{Power level comparison of TWIGL Benchmark with time adaptation}
+\label{fig:TWIGL_power_dt2}
+\end{figure}
+
 \newpage
@@
 
-The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor.
+The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.
 It is a super prompt-critical transient.
@@
 Figure \ref{fig:LRA_plots} show the IQS  solution as compared with the Brute Force solution.
-These plots show that IQS is consistent with Doppler feedback problems in RATTLESNAKE.
+These plots show that IQS is consistent with Doppler feedback problems in Rattlesnake.
 Figure \ref{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods.
@@
 \caption{Power Profile}
+\end{subfigure}
+\begin{subfigure}[!htbp]{0.49\textwidth}
+\includegraphics[width=\textwidth]{\FiguresDir/lra_power_profile2.png}
+\caption{Power Profile at Peak}
 \end{subfigure}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Appendix: aaaa}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\tcr{putting stuff here for now}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-It can be shown that the most appropriate weighting function $(w^g)$ is the initial adjoint flux $(\phi^{*g})$.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Appendix: aaaa}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\tcr{putting stuff here for now}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%It can be shown that the most appropriate weighting function $(w^g)$ is the initial adjoint flux $(\phi^{*g})$.
 For brevity, the following definition will be applied: $\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r=\left(\phi^{*g},f\right)$
-\begin{align}
-& \quad \frac{dp}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)+ p\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)= \nonumber \\
-& p\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)  \\
-& + \sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i) \nonumber
-\end{align}
-\be
-\frac{d}{dt}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)=\frac{1}{\keff}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})p - \sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \quad 1 \le i \le I
-\ee
-
-In order to impose uniqueness of the factorization, one imposes:
-\be
-\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=constant
-\ee
-
-Therefore the PRKE formulation reduces to:
-\begin{align}
-\frac{dp}{dt}= &\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g}  \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p \nonumber \\
-& +\sum_{i=1}^I\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i
-\end{align}
-\be
-\frac{d\xi_i}{dt}=\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p - \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i \quad 1 \le i \le I
-\ee
-Where:
-\be
-\xi_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g C_i)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-\ee
-
-It is convenient to define the effective reactivity, delay-neutron fraction, and delayed-neutron precursor decay constant:
-\be
-\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-%\label{eq:rmb}
-\ee
-\be
-\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-%\label{eq:b}
-\ee
-\be
-\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
-%\label{eq:l}
-\ee
-
-So:
-\be
-\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
-%\label{eq:p}
-\ee
-\be
-\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
-%\label{eq:c}
-\ee
-
-Equations (\ref{eq:p}) and (\ref{eq:c}) are a formulation of the point reactor kinetics equation (PRKE), but the parameters (Equations (\ref{eq:rmb})-(\ref{eq:l})) are dependent on the shape.
+%\begin{align}
+%& \quad \frac{dp}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)+ p\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)= \nonumber \\
+%& p\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)  \\
+%& + \sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i) \nonumber
+%\end{align}
+%\be
+%\frac{d}{dt}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)=\frac{1}{\keff}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})p - \sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \quad 1 \le i \le I
+%\ee
+%
+%In order to impose uniqueness of the factorization, one imposes:
+%\be
+%\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=constant
+%\ee
+%
+%Therefore the PRKE formulation reduces to:
+%\begin{align}
+%\frac{dp}{dt}= &\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g}  \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p \nonumber \\
+%& +\sum_{i=1}^I\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i
+%\end{align}
+%\be
+%\frac{d\xi_i}{dt}=\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p - \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i \quad 1 \le i \le I
+%\ee
+%Where:
+%\be
+%\xi_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g C_i)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
+%\ee
+%
+%It is convenient to define the effective reactivity, delay-neutron fraction, and delayed-neutron precursor decay constant:
+%\be
+%\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
+%%\label{eq:rmb}
+%\ee
+%\be
+%\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
+%%\label{eq:b}
+%\ee
+%\be
+%\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
+%%\label{eq:l}
+%\ee
+%
+%So:
+%\be
+%\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
+%%\label{eq:p}
+%\ee
+%\be
+%\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
+%%\label{eq:c}
+%\ee
+%
+%Equations (\ref{eq:p}) and (\ref{eq:c}) are a formulation of the point reactor kinetics equation (PRKE), but the parameters (Equations (\ref{eq:rmb})-(\ref{eq:l})) are dependent on the shape.
 If the assumption is made that the shape is time-independent, the shape is computed once at the first time step and used for the PRKE parameter evaluation at all other steps.
@@
 These differences are crucial for IQS implementation in Rattlesnake.
-
-\begin{align}
-\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\ 
-& -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
-%\label{eq:shape2}
-\end{align}
-\be
-\frac{dC_i}{dt} = \frac{\beta_i}{\keff}p \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
-%\label{eq:preq2}
-\ee
+%
+%\begin{align}
+%\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\
+%& -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
+%%\label{eq:shape2}
+%\end{align}
+%\be
+%\frac{dC_i}{dt} = \frac{\beta_i}{\keff}p \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
+%%\label{eq:preq2}
+%\ee
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{ans}
+\bibliographystyle{abbrvnat}
 \bibliography{references_IQS}

### 2016-06-13 doc/NEAMS_Report/NEAMS_Report.tex
NEAMS report: conclusion and other edits

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Because of the recent interest in reactor transient modeling and the restart of the Transient Reactor (TREAT) Facility, there has been a need for more efficient, robust methods in computation frameworks.
 This is the impetus of implementing the Improved Quasi-Static method (IQS) in the Rattlesnake framework.
-IQS has implemented with CFEM diffusion by factorizing flux into time-dependent amplitude and spacial- and weakly time-dependent shape.
-The shape evaluation is very similar to a flux diffusion solve and is computed at large (macro) time steps.
+IQS has implemented by factorizing flux into time-dependent amplitude and spacial- and weakly time-dependent shape.
+The shape evaluation is very similar to a flux solve and is computed at large (macro) time steps.
 While the amplitude evaluation is a PRKE solve where the parameters are dependent on the shape and is computed at small (micro) time steps.
@@
 $C_i$					&	$=$	&	Concentration of delayed neutron precursor $i$ \\
-$\keff$					&	$=$	&	Effective multiplication factor \\
 $\Sigma_f^{g}$			&	$=$	&	Fission cross section in energy group $g$ \\
@@
 To derive the amplitude equation (also known as the PRKE equations), the shape/precursors equations are weighed by a space-dependent function (typically the initial adjoint flux, $\phi^{*g}$) and integrated over the phase-space.
-The final expressions are given below (the derivations are given in appendix for the interested reader).
+The final expressions are given below:
 \begin{subequations}
@@
 However, deriving the IQS form for other neutron balance equations (e.g., transport, simplified transport, etc.) is very similar.
-To this end, we now re-write the neutron conservation equations in operator form in Equation \ref{eq:transport}.
-Factorization is defined as $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ and the resulting shape equation is defined by Equation \ref{eq:tshape}.
-Finally, the PRKE parameters are defined by Equations \ref{eq:trmb} and \ref{eq:tb}, where $\left(\Psi^{*g},f^g\right) = \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d\Omega$.
+To this end, we now re-write the neutron conservation equations in operator form in \eqt{eq:transport}.
+Factorization is defined as $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ and the resulting shape equation is defined by \eqt{eq:tshape}.
+Finally, the PRKE parameters are defined by Equations~\ref{eq:trmb}~and~\ref{eq:tb}, where $\left(\Psi^{*g},f^g\right) = \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d\Omega$.
 \be
@@
 In transients where the shape varies much less than the flux, IQS can thus be very computationally effective.
-The two time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in Figure \ref{fig:iqsviz}.
+The two time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.
 
@@
 
-By this design, the benefits of IQS is most realized when using time adaptive techniques, one of these is described in Section \ref{sect:dt2}.
+By this design, the benefits of IQS is most realized when using time adaptive techniques, one of these is described in  \sct{sect:dt2}.
 As we noted earlier, shape-PRKE equations are a nonlinear system and thus need to be solved in a iterative manner.
@@
 Picard iteration is predominantly utilized for multiphysics simulation, where neutronics is nonlinearly coupled to thermal-hydraulics.
-So the shape-PRKE couple is easily fit into this process, described by Figure \ref{fig:picard}.
+So the shape-PRKE couple is easily fit into this process, described by \fig{fig:picard}.
 
@@
 The Predictor-Corrector version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the evaluation of the coupled system is different.
-This version first solves the flux diffusion (represented by Equations \ref{eq:flux} and \ref{eq:precursor}) to get a predicted flux.
+This version first solves the flux diffusion (represented by Equations~\ref{eq:flux}~and~\ref{eq:precursor}) to get a predicted flux.
 The predicted flux at this step is then converted to shape by rescaling as follows:
@@
 
-The PRKE parameters are then computed with this shape using Equations \ref{eq:rmb} - \ref{eq:l} and interpolated over the macro step, then the PRKE is evaluated.
+The PRKE parameters are then computed with this shape using Equations~\ref{eq:rmb}~-~\ref{eq:l} and interpolated over the macro step, then the PRKE is evaluated.
 With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
@@
 
-This section presents two different time-integration methods to solve coupled IQS shape + precursor equations, recalled below using, for simplicity, a single neutron group and a single precursor group.
+This section presents two different time-integration methods to solve coupled IQS shape and precursor equations, recalled below using, for simplicity, a single neutron group and a single precursor group.
 
@@
 The method entails linearly interpolating the fission source ($S_f = \nu \Sigma_f \varphi$) and integrating the resulting continuous function analytically.
-Equation \ref{eq:c_dis} is the equation to solve for the precursors and Equations \ref{eq:a1} and \ref{eq:a2} are integrating factors.
+\eqt{eq:c_dis} is the equation to solve for the precursors and Equations~\ref{eq:a1}~and~\ref{eq:a2} are integrating factors.
 
@@
 If the step is too large, the step will be repeated with a smaller step calculated with the resulting error.
-The error of the time step is approximated by Equation \ref{eq:edt2}, where $\varphi^g_1$ and $\varphi^g_2$ are the solutions with the full step and half step, respectively.
+The error of the time step is approximated by \eqt{eq:edt2}, where $\varphi^g_1$ and $\varphi^g_2$ are the solutions with the full step and half step, respectively.
 $\varphi$ is changed to $\phi$ for regular flux evaluation and IQS P-C.
 If $e_n > e_{max}$ the time step is repeated; if $e_n < e_{max}$ the system moves to the next time step.
-The next $\Delta t$ is calculated using Equation \ref{eq:dt2}, where $\mu$ is the convergence order of the time integration scheme being used.
+The next $\Delta t$ is calculated using \eqt{eq:dt2}, where $\mu$ is the convergence order of the time integration scheme being used.
 $e_{max}$ and $e_{tol}$ are user defined parameters; $e_{max}$ is usually less than $e_{tol}$ to better guarantee that the calculated $\Delta t_{new}$ will pass the error criteria so time steps won't be repeated.
@@
 
-To be clear, each solution undergoes Picard iterations until $Error_{IQS}$ converges before the error is calculated; for IQS P-C, each solution is re-scaled by $p$.
+To be clear, each solution undergoes Picard iterations until convergence before the error is calculated; for IQS P-C, each solution is re-scaled by its corresponding $p$.
 Additionally, if the PJFNK iteration does not converge, the entire step doubling process is repeated with half the time original time step.
@@
 The Rattlesnake action system gives the capability of consolidating MOOSE input syntax, so that a user does not have to define every kernel, variable, etc.
-in the problem, which could be on the order of thousands with multi-group problems.
+in the problem, which could be on the order of thousands with multigroup problems.
 The user, instead, is able to input an equation description (Diffusion, S$_n$, P$_n$, etc.) and solve method (SAAF,LS,CFEM,DFEM,etc.), and the action system will incorporate all the necessary physics involved (kernels, boundary conditions, postprocessor, etc.).
@@
 The IQS executioner contains a loop over micro time steps that computes the PRKE and then passes $p$ and $\frac{dp}{dt}$ for the Transient executioner to evaluate the shape equation at each macro step.
-The PRKE is computed with a user specified option of backward-Euler, Crank-Nicolson, or SDIRK33.
+The PRKE is computed with a user specified option of backward-Euler, Crank-Nicholson, or SDIRK33.
 The IQS executioner also supplements Transient Picard iteration process by adding its own error criteria:
@@
 Also, the precursor kernel was modified to include the $\frac{1}{p}$ term.
-Finally, the precursor auxkernel that evaluates Equation \ref{eq:preq2} using the analytical integration method described in Section \ref{sect:prec}.
+Finally, the precursor auxkernel that evaluates \eqt{eq:preq2} using the analytical integration method described in \sct{sect:prec}.
 
@@
 
-In order to evaluate the PRKE coefficients, defined by Equations \ref{eq:rmb} - \ref{eq:l}, four postprocessors were created.
+In order to evaluate the PRKE coefficients, defined by Equations~\ref{eq:rmb}~-~\ref{eq:l}, four postprocessors were created.
 The parameter calculations were separated by $\frac{\bar{\beta}_i}{\Lambda}$ numerator, $\bar{\lambda}_i$ numerator/denominator, $\frac{\rho-\bar{\beta}}{\Lambda}/\frac{\bar{\beta}}{\Lambda}$ denominator, and $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator.
@@
 However, Rattlesnake has other action systems capable of transient simulation, where IQS can be implemented and be effective.
-One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equation \ref{eq:flux} and \ref{eq:shape2}.
-However, in the derivation for IQS is term is unaffected between shape and flux evaluation.
-So saving the residual for this diffusion kernel in the \texttt{save\_in} variable is the only alteration to this action for IQS to function.
-
-Additional action systems involve transport, but it is evident from Section \ref{sect:transport} that IQS implementation in these is straightforward.
+One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equations~\ref{eq:flux}~and~\ref{eq:shape2}.
+However, in the derivation for IQS, this term is unaffected between shape and flux evaluation.
+So saving the residual for this diffusion kernel in the \texttt{save\_in} variable is the only alteration to this action for IQS to function.\\
+
+Additional action systems involve transport, but it is evident from \sct{sect:transport} that IQS implementation in these is straightforward.
 The main differences between diffusion implementation in transport is outlined below:
@@
 In order to preserve the already implemented standard version of IQS, an option in the IQS executioner was created to specify which method is desired.
-Because the diffusion solve is flux instead of shape, when predictor-corrector option is specied, the IQS removal kernel ($\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$) and the modified precursor kernel are bypassed, while all the postprocessors are still executed.
+Because the diffusion solve is flux instead of shape, when predictor-corrector option is specified, the IQS removal kernel ($\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$) and the modified precursor kernel are bypassed, while all the postprocessors are still executed.
 However, it is difficult to rescale the flux to shape before the PRKE parameter postprocessor are executed.
@@
 As seen in parameter definitions, when shape is replaced with flux, the amplitude comes out of the integral and cancels out.
-So the conversion of the predicted flux in Equation \ref{eq:rescale} to shape is unnecessary if the corrected flux is solved with Equation \ref{eq:phicorr2}.
+So the conversion of the predicted flux in  \eqt{eq:rescale} to shape is unnecessary if the corrected flux is solved with \eqt{eq:phicorr2}.
 After obtaining the corrected flux, the precursors are re-evaluated using a \texttt{EXEC\_LINEAR}.
@@
 The \texttt{predictor\_corrector} logical determines whether to do IQS P-C or regular IQS.
-The \texttt{IQS\_error\_tol} is the tolerance for the IQS error represented by Equation \ref{eq:eiqs} and will at most \texttt{picard\_max\_its} iterations until convergence.
+The \texttt{IQS\_error\_tol} is the tolerance for the IQS error represented by \eqt{eq:eiqs} and will at most \texttt{picard\_max\_its} iterations until convergence.
 The \texttt{prke\_scheme} defines the time discretization used for the PRKE evaluation.
@@
 
-This section describes results of an examples that tests the IQS implementation and shows its effectiveness on computation speed and accuracy.
+This section describes results of examples that test the IQS implementation and shows its effectiveness on computation speed and accuracy.
 Three examples were selected for this purpose.
@@
 The performance for each method are represented by convergence plots with constant $\Delta t$ and by the number of time steps with step doubling time adaptation.
+For time adaptation, backward-Euler time discretization was used and, for a safety factor, $e_{tol} = 0.8e_{max}$.
 
@@
 The example is very simple and computes quickly, it entails a one dimensional, heterogeneous 400 cm slab with a varying absorption cross section.
-Figure \ref{fig:slab} how the regions of the slab are divided and Table \ref{tab:1Dmat} shows the initial material properties.
-Regions 2, 3, and 4 have slope perturbations at different points in time, Table \ref{tab:1Dslope} shows the values of the absorption cross-section in each region at the times of interest.
+\fig{fig:slab} shows how the regions of the slab are divided and \tbl{tab:1Dmat} shows the initial material properties.
+Regions 2, 3, and 4 have slope perturbations at different points in time, \tbl{tab:1Dslope} shows the values of the absorption cross-section in each region at the times of interest.
 The values of $\Sigma_a$ between these times of interest are linear interpolations between the given values.
@@
 
-Figure \ref{fig:power} shows the power at each macro time step as compared to the traditional brute force (full flux time discretization) method.
+\fig{fig:power} shows the power at each macro time step as compared to the traditional brute force (full flux time discretization) method.
 The strong correlation between the two curves shows that IQS is consistent with a proven method for a highly transient example.
-Figure \ref{fig:conv} shows that IQS is not only consistent for this example, but also has a better error constant in the convergence study.
-Figure \ref{fig:conv} shows the error convergence for brute force and IQS methods.
+\fig{fig:conv} shows that IQS is not only consistent for this example, but also has a better error constant in the convergence study.
+\fig{fig:conv} shows the error convergence for brute force and IQS methods.
 This plot shows that IQS has consistent convergence slope for first and second order time schemes and that IQS has a better error constant for the convergence.
-Table \ref{tab:1Ddt2} shows the results for time adaptation.
+\tbl{tab:1Ddt2} shows the results for time adaptation.
 These results show that traditional IQS does not perform well with step doubling time adaptation and IQS P-C performs moderately better than brute force.
@@
 This benchmark problem originates from the Argonne National Lab Benchmark Problem Book \cite{ANL_BPB}.
-It is a 2D, 2-group reactor core model with no reflector region shown in Figure \ref{fig:TWIGL_reg}.
-Table \ref{tab:TWIGL_mat} shows the material properties of each fuel region and the ramp perturbation of Material 1.
+It is a 2D, 2-group reactor core model with no reflector region shown in \fig{fig:TWIGL_reg}.
+\tbl{tab:TWIGL_mat} shows the material properties of each fuel region and the ramp perturbation of Material 1.
 
@@
 \begin{center}
-\caption{1-D heterogeneous slab absorption cross-section slope perturbation}
+\caption{TWIGL benchmark material properties and slope perturbation}
 \label{tab:TWIGL_mat}
@@
 
-Figures \ref{fig:TWIGL_power} show the IQS  solution as compared with the Brute Force solution.
+\fig{fig:TWIGL_power} show the IQS  solution as compared with the Brute Force solution.
 These plots show that IQS is consistent in more complex, higher dimensional problems in Rattlesnake.
-Finally, Figure \ref{fig:TWIGL_conv} plots the error convergence of IQS and the Brute Force methods.
+Finally, \fig{fig:TWIGL_conv} plots the error convergence of IQS and the Brute Force methods.
 The curves show the impressive convergence of IQS for the highly transient TWIGL example.
-Table \ref{tab:TWIGLdt2} and Figure \ref{fig:TWIGL_power_dt2} show the results for time adaptation.
+\tbl{tab:TWIGLdt2} and \fig{fig:TWIGL_power_dt2} show the results for time adaptation.
 The results show that both IQS methods perform exceptionally well compared to brute force.
@@
 
-Figure \ref{fig:LRA_plots} show the IQS  solution as compared with the Brute Force solution.
+The \fig{fig:LRA_plots} plots show the IQS  solution as compared with the Brute Force solution.
 These plots show that IQS is consistent with Doppler feedback problems in Rattlesnake.
-Figure \ref{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods.
-Table  \ref{tab:LRAdt2} shows the time adaptation performance of both methods.
+\fig{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods.
+\tbl{tab:LRAdt2} shows the time adaptation performance of brute force and IQS P-C.
 These results show that IQS, again, performs much better than the brute force method by taking significantly less diffusion solves while retaining similar accuracy.
@@
 \centering
-\includegraphics[width=5in]{\FiguresDir/lra_convergence.jpg}
+\includegraphics[width=0.6\linewidth]{\FiguresDir/lra_convergence.jpg}
 \caption{Error convergence comparison of LRA Benchmark}
@@
 
-\tcr{needed}
+The purpose of this paper was to present the improved quasi-static method with its derivation, its implementation into Rattlesnake, and its effectiveness with several test cases.
+The derivation of IQS was done with multigroup diffusion, but it was seen in \sct{sect:transport} that application to transport is relatively simple.
+Two different solving methods were introduced.
+The first was the traditional IQS method where the shape equation and PRKE were solved as a coupled nonlinear system with Picard iterations.
+The second was IQS P-C where flux is solved and then rescaled using results from evaluating the PRKE.
+IQS P-C is a much simpler method and, computationally, much less expensive to perform.
+\\
+
+IQS's implementation into Rattlesnake involved creating a new kernel, auxiliary kernel, executioner, and four postprocessors, as well as alterations to the action system.
+The new kernel implements the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term in the shape equation.
+The auxiliary kernel alters how the precursors are solved, the method is outlined by \sct{sect:prec}.
+The executioner is where the PRKE is evaluated after shape solves are performed.
+The postprocessors are used to evaluate the integrals for the PRKE parameters and a user object is used to pass them to the executioner.
+IQS's implementation the action system made sure that all the correct kernels and postprocessors are executed when the IQS executioner is specified in the input.
+\\
+
+IQS was tested by comparing its performance with traditional "brute force" method using error convergence plots and time adaptation.
+Three examples were used including a custom one-dimensional example, the TWIGL benchmark, and LRA benchmark.
+The one-dimensional example shows large shape changes in the flux and proves consistency of the shape evaluation.
+IQS shows proper error convergence for first and second order time discretization.
+IQS P-C shows marginal improvement with time adaptation, but traditional IQS shows poor performance.
+The TWIGL benchmark shows the effectiveness of the PRKE in IQS because there is very little shape change through the transient.
+IQS shows very impressive performance with both constant time steps and time adaptation.
+The LRA benchmark tests IQS when applied with multiphysics.
+Again, IQS shows impressive performance by requiring half the time steps as brute force for a given error.
 
@@
 \newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliographystyle{abbrvnat}
+\bibliography{references_IQS}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{abbrvnat}
-\bibliography{references_IQS}
+
 \end{document}

### 2016-06-14 doc/NEAMS_Report/NEAMS_Report.tex
updated abstract

@@
 
-Because of the recent interest in reactor transient modeling and the restart of the Transient Reactor (TREAT) Facility, there has been a need for more efficient, robust methods in computation frameworks.
-This is the impetus of implementing the Improved Quasi-Static method (IQS) in the Rattlesnake framework.
-IQS has implemented by factorizing flux into time-dependent amplitude and spacial- and weakly time-dependent shape.
-The shape evaluation is very similar to a flux solve and is computed at large (macro) time steps.
-While the amplitude evaluation is a PRKE solve where the parameters are dependent on the shape and is computed at small (micro) time steps.
-IQS has been tested with a custom one-dimensional example, the TWIGL ramp benchmark, and LRA benchmark.
-The results from each example include error convergence plots and performance with time adaptation.
-These examples prove it to be a viable and effective method for highly transient cases.
-More complex cases are intended to be applied to further test the method and its implementation.
+This report presents  the implementation of the Improved Quasi-Static method (IQS) in the Yak/Rattlesnake applications of the MOOSE framework.
+In the IQS method, the neutron flux is factored into a time-dependent amplitude and a spatial- and weakly time-dependent shape.
+The shape evaluation is very similar to a standard flux solve and is computed at large (macro) time steps.
+The amplitude is obtained by solving the PRKEs on a micro time scale.
+The PRKE parameters are regularly updated based on the shape changes.
+Two versions of the IQS methodology have been implemented: (i) the standard approach, which leads to a system of nonlinear equations between the shape and the amplitude, and (ii) a predictor corrector approach that linearizes the system of equations.
+A time-adaptation technique has been developed to automatically select an appropriate macro time step size based on user-specified tolerances.
+The implementation has been verified and tested using manufactured solutions, a custom one-dimensional example, the TWIGL ramp benchmark, and the LRA benchmark.
+The numerical results include error convergence plots and performance with time adaptation.
+These examples prove it to be a viable and effective method for  transient neutronic simulations.
+More complex simulations (e.g., TREAT transients) are planned to perform coupled neutronics/heat conduction transient simulations within the MAMMOTH application.
 

### 2016-06-14 doc/NEAMS_Report/NEAMS_Report.tex
worked on intro

@@
 The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates.
-The IQS method has primarily been developed in the context of the neutron diffusion approximation \cite{Ban_2012,Ferguson_1973}, but \cite{Devooght_1980} provides an extension of the technique to transport, implemented in TDTORT.
+The IQS method has primarily been developed in the context of the neutron diffusion approximation \cite{Ban_2012,Ferguson_1973}, but \cite{Goluoglu_2001} provides an extension of the technique to transport, implemented in TDTORT.
 The method has mostly being applied to neutron kinetics, without thermal-hydraulic feedback.
@@
 \ee
-The factorization is not an approximation. Reporting it in the flux and precursor equations yields, after some simple algebra, the following equations for the shape and the precursors:
+We stress that the factorization is not an approximation. Reporting it in the flux and precursor equations yields, after some simple algebra, the following equations for the shape and the precursors:
 \begin{subequations}
@@
 \frac{1}{v^g}\frac{\partial\varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\ 
-& -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
+& -  \left( -\div D^g \grad  + \Sigma_r^g + \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right) \varphi^g + \boxed{\frac{1}{p}}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
 \label{eq:shape2}
@@
 \be
-\frac{dC_i}{dt} = \frac{\beta_i}{\keff}p \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
+\frac{dC_i}{dt} = \frac{\beta_i}{\keff}\boxed{p} \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
 \label{eq:preq2}
@@
 However, deriving the IQS form for other neutron balance equations (e.g., transport, simplified transport, etc.) is very similar.
-To this end, we now re-write the neutron conservation equations in operator form in \eqt{eq:transport}.
-Factorization is defined as $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ and the resulting shape equation is defined by \eqt{eq:tshape}.
-Finally, the PRKE parameters are defined by Equations~\ref{eq:trmb}~and~\ref{eq:tb}, where $\left(\Psi^{*g},f^g\right) = \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d\Omega$.
+To this end, we re-write the neutron conservation equations in operator form, shown in \eqt{eq:transport}.
 \be
@@
 Using \eqt{eq:flux}, the reader may easily obtain the functional form for these operators in the case of a diffusion approximation.
-Next, the factorization $\Psi^g = p \, \psi^g$ is introduced, where the shape is denoted by $\psi^g$, leading to the following shape equations:
+%
+%Factorization is defined as $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ and the resulting shape equation is defined by \eqt{eq:tshape}.  
+Next, the factorization $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ is introduced, where the shape is denoted by $\psi^g$, leading to the following shape equations, \eqt{eq:tshape}:
 \be
@@
 \ee
-\be
-\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},\sum_{g'}(H^{' \to g}g+P_p^{g' \to g}-L^{g'}\delta_{g'g})\psi^{g'}\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
+Finally, the PRKE parameters are defined by Equations~\ref{eq:trmb}~and~\ref{eq:tb}, where 
+$\left(\Psi^{*g},f^g\right) =$\\ $ \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d^2\Omega$.
+\be
+\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},\sum_{g'}(H^{g' \to g}g+P_p^{g' \to g}-L^{g'}\delta_{g'g})\psi^{g'}\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
 \label{eq:trmb}
@@
 \be
-\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{\sum_{g=1}^G(\Psi^{*g}, P_{d,i}^g \psi^g)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
+\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{\sum_{g=1}^G(\Psi^{*g}, \sum_{g'} P_{d,i}^{g' \to g} \psi^{g'})}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
 \label{eq:tb}
 \ee
-where $P_{d,i}^g$ is the delayed-neutron operator for precursor group $i$.
+where $P_{d,i}^{g' \to g}$ is the delayed-neutron operator for precursor group $i$.
 

### 2016-06-14 doc/NEAMS_Report/NEAMS_Report.tex
small changes

@@
 
-Solving for the shape can become expensive, especially in two or three dimensions.
-Using IQS, one expects that the shape's time dependency is weaker than that of the flux itself,  thus allowing larger time step sizes in computing the shape.
+Solving for the shape can become expensive, especially in two or three dimensions, and even more so when using the transport equations in lieu of the diffusion equations.
+Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself,  thus allowing for larger time step sizes in updating the shape.
 The PRKE equations form a small ODE system and can be solved using a much smaller time step size.
@@
 This iteration process must converge the shape such that the uniqueness condition $(\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=0)$ is preserved.
-The most convenient method to implement in Rattlesnake are Picard iterations.
-Where initially, the shape is evaluated with a guessed amplitude, if the normalization condition changes significantly, then the PRKE is evaluated with parameters computed from the last solution.
-This process is repeated until convergence.
-Picard iteration is predominantly utilized for multiphysics simulation, where neutronics is nonlinearly coupled to thermal-hydraulics.
-So the shape-PRKE couple is easily fit into this process, described by \fig{fig:picard}.
+
+The most convenient method to implement IQS in RATTLESNAKE are Picard or fixed-point iterations.
+Over a given Picard iteration, the shape is evaluated first with the latest available amplitude.
+The updated shape is employed to compute the PRKE parameters (using a linear interpolation in time) and then the PRKEs are solved using micro time steps.
+This process is repeated until convergence of the normalization condition.
+Picard iterations can also be utilized for multiphysics simulations, where neutronics (e.g., IQS) is nonlinearly coupled to thermal-hydraulics, for instance.
+Thus, inlcuding the IQS nonlinear system solve into the Picard iteration of MOOSE is a straightforward process and is described in \fig{fig:picard}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\tcr{change eqs}
 
@@
 \be
-e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_2 - \sum_{g=1}^G\varphi^g_1}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_2},\norm{\sum_{g=1}^G\varphi^g_1}\right)}
+e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2} - \sum_{g=1}^G\varphi^g_{\Delta t}}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_2},\norm{\sum_{g=1}^G\varphi^g_1}\right)}
 \label{eq:edt2}
@@
 \be
-\Delta t_{new} = \Delta t_{old} \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+\mu}}
+\Delta t_{new} = \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+\mu}}
 \label{eq:dt2}
 \ee
+\tcr{change subscripts}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Rattlesnake Implementation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Rattlesnake is a MOOSE-based application developed INL specific to solving radiation transport problems with multiphysics capabilities.
-MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like Rattlesnake.
+\section{RATTLESNAKE Implementation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+RATTLESNAKE is a MOOSE-based application developed INL specific to solving radiation transport problems with multiphysics capabilities.
+MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like RATTLESNAKE.
 Because radiation transport is such a complex problem, Rattlesnake is a relatively large application with a lot of functionality.
@@
 \frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
-& - \underbrace{\frac{1}{v^g} \overbrace{\frac{1}{p}\frac{dp}{dt}}^{From Executioner}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
+& - \underbrace{\frac{1}{v^g} \boxed{\overbrace{\frac{1}{p}\frac{dp}{dt}}^{From Executioner}}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
 \end{align}
 
-To enable Rattlesnake to solve this equation, another kernel was created that evaluates $\sum_{g=1}^G\frac{1}{v^gp}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
+To enable Rattlesnake to solve this equation, another kernel was created that evaluates $\sum_{g=1}^G\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
 Also, the precursor kernel was modified to include the $\frac{1}{p}$ term.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\tcr{make sure boxes are nice int he final version}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\tcr{homogeneous}
 
@@
 \includegraphics[width=5in]{\FiguresDir/1D_het_convergence.jpg}
-\caption{Error convergence comparison of 1D hetergenous example}
+\caption{Error convergence comparison of 1D heterogeneous example}
 \label{fig:conv}
@@
 
+\tcr{not an easy problem because shape changes in time a lot. 1 figure, shape at 3 times: t1=0, cosine, t2=max deform, t3=end simulation}
+
 \newpage
@@
 
-\fig{fig:TWIGL_power} show the IQS  solution as compared with the Brute Force solution.
+\fig{fig:TWIGL_power} shows the IQS  solution as compared with the Brute Force solution.
 These plots show that IQS is consistent in more complex, higher dimensional problems in Rattlesnake.

### 2016-06-14 doc/NEAMS_Report/NEAMS_Report.tex
NEAMS: Addressing comments

@@
 % common reference commands
-\newcommand{\eqt}[1]{Equation~(\ref{#1})}                     % equation
+\newcommand{\eqt}[1]{Equation~\ref{#1}}                     % equation
 \newcommand{\fig}[1]{Figure~\ref{#1}}                      % figure
@@
 The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates.
-The IQS method has primarily been developed in the context of the neutron diffusion approximation \cite{Ban_2012,Ferguson_1973}, but \cite{Goluoglu_2001} provides an extension of the technique to transport, implemented in TDTORT.
+The IQS method has primarily been developed in the context of the neutron diffusion approximation \cite{Ban_2012,Ferguson_1973}, but reference \cite{Goluoglu_2001} provides an extension of the technique to transport, implemented in TDTORT.
 The method has mostly being applied to neutron kinetics, without thermal-hydraulic feedback.
@@
 \begin{align}
-\frac{1}{v^g} \frac{\partial \phi^g }{\partial t} =& \frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \phi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g  \nonumber \\
+\frac{\partial}{\partial t}\left(\frac{\phi^g}{v^g}\right)  =& \frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \phi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g  \nonumber \\
 &  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \phi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
@@
 \begin{align}
-\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\ 
+\frac{\partial}{\partial t}\left(\frac{\varphi^g}{v^g}\right) = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\
 & -  \left( -\div D^g \grad  + \Sigma_r^g + \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right) \varphi^g + \boxed{\frac{1}{p}}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
@@
 \be
-\frac{1}{v^g}\frac{\partial \Psi^g}{\partial t} = \sum_{g'} \left(H^{g'\to g} + P_p^{g' \to g} \right) \Psi^{g'} - L^g\Psi^g + S_d^g
+\frac{\partial}{\partial t}\left(\frac{\Psi^g}{v^g}\right) = \sum_{g'} \left(H^{g'\to g} + P_p^{g' \to g} \right) \Psi^{g'} - L^g\Psi^g + S_d^g
 \label{eq:transport}
@@
 \be
-\frac{1}{v^g}\frac{\partial \psi^g}{\partial t} = \sum_{g'} \left(H^{g' \to g} + P_p^{g' \to g}\right) \psi^{g'} - \left( L^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \psi^g + \frac{1}{p} S_d^g
+\frac{\partial}{\partial t}\left(\frac{\psi^g}{v^g}\right) = \sum_{g'} \left(H^{g' \to g} + P_p^{g' \to g}\right) \psi^{g'} - \left( L^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \psi^g + \frac{1}{p} S_d^g
 \label{eq:tshape}
@@
 
-\tcr{change eqs}
-
-This section presents two different time-integration methods to solve coupled IQS shape and precursor equations, recalled below using, for simplicity, a single neutron group and a single precursor group.
-
-\be
-\frac{1}{v}\frac{\partial\varphi}{\partial t}=\nu\Sigma_f(1-\beta)\varphi-\left(-\div D \grad + \Sigma_a + \frac{1}{v}\frac{1}{p}\frac{dp}{dt}\right)\varphi+\frac{1}{p}\lambda C 
-\ee
-\be
-\frac{dC}{dt} = \beta\nu \Sigma_f \varphi p - \lambda C
-\ee
-
+This section presents the time-integration method used to solve coupled IQS shape and precursor equations, represented by Equations~\ref{eq:shape2}~and~\ref{eq:preq2}.
 First, we note that we could keep this system of two time-dependent equations and solve it as a coupled system.
@@
 If the step is too large, the step will be repeated with a smaller step calculated with the resulting error.
-The error of the time step is approximated by \eqt{eq:edt2}, where $\varphi^g_1$ and $\varphi^g_2$ are the solutions with the full step and half step, respectively.
+The error of the time step is approximated by \eqt{eq:edt2}, where $\varphi^g_{\Delta t}$ and $\varphi^g_{\Delta t/2}$ are the solutions with the full step and half step, respectively.
 $\varphi$ is changed to $\phi$ for regular flux evaluation and IQS P-C.
@@
 \be
-e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2} - \sum_{g=1}^G\varphi^g_{\Delta t}}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_2},\norm{\sum_{g=1}^G\varphi^g_1}\right)}
+e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2} - \sum_{g=1}^G\varphi^g_{\Delta t}}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2}},\norm{\sum_{g=1}^G\varphi^g_{\Delta t}}\right)}
 \label{eq:edt2}
@@
 \ee
-\tcr{change subscripts}
 
@@
 \begin{align}
-\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
+\frac{\partial}{\partial t}\left(\frac{\varphi^g}{v^g}\right)=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
 & - \underbrace{\frac{1}{v^g} \boxed{\overbrace{\frac{1}{p}\frac{dp}{dt}}^{From Executioner}}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\tcr{homogeneous}
-
-The example is very simple and computes quickly, it entails a one dimensional, heterogeneous 400 cm slab with a varying absorption cross section.
+
+The example is very simple and computes quickly, it entails a one dimensional, homogeneous 400 cm slab with a heterogenous perturbation in absorption cross section.
 \fig{fig:slab} shows how the regions of the slab are divided and \tbl{tab:1Dmat} shows the initial material properties.
@@
 \end{tabular}
-\caption{1-D heterogeneous slab region identification}
+\caption{1-D slab region identification}
 \label{fig:slab}
@@
 \begin{center}
-\caption{1-D heterogeneous slab material properties and problem parameters}
+\caption{1-D slab material properties and problem parameters}
 \label{tab:1Dmat}
@@
 \begin{center}
-\caption{1-D heterogeneous slab absorption cross-section at times of interest}
+\caption{1-D slab absorption cross-section at times of interest}
 \label{tab:1Dslope}
@@
 These results show that traditional IQS does not perform well with step doubling time adaptation and IQS P-C performs moderately better than brute force.
-
-%Figures \ref{fig:initial} - \ref{fig:final} plots shape changes in the IQS method, showing where the shape solution is necessary and a simple PRKE evaluation is inadequate.
+The IQS performance in this example are not exceptional because it is not an easy problem.
+As seen in \fig{eq:D_shape}, the shape undergoes significant change through the transient, which requires many shape evaluations to retain accuracy.
 
@@
 \includegraphics[width=0.5\textwidth]{\FiguresDir/1D_het_power_plot2.png}
-\caption{Power level of 1D heterogeneous example}
+\caption{Power level of 1D example}
 \label{fig:power}
 \end{figure}
-
-%\begin{figure}[!htbp]
-%\begin{center}
-%\includegraphics[width=5in,height=2in]{\FiguresDir/initial_flux.jpg}
-%\caption{Initial Flux Plot}
-%\label{fig:initial}
-%\end{center}
-%\end{figure}
-%
-%\begin{figure}[!htbp]
-%\begin{center}
-%\includegraphics[width=5in,height=2in]{\FiguresDir/max_flux.jpg}
-%\caption{Flux Plot when Absorption Cross Section is at Minimum}
-%\label{fig:max}
-%\end{center}
-%\end{figure}
-%
-%\begin{figure}[!htbp]
-%\begin{center}
-%\includegraphics[width=5in,height=2in]{\FiguresDir/final_flux.jpg}
-%\caption{Final Flux Computation (not steady-state)}
-%\label{fig:final}
-%\end{center}
-%\end{figure}
 
@@
 \includegraphics[width=5in]{\FiguresDir/1D_het_convergence.jpg}
-\caption{Error convergence comparison of 1D heterogeneous example}
+\caption{Error convergence comparison of 1D example}
 \label{fig:conv}
@@
 \begin{center}
-\caption{1-D heterogeneous slab step doubling results}
+\caption{1-D slab step doubling results}
 \label{tab:1Ddt2}
@@
 
-\tcr{not an easy problem because shape changes in time a lot. 1 figure, shape at 3 times: t1=0, cosine, t2=max deform, t3=end simulation}
+\begin{figure}[!htbp]
+\begin{center}
+\includegraphics[width=\linewidth]{\FiguresDir/1D_shape.png}
+\caption{Flux profile of 1D slab at various times}
+\label{fig:1D_shape}
+\end{center}
+\end{figure}
 

### 2016-06-15 doc/NEAMS_Report/NEAMS_Report.tex
final review.

@@
 
-The most convenient method to implement IQS in RATTLESNAKE are Picard or fixed-point iterations.
+The most convenient method to implement IQS in RATTLESNAKE is through Picard or fixed-point iterations.
 Over a given Picard iteration, the shape is evaluated first with the latest available amplitude.
@@
 
-The Predictor-Corrector version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the evaluation of the coupled system is different.
-This version first solves the flux diffusion (represented by Equations~\ref{eq:flux}~and~\ref{eq:precursor}) to get a predicted flux.
-The predicted flux at this step is then converted to shape by rescaling as follows:
+The Predictor-Corrector (P-C) version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the solution of the coupled system of equations is different.
+In the IQS P-C version, the flux equations (not the shape equations) are solved (represented by Equations~\ref{eq:flux}~and~\ref{eq:precursor}) in order to obtain a predicted flux solution.
+This predicted flux is then converted to shape by rescaling as follows:
 \be
@@
 \ee
-Where:
+where the scaling factors are given by
 \be
@@
 \be
-K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
-\ee
-
-The PRKE parameters are then computed with this shape using Equations~\ref{eq:rmb}~-~\ref{eq:l} and interpolated over the macro step, then the PRKE is evaluated.
-With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
-\be
-\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1}
-\ee
-
-The advantage to the predictor-corrector method is there is no iteration necessary for this method and in turn is much simpler and faster than the standard IQS.
-The disadvantage is that the method assumes $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)$ is inherently constant.
+K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0} \,.\right)
+\ee
+
+The PRKE parameters are then computed with this shape using Equations~\ref{eq:rmb}~-~\ref{eq:l} and interpolated over the macro step, then the PRKE ODE system is solved on the micro time scale.
+With the newly computed amplitude, the shape is rescaled into a flux and the final corrected flux is given by:
+\be
+\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1} \,.
+\ee
+
+The advantage to the predictor-corrector method is there is no iteration necessary for this method and, in turn, is much simpler and faster than the standard IQS.  
+%The disadvantage is that the method assumes, by construction, that $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)$ is inherently constant.
 
@@
 
-A very important aspect of IQS is how it performs with time adaptation because this will most likely be used with IQS.
-The time adaptation used for quantifying IQS's ability is step doubling.
-The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step and a solution with two half steps.
-If the step is small enough, the error will be smaller than a user driven tolerance and the magnitude of the next step will be calculated based on the error.
-If the step is too large, the step will be repeated with a smaller step calculated with the resulting error.
-The error of the time step is approximated by \eqt{eq:edt2}, where $\varphi^g_{\Delta t}$ and $\varphi^g_{\Delta t/2}$ are the solutions with the full step and half step, respectively.
-$\varphi$ is changed to $\phi$ for regular flux evaluation and IQS P-C.
-If $e_n > e_{max}$ the time step is repeated; if $e_n < e_{max}$ the system moves to the next time step.
-The next $\Delta t$ is calculated using \eqt{eq:dt2}, where $\mu$ is the convergence order of the time integration scheme being used.
-$e_{max}$ and $e_{tol}$ are user defined parameters; $e_{max}$ is usually less than $e_{tol}$ to better guarantee that the calculated $\Delta t_{new}$ will pass the error criteria so time steps won't be repeated.
-
+IQS aims at reducing the time discretization error in the flux solution by splitting the flux into an amplitude (that is highly resolved with at a micro time scale) and a shape (whose time-dependence is weaker than that of the flux itself).
+Thus, by construction, the IQS approach may employ larger time-step sizes for comparable temporal error.
+Further enhancements can be gained by using time adaptation (or time step control) in order to increase or reduce the macro time step size for the
+shape evaluation, depending on error estimates. Here, we employ a step-doubling technique to carry out the time adaptation.
+The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps ($\varphi^g_{\Delta t/2}$).
+Note:
+$\varphi$ is changed to $\phi$ for regular flux evaluation (brute-force time discretization of the governing equations) and IQS P-C.
+
+The relative error is computed as follows:
 \be
@@
 \ee
-\be
-\Delta t_{new} = \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+\mu}}
+If the error is smaller than the user-specified tolerance, $e_{tol}$, the time step is accepted. In addition, a new time step size is estimated as follows:
+\be
+\Delta t_{new} = S \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}} \,,
 \label{eq:dt2}
 \ee
-
-To be clear, each solution undergoes Picard iterations until convergence before the error is calculated; for IQS P-C, each solution is re-scaled by its corresponding $p$.
-Additionally, if the PJFNK iteration does not converge, the entire step doubling process is repeated with half the time original time step.
+where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor.\\
+If the error is larger than the user-specified tolerance, the time step is rejected. A new time step size is estimated using  \eqt{eq:edt2} as well.
+
+To sum up, each solution undergoes Picard iterations until convergence before the error is calculated; for IQS P-C, each solution is re-scaled by its corresponding amplitude $p$.
+Additionally, if the nonlinear iteration does not converge, the entire step doubling process is repeated with half the time original time step.
 
@@
 MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like RATTLESNAKE.
-Because radiation transport is such a complex problem, Rattlesnake is a relatively large application with a lot of functionality.
-At the heart of Rattlesnake is the action system.
-The Rattlesnake action system gives the capability of consolidating MOOSE input syntax, so that a user does not have to define every kernel, variable, etc.
-in the problem, which could be on the order of thousands with multigroup problems.
-The user, instead, is able to input an equation description (Diffusion, S$_n$, P$_n$, etc.) and solve method (SAAF,LS,CFEM,DFEM,etc.), and the action system will incorporate all the necessary physics involved (kernels, boundary conditions, postprocessor, etc.).
+At the heart of RATTLESNAKE is the action system, which provides a means to consolidate the MOOSE input syntax (which can be quite larger for multigroup transport simulations), so that a user does not have to define every kernel, variable, etc., used in the problem.
+Rather, the user inputs an equation description
+(e.g., Diffusion, S$_n$, P$_n$, etc.) and a solution method (e.g., SAAF, LS, CFEM, DFEM, etc.), and the action system will incorporate all the necessary physics involved (kernels, boundary conditions, postprocessor, etc.).
 
@@
 
+An IQS excecutioner was created to implement the IQS method.
 The IQS executioner derives from the Transient executioner in MOOSE.
-The IQS executioner contains a loop over micro time steps that computes the PRKE and then passes $p$ and $\frac{dp}{dt}$ for the Transient executioner to evaluate the shape equation at each macro step.
-The PRKE is computed with a user specified option of backward-Euler, Crank-Nicholson, or SDIRK33.
+The IQS executioner contains a loop over micro time steps that computes the PRKE solution and then passes $p$ and $\frac{dp}{dt}$ for the Transient executioner to evaluate the shape equation at each macro step.
+The PRKE solve is performed with a user specified option of backward-Euler, Crank-Nicholson, or SDIRK33.
 The IQS executioner also supplements Transient Picard iteration process by adding its own error criteria:
@@
 
-IQS defines its uniqueness from its executioner type; however, many changes needed to be made in the Rattlesnake action system in order to support IQS execution.
+The IQS implementation mostly require the specific IQS executioner, described above.
+However, IQS additional changes are in the RATTLESNAKE action system in order to support the IQS execution.
 First, changes needed to be made in order to evaluate the shape equation.
-The shape equation, after some manipulation, is very similar to the time-dependent, which Rattlesnake is already set up to solve:
+The shape equation, after some manipulation, is very similar to the time-dependent governing laws that RATTLESNAKE already solves.
+Using multigroup diffusion as an example again, we show the various kernels employed and highlight the new or modified kernels.
 
@@
 
-To enable Rattlesnake to solve this equation, another kernel was created that evaluates $\sum_{g=1}^G\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
+To enable RATTLESNAKE to solve this equation, an IQS removal kernel was created to evaluate $\sum_{g=1}^G\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
 Also, the precursor kernel was modified to include the $\frac{1}{p}$ term.
@@
 After each of these postprocessors are evaluated, a user object pulls together all the values and performs the numerator/denominator divisions.
-The resulting values are then passed to the executioner for the PRKE evaluation.
+The resulting values are then passed to the executioner for the PRKE solve.
 
@@
 For simplicity, IQS implementation has only been described for CFEM diffusion.
-However, Rattlesnake has other action systems capable of transient simulation, where IQS can be implemented and be effective.
+However, RATTLESNAKE has other action systems capable of transient simulation, where IQS can be implemented and be effective.
 One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equations~\ref{eq:flux}~and~\ref{eq:shape2}.
@@
 
-Additional action systems involve transport, but it is evident from \sct{sect:transport} that IQS implementation in these is straightforward.
-The main differences between diffusion implementation in transport is outlined below:
+Additional action systems involve transport, but it is evident from \sct{sect:transport} that IQS implementation in these is straightforward as well.
+The main differences between a diffusion implementation and a transport implementation are outlined below:
 \begin{enumerate}
-\item The form of the operators in the shape equations is different, but Rattlesnake has already implemented all the kernels necessary to represent these operators.
+\item The form of the operators in the shape equations is different, but RATTLESNAKE has already implemented all the kernels necessary to represent these operators.
 So no change is necessary to these kernels is necessary for IQS.
-Additionally, the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ is the same, so both actions can use the same kernel.
+Additionally, the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ is the same, so both action systems can use the same kernel.
 \item The PRKE parameters also change because of the operators.
@@
 
-IQS plans to be available to any action system capable of transient simulation, which include:
+IQS will be made available to any action system capable of transient simulation, which includes:
 \begin{itemize}
@@
 So the conversion of the predicted flux in  \eqt{eq:rescale} to shape is unnecessary if the corrected flux is solved with \eqt{eq:phicorr2}.
-After obtaining the corrected flux, the precursors are re-evaluated using a \texttt{EXEC\_LINEAR}.
+After obtaining the corrected flux, the precursors are re-evaluated using a \texttt{EXEC\_LINEAR} statement.
 \be
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\tcr{make sure boxes are nice int he final version}
-
-The input deck for IQS is very similar to the current transient diffusion input file.
+
+The input file for IQS is very similar to the current transient diffusion input file.
 The IQS input has a different executioner type and parameters.
-The executioner type is simply IQS and input parameters include number of micro time steps, IQS error tolerance, and initial power.
-The Rattlesnake transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $\keff$, which is present in the transient input deck.
+The executioner type is simply IQS and input parameters include the number of micro time steps per macro step, the IQS error tolerance, and the initial power.
+The RATTLESNAKE transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $\keff$, which is present in the transient input deck.
 However, IQS also requires an initial evaluation of the adjoint flux, for the weighting function.
@@
 The \texttt{IQS\_error\_tol} is the tolerance for the IQS error represented by \eqt{eq:eiqs} and will at most \texttt{picard\_max\_its} iterations until convergence.
-The \texttt{prke\_scheme} defines the time discretization used for the PRKE evaluation.
+The \texttt{prke\_scheme} defines the time discretization used for the PRKE solution.
 Where \texttt{RK} is SDIRK33, \texttt{CN} is Crank-Nicholson, and \texttt{IE} is implicit-Euler.
@@
 The performance for each method are represented by convergence plots with constant $\Delta t$ and by the number of time steps with step doubling time adaptation.
-For time adaptation, backward-Euler time discretization was used and, for a safety factor, $e_{tol} = 0.8e_{max}$.
+For time adaptation, backward-Euler time discretization was used.
+% and, for a safety factor, $e_{tol} = 0.8e_{max}$. I found tol and max to be confusing to the reader. i have simplified a bit
 
@@
 \hline
-Test & $e_{max}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
+Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
 \hline
@@
 The results show that both IQS methods perform exceptionally well compared to brute force.
-It also shows that traditional IQS performed better with large $e_{max}$, while IQS P-C was better with smaller $e_{max}$.
+It also shows that traditional IQS performed better with large $e_{tol}$, while IQS P-C was better with smaller $e_{tol}$.
 
@@
 To have better understanding on the cross sections given later, we present the equations here:
+\begin{subequations}
 \begin{align}
--\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\grad D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 - \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
--\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\grad D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
+-\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 - \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
+-\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
 S_f &= \sum_{g=1}^2 \Sigma_{f,g} \phi_g, \\
@@
 \end{align}
-where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_6f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0=T(\vec{r}, t=0)$.
+\end{subequations}
+where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0=T(\vec{r}, t=0)$.
 The two-group diffusion equation are solved with zero flux boundary conditions on external surfaces, reflecting conditions at symmetry boundaries and steady state initial conditions which are obtained by solving
 \begin{align}
--\grad D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
--\grad D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1. \\
+-\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
+-\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1. \\
 \end{align}
@@
 The \fig{fig:LRA_plots} plots show the IQS  solution as compared with the Brute Force solution.
-These plots show that IQS is consistent with Doppler feedback problems in Rattlesnake.
+These plots show that IQS is consistent with Doppler feedback problems in RATTLESNAKE.
 \fig{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods.
 \tbl{tab:LRAdt2} shows the time adaptation performance of brute force and IQS P-C.
-These results show that IQS, again, performs much better than the brute force method by taking significantly less diffusion solves while retaining similar accuracy.
+These results show that IQS, again, performs much better than the brute force method by taking significantly fewer diffusion solves while retaining similar accuracy.
 
@@
 \hline
-Test & $e_{max}$ & Error & Steps & Solves & Error & Steps & Solves \\
+Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves \\
 \hline
@@
 
-The purpose of this paper was to present the improved quasi-static method with its derivation, its implementation into Rattlesnake, and its effectiveness with several test cases.
-The derivation of IQS was done with multigroup diffusion, but it was seen in \sct{sect:transport} that application to transport is relatively simple.
-Two different solving methods were introduced.
-The first was the traditional IQS method where the shape equation and PRKE were solved as a coupled nonlinear system with Picard iterations.
-The second was IQS P-C where flux is solved and then rescaled using results from evaluating the PRKE.
-IQS P-C is a much simpler method and, computationally, much less expensive to perform.
-\\
-
-IQS's implementation into Rattlesnake involved creating a new kernel, auxiliary kernel, executioner, and four postprocessors, as well as alterations to the action system.
+The purpose of this report is to briefly review the improved quasi-static method for neutronic transient simulations, to describe its implementation in RATTLESNAKE, the radiation transport application based on MOOSE, and to show its effectiveness using several test cases.
+%The derivation of IQS was done with multigroup diffusion, but it was seen in \sct{sect:transport} that application to transport is relatively simple.  
+Two IQS implementations are available in RATTLESNAKE.
+The first one is the traditional IQS approach, where the shape equation and PRKE are solved as a coupled nonlinear system with Picard iterations.
+The second approach, IQS P-C, is a predictor-corrector linearization, where the flux is solved and then rescaled using the amplitude obtained through the PRKE
+(Point Reactor Kinetic Equations).  IQS P-C is a simpler method and is computationally much less expensive to perform. \\
+
+IQS's implementation in RATTLESNAKE involves a new kernel, an auxiliary kernel, a new executioner derived from the current transient, four postprocessors, as well as alterations to the action system.
 The new kernel implements the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term in the shape equation.
 The auxiliary kernel alters how the precursors are solved, the method is outlined by \sct{sect:prec}.
-The executioner is where the PRKE is evaluated after shape solves are performed.
-The postprocessors are used to evaluate the integrals for the PRKE parameters and a user object is used to pass them to the executioner.
+The executioner is where the PRKE are evaluated after shape solves are performed.
+The postprocessors are used to evaluate the parameters for the PRKE and a user object is used to pass them to the executioner.
 IQS's implementation the action system made sure that all the correct kernels and postprocessors are executed when the IQS executioner is specified in the input.
@@
 
-IQS was tested by comparing its performance with traditional "brute force" method using error convergence plots and time adaptation.
+The IQS implementations have been verified and comparing against the traditional ``brute force'' temporal discretization of the governing laws.
+Error convergence plots and time adaptation performance are provided.
 Three examples were used including a custom one-dimensional example, the TWIGL benchmark, and LRA benchmark.
@@
 The TWIGL benchmark shows the effectiveness of the PRKE in IQS because there is very little shape change through the transient.
-IQS shows very impressive performance with both constant time steps and time adaptation.
-The LRA benchmark tests IQS when applied with multiphysics.
-Again, IQS shows impressive performance by requiring half the time steps as brute force for a given error.
+IQS shows very satisfactory performance with both constant time steps and time adaptation.
+The LRA benchmark tests IQS when utilized in the context of a multiphysics simulation.
+Again, IQS shows good performance by requiring half the time steps as brute force for a given error.
 

### 2016-06-15 doc/NEAMS_Report/NEAMS_Report.tex
NEAMS: Making Rattlesnake consistent and adding header and footer

@@
 \newcommand{\norm}[1]{\left\lVert#1\right\rVert_{L^2}}
+\newcommand{\rattlesnake}{Rattlesnake }
 % extra space
@@
 
-\title{Improved Quasi-Static Methods in Rattlesnake}
-\subtitle{Development and Implementation of Improved Quasi-Static (IQS) methods for time-dependent neutron diffusion and neutron transport solvers in Rattlesnake}
+\title{Improved Quasi-Static Methods in \rattlesnake}
+\subtitle{Development and Implementation of Improved Quasi-Static (IQS) methods for time-dependent neutron diffusion and neutron transport solvers in \rattlesnake}
 \author{ \normalsize
@@
 \lhead{INL/EXT-16-38059}
-\rfoot{DRAFT}
+\rfoot{FINAL}
 \cfoot{\thepage}
@@
 
-\pagestyle{plain} % on headers or footers on the first page
+%\pagestyle{plain} % on headers or footers on the first page
 
@@
 
-This report presents  the implementation of the Improved Quasi-Static method (IQS) in the Yak/Rattlesnake applications of the MOOSE framework.
+This report presents  the implementation of the Improved Quasi-Static method (IQS) in the \rattlesnake application of the MOOSE framework.
 In the IQS method, the neutron flux is factored into a time-dependent amplitude and a spatial- and weakly time-dependent shape.
@@
 
-The most convenient method to implement IQS in RATTLESNAKE is through Picard or fixed-point iterations.
+The most convenient method to implement IQS in \rattlesnake is through Picard or fixed-point iterations.
 Over a given Picard iteration, the shape is evaluated first with the latest available amplitude.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{RATTLESNAKE Implementation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-RATTLESNAKE is a MOOSE-based application developed INL specific to solving radiation transport problems with multiphysics capabilities.
-MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like RATTLESNAKE.
-At the heart of RATTLESNAKE is the action system, which provides a means to consolidate the MOOSE input syntax (which can be quite larger for multigroup transport simulations), so that a user does not have to define every kernel, variable, etc., used in the problem.
+\section{\rattlesnake Implementation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\rattlesnake is a MOOSE-based application developed INL specific to solving radiation transport problems with multiphysics capabilities.
+MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like \rattlesnake.
+At the heart of \rattlesnake is the action system, which provides a means to consolidate the MOOSE input syntax (which can be quite larger for multigroup transport simulations), so that a user does not have to define every kernel, variable, etc., used in the problem.
 Rather, the user inputs an equation description
@@
 The IQS implementation mostly require the specific IQS executioner, described above.
-However, IQS additional changes are in the RATTLESNAKE action system in order to support the IQS execution.
+However, IQS additional changes are in the \rattlesnake action system in order to support the IQS execution.
 First, changes needed to be made in order to evaluate the shape equation.
-The shape equation, after some manipulation, is very similar to the time-dependent governing laws that RATTLESNAKE already solves.
+The shape equation, after some manipulation, is very similar to the time-dependent governing laws that \rattlesnake already solves.
 Using multigroup diffusion as an example again, we show the various kernels employed and highlight the new or modified kernels.
@@
 
-To enable RATTLESNAKE to solve this equation, an IQS removal kernel was created to evaluate $\sum_{g=1}^G\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
+To enable \rattlesnake to solve this equation, an IQS removal kernel was created to evaluate $\sum_{g=1}^G\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
 Also, the precursor kernel was modified to include the $\frac{1}{p}$ term.
@@
 For simplicity, IQS implementation has only been described for CFEM diffusion.
-However, RATTLESNAKE has other action systems capable of transient simulation, where IQS can be implemented and be effective.
+However, \rattlesnake has other action systems capable of transient simulation, where IQS can be implemented and be effective.
 One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equations~\ref{eq:flux}~and~\ref{eq:shape2}.
@@
 \begin{enumerate}
-\item The form of the operators in the shape equations is different, but RATTLESNAKE has already implemented all the kernels necessary to represent these operators.
+\item The form of the operators in the shape equations is different, but \rattlesnake has already implemented all the kernels necessary to represent these operators.
 So no change is necessary to these kernels is necessary for IQS.
@@
 The executioner type is simply IQS and input parameters include the number of micro time steps per macro step, the IQS error tolerance, and the initial power.
-The RATTLESNAKE transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $\keff$, which is present in the transient input deck.
+The \rattlesnake transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $\keff$, which is present in the transient input deck.
 However, IQS also requires an initial evaluation of the adjoint flux, for the weighting function.
@@
 \fig{fig:TWIGL_power} shows the IQS  solution as compared with the Brute Force solution.
-These plots show that IQS is consistent in more complex, higher dimensional problems in Rattlesnake.
+These plots show that IQS is consistent in more complex, higher dimensional problems in \rattlesnake.
 Finally, \fig{fig:TWIGL_conv} plots the error convergence of IQS and the Brute Force methods.
@@
 The \fig{fig:LRA_plots} plots show the IQS  solution as compared with the Brute Force solution.
-These plots show that IQS is consistent with Doppler feedback problems in RATTLESNAKE.
+These plots show that IQS is consistent with Doppler feedback problems in \rattlesnake.
 \fig{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods.
@@
 
-The purpose of this report is to briefly review the improved quasi-static method for neutronic transient simulations, to describe its implementation in RATTLESNAKE, the radiation transport application based on MOOSE, and to show its effectiveness using several test cases.
+The purpose of this report is to briefly review the improved quasi-static method for neutronic transient simulations, to describe its implementation in \rattlesnake, the radiation transport application based on MOOSE, and to show its effectiveness using several test cases.
 %The derivation of IQS was done with multigroup diffusion, but it was seen in \sct{sect:transport} that application to transport is relatively simple.  
-Two IQS implementations are available in RATTLESNAKE.
+Two IQS implementations are available in \rattlesnake.
 The first one is the traditional IQS approach, where the shape equation and PRKE are solved as a coupled nonlinear system with Picard iterations.
@@
 
-IQS's implementation in RATTLESNAKE involves a new kernel, an auxiliary kernel, a new executioner derived from the current transient, four postprocessors, as well as alterations to the action system.
+IQS's implementation in \rattlesnake involves a new kernel, an auxiliary kernel, a new executioner derived from the current transient, four postprocessors, as well as alterations to the action system.
 The new kernel implements the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term in the shape equation.
@@
 Equation (\ref{eq:preq2}) is very similar to the normal precursor equation (\ref{eq:precursor}), except the fission source term is multiplied by $p$.
-These differences are crucial for IQS implementation in Rattlesnake.
+These differences are crucial for IQS implementation in \rattlesnake.
 %

### 2016-06-15 doc/NEAMS_Report/NEAMS_Report.tex
NEAMS: Added Mark to list of authors

@@
 \author{ \normalsize
-  \textbf{Zachary M. Prince$^\dagger$, Jean C. Ragusa$^\dagger$, Yaqi Wang$^\star$} \\
+  \textbf{Zachary M. Prince$^\dagger$, Jean C. Ragusa$^\dagger$, Yaqi Wang$^\star$, Mark D. DeHart$^\star$} \\
  \normalsize $^\dagger$Department of Nuclear Engineering \\
@@
  \normalsize $^\star$Idaho National Laboratory\\
- \normalsize \href{mailto:zachmprince@tamu.edu}{zachmprince@tamu.edu}, \href{jean.ragusa@tamu.edu}{jean.ragusa@tamu.edu}, \href{yaqi.wang@inl.gov}{yaqi.wang@inl.gov} 
+ \normalsize \href{mailto:zachmprince@tamu.edu}{zachmprince@tamu.edu}, \href{jean.ragusa@tamu.edu}{jean.ragusa@tamu.edu}, \href{yaqi.wang@inl.gov}{yaqi.wang@inl.gov}, \href{mark.dehart@inl.gov}{mark.dehart@inl.gov}
 }

### 2016-07-26 doc/note-to-yaqi/Multiphysics_updates.tex
Temperature update note

\documentclass[11pt]{scrartcl} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{graphics}
\usepackage{float}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsbsy}
\usepackage{xspace}
\usepackage{verbatim}
\usepackage{csvsimple}
\usepackage{sectsty}
\usepackage{tikz}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% operators
\renewcommand{\div}{\vec{\nabla}\! \cdot \!}
\newcommand{\grad}{\vec{\nabla}}
% latex shortcuts
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bal}{\begin{align}}
\newcommand{\eali}{\end{align}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\ben}{\begin{enumerate}}
\newcommand{\een}{\end{enumerate}}
% DGFEM commands
\newcommand{\jmp}[1]{[\![#1]\!]}                     % jump
\newcommand{\mvl}[1]{\{\!\!\{#1\}\!\!\}}             % mean value
\newcommand{\kef}{\ensuremath{k_{\textit{eff}}}}
%\newcommand{\keff}{{\text{k}$_\textit{eff}$}\xspace}
\newcommand{\keff}{\kef\xspace}
% shortcut for domain notation
%\newcommand{\D}{\mathcal{D}}
% vector shortcuts
\newcommand{\vo}{\vec{\Omega}}
\newcommand{\vr}{\vec{r}}
\newcommand{\vn}{\vec{n}}
\newcommand{\vnk}{\vec{\mathbf{n}}}
\newcommand{\vj}{\vec{J}}
% extra space
\newcommand{\qq}{\quad\quad}
% common reference commands
\newcommand{\eqt}[1]{Eq.~(\ref{#1})}                     % equation
\newcommand{\fig}[1]{Fig.~\ref{#1}}                      % figure
\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table

\newcommand{\ud}{\,\mathrm{d}}
\newcommand{\mt}[1]{\marginpar{\tiny #1}}
\newcommand{\D}{\ensuremath{\mathcal{D}}}
\allowdisplaybreaks

\newcommand{\subj}{\\\hspace{-5mm}\textbf{Subject: IQS with multiple PRKE Parameter Updates within Macro Steps}}
\newcommand{\from}{\vspace{-20mm}\hspace{-1mm} \normalfont From: Zachary Prince \\}
\newcommand{\toms}{\hspace{-5mm} To: Jean Ragusa and Yaqi Wang \\}
\newcommand{\datet}{\hspace{-5mm} Date: \today \\}

\allsectionsfont{\centering \large}
\setlength\parindent{0pt}
\usepackage{ccaption}
\setlength{\abovecaptionskip}{0mm} %lowers the distance of captions to the figure
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{1in}
\setlength{\textwidth}{6.5in}

% tikz stuff
\usetikzlibrary{shapes,arrows,positioning}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\tikzstyle{greenblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=green!20, ultra thick, minimum width=60mm, minimum height=15mm,]
\tikzstyle{blueblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=blue!20, ultra thick, minimum width=60mm, minimum height=15mm]
\tikzstyle{reddiamond}=[diamond, draw, align=center, top color=white, bottom color=red!20, ultra thick, minimum width=60mm, aspect=2]
\newcommand{\tikzback}[5]{
 \begin{pgfonlayer}{background}
  \path (#1.west |- #2.north)+(-0.5,0.75) node (a1) {};
  \path (#3.east |- #4.south)+(+0.5,-0.25) node (a2) {};
  \path[fill=yellow!10,rounded corners, draw=black!100, dashed] (a1) rectangle (a2);
   \path (#3.east |- #2.north)+(0,1.25)--(#1.west |- #2.north) node[midway] (#5-n) {};
   \path (#3.east |- #2.south)+(0,-0.35)--(#1.west |- #2.south) node[midway] (#5-s) {};
   \path (#1.west |- #2.north)+(-0.75,0)--(#1.west |- #4.south) node[midway] (#5-w) {};
 \end{pgfonlayer}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%   BEGIN DOCUMENT
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%---------------------------------------------------------------------------%%
%% OPTIONS FOR NOTE
%%---------------------------------------------------------------------------%%
\title{\normalfont \Large TEXAS A\&M UNIVERSITY}
\subtitle{\normalfont Dwight Look College of Engineering \\ Department of Nuclear Engineering \\ \Large \textbf{Research Memo}}
\date{}

%%---------------------------------------------------------------------------%%
%% BEGIN NOTE
%%---------------------------------------------------------------------------%%

\maketitle
\from
\toms
\datet
\subj

%%---------------------------------------------------------------------------%%
\section{\large Initial Implementation of Multiple PRKE Parameter Updates}
%%---------------------------------------------------------------------------%%

The purpose of this note is to describe how IQS updates the PRKE parameters within macro time steps and the results of the initial implementation.  The impetus of this initiative was the underwhelming results from applying IQS to the LRA benchmark.  It is strongly believed that the treatment of the coupling between temperature and the PRKE parameters were to blame for results.  Figure~\ref{fig:lra_bad} shows that IQS only improves the time step size by less than half at $t=1.44$ s.  This is the point where the power is at a max, so the temperatue is changing rapidly, but the shape is not changing significantly.  Figure~\ref{fig:lra_good} shows the error at $t=1.40$ s, where IQS performs significantly better.  At this point, the temperature is not changing significantly.  Therefore, IQS does not treat temperature with enough accuracy to see significant improvement over brute force.

\begin{figure}[!htpb]
\centering
\subfigure[LRA convergence results at $t=1.44s$]{
\label{fig:lra_bad}
\includegraphics[width=0.52\textwidth]{mp_figs/lra_bad.jpg} }
\hspace{-15mm}
\subfigure[LRA convergence results at $t=1.40$]{
\label{fig:lra_good}
\includegraphics[width=0.52\textwidth]{mp_figs/lra_good.png} }
\caption{LRA convergence plots with only one temperature updated each macro step}
\end{figure}

Temperature has the most effect when calculating the PRKE parameters to solve for amplitude.  So in order to more accurately capture the behavior of the flux, the PRKE parameters and temperature need to be evaluated multiple times within each macro step. To visualize this process the system is setup with three different time scales, as shown in Figures~\ref{fig:time}~and~\ref{fig:timepc}.  Figures~\ref{fig:proc}~and\ref{fig:procpc} show the same processes in more detail.  The occurrence that there are three updates in the figures is arbitrary, the number of temperature updates can be set by the user.

\begin{figure}[H]
\centering
\begin{tikzpicture}[scale=2]
%Shape
\draw[] (0,2) -- (6,2) ;
\foreach \x in  {0,6}
\draw[shift={(\x,2)},color=black] (0pt,6pt) -- (0pt,-6pt);
\draw[shift={(0,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(shape) at (-.25,2) {$\varphi:$};

%Temp/Params
\draw[] (0,1) -- (6,1) ;
\foreach \x in  {0,6}
\draw[shift={(\x,1)},color=black] (0pt,6pt) -- (0pt,-6pt);
\foreach \x in  {2,4}
\draw[shift={(\x,1)},color=black] (0pt,4pt) -- (0pt,-4pt);
\draw[shift={(0,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(temp) at (-.75,1) {$T \rightarrow \rho, \bar{\beta}, \bar{\lambda}, \Lambda$:};

% PRKE
\draw[] (0,0) -- (6,0) ;
\foreach \x in  {0,6}
\draw[shift={(\x,0)},color=black] (0pt,6pt) -- (0pt,-6pt);
\foreach \x in  {0,2,4,6}
\draw[shift={(\x,0)},color=black] (0pt,4pt) -- (0pt,-4pt);
\foreach \x in  {0,0.2,0.4,0.6,0.8,1,1.2,1.4,1.6,1.8,2,2.2,2.4,2.6,2.8,3,3.2,3.4,3.6,3.8,4,4.2,4.4,4.6,4.8,5,5.2,5.4,5.6,5.8,6}
\draw[shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt);
\draw[shift={(0,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(prke) at (-.5,0) {$p, \frac{dp}{dt}:$};

\draw (0,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 2} (2,0);
\draw (2,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 4} (4,0);
\draw (4,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 6} (6,0);
\draw (0,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 1} (2,1);
\draw (2,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 3} (4,1);
\draw (4,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 5} (6,1);
\draw (0,2) edge[out=15,in=165,->,thick] node[above,sloped] {Step 7} (6,2);

\draw (0,2) edge[out=-135,in=135,->,thick] node[below,sloped] {$\varphi$} (0,1);
\draw (2,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (2,0);
\draw (2,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (2,1);
\draw (4,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (4,0);
\draw (4,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (4,1);
\draw (6,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (6,0);
\draw (6,0) edge[out=45,in=0,->,thick] node[below,sloped] {$p, \frac{dp}{dt}$} (6,2);

\end{tikzpicture}
\caption{Time scales and process of regular IQS}
\label{fig:time}
\end{figure}

\begin{figure}[H]
\centering
\begin{tikzpicture}[scale=2]
%Shape
\draw[] (0,2) -- (6,2) ;
\foreach \x in  {0,6}
\draw[shift={(\x,2)},color=black] (0pt,6pt) -- (0pt,-6pt);
\draw[shift={(0,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(shape) at (-.25,2) {$\varphi:$};

%Temp/Params
\draw[] (0,1) -- (6,1) ;
\foreach \x in  {0,6}
\draw[shift={(\x,1)},color=black] (0pt,6pt) -- (0pt,-6pt);
\foreach \x in  {2,4}
\draw[shift={(\x,1)},color=black] (0pt,4pt) -- (0pt,-4pt);
\draw[shift={(0,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(temp) at (-.75,1) {$T \rightarrow \rho, \bar{\beta}, \bar{\lambda}, \Lambda$:};

% PRKE
\draw[] (0,0) -- (6,0) ;
\foreach \x in  {0,6}
\draw[shift={(\x,0)},color=black] (0pt,6pt) -- (0pt,-6pt);
\foreach \x in  {0,2,4,6}
\draw[shift={(\x,0)},color=black] (0pt,4pt) -- (0pt,-4pt);
\foreach \x in  {0,0.2,0.4,0.6,0.8,1,1.2,1.4,1.6,1.8,2,2.2,2.4,2.6,2.8,3,3.2,3.4,3.6,3.8,4,4.2,4.4,4.6,4.8,5,5.2,5.4,5.6,5.8,6}
\draw[shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt);
\draw[shift={(0,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(prke) at (-.25,0) {$p:$};

\draw (0,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 3} (2,0);
\draw (2,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 5} (4,0);
\draw (4,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 7} (6,0);
\draw (0,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 2} (2,1);
\draw (2,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 4} (4,1);
\draw (4,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 6} (6,1);
\draw (0,2) edge[out=15,in=165,->,thick] node[above,sloped] {Step 1} (6,2);

\draw (0,2) edge[out=-135,in=135,->,thick] node[below,sloped] {$\varphi$} (0,1);
\draw (2,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (2,0);
\draw (2,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (2,1);
\draw (4,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (4,0);
\draw (4,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (4,1);
\draw (6,0) edge[out=0,in=0,->,thick] node[below,sloped] {$p$} (6,2);
\draw (6,2) edge[out=-45,in=45,->,thick] node[above,sloped] {$\phi$} (6,1);

\end{tikzpicture}
\caption{Time scales and process of IQS P-C}
\label{fig:timepc}
\end{figure}

\begin{figure}[!htpb]
\centering
\begin{tikzpicture}[every node/.style = {font=\normalsize}]

\node[greenblock](p1) at (0,0) {Save Old Solution};
\node[greenblock](p2) at (0,-2.5) {Save Current Solution};
\node[greenblock](p3) at (0,-5) {Interpolate Solution};
\node[blueblock](p4) at (0,-7.5) {Solve Temperature};
\node[greenblock] (p5) at (0,-10) {Update PRKE Params};
\node[greenblock] (p6) at (0,-12.5) {Solve PRKE};
\node[reddiamond] (check1) at (0,-17.5) {Check for \\ convergence ($p$)};
\node[reddiamond] (check2) at (7,-12) {If $t=t_{n+1}$};
\node[greenblock] (p7) at (8,-6) {Shape Solve};
\node[reddiamond] (check3) at (8,-2.5) {Check for \\ convergence ($\varphi$)};
%\node [above =0.1mm of p1] {\large{IQS Solve:}};
%\node [above =0.1mm of p5] {\large{Multi-physics:}};

%\tikzback{p1}{p1}{p4}{p4}{bk1}
%\tikzback{p5}{p5}{p5}{p5}{bk2}

\draw[->,ultra thick](p1.south) -- node[right] {$t=t_{n+1}$}(p2.north);
\draw[->,ultra thick](p2.south) -- node[right] {$t=t_{n}+\Delta t/3$}(p3.north);
\draw[->,ultra thick](p3.south) -- (p4.north);
\draw [->,ultra thick] (p4.south)-- (p5.north);
\draw[->,ultra thick](p5.south) -- (p6.north);
\draw[->,ultra thick](p6.south) -- (check1.north);
\draw[->,ultra thick](check1.west) -- node[above,sloped] {\large{no}} (-5,-17.5) |-  (p4.west);
\draw[->,ultra thick](check1.east) -| node[right] {\large{yes}} (check2.south);
\draw[->,ultra thick](check2.west) -- node[above,sloped] {\large{no}\small \qq $t=t+\Delta t/3$} (4,-5) -- (p3.east);
\draw[->,ultra thick](check2.north) --  (8,-10) -- node[right] {\large{yes}} (p7.south);
\draw[->,ultra thick](p7.north) -- (check3.south);
\draw[->,ultra thick](check3.west) -- node[above] {no} (p2.east);
\draw[->,ultra thick](check3.north) -- node[right] {yes} (8,0) -- node[above] {$t_n=t_{n+1}$} (p1.east);

\end{tikzpicture}
\caption{Visualization of Picard iteration and temperature update process for regular IQS}
\label{fig:proc}
\end{figure}

\begin{figure}[!htpb]
\centering
\begin{tikzpicture}[every node/.style = {font=\normalsize}]

\node[greenblock](p1) at (0,0) {Save Old Solution};
\node[greenblock](p2) at (0,-2.5) {Shape Solve};
\node[greenblock](p3) at (0,-5) {Save Current Solution};
\node[greenblock](p4) at (0,-7.5) {Interpolate Solution};
\node[blueblock](p5) at (0,-10) {Solve Temperature};
\node[greenblock] (p6) at (0,-12.5) {Update PRKE Params};
\node[greenblock] (p7) at (0,-15) {Solve PRKE};
\node[reddiamond] (check1) at (0,-18) {Check for \\ convergence ($p$)};
\node[reddiamond] (check2) at (8,-7.5) {If $t=t_{n+1}$};

%\tikzback{p1}{p1}{p4}{p4}{bk1}
%\tikzback{p5}{p5}{p5}{p5}{bk2}

\draw[->,ultra thick](p1.south) -- node[right] {$t=t_{n+1}$}(p2.north);
\draw[->,ultra thick](p2.south) -- (p3.north);
\draw[->,ultra thick](p3.south) -- node[right] {$t=t_{n}+\Delta t/3$} (p4.north);
\draw [->,ultra thick] (p4.south)-- (p5.north);
\draw[->,ultra thick](p5.south) -- (p6.north);
\draw[->,ultra thick](p6.south) -- (p7.north);
\draw[->,ultra thick](p7.south) -- (check1.north);
\draw[->,ultra thick](check1.west) -- node[above,sloped] {\large{no}} (-5,-18) |-  (p5.west);
\draw[->,ultra thick](check1.east) -| node[right] {\large{yes}} (check2.south);
\draw[->,ultra thick](check2.west) -- node[above,sloped] {\large{no}} node[below] {\small \qq $t=t+\Delta t/3$} (p4.east);
\draw[->,ultra thick](check2.north) -- node[right] {\large{yes}} (8,0) --  node[above] {\small $t_n=t_{n+1}$} (p1.east);

\end{tikzpicture}
\caption{Visualization of Picard iteration and temperature update process for IQS P-C}
\label{fig:procpc}
\end{figure}

\newpage
%%---------------------------------------------------------------------------%%
\section{\large Results}
%%---------------------------------------------------------------------------%%

This section describes the results of this implementation of temperature updates with the LRA benchmark.  Figure~\ref{fig:conv} shows the error convergence with 5 temperature updates per macro step.  It shows that with this amount of updates, IQS needs approximately a fourth of the time steps.  Figure~\ref{fig:mp} shows how various number of temperature updates affect the error with the same macro step size.

\begin{figure}[H]
\centering
\includegraphics[height=4in]{mp_figs/lra_mp_convergence.png}
\caption{Error convergence plot with 5 temperature updates per macro step}
\label{fig:conv}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[height=4in]{mp_figs/lra_mp.png}
\caption{Error plot with various temperature updates per macro step}
\label{fig:mp}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\newpage
%\closing
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-07-29 matlab/assemble_buckled_mass.m
Initial temperature feedback implementation

@@ -1,0 +1,51 @@
+function A = assemble_buckled_mass(fct_ptr,curr_time,T)
+
+% make the problem-data a global variable
+global npar
+
+% shortcuts
+porder= npar.porder;
+gn    = npar.gn;
+nel   = npar.nel;
+% n: linear system size
+n=npar.n;
+% ideally, we would analyze the connectivity to determine nnz
+nnz=npar.nnz;
+% allocate memory
+A=spalloc(n,n,nnz);
+% initialize local matrices
+a=zeros(porder+1,porder+1);
+% shape set
+b=npar.b;
+% quadrature stuff
+xq=npar.xq;
+wq=npar.wq;
+
+
+% loop over elements
+for iel=1:npar.nel
+    % element extremities
+    x0=npar.x(iel);
+    x1=npar.x(iel+1);
+    % jacobian of the transformation to the ref. element
+    Jac=(x1-x0)/2;
+    % get x values in the interval
+    x=(x1+x0)/2+xq*(x1-x0)/2;
+    % b and dbdx(:,:) are of size (nbr of xq values) x (porder+1),
+    % 2/dx is the 1d jacobian
+    % old: s=fct_ptr(curr_time,x)*Jac;
+    imat=npar.elem_to_mat(iel);
+    s=evaluate_buckled_material_prop(fct_ptr{imat},curr_time,x,T(iel))*Jac;
+    
+    % assemble
+    for i=1:porder+1
+        for j=1:porder+1
+            a(i,j)= dot(s.*wq.*b(:,i)    , b(:,j));
+        end
+    end
+    A(gn(iel,:),gn(iel,:)) = A(gn(iel,:),gn(iel,:)) + a;
+end
+
+return
+end
+

### 2016-07-29 matlab/assemble_buckled_time_dependent_operator.m

@@ -1,0 +1,25 @@
+function M = assemble_buckled_time_dependent_operator(curr_time)
+
+global dat npar
+
+n   = npar.n;
+nnz = npar.nnz;
+M   = sparse(3*n,3*n,nnz+2*n);
+
+% flux-flux
+IV = assemble_mass(dat.inv_vel,curr_time);
+if ~npar.set_bc_last
+    IV = apply_BC_mat_only(IV,npar.add_ones_on_diagonal);
+end
+
+% prec-prec
+I=speye(2*n);
+
+% save tr2.mat IV I;
+
+% assemble time dep operator
+M(1:n    ,1:n    ) = IV;
+M(n+1:3*n,n+1:3*n) = I;
+
+return
+end

### 2016-07-29 matlab/assemble_buckled_transient_operator.m

@@ -1,0 +1,54 @@
+function M = assemble_buckled_transient_operator(curr_time,T)
+
+global dat npar
+
+n   = npar.n;
+nnz = npar.nnz;
+M   = sparse(3*n,3*n,3*(nnz+n));
+
+D    = assemble_stiffness(   dat.cdiff   ,curr_time);
+A    = assemble_buckled_mass(dat.siga    ,curr_time,T);
+NFIp = assemble_mass(        dat.nusigf_p,curr_time) / npar.keff;
+NFId = assemble_mass(        dat.nusigf_d,curr_time) / npar.keff;
+NFI = assemble_mass(         dat.nusigf  ,curr_time);
+
+% flux-flux matrix
+tmp=NFIp-(D+A);
+% prec-flux matrix
+if ~npar.set_bc_last
+    NFId=apply_BC_mat_only(NFId,npar.add_zero_on_diagonal);
+    tmp =apply_BC_mat_only(tmp ,npar.add_zero_on_diagonal);
+end
+
+% generic matrix for precursors (it is diagonal because the unknown
+% precursor concentrations are not the FEM expansion values but int b_i C
+L=dat.lambda*speye(n);
+% Temperature operator
+LT = dat.alpha*NFI*dat.Pnorm;
+
+% prec-prec matrix
+Ldiag=L;
+% flux-prec matrix
+Loffd=L;
+
+if ~npar.set_bc_last
+    Ldiag=apply_BC_mat_only(L,npar.add_zero_on_diagonal);
+    Loffd=apply_BC_mat_only(L,npar.add_zero_on_diagonal);
+end
+
+% finally, build the rhs operator for the transient problem
+M(1:n    ,1:n    ) = tmp;
+M(1:n    ,n+1:2*n) =  Loffd;
+M(n+1:2*n,1:n    ) = NFId;
+M(n+1:2*n,n+1:2*n) = -Ldiag;
+M(2*n+1:3*n,1:n)   = LT;
+
+
+% A=apply_BC_mat_only(A,npar.add_zero_on_diagonal);
+% D=apply_BC_mat_only(D,npar.add_zero_on_diagonal);
+% NFIp=apply_BC_mat_only(NFIp,npar.add_zero_on_diagonal);
+%
+% save tr.mat D A NFIp NFId L;
+
+return
+end

### 2016-07-29 matlab/buckled_reference_solution.m

@@ -1,0 +1,72 @@
+function  [varargout] = buckled_reference_solution( t_end, u0)
+
+global dat npar io
+
+% options = odeset('RelTol',rtol,'AbsTol',atol,'InitialStep',1e-10,'OutputFcn',@odephas2,'OutputSel',[1:npar.ndofs]);
+% options = odeset('RelTol',rtol,'AbsTol',atol,'InitialStep',1e-10,'OutputFcn',@odeplot,'OutputSel',[1:npar.ndofs]);
+% options = odeset('RelTol',rtol,'AbsTol',atol);
+options = odeset('RelTol',1e-8,'AbsTol',1e-8,'InitialStep',1e-8,'Stats','on','MaxStep',1e-1,'Stats','off');
+% options = odeset('RelTol',1e-4,'AbsTol',1e-4);
+[t_ref,u_arr]=ode15s(@buckled_residual_TD_diffusion,[0 t_end],u0,options);
+u_arr=u_arr';
+
+% prepare for movie
+if io.make_movie
+    %# figure
+    figure, set(gcf, 'Color','white')
+    axis([0 dat.width 0 dat.max_y_val_movie]);
+    set(gca, 'nextplot','replacechildren', 'Visible','off');
+    %# preallocate
+    mov(1:ntimes) = struct('cdata',[], 'colormap',[]);
+    % movie name
+    movie_name = sprintf('PbID%d_method%s_nx%d.avi',dat.PbID,'ref_matlab',npar.nel);
+end
+
+% initial power level
+amplitude_norm=ones(length(t_ref),1);
+Ptot = dat.Ptot*ones(length(t_ref),1);
+
+%%% post-process solution %%%
+if io.plot_transient_figure
+    figure;
+end
+
+for it=1:length(t_ref)
+    % plot/movie
+    if io.plot_transient_figure
+        plot(npar.x_dofs,u_arr(1:npar.n,it));drawnow;
+        if io.make_movie && it>1, mov(it) = getframe(gca); end
+    end
+    % compute end time power for plotting
+    Ptot(it) = compute_power(dat.nusigf,t_ref(it),u_arr(1:npar.n,it));
+    % ratio of <u*,IVu> to its initial value
+    amplitude_norm(it) = (npar.phi_adj'*npar.IV*u_arr(1:npar.n,it))/npar.K0;
+end
+
+% make movie
+if io.plot_transient_figure && io.make_movie
+    close(gcf)
+    % save as AVI file
[…]

### 2016-07-29 matlab/buckled_residual_TD_diffusion.m

@@ -1,0 +1,22 @@
+function F = buckled_residual_TD_diffusion(curr_time,u)
+
+global npar
+sav_value = npar.set_bc_last;
+npar.set_bc_last = false;
+
+phi = u(1:npar.n);
+C = u(npar.n+1:npar.n*2);
+T = u(npar.n*2+1:end);
+
+TR = assemble_buckled_transient_operator(curr_time,T);
+M  = assemble_buckled_time_dependent_operator(curr_time);
+
+% build SS residual 
+F = M\(TR*u);
+% apply bc
+F=apply_BC_vec_only(F);
+
+npar.set_bc_last = sav_value;
+
+end
+

### 2016-07-29 matlab/buckled_time_marching_BF.m

@@ -1,0 +1,115 @@
+function  [varargout] = buckled_time_marching_BF( dt, ntimes, u0, T0, FUNHANDLE)
+
+global dat npar io
+
+% prepare for movie
+if io.make_movie
+    %# figure
+    figure, set(gcf, 'Color','white')
+    axis([0 dat.width 0 dat.max_y_val_movie]);
+    set(gca, 'nextplot','replacechildren', 'Visible','off');
+    %# preallocate
+    mov(1:ntimes) = struct('cdata',[], 'colormap',[]);
+    % movie name
+    movie_name = sprintf('PbID%d_method%s_nx%d_dt%g.avi',dat.PbID,func2str(FUNHANDLE),npar.nel,dt);
+end
+
+% initial stuff
+if strcmp(func2str(FUNHANDLE),'solve_TD_diffusion_an_prec') && strcmp(npar.an_interp_type,'lagrange')
+    order=npar.interpolation_order;
+else
+    order=1;
+end
+u=zeros(length(u0),order+1); u(:,end)=u0;
+T = T0;
+t = 0:dt:ntimes*dt;
+amplitude_norm=ones(ntimes+1,1);
+Ptot = dat.Ptot*ones(ntimes+1,1);
+average_temp=ones(ntimes+1,1); average_temp(1)=compute_power(dat.cdiff,0,T)/dat.width;
+max_temp=ones(ntimes+1,1); max_temp(1)=max(T);
+
+dat.ode.f_beg=zeros(npar.n,1);
+dat.ode.f_end=zeros(npar.n,1);
+
+
+%%% loop on time steps %%%
+for it=1:ntimes
+    
+    time_end=it*dt;
+    if io.console_print, fprintf('time end = %g \n',time_end); end
+    
+    u(:,1:end-1) = u(:,2:end);
+    
+    % solve time-dependent diffusion for flux
+    if it<order;
+        od = it;
+    else
+        od = order;
+    end
+    [u(:,end), T] = FUNHANDLE(u(:,end-od:end-1),dt,t(it-od+2:it+1),T);
+    
+    % update data for hermite interpolation
+    dat.ode.f_beg=dat.ode.f_end;
+    
+    % plot/movie
+    if io.plot_transient_figure
+        figure(1);
+        plot(npar.x_dofs,u(1:npar.n,end));drawnow;
+        if io.make_movie, mov(it) = getframe(gca); end
+    end
+    % compute end time power for plotting
[…]

### 2016-07-29 matlab/buckled_time_marching_IQS.m

@@ -1,0 +1,143 @@
+function  [varargout] = buckled_time_marching_IQS( dt, ntimes, u0, T0, FUNHANDLE)
+
+global dat npar io
+
+% prepare for movie
+if io.make_movie
+    %# figure
+    figure, set(gcf, 'Color','white')
+    axis([0 dat.width 0 dat.max_y_val_movie]);
+    set(gca, 'nextplot','replacechildren', 'Visible','off');
+    %# preallocate
+    mov(1:ntimes) = struct('cdata',[], 'colormap',[]);
+    % movie name
+    movie_name = sprintf('PbID%d_method%s_nx%d_dt%g.avi',dat.PbID,func2str(FUNHANDLE),npar.nel,dt);
+end
+
+%initial stuff
+if strcmp(func2str(FUNHANDLE),'solve_IQS_diffusion_an_prec') && strcmp(npar.an_interp_type,'lagrange')
+    order=npar.interpolation_order;
+else
+    order=1;
+end
+u_shape=zeros(length(u0),order+1);
+T = T0;
+X = zeros(2,order+1);
+tn = 0:dt:ntimes*dt;
+amplitude_norm=ones(ntimes+1,1);
+Ptot = dat.Ptot*ones(ntimes+1,1);
+average_temp=ones(ntimes+1,1); average_temp(1)=compute_power(dat.cdiff,0,T)/dat.width;
+
+% initial solution vector
+ u_shape(:,end)=u0;
+[~,beff_MGT]=compute_prke_parameters(0.,u0(1:npar.n));
+X(:,end)=[1;beff_MGT/dat.lambda];
+
+    dat.ode.f_beg=zeros(npar.n,1);
+    dat.ode.f_end=zeros(npar.n,1);
+
+
+time_prke_iqs=[];
+power_prke_iqs=[];
+
+%%% loop on time steps %%%
+for it=1:ntimes
+    
+    time_end=it*dt;
+    if io.console_print, fprintf('time end = %g \n',time_end); end
+    
+    u_shape(:,1:end-1) = u_shape(:,2:end); 
+    X(:,1:end-1) = X(:,2:end);
+    % solve time-dependent diffusion for flux
+    if it<order;
+        od = it;
+    else
+        od = order;
+    end
+    
+    % solve time-dependent diffusion for flux
[…]

### 2016-07-29 matlab/compute_buckled_prke_parameters.m

@@ -1,0 +1,36 @@
+function [rho_MGT,beff_MGT]=compute_buckled_prke_parameters(curr_time,shape,T)
+
+global dat npar
+
+phi_adjoint = npar.phi_adj;
+
+D    = assemble_stiffness(dat.cdiff   ,curr_time);
+A    = assemble_buckled_mass(dat.siga ,curr_time,T);
+NFI  = assemble_mass(     dat.nusigf  ,curr_time) / npar.keff;
+NFId = assemble_mass(     dat.nusigf_d,curr_time) / npar.keff;
+IV   = assemble_mass(     dat.inv_vel ,curr_time);
+
+IV   = apply_BC_mat_only(IV,npar.add_zero_on_diagonal);
+NFId = apply_BC_mat_only(NFId,npar.add_zero_on_diagonal);
+
+PmM = apply_BC_mat_only(NFI-D-A,npar.add_zero_on_diagonal);
+% rho  = phi_adjoint' * (NFI-D-A) * shape;
+rho  = phi_adjoint' * (PmM) * shape;
+beff = phi_adjoint' * NFId * shape;
+% MGT  = phi_adjoint' * IV * shape; % obviously, MGT = npar.K0 as well
+
+
+% rho_MGT  = rho/MGT;
+% beff_MGT = beff/MGT;
+rho_MGT  = rho/npar.K0;
+beff_MGT = beff/npar.K0;
+
+% prec_MGT = phi_adjoint' * C / MGT; 
+% % obviously, this is the same as beff_MGT/dat.lambda because, at steady
+% % state, C = NFid * phi0 / lambda, thus: 
+% % phi_adjoint' * C = beff / lambda, and finally
+% % phi_adjoint' * C / MGT = beff_MGT /lambda 
+
+
+return
+end

### 2016-07-29 matlab/compute_temp.m

@@ -1,0 +1,27 @@
+function T = compute_temp(Told,phi,t)
+
+global dat
+
+[~,n] = size(phi);
+if n ~= length(t)
+    error('phi and time must be same length')
+end
+
+fissrc = zeros(size(phi));
+for i=1:n
+    NF = assemble_mass(dat.nusigf,t(i));
+    fissrc(:,i) = NF*phi(:,i)*dat.Pnorm;
+end
+dt = t(2)-t(1);
+
+switch n
+    case 1
+        T = Told + dat.alpha*dt*fissrc;
+    case 2
+        T = Told + dat.alpha*dt*(fissrc(:,2) + fissrc(:,1))/2;
+    otherwise
+        error('Integration scheme not known')
+end
+
+return
+end

### 2016-07-29 matlab/compute_temp_IQS.m

@@ -1,0 +1,29 @@
+function T = compute_temp_IQS(Told,phi,t,p)
+
+global dat
+
+[~,n] = size(phi);
+if length(p) ~= length(t)
+    error('p and time must be same length')
+elseif n~=2
+    error('must have two shape functions')    
+end
+t1 = t(1);
+t2 = t(end);
+dt = t2-t1;
+
+NF = assemble_mass(dat.nusigf,t1);
+fissrc_old = NF*phi(:,1);
+NF = assemble_mass(dat.nusigf,t2);
+fissrc_new = NF*phi(:,2);
+
+pp = interp1(t,p,'linear','pp');
+A1 = @(t) (t2-t)/dt.*ppval(pp,t);
+A2 = @(t) (t-t1)/dt.*ppval(pp,t);
+a1 = integral(A1,t1,t2);
+a2 = integral(A2,t1,t2);
+
+T = Told + dat.alpha*(a2*fissrc_new + a1*fissrc_old)*dat.Pnorm;
+
+return
+end

### 2016-07-29 matlab/create_buckled_material_prop.m

@@ -1,0 +1,4 @@
+function aux=create_buckled_material_prop(time_dep,values,times,space_dep,space_function_handle,gamma,T0)
+
+aux = create_material_prop(time_dep,values,times,space_dep,space_function_handle);
+aux.fT = @(T) (1 + gamma*(sqrt(T)-sqrt(T0)));

### 2016-07-29 matlab/create_material_prop.m

@@ -36,5 +36,7 @@ else
     error('unknown spatial dependence');
 end
 
+aux.fT = @(T) 0.*T+1;
+
 return
 end

### 2016-07-29 matlab/evaluate_buckled_material_prop.m

@@ -1,0 +1,6 @@
+function values=evaluate_buckled_material_prop(mat_prop,time,x,T)
+
+values = mat_prop.ft(time) * mat_prop.fx(x) * mat_prop.fT(T);
+
+return
+end

### 2016-07-29 matlab/kinetics_driver_conv.m

@@ -9,18 +9,18 @@ warning('OFF','MATLAB:interp1:ppGriddedInterpolant')
 global npar io
 
 % verbose/output parameters
-io.console_print         = false;
+io.console_print         = true;
 io.plot_transient_figure = false;
-io.plot_power_figure     = false;
+io.plot_power_figure     = true;
 io.make_movie            = false;
 io.save_flux             = false;
-io.print_progress        = true;
+io.print_progress        = false;
 io.figID = 99;
 % one of the two choices for applying BC
 npar.set_bc_last=true;
 
 % select problem
-pbID=11; refinements=1;
+pbID=11; refinements=2;
 problem_init(pbID,refinements);
 
 % compute fundamental eigenmode
@@ -34,20 +34,23 @@ C0 = kinetics_init(phi0,curr_time);
 u=[phi0;C0];
 % save a copy of it
 u0=u;
+T0 = ones(size(phi0))*300;
 
 % time steping data
-t_end = 1.28;
-% t_end = 1.3;
+t_end = 1.2;
+% t_end = 3;
 % t_end = 0.1;
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%% MATLAB time discretization of
 %%%   the TD neutron diffusion eq and precursors eq
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-amplitude_norm_ref = reference_solution( t_end, u0);
+% amplitude_norm_ref = reference_solution( t_end, u0);
+% amplitude_norm_ref = buckled_reference_solution( t_end, [u0; T0]);
+amplitude_norm_ref = buckled_time_marching_BF( 3/1000, 1000, u0, T0, @solve_buckled_TD_diffusion_elim_prec);
 
-nn=5;
-ntimes = 2.^(0:nn-1)*4;
+nn=3;
+ntimes = 2.^(0:nn-1)*50;
 dt = t_end./ntimes;
 
 % Interpolation type of shape for IQS prke parameters
@@ -74,22 +77,23 @@ npar.hermite_prec_update=true;
 
 % Used of IQS_PC, determines how precursors are solved after diffusion
[…]

### 2016-07-29 matlab/kinetics_init.m

@@ -25,6 +25,8 @@ npar.K0 = K0;
 
 % total power
 dat.Ptot = compute_power(dat.nusigf,curr_time,phi);
+power = dat.Ptot*dat.kappa;
+dat.Pnorm = 0.01/power;
 
 if ~npar.set_bc_last % 0=neumann, 1=robin, 2=dirichlet
     % add precursors BC values

### 2016-07-29 matlab/plot_power_level.m

@@ -14,6 +14,9 @@ switch method_name
         curr_leg = 'space-time-ANALY';
     
     case 'solve_TD_diffusion_elim_prec'
+        curr_leg = 'space-time-elim';    
+    
+    case 'solve_buckled_TD_diffusion_elim_prec'
         curr_leg = 'space-time-elim';
     
     case 'solve_IQS_diffusion_an_prec'
@@ -21,11 +24,17 @@ switch method_name
     
     case 'solve_IQS_diffusion_elim_prec'
         curr_leg = 'IQS-elim';
+
+    case 'solve_buckled_IQS_diffusion_elim_prec'
+        curr_leg = 'IQS-elim';
     
     case 'solve_IQS_PC_diffusion_an_prec'
         curr_leg = 'IQS-PC-an';
     
     case 'solve_IQS_PC_diffusion_elim_prec'
+        curr_leg = 'IQS-PC-elim';
+    
+    case 'solve_buckled_IQS_PC_diffusion_elim_prec'
         curr_leg = 'IQS-PC-elim';
     
     case 'solve_IQS_diffusion_td_prec'
@@ -58,16 +67,16 @@ if isempty(a)
 else
     leg=char(char(a),curr_leg);
 end
-plot(t,p,'o-');
+plot(t,log10(p),'o-');
 
 % if IQS
-nVarargs = length(varargin);
-if nVarargs ==2
-    t_iqs_fine = varargin{1};
-    p_iqs_fine = varargin{2};
-    plot(t_iqs_fine,p_iqs_fine,'c.-');
-    leg = char(leg, char(strcat(curr_leg, ' fine')) );
-end
+% nVarargs = length(varargin);
+% if nVarargs ==2
+%     t_iqs_fine = varargin{1};
+%     p_iqs_fine = varargin{2};
+%     plot(t_iqs_fine,log10(p_iqs_fine),'c.-');
+%     leg = char(leg, char(strcat(curr_leg, ' fine')) );
+% end
 
 legend(leg,'Location','Best');
 

### 2016-07-29 matlab/problem_init.m

@@ -15,6 +15,11 @@ dat.rod_mov.t_beg_2=1.0; dat.rod_mov.t_end_2=2.7;
 dat.beta_tot=600e-5;
 dat.lambda=0.1;
 dat.invvel=1e-3;
+dat.nu=2.43;
+
+% Temperature parameters
+dat.alpha = 0.;
+dat.T0 = 300;
 
 % assign function pointers to the various physical coeffcients
 switch problem_ID
@@ -118,16 +123,20 @@ switch problem_ID
     case 11
         % have material identifiers: 20 regions. 2 materials. 1 rod movement
         n_regions = 20; % assumption: each region has the same width
-        region_width=400/n_regions;
+        region_width=200/n_regions;
         dat.width = region_width * n_regions;
         
+        dat.kappa = 3.204e-11;
+        dat.alpha = 3.83e-11;
+        
         imat = ones(n_regions,1);
-        imat(5) = 2;
+        imat(3)  = 2;
+        imat(18) = 2;
         
         b=dat.beta_tot;
         iv=dat.invvel;
         dat.cdiff{1}   = create_material_prop('constant_in_time',1        ,[],'constant_in_space',0);
-        dat.siga{1}    = create_material_prop('constant_in_time',1.1      ,[],'constant_in_space',0);
+        dat.siga{1}    = create_buckled_material_prop('constant_in_time',1.1,[],'constant_in_space',0,3.034e-3,dat.T0);
         dat.nusigf{1}  = create_material_prop('constant_in_time',1.1      ,[],'constant_in_space',0);
         dat.nusigf_p{1}= create_material_prop('constant_in_time',1.1*(1-b),[],'constant_in_space',0);
         dat.nusigf_d{1}= create_material_prop('constant_in_time',1.1*b    ,[],'constant_in_space',0);
@@ -142,8 +151,8 @@ switch problem_ID
             dat.inv_vel{id}  = dat.inv_vel{1} ;
             dat.ext_src{id}  = dat.ext_src{1}    ;
         end
-        times = [dat.rod_mov.t_beg_1 dat.rod_mov.t_beg_2];
-        dat.siga{2} = create_material_prop('ramp_in_time',[1.1 1.08],times,'constant_in_space',0);
+        times = [0 2];
[…]

### 2016-07-29 matlab/solve_buckled_IQS_PC_diffusion_elim_prec.m

@@ -1,0 +1,178 @@
+function [u_end, X,t,y,T] = solve_buckled_IQS_PC_diffusion_elim_prec(u,X,dt_macro,time_end,Told)
+
+global io dat npar
+
+% shortcuts
+lambda = dat.lambda;
+C_old = u(npar.n+1:end);
+dt = dt_macro;
+time_beg = time_end-dt;
+
+max_iter_iqs = npar.max_iter_iqs;
+tol_iqs      = npar.tol_iqs;
+
+npar.theta_old=[];
+
+% save values at beginning of macro time step: they are needed in the IQS iteration
+X_beg=X;
+u_beg=u;
+
+% for clarity, I also single out the shape function at the beginning/end of the macro time step
+% IV   = assemble_mass(     dat.inv_vel ,time_end);
+z = (npar.phi_adj)'*npar.IV*u(1:npar.n)/npar.K0;
+shape_beg=u(1:npar.n)/z; dat.ode.shape_beg = shape_beg;
+
+for iter = 1: max_iter_iqs
+    
+    % solve time-dependent diffusion for flux
+    [u_end, auxphi] = solve_buckled_TD_diffusion_elim_prec(u_beg,dt_macro,time_end,Told);
+%     [u_end] = solve_TD_diffusion_an_prec(u_beg,dt_macro,time_end);
+
+    % get a predicted value of the end flux
+    flux_end = u_end(1:npar.n);
+    C_new_tmp= u_end(npar.n+1:end);
+    
+    % get a shape
+    z = (npar.phi_adj)'*npar.IV*flux_end/npar.K0;
+    shape_end = u_end(1:npar.n)/ z;
+    
+    % solve for amplitude function
+    if strcmpi(npar.prke_solve,'matlab')
+        [X,w,t,y] =  solve_prke_ode(X_beg,dt_macro,time_end,shape_beg,shape_end);
+%         [X,dpdt,t,y] =  solve_prke_ode_events(X_beg,dt_macro,time_end,shape_beg,shape_end);
+    else
+        [X,dpdt,t,y,T] =  solve_prke_iqs_buckled(X_beg,dt_macro,time_end,shape_beg,shape_end,Told,npar.n_micro,npar.freq_react);
+    end
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    % assemble IQS
+    
+    % interpolating polynomial
+    if strcmpi(npar.prke_solve,'matlab')
[…]

### 2016-07-29 matlab/solve_buckled_IQS_diffusion_elim_prec.m

@@ -1,0 +1,151 @@
+function [u_shape, X,t,y,T] = solve_buckled_IQS_diffusion_elim_prec(u_shape,X,dt_macro,time_end,Told)
+
+global io dat npar
+
+% shortcuts
+lambda = dat.lambda;
+dt = dt_macro;
+C_old = u_shape(npar.n+1:end);
+
+max_iter_iqs = npar.max_iter_iqs;
+tol_iqs      = npar.tol_iqs;
+
+npar.theta_old=[];
+% save values at beginning of macro time step: they are needed in the IQS iteration
+X_beg=X;
+% for clarity, I also single out the shape function at the beginning/end of the macro time step
+shape_beg=u_shape(1:npar.n);
+shape_end=shape_beg;
+
+% SDIRK solve
+% Yi = yn + h sum_j aij f(tj,Yj)
+% gives (f=Qy+z)
+%  [I-aii.h.Q(ti)]Yi = yn + h.aii zi + h sum(j=1..i-1) ( aij {Q(tj)Yj + zj} )
+rk=npar.rk;
+% beginning of the time interval
+time_beg = time_end - dt;
+% storage for temp SDIRK quantities (-lambda Ci + Bi Phii)
+f=zeros(length(shape_beg),rk.s);
+fC=f;
+NFId_old = assemble_mass(dat.nusigf_d,time_beg) / npar.keff;
+
+for iter = 1: max_iter_iqs
+    
+    % solve for amplitude function over the entire macro-step
+    if strcmpi(npar.prke_solve,'matlab')
+        [X,w,t,y] =  solve_prke_ode(X_beg,dt_macro,time_end,shape_beg,shape_end);
+    else
+        [X,dpdt,t,y,T] =  solve_prke_iqs_buckled(X_beg,dt_macro,time_end,shape_beg,shape_end,Told,npar.n_micro,npar.freq_react);
+    end
+    
+    % interpolating polynomial
+    if strcmpi(npar.prke_solve,'matlab')
+        if npar.int_order==3
+            pp = interp1(t,y,'spline','pp');
+        elseif npar.int_order==2
+            pp = interp1(t,y,'pchip','pp');
+        elseif npar.int_order==1
+            pp = interp1(t,y,'linear','pp');
+        end
+    else
+        pp = interp1(t,y,'linear','pp');
+    end
+    
+    % extract details from piece-wise polynomial by breaking it apart
+    [breaks,coefs,l,k,d] = unmkpp(pp);
[…]

### 2016-07-29 matlab/solve_buckled_TD_diffusion_elim_prec.m

@@ -1,0 +1,105 @@
+function [ varargout ] = solve_buckled_TD_diffusion_elim_prec(u,dt,time_end,T)
+
+global dat npar
+
+% split initial solution vector into flux and precursors:
+Phi_old = u(1:npar.n);
+C_old   = u(npar.n+1:end);
+
+% shortcut
+lambda = dat.lambda;
+% %
+
+% SDIRK solve
+% Yi = yn + h sum_j aij f(tj,Yj)
+% gives (f=Qy+z)
+%  [I-aii.h.Q(ti)]Yi = yn + h.aii zi + h sum(j=1..i-1) ( aij {Q(tj)Yj + zj} )
+rk=npar.rk;
+
+% beginning of the time interval
+time_beg = time_end - dt;
+% storage for temp SDIRK quantities (-lambda Ci + Bi Phii)
+f=zeros(length(Phi_old),rk.s);
+fC=f;
+% NFId_old = assemble_mass(dat.nusigf_d,time_beg) / npar.keff;
+auxPhi=zeros(length(Phi_old),rk.s);
+
+for i=1:rk.s
+    % compute stage time
+    ti = time_beg + rk.c(i)*dt;
+    
+    % build system matrix
+    D    = assemble_stiffness(   dat.cdiff   ,ti);
+    A    = assemble_buckled_mass(dat.siga    ,ti,T);
+    NFIp = assemble_mass(        dat.nusigf_p,ti) / npar.keff;
+    IV   = assemble_mass(        dat.inv_vel ,ti);
+    NFId_new = assemble_mass(    dat.nusigf_d,ti) / npar.keff;
+    
+    % flux-flux matrix
+    TR=NFIp-(D+A);
+    
+    % expressions for the precursors
+	% Ci = [ C_old + h aii FISdi.Phii + sum_{j<i} h aij(-lambda.Cj + FISdj.Phij)]/(1+lambda h aii) 
+    % let fci = -lambda.Ci + FISdi.Phii
+	Ci = C_old; 
+    for j=1:i-1
+        Ci = Ci + rk.a(i,j)*dt*fC(:,j);
+    end
+	deno = (1+lambda*rk.a(i,i)*dt);
+	Ci = Ci/deno; % this is Ci without: h.aii.FISdi.Phii/deno
+    
+    % build transient matrix
+    TR = TR + (lambda * ( rk.a(i,i)*dt / deno )) * NFId_new; 
+    
+    % build system matrix
+    A = IV-rk.a(i,i)*dt*TR;
[…]

### 2016-07-29 matlab/solve_prke_iqs_buckled.m

@@ -1,0 +1,65 @@
+function [X,dpdt,t,y,T] =  solve_prke_iqs_buckled(X,dt_macro,time_end,shape_beg,shape_end,Told,n_micro,freq_react)
+
+% make the problem-data a global variable
+global dat npar
+
+% time step for prke solve
+dt = dt_macro/n_micro;
+
+% number of reactivity updates during one macro time step
+n_react = n_micro / freq_react;
+if abs(floor(n_react)-n_react) > 1e-10
+    error('n_micro / freq_react must be an integer');
+end
+% create a list of times when reactivity will be recomputed for this macro
+% time step
+time_beg = time_end - dt_macro;
+times_react_update = time_beg + linspace(0,n_react,n_react+1) * dt_macro/n_react;
+
+% storage for t and y (for post-processing)
+t(1) = time_beg;
+y(1) = X(1);
+etol = 1e-12;
+
+% loop over micro time steps
+for i=1:n_react
+    t_beg = times_react_update(i);
+    t_end = times_react_update(i+1);
+    % weights for linear interpolation of the shape
+    w1 = (time_end-t_beg)/dt_macro;
+    w2 = (t_beg-time_beg)/dt_macro;
+    % shape interpolation
+    shape = shape_beg * w1 + shape_end * w2;
+    X_old = X;
+    t = [t t_end];
+    y = [y X(1)];
+    err = etol+1;
+    mp_iter = 0;
+    while err > etol
+        mp_iter = mp_iter+1;
+        X_prev = X;
+        X = X_old;
+        T = compute_temp_IQS(Told,[shape_beg shape],t,y);
+        [rho_MGT(i+1),beff_MGT(i+1)]=compute_buckled_prke_parameters(t_end,shape,T);
+        rho_MGT_iqs  = linspace(rho_MGT(i) ,rho_MGT(i+1) ,freq_react+1);
+        beff_MGT_iqs = linspace(beff_MGT(i),beff_MGT(i+1),freq_react+1);
+        for it=1:freq_react
+            % build prke matrix
+            A=[(rho_MGT_iqs(it+1)-beff_MGT_iqs(it+1)) dat.lambda ; ...
+                beff_MGT_iqs(it+1)                   -dat.lambda];
+            % solve
+            X=(eye(2)-dt*A)\X;
+            % storage
+            t((i-1)*freq_react+it+1)=t((i-1)*freq_react+it)+dt;
[…]

### 2016-08-15 doc/ANS_winter_2016/ANS_summary.tex
ANS Summary: Final Submission

@@
 \frac{1}{v^g}\frac{\partial \Psi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \Psi^{g'} - L^g\Psi^g + S_{d}^g
+\label{eq:flux}
 \ee 
-\label{eq:flux}
 \be
@@
 \be
-\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
+\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}p-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
 \ee
@@
 \be
-\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},\sum_{g'}(H^{' \to g}g+P_p^{g' \to g}-L^{g'}\delta_{g'g})\psi^{g'}\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
+\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},\sum_{g'}(H^{g' \to g}+P_p^{g' \to g}-L^{g'}\delta_{g'g})\psi^{g'}\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
 \label{eq:rmb}
@@
 If the step is too large, the step will be repeated with a smaller step calculated with the resulting error.
-The error of the time step is approximated by Equation \ref{eq:edt2}, where $\varphi^g_1$ and $\varphi^g_2$ are the solutions with the full step and half step, respectively.
-$\varphi$ is changed to $\phi$ for regular flux evaluation and IQS P-C.
+The error of the time step is approximated by Equation \ref{eq:edt2}, where $\Psi^g_{\Delta t}$ and $\Psi^g_{\Delta t/2}$ are the solutions with the full step and half step, respectively.
 If $e_n > e_{max}$ the time step is repeated; if $e_n < e_{max}$ the system moves to the next time step.
@@
 \be
-e_n = \frac{\norm{\sum_{g=1}^G\int_{4\pi}\Psi^g_2 d^2\Omega - \sum_{g=1}^G\int_{4\pi}\Psi^g_1 d^2\Omega}}{\text{max}\left(\norm{\sum_{g=1}^G\int_{4\pi}\Psi^g_2 d^2\Omega},\norm{\sum_{g=1}^G \int_{4\pi}\Psi^g_1 d^2\Omega}\right)}
+e_n = \frac{\norm{\sum_{g=1}^G\int_{4\pi}\Psi^g_{\Delta t/2} d^2\Omega - \sum_{g=1}^G\int_{4\pi}\Psi^g_{\Delta t} d^2\Omega}}{\text{max}\left(\norm{\sum_{g=1}^G\int_{4\pi}\Psi^g_{\Delta t/2} d^2\Omega},\norm{\sum_{g=1}^G \int_{4\pi}\Psi^g_{\Delta t} d^2\Omega}\right)}
 \label{eq:edt2}
@@
 \be
-\Delta t_{new} = \Delta t_{old} \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{\mu+1}}
+\Delta t_{new} = \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{\mu+1}}
 \label{eq:dt2}
@@
 
-This benchmark problem originates from the Argonne National Lab Benchmark Problem Book.
-It is a 2D, 2-group reactor core model with no reflector region, the geometry, material properties, and transient perturbation can be seen in \cite{BPB}.
+This benchmark problem originates from the Argonne National Lab Benchmark Problem Book (8-A1-1) \cite{ANL_BPB}.
+It is a 2D, 2-group reactor core model with no reflector region, the geometry, material properties, and transient perturbation can be seen in \cite{TWIGL_benchmark}.
 Figure \ref{fig:TWIGL_power} shows the IQS  solution as compared with the Brute Force solution with minimal time steps.
@@
 The results show that IQS performs exceptionally well compared to brute force for this highly transient example.
+
+%\begin{figure}[!htbp]
+%\begin{center}
+%\includegraphics[width=0.45\textwidth]{TWIGL_regions.jpg}
+%\caption{TWIGL benchmark problem description}
+%\label{fig:TWIGL_reg}
+%\end{center}
+%\end{figure}
+%
+%\begin{table}[!htbp]
+%\begin{center}
+%\resizebox{0.45\textwidth}{!}{
+%\begin{tabular}{llllllll}
+%\hline
+%  &  &  &  &  &  &  \multicolumn{2}{c}{$\underline{\Sigma_s (cm^{-1})} $} \\
+%Material & Group & $D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu\Sigma_f (cm^{-1})$ & $\chi$ & $g \rightarrow 1$ & $g \rightarrow 2$ \\
+%\hline
+%1 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01 \\
+%  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
+%2 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01  \\
+%  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
+%3 & 1 & 1.3 & 0.008 & 0.003 & 1.0 & 0.0 & 0.01  \\
+%  & 2 & 0.5 & 0.050 & 0.060 & 0.0 & 0.0 & 0.00  \\
+%\hline
+%  & $\nu$ & $v_1 (cm/s)$ & $v_2 (cm/s)$ & $\beta$ & $\lambda (1/s)$ &   &   \\
+%\hline
+%  & 2.43 & 1.0E7 & 2.0E5 & 0.0075 & 0.08 &   &   \\
+%\hline
+% \multicolumn{8}{l}{Material 1 ramp perturbation:} \\
+%\multicolumn{8}{l}{$\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (1-0.11667t) \quad t \leq 0.2 s$} \\
+%\multicolumn{8}{l}{$\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (0.97666t) \quad t > 0.2 s$} \normalsize
+%\end{tabular}}
+%\end{center}
+%\caption{TWIGL benchmark material properties and slope perturbation}
+%\label{tab:TWIGL_mat}
+%\end{table}
 
@@
 1 &	0.05    &	0.00012677 &	9    &	29   &	0.03380433 &	4  &	9   \\
-2 &	0.01    &	3.5555e-05 &	5   &	40   &	0.00263068 &	5  &	12  \\
-3 &	0.005   &	4.0364e-05 &	5   &	40   &	0.00160486 &	6  &	21  \\
-4 &	0.001   &	0.00294822 &	5   &	36   &	1.7527e-05 &	10 &	35  \\
-5 &	0.0005  &	0.00099778 &	6   &	55   &	1.4185e-05 &	16 &	74  \\
-6 &	0.0001  &	0.00019510 &	7   &	65   &	6.2903e-06 &	19 &	78  \\
-7 &	5.0e-05 &	0.00018372 &	12  &	163  &	1.5247e-06 &	24 &	92  \\
-8 &	1.0e-05 &	8.0564e-05 &	379 &	5729 &	9.8321e-07 &	48 &	210 \\
+2 &	0.01    &	3.5555e-05 &	11   &	35   &	0.00263068 &	5  &	12  \\
+3 &	0.005   &	4.0364e-05 &	11   &	31   &	0.00160486 &	6  &	21  \\
+4 &	0.001   &	0.00294822 &	33   &	122  &	1.7527e-05 &	10 &	35  \\
+5 &	0.0005  &	0.00099778 &	39   &	131  &	1.4185e-05 &	16 &	74  \\
+6 &	0.0001  &	0.00019510 &	78   &	236  &	6.2903e-06 &	19 &	78  \\
+7 &	5.0e-05 &	0.00018372 &	112  &	342  &	1.5247e-06 &	24 &	92  \\
+8 &	1.0e-05 &	8.0564e-05 &	263  &	794  &	9.8321e-07 &	48 &	210 \\
 \hline

### 2016-09-02 doc/NEAMS_Report/NEAMS_Report_Uncollided_Flux.tex
update to report

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{scrartcl} 
%########################### Preferences #################################

\usepackage{vmargin}
\usepackage{color}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amstext}
\usepackage{amsbsy}
%\usepackage{mathbbol}
\usepackage{graphicx} 
\usepackage{verbatim}
\usepackage{csvsimple} 
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{listings}
\usepackage[numbers,sort&compress]{natbib}[2010/09/13]

%  new definitions
\renewcommand{\div}{\vec{\nabla}\! \cdot \!}
\newcommand{\grad}{\vec{\nabla}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert_{L^2}}
\newcommand{\rattlesnake}{Rattlesnake }
% extra space
\newcommand{\qq}{\quad\quad}
% common reference commands`
\newcommand{\eqt}[1]{Equation~\ref{#1}}                     % equation
\newcommand{\fig}[1]{Figure~\ref{#1}}                      % figure
\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table
\newcommand{\sct}[1]{Section~\ref{#1}}                   % section
\newcommand{\app}[1]{Appendix~\ref{#1}}                   % appendix

\newcommand{\bs}[1]{\mathbf{#1}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\keff}{k_\textit{eff}}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\vn}{\vec{n}}
\newcommand{\vel}{\vec{\mathrm{v}}}
\newcommand{\adj}{\Phi^\dagger_0}
\newcommand{\tcr}[1]{\textcolor{red}{#1}}
\newcommand{\tcb}[1]{\textcolor{blue}{#1}}

% tikz stuff
\usetikzlibrary{shapes,arrows,positioning}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\tikzstyle{greenblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=green!20, ultra thick, minimum width=60mm, minimum height=15mm,]
\tikzstyle{blueblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=blue!20, ultra thick, minimum width=60mm, minimum height=15mm]
\tikzstyle{reddiamond}=[diamond, draw, align=center, top color=white, bottom color=red!20, ultra thick, minimum width=60mm, aspect=2]
\newcommand{\tikzback}[5]{
 \begin{pgfonlayer}{background}
  \path (#1.west |- #2.north)+(-0.5,0.75) node (a1) {};
  \path (#3.east |- #4.south)+(+0.5,-0.25) node (a2) {};
  \path[fill=yellow!10,rounded corners, draw=black!100, dashed] (a1) rectangle (a2);
   \path (#3.east |- #2.north)+(0,1.25)--(#1.west |- #2.north) node[midway] (#5-n) {};
   \path (#3.east |- #2.south)+(0,-0.35)--(#1.west |- #2.south) node[midway] (#5-s) {};
   \path (#1.west |- #2.north)+(-0.75,0)--(#1.west |- #4.south) node[midway] (#5-w) {};
 \end{pgfonlayer}}

% input stuff
\lstset{%
%  language=C++,
  showstringspaces=false,
% basicstyle=\scriptsize\ttfamily,
  basicstyle=\footnotesize\ttfamily,
  commentstyle=\color{inl@green},
%  keywordstyle=\bfseries,
%  escapeinside=$$
%  framesep=0pt,
%  rulesep=0pt,
  frame=single
}
% Make nice big tildes inside lstlistings instead of
% tiny ones.
\lstset{
    literate={~} {$\sim$}{1}
}

% ********* Caption Layout ************
%\usepackage{ccaption} % allows special formating of the captions
%\captionnamefont{\bf\footnotesize\sffamily} % defines the font of the caption name (e.g. Figure: or Table:)
%\captiontitlefont{\footnotesize\sffamily} % defines the font of the caption text (same as above, but not bold)
%\setlength{\abovecaptionskip}{0mm} %lowers the distance of captions to the figure
\setlength\parindent{0pt}
\setlength{\oddsidemargin}{1in}
\setlength{\evensidemargin}{1in}
\setlength{\textwidth}{6.5in}
% ********* Header and Footer **********
% This is something to play with forever. I use here the advanced settings of the KOMA script

\title{Uncollided Flux Techniques for Discrete-Ordinate Radiation Transport Solutions in \rattlesnake}
\subtitle{Uncollided Flux Treatment \rattlesnake}
\author{ \normalsize
  \textbf{Jean C. Ragusa$^\dagger$, Derek Gaston$^\star$, Mark D. DeHart$^\star$} \\
 \normalsize $^\dagger$Texas A\&M University, College Station, TX, USA\\
 \normalsize $^\star$Idaho National Laboratory, Idaho Falls, ID, USA\\
 \normalsize \href{jean.ragusa@tamu.edu}{jean.ragusa@tamu.edu}, \href{derek.gaston@inl.gov}{derek.gaston@inl.gov}, \href{mark.dehart@inl.gov}{mark.dehart@inl.gov}
}

\hypersetup{
  colorlinks=true,
  urlcolor=blue,
}

\pagestyle{fancy}
\fancyhf{}
\lhead{INL/EXT-16-xxxxx}
\rfoot{DRAFT}
\cfoot{\thepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% boxes
\usepackage{color}
\definecolor{myblue}{rgb}{.8, .8, 1}
\usepackage{empheq}

\newlength\mytemplen
\newsavebox\mytempbox

\makeatletter
\newcommand\mybluebox{%
    \@ifnextchar[%]
       {\@mybluebox}%
       {\@mybluebox[0pt]}}

\def\@mybluebox[#1]{%
    \@ifnextchar[%]
       {\@@mybluebox[#1]}%
       {\@@mybluebox[#1][0pt]}}

\def\@@mybluebox[#1][#2]#3{
    \sbox\mytempbox{#3}%
    \mytemplen\ht\mytempbox
    \advance\mytemplen #1\relax
    \ht\mytempbox\mytemplen
    \mytemplen\dp\mytempbox
    \advance\mytemplen #2\relax
    \dp\mytempbox\mytemplen
    \colorbox{myblue}{\hspace{1em}\usebox{\mytempbox}\hspace{1em}}}

\makeatother

%################ End Preferences, Begin Document #####################

%\pagestyle{plain} % on headers or footers on the first page

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% set relative directory for figures. can be changed at any time later with \renewcommand
\newcommand{\FiguresDir}{./figs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
\tableofcontents
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{Abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This report reviews uncollided flux techniques (first and last collision methods) to be implemented in 
the Rattlesnake $S_N$ code in order to mitigate ray effects in modeling the TREAT reactor+hodoscope system. Angular
discretization techniques ($S_N$ and $P_N$) for the transport equation are notoriously poor at capturing effectively
streaming effects.

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\par
This report reviews uncollided flux techniques (first and last collision methods) to be implemented in 
the Rattlesnake $S_N$ code in order to mitigate ray effects in modeling the TREAT reactor+hodoscope system. Angular
discretization techniques ($S_N$ and $P_N$) for the transport equation are notoriously poor at capturing accurately
streaming effects. The uncollided component of the angular flux solution is the most anisotropic part and it can be
difficult for discrete-ordinate methods to accurately represent it; this phenomena is well-known and termed ``ray effects''.
However, it has long been recognized that an analytical or semi-analytical treatment of the uncollided flux, coupled with a discrete-ordinate treatment of the collided flux, can yield dramatic improvements in accuracy and computational efficiency. In this report, we present an algorithm for semi-analytical calculation of the uncollided flux. The algorithm seeks to compute
\begin{enumerate}
\item The existing uncollided flux at each point in a collect of point on the surface of the problem domain, with the angular flux reported for each ``source point''. The direction is fro the source point to the surface point. The source points are chosen to be spatial quadrature points that integrate the source volume.
\item Spatial and angular distributions of the uncollided flux in every cell in the problem. These provide all of the information necessary to form the first-collision source in each cell.
\end{enumerate}
The proposed algorithm is tailored for parallel efficiency given a spatial domain decomposition.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Basic Approach}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For simplicity, we describe the uncollided flux technique using a single-speed description. Generalization to multigroup is straightforward.
We recall the one-group transport equation below:
\begin{equation}
\label{eq:transport_eq}
\vec{\Omega} \cdot \grad \Psi(\vec{r},\vec{\Omega}) + \sigma_t(\vec{r}) \Psi(\vec{r},\vec{\Omega}) = 
\sum_\ell \frac{2\ell+1}{4\pi}\sigma_{s,\ell}(\vec{r}) \sum_{m=-\ell}^{\ell} \Phi_{\ell,m}(\vec{r})Y_{\ell,m}(\vec{\Omega}) 
+ q(\vec{r},\vec{\Omega}) \,,
\end{equation}
where $\Psi(\vec{r},\vec{\Omega})$ is the angular flux at position $\vec{r}$ and in direction $\vec{\Omega}$, $Y_{\ell,m}(\vec{\Omega}) $ is the spherical harmonic function of degree $\ell$ and order $m$, $\Phi_{\ell,m}$ is the flux moment of degree $\ell$ and order $m$
\[
 \Phi_{\ell,m}(\vec{r}) = \int_{4\pi} d\Omega  \Psi(\vec{r},\vec{\Omega}) Y_{\ell,m}(\vec{\Omega}) \,.
\]
It helps to introduce an operator notation for brevity:
\begin{equation}
\label{eq:transport_eq_operator}
L\Psi = H\Psi +q \,.
\end{equation}
where $L$ is the streaming and total interaction operator and $H$ the scattering operator. 
Typically, the transport equation is solved iteratively (index $k$)
\begin{equation}
L\Psi^{(k+1)} = H\Psi^{(k)} +q \,.
\end{equation}

Let us now introduce a decomposition of the angular flux into collided and uncollided components:
\[
\Psi = \Psi^u + \Psi^c
\]
Then, \eqt{eq:transport_eq_operator} can be re-cast as
\begin{subequations}
\begin{equation}
\label{eq:_transport_eq_uncoll}
L^{RT}\Psi^u = q 
\end{equation}
\begin{equation}
\label{eq:_transport_eq_coll}
L^{S_N}\Psi^c = H\Psi^c + H\Psi^u 
\end{equation}
\end{subequations}
where we have emphasized how the transport operator $L$ will be solved in each case. The subscript $RT$ denotes ray-tracing
while  $S_N$ stands for discrete-ordinate techniques.

\subsubsection{Uncollided Flux Equation}
%---------------------------------------
The uncollided flux equation (\eqt{eq:_transport_eq_uncoll})
\[
L^{RT}\Psi^u = q \quad \text{or} \quad 
\vec{\Omega} \cdot \grad \Psi^u(\vec{r},\vec{\Omega}) + \sigma_t(\vec{r}) \Psi^u(\vec{r},\vec{\Omega}) = 
 q(\vec{r},\vec{\Omega}) \,,
\]
can be re-written, along a given direction $\vec{\Omega}$, as
\begin{equation}
\label{eq:transport_eq_attenuation}
\frac{d\Psi^u(\vec{r},\vec{\Omega})}{ds} + \sigma_t(\vec{r}) \Psi^u(\vec{r},\vec{\Omega}) = 
 q(\vec{r},\vec{\Omega}) \,,
\end{equation}
where $\vec{r} = \vec{r}_0+s\vec{\Omega}$ and $\vec{r}_0$ is an origin point (i.e., a source point). 
\eqt{eq:transport_eq_attenuation} can be solved analytically for simple geometries or semi-analytically 
using ray-tracing, for more complicated geometries.

\subsubsection{Collided Flux Equation}
%---------------------------------------

Once \eqt{eq:transport_eq_attenuation} has been solved, the uncollided flux solution $\Psi^u$ 
is thus available throughout the domain and the first collision source $q^{1st}$ can be computed as follows:
\[
q^{1st}(\vec{r},\vec{\Omega}) = H \Psi^u = \sum_\ell \frac{2\ell+1}{4\pi}\sigma_{s,\ell}(\vec{r}) \sum_{m=-\ell}^{\ell} \Phi^u_{\ell,m}(\vec{r})Y_{\ell,m}(\vec{\Omega}) 
\]
where $\Phi^u$ denote the flux moments computed from the uncollided angular flux $\Psi^u$. 
Then, one simply needs to solve \eqt{eq:_transport_eq_coll}, the equation for the collided component $\Psi^c$:
\begin{equation}
\label{eq:transport_eq_collided}
L^{S_N}\Psi^c = H\Psi^c + H\Psi^u = L^{S_N}\Psi^c = H\Psi^c + q^{1st} \,.
\end{equation}
Note that \eqt{eq:transport_eq_collided} is similar in nature to \eqt{eq:transport_eq}. An iterative technique (Source Iteration) is employed:
\[
L^{S_N}\Psi^{c,(k+1)} = H\Psi^{c,(k)} + q^{1st} \,.
\]
\eqt{eq:transport_eq_collided} gives the collided component of the angular flux. Standard discrete-ordinates methods are used to solve \eqt{eq:transport_eq_collided}.

\subsubsection{From First-collision Source Treatment to Last-collision Source Treatment}
%---------------------------------------

The above uncollided flux treatment addresses ray effect issues in the uncollided flux solution. However, ray effects can also occur in the collision flux solution, for instance, when there is a large distance from a small scatter source to a detector (e.g., neutrons streaming in a duct). TREAT's fuel motion monitoring systems clearly falls into this category. In a manner analogous to the uncollided flux treatment, a last collided flux treatment can be employed to remove ray effects in a detector response. The last collision source treatment is simply expressed as the original transport equation, where the first-collided source due to the uncollided angular flux and the collided source (obtained from a discrete ordinate solution) 
are known (they come from the ray-tracing and the discrete-ordinate solution, previously described):
\begin{equation}
\label{eq_transport_last}
L^{RT}\Psi^{last} = H\Psi^c + H\Psi^u + q = q^{last}\,,
\end{equation}
That is, instead of using $\Psi^c + \Psi^u$ as the final angular flux at the detector locations, we perform one
more ray-tracing using the total source $q^{last}$ and then employ $\Psi^{last}$ as the final answer.
%given a first-collided and a multiple collided source (and the external source), one ray-traces the angular flux solution, from any point in the domain to the detector surface.
Note, we can also solve for the last collision angular flux $\Psi^{c,last}$ as follows
\begin{equation}
\label{eq_transport_last_coll}
L^{RT}\Psi^{c,last} = H\Psi^c + H\Psi^u  = q^{c,last}\,.
\end{equation}
\eqt{eq_transport_last_coll} is simply obtained by subtracting \eqt{eq_transport_last} and \eqt{eq:_transport_eq_uncoll}. Then
\[
\Psi^{last} = \Psi^{c,last}  + \Psi^u \,.
\]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parallel Aspects of the Traditional Uncollided Flux Treatment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Most uncollided-flux algorithms track directly form each source point to a set of a few points in each cell 
of the domain. The source volume assigned to each source point is effectively treated as vanishingly small, so that 
the angular flux at any other spatial point is treated as a $\delta$-function in the direction $\vec{\Omega}$ subtended 
by the two points. 
This traditional algorithm does not scale well given a spatial domain decomposition: the process (or processor) that owns
a source point must execute $\mathcal{O}(N)$ work given $N$ spatial cells. If the process that owns the  source point
has knowledge of the entire domain, its works grows proportionally to $N$. When sources are localized (as this is the case very the experimental vehicle in TREAT or the detectors in the hodoscope n the case of the last-collision source treatment), 
then many processes not owning any source points stay idle. A simple workaround would be for processes not owning source points to ray-trace once a ray enters their subdomains (this would be mandatory is the mesh is truly parallel and not reproduced on any one processor). However, there are two deficiencies with this approach:
\begin{enumerate}
\item The algorithm is partially sequential. Processes not owning a source point stay idle until a ray has reached them.
\item As one moves further away from the source point, the ray density decreases, which can cause inaccuracies.
\end{enumerate}
\tcr{if time, I add a figure to illustrate this}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SQUID: A MOOSE-based Ray-Tracing Application}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Overview}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tcr{a brief overview of Squid. What it is used for today (MOC, others?). many some scalability results.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Enhancements Required for the Uncollided Flux Treatment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tcr{check}
Current, SQUID ray-traces through a domain (mesh) and computes the angular flux attenuation using a given direction $\vec{\Omega}$, a starting point $\vec{r}_0$, and a termination distance $D:
\[
\Psi^u(\vec{r}_0+s\vec{\Omega},\vec{\Omega}) = 
\Psi^u(\vec{r}_0,\vec{\Omega}) 
\]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Finite-element Evaluation of the First-collided Scattering Source}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\clearpage
%\pagebreak
%\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\bibliographystyle{abbrvnat}
%\bibliography{references_IQS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2016-09-03 doc/August2016_Report/August2016_Report.tex
August IQS blurb for Mark

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[11pt]{scrartcl} 
%########################### Preferences #################################

\usepackage{vmargin}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage[numbers,sort&compress]{natbib}[2010/09/13]

% ********* Caption Layout ************
%\usepackage{ccaption} % allows special formating of the captions
%\captionnamefont{\bf\footnotesize\sffamily} % defines the font of the caption name (e.g. Figure: or Table:)
%\captiontitlefont{\footnotesize\sffamily} % defines the font of the caption text (same as above, but not bold)
%\setlength{\abovecaptionskip}{0mm} %lowers the distance of captions to the figure
\setlength\parindent{0pt}
\setlength{\oddsidemargin}{1in}
\setlength{\evensidemargin}{1in}
\setlength{\textwidth}{6.5in}
% ********* Header and Footer **********
% This is something to play with forever. I use here the advanced settings of the KOMA script

\title{\vspace{-30mm}\Large Improved Quasi-Static Methods in Rattlesnake}
\subtitle{\large August 2016 Status Report}
\author{ \normalsize \textbf{Zachary M. Prince, Jean C. Ragusa, Yaqi Wang, Mark D. DeHart}}

%\author{ \normalsize
%  \textbf{Zachary M. Prince$^\dagger$, Jean C. Ragusa$^\dagger$, Yaqi Wang$^\star$, Mark D. DeHart$^\star$} \\
% \normalsize $^\dagger$Department of Nuclear Engineering \\
% \normalsize Texas A\&M University, College Station, TX, USA\\
% \normalsize $^\star$Idaho National Laboratory\\
% \normalsize \href{mailto:zachmprince@tamu.edu}{zachmprince@tamu.edu}, \href{jean.ragusa@tamu.edu}{jean.ragusa@tamu.edu}, \href{yaqi.wang@inl.gov}{yaqi.wang@inl.gov}, \href{mark.dehart@inl.gov}{mark.dehart@inl.gov}
%}

%\hypersetup{
%  colorlinks=true,
%  urlcolor=blue,
%}

%\pagestyle{fancy}
%\fancyhf{}
%\lhead{INL/EXT-16-38059}
%\rfoot{FINAL}
%\cfoot{\thepage}

%################ End Preferences, Begin Document #####################

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{\large Development for August 2016}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In August 2016, the major development of IQS in Rattlesnake was to improve its capability for multi-physics simulations.  Benchmark testing shows IQS has impressive performance for purely neutronics simulations, including the TWIGL and C5G7 transient benchmarks.  However, TREAT simulation requires accurate treatment of temperature feedback in the fuel, improvements to this aspect of IQS's capability is essential. The LRA benchmark is an adequate substitute to a full core TREAT model for testing purposes; the benchmark includes similar temperature feedback is much quicker, computationally, to run.  Previous reports have IQS's preliminary performance with the LRA benchmark, which show a marginal improvement over a full diffusion simulation.  Therefore, the primary goal for August was to improve this performance.
\\

Several techniques were employed in an effort to improve IQS's performance with temperature feedback, testing with the LRA benchmark. The first technique was to employ an analytical time integration for the temperature evaluation.  For this benchmark, Rattlesnake uses its auxiliary system to evaluate temperature using auxkernels.  Since IQS has much more information about power for a given time step, due to multitude of PRKE evaluations per diffusion solve, a new auxkernel was created more accurately describe the temperature change over a given time step.  This analytical integration is very similar to the one applied to the delayed neutron precursor evaluation.  The second technique employed was to compute temperature at a smaller time scale than the shape/flux.  In theory, this means that the PRKE parameters, which are highly dependent on temperature, are updated more frequently.  Better approximations to the PRKE parameters leads to a more accurate amplitude evaluation, thus a more accurate temperature and shape/flux profile.  Since there is an obvious coupling between PRKE and temperature evaluations, the final technique was to apply pseudo-Picard iterations between temperature and amplitude.  Compared to a diffusion evaluation, these iterations take very little computation time.  This technique improves both the amplitude and temperature solution and reduced the required iterations of the shape/flux, significantly reducing computation time.  These three techniques were all applied to the IQS executioner and tested with the LRA benchmark.  The results showed that the application improved the performance of IQS by about 800\%.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{\large Future Development for September 2016}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The next step for IQS development in Rattlesnake is to expand its capability to SAAF-SN.  SAAF-SN is Rattlesnake's most robust transport solver. After investigation of the SAAF-SN application, its implementation in IQS is more complex than previously predicted. Due to the time dependence of the test function, a new derivation of IQS had to be drafted.  After this derivation, it is apparent that a new suite of post-processors and user-object need to be created.
\\

Another interest is to integrate IQS into the PJFNK iteration process of MOOSE. This involves creating a master user-object that includes the evaluation of the PRKE and its parameters.  This implementation is more in line with MOOSE architecture and make more developer sense. However, the \texttt{save$\_$in} capability can no longer be utilized; so each PRKE parameter will need to be evaluated explicitly instead of using specific residual contributions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2016-09-04 doc/August2016_Report/August2016_Report.tex
small changes. do you concur?

@@
 Benchmark testing shows IQS has impressive performance for purely neutronics simulations, including the TWIGL and C5G7 transient benchmarks.
-However, TREAT simulation requires accurate treatment of temperature feedback in the fuel, improvements to this aspect of IQS's capability is essential.
-The LRA benchmark is an adequate substitute to a full core TREAT model for testing purposes; the benchmark includes similar temperature feedback is much quicker, computationally, to run.
-Previous reports have IQS's preliminary performance with the LRA benchmark, which show a marginal improvement over a full diffusion simulation.
-Therefore, the primary goal for August was to improve this performance.
-\\
+However, TREAT simulations require an accurate treatment of temperature feedback in the fuel, so testing IQS in a multiphysics setting was necessary.
+The LRA benchmark (adiabatic temperature feedback) is an adequate substitute to a full core TREAT model for testing purposes; the benchmark includes similar temperature feedback is much quicker, computationally, to run.
+Previous reports have IQS's preliminary performance with the LRA benchmark and these show only a marginal improvement over a full diffusion simulation.
+Therefore, the primary goal for August was to improve the performance of IQS for multiphysics simulations.
 
 Several techniques were employed in an effort to improve IQS's performance with temperature feedback, testing with the LRA benchmark.
-The first technique was to employ an analytical time integration for the temperature evaluation.
+The first technique was to employ an analytical time integration for the temperature evaluation over the IQS macro time-step.
 For this benchmark, Rattlesnake uses its auxiliary system to evaluate temperature using auxkernels.
-Since IQS has much more information about power for a given time step, due to multitude of PRKE evaluations per diffusion solve, a new auxkernel was created more accurately describe the temperature change over a given time step.
+Since IQS contains power amplitude at every micro time-step, one can resolve the adiabatic temperature model on that time-scale, using the fine-scale power amplitude and the macro-scale flux shape.
+A new auxkernel was created to more accurately describe the temperature change over a given time step.
 This analytical integration is very similar to the one applied to the delayed neutron precursor evaluation.
@@
 Better approximations to the PRKE parameters leads to a more accurate amplitude evaluation, thus a more accurate temperature and shape/flux profile.
-Since there is an obvious coupling between PRKE and temperature evaluations, the final technique was to apply pseudo-Picard iterations between temperature and amplitude.
+Since there is an obvious coupling between PRKE and temperature evaluations, the final technique was to apply Picard iterations between temperature and amplitude.
 Compared to a diffusion evaluation, these iterations take very little computation time.

### 2016-09-23 doc/Thesis_proposal/thesis_proposal.tex
Initial Draft of Thesis Proposal

--- 2016-06-15 doc/NEAMS_Report/NEAMS_Report.tex
@@
 \usepackage{amsbsy}
-%\usepackage{mathbbol}
 \usepackage{graphicx} 
-\usepackage{verbatim}
-\usepackage{csvsimple} 
 \usepackage{subcaption}
@@
 \usepackage{fancyhdr}
-\usepackage{multirow}
-\usepackage{tikz}
 \usepackage{listings}
 \usepackage[numbers,sort&compress]{natbib}[2010/09/13]
+\usepackage{titlesec}
+\usepackage{tikz}
+\usepackage{enumitem}
 
@@
 \newcommand{\adj}{\Phi^\dagger_0}
-\newcommand{\tcr}[1]{\textcolor{red}{#1}}
-\newcommand{\tcb}[1]{\textcolor{blue}{#1}}
-
-% tikz stuff
-\usetikzlibrary{shapes,arrows,positioning}
-\pgfdeclarelayer{background}
-\pgfdeclarelayer{foreground}
-\pgfsetlayers{background,main,foreground}
-\tikzstyle{greenblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=green!20, ultra thick, minimum width=60mm, minimum height=15mm,]
-\tikzstyle{blueblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=blue!20, ultra thick, minimum width=60mm, minimum height=15mm]
-\tikzstyle{reddiamond}=[diamond, draw, align=center, top color=white, bottom color=red!20, ultra thick, minimum width=60mm, aspect=2]
-\newcommand{\tikzback}[5]{
- \begin{pgfonlayer}{background}
-  \path (#1.west |- #2.north)+(-0.5,0.75) node (a1) {};
-  \path (#3.east |- #4.south)+(+0.5,-0.25) node (a2) {};
-  \path[fill=yellow!10,rounded corners, draw=black!100, dashed] (a1) rectangle (a2);
-   \path (#3.east |- #2.north)+(0,1.25)--(#1.west |- #2.north) node[midway] (#5-n) {};
-   \path (#3.east |- #2.south)+(0,-0.35)--(#1.west |- #2.south) node[midway] (#5-s) {};
-   \path (#1.west |- #2.north)+(-0.75,0)--(#1.west |- #4.south) node[midway] (#5-w) {};
- \end{pgfonlayer}}
-
-% input stuff
-\lstset{%
-%  language=C++,
-  showstringspaces=false,
-% basicstyle=\scriptsize\ttfamily,
-  basicstyle=\small\ttfamily,
-  commentstyle=\color{inl@green},
-%  keywordstyle=\bfseries,
-%  escapeinside=$$
-%  framesep=0pt,
-%  rulesep=0pt,
-  frame=single
-}
-% Make nice big tildes inside lstlistings instead of
-% tiny ones.
-\lstset{
-    literate={~} {$\sim$}{1}
-}
 
@@
 % This is something to play with forever. I use here the advanced settings of the KOMA script
+\pagestyle{fancy}
+\fancyhf{}
+\lhead{Prince 2016 - Thesis Proposal}
+\rfoot{}
+\cfoot{\thepage}
 
-\title{Improved Quasi-Static Methods in \rattlesnake}
-\subtitle{Development and Implementation of Improved Quasi-Static (IQS) methods for time-dependent neutron diffusion and neutron transport solvers in \rattlesnake}
+% ********* Title **********
+\title{\normalfont \LARGE Improved Quasi-Static Methods for Time-Dependent Neutron Diffusion and Implementation in Rattlesnake}
+\subtitle{\normalfont Thesis Proposal}
 \author{ \normalsize
-  \textbf{Zachary M. Prince$^\dagger$, Jean C. Ragusa$^\dagger$, Yaqi Wang$^\star$} \\
- \normalsize $^\dagger$Department of Nuclear Engineering \\
- \normalsize Texas A\&M University, College Station, TX, USA\\
- \normalsize $^\star$Idaho National Laboratory\\
- \normalsize \href{mailto:zachmprince@tamu.edu}{zachmprince@tamu.edu}, \href{jean.ragusa@tamu.edu}{jean.ragusa@tamu.edu}, \href{yaqi.wang@inl.gov}{yaqi.wang@inl.gov} 
+  \textbf{Zachary M. Prince} \\ \vspace{-3mm}
+ \normalsize Department of Nuclear Engineering \\ \vspace{-3mm}
+ \normalsize Texas A\&M University\\ \vspace{-3mm}
+ \normalsize College Station, TX, 77843-3133\\
+ \normalsize \href{mailto:zachmprince@tamu.edu}{zachmprince@tamu.edu}
 }
@@
   colorlinks=true,
+  citecolor=black,
+  linkcolor=black,
   urlcolor=blue,
@@
 
-\pagestyle{fancy}
-\fancyhf{}
-\lhead{INL/EXT-16-38059}
-\rfoot{FINAL}
-\cfoot{\thepage}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% boxes
-\usepackage{color}
-\definecolor{myblue}{rgb}{.8, .8, 1}
-\usepackage{empheq}
-
-\newlength\mytemplen
-\newsavebox\mytempbox
-
-\makeatletter
-\newcommand\mybluebox{%
-    \@ifnextchar[%]
-       {\@mybluebox}%
-       {\@mybluebox[0pt]}}
-
-\def\@mybluebox[#1]{%
-    \@ifnextchar[%]
-       {\@@mybluebox[#1]}%
-       {\@@mybluebox[#1][0pt]}}
-
-\def\@@mybluebox[#1][#2]#3{
-    \sbox\mytempbox{#3}%
-    \mytemplen\ht\mytempbox
-    \advance\mytemplen #1\relax
-    \ht\mytempbox\mytemplen
-    \mytemplen\dp\mytempbox
-    \advance\mytemplen #2\relax
-    \dp\mytempbox\mytemplen
-    \colorbox{myblue}{\hspace{1em}\usebox{\mytempbox}\hspace{1em}}}
-
-\makeatother
+\titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}
+\titleformat{\subsection}{\normalfont\large\bfseries}{\thesubsection}{1em}{}
 
 %################ End Preferences, Begin Document #####################
-
-%\pagestyle{plain} % on headers or footers on the first page
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{Abstract}
+\section{Introduction}
+\label{sect:introduction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-This report presents  the implementation of the Improved Quasi-Static method (IQS) in the \rattlesnake application of the MOOSE framework.
-In the IQS method, the neutron flux is factored into a time-dependent amplitude and a spatial- and weakly time-dependent shape.
-The shape evaluation is very similar to a standard flux solve and is computed at large (macro) time steps.
-The amplitude is obtained by solving the PRKEs on a micro time scale.
-The PRKE parameters are regularly updated based on the shape changes.
-Two versions of the IQS methodology have been implemented: (i) the standard approach, which leads to a system of nonlinear equations between the shape and the amplitude, and (ii) a predictor corrector approach that linearizes the system of equations.
-A time-adaptation technique has been developed to automatically select an appropriate macro time step size based on user-specified tolerances.
-The implementation has been verified and tested using manufactured solutions, a custom one-dimensional example, the TWIGL ramp benchmark, and the LRA benchmark.
-The numerical results include error convergence plots and performance with time adaptation.
-These examples prove it to be a viable and effective method for  transient neutronic simulations.
-More complex simulations (e.g., TREAT transients) are planned to perform coupled neutronics/heat conduction transient simulations within the MAMMOTH application.
+Transient modeling of nuclear reactors has been a chronically foreboding task due to its computationally expensive nature.
+However, recent developments in nuclear testing, including the revitalization of the Transient Reactor Testing (TREAT) Facility at Idaho National laboratory (INL), have brought significant attention to the development of transient reactor modeling.
+Transient computational methods currently in laboratory frameworks have proven to be overbearingly slow for complex, real-world scenarios.
+Therefore, more efficient and faster methods for computational frameworks are highly desired.
+The Department of Energy (DOE) has especially sought to develop in the MOOSE framework.
+\\
 
-\pagebreak
+MOOSE is a multiphysics framework being developed at INL that presents the architecture for physics-based applications \cite{moose}.
+Rattlesnake is a large application in MOOSE that involves deterministic radiation transport physics.
+Rattlesnake is able to solve steady-state, transient, and k-eigenvalue neutron transport and diffusion problems using finite element method (FEM) \cite{wang2013}.
+\\
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Theory}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The difficulty in transient reactor modeling is due to the neutron transport equation and the complicated nature of neutron motion that the equation represents.
+The equation has seven independent variables: space ($\vec{r}$), energy ($E$), direction ($\vec{\Omega}$), and time ($t$) \cite{duderstadt1976nuclear}.
+Nuclear reactor simulations often utilize the neutron diffusion approximation, which carefully eliminates the dependence on direction.
+For the purposes of this research, neutron diffusion will be discussed exclusively.
+Additionally, simulations can often neglect temporal dependence due to the steady-state nature of reactors.
+However, simulations for facilities such as TREAT, must retain time dependence in order to obtain relevant information.
+\\
 
-\par
-The improved quasi-static (IQS) method is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (the shape) and a time-dependent component (the amplitude) \cite{Ott_1966,Dulla2008}.
-The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates.
-The IQS method has primarily been developed in the context of the neutron diffusion approximation \cite{Ban_2012,Ferguson_1973}, but reference \cite{Goluoglu_2001} provides an extension of the technique to transport, implemented in TDTORT.
-The method has mostly being applied to neutron kinetics, without thermal-hydraulic feedback.
-In this report, we extend the IQS method in the context of transient multiphysics simulations.
+For computational purposes, the neutron diffusion equation is evaluated with discretized space, energy, and time.
+There are many discretization schemes for each of these variables, but FEM for space and multigroup method for energy were used for the development of this research.
+Since the purpose of this research involves temporal dependence, discretization in time was kept general.
+The time dependent neutron diffusion equation with delayed neutron precursors can be seen in Equations~\ref{eq:flux}~and~\ref{eq:precursor}.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{IQS for Multigroup Diffusion Equations}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In this Section, we recall the equations for the IQS method, using the standard multigroup diffusion equations written below:
 \begin{subequations}
 \begin{align}
-\frac{\partial}{\partial t}\left(\frac{\phi^g}{v^g}\right)  =& \frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \phi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g  \nonumber \\
+\frac{1}{v^g}\frac{\partial \phi^g}{\partial t}  =& \frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \phi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g  \nonumber \\
 &  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \phi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
@@
 \end{subequations}
-%
-Extension to the transport equations will be provided subsequently, once a concise operator notation has been introduced. 
-In the above equations,  \\
+
+Where,
 
@@
 
-Note that the prompt fission terms have been normalized by $\keff$, the eigenvalue of the initial state, as transient simulations are often carried out for initially critical reactors.
-The expressions for subcritical and source-driven problems are straightforward to derive but omitted in this report for brevity.
+Most reactor computation frameworks, including Rattlesnake, discretize the time variable directly with these equations using a multitude of schemes (Implicit Euler, Crank-Nicholson, Runge-Kutta, etc.).
+In this paper, this method of discretizing Equations~\ref{eq:flux}~and~\ref{eq:precursor} is considered brute-force method.
+This research intends to improve upon this method by instead implementing the improved quasi-static method (IQS).
 \\
 
-The factorization approach in IQS leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent multigroup shape ($\varphi^g$):
-\be
-\phi^g(\vec{r},t)=p(t)\varphi^g(\vec{r},t)
-\ee
-We stress that the factorization is not an approximation. Reporting it in the flux and precursor equations yields, after some simple algebra, the following equations for the shape and the precursors:
+IQS is a spatial kinetics method that involves factorizing the flux solution into space- and time-dependent components \cite{Ott_1966,Dulla2008}.
+These components are the flux's amplitude and its shape.
+Amplitude is only time-dependent, while the shape is both space- and time-dependent.
+However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the shape may not require an update at the same frequency as the amplitude function, but only on larger macro-time steps.
+The proposal for this research contains three sections: \sct{sect:back} explains the current theory of IQS and its implementation, \sct{sect:research} describes the proposed research in IQS, and \sct{sect:results} shows preliminary results of the proposed research.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{IQS Theory and Status}
+\label{sect:back}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+This section describes the theory of IQS and the current status of its development.
+The theory and status are broken up into three section: \sct{sect:theory} analytically derives the method, \sct{sect:iter} and \sct{sect:pc} present two solution techniques and describes their application and development.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Theory}
+\label{sect:theory}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+IQS involves factorizing the flux from \eqt{eq:flux} into time-dependent amplitude ($p$) and space- and time- dependent shape ($\varphi$).
+The resulting equation is the shape-diffusion equation with precursors represented by Equations~\ref{eq:shape}~and~\ref{eq:prec}.
+
 \begin{subequations}
 \begin{align}
-\frac{\partial}{\partial t}\left(\frac{\varphi^g}{v^g}\right) = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\
+\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\
 & -  \left( -\div D^g \grad  + \Sigma_r^g + \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right) \varphi^g + \boxed{\frac{1}{p}}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
-\label{eq:shape2}
+\label{eq:shape}
 \end{align}
@@
 \frac{dC_i}{dt} = \frac{\beta_i}{\keff}\boxed{p} \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
-\label{eq:preq2}
+\label{eq:prec}
 \ee
@@
 
-To derive the amplitude equation (also known as the PRKE equations), the shape/precursors equations are weighed by a space-dependent function (typically the initial adjoint flux, $\phi^{*g}$) and integrated over the phase-space.
+To derive the amplitude equation, the shape/precursors equations are weighted by a space-dependent function and integrated over the phase-space.
+The weight function is typically the adjoint flux $\phi^{*g}$, which can be proven to minimize truncation error \cite{duderstadt1976nuclear}.
 The final expressions are given below:
@@
 \end{subequations}
-where we have defined the reactivity, effective delayed-neutron fraction, and delayed-neutron precursor decay constant as follows:
+This equation is also known as the point reactor kinetics equation (PRKE); where the reactivity, effective delayed-neutron fraction, and delayed-neutron precursor decay constant are defined as follows:
 \begin{subequations}
@@
 The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.\\
-Additionally, in order to impose uniqueness on the factorization and to derive the PRKE, the following normalization condition is imposed: $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}$.
-This condition is later used  as the convergence criteria when solving the nonlinear system of equations resulting from the IQS method.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Operator Notation and Extension of IQS to Multigroup Transport Equations}
-\label{sect:transport}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Additionally, in order to impose uniqueness on the factorization and to derive the PRKE, the following normalization condition is imposed: $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}$ \citep{Dulla2008}.
+\\
 
-For simplicity, the previous section only described the IQS implementation using the neutron diffusion equation.
-However, deriving the IQS form for other neutron balance equations (e.g., transport, simplified transport, etc.) is very similar.
-To this end, we re-write the neutron conservation equations in operator form, shown in \eqt{eq:transport}.
-\be
-\frac{\partial}{\partial t}\left(\frac{\Psi^g}{v^g}\right) = \sum_{g'} \left(H^{g'\to g} + P_p^{g' \to g} \right) \Psi^{g'} - L^g\Psi^g + S_d^g
-\label{eq:transport}
-\ee
-where $\Psi^g$ is the multigroup neutron flux (angular flux in the case of transport), $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt neutron production operator, $L^g$ is the loss operator, and $S_d^g$ is the delayed neutron source operator.
-Using \eqt{eq:flux}, the reader may easily obtain the functional form for these operators in the case of a diffusion approximation.
-%
-%Factorization is defined as $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ and the resulting shape equation is defined by \eqt{eq:tshape}.  
-Next, the factorization $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ is introduced, where the shape is denoted by $\psi^g$, leading to the following shape equations, \eqt{eq:tshape}:
-\be
-\frac{\partial}{\partial t}\left(\frac{\psi^g}{v^g}\right) = \sum_{g'} \left(H^{g' \to g} + P_p^{g' \to g}\right) \psi^{g'} - \left( L^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \psi^g + \frac{1}{p} S_d^g
-\label{eq:tshape}
-\ee
-Finally, the PRKE parameters are defined by Equations~\ref{eq:trmb}~and~\ref{eq:tb}, where 
-$\left(\Psi^{*g},f^g\right) =$\\ $ \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d^2\Omega$.
-\be
-\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},\sum_{g'}(H^{g' \to g}g+P_p^{g' \to g}-L^{g'}\delta_{g'g})\psi^{g'}\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
-\label{eq:trmb}
-\ee
-\be
-\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{\sum_{g=1}^G(\Psi^{*g}, \sum_{g'} P_{d,i}^{g' \to g} \psi^{g'})}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
-\label{eq:tb}
-\ee
-where $P_{d,i}^{g' \to g}$ is the delayed-neutron operator for precursor group $i$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Solving the IQS System of Equations}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Solving for the shape can become expensive, especially in two or three dimensions, and even more so when using the transport equations in lieu of the diffusion equations.
+Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three dimensions, and even more so when using the transport equations in lieu of the diffusion equations.
 Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself,  thus allowing for larger time step sizes in updating the shape.
@@
 In transients where the shape varies much less than the flux, IQS can thus be very computationally effective.
-The two time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.
+The two time scale solution process, a micro scale for the PRKE and a macro scale for the shape, are illustrated in \fig{fig:iqsviz}.
 
@@
 
-By this design, the benefits of IQS is most realized when using time adaptive techniques, one of these is described in  \sct{sect:dt2}.
-As we noted earlier, shape-PRKE equations are a nonlinear system and thus need to be solved in a iterative manner.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Iterative Solution Techniques}
+\label{sect:iter}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+As we noted in \sct{sect:theory}, shape-PRKE equations are a nonlinear system and thus may be solved in a iterative manner.
 Each macro time step can be iterated so the best shape is used to compute power at the micro time steps.
-This iteration process must converge the shape such that the uniqueness condition $(\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=0)$ is preserved.
+Sissaoui et al.
+from \cite{Sissaoui_1995}, Koclas et al.
+from \cite{Koclas_1996}, Devooght et al.
+from \cite{Devooght_1984}, and Monier from \cite{Monier_diss} all use iterative techniques for their quasi-static simulations.
+They all undergo a similar process:
+\begin{itemize}
+\item[\textit{Step 1:}] Compute the PRKE parameters at the end of the macro step using the last computed shape
+\item[\textit{Step 2:}] Linearly interpolate the computed PRKE parameters over the macro step
+\item[\textit{Step 3:}] Solve the PRKE on micro steps over the entire macro step
+\item[\textit{Step 4:}] Solve the shape equation on the macro step using the computed values of $p$ and $dp/dt$.
+\item[\textit{Step 5:}] Check if the shape solution has converged:
+	\begin{itemize}
+	\item \textit{No:} Repeat the same macro time step
+	\item \textit{Yes:} Move on to the next macro time step
+	\end{itemize}
+\end{itemize}
 
-The most convenient method to implement IQS in \rattlesnake is through Picard or fixed-point iterations.
-Over a given Picard iteration, the shape is evaluated first with the latest available amplitude.
-The updated shape is employed to compute the PRKE parameters (using a linear interpolation in time) and then the PRKEs are solved using micro time steps.
-This process is repeated until convergence of the normalization condition.
-Picard iterations can also be utilized for multiphysics simulations, where neutronics (e.g., IQS) is nonlinearly coupled to thermal-hydraulics, for instance.
-Thus, inlcuding the IQS nonlinear system solve into the Picard iteration of MOOSE is a straightforward process and is described in \fig{fig:picard}.
+The major difference between the methods of these authors is the convergence criteria used.
+Sissaoui and Koclas \cite{Sissaoui_1995, Koclas_1996} use fixed point iteration where the criteria is the simply the normalized difference between the last two computed shapes.
+Monier in \cite{Monier_diss} also does fixed point iteration with the same criteria, except the solution is scaled by $\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n})\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n+1})\right)}$ after each iteration.
+Devooght in \cite{Devooght_1984} does a Newton-SOR iteration where the residual of the shape function evaluation is the convergence criteria and next iteration's solution is computed using Newton-Raphson method.\\
 
-\begin{figure}[!htpb]
-\centering
-\begin{tikzpicture}[every node/.style = {font=\normalsize}]
-
-\node[greenblock](p1) at (0,0) {Solve shape equations \\ over macro time step};
-\node[greenblock](p2) at (0,-2.5) {Compute PRKE parameters \\ over macro time step};
-\node[greenblock](p3) at (0,-5) {Solve PRKE using micro \\ steps};
-\node[greenblock](p4) at (0,-7.5) {Update $p$ and $\frac{dp}{dt}$};
-\node[blueblock] (p5) at (0,-11) {Solve other physics \\ components};
-\node[reddiamond] (check) at (0,-15) {Check for \\ convergence};
-\node [above =0.1mm of p1] {\large{IQS Solve:}};
-\node [above =0.1mm of p5] {\large{Multi-physics:}};
-
-\tikzback{p1}{p1}{p4}{p4}{bk1}
-\tikzback{p5}{p5}{p5}{p5}{bk2}
-
-\draw[->,ultra thick](p1.south) -- (p2.north);
-\draw[->,ultra thick](p2.south) -- (p3.north);
-\draw[->,ultra thick](p3.south) -- (p4.north);
-\draw [->,ultra thick] (p4.south)+(0,-0.25) -- node [above] {} (bk2-n);
-\draw[->,ultra thick](p5.south)+(0,-0.25) -- (check.north);
-%\draw[->,ultra thick](check.west) -- (-5,-15) -- node[above,sloped] {\large{no}}(-5,0) -- node[above] {}(bk1-w);
-\draw[->,ultra thick](check.west) -- node[above,sloped] {\large{no}} (-5,-15) |-  node[above] {}(bk1-w);
-\draw[->,ultra thick](check.south) -- node[right] {\large{yes}} ++(0,-2);
-
-\end{tikzpicture}
-\caption{Visualization of Picard iteration process}
-\label{fig:picard}
-\end{figure}
+These techniques are by no means an exhaustive list of the possible iteration techniques for IQS.
+Dulla et al.
+in \cite{Dulla2008} does an in depth analysis of the fixed point iteration technique most similar to Sissaoui and Koclas, involving convergence rates and solution results.
+However, no comprehensive analysis of iteration techniques exists.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{A Predictor-Corrector version of IQS (IQS P-C)}
+\subsection{Predictor-Corrector version of IQS (IQS P-C)}
+\label{sect:pc}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The advantage to the predictor-corrector method is there is no iteration necessary for this method and, in turn, is much simpler and faster than the standard IQS.
-%The disadvantage is that the method assumes, by construction, that $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)$ is inherently constant.
+Ikeda et al.
+in \cite{Ikeda_2001} and Goluoglu et al.
+in \cite{Goluoglu_2001} both use IQS P-C for complex, three-dimensional problems.
+Their results prove IQS P-C to be impressively effective, despite the de-coupling of the system.
+Dulla et al.
+in \cite{Dulla2008} also describes an in depth comparison of IQS P-C with traditional IQS.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Precursor Integration}
-\label{sect:prec}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Proposed Research}
+\label{sect:research}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-This section presents the time-integration method used to solve coupled IQS shape and precursor equations, represented by Equations~\ref{eq:shape2}~and~\ref{eq:preq2}.
-First, we note that we could keep this system of two time-dependent equations and solve it as a coupled system.
-However, this is unnecessary and a memory expensive endeavor because the precursor equation is only an ODE and not a PDE.
-Instead, one may discretize in time the shape equation, which typically requires the knowledge of the precursor concentrations at the end of the time step.
-This precursor value is taken from the solution, numerical or analytical, of the precursors ODE.
-\\
+The proposed research intends to continue the investigation and development of the improved quasi-static method.
 
-Through prototyping, it has been found that typical implicit nor Crank-Nicholson time discretizations of precursors are not preferable methods for solving the shape equation in IQS.
-It has been found that these discretizations result in a lack of convergence of the shape over the IQS iteration process.
-In order to remedy the error, a analytical representation of the precursors was implemented in the prototype and the shape solution was able to converge.
-The method entails linearly interpolating the fission source ($S_f = \nu \Sigma_f \varphi$) and integrating the resulting continuous function analytically.
-\eqt{eq:c_dis} is the equation to solve for the precursors and Equations~\ref{eq:a1}~and~\ref{eq:a2} are integrating factors.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Investigation of Iteration Convergence}
+\label{sect:inv_iter}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\sct{sect:iter} describes several iterative techniques employed for IQS, with different convergence criteria and solution scaling.
+A new convergence criteria, not previously attempted, can be applied.
+This criteria involves checking that the uniqueness criteria described in \sct{sect:theory} is preserved.
+The error of a certain iteration is described by \eqt{eq:eiqs}.
+Where $\varphi^{g,n}$ corresponds to the solution at the most recent time step and $\varphi^{g,0}$ corresponds to the initial solution.
+This technique is intended to ensure that the solution retains the purity of the analytical derivation.
 
 \be
-C^{n+1} = C^n e^{-\lambda \Delta t} 
-+ \beta_i \left(a_2 S_f^{n+1} + a_1 S_f^{n}\right)
-\label{eq:c_dis}
+Error_{IQS}=\left|\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,n}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,0}\right)}-1\right|
+\label{eq:eiqs}
 \ee
 
-\be
-a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)p(t')e^{-\lambda(t_{n+1}-t')}dt'
-\label{eq:a1}
-\ee
-\be
-a_2 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)p(t')e^{-\lambda(t_{n+1}-t')}dt'
-\label{eq:a2}
-\ee
-
-The amplitude $(p)$ is included in the integration coefficient because it has been highly accurately calculated in the micro step scheme, so a piecewise interpolation between those points can be done to maximize accuracy.
+This research intends to do investigate the performance of the previously implemented iterative techniques and compare them to this newly implemented technique.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Time Adaptation}
-\label{sect:dt2}
+\subsection{Investigation of Various Time Discretization Schemes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A vital part of the verification and validation for IQS is analyzing error convergence.
+Since any time discretization scheme is capable of being applied to the shape equation of IQS, it is important to investigate IQS's performance to a variety of these schemes.
+There is lack of literature that applies IQS to schemes other than implicit Euler; higher order schemes are never rigorously tested.
+This research intends to apply a variety of schemes, including implicit Euler, Crank-Nicholson, backward difference formula (BDF), and diagonally implicit Runge-Kutta (DIRK), to test stability and error convergence.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Application of Time Adaptation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Thus, by construction, the IQS approach may employ larger time-step sizes for comparable temporal error.
-Further enhancements can be gained by using time adaptation (or time step control) in order to increase or reduce the macro time step size for the
-shape evaluation, depending on error estimates. Here, we employ a step-doubling technique to carry out the time adaptation.
+Further enhancements can be gained by using time adaptation (or time step control) in order to increase or reduce the macro time step size for the shape evaluation, depending on error estimates.
+A step-doubling technique is chosen as the time adaptation technique.
 The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps ($\varphi^g_{\Delta t/2}$).
-Note:
-$\varphi$ is changed to $\phi$ for regular flux evaluation (brute-force time discretization of the governing equations) and IQS P-C.
+Note: $\varphi$ is changed to $\phi$ for brute-force and IQS P-C.
+\\
 
@@
 \be
-\Delta t_{new} = S \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}} \,,
+\Delta t_{new} = S \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}}
 \label{eq:dt2}
 \ee
-where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor.\\
-If the error is larger than the user-specified tolerance, the time step is rejected. A new time step size is estimated using  \eqt{eq:edt2} as well.
+Where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor.
+If the error is larger than the user-specified tolerance, the time step is rejected.
+A new time step size is estimated using  \eqt{eq:edt2} as well.
+\\
 
-To sum up, each solution undergoes Picard iterations until convergence before the error is calculated; for IQS P-C, each solution is re-scaled by its corresponding amplitude $p$.
-Additionally, if the nonlinear iteration does not converge, the entire step doubling process is repeated with half the time original time step.
+To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to brute force method, traditional IQS, and IQS P-C.
+Each of these methods will be applied to several diffusion problems; the number of time steps taken and the resulting error will be used to compare the methods.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Application of Temperature Feedback}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+IQS is first and foremost a nuclear reactor simulation method.
+In nuclear reactors, multiple physics affect the profile of the neutron flux.
+The most simple example of mulitphysics reactor simulations is adiabatic heat up with Doppler feedback.
+The principle of Doppler feedback is that fission in a fuel causes the material to increase temperature and induces a change in the neutronics properties.
+The material heat up is described by \eqt{eq:temp}; where $\rho$ is the material density, $c_p$ is the specific heat, $T$ is temperature, and $\kappa_f$ is the energy released per fission \cite{ANL_BPB}.
+The change in temperature of the material mainly affects the thermal macroscopic absorption cross section described by \eqt{eq:dopp} \cite{ANL_BPB}.
+
+\be
+\rho c_p \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
+\label{eq:temp}
+\ee
+
+\be
+\Sigma_a^{thermal}(\vec{r},t) = \Sigma_a^{thermal}(\vec{r},0)\left[1+\gamma\left(\sqrt{T}-\sqrt{T_0}\right)\right]
+\label{eq:dopp}
+\ee
+
+For IQS, this temperature feedback affects both the shape equation and the reactivity of the PRKE; thus, it is an additional nonlinear component to the already coupled shape-amplitude equations.
+In foresight to the application of this component, temperature is much more time dependent than the shape, but less so than the amplitude.
+Therefore, the evaluation of temperature will have its own time scale.
+A possible solution process for a problem with temperature feedback will have time three time scales portrayed in \fig{fig:time}.
+The first time scale is the shape solve, the second is the temperature evaluation as well as the computation of PRKE parameters, and the third is the PRKE scale.
+
+\begin{figure}[htpb!]
+\centering
+\begin{tikzpicture}[scale=2]
+%Shape
+\draw[] (0,2) -- (6,2) ;
+\foreach \x in  {0,6}
+\draw[shift={(\x,2)},color=black] (0pt,6pt) -- (0pt,-6pt);
+\draw[shift={(0,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
+\draw[shift={(6,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
+\node(shape) at (-.25,2) {$\varphi:$};
+
+%Temp/Params
+\draw[] (0,1) -- (6,1) ;
+\foreach \x in  {0,6}
+\draw[shift={(\x,1)},color=black] (0pt,6pt) -- (0pt,-6pt);
+\foreach \x in  {2,4}
+\draw[shift={(\x,1)},color=black] (0pt,4pt) -- (0pt,-4pt);
+\draw[shift={(0,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
+\draw[shift={(6,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
+\node(temp) at (-.75,1) {$T \rightarrow \rho, \bar{\beta}, \bar{\lambda}, \Lambda$:};
+
+% PRKE
+\draw[] (0,0) -- (6,0) ;
+\foreach \x in  {0,6}
+\draw[shift={(\x,0)},color=black] (0pt,6pt) -- (0pt,-6pt);
+\foreach \x in  {0,2,4,6}
+\draw[shift={(\x,0)},color=black] (0pt,4pt) -- (0pt,-4pt);
+\foreach \x in  {0,0.2,0.4,0.6,0.8,1,1.2,1.4,1.6,1.8,2,2.2,2.4,2.6,2.8,3,3.2,3.4,3.6,3.8,4,4.2,4.4,4.6,4.8,5,5.2,5.4,5.6,5.8,6}
+\draw[shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt);
+\draw[shift={(0,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
+\draw[shift={(6,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
+\node(prke) at (-.5,0) {$p, \frac{dp}{dt}:$};
+
+\draw (0,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 2} (2,0);
+\draw (2,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 4} (4,0);
+\draw (4,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 6} (6,0);
+\draw (0,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 1} (2,1);
+\draw (2,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 3} (4,1);
+\draw (4,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 5} (6,1);
+\draw (0,2) edge[out=15,in=165,->,thick] node[above,sloped] {Step 7} (6,2);
+
+\draw (0,2) edge[out=-135,in=135,->,thick] node[below,sloped] {$\varphi$} (0,1);
+\draw (2,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (2,0);
+\draw (2,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (2,1);
+\draw (4,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (4,0);
+\draw (4,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (4,1);
+\draw (6,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (6,0);
+\draw (6,0) edge[out=45,in=0,->,thick] node[below,sloped] {$p, \frac{dp}{dt}$} (6,2);
+
+\end{tikzpicture}
+\caption{Time scales and process of IQS with temperature feedback}
+\label{fig:time}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Implementation into Rattlesnake/MOOSE Framework}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A one-dimensional, single group FEM prototype will be created for IQS in order to test validation for simple problems.
+However, testing complex, real-world problems is most easily performed by implementing IQS in a FEM framework.
+The Rattlesnake radiation transport framework is a excellent option for performing these problems.
+Rattlesnake already contains a neutron transport method for CFEM diffusion equivalent to the brute-force method.
+Additionally, Rattlesnake is well documented and well organized, so implementing IQS is relatively straightforward.
+With this implementation, benchmark testing such as 16-A1 \cite{ANL_BPB}, TWIGL, LRA, C5G7-2D, etc.
+can be performed, as well as a full core simulation of TREAT.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{\rattlesnake Implementation}
+\newpage
+\section{Preliminary Results}
+\label{sect:results}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\rattlesnake is a MOOSE-based application developed INL specific to solving radiation transport problems with multiphysics capabilities.
-MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like \rattlesnake.
-At the heart of \rattlesnake is the action system, which provides a means to consolidate the MOOSE input syntax (which can be quite larger for multigroup transport simulations), so that a user does not have to define every kernel, variable, etc., used in the problem.
-Rather, the user inputs an equation description
-(e.g., Diffusion, S$_n$, P$_n$, etc.) and a solution method (e.g., SAAF, LS, CFEM, DFEM, etc.), and the action system will incorporate all the necessary physics involved (kernels, boundary conditions, postprocessor, etc.).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Executioner}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-An IQS excecutioner was created to implement the IQS method.
-The IQS executioner derives from the Transient executioner in MOOSE.
-The IQS executioner contains a loop over micro time steps that computes the PRKE solution and then passes $p$ and $\frac{dp}{dt}$ for the Transient executioner to evaluate the shape equation at each macro step.
-The PRKE solve is performed with a user specified option of backward-Euler, Crank-Nicholson, or SDIRK33.
-The IQS executioner also supplements Transient Picard iteration process by adding its own error criteria:
-\be
-Error_{IQS}=\left|\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,n}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,0}\right)}-1\right|
-\label{eq:eiqs}
-\ee
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Action System}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The IQS implementation mostly require the specific IQS executioner, described above.
-However, IQS additional changes are in the \rattlesnake action system in order to support the IQS execution.
-First, changes needed to be made in order to evaluate the shape equation.
-The shape equation, after some manipulation, is very similar to the time-dependent governing laws that \rattlesnake already solves.
-Using multigroup diffusion as an example again, we show the various kernels employed and highlight the new or modified kernels.
-
-\begin{align}
-\frac{\partial}{\partial t}\left(\frac{\varphi^g}{v^g}\right)=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
-& - \underbrace{\frac{1}{v^g} \boxed{\overbrace{\frac{1}{p}\frac{dp}{dt}}^{From Executioner}}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
-\end{align}
-
-To enable \rattlesnake to solve this equation, an IQS removal kernel was created to evaluate $\sum_{g=1}^G\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
-Also, the precursor kernel was modified to include the $\frac{1}{p}$ term.
-Finally, the precursor auxkernel that evaluates \eqt{eq:preq2} using the analytical integration method described in \sct{sect:prec}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{PRKE coefficients}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In order to evaluate the PRKE coefficients, defined by Equations~\ref{eq:rmb}~-~\ref{eq:l}, four postprocessors were created.
-The parameter calculations were separated by $\frac{\bar{\beta}_i}{\Lambda}$ numerator, $\bar{\lambda}_i$ numerator/denominator, $\frac{\rho-\bar{\beta}}{\Lambda}/\frac{\bar{\beta}}{\Lambda}$ denominator, and $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator.
-The first three are relatively simple, only relying on material properties and solution quantities, then computing the elemental integral.
-The $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator requires the use of the MOOSE \texttt{save\_in} feature.
-This feature saves the residual from a calculated kernel or boundary contribution in the shape evaluation to an auxiliary variable.
-The postprocessor then computes the inner product of this variable and the initial adjoint solution.
-After each of these postprocessors are evaluated, a user object pulls together all the values and performs the numerator/denominator divisions.
-The resulting values are then passed to the executioner for the PRKE solve.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Other Action Systems}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-For simplicity, IQS implementation has only been described for CFEM diffusion.
-However, \rattlesnake has other action systems capable of transient simulation, where IQS can be implemented and be effective.
-One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equations~\ref{eq:flux}~and~\ref{eq:shape2}.
-However, in the derivation for IQS, this term is unaffected between shape and flux evaluation.
-So saving the residual for this diffusion kernel in the \texttt{save\_in} variable is the only alteration to this action for IQS to function.\\
-
-Additional action systems involve transport, but it is evident from \sct{sect:transport} that IQS implementation in these is straightforward as well.
-The main differences between a diffusion implementation and a transport implementation are outlined below:
-\begin{enumerate}
-\item The form of the operators in the shape equations is different, but \rattlesnake has already implemented all the kernels necessary to represent these operators.
-So no change is necessary to these kernels is necessary for IQS.
-Additionally, the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ is the same, so both action systems can use the same kernel.
-\item The PRKE parameters also change because of the operators.
-For the $(\rho-\bar{\beta})$ parameter, the same post-processor can be used with the \texttt{save\_in} functionality in MOOSE.
-The post-processor for $\bar{\beta}_i$ must be re-written for transport, but takes a very similar form as diffusion.
-\end{enumerate}
-
-IQS will be made available to any action system capable of transient simulation, which includes:
-\begin{itemize}
-\item CFEM-Diffusion
-\item DFEM-Diffusion
-\item SAAF-CFEM-SN
-\item SAAF-CFEM-PN
-\item LS-CFEM-SN
-\item DFEM-SN
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Predictor-Corrector Modification}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In order to preserve the already implemented standard version of IQS, an option in the IQS executioner was created to specify which method is desired.
-Because the diffusion solve is flux instead of shape, when predictor-corrector option is specified, the IQS removal kernel ($\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$) and the modified precursor kernel are bypassed, while all the postprocessors are still executed.
-However, it is difficult to rescale the flux to shape before the PRKE parameter postprocessor are executed.
-So the parameters are computed using the full flux, but amplitude is space independent and comes out of the integrals.
-As seen in parameter definitions, when shape is replaced with flux, the amplitude comes out of the integral and cancels out.
-So the conversion of the predicted flux in  \eqt{eq:rescale} to shape is unnecessary if the corrected flux is solved with \eqt{eq:phicorr2}.
-After obtaining the corrected flux, the precursors are re-evaluated using a \texttt{EXEC\_LINEAR} statement.
-\be
-\underbrace{\phi^g_{n+1}}_{\text{corrected}} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}} p_{n+1}
-\label{eq:phicorr2}
-\ee
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Input}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The input file for IQS is very similar to the current transient diffusion input file.
-The IQS input has a different executioner type and parameters.
-The executioner type is simply IQS and input parameters include the number of micro time steps per macro step, the IQS error tolerance, and the initial power.
-The \rattlesnake transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $\keff$, which is present in the transient input deck.
-However, IQS also requires an initial evaluation of the adjoint flux, for the weighting function.
-So another input file and multi-app transfer was made for the adjoint calculation.
-\\
-
-Below is the syntax for the executioner block for an IQS input file.
-The \texttt{predictor\_corrector} logical determines whether to do IQS P-C or regular IQS.
-The \texttt{IQS\_error\_tol} is the tolerance for the IQS error represented by \eqt{eq:eiqs} and will at most \texttt{picard\_max\_its} iterations until convergence.
-The \texttt{prke\_scheme} defines the time discretization used for the PRKE solution.
-Where \texttt{RK} is SDIRK33, \texttt{CN} is Crank-Nicholson, and \texttt{IE} is implicit-Euler.
-\begin{lstlisting}
-[Executioner]
-  type = IQS
-  predictor_corrector = true/false
-  picard_max_its = 5
-  ...
-  n_micro = 10000
-  IQS_error_tol = 1e-7
-  prke_scheme = 'RK'
-[]
-\end{lstlisting}
-
-Since IQS needs to use the adjoint solution for the PRKE parameter evaluation, auxiliary variables need to be created for each group in the input file.  Below is an example of their definition: 
-\begin{lstlisting}
-[AuxVariables]
-  [./adjoint_flux_g0] 
-    family = LAGRANGE
-    order = FIRST
-  [../]
-  [./adjoint_flux_g1]
-    family = LAGRANGE
-    order = FIRST
-  [../]
-  ...
-[]
-\end{lstlisting}
-
-Below is the syntax for the multi-app block to perform forward and adjoint steady-state evaluations:
-\begin{lstlisting}
-[MultiApps]
- [./initial_solve]
-   type = FullSolveMultiApp
-   execute_on = initial
-   input_files = initial.i
- [../]
- [./adjoint_solve]
-   type = FullSolveMultiApp
-   execute_on = initial
-   input_files = adjoint.i
- [../]
-[]
-\end{lstlisting}
-
-Below is the syntax for the Transfer block to copy the initial and adjoint solutions from the multi-apps.
-\begin{lstlisting}
-[Transfers]
- [./copy_solution]
-   type = TransportSystemVariableTransfer
-   direction = from_multiapp
-   multi_app = initial_solve
-   execute_on = initial
-   from_transport_system = diff
-   to_transport_system = diff
- [../]
- [./copy_adjoint]
-   type = MultiAppVariableTransfer
-   execute_on = initial
-   direction = from_multiapp
-   multi_app = adjoint_solve
-   from_variables = 'sflux_g0 sflux_g1 ...'
-   to_variables = 'adjoint_flux_g0 adjoint_flux_g1 ...'
- [../]
- [./copy_eigenvalue]
-   type = EigenvalueTransfer
-   execute_on = initial
-   direction = from_multiapp
-   multi_app = initial_solve
- [../]
-[]
-\end{lstlisting}
-
-\newpage
-%------------------------------------------------------------------------------
-%
-%------------------------------------------------------------------------------
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{RESULTS} 
-\label{sect::results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This section describes results of examples that test the IQS implementation and shows its effectiveness on computation speed and accuracy.
-Three examples were selected for this purpose.
-The first is a homogeneous one-group problem, subjected to a heterogeneous material change (absorption cross-section change as a ramp in time for a subset of the geometry).
-The second is the two-dimensional TWIGL ramp transient benchmark, described further.
-The final is the LRA benchmark, which a two-dimensional, temperature feedback example.
-Each example was solved with regular diffusion (brute force), IQS, and IQS P-C.
-Each method was tested with backward-Euler and BDF2 for the first two examples; just Crank-Nicholson was used for third example.
-The performance for each method are represented by convergence plots with constant $\Delta t$ and by the number of time steps with step doubling time adaptation.
-For time adaptation, backward-Euler time discretization was used.
-% and, for a safety factor, $e_{tol} = 0.8e_{max}$. I found tol and max to be confusing to the reader. i have simplified a bit
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{One-Dimensional Custom Example}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+This section describes preliminary results of an example that tests an IQS implementation and shows its effectiveness on computation speed and accuracy.
 The example is very simple and computes quickly, it entails a one dimensional, homogeneous 400 cm slab with a heterogenous perturbation in absorption cross section.
@@
 The values of $\Sigma_a$ between these times of interest are linear interpolations between the given values.
+\\
 
@@
 The strong correlation between the two curves shows that IQS is consistent with a proven method for a highly transient example.
-\fig{fig:conv} shows that IQS is not only consistent for this example, but also has a better error constant in the convergence study.
 \fig{fig:conv} shows the error convergence for brute force and IQS methods.
@@
 The IQS performance in this example are not exceptional because it is not an easy problem.
-As seen in \fig{eq:D_shape}, the shape undergoes significant change through the transient, which requires many shape evaluations to retain accuracy.
+As seen in \fig{fig:1D_shape}, the shape undergoes significant change through the transient, which requires many shape evaluations to retain accuracy.
 
@@
 
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TWIGL Benchmark}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This benchmark problem originates from the Argonne National Lab Benchmark Problem Book \cite{ANL_BPB}.
-It is a 2D, 2-group reactor core model with no reflector region shown in \fig{fig:TWIGL_reg}.
-\tbl{tab:TWIGL_mat} shows the material properties of each fuel region and the ramp perturbation of Material 1.
-
-\begin{figure}[!htbp]
-\begin{center}
-\includegraphics[width=0.75\textwidth]{\FiguresDir/TWIGL_regions.jpg}
-\caption{TWIGL benchmark problem description}
-\label{fig:TWIGL_reg}
-\end{center}
-\end{figure}
-
-\begin{table}[!htbp]
-\begin{center}
-\caption{TWIGL benchmark material properties and slope perturbation}
-\label{tab:TWIGL_mat}
-\begin{tabular}{llllllll}
-\hline
-  &  &  &  &  &  &  \multicolumn{2}{c}{$\underline{\Sigma_s (cm^{-1})} $} \\
-Material & Group & $D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu\Sigma_f (cm^{-1})$ & $\chi$ & $g \rightarrow 1$ & $g \rightarrow 2$ \\
-\hline
-1 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01 \\
-  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
-2 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01  \\
-  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
-3 & 1 & 1.3 & 0.008 & 0.003 & 1.0 & 0.0 & 0.01  \\
-  & 2 & 0.5 & 0.050 & 0.060 & 0.0 & 0.0 & 0.00  \\
-\hline
-  & $\nu$ & $v_1 (cm/s)$ & $v_2 (cm/s)$ & $\beta$ & $\lambda (1/s)$ &   &   \\
-\hline
-  & 2.43 & 1.0E7 & 2.0E5 & 0.0075 & 0.08 &   &   \\
-\hline
- \multicolumn{8}{l}{\footnotesize Material 1 ramp perturbation:} \\
-\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (1-0.11667t) \quad t \leq 0.2 s$} \\
-\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (0.97666t) \quad t > 0.2 s$} \normalsize
-\end{tabular}
-\end{center}
-\end{table}
-
-\fig{fig:TWIGL_power} shows the IQS  solution as compared with the Brute Force solution.
-These plots show that IQS is consistent in more complex, higher dimensional problems in \rattlesnake.
-Finally, \fig{fig:TWIGL_conv} plots the error convergence of IQS and the Brute Force methods.
-The curves show the impressive convergence of IQS for the highly transient TWIGL example.
-\tbl{tab:TWIGLdt2} and \fig{fig:TWIGL_power_dt2} show the results for time adaptation.
-The results show that both IQS methods perform exceptionally well compared to brute force.
-It also shows that traditional IQS performed better with large $e_{tol}$, while IQS P-C was better with smaller $e_{tol}$.
-
-%It is important to note the IQS shape plot is scaled differently than the Brute Force flux plot because the amplitude term is not included, but the gradients of colors is comparable.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.49\linewidth]{\FiguresDir/TWIGL_power_plot.png}
-\includegraphics[width=0.5\linewidth]{\FiguresDir/TWIGL_power_plot2.png}
-\caption{Power level comparison of TWIGL Benchmark}
-\label{fig:TWIGL_power}
-\end{figure}
-
-%\begin{figure}[!htbp]
-%\begin{center}
-%\begin{subfigure}[!htbp]{0.4\textwidth}
-%\includegraphics[width=\textwidth]{\FiguresDir/ndiff_ramp_flux.jpg}
-%\caption{Brute force flux}
-%\end{subfigure}
-%\quad
-%\begin{subfigure}[!htbp]{0.4\textwidth}
-%\includegraphics[width=\textwidth]{\FiguresDir/iqs_ramp_shape.jpg}
-%\caption{IQS Shape}
-%\end{subfigure}
-%\caption{TWIGL Benchmark flux/shape comparison at $t=0.2$}
-%\label{fig:TWIGL_plots}
-%\end{center}
-%\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=5in]{\FiguresDir/TWIGL_convergence.jpg}
-\caption{Error convergence comparison of TWIGL Benchmark}
-\label{fig:TWIGL_conv}
-\end{figure}
-
-\begin{table}[!htbp]
-\begin{center}
-\caption{TWIGL step doubling results}
-\label{tab:TWIGLdt2}
-\resizebox{\textwidth}{!}{
-\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
-\hline
-  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
-\hline
-Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
-\hline
-1 &	0.05    &	0.00012677 &	9   &	29  &	0.03380433 &	4   &	20   &	0.00323100 &	4  &	9   \\
-2 &	0.01    &	3.5555e-05 &	11  &	35  &	0.00166991 &	5   &	40   &	0.00263068 &	5  &	12  \\
-3 &	0.005   &	4.0364e-05 &	11  &	31  &	0.00886584 &	5   &	40   &	0.00160486 &	6  &	21  \\
-4 &	0.001   &	0.00294822 &	33  &	122 &	0.02976305 &	5   &	36   &	1.7527e-05 &	10 &	35  \\
-5 &	0.0005  &	0.00099778 &	39  &	131 &	0.00143781 &	6   &	55   &	1.4185e-05 &	16 &	74  \\
-6 &	0.0001  &	0.00019510 &	78  &	236 &	0.00016175 &	7   &	65   &	6.2903e-06 &	19 &	78  \\
-7 &	5.0e-05 &	0.00018372 &	112 &	342 &	6.0328e-05 &	12  &	163  &	1.5247e-06 &	24 &	92  \\
-8 &	1.0e-05 &	8.0564e-05 &	263 &	794 &	7.7103e-05 &	379 &	5729 &	9.8321e-07 &	48 &	210 \\
-\hline
-
-\end{tabular}}
-\end{center}
-\end{table}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.8\linewidth]{\FiguresDir/TWIGL_power_plot_dt2.png}
-\caption{Power level comparison of TWIGL Benchmark with time adaptation}
-\label{fig:TWIGL_power_dt2}
-\end{figure}
-
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{LRA Benchmark}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.
-It is a super prompt-critical transient.
-To have better understanding on the cross sections given later, we present the equations here:
-\begin{subequations}
-\begin{align}
--\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 - \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
--\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
-S_f &= \sum_{g=1}^2 \Sigma_{f,g} \phi_g, \\
-\frac{\partial C_i}{\partial t} &= \nu\beta_i f - \lambda_i C_i, \quad i=1,2, \\
-\frac{\partial T}{\partial t} &= \alpha f, \label{eq:lra-temp} \\
-\Sigma_{a,1} &= \Sigma_{a,1}(\vec{r}, t=0) \left[1+\gamma\left(\sqrt{T} - \sqrt{T_0}\right)\right], \\
-P &= \kappa S_f,
-\end{align}
-\end{subequations}
-where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0=T(\vec{r}, t=0)$.
-The two-group diffusion equation are solved with zero flux boundary conditions on external surfaces, reflecting conditions at symmetry boundaries and steady state initial conditions which are obtained by solving
-\begin{align}
--\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
--\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1. \\
-\end{align}
-The eigenvalue $k$ is used to modify the fission cross section for the transient simulations with $\frac{1}{k}\Sigma_{f,g}, g=1,2$.
-The initial flux distribution shall be normalized such that the averaged power density
-\begin{align}
-\bar{P} \equiv \frac{\int_{V_{core}} P(\vec{r}, t=0) d\vec{r}}{\int_{V_{core}} d\vec{r}},
-\end{align}
-where $V_{core}$ is the core region with fuels, is equal to $10^{-6} W\cdot cm^{-3}$.
-The initial precursor concentrations are in equilibrium with the initial critical flux distribution.\\
-
-The geometry is illustrated in~\fig{fig:lra-geometry}.\\
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.8\linewidth]{\FiguresDir/lra.png}
-\caption{LRA benchmark geometry with region assignment.}
-\label{fig:lra-geometry}
-\end{figure}
-
-Initial two-group constants are presented in~\tbl{tab:lra-xs}.
-$\nu$ is equal to 2.43.
-Axial bulking $B^2 = 10^{-4}$ is applied for both energy groups.
-Delayed neutron data are presented in~\tbl{tab:lra-dnp}.
-All fuel materials have the same delayed neutron data.
-Some scalar data are listed in~\tbl{tab:lra-scalar}.\\
-\begin{table}[!htbp]
-    \centering
-    \caption{LRA benchmark initial two-group constants.\label{tab:lra-xs}}
-    \resizebox{\textwidth}{!}{
-      \begin{tabular}{|c|c|c|c|l|l|l|l|l|}
-      \hline
-             &                 & Group   & $D_g$      & $\Sigma_{a,g}$ & $\nu\Sigma_{f,g}$ & $\Sigma_{s,1\rightarrow 2}$ & $\chi_g$ & $v_g$              \\
-      Region & Material        &       g & (cm)       &  ($cm^{-1}$)   &  ($cm^{-1}$)      &  ($cm^{-1}$)                &          & ($cm\cdot s^{-1}$) \\
-      \hline
-      1      & Fuel 1 with rod & 1       & 1.255      & 0.008252       & 0.004602          &                             & 1        & $3.0\times10^7$    \\
-             &                 & 2       & 0.211      & 0.1003         & 0.1091            & 0.02533                     & 0        & $3.0\times10^5$    \\
-      \hline
-      2      & Fuel 1 without rod & 1    & 1.268      & 0.007181       & 0.004609          &                             & 1        & $3.0\times10^7$    \\
-             &                    & 2    & 0.1902     & 0.07047        & 0.08675           & 0.02767                     & 0        & $3.0\times10^5$    \\
-      \hline
-      3      & Fuel 2 with rod & 1       & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
-             &                 & 2       & 0.2091     & 0.08344        & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
-      \hline
-      4      & Fuel 2 without rod & 1    & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
-             &                    & 2    & 0.2091     & 0.073324       & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
-      \hline
-      5      & Reflector        & 1      & 1.257      & 0.0006034      & -                 &                             & -        & $3.0\times10^7$    \\
-             &                  & 2      & 0.1592     & 0.01911        & -                 & 0.04754                     & -        & $3.0\times10^5$    \\
-      \hline
-      \end{tabular}}
-\end{table}
-
-\begin{table}[!htbp]
-    \centering
-    \caption{LRA benchmark delayed neutron data.\label{tab:lra-dnp}}
-      \begin{tabular}{|c|l|l|l|l|}
-      \hline
-      Group i & $\beta_i$ & $\lambda_i$ ($s^{-1}$) & $\chi_{d,i,1}$ & $\chi_{d,i,2}$ \\
-      \hline
-      1       & 0.0054    & 0.0654  & 1 & 0 \\
-      2       & 0.001087  & 1.35    & 1 & 0 \\
-      \hline
-      \end{tabular}
-\end{table}
-
-\begin{table}[!htbp]
-    \centering
-    \caption{LRA benchmark scalar values.\label{tab:lra-scalar}}
-      \begin{tabular}{|l|c|l|}
-      \hline
-      Meaning & Notation & value \\
-      \hline
-      Axial buckling for both energy groups & $B_g^2$  & $10^{-4}$ ($cm^{-2}$)\\
-      Mean number of neutrons per fission   & $\nu$    & 2.43 \\
-      Conversion factor                     & $\alpha$ & $3.83\times 10^{-11}$ ($K\cdot cm^{3}$) \\
-      Feedback constant                     & $\gamma$ & $3.034\times 10^{-3}$ ($K^{1/2}$) \\
-      Energy released per fission           & $\kappa$ & $3.204\times 10^{-11} $ ($J/fission$) \\
-      Initial and reference temperature     & $T_0$    & 300 (K) \\
-      Active core volume                    & $V_{core}$ & 17550 ($cm^2$)\\
-      \hline
-      \end{tabular}
-\end{table}
-
-The transient is initiated by changing the thermal absorption cross section as the following:
-\begin{align}
-\Sigma_{a,2}(t) = \Sigma_{a,2}(t=0) \left\{\begin{array}{lr} 1-0.0606184t, & t\leq 2 \\
-                                                             0.8787631, & t>2
-                                           \end{array}\right.
-\end{align}
-where $t$ is time in seconds.\\
-
-The \fig{fig:LRA_plots} plots show the IQS  solution as compared with the Brute Force solution.
-These plots show that IQS is consistent with Doppler feedback problems in \rattlesnake.
-\fig{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods.
-\tbl{tab:LRAdt2} shows the time adaptation performance of brute force and IQS P-C.
-These results show that IQS, again, performs much better than the brute force method by taking significantly fewer diffusion solves while retaining similar accuracy.
-
-\begin{figure}[!htbp]
-\begin{center}
-\begin{subfigure}[!htbp]{0.49\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/lra_power_profile.png}
-\caption{Power Profile}
-\end{subfigure}
-\begin{subfigure}[!htbp]{0.49\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/lra_power_profile2.png}
-\caption{Power Profile at Peak}
-\end{subfigure}
-\begin{subfigure}[!htbp]{0.49\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/lra_temp_profile.png}
-\caption{Temperature Profile}
-\end{subfigure}
-\caption{LRA Benchmark power and temperature comparison}
-\label{fig:LRA_plots}
-\end{center}
-\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.6\linewidth]{\FiguresDir/lra_convergence.jpg}
-\caption{Error convergence comparison of LRA Benchmark}
-\label{fig:lra_conv}
-\end{figure}
-
-\begin{table}[!htbp]
-\caption{LRA step doubling results}
-\label{tab:LRAdt2}
-\begin{center}
-\begin{tabular}{|l|l|l|l|l|l|l|l|}
-\hline
-  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS P-C} \\
-\hline
-Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves \\
-\hline
-1 &	 0.01 &	 0.0437566 &	 145 &	 735 &	 0.143855 &	 105 &	 616 \\ 
-2 &	 0.005 &	 0.0309732 &	 266 &	 948 &	 0.104782 &	 157 &	 814 \\ 
-3 &	 0.001 &	 0.158242 &	 585 &	 2046 &	 0.155676 &	 218 &	 1062 \\ 
-4 &	 0.0005 &	 0.270949 &	 833 &	 2908 &	 0.211994 &	 480 &	 2038 \\ 
-5 &	 0.0001 &	 0.280468 &	 1793 &	 6155 &	 0.0437761 &	 434 &	 1458 \\ 
-6 &	 5e-05 &	 0.212274 &	 2485 &	 8511 &	 0.0385312 &	 542 &	 1719 \\ 
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-The purpose of this report is to briefly review the improved quasi-static method for neutronic transient simulations, to describe its implementation in \rattlesnake, the radiation transport application based on MOOSE, and to show its effectiveness using several test cases.
-%The derivation of IQS was done with multigroup diffusion, but it was seen in \sct{sect:transport} that application to transport is relatively simple.  
-Two IQS implementations are available in \rattlesnake.
-The first one is the traditional IQS approach, where the shape equation and PRKE are solved as a coupled nonlinear system with Picard iterations.
-The second approach, IQS P-C, is a predictor-corrector linearization, where the flux is solved and then rescaled using the amplitude obtained through the PRKE
-(Point Reactor Kinetic Equations).  IQS P-C is a simpler method and is computationally much less expensive to perform. \\
-
-IQS's implementation in \rattlesnake involves a new kernel, an auxiliary kernel, a new executioner derived from the current transient, four postprocessors, as well as alterations to the action system.
-The new kernel implements the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term in the shape equation.
-The auxiliary kernel alters how the precursors are solved, the method is outlined by \sct{sect:prec}.
-The executioner is where the PRKE are evaluated after shape solves are performed.
-The postprocessors are used to evaluate the parameters for the PRKE and a user object is used to pass them to the executioner.
-IQS's implementation the action system made sure that all the correct kernels and postprocessors are executed when the IQS executioner is specified in the input.
-\\
-
-The IQS implementations have been verified and comparing against the traditional ``brute force'' temporal discretization of the governing laws.
-Error convergence plots and time adaptation performance are provided.
-Three examples were used including a custom one-dimensional example, the TWIGL benchmark, and LRA benchmark.
-The one-dimensional example shows large shape changes in the flux and proves consistency of the shape evaluation.
-IQS shows proper error convergence for first and second order time discretization.
-IQS P-C shows marginal improvement with time adaptation, but traditional IQS shows poor performance.
-The TWIGL benchmark shows the effectiveness of the PRKE in IQS because there is very little shape change through the transient.
-IQS shows very satisfactory performance with both constant time steps and time adaptation.
-The LRA benchmark tests IQS when utilized in the context of a multiphysics simulation.
-Again, IQS shows good performance by requiring half the time steps as brute force for a given error.
-
+\newpage
+\section{Goals of Proposed Research}
+\label{sect:goals}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\clearpage
-\pagebreak
-\newpage
+\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
+\item Investigate iteration convergence for previously implemented iterative schemes and newly conceived one described in \sct{sect:inv_iter}.
+\item Validate and verify IQS method with several low and high order time discretization schemes.
+\item Implement step doubling time adaptation with IQS and analyze its performance compared to brute-force.
+\item Apply temperature feedback with IQS and analyze its performance with multiphysics benchmarks.
+\item Implement IQS into the Rattlesnake framework at INL.
+\item Perform testing with IQS to benchmarks including TWIGL, LRA, and C5G7-2D.
+\end{enumerate}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Other Possible Outlooks}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
+\item Perform a full core simulation of TREAT with IQS.
+\item Investigate other multiphysics coupling with IQS including structural mechanics and thermal hydraulics.
+\item Investigate IQS application to self-adjoint angular flux S$_\text{N}$ (SAAF-SN).
+\end{enumerate}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
 \bibliographystyle{abbrvnat}
 \bibliography{references_IQS}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Appendix: aaaa}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\tcr{putting stuff here for now}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%It can be shown that the most appropriate weighting function $(w^g)$ is the initial adjoint flux $(\phi^{*g})$.
-For brevity, the following definition will be applied: $\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r=\left(\phi^{*g},f\right)$
-%\begin{align}
-%& \quad \frac{dp}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)+ p\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)= \nonumber \\
-%& p\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)  \\
-%& + \sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i) \nonumber
-%\end{align}
-%\be
-%\frac{d}{dt}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)=\frac{1}{\keff}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})p - \sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \quad 1 \le i \le I
-%\ee
-%
-%In order to impose uniqueness of the factorization, one imposes:
-%\be
-%\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=constant
-%\ee
-%
-%Therefore the PRKE formulation reduces to:
-%\begin{align}
-%\frac{dp}{dt}= &\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g}  \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p \nonumber \\
-%& +\sum_{i=1}^I\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i
-%\end{align}
-%\be
-%\frac{d\xi_i}{dt}=\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p - \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i \quad 1 \le i \le I
-%\ee
-%Where:
-%\be
-%\xi_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g C_i)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-%\ee
-%
-%It is convenient to define the effective reactivity, delay-neutron fraction, and delayed-neutron precursor decay constant:
-%\be
-%\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-%%\label{eq:rmb}
-%\ee
-%\be
-%\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-%%\label{eq:b}
-%\ee
-%\be
-%\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
-%%\label{eq:l}
-%\ee
-%
-%So:
-%\be
-%\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
-%%\label{eq:p}
-%\ee
-%\be
-%\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
-%%\label{eq:c}
-%\ee
-%
-%Equations (\ref{eq:p}) and (\ref{eq:c}) are a formulation of the point reactor kinetics equation (PRKE), but the parameters (Equations (\ref{eq:rmb})-(\ref{eq:l})) are dependent on the shape.
-If the assumption is made that the shape is time-independent, the shape is computed once at the first time step and used for the PRKE parameter evaluation at all other steps.
-However, if the shape is dependent on time, the shape needs to be computed in transient using equation (\ref{eq:shape}) and (\ref{eq:prec}) in order retain accuracy.
-Equations (\ref{eq:shape2}) and (\ref{eq:preq2}) shows the usual form of the shape and precursor equations with amplitude put on the right hand side.
-Equation (\ref{eq:shape2}) is very similar to the multigroup flux equation (\ref{eq:flux}), except the removal cross-section term is augmented by a $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term and the precursor contribution has a $\frac{1}{p}$ multiplier.
-Equation (\ref{eq:preq2}) is very similar to the normal precursor equation (\ref{eq:precursor}), except the fission source term is multiplied by $p$.
-These differences are crucial for IQS implementation in \rattlesnake.
-%
-%\begin{align}
-%\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\
-%& -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
-%%\label{eq:shape2}
-%\end{align}
-%\be
-%\frac{dC_i}{dt} = \frac{\beta_i}{\keff}p \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
-%%\label{eq:preq2}
-%\ee
-
-%\begin{align}
-%& \quad \sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}\right)= \nonumber \\ 
-%&\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g\right) \\
-%&+\frac{1}{p}\sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \nonumber
-%\end{align}
 

### 2016-09-28 doc/Thesis_proposal/thesis_proposal.tex
jcr comments to the ms proposal

@@
 \newcommand{\adj}{\Phi^\dagger_0}
+\newcommand{\tcr}[1]{\textcolor{red}{#1}}
 
@@
 However, recent developments in nuclear testing, including the revitalization of the Transient Reactor Testing (TREAT) Facility at Idaho National laboratory (INL), have brought significant attention to the development of transient reactor modeling.
-Transient computational methods currently in laboratory frameworks have proven to be overbearingly slow for complex, real-world scenarios.
-Therefore, more efficient and faster methods for computational frameworks are highly desired.
+Transient computational methods currently in \tcr{laboratory frameworks--people may not understand what you mean here} have proven to be overbearingly \tcr{slow-CPU-expensive} for complex, real-world scenarios.
+Therefore, more efficient and faster methods for computational \tcr{frameworks} are highly desired.
 The Department of Energy (DOE) has especially sought to develop in the MOOSE framework.
@@
 
-The difficulty in transient reactor modeling is due to the neutron transport equation and the complicated nature of neutron motion that the equation represents.
+The difficulty in transient reactor modeling is due to the neutron transport equation and \tcr{the complicated nature of neutron motion that the equation represents-- ????
+it is the dimensionality of the phase-space that is the issue, and the fact that the time discretization has to be implicit, which  leads to kinear system solves}.
 The equation has seven independent variables: space ($\vec{r}$), energy ($E$), direction ($\vec{\Omega}$), and time ($t$) \cite{duderstadt1976nuclear}.
@@
 For the purposes of this research, neutron diffusion will be discussed exclusively.
-Additionally, simulations can often neglect temporal dependence due to the steady-state nature of reactors.
+\tcr{Additionally, simulations can often neglect temporal dependence due to the steady-state nature of reactors--???}.
 However, simulations for facilities such as TREAT, must retain time dependence in order to obtain relevant information.
@@
 
-For computational purposes, the neutron diffusion equation is evaluated with discretized space, energy, and time.
-There are many discretization schemes for each of these variables, but FEM for space and multigroup method for energy were used for the development of this research.
+For computational purposes, the neutron diffusion equation is discretized in space, energy, and time.
+There are many discretization schemes for each of these variables, but FEM for space and multigroup method for energy were used for the development of this research \tcr{I do not know of ``many'' discretizations in the energy variable ...}.
 Since the purpose of this research involves temporal dependence, discretization in time was kept general.
@@
 
-Where,
+where,
 
@@
 Most reactor computation frameworks, including Rattlesnake, discretize the time variable directly with these equations using a multitude of schemes (Implicit Euler, Crank-Nicholson, Runge-Kutta, etc.).
-In this paper, this method of discretizing Equations~\ref{eq:flux}~and~\ref{eq:precursor} is considered brute-force method.
+In this paper, this method of discretizing Equations~\ref{eq:flux}~and~\ref{eq:precursor} is considered \tcr{brute-force--I wonder if we should come up with a better terminology...} method.
 This research intends to improve upon this method by instead implementing the improved quasi-static method (IQS).
@@
 This section describes the theory of IQS and the current status of its development.
-The theory and status are broken up into three section: \sct{sect:theory} analytically derives the method, \sct{sect:iter} and \sct{sect:pc} present two solution techniques and describes their application and development.
+The theory and status are broken up into three sections: \sct{sect:theory} analytically derives the method, \sct{sect:iter} and \sct{sect:pc} present two solution techniques and describes their application and development.
 
@@
 
-We note that the time-dependent shape equation is similar to the time-dependent flux, with the following modifications:
+We note that the time-dependent shape equation is similar to the time-dependent flux equation, with the following modifications:
 \begin{enumerate}
@@
 The weight function is typically the adjoint flux $\phi^{*g}$, which can be proven to minimize truncation error \cite{duderstadt1976nuclear}.
+\tcr{Check Cristian Rabitti's PhD work.
+There is more to it.
+He used a new adjoint at each time step I believe.
+It is interesting to read}
 The final expressions are given below:
@@
 In transients where the shape varies much less than the flux, IQS can thus be very computationally effective.
-The two time scale solution process, a micro scale for the PRKE and a macro scale for the shape, are illustrated in \fig{fig:iqsviz}.
+The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.
 
@@
 Sissaoui and Koclas \cite{Sissaoui_1995, Koclas_1996} use fixed point iteration where the criteria is the simply the normalized difference between the last two computed shapes.
-Monier in \cite{Monier_diss} also does fixed point iteration with the same criteria, except the solution is scaled by $\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n})\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n+1})\right)}$ after each iteration.
+Monier in \cite{Monier_diss} also does fixed point iterations with the same criteria, except the solution is scaled by $\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n})\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n+1})\right)}$ after each iteration.
 Devooght in \cite{Devooght_1984} does a Newton-SOR iteration where the residual of the shape function evaluation is the convergence criteria and next iteration's solution is computed using Newton-Raphson method.\\
@@
 In the IQS P-C version, the flux equations (not the shape equations) are solved (represented by Equations~\ref{eq:flux}~and~\ref{eq:precursor}) in order to obtain a predicted flux solution.
-This predicted flux is then converted to shape by rescaling as follows:
+This predicted flux is then converted to a shape by rescaling it as follows:
 \be
@@
 
+\tcr{is it interesting to try to converge the nonlinearities in $T$ and $\rho$ before going back to the shape?}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The Rattlesnake radiation transport framework is a excellent option for performing these problems.
-Rattlesnake already contains a neutron transport method for CFEM diffusion equivalent to the brute-force method.
-Additionally, Rattlesnake is well documented and well organized, so implementing IQS is relatively straightforward.
+Rattlesnake already contains a neutron transport method for CFEM diffusion equivalent to the brute-force method \tcr{Are you referring to SAAF?
+or did you mean diffusion in lieu of transport?}.
+Additionally, Rattlesnake is well documented and well organized, so implementing IQS may be relatively straightforward.
 With this implementation, benchmark testing such as 16-A1 \cite{ANL_BPB}, TWIGL, LRA, C5G7-2D, etc.
-can be performed, as well as a full core simulation of TREAT.
+can be performed, as well as a full core simulation of the TREAT reactor.
 
@@
 
+\tcr{for a paper, can we think of testing Embeded RK instead of/in addition to step-doubling?}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Other Possible Outlooks}
+\subsection{Possible Extensions}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Some extensions, beyond the proposed scope of this research, have been identified as follows:
 

### 2016-10-13 doc/Thesis_proposal/thesis_proposal.tex
Thesis: Addressing some comments

@@
 However, recent developments in nuclear testing, including the revitalization of the Transient Reactor Testing (TREAT) Facility at Idaho National laboratory (INL), have brought significant attention to the development of transient reactor modeling.
-Transient computational methods currently in \tcr{laboratory frameworks--people may not understand what you mean here} have proven to be overbearingly \tcr{slow-CPU-expensive} for complex, real-world scenarios.
+Transient computational methods currently in \tcr{laboratory frameworks--people may not understand what you mean here} have proven to be overbearingly computationally expensive for complex, real-world scenarios.
 Therefore, more efficient and faster methods for computational \tcr{frameworks} are highly desired.
@@
 
-The difficulty in transient reactor modeling is due to the neutron transport equation and \tcr{the complicated nature of neutron motion that the equation represents-- ????
-it is the dimensionality of the phase-space that is the issue, and the fact that the time discretization has to be implicit, which  leads to kinear system solves}.
+The difficulty in transient reactor modeling is due to the neutron transport equation, which describes the flux of neutrons.
+The difficulty in solving or evaluating the equation comes from the dimensionality of the phase-space and the fact that the time discretization has to be implicit, which  leads to kinear system solves.
 The equation has seven independent variables: space ($\vec{r}$), energy ($E$), direction ($\vec{\Omega}$), and time ($t$) \cite{duderstadt1976nuclear}.
@@
 For computational purposes, the neutron diffusion equation is discretized in space, energy, and time.
-There are many discretization schemes for each of these variables, but FEM for space and multigroup method for energy were used for the development of this research \tcr{I do not know of ``many'' discretizations in the energy variable ...}.
+There are several viable discretization schemes for each of these variables, but FEM for space and multigroup method for energy were used for the development of this research.
 Since the purpose of this research involves temporal dependence, discretization in time was kept general.
@@
 
-\tcr{is it interesting to try to converge the nonlinearities in $T$ and $\rho$ before going back to the shape?}
+The couplings between temperature, amplitude, and shape are all nonlinear, so iteration processes are need for each time scale.
+The amplitude and temperature need to be iterated on the middle time scale until convergence on each temperature step.
+Then another iterative process needs to occur in the shape time scale scale on all three variables.
 
@@
 The Rattlesnake radiation transport framework is a excellent option for performing these problems.
-Rattlesnake already contains a neutron transport method for CFEM diffusion equivalent to the brute-force method \tcr{Are you referring to SAAF?
-or did you mean diffusion in lieu of transport?}.
+Rattlesnake already contains a neutron diffusion method for continuous-FEM equivalent to the brute-force method.
 Additionally, Rattlesnake is well documented and well organized, so implementing IQS may be relatively straightforward.

### 2016-10-14 doc/M&C_2017/MC2017_Summary.tex
Half a draft

\documentclass{anstrans}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Multiphysics Techniques for Neutron Diffusion Using Improved Quasi-Static Method}
\author{Zachary M.~Prince, Jean C. Ragusa}

\institute{
Department of Nuclear Engineering, Texas A\&M University, College Station, TX
}

\email{zachmprince@tamu.edu \and jean.ragusa@tamu.edu}

%%%% packages and definitions (optional)
\usepackage{color} % allows inclusion of graphics
\usepackage{graphicx} % allows inclusion of graphics
\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{microtype} % improves typography for PDF
\usepackage{subcaption}

\newcommand{\SN}{S$_N$}
\renewcommand{\vec}[1]{\bm{#1}} %vector is bold italic
\newcommand{\vd}{\bm{\cdot}} % slightly bold vector dot
\newcommand{\ud}{\mathop{}\!\mathrm{d}} % upright derivative symbol
%  new definitions
\newcommand{\bs}[1]{\mathbf{#1}}
\renewcommand{\div}{\bs{\nabla}\! \cdot \!}
\newcommand{\grad}{\bs{\nabla}}
% extra space
\newcommand{\qq}{\quad\quad}
% common reference commands
\newcommand{\eqt}[1]{Eq.~(\ref{#1})}                     % equation
\newcommand{\fig}[1]{Fig.~\ref{#1}}                      % figure
\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table
\newcommand{\sct}[1]{Section~\ref{#1}}                   % section
\newcommand{\app}[1]{Appendix~\ref{#1}}                   % appendix

\newcommand{\keff}{k_\textit{eff}}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\vn}{\vec{n}}
\newcommand{\vel}{\vec{\mathrm{v}}}
\newcommand{\adj}{\Phi^\dagger_0}
\newcommand{\tcr}[1]{\textcolor{red}{#1}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert_{L^2}}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
The purpose of this summary is to introduce several techniques in dealing with temperature feedback for quasi-static neutron diffusion calculations.  Temperature is tightly coupled with neutron flux when simulating a nuclear reactor, thus it is a very important quantity to evaluate accurately and efficiently.  The improved quasi-static method (IQS) is an effective technique in simulating flux in reactors and the effect of multiphysics with IQS is worth exploring..

IQS is a transient spatial kinetics method that involves factorizing neutron flux into a space- and time-dependent component (shape) and a time-dependent component (amplitude) \cite{Ott_1966,Dulla2008}. The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates. IQS has mostly been applied to neutron kinetics, without thermal-hydraulic feedback. This summary discusses the application of temperature feedback with IQS and analyzes performance with benchmark problems

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
In this Section, we recall the equation for the IQS method, starting from the standard multi-group transport equations with delayed neutron precursors in operator form:

\be
\frac{1}{v^g}\frac{\partial \phi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \phi^{g'} - L^g\phi^g + S_{d}^g
\label{eq:flux}
\ee 
\be
\frac{dC_i}{dt} = \sum_{g=1}^G P_{d,i}^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:precursor}
\ee

Where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the diffusion and removal operator, $S_{d}^g$ is the delay source, and $P_{d,i}^g$ is the delay fission operator.

Factorization is an important step in the derivation of the IQS method. The factorization approach leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent multigroup shape ($\psi$):
\be
\phi^g(\vec{r},\Omega,t)=p(t)\varphi^g(\vec{r},\Omega,t)
\ee
After implementing the factorization, the shape diffusion equations result:
\be
\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \phi^{g'} - L^g\phi^g + S_{d}^g
\label{eq:flux}
\ee 
\be
\frac{dC_i}{dt} = \sum_{g=1}^G P_{d,i}^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:precursor}
\ee

To obtain the amplitude equations, the multigroup equations are multiplied by a weighting function, typically the initial adjoint flux ($\phi^{*g}$), and then integrated over phase-space.  For brevity, the inner product over space will be represented with parenthetical notation ($\left(\phi^{*g},f^g\right) = \int_D \phi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r\Omega
$). In order to impose uniqueness of the factorization, one requires $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)$ to be constant.  And after some manipulation, the standard point reactor kinetics equations (PRKE) for the amplitude solution are obtained:
\be
\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
\ee
\be
\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}p-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
\ee
Where the functional coefficients are calculated using the space-/time-dependent shape function as follows:
\be
\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\sum_{g'}(H^{g' \to g}+P_p^{g' \to g}-L^{g'}\delta_{g'g})\varphi^{g'}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:rmb}
\ee
\be
\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{\sum_{g=1}^G(\phi^{*g}, P_{d,i}^g \varphi^g)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\ee
\be
\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
\label{eq:l}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IQS Predictor-Corrector (IQS P-C)}

This version of IQS first solves the flux diffusion (represented by Equations \ref{eq:flux} and \ref{eq:precursor}) to get a predicted flux.  The predicted flux at this step is then converted to shape by rescaling as follows:
\be
\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
\label{eq:rescale}
\ee
Where:
\be
K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
\ee
\be
K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
\ee

The PRKE parameters are then computed with this shape using Equations \ref{eq:rmb} - \ref{eq:l} and interpolated over the macro step, then the PRKE is evaluated.  With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Temperature Feedback}

In nuclear reactors, multiple physics affect the profile of the neutron flux.  The most simple example of mulitphysics reactor simulations is adiabatic heat up with Doppler feedback. The principle of Doppler feedback is that fission in a fuel causes the material to increase temperature and induces a change in the neutronics properties.  The material heat up is described by \eqt{eq:temp}; where $\rho$ is the material density, $c_p$ is the specific heat, $T$ is temperature, and $\kappa_f$ is the energy released per fission \cite{ANL_BPB}. The change in temperature of the material mainly affects the thermal macroscopic absorption cross section described by \eqt{eq:dopp} \cite{ANL_BPB}.

\be
\rho c_p \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
\label{eq:temp}
\ee

\be
\Sigma_a^{thermal}(\vec{r},t) = \Sigma_a^{thermal}(\vec{r},0)\left[1+\gamma\left(\sqrt{T}-\sqrt{T_0}\right)\right]
\label{eq:dopp}
\ee

For IQS, this temperature feedback affects both the shape equation and the reactivity of the PRKE; thus, it is an additional nonlinear component to the already coupled shape-amplitude equations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Analysis}
The results were interesting, so interesting in fact that we have decided to
present them here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Subsection Goes Here}
The user must manually capitalize initial letters of a subsection heading.

For those who like equations in their papers, \LaTeX\ is a good choice. Here is
an equation for the Marshak diffusion boundary condition:
\begin{equation} \label{eq:marshak}
  4 J^- = \phi + 2 D \vec{n} \vd \grad \phi \,.
\end{equation}
If we so choose, we can effortlessly reference the equation later.

Another paragraph starts with Eq.~\eqref{eq:marshak} and sets $J^-$ to zero, a
vacuum boundary condition:
\begin{equation*}
  0 = \phi + \frac{2}{3} \frac{1}{\sigma} \vec{n} \vd \grad \phi \,.
\end{equation*}
The extrapolation distance is $2/3$. A more detailed asymptotic analysis yields
an extrapolation distance of about $0.71045$.

Figure~\ref{fig:voltage} shows how a plot might conceivably look in your
document. Always place figures after they are referenced so as not to throw
off the reader. You can use symbols and different line styles to help
differentiate your results, especially if they are printed in black and white.
Note how Fig.~\ref{fig:voltage} uses dashed lines \verb|--| for the exact
solution, solid lines \verb|-| for the new method's solutions, and dotted lines
\verb|:| for existing inaccurate methods.
\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics{example_figure}
  \caption{Captions are flush with the left.}
  \label{fig:voltage}
\end{figure}

Later on, we can include a table, even one that spans two columns such as
Table~\ref{tab:widetable}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[htb]
  \centering
\begin{tabular}{llllllllll}\toprule
      & $\phi_T(0)$      & $\phi_T(10)$      & $\phi_T(20)$      &
      $\phi_D(0)$      & $\phi_D(10)$      & $\phi_D(20)$      & $\rho$      &
      $\varepsilon$      & $N_\text{it}$
\\ \midrule
$c=0.999$  & 0.9038 & 20.63 & 31.24 & 0.9087 & 20.63 & 31.23 & 0.2192 & $10^{-7}$ & 15
\\
$c=0.990$  & 0.3675 & 13.04 & 24.7 & 0.3696 & 13.04 & 24.69 & 0.2184 & $10^{-7}$ & 15
\\
$c=0.900$  & 0.009909 & 4.776 & 17.64 & 0.009984 & 4.786 & 17.63 & 0.2118 & $10^{-7}$ & 14
\\
$c=0.500$  & $6.069\times 10^{-5}$ & 2.212 & 15.53 & 6.213$\times 10^{-5}$ & 2.239 & 15.53 & 0.2068 & $10^{-7}$ & 13
\\
\bottomrule
\end{tabular}
  \caption{This is an example of a really wide table which might not normally
  fit in the document.}
  \label{tab:widetable}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Notice how the table reference uses a Roman numeral
for its numbering scheme, whereas the figure reference uses an Arabic numeral.
For one-column tables, use the \verb|table| environment; two-column tables use
\verb|table*|. The same applies to figures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Another Subsection}
Excessive sectioning in a three-page document is discouraged, but here are more
subsections to demonstrate compliance with the ANS formatting guidelines.

\subsubsection{Third-level Heading}
This subsubsection shows compliance with the ANS-specified standard. This level
of heading should be used rarely.

\subsubsection{Another Such Heading}
And, if you really think you need a third-level heading, you should make sure
that your subsection needs at least two of them.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

The included ANS style file and this clear example file are a panacea for
the hours of headache that invariably results from formatting a document in
Microsoft Word.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Appendix}

Numbering in the appendix is different:
\begin{equation} \label{eq:appendix}
  2 + 2 = 5\,.
\end{equation}
and another equation:
\begin{equation} \label{eq:appendix2}
  a + b = c\,.
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
This material is based upon work supported a Department of Energy Nuclear
Energy University Programs Graduate Fellowship.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{ans}
\bibliography{bibliography}
\end{document}

### 2016-10-15 doc/M&C_2017/MC2017_Summary.tex
A more complete draft

@@
 \usepackage{subcaption}
+\usepackage{tikz}
 
@@
 \section{Introduction}
-The purpose of this summary is to introduce several techniques in dealing with temperature feedback for quasi-static neutron diffusion calculations.
-Temperature is tightly coupled with neutron flux when simulating a nuclear reactor, thus it is a very important quantity to evaluate accurately and efficiently.
-The improved quasi-static method (IQS) is an effective technique in simulating flux in reactors and the effect of multiphysics with IQS is worth exploring..
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The purpose of this summary is to introduce several techniques in dealing with multiphysics feedback for quasi-static neutron diffusion calculations.
+Other physical variables, including temperature, are tightly coupled with neutron flux when simulating a nuclear reactor, thus are very important quantities to evaluate accurately and efficiently.
+The improved quasi-static method (IQS) is an effective technique in simulating flux in reactors and the effect of multiphysics with IQS is worth exploring.
 
@@
 IQS has mostly been applied to neutron kinetics, without thermal-hydraulic feedback.
-This summary discusses the application of temperature feedback with IQS and analyzes performance with benchmark problems
+This summary discusses the application of multiphysics feedback with IQS and analyzes performance with temperature feedback benchmark problems
 
@@
 \section{Theory}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 In this Section, we recall the equation for the IQS method, starting from the standard multi-group transport equations with delayed neutron precursors in operator form:
@@
 \be
-\phi^g(\vec{r},\Omega,t)=p(t)\varphi^g(\vec{r},\Omega,t)
+\phi^g(\vec{r},t)=p(t)\varphi^g(\vec{r},t)
 \ee
@@
 \be
-\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \phi^{g'} - L^g\phi^g + S_{d}^g
+\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \varphi^{g'} - \left(L^g - \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right)\varphi^g + \boxed{\frac{1}{p}}S_{d}^g
 \label{eq:flux}
@@
 \be
-\frac{dC_i}{dt} = \sum_{g=1}^G P_{d,i}^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
+\frac{dC_i}{dt} = \boxed{p}\sum_{g=1}^G P_{d,i}^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
 \label{eq:precursor}
@@
 
-To obtain the amplitude equations, the multigroup equations are multiplied by a weighting function, typically the initial adjoint flux ($\phi^{*g}$), and then integrated over phase-space.
-For brevity, the inner product over space will be represented with parenthetical notation ($\left(\phi^{*g},f^g\right) = \int_D \phi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r\Omega
+We note that the time-dependent shape equation is similar to the time-dependent flux.
+The main differences is that there is a nonlinear coupling with amplitude contained in the addition of the boxed terms.
+
+To obtain the amplitude equations, the multigroup shape equations are multiplied by a weighting function, typically the initial adjoint flux ($\phi^{*g}$), and then integrated over the domain.
+For brevity, the inner product over space will be represented with parenthetical notation ($\left(\phi^{*g},f^g\right) = \int_D \phi^{*g}(\vec{r})f^g(\vec{r})d^3r
 $).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Temperature Feedback}
+\subsection{IQS Solution Process with Multiphysics}
 
-In nuclear reactors, multiple physics affect the profile of the neutron flux.
-The most simple example of mulitphysics reactor simulations is adiabatic heat up with Doppler feedback.
-The principle of Doppler feedback is that fission in a fuel causes the material to increase temperature and induces a change in the neutronics properties.
-The material heat up is described by \eqt{eq:temp}; where $\rho$ is the material density, $c_p$ is the specific heat, $T$ is temperature, and $\kappa_f$ is the energy released per fission \cite{ANL_BPB}.
-The change in temperature of the material mainly affects the thermal macroscopic absorption cross section described by \eqt{eq:dopp} \cite{ANL_BPB}.
+Other physical quantities, such as temperature, are affected by reaction rates and subsequently affect the operators of the flux equations.
+For IQS, this feedback affects both the shape equation and the parameters of the PRKE; thus, it is an additional nonlinear component to the already coupled shape-amplitude equations.
+Each of these components have different temporal dependencies; so it is beneficial for efficiency to evaluate them on different time scales.
+Amplitude is much more time dependent than shape, and shape is much more computationally expensive to evaluate.
+So amplitude is evaluated on a much smaller time scale than shape, while it may be appropriate so solve quantities, such as temperature, on middle time scale.
+Figure~\ref{fig:time} shows a solution process for temperature feedback.
+
+\begin{figure}[htbp!]
+\centering
+\begin{tikzpicture}[scale=1]
+\tikzset{font={\footnotesize}}
+%Shape
+\draw[] (0,2) -- (6,2) ;
+\foreach \x in  {0,6}
+\draw[shift={(\x,2)},color=black] (0pt,6pt) -- (0pt,-6pt);
+\draw[shift={(0,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
+\draw[shift={(6,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
+\node(shape) at (-.25,2) {$\varphi:$};
+
+%Temp/Params
+\draw[] (0,1) -- (6,1) ;
+\foreach \x in  {0,6}
+\draw[shift={(\x,1)},color=black] (0pt,6pt) -- (0pt,-6pt);
+\foreach \x in  {2,4}
+\draw[shift={(\x,1)},color=black] (0pt,4pt) -- (0pt,-4pt);
+\draw[shift={(0,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
+\draw[shift={(6,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
+\node(temp) at (-.75,1) {$T \rightarrow \rho$:};
+
+% PRKE
+\draw[] (0,0) -- (6,0) ;
+\foreach \x in  {0,6}
+\draw[shift={(\x,0)},color=black] (0pt,6pt) -- (0pt,-6pt);
+\foreach \x in  {0,2,4,6}
+\draw[shift={(\x,0)},color=black] (0pt,4pt) -- (0pt,-4pt);
+\foreach \x in  {0,0.2,0.4,0.6,0.8,1,1.2,1.4,1.6,1.8,2,2.2,2.4,2.6,2.8,3,3.2,3.4,3.6,3.8,4,4.2,4.4,4.6,4.8,5,5.2,5.4,5.6,5.8,6}
+\draw[shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt);
+\draw[shift={(0,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
+\draw[shift={(6,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
+\node(prke) at (-.5,0) {$p, \frac{dp}{dt}:$};
+
+\draw (0,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 2} (2,0);
+\draw (2,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 4} (4,0);
+\draw (4,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 6} (6,0);
+\draw (0,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 1} (2,1);
+\draw (2,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 3} (4,1);
+\draw (4,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 5} (6,1);
+\draw (0,2) edge[out=15,in=165,->,thick] node[above,sloped] {Step 7} (6,2);
+
+\draw (0,2) edge[out=-135,in=135,->,thick] node[below,sloped] {$\varphi$} (0,1);
+\draw (2,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (2,0);
+\draw (2,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (2,1);
+\draw (4,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (4,0);
+\draw (4,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (4,1);
+\draw (6,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (6,0);
+\draw (6,0) edge[out=45,in=0,->,thick] node[below,sloped] {$p, \frac{dp}{dt}$} (6,2);
+
+\end{tikzpicture}
+\caption{Time scales and process of IQS}
+\label{fig:time}
+\end{figure}
+
+The top time scale represents a shape diffusion evaluation on a macro step, the middle as an arbitrary three steps within the macro step where temperature and the PRKE parameters are evaluated, and the bottom one represents the PRKE evaluation on micro steps.
+The shape is linearly interpolated within the macro step for the temperature and PRKE parameter evaluation, and the parameters are interpolated within the temperature step for the PRKE evaluation.
+Since there is a nonlinear coupling between all these components, each temperature step is iterated until amplitude has converged and the macro step is iterated until shape has converged.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results and Analysis}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To test the multiple time scale implementation of IQS, the LRA benchmark was chosen for reference.
+The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.
+It is a super prompt-critical transient.
+Three different evaluation techniques were applied: full discretization of the flux equation (Brute force), IQS, and IQS-PC.
+Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.
+The performance of IQS and the temperature updates was measured by its improvement in accuracy at peak power over the Brute force technique.
+The results show the effectiveness of multiple temperature updates along the macro step.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{LRA Temperature Feedback}
+
+The heat up is described by \eqt{eq:temp} and the feedback is described by \eqt{eq:dopp}.
 
@@
 
-For IQS, this temperature feedback affects both the shape equation and the reactivity of the PRKE; thus, it is an additional nonlinear component to the already coupled shape-amplitude equations.
+In the temperature evaluation, a typical implicit solver would simply use the interpolated flux at end of the temperature time step for the right hand side of the equation.
+However, IQS has much more information about the profile of the flux along the time step because of the micro-step amplitude evaluation.
+Therefore, it is possible to solve for temperature using a quasi-analytical approach, shown by \eqt{eq:temp_an}.
+
+\be
+T^{n+1} = T^n + \frac{\kappa_f}{\rho c_p} \left(a_2 \varphi^{n+1} + a_1 \varphi^{n}\right)
+\label{eq:temp_an}
+\ee
+Where $n$ corresponds to the beginning of the temperature step.  $a_1$ and $a_2$ are integration coefficients defined by \eqt{eq:a1} and \eqt{eq:a2}.
+\be
+a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)p(t')dt'
+\label{eq:a1}
+\ee
+\be
+a_2 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)p(t')dt'
+\label{eq:a2}
+\ee
+
+Any interpolation of the amplitude along the micro steps is possible for the integration, this application uses piece-wise linear.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results and Analysis}
-The results were interesting, so interesting in fact that we have decided to
-present them here.
+\subsection{LRA Convergence Results}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Subsection Goes Here}
-The user must manually capitalize initial letters of a subsection heading.
+\fig{fig:lra_profile} shows the baseline power and temperature profile for the LRA benchmark.
+The relative difference in the magnitude of the peak power ($t\approx1.44 s$) from the baseline was used for error comparison.
+\fig{fig:lra_bad} is an error convergence plot comparing the three techniques where temperature is evaluated only on the macro step (1 temperature update).
+\fig{fig:conv} is an error convergence plot comparing the three techniques where temperature is evaluated 5 times within a macro step (5 temperature updates).
+Finally, \fig{fig:mp} shows the affect of various temperature updates.
+The dotted lines correspond to brute force at different flux step sizes, while the IQS macro step size is kept constant.
 
-For those who like equations in their papers, \LaTeX\ is a good choice. Here is
-an equation for the Marshak diffusion boundary condition:
-\begin{equation} \label{eq:marshak}
-  4 J^- = \phi + 2 D \vec{n} \vd \grad \phi \,.
-\end{equation}
-If we so choose, we can effortlessly reference the equation later.
-
-Another paragraph starts with Eq.~\eqref{eq:marshak} and sets $J^-$ to zero, a
-vacuum boundary condition:
-\begin{equation*}
-  0 = \phi + \frac{2}{3} \frac{1}{\sigma} \vec{n} \vd \grad \phi \,.
-\end{equation*}
-The extrapolation distance is $2/3$. A more detailed asymptotic analysis yields
-an extrapolation distance of about $0.71045$.
-
-Figure~\ref{fig:voltage} shows how a plot might conceivably look in your
-document. Always place figures after they are referenced so as not to throw
-off the reader. You can use symbols and different line styles to help
-differentiate your results, especially if they are printed in black and white.
-Note how Fig.~\ref{fig:voltage} uses dashed lines \verb|--| for the exact
-solution, solid lines \verb|-| for the new method's solutions, and dotted lines
-\verb|:| for existing inaccurate methods.
-\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
-  \centering
-  \includegraphics{example_figure}
-  \caption{Captions are flush with the left.}
-  \label{fig:voltage}
+\begin{figure}[htbp!]
+\includegraphics[width=\linewidth]{lra_profile.png}
+\caption{LRA baseline temperature and power profile}
+\label{fig:lra_profile}
 \end{figure}
 
-Later on, we can include a table, even one that spans two columns such as
-Table~\ref{tab:widetable}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table*}[htb]
+\begin{figure}[!htpb]
+\includegraphics[width=\linewidth]{lra_bad.jpg}
+\caption{LRA convergence plots with only one temperature updated each macro step}
+\label{fig:lra_bad}
+\end{figure}
+
+\begin{figure}[htbp!]
 \centering
-\begin{tabular}{llllllllll}\toprule
-      & $\phi_T(0)$      & $\phi_T(10)$      & $\phi_T(20)$      &
-      $\phi_D(0)$      & $\phi_D(10)$      & $\phi_D(20)$      & $\rho$      &
-      $\varepsilon$      & $N_\text{it}$
-\\ \midrule
-$c=0.999$  & 0.9038 & 20.63 & 31.24 & 0.9087 & 20.63 & 31.23 & 0.2192 & $10^{-7}$ & 15
-\\
-$c=0.990$  & 0.3675 & 13.04 & 24.7 & 0.3696 & 13.04 & 24.69 & 0.2184 & $10^{-7}$ & 15
-\\
-$c=0.900$  & 0.009909 & 4.776 & 17.64 & 0.009984 & 4.786 & 17.63 & 0.2118 & $10^{-7}$ & 14
-\\
-$c=0.500$  & $6.069\times 10^{-5}$ & 2.212 & 15.53 & 6.213$\times 10^{-5}$ & 2.239 & 15.53 & 0.2068 & $10^{-7}$ & 13
-\\
-\bottomrule
-\end{tabular}
-  \caption{This is an example of a really wide table which might not normally
-  fit in the document.}
-  \label{tab:widetable}
-\end{table*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Notice how the table reference uses a Roman numeral
-for its numbering scheme, whereas the figure reference uses an Arabic numeral.
-For one-column tables, use the \verb|table| environment; two-column tables use
-\verb|table*|. The same applies to figures.
+\includegraphics[width=\linewidth]{lra_mp_convergence.png}
+\caption{Error convergence plot with 5 temperature updates per macro step}
+\label{fig:conv}
+\end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Another Subsection}
-Excessive sectioning in a three-page document is discouraged, but here are more
-subsections to demonstrate compliance with the ANS formatting guidelines.
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=\linewidth]{lra_mp.png}
+\caption{Error plot with various temperature updates per macro step}
+\label{fig:mp}
+\end{figure}
 
-\subsubsection{Third-level Heading}
-This subsubsection shows compliance with the ANS-specified standard. This level
-of heading should be used rarely.
-
-\subsubsection{Another Such Heading}
-And, if you really think you need a third-level heading, you should make sure
-that your subsection needs at least two of them.
+The convergence plots show that updating temperature and the PRKE parameters within a macro step has a significant affect on the performance of IQS.
+With only one update, IQS was only slightly better than brute force, brute force required about 1.5X more time steps than IQS for the same error.
+While 5 temperature updates showed a much more significant IQS performcance, brute force required about 4X more time steps than IQS for the same error.
+\fig{fig:mp} shows that there is an exponential convergence of error for the number of temperature updates.
+This convergence makes sense because temperature can only be so accurate before the error in shape is dominating.
 
@@
 
-The included ANS style file and this clear example file are a panacea for
-the hours of headache that invariably results from formatting a document in
-Microsoft Word.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\section{Appendix}
-
-Numbering in the appendix is different:
-\begin{equation} \label{eq:appendix}
-  2 + 2 = 5\,.
-\end{equation}
-and another equation:
-\begin{equation} \label{eq:appendix2}
-  a + b = c\,.
-\end{equation}
+%\appendix
+%\section{Appendix}
 
@@
 \section{Acknowledgments}
-This material is based upon work supported a Department of Energy Nuclear
-Energy University Programs Graduate Fellowship.
+This work was supported by the Department of Energy, Idaho National Laboratory, and the Integrated University Program Fellowship.
+Special thanks to Mark Dehart, Yaqi Wang, NEAMS, and the Moose/Rattlesnake team at INL.
 

### 2016-10-17 doc/M&C_2017/MC2017_Summary.tex
comments for mc 2017 summary

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Multiphysics Techniques for Neutron Diffusion Using Improved Quasi-Static Method}
-\author{Zachary M.~Prince, Jean C. Ragusa}
+% \title{Multiphysics Techniques for Neutron Diffusion Using Improved Quasi-Static Method}
+\title{Multiphysics Core Dynamics Simulation Using the Improved Quasi-Static Method}
+\author{Zachary M.~Prince, Jean C.~Ragusa}
 
@@
 
-The purpose of this summary is to introduce several techniques in dealing with multiphysics feedback for quasi-static neutron diffusion calculations.
-Other physical variables, including temperature, are tightly coupled with neutron flux when simulating a nuclear reactor, thus are very important quantities to evaluate accurately and efficiently.
-The improved quasi-static method (IQS) is an effective technique in simulating flux in reactors and the effect of multiphysics with IQS is worth exploring.
+The purpose of this summary is to introduce several techniques in dealing with multiphysics feedback for transient neutron diffusion calculations.  
+In dynamic simulation, the neutron temporal distribution in a nuclear core can be strongly influenced by non-neutronics variables, e.g., temperature.
+The improved quasi-static method (IQS) is an effective technique for simulating the temporal behavior of the neutron flux in reactors. Here, we present a study
+combining the IQS method with multiphysics solution techniques for coupled transient calculations.
 
-IQS is a transient spatial kinetics method that involves factorizing neutron flux into a space- and time-dependent component (shape) and a time-dependent component (amplitude) \cite{Ott_1966,Dulla2008}.
-The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates.
-IQS has mostly been applied to neutron kinetics, without thermal-hydraulic feedback.
-This summary discusses the application of multiphysics feedback with IQS and analyzes performance with temperature feedback benchmark problems
+IQS is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (shape) and a time-dependent component 
+(amplitude) \cite{Ott_1966,Dulla2008}. The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape 
+computations or updates. IQS has \tcr{mostly} been applied to neutron kinetics, without thermal-hydraulic feedback. This summary discusses the application of 
+multiphysics feedback with IQS and analyzes performance with temperature feedback benchmark problems
 
@@
 
-In this Section, we recall the equation for the IQS method, starting from the standard multi-group transport equations with delayed neutron precursors in operator form:
+In this Section, we recall the equation for the IQS method, starting from the standard multi-group \tcr{transport} equations with delayed neutron precursors in operator form:
 
@@
 
-Where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the diffusion and removal operator, $S_{d}^g$ is the delay source, and $P_{d,i}^g$ is the delay fission operator.
+Where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the streaming and interaction operator, 
+$S_{d}^g$ is the delay source, and $P_{d,i}^g$ is the delayed-neutron fission operator.
 
-Factorization is an important step in the derivation of the IQS method.
-The factorization approach leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent multigroup shape ($\psi$):
+The flux factorization approach leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent 
+multigroup shape ($\psi$): \tcr{your notation is not consistent. psi. phi?}
 \be
@@
 
-We note that the time-dependent shape equation is similar to the time-dependent flux.
-The main differences is that there is a nonlinear coupling with amplitude contained in the addition of the boxed terms.
+Note that the time-dependent shape equation is similar to the time-dependent flux equation. 
+However, the shape/amplitude equations are now non-linearly coupled (boxed terms).
 
@@
 
-This version of IQS first solves the flux diffusion (represented by Equations \ref{eq:flux} and \ref{eq:precursor}) to get a predicted flux.
-The predicted flux at this step is then converted to shape by rescaling as follows:
+To avoid performing the IQS non-linear solve, a predictor-corrector of IQS has been derived in \cite{Dulla2008}. 
+One first solves the neutron flux (represented by Equations \ref{eq:flux} and \ref{eq:precursor}) to obtain a predicted flux.  
+The predicted flux is then converted to a shape through a rescaling argument :
 \be
@@
 
-The PRKE parameters are then computed with this shape using Equations \ref{eq:rmb} - \ref{eq:l} and interpolated over the macro step, then the PRKE is evaluated.
-With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
+The PRKE parameters are then computed with this shape using Equations \eqref{eq:rmb}-\eqref{eq:l} and interpolated over the macro step
+for the solution of the PRKE equations.  With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
 \be
-\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1}
+\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1} \,.
 \ee
@@
 Other physical quantities, such as temperature, are affected by reaction rates and subsequently affect the operators of the flux equations.  
-For IQS, this feedback affects both the shape equation and the parameters of the PRKE; thus, it is an additional nonlinear component to the already coupled shape-amplitude equations.
-Each of these components have different temporal dependencies; so it is beneficial for efficiency to evaluate them on different time scales.
-Amplitude is much more time dependent than shape, and shape is much more computationally expensive to evaluate.
-So amplitude is evaluated on a much smaller time scale than shape, while it may be appropriate so solve quantities, such as temperature, on middle time scale.
-Figure~\ref{fig:time} shows a solution process for temperature feedback.
+For IQS, this feedback affects both the shape equation and the parameters of the PRKE; thus, it is an additional nonlinear component to the 
+already non-linear shape-amplitude equations.  Each of these components have different temporal dependencies; so it may be beneficial for 
+efficiency to evaluate them on different time scales.  The amplitude is more rapidly varying than the shape which is computationally expensive to evaluate
+and is evaluated only on macro-time steps. In multiphysics simulations, one may take advantage of a fine-scale power distribution in the coupled
+physics components (temperature). Figure~\ref{fig:time} shows a such a solution process for temperature feedback.
 
@@
 
-The top time scale represents a shape diffusion evaluation on a macro step, the middle as an arbitrary three steps within the macro step where temperature and the PRKE parameters are evaluated, and the bottom one represents the PRKE evaluation on micro steps.
-The shape is linearly interpolated within the macro step for the temperature and PRKE parameter evaluation, and the parameters are interpolated within the temperature step for the PRKE evaluation.
-Since there is a nonlinear coupling between all these components, each temperature step is iterated until amplitude has converged and the macro step is iterated until shape has converged.
+The top time scale represents a shape diffusion evaluation on a macro step, the middle as an arbitrary 
+three steps within the macro step where temperature and the PRKE parameters are evaluated, and the 
+bottom one represents the PRKE evaluation on micro steps.  The shape is linearly interpolated within 
+the macro step for the temperature and PRKE parameter evaluation, and the parameters are interpolated 
+within the temperature step for the PRKE evaluation.  Since there is a nonlinear coupling between all 
+these components, each temperature step is iterated until amplitude has converged and the macro step is iterated until shape has converged.
 
@@
 To test the multiple time scale implementation of IQS, the LRA benchmark was chosen for reference. 
-The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.
-It is a super prompt-critical transient.
-Three different evaluation techniques were applied: full discretization of the flux equation (Brute force), IQS, and IQS-PC.
-Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.
-The performance of IQS and the temperature updates was measured by its improvement in accuracy at peak power over the Brute force technique.
-The results show the effectiveness of multiple temperature updates along the macro step.
+The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and 
+Doppler feedback in thermal reactor \cite{ANL_BPB}.  It is a super prompt-critical transient. 
+Three different evaluation techniques were applied: temporal discretization of the flux equation (Brute force), 
+IQS, and IQS-PC. Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.  
+The performance of IQS and the temperature updates was measured by its improvement in accuracy at peak power 
+over the Brute force technique.  The results show the effectiveness of multiple temperature updates along the macro step.
+\tcr{you comment on results too early}
 
@@
 \fig{fig:lra_profile} shows the baseline power and temperature profile for the LRA benchmark.  
-The relative difference in the magnitude of the peak power ($t\approx1.44 s$) from the baseline was used for error comparison.
-\fig{fig:lra_bad} is an error convergence plot comparing the three techniques where temperature is evaluated only on the macro step (1 temperature update).
-\fig{fig:conv} is an error convergence plot comparing the three techniques where temperature is evaluated 5 times within a macro step (5 temperature updates).
-Finally, \fig{fig:mp} shows the affect of various temperature updates.
+The relative difference in the magnitude of the peak power ($t\approx1.44 s$) from the baseline 
+was used for error comparison.  \fig{fig:lra_bad} is an error convergence plot comparing the 
+three techniques where temperature is evaluated only on the macro step (1 temperature update).  
+\fig{fig:conv} is an error convergence plot comparing the three techniques where temperature is evaluated 
+5 times within a macro step (5 temperature updates).  Finally, \fig{fig:mp} shows the affect of various temperature updates.  
 The dotted lines correspond to brute force at different flux step sizes, while the IQS macro step size is kept constant.
@@
 
-The convergence plots show that updating temperature and the PRKE parameters within a macro step has a significant affect on the performance of IQS.
-With only one update, IQS was only slightly better than brute force, brute force required about 1.5X more time steps than IQS for the same error.
-While 5 temperature updates showed a much more significant IQS performcance, brute force required about 4X more time steps than IQS for the same error.
-\fig{fig:mp} shows that there is an exponential convergence of error for the number of temperature updates.
+The convergence plots show that updating temperature and the PRKE parameters within a macro step has 
+a significant affect on the performance of IQS.  With only one update, IQS was only slightly better than brute force, 
+brute force required about 1.5x more time steps than IQS for the same error.  While 5 temperature updates showed 
+a much more significant IQS performance, brute force required about 4x more time steps than IQS for the same error.  
+\fig{fig:mp} shows that there is an \tcr{exponential} convergence of error for the number of temperature updates.  
 This convergence makes sense because temperature can only be so accurate before the error in shape is dominating.
@@
 This work was supported by the Department of Energy, Idaho National Laboratory, and the Integrated University Program Fellowship.  
-Special thanks to Mark Dehart, Yaqi Wang, NEAMS, and the Moose/Rattlesnake team at INL.
+We thank Mark Dehart, Yaqi Wang, the NEAMS program, and INL's Moose/Rattlesnake team for their support.
 

### 2016-10-25 doc/M&C_2017/MC2017_Summary.tex
Pretty good draft

@@
 
-The purpose of this summary is to introduce several techniques in dealing with multiphysics feedback for transient neutron diffusion calculations.  
-In dynamic simulation, the neutron temporal distribution in a nuclear core can be strongly influenced by non-neutronics variables, e.g., temperature.
-The improved quasi-static method (IQS) is an effective technique for simulating the temporal behavior of the neutron flux in reactors. Here, we present a study
-combining the IQS method with multiphysics solution techniques for coupled transient calculations.
+The purpose of this summary is to introduce several techniques in dealing with multi-physics feedback for transient neutron diffusion calculations.
+In dynamic simulation, the neutron temporal distribution in a nuclear core can be strongly influenced by non-neutronics variables, e.g., temperature.The improved quasi-static method (IQS) is an effective technique for simulating the temporal behavior of the neutron flux in reactors.
+Here, we present a study combining the IQS method with multi-physics solution techniques for coupled transient calculations.
 
-IQS is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (shape) and a time-dependent component 
-(amplitude) \cite{Ott_1966,Dulla2008}. The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape 
-computations or updates. IQS has \tcr{mostly} been applied to neutron kinetics, without thermal-hydraulic feedback. This summary discusses the application of 
-multiphysics feedback with IQS and analyzes performance with temperature feedback benchmark problems
+IQS is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (shape) and a time-dependent component (amplitude) \cite{Ott_1966,Dulla2008}.
+The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates.
+IQS has largely been applied to neutron kinetics, without thermal-hydraulic feedback.
+This summary discusses the application of multi-physics feedback with IQS and analyzes performance with temperature feedback benchmark problems
 
@@
 
-In this Section, we recall the equation for the IQS method, starting from the standard multi-group \tcr{transport} equations with delayed neutron precursors in operator form:
+In this Section, we recall the equation for the IQS method, starting from the standard multi-group diffusion equations with delayed neutron precursors in operator form:
 
@@
 The flux factorization approach leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent 
-multigroup shape ($\psi$): \tcr{your notation is not consistent. psi. phi?}
+multigroup shape ($\varphi$):
 \be
@@
 To test the multiple time scale implementation of IQS, the LRA benchmark was chosen for reference.
-The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and 
-Doppler feedback in thermal reactor \cite{ANL_BPB}.  It is a super prompt-critical transient. 
-Three different evaluation techniques were applied: temporal discretization of the flux equation (Brute force), 
-IQS, and IQS-PC. Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.  
-The performance of IQS and the temperature updates was measured by its improvement in accuracy at peak power 
-over the Brute force technique.  The results show the effectiveness of multiple temperature updates along the macro step.
-\tcr{you comment on results too early}
+The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.
+It is a super prompt-critical transient.
+Three different evaluation techniques were applied: temporal discretization of the flux equation (Brute force), IQS, and IQS-PC.
+Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.
+The performance of IQS and the temperature updates was measured by its improvement in accuracy at peak power over the Brute force technique.
 
@@
 \fig{fig:lra_profile} shows the baseline power and temperature profile for the LRA benchmark.
-The relative difference in the magnitude of the peak power ($t\approx1.44 s$) from the baseline 
-was used for error comparison.  \fig{fig:lra_bad} is an error convergence plot comparing the 
-three techniques where temperature is evaluated only on the macro step (1 temperature update).  
-\fig{fig:conv} is an error convergence plot comparing the three techniques where temperature is evaluated 
-5 times within a macro step (5 temperature updates).  Finally, \fig{fig:mp} shows the affect of various temperature updates.  
+The relative difference in the magnitude of the peak power ($t\approx1.44 s$) from the baseline was used for error comparison.
+\fig{fig:lra_bad} is an error convergence plot comparing the three techniques where temperature is evaluated only on the macro step (1 temperature update).
+\fig{fig:conv} is an error convergence plot comparing the three techniques where temperature is evaluated 5 times within a macro step (5 temperature updates).
+Finally, \fig{fig:mp} shows the affect of various temperature updates.
 The dotted lines correspond to brute force at different flux step sizes, while the IQS macro step size is kept constant.
@@
 
-The convergence plots show that updating temperature and the PRKE parameters within a macro step has 
-a significant affect on the performance of IQS.  With only one update, IQS was only slightly better than brute force, 
-brute force required about 1.5x more time steps than IQS for the same error.  While 5 temperature updates showed 
-a much more significant IQS performance, brute force required about 4x more time steps than IQS for the same error.  
-\fig{fig:mp} shows that there is an \tcr{exponential} convergence of error for the number of temperature updates.  
+The convergence plots show that updating temperature and the PRKE parameters within a macro step has a significant affect on the performance of IQS.
+With only one update, IQS was only slightly better than brute force, brute force required about 1.5x more time steps than IQS for the same error.
+While 5 temperature updates showed a much more significant IQS performance, brute force required about 4x more time steps than IQS for the same error.
+\fig{fig:mp} shows that error has a convergent behavior for the number of temperature updates.
 This convergence makes sense because temperature can only be so accurate before the error in shape is dominating.
@@
 \section{Conclusions}
+
+IQS is a powerful tool for reactor simulation, it has the capability of greatly reducing computation time while retaining accuracy of relevant quantities.
+The goal of this summary is to showcase IQS's performance with multi-physics simulation, as well as techniques to improve its performance.
+The technique focused on incorporating a new time scale in IQS for feedback quantities and PRKE parameters.
+The results when IQS is applied to the temperature feedback model of the LRA benchmark showed that incorporating a middle time scale for temperature had significant impact on IQS performance.
+
+A full report of this topic will include applying step doubling time adaptation to the LRA benchmark and analyzing IQS performance.
+It is also intended to include IQS application to a full core TREAT model to analyze performance with a more complex problem.
+It is also possible to include IQS application to the Godiva model as to analyze performance with a structural feedback problem \cite{CritSafetyHandbook}.
 
@@
 \bibliographystyle{ans}
-\bibliography{bibliography}
+\bibliography{references_IQS}
 \end{document}

### 2016-10-28 doc/M&C_2017/MC2017_Summary.tex
final review

@@
 IQS is a powerful tool for reactor simulation, it has the capability of greatly reducing computation time while retaining accuracy of relevant quantities.
-The goal of this summary is to showcase IQS's performance with multi-physics simulation, as well as techniques to improve its performance.
+The goal of this summary is to demonstrate IQS's performance with multi-physics simulations, as well as techniques to improve its performance.
 The technique focused on incorporating a new time scale in IQS for feedback quantities and PRKE parameters.
-The results when IQS is applied to the temperature feedback model of the LRA benchmark showed that incorporating a middle time scale for temperature had significant impact on IQS performance.
+The results when IQS is applied to the temperature feedback model of the LRA benchmark showed that incorporating a intermediate time scale for temperature has significant impact on IQS performance.
 
-A full report of this topic will include applying step doubling time adaptation to the LRA benchmark and analyzing IQS performance.
-It is also intended to include IQS application to a full core TREAT model to analyze performance with a more complex problem.
-It is also possible to include IQS application to the Godiva model as to analyze performance with a structural feedback problem \cite{CritSafetyHandbook}.
+The full paper version will include time adaptive simulations for multiphysics dynamic transients: LRA benchmark, a full core TREAT model, and the Lady Godiva benchmark with structural feedback \cite{CritSafetyHandbook}.
+Performance of the IQS method with intermediate time scale for physics feedback will be compared to that of an implicit discretization of the flux equations with feedback.
 

### 2016-10-30 doc/M&C_2017/MC2017_Summary.tex
Final Draft

@@
 
-To avoid performing the IQS non-linear solve, a predictor-corrector of IQS has been derived in \cite{Dulla2008}. 
+To avoid performing the IQS non-linear solve, a predictor-corrector version of IQS has been derived in \cite{Dulla2008}.
 One first solves the neutron flux (represented by Equations \ref{eq:flux} and \ref{eq:precursor}) to obtain a predicted flux.
@@
 
-The PRKE parameters are then computed with this shape using Equations \eqref{eq:rmb}-\eqref{eq:l} and interpolated over the macro step
-for the solution of the PRKE equations.  With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
+The PRKE parameters are then computed with this shape using Equations \eqref{eq:rmb}-\eqref{eq:l} and interpolated over the macro step for the solution of the PRKE equations.
+With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
 \be
@@
 
-The top time scale represents a shape diffusion evaluation on a macro step, the middle as an arbitrary 
-three steps within the macro step where temperature and the PRKE parameters are evaluated, and the 
-bottom one represents the PRKE evaluation on micro steps.  The shape is linearly interpolated within 
-the macro step for the temperature and PRKE parameter evaluation, and the parameters are interpolated 
-within the temperature step for the PRKE evaluation.  Since there is a nonlinear coupling between all 
-these components, each temperature step is iterated until amplitude has converged and the macro step is iterated until shape has converged.
+The top time scale represents a shape diffusion evaluation on a macro step, the middle has an arbitrary three steps within the macro step where temperature and the PRKE parameters are evaluated, and the bottom one represents the PRKE evaluation on micro steps.
+The shape is linearly interpolated within the macro step for the temperature and PRKE parameter evaluation, and the parameters are interpolated within the temperature step for the PRKE evaluation.
+Since there is a nonlinear coupling between all these components, each temperature step is iterated until amplitude has converged and the macro step is iterated until shape has converged.
 
@@
 It is a super prompt-critical transient.
-Three different evaluation techniques were applied: temporal discretization of the flux equation (Brute force), IQS, and IQS-PC.
+Three different evaluation techniques were applied: implicit discretization of the flux equation (Brute force), IQS, and IQS-PC.
 Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.
-The performance of IQS and the temperature updates was measured by its improvement in accuracy at peak power over the Brute force technique.
+Third order Runge–Kutta discretization with step doubling adaptation was used for the PRKE evaluation.
+The performance of IQS and the temperature updates were measured by its improvement in accuracy at peak power over the Brute force technique.
 
@@
 However, IQS has much more information about the profile of the flux along the time step because of the micro-step amplitude evaluation.
-Therefore, it is possible to solve for temperature using a quasi-analytical approach, shown by \eqt{eq:temp_an}.
+Therefore, it is possible to solve for temperature using a semi-analytical approach, shown by \eqt{eq:temp_an}.
 
@@
 Finally, \fig{fig:mp} shows the affect of various temperature updates.
-The dotted lines correspond to brute force at different flux step sizes, while the IQS macro step size is kept constant.
+The dashed lines correspond to brute force at different flux step sizes, while the IQS macro step size is kept constant.
 
@@
 \includegraphics[width=\linewidth]{lra_bad.jpg}
-\caption{LRA convergence plots with only one temperature updated each macro step}
+\caption{LRA convergence plots with only one temperature update each macro step}
 \label{fig:lra_bad}
@@
 The convergence plots show that updating temperature and the PRKE parameters within a macro step has a significant affect on the performance of IQS.
-With only one update, IQS was only slightly better than brute force, brute force required about 1.5x more time steps than IQS for the same error.
-While 5 temperature updates showed a much more significant IQS performance, brute force required about 4x more time steps than IQS for the same error.
+With only one update, IQS was only slightly better than brute force, brute force required about 150\% more time steps than IQS for the same error.
+While 5 temperature updates showed a much more significant IQS performance, brute force required about 400\% more time steps than IQS for the same error.
 \fig{fig:mp} shows that error has a convergent behavior for the number of temperature updates.

### 2016-11-21 PetSc_mat_view/rattlesnake_test_1.m
adding petsc mat view

@@ -1,0 +1,352 @@
+close all;
+
+fd = PetscOpenFile('A');
+A = PetscBinaryRead(fd);
+
+fd = PetscOpenFile('D');
+D = PetscBinaryRead(fd);
+
+
+% % condest(A)
+% % condest(D)
+% % 
+
+figure(1)
+subplot(1,4,1); spy(A,5);
+subplot(1,4,2); spy(D,5);
+
+G=11;
+[N,~]=size(A);
+n=N/G;
+
+k=1:G:N;
+if length(k)~=n
+    error('lengths');
+end
+
+AA = spalloc(N,N,nnz(A));
+row_id=0;
+for g=1:G
+    for i=1:n
+        ia=k(i)+(g-1);
+        row=[];
+        for gg=1:G
+            ja=k+(gg-1);
+            row = [row A(ia,ja)];
+        end
+        row_id=row_id+1;
+        AA(row_id,:)=row;
+    end
+end
+subplot(1,4,3); spy(AA,5);
+
+DD = spalloc(N,N,nnz(D));
+row_id=0;
+for g=1:G
+    for i=1:n
+        ia=k(i)+(g-1);
+        row=[];
+        for gg=1:G
+            ja=k+(gg-1);
+            row = [row D(ia,ja)];
+        end
+        row_id=row_id+1;
+        DD(row_id,:)=row;
+    end
+end
+subplot(1,4,4); spy(AA-DD,5);
+
+Sparsity = spones(AA);
+sss=Sparsity(1:n,1:n);
+% xs = eye(G,G);
+% xs(7:G,7:G)=1;
+% xs(6,7)=1;
+% xs(5:8,6)=1;
+% xs(2:4,1)=1;
+% xs(3:4,2)=1;
+% xs(4,3)=1;
+% xs(5,4)=1;
+% xs(6,5)=1;
+% spy(kron(xs,sss)-Sparsity)
+
+figure(99)
+subplot(2,3,1); spy(A,5); title('full system')
+subplot(2,3,2); spy(D,5); title('block diag precondtioner')
+subplot(2,3,3); spy(A-D,5); title('full minus block diag')
+subplot(2,3,4); spy(AA,5); title('full system re-ordered')
+subplot(2,3,5); spy(DD,5); title('block diag precondtioner re-ordered')
+subplot(2,3,6); spy(AA-DD,5); title('full minus block diag re-ordered')
+
[…]

### 2016-12-17 doc/Thesis_proposal/thesis_proposal.tex
Working on thesis proposal

@@
 However, recent developments in nuclear testing, including the revitalization of the Transient Reactor Testing (TREAT) Facility at Idaho National laboratory (INL), have brought significant attention to the development of transient reactor modeling.
-Transient computational methods currently in \tcr{laboratory frameworks--people may not understand what you mean here} have proven to be overbearingly computationally expensive for complex, real-world scenarios.
-Therefore, more efficient and faster methods for computational \tcr{frameworks} are highly desired.
+Transient computational methods currently in use at national laboratories, including INL, have proven to be overbearingly computationally expensive for complex, real-world scenarios.
+Therefore, implementation of more efficient and faster methods is highly desired.
 The Department of Energy (DOE) has especially sought to develop in the MOOSE framework.
@@
 For the purposes of this research, neutron diffusion will be discussed exclusively.
-\tcr{Additionally, simulations can often neglect temporal dependence due to the steady-state nature of reactors--???}.
-However, simulations for facilities such as TREAT, must retain time dependence in order to obtain relevant information.
 \\
@@
 Most reactor computation frameworks, including Rattlesnake, discretize the time variable directly with these equations using a multitude of schemes (Implicit Euler, Crank-Nicholson, Runge-Kutta, etc.).
-In this paper, this method of discretizing Equations~\ref{eq:flux}~and~\ref{eq:precursor} is considered \tcr{brute-force--I wonder if we should come up with a better terminology...} method.
+In this paper, this method of discretizing Equations~\ref{eq:flux}~and~\ref{eq:precursor} is considered implicit discretization.
 This research intends to improve upon this method by instead implementing the improved quasi-static method (IQS).
@@
 The weight function is typically the adjoint flux $\phi^{*g}$, which can be proven to minimize truncation error \cite{duderstadt1976nuclear}.
-\tcr{Check Cristian Rabitti's PhD work.
-There is more to it.
-He used a new adjoint at each time step I believe.
-It is interesting to read}
 The final expressions are given below:
@@
 The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps ($\varphi^g_{\Delta t/2}$).
-Note: $\varphi$ is changed to $\phi$ for brute-force and IQS P-C.
+Note: $\varphi$ is changed to $\phi$ for implicit discretization and IQS P-C.
 \\
@@
 
-To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to brute force method, traditional IQS, and IQS P-C.
+To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to implicit discretization method, traditional IQS, and IQS P-C.
 Each of these methods will be applied to several diffusion problems; the number of time steps taken and the resulting error will be used to compare the methods.
@@
 The Rattlesnake radiation transport framework is a excellent option for performing these problems.
-Rattlesnake already contains a neutron diffusion method for continuous-FEM equivalent to the brute-force method.
+Rattlesnake already contains a neutron diffusion method for continuous-FEM equivalent to the implicit discretization method.
 Additionally, Rattlesnake is well documented and well organized, so implementing IQS may be relatively straightforward.
@@
 
-\fig{fig:power} shows the power at each macro time step as compared to the traditional brute force (full flux time discretization) method.
+\fig{fig:power} shows the power at each macro time step as compared to the traditional brute force (full flux implicit discretization) method.
 The strong correlation between the two curves shows that IQS is consistent with a proven method for a highly transient example.
-\fig{fig:conv} shows the error convergence for brute force and IQS methods.
+\fig{fig:conv} shows the error convergence for implicit discretization and IQS methods.
 This plot shows that IQS has consistent convergence slope for first and second order time schemes and that IQS has a better error constant for the convergence.
 \tbl{tab:1Ddt2} shows the results for time adaptation.
-These results show that traditional IQS does not perform well with step doubling time adaptation and IQS P-C performs moderately better than brute force.
+These results show that traditional IQS does not perform well with step doubling time adaptation and IQS P-C performs moderately better than implicit discretization.
 The IQS performance in this example are not exceptional because it is not an easy problem.
@@
 \hline
-  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
+  &  & \multicolumn{3}{|c|}{Implicit Discretization} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
 \hline
@@
 \item Validate and verify IQS method with several low and high order time discretization schemes.
-\item Implement step doubling time adaptation with IQS and analyze its performance compared to brute-force.
+\item Implement step doubling time adaptation with IQS and analyze its performance compared to implicit discretization.
 \item Apply temperature feedback with IQS and analyze its performance with multiphysics benchmarks.
@@
 \end{enumerate}
-
-\tcr{for a paper, can we think of testing Embeded RK instead of/in addition to step-doubling?}
 

### 2016-12-20 doc/Thesis_proposal/thesis_proposal.tex
final comments on the MS proposal

@@
 However, recent developments in nuclear testing, including the revitalization of the Transient Reactor Testing (TREAT) Facility at Idaho National laboratory (INL), have brought significant attention to the development of transient reactor modeling.
-Transient computational methods currently in use at national laboratories, including INL, have proven to be overbearingly computationally expensive for complex, real-world scenarios.
-Therefore, implementation of more efficient and faster methods is highly desired.
-The Department of Energy (DOE) has especially sought to develop in the MOOSE framework.
-\\
+3-D transient computational methods currently in use at national laboratories, including INL, have proven to be overbearingly computationally expensive for complex, real-world scenarios.
+Therefore, implementation of more efficient and faster methods for 3-D transient simulations is highly desired.
+The Department of Energy (DOE), through its NEAMS (Nuclear Energy Advanced Modeling and Simulation) program, has especially sought the development of transient multiphysics capability in
+INL's MOOSE framework.  \\
 
@@
 Rattlesnake is a large application in MOOSE that involves deterministic radiation transport physics.
-Rattlesnake is able to solve steady-state, transient, and k-eigenvalue neutron transport and diffusion problems using finite element method (FEM) \cite{wang2013}.
+Currently, Rattlesnake is able to solve steady-state, transient, and k-eigenvalue neutron transport and diffusion problems using finite element methods (FEM) \cite{wang2013}.
 \\
 
-The difficulty in transient reactor modeling is due to the neutron transport equation, which describes the flux of neutrons.
-The difficulty in solving or evaluating the equation comes from the dimensionality of the phase-space and the fact that the time discretization has to be implicit, which  leads to kinear system solves.
-The equation has seven independent variables: space ($\vec{r}$), energy ($E$), direction ($\vec{\Omega}$), and time ($t$) \cite{duderstadt1976nuclear}.
+The difficulty in transient reactor modeling is due to 
+\begin{enumerate}
+\item
+the high-dimensionality of the phase-space for the governing equations that describe the flux of neutrons (6-D+time for multigroup neutron transport and 4-D+time for multigroup neutron diffusion), and
+\item
+the fact that the time discretization has to be implicit, which  leads to stiff system of equations.  
+\end{enumerate}
+The transport equation has seven independent variables: space ($\vec{r}$), energy ($E$), direction ($\vec{\Omega}$), and time ($t$) \cite{duderstadt1976nuclear}.
 Nuclear reactor simulations often utilize the neutron diffusion approximation, which carefully eliminates the dependence on direction.
-For the purposes of this research, neutron diffusion will be discussed exclusively.
+For the purpose of this research, neutron diffusion will be discussed exclusively.
 \\
@@
 For computational purposes, the neutron diffusion equation is discretized in space, energy, and time.
-There are several viable discretization schemes for each of these variables, but FEM for space and multigroup method for energy were used for the development of this research.
+There are several viable discretization schemes for each of these variables; FEM for space and multigroup in energy are used for the development of this research.
 Since the purpose of this research involves temporal dependence, discretization in time was kept general.
@@
 \begin{subequations}
-\begin{align}
-\frac{1}{v^g}\frac{\partial \phi^g}{\partial t}  =& \frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \phi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g  \nonumber \\
-&  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \phi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
+\begin{multline}
+\frac{1}{v^g}\frac{\partial \phi^g}{\partial t}  = 
+\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \phi^{g'} 
++ \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \phi^{g'}  \\ + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i 
+-  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g   
+\ , \quad 1 \le g \le G 
 \label{eq:flux}
-\end{align}
+\end{multline}
 \be
@@
 
-Most reactor computation frameworks, including Rattlesnake, discretize the time variable directly with these equations using a multitude of schemes (Implicit Euler, Crank-Nicholson, Runge-Kutta, etc.).
-In this paper, this method of discretizing Equations~\ref{eq:flux}~and~\ref{eq:precursor} is considered implicit discretization.
-This research intends to improve upon this method by instead implementing the improved quasi-static method (IQS).
-\\
+Most reactor computation frameworks, including Rattlesnake, discretize the time variable directly with these equations using a multitude of schemes (Implicit Euler, Crank-Nicholson, implicit Runge-Kutta, etc.).
+In this paper, this method of discretizing Equations~\ref{eq:flux}~and~\ref{eq:precursor} is generally referred to as ``implicit discretization''.
+This research intends to improve upon this method by instead implementing the improved quasi-static method (IQS) for neutron kinetics and
+implement it within a multiphysics setting. \\
 
@@
 However, no comprehensive analysis of iteration techniques exists.
+\tcr{last sentence is a bit vague}
 
@@
 \item Investigate iteration convergence for previously implemented iterative schemes and newly conceived one described in \sct{sect:inv_iter}.
-\item Validate and verify IQS method with several low and high order time discretization schemes.
-\item Implement step doubling time adaptation with IQS and analyze its performance compared to implicit discretization.
+\item Verify IQS method with several low and high order time discretization schemes.
+\item Implement step doubling time adaptation with IQS and analyze its performance compared to implicit discretization of the flux equations.
 \item Apply temperature feedback with IQS and analyze its performance with multiphysics benchmarks.
 \item Implement IQS into the Rattlesnake framework at INL.
-\item Perform testing with IQS to benchmarks including TWIGL, LRA, and C5G7-2D.
+\item Perform testing with IQS to benchmarks including TWIGL and LRA. \tcr{what about the LMR benchmark? needed for paper}
 \end{enumerate}

### 2016-12-20 PetSc_mat_view/rattlesnake_test_1.m
Some changes to see how restart and precond affect gmres

@@ -16,9 +16,9 @@ F = PetscBinaryRead(fd);
 % % condest(D)
 % % 
 
-figure(1)
-subplot(1,4,1); spy(A,5);
-subplot(1,4,2); spy(D,5);
+% figure(1)
+% subplot(1,4,1); spy(A,5);
+% subplot(1,4,2); spy(D,5);
 
 G=11;
 [N,~]=size(A);
@@ -43,7 +43,7 @@ for g=1:G
         AA(row_id,:)=row;
     end
 end
-subplot(1,4,3); spy(AA,5);
+% subplot(1,4,3); spy(AA,5);
 
 DD = spalloc(N,N,nnz(D));
 row_id=0;
@@ -59,7 +59,7 @@ for g=1:G
         DD(row_id,:)=row;
     end
 end
-subplot(1,4,4); spy(AA-DD,5);
+% subplot(1,4,4); spy(AA-DD,5);
 
 SS = spalloc(N,N,nnz(S));
 row_id=0;
@@ -104,24 +104,24 @@ sss=Sparsity(1:n,1:n);
 % xs(6,5)=1;
 % spy(kron(xs,sss)-Sparsity)
 
-figure(99)
-subplot(2,3,1); spy(A,5); title('full system')
-subplot(2,3,2); spy(D,5); title('block diag precondtioner')
-subplot(2,3,3); spy(A-D,5); title('full minus block diag')
-subplot(2,3,4); spy(AA,5); title('full system re-ordered')
-subplot(2,3,5); spy(DD,5); title('block diag precondtioner re-ordered')
-subplot(2,3,6); spy(AA-DD,5); title('full minus block diag re-ordered')
-
-figure(69)
-subplot(3,3,1); spy(AA,5); title('Full system')
-subplot(3,3,2); spy(DD,5); title('Block Diag')
-subplot(3,3,3); spy(AA-DD,5); title('full minus block diag')
-subplot(3,3,4); spy(AA,5); title('Full system')
-subplot(3,3,5); spy(SS,5); title('Diag with Scattering')
-subplot(3,3,6); spy(AA-SS,5); title('full minus diag+scat')
-subplot(3,3,7); spy(AA,5); title('Full system')
-subplot(3,3,8); spy(FF,5); title('Block Diag with Fission')
-subplot(3,3,9); spy(AA-FF,5); title('full minus diag+fiss')
+% figure(99)
[…]

### 2016-12-20 doc/Thesis_proposal/thesis_proposal.tex
Addressing final MS proposal comments

@@
 in \cite{Dulla2008} does an in depth analysis of the fixed point iteration technique most similar to Sissaoui and Koclas, involving convergence rates and solution results.
-However, no comprehensive analysis of iteration techniques exists.
-\tcr{last sentence is a bit vague}
+However, no comprehensive analysis of iteration techniques exists, comparing both Newton and fixed-point convergence rates.
 
@@
 \item Implement IQS into the Rattlesnake framework at INL.
-\item Perform testing with IQS to benchmarks including TWIGL and LRA. \tcr{what about the LMR benchmark? needed for paper}
+\item Perform testing with IQS to benchmarks including TWIGL, LRA, and LMW.
 \end{enumerate}

### 2017-01-03 PetSc_mat_view/rattlesnake_test_1.m
More testing

@@ -112,16 +112,16 @@ sss=Sparsity(1:n,1:n);
 % subplot(2,3,5); spy(DD,5); title('block diag precondtioner re-ordered')
 % subplot(2,3,6); spy(AA-DD,5); title('full minus block diag re-ordered')
 % 
-% figure(69)
-% subplot(3,3,1); spy(AA,5); title('Full system')
-% subplot(3,3,2); spy(DD,5); title('Block Diag')
-% subplot(3,3,3); spy(AA-DD,5); title('full minus block diag')
-% subplot(3,3,4); spy(AA,5); title('Full system')
-% subplot(3,3,5); spy(SS,5); title('Diag with Scattering')
-% subplot(3,3,6); spy(AA-SS,5); title('full minus diag+scat')
-% subplot(3,3,7); spy(AA,5); title('Full system')
-% subplot(3,3,8); spy(FF,5); title('Block Diag with Fission')
-% subplot(3,3,9); spy(AA-FF,5); title('full minus diag+fiss')
+figure(69)
+subplot(3,3,1); spy(AA,5); title('Full system')
+subplot(3,3,2); spy(DD,5); title('Block Diag')
+subplot(3,3,3); spy(AA-DD,5); title('full minus block diag')
+subplot(3,3,4); spy(AA,5); title('Full system')
+subplot(3,3,5); spy(SS,5); title('Diag with Scattering')
+subplot(3,3,6); spy(AA-SS,5); title('full minus diag+scat')
+subplot(3,3,7); spy(AA,5); title('Full system')
+subplot(3,3,8); spy(FF,5); title('Block Diag with Fission')
+subplot(3,3,9); spy(AA-FF,5); title('full minus diag+fiss')
 
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-01-03 PetSc_mat_view/rattlesnake_test_2.m

@@ -1,0 +1,55 @@
+clear;close all
+
+fd = PetscOpenFile('A');
+A = PetscBinaryRead(fd);
+[N,~]=size(A);
+b=ones(N,1);
+restart=[];
+tol=1e-8;
+maxit=120;
+[x,flag,relres,iter,resvec] = gmres(A,b,restart,tol,maxit,A);
+semilogy(1:length(resvec),resvec/norm(A\b));
+leg = char('PJFNK+LU SMP full ltol=1e-8');
+hold on;
+
+fd = PetscOpenFile('PJFNK_GMRES_SMPfull_l8');
+A1 = PetscBinaryRead(fd);
+restart=[];
+tol=1e-8;
+maxit=120;
+[x,flag,relres,iter,resvec1] = gmres(A,b,restart,tol,maxit,A1);
+semilogy(1:length(resvec1),resvec1/norm(A1\b));
+leg = char(leg,'PJFNK+GMRES SMP full ltol=1e-8');
+
+fd = PetscOpenFile('PJFNK_GMRES_SMPfull_l10');
+A2 = PetscBinaryRead(fd);
+restart=[];
+tol=1e-10;
+maxit=120;
+[x,flag,relres,iter,resvec2] = gmres(A,b,restart,tol,maxit,A2);
+semilogy(1:length(resvec2),resvec2/norm(A2\b));
+leg = char(leg,'PJFNK+GMRES SMP full ltol=1e-10');
+
+fd = PetscOpenFile('NEWTON_GMRES_SMPfull_l8');
+A3 = PetscBinaryRead(fd);
+restart=[];
+tol=1e-8;
+maxit=120;
+[x,flag,relres,iter,resvec3] = gmres(A,b,restart,tol,maxit,A3);
+semilogy(1:length(resvec3),resvec3/norm(A3\b));
+leg = char(leg,'NEWTON+GMRES SMP full ltol=1e-8');
+
+fd = PetscOpenFile('PJFNK_GMRES_SMPdiag_l8');
+D1 = PetscBinaryRead(fd);
+restart=[];
+tol=1e-8;
+maxit=120;
+% [x,flag,relres,iter,resvec4] = gmres(A,b,restart,tol,maxit,eye(size(D1)),D1);
+[x,flag,relres,iter,resvec4] = gmres(A,b,restart,tol,maxit,D1);
+semilogy(1:length(resvec4),resvec4/norm(D1\b));
[…]

### 2017-01-03 doc/Thesis_proposal/thesis_proposal.tex
Addressing comments from peer edits

@@
 Transient modeling of nuclear reactors has been a chronically foreboding task due to its computationally expensive nature.
-However, recent developments in nuclear testing, including the revitalization of the Transient Reactor Testing (TREAT) Facility at Idaho National laboratory (INL), have brought significant attention to the development of transient reactor modeling.
+However, recent developments in nuclear testing, including the revitalization of the Transient Reactor Testing (TREAT) Facility at Idaho National Laboratory (INL), have brought significant attention to the development of transient reactor modeling.
 3-D transient computational methods currently in use at national laboratories, including INL, have proven to be overbearingly computationally expensive for complex, real-world scenarios.
 Therefore, implementation of more efficient and faster methods for 3-D transient simulations is highly desired.
-The Department of Energy (DOE), through its NEAMS (Nuclear Energy Advanced Modeling and Simulation) program, has especially sought the development of transient multiphysics capability in
-INL's MOOSE framework.  \\
+The Department of Energy (DOE), through its NEAMS (Nuclear Energy Advanced Modeling and Simulation) program, has especially sought the development of transient multiphysics capability in INL's MOOSE (Multiphysics Object-Oriented Simulation Environment).
+\\
 
@@
 Most reactor computation frameworks, including Rattlesnake, discretize the time variable directly with these equations using a multitude of schemes (Implicit Euler, Crank-Nicholson, implicit Runge-Kutta, etc.).
-In this paper, this method of discretizing Equations~\ref{eq:flux}~and~\ref{eq:precursor} is generally referred to as ``implicit discretization''.
+In this paper, the method of discretizing Equations~\ref{eq:flux}~and~\ref{eq:precursor} is generally referred to as ``implicit discretization''.
 This research intends to improve upon this method by instead implementing the improved quasi-static method (IQS) for neutron kinetics and
@@
 
-This research intends to do investigate the performance of the previously implemented iterative techniques and compare them to this newly implemented technique.
+This research intends to investigate the performance of the previously implemented iterative techniques and compare them to this newly implemented technique.
 
@@
 
-IQS aims at reducing the time discretization error in the flux solution by splitting the flux into an amplitude (that is highly resolved with at a micro time scale) and a shape (whose time-dependence is weaker than that of the flux itself).
+IQS aims at reducing the time discretization error in the flux solution by splitting the flux into an amplitude (highly resolved at a micro time scale) and a shape (whose time-dependence is weaker than that of the flux itself).
 Thus, by construction, the IQS approach may employ larger time-step sizes for comparable temporal error.
@@
 
-The couplings between temperature, amplitude, and shape are all nonlinear, so iteration processes are need for each time scale.
+The couplings between temperature, amplitude, and shape are all nonlinear, so iteration processes are needed for each time scale.
 The amplitude and temperature need to be iterated on the middle time scale until convergence on each temperature step.
-Then another iterative process needs to occur in the shape time scale scale on all three variables.
+Then another iterative process needs to occur in the shape time scale on all three variables.
 
@@
 This section describes preliminary results of an example that tests an IQS implementation and shows its effectiveness on computation speed and accuracy.
-The example is very simple and computes quickly, it entails a one dimensional, homogeneous 400 cm slab with a heterogenous perturbation in absorption cross section.
+The example is very simple and computes quickly; it entails a one dimensional, homogeneous 400 cm slab with a heterogenous perturbation in absorption cross section.
 \fig{fig:slab} shows how the regions of the slab are divided and \tbl{tab:1Dmat} shows the initial material properties.
@@
 These results show that traditional IQS does not perform well with step doubling time adaptation and IQS P-C performs moderately better than implicit discretization.
-The IQS performance in this example are not exceptional because it is not an easy problem.
+The IQS performance in this example is not exceptional because the problem is difficult.
 As seen in \fig{fig:1D_shape}, the shape undergoes significant change through the transient, which requires many shape evaluations to retain accuracy.

### 2017-01-23 matlab/kinetics_driver_conv.m
Some small changes for MMS

@@ -3,6 +3,7 @@ function kinetics_driver_conv
 clc;
 % clear all;
 close all;
+clear global;
 % turn off warning due to interp1
 warning('OFF','MATLAB:interp1:ppGriddedInterpolant')
 
@@ -11,7 +12,7 @@ global npar io
 % verbose/output parameters
 io.console_print         = false;
 io.plot_transient_figure = false;
-io.plot_power_figure     = false;
+io.plot_power_figure     = true;
 io.make_movie            = false;
 io.save_flux             = false;
 io.print_progress        = false;
@@ -32,6 +33,7 @@ if pbID~=12
 else
     [phi0] = npar.phi_exact(npar.x_dofs',curr_time);
     [C0]   = assemble_source(npar.C_exact,curr_time);
+    C0 = kinetics_init(phi0,curr_time);
 end
 
 
@@ -42,7 +44,7 @@ u=[phi0;C0];
 u0=u;
 
 % time steping data
-t_end = 1.28;
+t_end = 1.1;
 % t_end = 1.3;
 % t_end = 0.1;
 
@@ -56,8 +58,8 @@ else
     amplitude_norm_ref = reference_solution( t_end, u0);
 end
 
-nn=9;
-ntimes = 2.^(3:nn-1)*4;
+nn=5;
+ntimes = 2.^(0:nn-1)*10;
 dt = t_end./ntimes;
 
 % Interpolation type of shape for IQS prke parameters
@@ -76,11 +78,11 @@ npar.an_interp_type='lagrange'; % This is just a place holder, actual definition
     
 % Number of steps in history for lagrange interpolation of precursors for
 % analytical elimination precursors (see solve_TD_diffusion_an_prec)
-npar.interpolation_order=1;
+npar.interpolation_order=3;
 
 % If this is true, precursors are recalculated using a cubic hermite
 % interpolation of shape (see solve_TD_diffusion_an_prec line 192)
-npar.hermite_prec_update=true;
+npar.hermite_prec_update=false;
 
 % Used of IQS_PC, determines how precursors are solved after diffusion
 % evaluation (see solve_IQS_PC_diffusion_elim_prec line 80)
@@ -94,9 +96,9 @@ npar.prec_solve_type = 'linear';
 
 i=0;
[…]

### 2017-01-23 matlab/problem_init.m

@@ -17,6 +17,8 @@ dat.lambda=0.1;
 dat.invvel=1.;
 
 dat.source_phi = @(x,t) 0;
+dat.source_phi = @(x,t) 0;
+dat.source_c   = @(x,t) 0;
 
 dat.source_phi = @(x,t) 0;
 dat.source_c   = @(x,t) 0;
@@ -153,7 +155,7 @@ switch problem_ID
     case 12
         % manufactured solution, one material, constant in space and time,
         % no precursors
-        do_steady = true;
+        do_steady = false;
         if do_steady
             dat.invvel=dat.invvel*0; % use this for steady state
         end
@@ -191,22 +193,26 @@ switch problem_ID
 %         A(t) = int(a(x,t),x,0,1);
 %         phi(x,t) = f(t)*a(x,t)/A(t);
 
-        phi(x,t) = x*(1-x)*(1+t)^2;
-        C(x,t) = x*(1-x)*(1+t)^1.2;
+        phi(x,t) = x*(1-x)*(1+t)^3;
+        C(x,t) = x*(1-x)*(1+t)^2;
         if do_steady
             phi(x,t) = x*(1-x); % use this for steady state
         end        
         npar.phi_exact = matlabFunction(phi);
         npar.C_exact = matlabFunction(C);
 %         npar.C_exact = @(x,t) 0.*x.*t;                
+%         npar.C_exact = matlabFunction(C);
+        npar.C_exact = @(x,t) 0.0*x.*t;
+                
 
-%         S_p(x,t) = iv*diff(phi,t) + (Sa-nfSf*(1-b))*phi - cdiff*diff(diff(phi,x),x) ;
-        S_p(x,t) = iv*diff(phi,t) + (Sa-nfSf*(1-b))*phi - cdiff*diff(diff(phi,x),x) -dat.lambda*C;
+        S_p(x,t) = iv*diff(phi,t) + (Sa-nfSf*(1-b))*phi - cdiff*diff(diff(phi,x),x) ;
+%         S_p(x,t) = iv*diff(phi,t) + (Sa-nfSf*(1-b))*phi - cdiff*diff(diff(phi,x),x) -dat.lambda*C;
         S_c(x,t) = diff(C,t) + dat.lambda *C -nfSf*b*phi;
         dat.source_phi = matlabFunction(S_p);
-        dat.source_c = matlabFunction(S_c);
-%         dat.source_c = @(x,t) 0.*x.*t;
[…]

### 2017-01-23 matlab/solve_IQS_diffusion_an_prec.m

@@ -36,7 +36,7 @@ for iter = 1: max_iter_iqs
     
     % solve for amplitude function
     if strcmpi(npar.prke_solve,'matlab')
-        [X,w,t,y]    = solve_prke_ode(X_beg(:,end),dt_macro,time_end,shape_beg(:,end),shape_end);
+        [X,dXdt,w,t,y]    = solve_prke_ode(X_beg(:,end),dt_macro,time_end,shape_beg(:,end),shape_end);
     else
         [X,dpdt,t,y] = solve_prke_iqs(X_beg(:,end),dt_macro,time_end,shape_beg(:,end),shape_end,npar.n_micro,npar.freq_react);
     end
@@ -97,7 +97,8 @@ for iter = 1: max_iter_iqs
         % OLD shortcut
         % OLD p=X(1);
         pi = ppval(pp,ti);
-        dpdti = ppval(pp_der,ti);
+%         dpdti = ppval(pp_der,ti);
+        dpdti = dXdt(1);
         
         % assemble IQS
         D    = assemble_stiffness(dat.cdiff   ,ti);

### 2017-01-23 matlab/solve_IQS_diffusion_elim_prec.m

@@ -40,7 +40,7 @@ for iter = 1: max_iter_iqs
     
     % solve for amplitude function over the entire macro-step
     if strcmpi(npar.prke_solve,'matlab')
-        [X,w,t,y] =  solve_prke_ode(X_beg(:,end),dt_macro,time_end,shape_beg(:,end),shape_end);
+        [X,dXdt,w,t,y] =  solve_prke_ode(X_beg(:,end),dt_macro,time_end,shape_beg(:,end),shape_end);
     else
         [X,dpdt,t,y] =  solve_prke_iqs(X_beg(:,end),dt_macro,time_end,shape_beg(:,end),shape_end,npar.n_micro,npar.freq_react);
     end
@@ -70,8 +70,13 @@ for iter = 1: max_iter_iqs
         
         % OLD shortcut
         % OLD p=X(1);
+%         pi = (1+ti)^2;
+%         dpdti = 2*(1+ti);
         pi = ppval(pp,ti);
-        dpdti = ppval(pp_der,ti);
+%         dpdti = ppval(pp_der,ti);
+%         tmp = funprke(ti,X);
+%         dpdti = tmp(1);
+        dpdti = dXdt(1);
         
         % assemble IQS
         D    = assemble_stiffness(dat.cdiff   ,ti);
@@ -137,6 +142,7 @@ for iter = 1: max_iter_iqs
         end
         % solve for new shape_end
         shape_end = A\rhs;
+%         shape_end = shape_beg(:,end);
         % finish Ci
         if strcmp(npar.method,'BDF')
             Ci = Ci +  bdf.b(order)*dt / deno * NFId_new*shape_end*pi;

### 2017-01-23 matlab/solve_prke_ode.m

@@ -1,4 +1,4 @@
-function [X,w,t,pow] =  solve_prke_ode(X,dt_macro,time_end,shape_beg,shape_end)
+function [X,dXdt,w,t,pow] =  solve_prke_ode(X,dt_macro,time_end,shape_beg,shape_end)
 
 % make the problem-data a global variable
 global dat npar
@@ -22,6 +22,7 @@ if npar.solve_prke_compute_rho_each_time
     dat.ode.shape_end = shape_end;
     % call ode solver with prke function that re-compute rho/beta at each time (expensive)
     [t,y]=ode15s(@funprke,[time_beg time_end],X,options);
+    dXdt = funprke(time_end,y(end,:)');
     w = 0;
 else
     switch npar.iqs_prke_interpolation_method
@@ -31,6 +32,7 @@ else
             [dat.ode.rho_MGT_end,dat.ode.beff_MGT_end,dat.ode.q_MGT_end]=compute_prke_parameters(time_end,shape_end);
             % call ode solver with prke function that linearly interpolates
             [t,y]=ode15s(@funprke_lin_interp,[time_beg time_end],X,options);
+            dXdt = funprke_lin_interp(time_end,y(end,:)');
         case 2 % linear variation for XS, linear for shape
             w = [ shape_beg shape_end]; 
             for k=1:size(w,2)
@@ -39,6 +41,7 @@ else
             end
             % call ode solver with prke function that linearly interpolates
             [t,y]=ode15s(@funprke_linlin_interp,[time_beg time_end],X,options);
+            dXdt = funprke_linlin_interp(time_end,y(end,:)');
         case 3 % linear variation for XS, cubic hermite for shape
             % hermite interpolant
             t1=time_beg;
@@ -54,6 +57,7 @@ else
             end
             % call ode solver with prke function that linearly interpolates
             [t,y]=ode15s(@funprke_linhermite_interp,[time_beg time_end],X,options);
+            dXdt = funprke_linhermite_interp(time_end,y(end,:)');
         case 4 % linear variation for XS, quadratic hermite for shape
             % hermite interpolant
             t1=time_beg;
@@ -69,6 +73,7 @@ else
             end
             % call ode solver with prke function that linearly interpolates
             [t,y]=ode15s(@funprke_linhermite_interp,[time_beg time_end],X,options);
+            dXdt = funprke_linhermite_interp(time_end,y(end,:)');
         otherwise
             error('interpolation_method method implemented')
     end

### 2017-01-23 matlab/time_marching_BF.m

@@ -97,10 +97,11 @@ if io.plot_power_figure
 end
 
 if dat.PbID==12
-    amplitude_norm = compute_L2norm(@(x) npar.phi_exact(x,time_end),u(1:npar.n,end))
-    [Cend] = assemble_source(npar.C_exact,time_end);
+%     amplitude_norm = compute_L2norm(@(x) npar.phi_exact(x,time_end),u(1:npar.n,end))
+    amplitude_norm = abs(npar.phi_exact(npar.x_dofs(2),time_end) - u(2,end))
+%     [Cend] = assemble_source(npar.C_exact,time_end);
 
-    amplitude_norm = norm(Cend-u(npar.n+1:end,end),2)
+%     amplitude_norm = norm(Cend-u(npar.n+1:end,end),2)
 end
     
 % output

### 2017-01-23 matlab/time_marching_IQS.m

@@ -115,7 +115,8 @@ end
 if dat.PbID==12
     if (strcmp(func2str(FUNHANDLE),'solve_IQS_PC_diffusion_an_prec') || ...
     strcmp(func2str(FUNHANDLE),'solve_IQS_PC_diffusion_elim_prec'))
-        amplitude_norm = compute_L2norm(@(x) npar.phi_exact(x,time_end),         u_shape(1:npar.n,end))
+%         amplitude_norm = compute_L2norm(@(x) npar.phi_exact(x,time_end),         u_shape(1:npar.n,end))
+        amplitude_norm = abs(npar.phi_exact(npar.x_dofs(2),time_end) - u_shape(2,end))
     else
         amplitude_norm = compute_L2norm(@(x) npar.phi_exact(x,time_end),X(1,end)*u_shape(1:npar.n,end))
     end

### 2017-01-31 matlab/funprke_lag_interp.m
Getting third order now

@@ -1,0 +1,37 @@
+function dydt=funprke_lag_interp(time,y)
+
+% global
+global dat npar
+
+tn = dat.ode.tn;
+
+% reactivity coef.  interpolation
+rho_MGT = 0.0;
+beff_MGT = 0.0;
+for i=1:length(tn)
+    rho = dat.ode.rho_MGT(i);
+    beff = dat.ode.beff_MGT(i);
+    for j=1:length(tn)
+        if i~=j
+            rho = rho.*(time-tn(j))/(tn(i)-tn(j));
+            beff = beff.*(time-tn(j))/(tn(i)-tn(j));
+        end
+    end
+    rho_MGT = rho_MGT + rho;
+    beff_MGT = beff_MGT + beff;
+end
+
+% compute PRKE matrix at current time
+J=[(rho_MGT-beff_MGT)   dat.lambda ; ...
+    beff_MGT           -dat.lambda ];
+
+S    = assemble_source(   dat.source_phi,time);
+q_MGT    = (npar.phi_adj' * S)/npar.K0;
+
+q = [q_MGT; 0];
+
+% provide dydt to ode solver
+dydt=J*y + q;
+
+return
+end

### 2017-01-31 matlab/funprke_lin_interp.m

@@ -1,7 +1,7 @@
 function dydt=funprke_lin_interp(time,y)
 
 % global
-global dat
+global dat npar
 
 time_beg = dat.ode.time_beg;
 time_end = dat.ode.time_end;
@@ -19,6 +19,9 @@ q_MGT    = dat.ode.q_MGT_beg    * w1 + dat.ode.q_MGT_end    * w2;
 % compute PRKE matrix at current time
 J=[(rho_MGT-beff_MGT)   dat.lambda ; ...
     beff_MGT           -dat.lambda ];
+
+S    = assemble_source(   dat.source_phi,time);
+q_MGT    = (npar.phi_adj' * S)/npar.K0;
 
 q = [q_MGT; 0];
 

### 2017-01-31 matlab/kinetics_driver_conv.m

@@ -12,7 +12,7 @@ global npar io
 % verbose/output parameters
 io.console_print         = false;
 io.plot_transient_figure = false;
-io.plot_power_figure     = true;
+io.plot_power_figure     = false;
 io.make_movie            = false;
 io.save_flux             = false;
 io.print_progress        = false;
@@ -21,7 +21,7 @@ io.figID = 99;
 npar.set_bc_last=true;
 
 % select problem
-pbID=12; refinements=1;
+pbID=10; refinements=1;
 problem_init(pbID,refinements);
 
 % compute fundamental eigenmode
@@ -69,7 +69,8 @@ dt = t_end./ntimes;
 %   2      = linear interpolation of shape
 %   3      = cubic interpolation of shape
 %   4      = quadratic interpolation of shape
-npar.iqs_prke_interpolation_method=2;
+npar.iqs_prke_interpolation_method=5;
+npar.solve_prke_compute_rho_each_time = false;
 
 % Type of precursor interpolation for analytical elimination
 % (see solve_TD_diffusion_an_prec)
@@ -78,7 +79,7 @@ npar.an_interp_type='lagrange'; % This is just a place holder, actual definition
     
 % Number of steps in history for lagrange interpolation of precursors for
 % analytical elimination precursors (see solve_TD_diffusion_an_prec)
-npar.interpolation_order=3;
+npar.interpolation_order=1;
 
 % If this is true, precursors are recalculated using a cubic hermite
 % interpolation of shape (see solve_TD_diffusion_an_prec line 192)
@@ -97,12 +98,12 @@ npar.prec_solve_type = 'linear';
 i=0;
 % not to be used for conv. studies % i=i+1; list_runs{i}= 'brute_force_matlab';
 % i=i+1; list_runs{i}= 'brute_force';
-% i=i+1; list_runs{i}= 'brute_force_elim_prec';
-i=i+1; list_runs{i}= 'brute_force_an_prec';
+i=i+1; list_runs{i}= 'brute_force_elim_prec';
+% i=i+1; list_runs{i}= 'brute_force_an_prec';
 % i=i+1; list_runs{i}= 'iqs_an_prec';
-% i=i+1; list_runs{i}= 'iqs_elim_prec';
+i=i+1; list_runs{i}= 'iqs_elim_prec';
[…]

### 2017-01-31 matlab/kinetics_init.m

@@ -7,8 +7,6 @@ NFId = assemble_mass(dat.nusigf_d,curr_time) /npar.keff;
 if ~npar.set_bc_last
     NFId = apply_BC_mat_only(NFId,npar.add_zero_on_diagonal);
 end
-size(NFId)
-size(phi)
 C = NFId*phi/dat.lambda;
 
 % verif

### 2017-01-31 matlab/problem_init.m

@@ -12,9 +12,9 @@ dat.rod_mov.t_beg_1=0.1; dat.rod_mov.t_end_1=0.6;
 dat.rod_mov.t_beg_2=1.0; dat.rod_mov.t_end_2=2.7;
 
 % kinetic parameters
-dat.beta_tot=600e-5;
+dat.beta_tot=600e-5*0;
 dat.lambda=0.1;
-dat.invvel=1.;
+dat.invvel=1e-3;
 
 dat.source_phi = @(x,t) 0;
 dat.source_phi = @(x,t) 0;
@@ -182,24 +182,31 @@ switch problem_ID
         
         syms x t
         syms S_p(x,t) phi(x,t) C(x,t) S_c(x,t)
+        syms varphi(x,t) amp(t) dvarphi(x,t)
         
-%         syms f(t) a(x,t) A(t)
+        syms f(t) a(x,t) A(t)
 %         f(t) = (t+1)^5
 %         a(x,t) = sin(x*(1-x)*(t+1))
 %         A(t) = int(a(x,t),x,0,1)
 %         phi(x,t) = f(t)*a(x,t)/A(t);
-%         f(t) = (t+1)^0;
-%         a(x,t) = x*(1-x)*(x+t);
-%         A(t) = int(a(x,t),x,0,1);
-%         phi(x,t) = f(t)*a(x,t)/A(t);
+        f(t) = (t+1)^1;
+        a(x,t) = x*(1-x)*(x+t);
+        A(t) = int(a(x,t)*a(x,0),x,0,1);
+        phi(x,t) = f(t)*a(x,t)/A(t);
+        varphi(x,t) = a(x,t)/A(t);
+        amp(t) = f(t);
 
-        phi(x,t) = x*(1-x)*(1+t)^3;
-        C(x,t) = x*(1-x)*(1+t)^2;
+%         phi(x,t) = x*(1-x)*(1+t)^3;
+%         C(x,t) = x*(1-x)*(1+t)^2;
+        dvarphi = diff(varphi,t);
         if do_steady
             phi(x,t) = x*(1-x); % use this for steady state
         end        
         npar.phi_exact = matlabFunction(phi);
-        npar.C_exact = matlabFunction(C);
+        npar.shape_exact = matlabFunction(varphi);
+        npar.p_exact = matlabFunction(amp);
+        npar.dshape_exact = matlabFunction(dvarphi);
+%         npar.C_exact = matlabFunction(C);
 %         npar.C_exact = @(x,t) 0.*x.*t;                
[…]

### 2017-01-31 matlab/solve_IQS_PC_diffusion_elim_prec.m

@@ -19,8 +19,11 @@ X_beg=X(:,end);
 u_beg=u;
 % for clarity, I also single out the shape function at the beginning/end of the macro time step
 % IV   = assemble_mass(     dat.inv_vel ,time_end);
-z = (npar.phi_adj)'*npar.IV*u(1:npar.n,end)/npar.K0;
-shape_beg=u(1:npar.n,end)/z; dat.ode.shape_beg=shape_beg;
+for i=1:length(u(1,:))
+    z = (npar.phi_adj)'*npar.IV*u(1:npar.n,i)/npar.K0;
+    shape_old(:,i)=u(1:npar.n,i)/z; 
+end
+shape_beg = shape_old(:,end); dat.ode.shape_beg=shape_beg;
 
 for iter = 1: max_iter_iqs
     
@@ -37,7 +40,7 @@ for iter = 1: max_iter_iqs
     
     % solve for amplitude function
     if strcmpi(npar.prke_solve,'matlab')
-        [X,w,t,y] =  solve_prke_ode(X_beg,dt_macro,time_end,shape_beg,shape_end);
+        [X,dXdt,w,t,y] =  solve_prke_ode(X_beg(:,end),dt_macro,tn,shape_old(:,1:end),shape_end);
     else
         [X,dpdt,t,y] =  solve_prke_iqs(X_beg,dt_macro,time_end,shape_beg,shape_end,npar.n_micro,npar.freq_react);
     end
@@ -186,5 +189,5 @@ for iter = 1: max_iter_iqs
 end
 
 if err>=tol_iqs
-    warning('IQS did not converge in %s',mfilename);
+%     warning('IQS did not converge in %s',mfilename);
 end

### 2017-01-31 matlab/solve_IQS_diffusion_elim_prec.m

@@ -24,7 +24,7 @@ shape_end=shape_beg(:,end);
 rk=npar.rk;
 if strcmp(npar.method,'BDF')
     bdf = npar.bdf;
-    order = length(tn);
+    order = length(tn)-1;
 end
 
 % beginning of the time interval
@@ -40,11 +40,11 @@ for iter = 1: max_iter_iqs
     
     % solve for amplitude function over the entire macro-step
     if strcmpi(npar.prke_solve,'matlab')
-        [X,dXdt,w,t,y] =  solve_prke_ode(X_beg(:,end),dt_macro,time_end,shape_beg(:,end),shape_end);
+        [X,dXdt,w,t,y] =  solve_prke_ode(X_beg(:,end),dt_macro,tn,shape_beg(:,1:end),shape_end);
     else
         [X,dpdt,t,y] =  solve_prke_iqs(X_beg(:,end),dt_macro,time_end,shape_beg(:,end),shape_end,npar.n_micro,npar.freq_react);
     end
-    
+%     X(1) = (1+time_end);
     % interpolating polynomial
     if strcmpi(npar.prke_solve,'matlab')
         if npar.int_order==3
@@ -70,8 +70,8 @@ for iter = 1: max_iter_iqs
         
         % OLD shortcut
         % OLD p=X(1);
-%         pi = (1+ti)^2;
-%         dpdti = 2*(1+ti);
+%         pi = (1+ti);
+%         dpdti = 1;
         pi = ppval(pp,ti);
 %         dpdti = ppval(pp_der,ti);
 %         tmp = funprke(ti,X);
@@ -142,6 +142,7 @@ for iter = 1: max_iter_iqs
         end
         % solve for new shape_end
         shape_end = A\rhs;
+%         shape_end = npar.shape_exact(npar.x_dofs',ti);
 %         shape_end = shape_beg(:,end);
         % finish Ci
         if strcmp(npar.method,'BDF')
@@ -156,7 +157,8 @@ for iter = 1: max_iter_iqs
     
     % save for hermite interp
     if npar.iqs_prke_interpolation_method>=3
-        dat.ode.f_end=IV\f(:,rk.s);
+%         dat.ode.f_end=IV\f(:,rk.s);
+%         dat.ode.f_end=npar.dshape_exact(npar.x_dofs',time_end);
     end
     
     % re-package as single solution vector

### 2017-01-31 matlab/solve_TD_diffusion_elim_prec.m

@@ -17,7 +17,7 @@ lambda = dat.lambda;
 rk=npar.rk;
 if strcmp(npar.method,'BDF')
     bdf = npar.bdf;
-    order = length(tn);
+    order = length(tn)-1;
 end
 
 % beginning of the time interval

### 2017-01-31 matlab/solve_prke_ode.m

@@ -1,11 +1,15 @@
-function [X,dXdt,w,t,pow] =  solve_prke_ode(X,dt_macro,time_end,shape_beg,shape_end)
+function [X,dXdt,w,t,pow] =  solve_prke_ode(X,dt_macro,tn,shape_beg,shape_end)
 
 % make the problem-data a global variable
 global dat npar
 
 % create a list of times when reactivity will be recomputed for this macro
 % time step
+time_end = tn(end);
 time_beg = time_end - dt_macro;
+if npar.iqs_prke_interpolation_method~=5 || npar.solve_prke_compute_rho_each_time
+    shape_beg = shape_beg(:,end);
+end
 
 dat.ode.time_beg = time_beg;
 dat.ode.time_end = time_end;
@@ -68,12 +72,20 @@ else
             %  shape(t) = w(1) t^2 + w(2) t + w(3)
             w = mat\rhs; w=w'; dat.ode.nw = size(w,2);
             for k=1:dat.ode.nw
-                [dat.ode.rho_MGT_beg(k),dat.ode.beff_MGT_beg(k)]=compute_prke_parameters(time_beg,w(:,k));
-                [dat.ode.rho_MGT_end(k),dat.ode.beff_MGT_end(k)]=compute_prke_parameters(time_end,w(:,k));
+                [dat.ode.rho_MGT_beg(k),dat.ode.beff_MGT_beg(k),dat.ode.q_MGT_beg(k)]=compute_prke_parameters(time_beg,w(:,k));
+                [dat.ode.rho_MGT_end(k),dat.ode.beff_MGT_end(k),dat.ode.q_MGT_end(k)]=compute_prke_parameters(time_end,w(:,k));
             end
             % call ode solver with prke function that linearly interpolates
             [t,y]=ode15s(@funprke_linhermite_interp,[time_beg time_end],X,options);
             dXdt = funprke_linhermite_interp(time_end,y(end,:)');
+        case 5 % lagrange interpolation of shape using previous shapes
+            shape = [shape_beg shape_end];
+            dat.ode.tn = tn;
+            for k=1:length(tn)
+                [dat.ode.rho_MGT(k),dat.ode.beff_MGT(k),~]=compute_prke_parameters(tn(k),shape(:,k));
+            end
+            [t,y]=ode15s(@funprke_lag_interp,[time_beg time_end],X,options);
+            w=0;
         otherwise
[…]

### 2017-01-31 matlab/time_marching_IQS.m

@@ -55,7 +55,7 @@ for it=1:ntimes
     end
     
     % solve time-dependent diffusion for flux
-    [u_shape(:,end),X(:,end),t,y] = FUNHANDLE(u_shape(:,end-od:end-1),X(:,end-od:end-1),dt,tn(it-od+2:it+1));
+    [u_shape(:,end),X(:,end),t,y] = FUNHANDLE(u_shape(:,end-od:end-1),X(:,end-od:end-1),dt,tn(it-od+1:it+1));
     
     % update data for hermite interpolation
         dat.ode.f_beg=dat.ode.f_end;
@@ -115,8 +115,8 @@ end
 if dat.PbID==12
     if (strcmp(func2str(FUNHANDLE),'solve_IQS_PC_diffusion_an_prec') || ...
     strcmp(func2str(FUNHANDLE),'solve_IQS_PC_diffusion_elim_prec'))
-%         amplitude_norm = compute_L2norm(@(x) npar.phi_exact(x,time_end),         u_shape(1:npar.n,end))
-        amplitude_norm = abs(npar.phi_exact(npar.x_dofs(2),time_end) - u_shape(2,end))
+        amplitude_norm = compute_L2norm(@(x) npar.phi_exact(x,time_end),         u_shape(1:npar.n,end))
+%         amplitude_norm = abs(npar.phi_exact(npar.x_dofs(2),time_end) - u_shape(2,end))
     else
         amplitude_norm = compute_L2norm(@(x) npar.phi_exact(x,time_end),X(1,end)*u_shape(1:npar.n,end))
     end

### 2017-02-03 doc/M&C_2017/FullPaper/MC2017_FullPaper.tex
Intial draft of full paper

--- 2016-10-30 doc/M&C_2017/MC2017_Summary.tex
@@
 \documentclass{anstrans}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \title{Multiphysics Techniques for Neutron Diffusion Using Improved Quasi-Static Method}
-\title{Multiphysics Core Dynamics Simulation Using the Improved Quasi-Static Method}
-\author{Zachary M.~Prince, Jean C.~Ragusa}
-
-\institute{
-Department of Nuclear Engineering, Texas A\&M University, College Station, TX
-}
-
-\email{zachmprince@tamu.edu \and jean.ragusa@tamu.edu}
 
@@
 
+%%%% changes from the original 'anstrans' class
+\usepackage{fancyhdr} % allows headers and footers
+
+\renewcommand\headrule{} % remove underline in the header
+\setcounter{secnumdepth}{1}
+\renewcommand{\thesection}{\Roman{section}.}
+\renewcommand{\thesubsection}{\arabic{subsection}.}
+\renewcommand{\thesubsubsection}{\Alph{subsubsection}.}
+\makeatletter
+\renewcommand*{\@seccntformat}[1]{\csname the#1\endcsname\hspace{1mm}}
+\makeatother
+
+%%%% Header
+\pagestyle{fancy}
+\fancyhf{}
+\fancyhead[L]{\fontsize{9}{9} \itshape
+M\&C 2017 - International Conference on Mathematics \& Computational Methods Applied to Nuclear Science \& Engineering,
+\\ Jeju, Korea, April 16-20, 2017, on USB (2017)
+}
+
+%%%% Maketitle
+\title{Multiphysics Core Dynamics Simulation Using the Improved Quasi-Static Method}
+\author{Zachary M.~Prince, Jean C.~Ragusa}
+
+\institute{Department of Nuclear Engineering, Texas A\&M University, College Station, TX}
+
+\email{zachmprince@tamu.edu \and jean.ragusa@tamu.edu}
+
+% tikz stuff
+\usetikzlibrary{shapes,arrows,positioning}
+\pgfdeclarelayer{background}
+\pgfdeclarelayer{foreground}
+\pgfsetlayers{background,main,foreground}
+\tikzstyle{greenblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=green!20, ultra thick, minimum width=60mm, minimum height=15mm,]
+\tikzstyle{blueblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=blue!20, ultra thick, minimum width=60mm, minimum height=15mm]
+\tikzstyle{reddiamond}=[diamond, draw, align=center, top color=white, bottom color=red!20, ultra thick, minimum width=60mm, aspect=2]
+
+%%%% Abstract
 \begin{document}
+\vspace*{-42pt}
+\begin{strip}
+\centering{\parbox{153mm}{{\bf Abstract} \itshape - 
+This template provides the standard format and styles of the full papers to be submitted to M\&C 2017.
+It is basically the same as the extended summary template except that the abstract is given and the section numbers are specified.
+}\par}
+\vspace*{14pt}
+\end{strip}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The purpose of this summary is to introduce several techniques in dealing with multi-physics feedback for transient neutron diffusion calculations.
-In dynamic simulation, the neutron temporal distribution in a nuclear core can be strongly influenced by non-neutronics variables, e.g., temperature.The improved quasi-static method (IQS) is an effective technique for simulating the temporal behavior of the neutron flux in reactors.
+The purpose of this paper is to introduce several techniques in dealing with multi-physics feedback for transient neutron diffusion calculations.
+In dynamic simulations, the neutron temporal distribution in a nuclear core can be strongly influenced by non-neutronics variables, e.g., temperature.The improved quasi-static method (IQS) is an effective technique for simulating the temporal behavior of the neutron flux in reactors.
 Here, we present a study combining the IQS method with multi-physics solution techniques for coupled transient calculations.
@@
 The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates.
-IQS has largely been applied to neutron kinetics, without thermal-hydraulic feedback.
-This summary discusses the application of multi-physics feedback with IQS and analyzes performance with temperature feedback benchmark problems
+IQS has largely been applied to neutron kinetics, without other feedback mechanisms.
+This paper presents the application of multi-physics feedback with IQS and analyzes performance with temperature feedback benchmark problems.
+
+The majority of publicized applications of IQS involve purely neutron kinetics, with several exceptions.
+Devooght et.
+al.
+discusses the application of the generalized quasi-static method with thermal hydraulic feedback within a newton iteration scheme in \cite{Devooght_1984}.
+This implementation is not ideally efficient for IQS because, theoretically, the shape is less variant in time than temperature feedback and temperature is more closely coupled with amplitude.
+Other references such as \cite{Banfield_2013} and \cite{KIKO3D_2003} apply IQS to multi-physics feedback problems, but are ambiguous on their coupling treatment.
+
+This paper presents a new approach to multi-physics feedback with IQS involving inclusion of feedback mechanisms in the quasi-static process.
+The intention of this implementation is to further optimize solution accuracy with computational effort.
+In order to evaluate the performance of the methodology, it will be tested with two temperature feedback problems: the LRA benchmark and an example from Transient Reactor Testing Facility (TREAT) reactor.
+The performance is quantified by comparing accuracies with traditional implementations of neutron dynamics and IQS.
 
@@
 \frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \varphi^{g'} - \left(L^g - \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right)\varphi^g + \boxed{\frac{1}{p}}S_{d}^g
-\label{eq:flux}
+\label{eq:shape}
 \ee 
@@
 \frac{dC_i}{dt} = \boxed{p}\sum_{g=1}^G P_{d,i}^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
-\label{eq:precursor}
+\label{eq:prec}
 \ee
@@
 \ee
+
+Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three dimensions, and even more so when using the transport equations in lieu of the diffusion equations.
+Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself,  thus allowing for larger time step sizes in updating the shape.
+The PRKE equations form a small ODE system and can be solved using a much smaller time step size.
+In transients where the shape varies much less than the flux, IQS can thus be very computationally effective.
+The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.
+
+\begin{figure}[!htbp]
+\includegraphics[width=\linewidth]{IQS_visualization.jpg}
+\caption{IQS method visualization}
+\label{fig:iqsviz}
+\end{figure}
+
+It is important to note that the PRKE parameters are evaluated at each macro step and linearly interpolated for the PRKE evaluation.
+In order to preserve the error convergence of high order discretization schemes for shape, higher order interpolation of the parameters is required.
 
@@
 Since there is a nonlinear coupling between all these components, each temperature step is iterated until amplitude has converged and the macro step is iterated until shape has converged.
+\fig{fig:proc} further visualizes this process as a programming structure.
+
+\begin{figure}[!htpb]
+\centering
+\resizebox{\linewidth}{!}{%
+\begin{tikzpicture}[every node/.style = {font=\normalsize}]
+
+\node[greenblock](p1) at (0,0) {Save Old Solution};
+\node[greenblock](p2) at (0,-2.5) {Save Current Solution};
+\node[greenblock](p3) at (0,-5) {Interpolate Solution};
+\node[blueblock](p4) at (0,-7.5) {Solve Temperature};
+\node[greenblock] (p5) at (0,-10) {Update PRKE Params};
+\node[greenblock] (p6) at (0,-12.5) {Solve PRKE};
+\node[reddiamond] (check1) at (0,-17.5) {Check for \\ convergence ($p$)};
+\node[reddiamond] (check2) at (7,-12) {If $t=t_{n+1}$};
+\node[greenblock] (p7) at (8,-6) {Shape Solve};
+\node[reddiamond] (check3) at (8,-2.5) {Check for \\ convergence ($\varphi$)};
+%\node [above =0.1mm of p1] {\large{IQS Solve:}};
+%\node [above =0.1mm of p5] {\large{Multi-physics:}};
+
+%\tikzback{p1}{p1}{p4}{p4}{bk1}
+%\tikzback{p5}{p5}{p5}{p5}{bk2}
+
+\draw[->,ultra thick](p1.south) -- node[right] {$t=t_{n+1}$}(p2.north);
+\draw[->,ultra thick](p2.south) -- node[right] {$t=t_{n}+\Delta t/3$}(p3.north);
+\draw[->,ultra thick](p3.south) -- (p4.north);
+\draw [->,ultra thick] (p4.south)-- (p5.north);
+\draw[->,ultra thick](p5.south) -- (p6.north);
+\draw[->,ultra thick](p6.south) -- (check1.north);
+\draw[->,ultra thick](check1.west) -- node[above,sloped] {\large{no}} (-5,-17.5) |-  (p4.west);
+\draw[->,ultra thick](check1.east) -| node[right] {\large{yes}} (check2.south);
+\draw[->,ultra thick](check2.west) -- node[above,sloped] {\large{no}\small \qq $t=t+\Delta t/3$} (4,-5) -- (p3.east);
+\draw[->,ultra thick](check2.north) --  (8,-10) -- node[right] {\large{yes}} (p7.south);
+\draw[->,ultra thick](p7.north) -- (check3.south);
+\draw[->,ultra thick](check3.west) -- node[above] {no} (p2.east);
+\draw[->,ultra thick](check3.north) -- node[right] {yes} (8,0) -- node[above] {$t_n=t_{n+1}$} (p1.east);
+
+\end{tikzpicture}}
+\caption{Visualization of shape iteration and temperature update process for IQS}
+\label{fig:proc}
+\end{figure}
 
@@
 
-To test the multiple time scale implementation of IQS, the LRA benchmark was chosen for reference.
+To test the multiple time scale implementation of IQS, the LRA benchmark was chosen for reference and a TREAT example was employed.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{LRA Benchmark}
+
 The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.
 It is a super prompt-critical transient.
-Three different evaluation techniques were applied: implicit discretization of the flux equation (Brute force), IQS, and IQS-PC.
+The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL).
+The spacial discretization was performed using continuous finite element method with first order Lagrangian basis functions.
+The mesh consisted of blocks 11X11 with five uniform refinements, totaling 165,165 elements and 124,609 nodes.
+Three different temporal techniques were applied: implicit discretization of the flux equation (Brute force), IQS, and IQS-PC.
 Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{LRA Temperature Feedback}
+\subsubsection{LRA Temperature Feedback}
+\label{sec:LRA_T}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{LRA Convergence Results}
+\subsubsection{LRA Results}
 
 \fig{fig:lra_profile} shows the baseline power and temperature profile for the LRA benchmark.
+The baseline results are compared to the results achieved by Sutton and Aviles in \cite{Sutton_1996} and presented in \tbl{tab:base}.
 The relative difference in the magnitude of the peak power ($t\approx1.44 s$) from the baseline was used for error comparison.
@@
 The dashed lines correspond to brute force at different flux step sizes, while the IQS macro step size is kept constant.
+\\
 
@@
 \end{figure}
+
+\begin{table}[!htbp]
+\begin{center}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{|l|cc|}
+\hline
+Calculation  &  Baseline & Sutton (Spandex 1936) \\
+\hline
+No. of Mesh 			& 3872 		& 1936 \\
+Eigenvalue 				& 0.99637	& 0.99637 \\
+No. of Time Steps 		& 6000 		& 23,890 \\
+Time to Peak Power (s) 	& 1.441 	& 1.441 \\
+Peak Power (W/cm$^3$) 	& 5456 		& 5461 \\
+\hline
+\end{tabular}
+}
+\end{center}
+\caption{LRA baseline verification}
+\label{tab:base}
+\end{table}
 
@@
 This convergence makes sense because temperature can only be so accurate before the error in shape is dominating.
+Table \ref{tab:ndiff_lra} shows the run time results for the implicit discretization calculations.
+Tables \ref{tab:iqs_lra} and \ref{tab:iqspc_lra} present the IQS run-times with various numbers of temperature updates.
+These run-times are based on total alive time of the execution where the diffusion evaluation is distributed over 24 processors.
+\\
+
+\begin{table}[!htbp]
+\begin{center}
+%\resizebox{\linewidth}{!}{
+\begin{tabular}{|l|l|cc|}
+\hline
+Run  &  $\Delta t$ & Error & Runtime (hr) \\
+\hline
+1	& 4.0e-3	& 1.407e-2 	& 4.11	\\
+2	& 2.0e-3	& 3.174e-3 	& 5.68 	\\
+3 	& 1.0e-3 	& 7.690e-4 	& 5.55	\\
+4 	& 5.0e-4 	& 1.892e-4 	& 13.57	\\
+5 	& 2.5e-4	& 4.590e-5 	& 23.81	\\
+\hline
+\end{tabular}
+%}
+\end{center}
+\caption{Implicit discretization run time results}
+\label{tab:ndiff_lra}
+\end{table}
+
+\begin{table}[!htbp]
+\begin{center}
+%\resizebox{\linewidth}{!}{
+\begin{tabular}{|l|l|ccc|}
+\hline
+	&  Temperature 	&  		& Runtime 	& \% Increase	\\
+Run	&  Updates 	& Error & (hr)		& in Runtime$^*$\\
+\hline
+1	& 1		& 2.612e-3 	&  `&	\\
+2	& 2		& 9.893e-4 	& 	& 	\\
+3 	& 4 	& 5.796e-4 	& 	&	\\
+4 	& 8 	& 4.772e-4 	& 	&	\\
+5 	& 16	& 4.516e-4 	& 	&	\\
+\hline
+\end{tabular}
+%}
+\end{center}
+\caption{Traditional IQS run time results with $\Delta t = 0.004$}
+\label{tab:iqs_lra}
+\end{table}
+
+\begin{table}[!htbp]
+\begin{center}
+%\resizebox{\linewidth}{!}{
+\begin{tabular}{|l|l|ccc|}
+\hline
+	&  Temperature 	&  		& Runtime 	& \% Increase	\\
+Run	&  Updates 	& Error & (hr)		& in Runtime$^*$\\
+\hline
+1	& 1		& 3.488e-3 	& 6.24 `&	\\
+2	& 2		& 1.349e-3 	& 6.18	& 	\\
+3 	& 4 	& 9.161e-4 	& 7.05	&	\\
+4 	& 8 	& 8.052e-4 	& 7.72	&	\\
+5 	& 16	& 7.905e-4 	& 10.82	&	\\
+\hline
+\end{tabular}
+%}
+\end{center}
+\caption{IQS PC run time results with $\Delta t = 0.004$}
+\label{tab:iqspc_lra}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Transient-15 TREAT Example}
+
+Transient 15 is a small core TREAT experiment that was modeled by the Rattlesnake group at INL for demonstrating their model's fidelity to TREAT’s feedback mechanisms.
+The goal of the simulations is to obtain a power peak whose position and magnitude that sufficiently matches experimental results.
+The model involves an 11 energy group diffusion approximation and is discretized into 355,712 hexahedral continuous finite elements totaling 4,109,523 degrees of freedom.
+The three second transient involves a linear ramp decrease in boron content throughout the fuel region.
+\fig{fig:Tran15} shows a visualization of the flux profile within the core, hidden is the massive amount of graphite surrounding the core.
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=\linewidth]{Tran15_core2.png}
+\caption{Transient 15 core power profile at peak power}
+\label{fig:Tran15}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Transient-15 Temperature Feedback}
+
+The Transient-15 model uses a adiabatic temperature feedback mechanism, similar to the one explored by the LRA.
+\eqt{eq:temp2} describes the heat up of the fuel.
+It is very similar, except the specific heat is now dependent on temperature is described by \eqt{eq:cp}.
+The temperature evaluation is identical to the one described in LRA section, except a Newton iteration process is employed to resolve the nonlinearity from the specific heat term.
+The feedback to the cross-sections are applied using linear interpolation of tabular data provided by INL.
+
+\be
+\rho c_p(T) \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
+\label{eq:temp2}
+\ee
+
+\be
+c_p = -5.8219e-10T^3 - 4.3694e-7T^2 + 2.8369e-3T -1.009e-2
+\label{eq:cp}
+\ee
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Transient-15 Results}
+
+\fig{fig:Tran15_profile} shows the baseline power and temperature profile for the Transient-15 example.
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=\linewidth]{Tran15_profile.png}
+\caption{Transient 15 total power and average temperature profile during transient}
+\label{fig:Tran15_profile}
+\end{figure}
 
@@
 \bibliography{references_IQS}
+
 \end{document}

### 2017-02-04 doc/Thesis/data/abstract.tex
Initial thesis draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2016.  
%
%  Author: Sean Zachary Roberson
%	 Version 3.16.09
%  Last updated 9/12/2016
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           ABSTRACT 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter*{ABSTRACT}
\addcontentsline{toc}{chapter}{ABSTRACT} % Needs to be set to part, so the TOC doesnt add 'CHAPTER ' prefix in the TOC.

\pagestyle{plain} % No headers, just page numbers
\pagenumbering{roman} % Roman numerals
\setcounter{page}{2}

\indent Going to finish this when I finish the rest of the paper.

\pagebreak{}

### 2017-02-04 doc/Thesis/data/acknowledgements.tex
Initial thesis draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2016.
%
%  Author: Sean Zachary Roberson
%	 Version 3.16.09
%  Last updated 9/12/2016
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           ACKNOWLEDGMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{ACKNOWLEDGMENTS}
\addcontentsline{toc}{chapter}{ACKNOWLEDGMENTS}  % Needs to be set to part, so the TOC doesnt add 'CHAPTER ' prefix in the TOC.

\indent This section is also optional, limited to four pages. It must follow the Dedication Page (or Abstract, if no Dedication). If listing preliminary pages in Table of Contents, include Acknowledgments. Heading (\MakeUppercase{Acknowledgments}) is bold if major headings are bold. It should be in same type size and style as text. So does vertical spacing, paragraph style, and margins. Also, ensure that the spelling of ``acknowledgments'' matches throughout the text and the table of contents.

I would like to thank the Texas A\&M University Office of Graduate and Professional Studies to allow me to construct this \LaTeX\ thesis template. Special thanks to JaeCee Crawford, Amy Motquin, Ashley Schmitt, Rachel Krolczyk, and Roberta Caton for carefully reviewing this material.  % use A\&M instead of A$\&$M, not use $A\&M$ as well, the last one won't be bold.

\pagebreak{}

### 2017-02-04 doc/Thesis/data/appendix2.tex
Initial thesis draft

--- 2017-02-04 doc/Thesis/data/acknowledgements.tex
@@
 %  New template code for TAMU Theses and Dissertations starting Fall 2016.
+%
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                           ACKNOWLEDGMENTS
+%%                           APPENDIX B
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter*{ACKNOWLEDGMENTS}
-\addcontentsline{toc}{chapter}{ACKNOWLEDGMENTS}  % Needs to be set to part, so the TOC doesnt add 'CHAPTER ' prefix in the TOC.
 
-\indent This section is also optional, limited to four pages.
-It must follow the Dedication Page (or Abstract, if no Dedication).
-If listing preliminary pages in Table of Contents, include Acknowledgments.
-Heading (\MakeUppercase{Acknowledgments}) is bold if major headings are bold.
-It should be in same type size and style as text.
-So does vertical spacing, paragraph style, and margins.
-Also, ensure that the spelling of ``acknowledgments'' matches throughout the text and the table of contents.
+\chapter{\uppercase {A Second Appendix Whose Title Is Much Longer Than The First}}
 
-I would like to thank the Texas A\&M University Office of Graduate and Professional Studies to allow me to construct this \LaTeX\ thesis template.
-Special thanks to JaeCee Crawford, Amy Motquin, Ashley Schmitt, Rachel Krolczyk, and Roberta Caton for carefully reviewing this material.
-% use A\&M instead of A$\&$M, not use $A\&M$ as well, the last one won't be bold.
+Text for the Appendix follows.
+
+\section{Appendix Section}
+
+\section{Second Appendix Section}
 

### 2017-02-04 doc/Thesis/data/dedication.tex
Initial thesis draft

--- 2017-02-04 doc/Thesis/data/appendix2.tex
@@
 %  New template code for TAMU Theses and Dissertations starting Fall 2016.  
-%
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                           APPENDIX B
+%%                           DEDICATION
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter*{DEDICATION}
+\addcontentsline{toc}{chapter}{DEDICATION}  % Needs to be set to part, so the TOC doesnt add 'CHAPTER ' prefix in the TOC.
 
-\chapter{\uppercase {A Second Appendix Whose Title Is Much Longer Than The First}}
-
-Text for the Appendix follows.
-
-\section{Appendix Section}
-
-\section{Second Appendix Section}
+\begin{center}
+\vspace*{\fill}
+To my mother, my father, my grandfather, and my grandmother. To see what happens with multiple lines, I extend this next part into a second line.
+\vspace*{\fill}
+\end{center}
 

### 2017-02-04 doc/Thesis/data/section1.tex
Initial thesis draft

--- 2017-01-03 doc/Thesis_proposal/thesis_proposal.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[12pt]{scrartcl} 
-%########################### Preferences #################################
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%  New template code for TAMU Theses and Dissertations starting Fall 2016.
+%
+%  Author: Sean Zachary Roberson 
+%	 Version 3.08.16
+%  Last updated 8/19/2016
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\usepackage{vmargin}
-\usepackage{color}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amstext}
-\usepackage{amsbsy}
-\usepackage{graphicx} 
-\usepackage{subcaption}
-\usepackage{hyperref}
-\usepackage{fancyhdr}
-\usepackage{listings}
-\usepackage[numbers,sort&compress]{natbib}[2010/09/13]
-\usepackage{titlesec}
-\usepackage{tikz}
-\usepackage{enumitem}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                           SECTION I
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%  new definitions
-\renewcommand{\div}{\bs{\nabla}\! \cdot \!}
-\newcommand{\grad}{\bs{\nabla}}
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert_{L^2}}
-\newcommand{\rattlesnake}{Rattlesnake }
-% extra space
-\newcommand{\qq}{\quad\quad}
-% common reference commands
-\newcommand{\eqt}[1]{Equation~\ref{#1}}                     % equation
-\newcommand{\fig}[1]{Figure~\ref{#1}}                      % figure
-\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table
-\newcommand{\sct}[1]{Section~\ref{#1}}                   % section
-\newcommand{\app}[1]{Appendix~\ref{#1}}                   % appendix
+\pagestyle{plain} % No headers, just page numbers
+\pagenumbering{arabic} % Arabic numerals
+\setcounter{page}{1}
 
-\newcommand{\bs}[1]{\mathbf{#1}}
-\newcommand{\dd}{\mathrm{d}}
-\newcommand{\keff}{k_\textit{eff}}
-
-\newcommand{\be}{\begin{equation}}
-\newcommand{\ee}{\end{equation}}
-\newcommand{\vn}{\vec{n}}
-\newcommand{\vel}{\vec{\mathrm{v}}}
-\newcommand{\adj}{\Phi^\dagger_0}
-\newcommand{\tcr}[1]{\textcolor{red}{#1}}
-
-% ********* Caption Layout ************
-%\usepackage{ccaption} % allows special formating of the captions
-%\captionnamefont{\bf\footnotesize\sffamily} % defines the font of the caption name (e.g. Figure: or Table:)
-%\captiontitlefont{\footnotesize\sffamily} % defines the font of the caption text (same as above, but not bold)
-%\setlength{\abovecaptionskip}{0mm} %lowers the distance of captions to the figure
-\setlength\parindent{0pt}
-\setlength{\oddsidemargin}{1in}
-\setlength{\evensidemargin}{1in}
-\setlength{\textwidth}{6.5in}
-% ********* Header and Footer **********
-% This is something to play with forever. I use here the advanced settings of the KOMA script
-\pagestyle{fancy}
-\fancyhf{}
-\lhead{Prince 2016 - Thesis Proposal}
-\rfoot{}
-\cfoot{\thepage}
-
-% ********* Title **********
-\title{\normalfont \LARGE Improved Quasi-Static Methods for Time-Dependent Neutron Diffusion and Implementation in Rattlesnake}
-\subtitle{\normalfont Thesis Proposal}
-\author{ \normalsize
-  \textbf{Zachary M. Prince} \\ \vspace{-3mm}
- \normalsize Department of Nuclear Engineering \\ \vspace{-3mm}
- \normalsize Texas A\&M University\\ \vspace{-3mm}
- \normalsize College Station, TX, 77843-3133\\
- \normalsize \href{mailto:zachmprince@tamu.edu}{zachmprince@tamu.edu}
-}
-
-\hypersetup{
-  colorlinks=true,
-  citecolor=black,
-  linkcolor=black,
-  urlcolor=blue,
-}
-
-\titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}
-\titleformat{\subsection}{\normalfont\large\bfseries}{\thesubsection}{1em}{}
-
-%################ End Preferences, Begin Document #####################
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% set relative directory for figures. can be changed at any time later with \renewcommand
-\newcommand{\FiguresDir}{./figs}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\maketitle
-\pagenumbering{arabic}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{sect:introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{\uppercase{Introduction}}
 
@@
 \\
+
+\section{Neutron Diffusion Solution Process}
 
@@
 
+\subsection{Point Reactor Kinetics Equation}
+
+A common solution process for neutron kinetics is the evaluation of the point reactor kinetics equation (PRKE).
+The derivation of this equation involves factorization flux into a space-dependent shape and time-dependent amplitude.
+This factorization immediately creates a large assumption that the spacial variance of the flux is time-independent.
+However, this takes very little computational effort to evaluate because the multitudinous variables from the spacial discretization only needs to be evaluated on the initial time step.
+The rest of the time steps only evaluated one variable: amplitude.
+
+\subsection{Improved Quasi-Static Method}
+
 IQS is a spatial kinetics method that involves factorizing the flux solution into space- and time-dependent components \cite{Ott_1966,Dulla2008}.
@@
 However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the shape may not require an update at the same frequency as the amplitude function, but only on larger macro-time steps.
-The proposal for this research contains three sections: \sct{sect:back} explains the current theory of IQS and its implementation, \sct{sect:research} describes the proposed research in IQS, and \sct{sect:results} shows preliminary results of the proposed research.
+\\
+\indent Since shape and amplitude share a independent variable, they are temporally coupled.
+Additionally, the resulting system of equations is heavily nonlinear.
+The are a number of published strategies that evaluate this system, but a in depth comparison of the techniques has not been found.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{IQS Theory and Status}
-\label{sect:back}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Objective}
 
-This section describes the theory of IQS and the current status of its development.
-The theory and status are broken up into three sections: \sct{sect:theory} analytically derives the method, \sct{sect:iter} and \sct{sect:pc} present two solution techniques and describes their application and development.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Theory}
-\label{sect:theory}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-IQS involves factorizing the flux from \eqt{eq:flux} into time-dependent amplitude ($p$) and space- and time- dependent shape ($\varphi$).
-The resulting equation is the shape-diffusion equation with precursors represented by Equations~\ref{eq:shape}~and~\ref{eq:prec}.
-
-\begin{subequations}
-\begin{align}
-\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\
-& -  \left( -\div D^g \grad  + \Sigma_r^g + \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right) \varphi^g + \boxed{\frac{1}{p}}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
-\label{eq:shape}
-\end{align}
-\be
-\frac{dC_i}{dt} = \frac{\beta_i}{\keff}\boxed{p} \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
-\label{eq:prec}
-\ee
-\end{subequations}
-
-We note that the time-dependent shape equation is similar to the time-dependent flux equation, with the following modifications:
-\begin{enumerate}
-\item The shape equation contains an additional term equivalent to a removal cross section,  $ \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$.
-\item The delayed neutron source term is divided by $p$.
-\item The system of equations is now nonlinear due to the factorization.
-\item An equation is needed to obtain the amplitude $p$.
-\end{enumerate}
-
-To derive the amplitude equation, the shape/precursors equations are weighted by a space-dependent function and integrated over the phase-space.
-The weight function is typically the adjoint flux $\phi^{*g}$, which can be proven to minimize truncation error \cite{duderstadt1976nuclear}.
-The final expressions are given below:
-\begin{subequations}
-\be
-\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
-\label{eq:p}
-\ee
-\be
-\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
-\label{eq:c}
-\ee
-\end{subequations}
-This equation is also known as the point reactor kinetics equation (PRKE); where the reactivity, effective delayed-neutron fraction, and delayed-neutron precursor decay constant are defined as follows:
-\begin{subequations}
-\be
-\frac{\rho-\bar{\beta}}{\Lambda} = 
-\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}
-{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-\label{eq:rmb}
-\ee
-\be
-\frac{\bar{\beta}}{\Lambda} = \sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda} = 
-\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}
-{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-\label{eq:b}
-\ee
-\be
-\bar{\lambda}_i = \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
-\label{eq:l}
-\ee 
-\end{subequations}
-The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.\\
-
-Additionally, in order to impose uniqueness on the factorization and to derive the PRKE, the following normalization condition is imposed: $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}$ \citep{Dulla2008}.
+The goal of this research is to continue the investigation and development of the improved quasi-static method.
+This goal requires the development of IQS in a easily modifiable, prototype-like code, as well as a robust FEM framework.
 \\
-
-Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three dimensions, and even more so when using the transport equations in lieu of the diffusion equations.
-Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself,  thus allowing for larger time step sizes in updating the shape.
-The PRKE equations form a small ODE system and can be solved using a much smaller time step size.
-In transients where the shape varies much less than the flux, IQS can thus be very computationally effective.
-The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.
-
-\begin{figure}[!htbp]
-\includegraphics[width=\linewidth]{\FiguresDir/IQS_visualization.jpg}
-\caption{IQS method visualization}
-\label{fig:iqsviz}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Iterative Solution Techniques}
-\label{sect:iter}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-As we noted in \sct{sect:theory}, shape-PRKE equations are a nonlinear system and thus may be solved in a iterative manner.
-Each macro time step can be iterated so the best shape is used to compute power at the micro time steps.
-Sissaoui et al.
-from \cite{Sissaoui_1995}, Koclas et al.
-from \cite{Koclas_1996}, Devooght et al.
-from \cite{Devooght_1984}, and Monier from \cite{Monier_diss} all use iterative techniques for their quasi-static simulations.
-They all undergo a similar process:
-\begin{itemize}
-\item[\textit{Step 1:}] Compute the PRKE parameters at the end of the macro step using the last computed shape
-\item[\textit{Step 2:}] Linearly interpolate the computed PRKE parameters over the macro step
-\item[\textit{Step 3:}] Solve the PRKE on micro steps over the entire macro step
-\item[\textit{Step 4:}] Solve the shape equation on the macro step using the computed values of $p$ and $dp/dt$.
-\item[\textit{Step 5:}] Check if the shape solution has converged:
-	\begin{itemize}
-	\item \textit{No:} Repeat the same macro time step
-	\item \textit{Yes:} Move on to the next macro time step
-	\end{itemize}
-\end{itemize}
-
-The major difference between the methods of these authors is the convergence criteria used.
-Sissaoui and Koclas \cite{Sissaoui_1995, Koclas_1996} use fixed point iteration where the criteria is the simply the normalized difference between the last two computed shapes.
-Monier in \cite{Monier_diss} also does fixed point iterations with the same criteria, except the solution is scaled by $\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n})\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n+1})\right)}$ after each iteration.
-Devooght in \cite{Devooght_1984} does a Newton-SOR iteration where the residual of the shape function evaluation is the convergence criteria and next iteration's solution is computed using Newton-Raphson method.\\
-
-These techniques are by no means an exhaustive list of the possible iteration techniques for IQS.
-Dulla et al.
-in \cite{Dulla2008} does an in depth analysis of the fixed point iteration technique most similar to Sissaoui and Koclas, involving convergence rates and solution results.
-However, no comprehensive analysis of iteration techniques exists, comparing both Newton and fixed-point convergence rates.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Predictor-Corrector version of IQS (IQS P-C)}
-\label{sect:pc}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The Predictor-Corrector (P-C) version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the solution of the coupled system of equations is different.
-In the IQS P-C version, the flux equations (not the shape equations) are solved (represented by Equations~\ref{eq:flux}~and~\ref{eq:precursor}) in order to obtain a predicted flux solution.
-This predicted flux is then converted to a shape by rescaling it as follows:
-\be
-\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
-\label{eq:rescale}
-\ee
-where the scaling factors are given by
-\be
-K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
-\ee
-\be
-K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0} \,.\right)
-\ee
-
-The PRKE parameters are then computed with this shape using Equations~\ref{eq:rmb}~-~\ref{eq:l} and interpolated over the macro step, then the PRKE ODE system is solved on the micro time scale.
-With the newly computed amplitude, the shape is rescaled into a flux and the final corrected flux is given by:
-\be
-\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1} \,.
-\ee
-
-The advantage to the predictor-corrector method is there is no iteration necessary for this method and, in turn, is much simpler and faster than the standard IQS.
-Ikeda et al.
-in \cite{Ikeda_2001} and Goluoglu et al.
-in \cite{Goluoglu_2001} both use IQS P-C for complex, three-dimensional problems.
-Their results prove IQS P-C to be impressively effective, despite the de-coupling of the system.
-Dulla et al.
-in \cite{Dulla2008} also describes an in depth comparison of IQS P-C with traditional IQS.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Proposed Research}
-\label{sect:research}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The proposed research intends to continue the investigation and development of the improved quasi-static method.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Investigation of Iteration Convergence}
-\label{sect:inv_iter}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\sct{sect:iter} describes several iterative techniques employed for IQS, with different convergence criteria and solution scaling.
-A new convergence criteria, not previously attempted, can be applied.
-This criteria involves checking that the uniqueness criteria described in \sct{sect:theory} is preserved.
-The error of a certain iteration is described by \eqt{eq:eiqs}.
-Where $\varphi^{g,n}$ corresponds to the solution at the most recent time step and $\varphi^{g,0}$ corresponds to the initial solution.
-This technique is intended to ensure that the solution retains the purity of the analytical derivation.
-
-\be
-Error_{IQS}=\left|\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,n}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,0}\right)}-1\right|
-\label{eq:eiqs}
-\ee
-
-This research intends to investigate the performance of the previously implemented iterative techniques and compare them to this newly implemented technique.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Investigation of Various Time Discretization Schemes}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A vital part of the verification and validation for IQS is analyzing error convergence.
-Since any time discretization scheme is capable of being applied to the shape equation of IQS, it is important to investigate IQS's performance to a variety of these schemes.
-There is lack of literature that applies IQS to schemes other than implicit Euler; higher order schemes are never rigorously tested.
-This research intends to apply a variety of schemes, including implicit Euler, Crank-Nicholson, backward difference formula (BDF), and diagonally implicit Runge-Kutta (DIRK), to test stability and error convergence.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Application of Time Adaptation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-IQS aims at reducing the time discretization error in the flux solution by splitting the flux into an amplitude (highly resolved at a micro time scale) and a shape (whose time-dependence is weaker than that of the flux itself).
-Thus, by construction, the IQS approach may employ larger time-step sizes for comparable temporal error.
-Further enhancements can be gained by using time adaptation (or time step control) in order to increase or reduce the macro time step size for the shape evaluation, depending on error estimates.
-A step-doubling technique is chosen as the time adaptation technique.
-The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps ($\varphi^g_{\Delta t/2}$).
-Note: $\varphi$ is changed to $\phi$ for implicit discretization and IQS P-C.
-\\
-
-The relative error is computed as follows:
-\be
-e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2} - \sum_{g=1}^G\varphi^g_{\Delta t}}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2}},\norm{\sum_{g=1}^G\varphi^g_{\Delta t}}\right)}
-\label{eq:edt2}
-\ee
-If the error is smaller than the user-specified tolerance, $e_{tol}$, the time step is accepted. In addition, a new time step size is estimated as follows:
-\be
-\Delta t_{new} = S \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}}
-\label{eq:dt2}
-\ee
-Where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor.
-If the error is larger than the user-specified tolerance, the time step is rejected.
-A new time step size is estimated using  \eqt{eq:edt2} as well.
-\\
-
-To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to implicit discretization method, traditional IQS, and IQS P-C.
-Each of these methods will be applied to several diffusion problems; the number of time steps taken and the resulting error will be used to compare the methods.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Application of Temperature Feedback}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-IQS is first and foremost a nuclear reactor simulation method.
-In nuclear reactors, multiple physics affect the profile of the neutron flux.
-The most simple example of mulitphysics reactor simulations is adiabatic heat up with Doppler feedback.
-The principle of Doppler feedback is that fission in a fuel causes the material to increase temperature and induces a change in the neutronics properties.
-The material heat up is described by \eqt{eq:temp}; where $\rho$ is the material density, $c_p$ is the specific heat, $T$ is temperature, and $\kappa_f$ is the energy released per fission \cite{ANL_BPB}.
-The change in temperature of the material mainly affects the thermal macroscopic absorption cross section described by \eqt{eq:dopp} \cite{ANL_BPB}.
-
-\be
-\rho c_p \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
-\label{eq:temp}
-\ee
-
-\be
-\Sigma_a^{thermal}(\vec{r},t) = \Sigma_a^{thermal}(\vec{r},0)\left[1+\gamma\left(\sqrt{T}-\sqrt{T_0}\right)\right]
-\label{eq:dopp}
-\ee
-
-For IQS, this temperature feedback affects both the shape equation and the reactivity of the PRKE; thus, it is an additional nonlinear component to the already coupled shape-amplitude equations.
-In foresight to the application of this component, temperature is much more time dependent than the shape, but less so than the amplitude.
-Therefore, the evaluation of temperature will have its own time scale.
-A possible solution process for a problem with temperature feedback will have time three time scales portrayed in \fig{fig:time}.
-The first time scale is the shape solve, the second is the temperature evaluation as well as the computation of PRKE parameters, and the third is the PRKE scale.
-
-\begin{figure}[htpb!]
-\centering
-\begin{tikzpicture}[scale=2]
-%Shape
-\draw[] (0,2) -- (6,2) ;
-\foreach \x in  {0,6}
-\draw[shift={(\x,2)},color=black] (0pt,6pt) -- (0pt,-6pt);
-\draw[shift={(0,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
-\draw[shift={(6,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
-\node(shape) at (-.25,2) {$\varphi:$};
-
-%Temp/Params
-\draw[] (0,1) -- (6,1) ;
-\foreach \x in  {0,6}
-\draw[shift={(\x,1)},color=black] (0pt,6pt) -- (0pt,-6pt);
-\foreach \x in  {2,4}
-\draw[shift={(\x,1)},color=black] (0pt,4pt) -- (0pt,-4pt);
-\draw[shift={(0,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
-\draw[shift={(6,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
-\node(temp) at (-.75,1) {$T \rightarrow \rho, \bar{\beta}, \bar{\lambda}, \Lambda$:};
-
-% PRKE
-\draw[] (0,0) -- (6,0) ;
-\foreach \x in  {0,6}
-\draw[shift={(\x,0)},color=black] (0pt,6pt) -- (0pt,-6pt);
-\foreach \x in  {0,2,4,6}
-\draw[shift={(\x,0)},color=black] (0pt,4pt) -- (0pt,-4pt);
-\foreach \x in  {0,0.2,0.4,0.6,0.8,1,1.2,1.4,1.6,1.8,2,2.2,2.4,2.6,2.8,3,3.2,3.4,3.6,3.8,4,4.2,4.4,4.6,4.8,5,5.2,5.4,5.6,5.8,6}
-\draw[shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt);
-\draw[shift={(0,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
-\draw[shift={(6,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
-\node(prke) at (-.5,0) {$p, \frac{dp}{dt}:$};
-
-\draw (0,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 2} (2,0);
-\draw (2,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 4} (4,0);
-\draw (4,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 6} (6,0);
-\draw (0,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 1} (2,1);
-\draw (2,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 3} (4,1);
-\draw (4,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 5} (6,1);
-\draw (0,2) edge[out=15,in=165,->,thick] node[above,sloped] {Step 7} (6,2);
-
-\draw (0,2) edge[out=-135,in=135,->,thick] node[below,sloped] {$\varphi$} (0,1);
-\draw (2,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (2,0);
-\draw (2,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (2,1);
-\draw (4,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (4,0);
-\draw (4,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (4,1);
-\draw (6,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (6,0);
-\draw (6,0) edge[out=45,in=0,->,thick] node[below,sloped] {$p, \frac{dp}{dt}$} (6,2);
-
-\end{tikzpicture}
-\caption{Time scales and process of IQS with temperature feedback}
-\label{fig:time}
-\end{figure}
-
-The couplings between temperature, amplitude, and shape are all nonlinear, so iteration processes are needed for each time scale.
-The amplitude and temperature need to be iterated on the middle time scale until convergence on each temperature step.
-Then another iterative process needs to occur in the shape time scale on all three variables.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Implementation into Rattlesnake/MOOSE Framework}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A one-dimensional, single group FEM prototype will be created for IQS in order to test validation for simple problems.
-However, testing complex, real-world problems is most easily performed by implementing IQS in a FEM framework.
-The Rattlesnake radiation transport framework is a excellent option for performing these problems.
-Rattlesnake already contains a neutron diffusion method for continuous-FEM equivalent to the implicit discretization method.
-Additionally, Rattlesnake is well documented and well organized, so implementing IQS may be relatively straightforward.
-With this implementation, benchmark testing such as 16-A1 \cite{ANL_BPB}, TWIGL, LRA, C5G7-2D, etc.
-can be performed, as well as a full core simulation of the TREAT reactor.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\section{Preliminary Results}
-\label{sect:results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This section describes preliminary results of an example that tests an IQS implementation and shows its effectiveness on computation speed and accuracy.
-The example is very simple and computes quickly; it entails a one dimensional, homogeneous 400 cm slab with a heterogenous perturbation in absorption cross section.
-\fig{fig:slab} shows how the regions of the slab are divided and \tbl{tab:1Dmat} shows the initial material properties.
-Regions 2, 3, and 4 have slope perturbations at different points in time, \tbl{tab:1Dslope} shows the values of the absorption cross-section in each region at the times of interest.
-The values of $\Sigma_a$ between these times of interest are linear interpolations between the given values.
-\\
-
-\begin{figure}[!htbp]
-\begin{center}
-\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l |}
-\hline \hline \hline
-  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
-1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 4 & 4 & 1 & 1 & 1 & 1 \\
-  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
-\hline \hline \hline
-\end{tabular}
-\caption{1-D slab region identification}
-\label{fig:slab}
-\end{center}
-\end{figure}
-
-\begin{table}[!htbp]
-\begin{center}
-\caption{1-D slab material properties and problem parameters}
-\label{tab:1Dmat}
-\begin{tabular}{llllll}
-\hline
-$D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu \Sigma_f (cm^{-1})$ & $v (cm/s)$ & $\beta$ & $\lambda (s^{-1})$ \\
-\hline
-1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\
-
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[!htbp]
-\begin{center}
-\caption{1-D slab absorption cross-section at times of interest}
-\label{tab:1Dslope}
-\begin{tabular}{llllllll}
-\hline
-Region & Material Property & 0.0 s & 0.1 s & 0.6 s & 1.0 s & 1.7 s \\
-\hline
-2 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.095 & 1.095 & 1.095 \\
-3 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.09 & 1.09 & 1.1 \\
-4 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.105 & 1.105 & 1.105 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\fig{fig:power} shows the power at each macro time step as compared to the traditional brute force (full flux implicit discretization) method.
-The strong correlation between the two curves shows that IQS is consistent with a proven method for a highly transient example.
-\fig{fig:conv} shows the error convergence for implicit discretization and IQS methods.
-This plot shows that IQS has consistent convergence slope for first and second order time schemes and that IQS has a better error constant for the convergence.
-\tbl{tab:1Ddt2} shows the results for time adaptation.
-These results show that traditional IQS does not perform well with step doubling time adaptation and IQS P-C performs moderately better than implicit discretization.
-The IQS performance in this example is not exceptional because the problem is difficult.
-As seen in \fig{fig:1D_shape}, the shape undergoes significant change through the transient, which requires many shape evaluations to retain accuracy.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.49\textwidth]{\FiguresDir/1D_het_power_plot.png}
-\includegraphics[width=0.5\textwidth]{\FiguresDir/1D_het_power_plot2.png}
-\caption{Power level of 1D example}
-\label{fig:power}
-\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=5in]{\FiguresDir/1D_het_convergence.jpg}
-\caption{Error convergence comparison of 1D example}
-\label{fig:conv}
-\end{figure}
-
-\begin{table}[!htbp]
-\begin{center}
-\caption{1-D slab step doubling results}
-\label{tab:1Ddt2}
-\resizebox{\textwidth}{!}{
-\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
-\hline
-  &  & \multicolumn{3}{|c|}{Implicit Discretization} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
-\hline
-Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
-\hline
-1 &	0.05 	&	0.021596 &	10  &	32	 &	0.10084 &	12	  &	175		&	0.028019 &	10	&	33 \\
-2 &	0.01 	&	0.032864 &	21  &	80	 &	0.08030 &	41	  &	850	 	&	0.044721 &	20	&	74 \\
-3 &	0.005 	&	0.053159 &	27  &	96	 &	0.01215 &	115	  &	2220 	&	0.052095 &	25	&	96 \\
-4 & 0.001 	&	0.056546 &	56  &	188	 &	0.03242 &	848	  &	12511 	&	0.061274 &	50	&	162 \\
-5 &	0.0005 	&	0.062882 &	77  &	245	 &	0.03491 &	1718  &	25841 	&	0.061240 &	69	&	225 \\
-6 &	0.0001 	&	0.060089 &	177 &	537	 &	0.03554 &	8702  &	129985 	&	0.060824 &	159	&	480 \\
-7 &	5.0e-05	&	0.059513 &	252 &	767	 &	0.03622 &	17282 &	256463 	&	0.061078 &	224	&	680 \\
-8 &	1.0e-05	&	0.061063 &	561 &	1691 &	0.03142 &	79988 &	1104227 &	0.061901 &	501	&	1509 \\
-\hline
-\end{tabular}}
-\end{center}
-\end{table}
-
-\begin{figure}[!htbp]
-\begin{center}
-\includegraphics[width=\linewidth]{\FiguresDir/1D_shape.png}
-\caption{Flux profile of 1D slab at various times}
-\label{fig:1D_shape}
-\end{center}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\section{Goals of Proposed Research}
-\label{sect:goals}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
-\item Investigate iteration convergence for previously implemented iterative schemes and newly conceived one described in \sct{sect:inv_iter}.
-\item Verify IQS method with several low and high order time discretization schemes.
-\item Implement step doubling time adaptation with IQS and analyze its performance compared to implicit discretization of the flux equations.
-\item Apply temperature feedback with IQS and analyze its performance with multiphysics benchmarks.
-\item Implement IQS into the Rattlesnake framework at INL.
-\item Perform testing with IQS to benchmarks including TWIGL, LRA, and LMW.
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Possible Extensions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Some extensions, beyond the proposed scope of this research, have been identified as follows:
-
-\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
-\item Perform a full core simulation of TREAT with IQS.
-\item Investigate other multiphysics coupling with IQS including structural mechanics and thermal hydraulics.
-\item Investigate IQS application to self-adjoint angular flux S$_\text{N}$ (SAAF-SN).
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\bibliographystyle{abbrvnat}
-\bibliography{references_IQS}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}
+\indent The prototyping code is used to investigate time discretization and nonlinear iteration schemes.
+The prototype will also guide the development of IQS in the large FEM framework of Rattlesnake/MOOSE, where more complex examples and benchmarks can be tested.
+Development in Rattlesnake also spawns the opportunity to test IQS with multiphysics, namely temperature feedback, and time adaptation.

### 2017-02-04 doc/Thesis/data/section2.tex
Initial thesis draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2016.
%
%  Author: Sean Zachary Roberson
%	 Version 3.16.09
%  Last updated 9/12/2016
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                           SECTION II
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Improved Quasi-Static Method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
\label{sect:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

IQS involves factorizing the flux from \eqt{eq:flux} into time-dependent amplitude ($p$) and space- and time- dependent shape ($\varphi$).  The resulting equation is the shape-diffusion equation with precursors represented by Equations~\ref{eq:shape}~and~\ref{eq:prec}.

\begin{subequations}
\begin{align}
\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\ 
& -  \left( -\div D^g \grad  + \Sigma_r^g + \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right) \varphi^g + \boxed{\frac{1}{p}}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
\label{eq:shape}
\end{align}
\be
\frac{dC_i}{dt} = \frac{\beta_i}{\keff}\boxed{p} \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:prec}
\ee
\end{subequations}

We note that the time-dependent shape equation is similar to the time-dependent flux equation, with the following modifications:
\begin{enumerate}
\item The shape equation contains an additional term equivalent to a removal cross section,  $ \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$.
\item The delayed neutron source term is divided by $p$.
\item The system of equations is now nonlinear due to the factorization.
\item An equation is needed to obtain the amplitude $p$.
\end{enumerate}

To derive the amplitude equation, the shape/precursors equations are weighted by a space-dependent function and integrated over the phase-space. The weight function is typically the adjoint flux $\phi^{*g}$, which can be proven to minimize truncation error \cite{duderstadt1976nuclear}. The final expressions are given below:
\begin{subequations}
\be
\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
\label{eq:p}
\ee
\be
\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
\label{eq:c}
\ee
\end{subequations}
This equation is also known as the point reactor kinetics equation (PRKE); where the reactivity, effective delayed-neutron fraction, and delayed-neutron precursor decay constant are defined as follows:
\begin{subequations}
\be
\frac{\rho-\bar{\beta}}{\Lambda} = 
\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}
{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:rmb}
\ee
\be
\frac{\bar{\beta}}{\Lambda} = \sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda} = 
\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}
{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:b}
\ee
\be
\bar{\lambda}_i = \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
\label{eq:l}
\ee 
\end{subequations}
The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.\\

Additionally, in order to impose uniqueness on the factorization and to derive the PRKE, the following normalization condition is imposed: $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}$ \cite{Dulla2008}. \\

Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three dimensions, and even more so when using the transport equations in lieu of the diffusion equations.  Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself,  thus allowing for larger time step sizes in updating the shape. The PRKE equations form a small ODE system and can be solved using a much smaller time step size. In transients where the shape varies much less than the flux, IQS can thus be very computationally effective. The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.

\begin{figure}[!htbp]
\includegraphics[width=\linewidth]{\FiguresDir/IQS_visualization.jpg}
\caption{IQS method visualization}
\label{fig:iqsviz}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Operator Notation and Extension of IQS to Multigroup Transport Equations}
\label{sect:transport}

For simplicity, the previous section only described the IQS implementation using the neutron diffusion equation. However, deriving the IQS form for other neutron balance equations (e.g., transport, simplified transport, etc.) is very similar. To this end, we re-write the neutron conservation equations in operator form, shown in \eqt{eq:transport}. 
\be
\frac{\partial}{\partial t}\left(\frac{\Psi^g}{v^g}\right) = \sum_{g'} \left(H^{g'\to g} + P_p^{g' \to g} \right) \Psi^{g'} - L^g\Psi^g + S_d^g
\label{eq:transport}
\ee
where $\Psi^g$ is the multigroup neutron flux (angular flux in the case of transport), $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt neutron production operator, $L^g$ is the loss operator, and $S_d^g$ is the delayed neutron source operator. Using \eqt{eq:flux}, the reader may easily obtain the functional form for these operators in the case of a diffusion approximation.
%
%Factorization is defined as $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ and the resulting shape equation is defined by \eqt{eq:tshape}.  
Next, the factorization $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ is introduced, where the shape is denoted by $\psi^g$, leading to the following shape equations, \eqt{eq:tshape}:
\be
\frac{\partial}{\partial t}\left(\frac{\psi^g}{v^g}\right) = \sum_{g'} \left(H^{g' \to g} + P_p^{g' \to g}\right) \psi^{g'} - \left( L^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \psi^g + \frac{1}{p} S_d^g
\label{eq:tshape}
\ee
Finally, the PRKE parameters are defined by Equations~\ref{eq:trmb}~and~\ref{eq:tb}, where 
$\left(\Psi^{*g},f^g\right) =$\\ $ \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d^2\Omega$.
\be
\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},\sum_{g'}(H^{g' \to g}g+P_p^{g' \to g}-L^{g'}\delta_{g'g})\psi^{g'}\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
\label{eq:trmb}
\ee
\be
\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{\sum_{g=1}^G(\Psi^{*g}, \sum_{g'} P_{d,i}^{g' \to g} \psi^{g'})}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
\label{eq:tb}
\ee
where $P_{d,i}^{g' \to g}$ is the delayed-neutron operator for precursor group $i$. \\
\indent This section is simply meant to show the theoretical expandability of IQS to transport problems and its derivation in operator notation.  As stated previously, this research only applies and tests IQS with diffusion problems.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Iterative Solution Techniques}
\label{sect:iter}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

As we noted in \sct{sect:theory}, shape-PRKE equations are a nonlinear system and thus may be solved in a iterative manner.  Each macro time step can be iterated so the best shape is used to compute power at the micro time steps. Sissaoui et al. from \cite{Sissaoui_1995}, Koclas et al. from \cite{Koclas_1996}, Devooght et al. from \cite{Devooght_1984}, and Monier from \cite{Monier_diss} all use iterative techniques for their quasi-static simulations.  They all undergo a similar process:
\begin{itemize}
\item[\textit{Step 1:}] Compute the PRKE parameters at the end of the macro step using the last computed shape
\item[\textit{Step 2:}] Linearly interpolate the computed PRKE parameters over the macro step
\item[\textit{Step 3:}] Solve the PRKE on micro steps over the entire macro step
\item[\textit{Step 4:}] Solve the shape equation on the macro step using the computed values of $p$ and $dp/dt$.
\item[\textit{Step 5:}] Check if the shape solution has converged:
	\begin{itemize}
	\item \textit{No:} Repeat the same macro time step
	\item \textit{Yes:} Move on to the next macro time step
	\end{itemize}
\end{itemize}
This process can be visualized by \fig{fig:picard}.
\begin{figure}[!htpb]
\centering
\begin{tikzpicture}[every node/.style = {font=\normalsize}]

\node[greenblock](p1) at (0,0) {Solve shape equations \\ over macro time step};
\node[greenblock](p2) at (0,-2.5) {Compute PRKE parameters \\ over macro time step};
\node[greenblock](p3) at (0,-5) {Solve PRKE using micro \\ steps};
\node[greenblock](p4) at (0,-7.5) {Update $p$ and $\frac{dp}{dt}$};
\node[blueblock2] (p5) at (0,-11) {Solve other physics \\ components};
\node[reddiamond] (check) at (0,-15) {Check for \\ convergence};
\node [above =0.1mm of p1] {\large{IQS Solve:}};
\node [above =0.1mm of p5] {\large{Multi-physics:}};

\tikzback{p1}{p1}{p4}{p4}{bk1}
\tikzback{p5}{p5}{p5}{p5}{bk2}

\draw[->,ultra thick](p1.south) -- (p2.north);
\draw[->,ultra thick](p2.south) -- (p3.north);
\draw[->,ultra thick](p3.south) -- (p4.north);
\draw [->,ultra thick] (p4.south)+(0,-0.25) -- node [above] {} (bk2-n);
\draw[->,ultra thick](p5.south)+(0,-0.25) -- (check.north);
%\draw[->,ultra thick](check.west) -- (-5,-15) -- node[above,sloped] {\large{no}}(-5,0) -- node[above] {}(bk1-w);
\draw[->,ultra thick](check.west) -- node[above,sloped] {\large{no}} (-5,-15) |-  node[above] {}(bk1-w);
\draw[->,ultra thick](check.south) -- node[right] {\large{yes}} ++(0,-2);

\end{tikzpicture}
\caption{Visualization of IQS fixed-point iteration process}
\label{fig:picard}
\end{figure}

The major difference between the methods of these authors is the convergence criteria used.  Sissaoui and Koclas \cite{Sissaoui_1995, Koclas_1996} use fixed point iteration where the criteria is the simply the normalized difference between the last two computed shapes.  Monier in \cite{Monier_diss} also does fixed point iterations with the same criteria, except the solution is scaled by $\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n})\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n+1})\right)}$ after each iteration.  Devooght in \cite{Devooght_1984} does a Newton-SOR iteration where the residual of the shape function evaluation is the convergence criteria and next iteration's solution is computed using Newton-Raphson method.\\

These techniques are by no means an exhaustive list of the possible iteration techniques for IQS. Dulla et al. in \cite{Dulla2008} does an in depth analysis of the fixed point iteration technique most similar to Sissaoui and Koclas, involving convergence rates and solution results.  However, no comprehensive analysis of iteration techniques exists, comparing both Newton and fixed-point convergence rates.  The following sections describes each iteration technique investigated by this research.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Shape Convergence}

The most obvious convergence criteria is to observe the change in the shape from one iteration to the next.  Monier in \cite{Monier_diss} observes the $L^{\infty}$ norm of the shape for the convergence criteria, described in \eqt{eq:shape_Linf}.  However, any norm can be used for the criteria, so a $L^2$ norm is another possible criteria, described by \eqt{eq:shape_L2}.
\be
\frac{\text{max}\left|\varphi_{n}^{(k+1)} - \varphi_{n}^{(k)}\right|}{\text{max}\left|\varphi_{n}^{(k+1)}\right|} < \epsilon_{\varphi}
\label{eq:shape_Linf}
\ee 
\be
\frac{\norm{\varphi_{n}^{(k+1)} - \varphi_{n}^{(k)}}}{\norm{\varphi_{n}^{(k+1)}}} < \epsilon_{\varphi}
\label{eq:shape_L2}
\ee

Where $n$ is the time step, $k$ is the iteration number, and $\epsilon_{\varphi}$ is the numerical criteria provided by a user.  There is no guarantee that the shape converges, so max number of iterations is typically enforced.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Property Convergence}

Monier in \cite{Monier_diss} describes other properties, other than shape, to observe for convergence, shown in Equations \eqref{eq:rho_conv} - \eqref{eq:K_conv}.  These criteria can be added constraints to \eqt{eq:shape_Linf} or be in supplement to.
\be
\left(\frac{\rho}{\Lambda}\right)^{(k+1)} - \left(\frac{\rho}{\Lambda}\right)^{(k)} < \epsilon_{\rho}
\label{eq:rho_conv}
\ee
\be 
p_n^{k+1} - p_n^{k} < \epsilon_p
\label{eq:p_conv}
\ee
\be 
\frac{K_n^{(k+1)} - K_0}{K_0} < \epsilon_K
\label{eq:K_conv}
\ee

Where $K$ is the IQS uniqueness expression:
\be 
K_n^{(k+1)} = \sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi_n^{g,(k+1)}\right)
\label{eq:K}
\ee

$\epsilon_{\rho}$ is the reactivity convergence criteria, $\epsilon_{p}$ is the amplitude convergence criteria, and $\epsilon_{K}$ is the constraint convergence criteria.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solution Scaling}

In order to preserve the uniqueness criteria, it is beneficial to scale the shape such that the $K_n$ is constant, shown in \eqt{eq:shape_scale}.  This scaling can also be done after each iteration, to insure that the uniqueness criteria is satisfied whenever the shape is evaluated.
\be 
\varphi^{g}_n = \varphi^{g,(\text{last})}_n \frac{K_0}{K^{(\text{last})}_{n}}
\label{eq:shape_scale}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Preconditioned Jacobian-Free Newton-Krylov}

By far, the most common nonlinear system iteration method in MOOSE is Preconditioned Jacobian-Free Newton-Krylov with Generalized Minimal RESidual method (GMRES) as the linear system solver.  This section only descirbes the methods' application to IQS; a very detailed description of PJFNK is transcribed by Knoll in \cite{PJFNK_Knoll}. Essentially, the IQS system of equations can be described by \eqt{eq:IQSsys}.  Where $A$ is a matrix operator, $\varphi$ is the solution vector, and $F$ is the forcing vector.
\be
A(\varphi(p))\varphi=F(\varphi,p,t)
\label{eq:IQSsys}
\ee
Applying the residual based Newton-method yields:
\be
J\delta\varphi = -R(\varphi,p)
\ee 
Where $J$ is the Jacobian matrix defined as $J_{ij}=\partial R_i/\partial\varphi_j$, $\delta\varphi$ is the error in the iterating solution, and $R$ is the residual vector defined as $A\varphi-F$ (most methods simply define the residual instead of evaluating this expression).  For IQS, a residual evaluation entails a PRKE parameter evaluation and the PRKE evaluation over the entire macro-step.  Applying a preconditioner $P$ yields:
\be 
(JP^{-1})(P\delta\varphi)=-R(\varphi,p)
\ee
This resulting system can be split into two systems, represented by Equations~\eqref{eq:PJFNK_1}~and~\eqref{eq:PJFNK_2}.
\be 
(JP^{-1})w = -R(\varphi,p)
\label{eq:PJFNK_1}
\ee
\be 
\delta\varphi = P^{-1}w
\label{eq:PJFNK_2}
\ee
Solving \eqt{eq:PJFNK_1} requires the evaluation of Jacobian operator, which is done each GMRES iteration in two steps:
\begin{enumerate}
\item Approximately solve the preconditioner system for $y$: $Py=w$
\item Perform matrix-free Jacobian operation: $Jy\approx[R(\varphi+\epsilon y,p')-R(\varphi,p)]/\epsilon$
\end{enumerate}
This Jacobian operation is a finite difference approach.  $\epsilon$ is a perturbation scalar and $p'$ is the amplitude computed with PRKE parameters calculated from the perturbed shape.  For IQS, the PRKE and its parameters must be evaluated for the perturbed and original system, although the unperturbed residual is stored from the beginning of the PJFNK iteration.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Predictor-Corrector version of IQS (IQS P-C)}
\label{sect:pc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Predictor-Corrector (P-C) version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the solution of the coupled system of equations is different.  In the IQS P-C version, the flux equations (not the shape equations) are solved (represented by Equations~\ref{eq:flux}~and~\ref{eq:precursor}) in order to obtain a predicted flux solution. This predicted flux is then converted to a shape by rescaling it as follows:
\be
\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
\label{eq:rescale}
\ee
where the scaling factors are given by
\be
K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
\ee
\be
K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
\ee

The PRKE parameters are then computed with this shape using Equations~\ref{eq:rmb}~-~\ref{eq:l} and interpolated over the macro step, then the PRKE ODE system is solved on the micro time scale.  With the newly computed amplitude, the shape is rescaled into a flux and the final corrected flux is given by:
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1} \,.
\ee

The advantage to the predictor-corrector method is there is no iteration necessary for this method and, in turn, is much simpler and faster than the standard IQS.  Ikeda et al. in \cite{Ikeda_2001} and Goluoglu et al. in \cite{Goluoglu_2001} both use IQS P-C for complex, three-dimensional problems.  Their results prove IQS P-C to be impressively effective, despite the de-coupling of the system.  Dulla et al. in \cite{Dulla2008} also describes an in depth comparison of IQS P-C with traditional IQS.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Temperature Feedback}
\label{sect:mp}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

IQS is first and foremost a nuclear reactor simulation method.  In nuclear reactors, multiple physics affect the profile of the neutron flux.  The most simple example of mulitphysics reactor simulations is adiabatic heat up with Doppler feedback. The principle of Doppler feedback is that fission in a fuel causes the material to increase temperature and induces a change in the neutronics properties.  The material heat up is described by \eqt{eq:temp}; where $\rho$ is the material density, $c_p$ is the specific heat, $T$ is temperature, and $\kappa_f$ is the energy released per fission \cite{ANL_BPB}. The change in temperature of the material mainly affects the thermal macroscopic absorption cross section described by \eqt{eq:dopp} \cite{ANL_BPB}.

\be
\rho c_p \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
\label{eq:temp}
\ee

\be
\Sigma_a^{thermal}(\vec{r},t) = \Sigma_a^{thermal}(\vec{r},0)\left[1+\gamma\left(\sqrt{T}-\sqrt{T_0}\right)\right]
\label{eq:dopp}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Temperature Evaluation}

Temperature evaluation (\eqt{eq:temp}) is quite similar to the evaluation of the delayed neutron precursors from \sct{sect:dnp}. A typical implicit solver would simply use the interpolated flux at end of the temperature time step for the right hand side of the equation, a theta-method discretization.  However, IQS has much more information about the profile of the flux along the time step because of the micro-step amplitude evaluation.  Therefore, it is possible to solve for temperature using a semi-analytical approach, shown by \eqt{eq:temp_an}.
\be
T^{n+1} = T^n + \frac{\kappa_f}{\rho c_p} \left(a_2 \varphi^{n+1} + a_1 \varphi^{n}\right)
\label{eq:temp_an}
\ee
Where $n$ corresponds to the beginning of the temperature step.  $a_1$ and $a_2$ are integration coefficients defined by \eqt{eq:a1} and \eqt{eq:a2}.  Any interpolation of the amplitude along the micro steps is possible for the integration, this application uses piece-wise linear.
\be
a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)p(t')dt'
\label{eq:a1}
\ee
\be
a_2 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)p(t')dt'
\label{eq:a2}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Intermediate Time Scale}

For IQS, this temperature feedback affects both the shape equation and the reactivity of the PRKE; thus, it is an additional nonlinear component to the already coupled shape-amplitude equations. In foresight to the application of this component, temperature is much more time dependent than the shape, but less so than the amplitude.  Therefore, the evaluation of temperature will have its own time scale.  A possible solution process for a problem with temperature feedback will have time three time scales portrayed in \fig{fig:time}.  The first time scale is the shape solve, the second is the temperature evaluation as well as the computation of PRKE parameters, and the third is the PRKE scale.  It is important to note that the number of time steps in each scale is arbitrary and meant only for visual purposes.

\begin{figure}[htpb!]
\centering
\begin{tikzpicture}[scale=2]
%Shape
\draw[] (0,2) -- (6,2) ;
\foreach \x in  {0,6}
\draw[shift={(\x,2)},color=black] (0pt,6pt) -- (0pt,-6pt);
\draw[shift={(0,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(shape) at (-.25,2) {$\varphi:$};

%Temp/Params
\draw[] (0,1) -- (6,1) ;
\foreach \x in  {0,6}
\draw[shift={(\x,1)},color=black] (0pt,6pt) -- (0pt,-6pt);
\foreach \x in  {2,4}
\draw[shift={(\x,1)},color=black] (0pt,4pt) -- (0pt,-4pt);
\draw[shift={(0,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(temp) at (-.75,1) {$T \rightarrow \rho, \bar{\beta}, \bar{\lambda}, \Lambda$:};

% PRKE
\draw[] (0,0) -- (6,0) ;
\foreach \x in  {0,6}
\draw[shift={(\x,0)},color=black] (0pt,6pt) -- (0pt,-6pt);
\foreach \x in  {0,2,4,6}
\draw[shift={(\x,0)},color=black] (0pt,4pt) -- (0pt,-4pt);
\foreach \x in  {0,0.2,0.4,0.6,0.8,1,1.2,1.4,1.6,1.8,2,2.2,2.4,2.6,2.8,3,3.2,3.4,3.6,3.8,4,4.2,4.4,4.6,4.8,5,5.2,5.4,5.6,5.8,6}
\draw[shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt);
\draw[shift={(0,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(prke) at (-.5,0) {$p, \frac{dp}{dt}:$};

\draw (0,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 2} (2,0);
\draw (2,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 4} (4,0);
\draw (4,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 6} (6,0);
\draw (0,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 1} (2,1);
\draw (2,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 3} (4,1);
\draw (4,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 5} (6,1);
\draw (0,2) edge[out=15,in=165,->,thick] node[above,sloped] {Step 7} (6,2);

\draw (0,2) edge[out=-135,in=135,->,thick] node[below,sloped] {$\varphi$} (0,1);
\draw (2,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (2,0);
\draw (2,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (2,1);
\draw (4,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (4,0);
\draw (4,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (4,1);
\draw (6,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (6,0);
\draw (6,0) edge[out=45,in=0,->,thick] node[below,sloped] {$p, \frac{dp}{dt}$} (6,2);

\end{tikzpicture}
\caption{Time scales and process of IQS with temperature feedback}
\label{fig:time}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Programming Logic}

The couplings between temperature, amplitude, and shape are all nonlinear, so iteration processes are needed for each time scale.  The amplitude and temperature need to be iterated on the middle time scale until convergence on each temperature step.  Then another iterative process needs to occur in the shape time scale on all three variables. \fig{fig:proc} shows the programming diagram implement to execute this process. The time increment of $\Delta t/3$ for the temperature solve is arbitrary and is meant only to match \fig{fig:time}.

\begin{figure}[!htpb]
\centering
\begin{tikzpicture}[every node/.style = {font=\normalsize}]

\node[greenblock](p1) at (0,0) {Save Old Solution};
\node[greenblock](p2) at (0,-2.5) {Save Current Solution};
\node[greenblock](p3) at (0,-5) {Interpolate Solution};
\node[blueblock2](p4) at (0,-7.5) {Solve Temperature};
\node[greenblock] (p5) at (0,-10) {Update PRKE Params};
\node[greenblock] (p6) at (0,-12.5) {Solve PRKE};
\node[reddiamond] (check1) at (0,-17.5) {Check for \\ convergence ($p$)};
\node[reddiamond] (check2) at (7,-12) {If $t=t_{n+1}$};
\node[greenblock] (p7) at (8,-6) {Shape Solve};
\node[reddiamond] (check3) at (8,-2.5) {Check for \\ convergence ($\varphi$)};
%\node [above =0.1mm of p1] {\large{IQS Solve:}};
%\node [above =0.1mm of p5] {\large{Multi-physics:}};

%\tikzback{p1}{p1}{p4}{p4}{bk1}
%\tikzback{p5}{p5}{p5}{p5}{bk2}

\draw[->,ultra thick](p1.south) -- node[right] {$t=t_{n+1}$}(p2.north);
\draw[->,ultra thick](p2.south) -- node[right] {$t=t_{n}+\Delta t/3$}(p3.north);
\draw[->,ultra thick](p3.south) -- (p4.north);
\draw [->,ultra thick] (p4.south)-- (p5.north);
\draw[->,ultra thick](p5.south) -- (p6.north);
\draw[->,ultra thick](p6.south) -- (check1.north);
\draw[->,ultra thick](check1.west) -- node[above,sloped] {\large{no}} (-5,-17.5) |-  (p4.west);
\draw[->,ultra thick](check1.east) -| node[right] {\large{yes}} (check2.south);
\draw[->,ultra thick](check2.west) -- node[above,sloped] {\large{no}\small \qq $t=t+\Delta t/3$} (4,-5) -- (p3.east);
\draw[->,ultra thick](check2.north) --  (8,-10) -- node[right] {\large{yes}} (p7.south);
\draw[->,ultra thick](p7.north) -- (check3.south);
\draw[->,ultra thick](check3.west) -- node[above] {no} (p2.east);
\draw[->,ultra thick](check3.north) -- node[right] {yes} (8,0) -- node[above] {$t_n=t_{n+1}$} (p1.east);

\end{tikzpicture}
\caption{Visualization of fixed-point iteration and temperature update process for IQS}
\label{fig:proc}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Time Discretization Schemes}
\label{sect:dt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A vital part of the verification and validation for IQS is analyzing error convergence.  Since any time discretization scheme is capable of being applied to the shape equation of IQS, it is important to investigate IQS's performance to a variety of these schemes.  There is lack of literature that applies IQS to schemes other than implicit Euler; higher order schemes are never rigorously tested.  This research intends to apply a variety of schemes, including implicit Euler, Crank-Nicholson, backward difference formula (BDF), and diagonally implicit Runge-Kutta (DIRK), to test stability and error convergence.  For brevity in the following sections, the shape-diffusion equation is represented by a general operator notation, described by \eqt{eq:shape_mat}.
\be 
IV\frac{\partial \varphi}{\partial t} = A\varphi+b
\label{eq:shape_mat}
\ee
Where $IV$ is the inverse velocity operator, $A$ contains all the operations on $\varphi$ from the left-hand-side of \eqt{eq:shape}, and $b$ is the source from the precursors.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Theta Method Time Discretization}
\label{sect:theta}

A fairly simple way to evaluate the shape equation is to employ the $\theta$-scheme ($0\le\theta\le1$, explicit when $\theta=0$, implicit when $\theta=1$, and Crank-Nicholson when $\theta=1/2$) \cite{Ferziger}.  Generally, if there is a function $u$ whose governing equation is $\frac{du}{dt}=f(u,t)$, then the $\theta$-discretization is:
\be
\frac{u^{n+1}-u^n}{\Delta t}=(1-\theta)f(u^n,t) + \theta f(u^{n+1},t) \,.
\ee
Where $n$ is the previous time step and $n+1$ is the time step being evaluated. Applying this to \eqt{eq:shape_mat} yields:
\be
\varphi_{n+1} = (IV-\Delta t A_{n+1})^{-1}\left[\Delta t(1-\theta)(A_n\varphi_n+b_n) + \Delta t\theta b_{n+1} + IV\varphi_n\right]
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Backward Difference Formulas}

An extension of implicit discretization is the backward difference formulas (BDF) \cite{Gear:2007}.  BDF's can increase the order of error convergence by interpolating solutions from previous time time steps. The general formula for BDF is discribed by \eqt{eq:bdf}.
\be 
\varphi_{n+1} = (IV-\Delta t A_{n+1})^{-1}\left[IV\sum_{j=0}^{k-1}\alpha_{jk}\varphi_{n-j} +
\Delta t b_{n+1}\right]
\label{eq:bdf}
\ee
Where $k$ is the order of the error convergence and $\alpha_{ij}$ are coefficients chosen such that the temporal truncation error is minimized.  The benifit of using BDF is that any order of convergence can be applied and are all unconditionally stable, so the IQS approximation can easily be validated for high order discretization.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Singly-Diagonally-Implicit Runge-Kutta Method}

Singly-Diagonally-Implicit Runge-Kutta Method (SDIRK) is a powerful discretization method that involves solving the linear system in stages to reach a high order solution \cite{SDIRK}.  Generally, the method can be depicted by using a system where $dy/dt=f(t,y)$ and \eqt{eq:sdirk}.
\be 
y_{n+1} = y_n + \Delta t\sum_{i=1}^s b_ik_i
\label{eq:sdirk}
\ee
Where,
\be 
k_i = f(t_n+c_i\Delta t, y_n+\Delta t \sum_{j=1}^{i} a_{ij}k_j)
\ee
The coefficients $a_{ij}$, $b_i$, and $c_i$ can be represented by a Butcher tableau:
\begin{center}
\begin{tabular}{c|cccc}
$c_1$ & $a_{11}$ & &  & \\
'$c_2$ & $a_{21}$ & $a_{22}$ & & \\
$\vdots$ & $\vdots$ & $\vdots$ & $\ddots$ & \\
$c_s$ & $a_{s1}$ & $a_{s2}$ & $\ldots$ & $a_{ss}$\\
\hline
&$b_1$ & $b_2$ & $\ldots$ & $b_s$
\end{tabular}
\end{center}
An example used extensively in this research is SDIRK33, which is a third-order method with three stages.
\begin{center}
\begin{tabular}{c|ccc}
$\lambda$ & $\lambda$ & & \\
$\frac{1}{2}(1+\lambda)$ & $\frac{1}{2}(1-\lambda)$ & $\lambda$ & \\
$1$ & $\frac{1}{4}(-1+16\lambda-6\lambda^2)$ & $\frac{1}{4}(5-20\lambda+6\lambda^2)$ & $\lambda$\\
\hline
 & $\frac{1}{4}(-1+16\lambda-6\lambda^2)$ & $\frac{1}{4}(5-20\lambda+6\lambda^2)$ & $\lambda$
\end{tabular}
\end{center}
Where $\lambda\approx0.4358665215$ satisfies $1-9\lambda+18\lambda^2-6\lambda^3=0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Adaptation}

IQS aims at reducing the time discretization error in the flux solution by splitting the flux into an amplitude (highly resolved at a micro time scale) and a shape (whose time-dependence is weaker than that of the flux itself). Thus, by construction, the IQS approach may employ larger time-step sizes for comparable temporal error. Further enhancements can be gained by using time adaptation (or time step control) in order to increase or reduce the macro time step size for the shape evaluation, depending on error estimates. A step-doubling technique is chosen as the time adaptation technique \cite{NumC}. The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps ($\varphi^g_{\Delta t/2}$). Note: $\varphi$ is changed to $\phi$ for implicit discretization and IQS P-C.

The relative error is computed as follows:
\be
e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2} - \sum_{g=1}^G\varphi^g_{\Delta t}}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2}},\norm{\sum_{g=1}^G\varphi^g_{\Delta t}}\right)}
\label{eq:edt2}
\ee
If the error is smaller than the user-specified tolerance, $e_{tol}$, the time step is accepted. In addition, a new time step size is estimated as follows:
\be
\Delta t_{new} = S \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}}
\label{eq:dt2}
\ee
Where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor. If the error is larger than the user-specified tolerance, the time step is rejected. A new time step size is estimated using  \eqt{eq:edt2} as well. This process can be visualized by Figs.~\ref{fig:dt2_1}~and~\ref{fig:dt2_2}.  Where a step involves a full convergence of shape, amplitude, and any multiphysics on the respective time step.

To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to implicit discretization method, traditional IQS, and IQS P-C.  Each of these methods will be applied to several diffusion problems; the number of time steps taken and the resulting error will be used to compare the methods.

\begin{figure}[htpb!]
\centering
\begin{tikzpicture}[scale=1.5]
\draw[] (0,2) -- (6,2) ;
\foreach \x in  {0,3,6}
\draw[shift={(\x,2)},color=black] (0pt,6pt) -- (0pt,-6pt);
\draw[shift={(0,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(3,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1/2}$};
\draw[shift={(6,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\draw (0,2) edge[out=45,in=135,->,thick,blue] node[above,sloped] {\tcb{Step 1}} (6,2);
\draw (0,2) edge[out=25,in=155,->,thick,red] node[above,sloped] {\tcr{Step 2}} (3,2);
\draw (3,2) edge[out=25,in=155,->,thick,red] node[above,sloped] {\tcr{Step 3}} (6,2);
\node[anchor=west](shape) at (6.1,2.5) {$\tcb{\phi_{\Delta t}^g}$};
\node[anchor=west](shape) at (6.1,2) {$\tcr{\phi_{\Delta t/2}^g}$};

\end{tikzpicture}
\caption{Visualization of step doubling process on time-line}
\label{fig:dt2_1}
\end{figure}

\begin{figure}[!htpb]
\centering
\begin{tikzpicture}[every node/.style = {font=\large},scale=0.75]

\node[blueblock](p1) at (-6,3) {Step 1} ;
\node[redblock](p2) at (-8,0) {Step 2} ;
\node[redblock](p3) at (-4,0) {Step 3} ;
\node[purpleblock](p4) at (0,0) {Compute \\ $e_n$};
\node[purpleblock](p5) at (4,0) {Compute  \\ $\Delta t_{new}$};
\node[greendiamond](p6) at (8,0) {$e_n < e_{tol}$};
\node[orangeblock](p7) at (0,-3) {Restore \\ Solutions};

\draw[->,thick](p1.east) -| (p4.north);
\draw[->,thick](p2.east) -- (p3.west);
\draw[->,thick](p3.east) -- (p4.west);
\draw[->,thick](p4.east) -- (p5.west);
\draw[->,thick](p5.east) -- (p6.west);
\draw[->,thick](p6.south) |- node[right] {no} (p7.east);
\draw[->,thick](p6.east) -- node[above] {yes} (11,0);
\draw[->,thick](p7.west) -| (-10,0) -- (p2.west);
\draw[->,thick](p7.west) -| (-10,0) |- (p1.west);

\end{tikzpicture}
\caption{Visualization of step doubling process with coding logic}
\label{fig:dt2_2}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Delayed Neutron Precursor Updates}
\label{sect:dnp}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section presents the time-integration method used to solve coupled flux/shape and precursor equations, represented by Equations~\eqref{eq:flux}/\eqref{eq:shape}~and~\eqref{eq:precursor}/\eqref{eq:prec}. First, we note that we could keep this system of two time-dependent equations and solve it as a coupled system. However, this is unnecessary and a memory expensive endeavor because the precursor equation is only an ODE and not a PDE. Instead, one may discretize in time the shape equation, which typically requires the knowledge of the precursor concentrations at the end of the time step. This precursor value is taken from the solution, numerical or analytical, of the precursors ODE.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Precursor Evaluation Using Theta Method Time Discretization}

A common precursor evaluation technique is using the theta method, described in \sct{sect:theta}. Applying this to \eqt{eq:prec}:
\be
\frac{C^{n+1}-C^n}{\Delta t}=(1-\theta)\beta S_f^np^n-(1-\theta)\lambda C^n + \theta\beta S_f^{n+1}p^{n+1}-\theta\lambda C^{n+1}
\ee
Where $S_f$ is the fission source equivalent for shape ($S_f^n=(\nu\Sigma_f)^n\varphi^n$). Rearranging to solve for the precursor at the end of the time step yields
\be
C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta}{1+\theta\Delta t\lambda}S_f^n p^n +  \frac{\theta\Delta t \beta}{1+\theta\Delta t\lambda}S_f^{n+1} p^{n+1}
\ee
Reporting this value of $C^{n+1}$, one can solve for the shape $\varphi^{n+1}$ as a function of $\varphi^n$ and $C^n$ (and $p^n$, $p^{n+1}$, $dp/dt|_n$ and  $dp/dt|_{n+1}$).
Once $\varphi^{n+1}$ has been determined, $C^{n+1}$ is updated.  Applying this technique to implicit discretization is done by changing the definition the fission source ($S_f^n=(\nu\Sigma_f)^n\varphi^n$) and eliminating all $p$ terms.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Analytical Precursor Integration}

Another technique to evaluating the precursor equation is to use a exponential operator and integrate the time derivative analytically.  Applying this operation to \eqt{eq:prec} yields:
\be
C^{n+1} =  C^n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
\label{eq:prec_an}
\ee
Again, this can be applied to implicit discretization by altering the definition of the fission source and eliminating $p$.  Because $S_f$ is not known continuously over the time step, the integration can be done using any schemet (Riemann, trapezoid, Simpson's, etc.). However, there is a very accurate representation of $p(t)$ over the macro step from the PRKE solve. In order to utilize this information, another possibility is to interpolate $S_f$ linearly over the macro step.  Such that:
\be
S_f(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}S_f^n  + \frac{t-t_n}{t_{n+1}-t_n}S_f^{n+1}  \quad t_n \le t \le t_{n+1}
\ee
Applying this to \eqt{eq:prec_an} yields:
\be
C^{n+1} = C^n e^{-\lambda \Delta t} + \left(\hat{a}_2 S_f^{n+1}+\hat{a}_1 S_f^n\right)\beta
\ee
With integration coefficients defined as:
\begin{align}
&\hat{a}_1= \int_{t_n}^{t_{n+1}}\frac{t_{n+1}-t'}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
&\hat{a}_2 = \int_{t_n}^{t_{n+1}}\frac{t'-t_n}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt'
\end{align}
The amplitude $(p)$ is included in the integration coefficient because it has been highly accurately calculated in the micro step scheme, so a piecewise interpolation (linear, cubic, etc.) between those points can be done to maximize accuracy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rattlesnake Implementation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Rattlesnake is a MOOSE-based application developed INL specific to solving radiation transport problems with multiphysics capabilities.  MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like Rattlesnake.   At the heart of Rattlesnake is the action system, which provides a means to consolidate the MOOSE input syntax (which can be quite larger for multigroup transport simulations), so that a user does not have to define every kernel, variable, etc., used in the problem. Rather, the user inputs an equation description 
(e.g., Diffusion, S$_n$, P$_n$, etc.) and a solution method (e.g., SAAF, LS, CFEM, DFEM, etc.), and the action system will incorporate all the necessary physics involved (kernels, boundary conditions, postprocessor, etc.).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Executioner}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

An IQS excecutioner was created to implement the IQS method. The IQS executioner derives from the Transient executioner in MOOSE.  The IQS executioner contains a loop over micro time steps that computes the PRKE solution and then passes $p$ and $\frac{dp}{dt}$ for the Transient executioner to evaluate the shape equation at each macro step.  The PRKE solve is performed with a user specified option of backward-Euler, Crank-Nicholson, or SDIRK33.   The IQS executioner also supplements Transient Picard iteration process by adding its own error criteria:
\be
Error_{IQS}=\left|\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,n}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,0}\right)}-1\right|
\label{eq:eiqs}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Action System}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The IQS implementation mostly require the specific IQS executioner, described above. However, IQS additional changes are in the Rattlesnake action system in order to support the IQS execution.  First, changes needed to be made in order to evaluate the shape equation.  The shape equation, after some manipulation, is very similar to the time-dependent governing laws that Rattlesnake already solves. Using multigroup diffusion as an example again, we show the various kernels employed and highlight the new or modified kernels.

\begin{align}
\frac{\partial}{\partial t}\left(\frac{\varphi^g}{v^g}\right)=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
& - \underbrace{\frac{1}{v^g} \boxed{\overbrace{\frac{1}{p}\frac{dp}{dt}}^{From Executioner}}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
\end{align}

To enable Rattlesnake to solve this equation, an IQS removal kernel was created to evaluate $\sum_{g=1}^G\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.  Also, the precursor kernel was modified to include the $\frac{1}{p}$ term. Finally, the precursor auxkernel that evaluates \eqt{eq:prec} using the analytical integration method described in \sct{sect:dnp}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{PRKE coefficients}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In order to evaluate the PRKE coefficients, defined by Equations~\ref{eq:rmb}~-~\ref{eq:l}, four postprocessors were created.  The parameter calculations were separated by $\frac{\bar{\beta}_i}{\Lambda}$ numerator, $\bar{\lambda}_i$ numerator/denominator, $\frac{\rho-\bar{\beta}}{\Lambda}/\frac{\bar{\beta}}{\Lambda}$ denominator, and $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator. The first three are relatively simple, only relying on material properties and solution quantities, then computing the elemental integral.  The $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator requires the use of the MOOSE \texttt{save\_in} feature. This feature saves the residual from a calculated kernel or boundary contribution in the shape evaluation to an auxiliary variable.  The postprocessor then computes the inner product of this variable and the initial adjoint solution.  After each of these postprocessors are evaluated, a user object pulls together all the values and performs the numerator/denominator divisions.  The resulting values are then passed to the executioner for the PRKE solve.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Other Action Systems}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For simplicity, IQS implementation has only been described for CFEM diffusion.  However, Rattlesnake has other action systems capable of transient simulation, where IQS can be implemented and be effective.  One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equations~\eqref{eq:flux}~and~\eqref{eq:shape}.  However, in the derivation for IQS, this term is unaffected between shape and flux evaluation.  So saving the residual for this diffusion kernel in the \texttt{save\_in} variable is the only alteration to this action for IQS to function.\\
Additional action systems involve transport, but it is evident from \sct{sect:transport} that IQS implementation in these is straightforward as well.  The main differences between a diffusion implementation and a transport implementation are outlined below:
\begin{enumerate}
\item The form of the operators in the shape equations is different, but Rattlesnake has already implemented all the kernels necessary to represent these operators.  So no change is necessary to these kernels is necessary for IQS.  Additionally, the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ is the same, so both action systems can use the same kernel.
\item The PRKE parameters also change because of the operators.  For the $(\rho-\bar{\beta})$ parameter, the same post-processor can be used with the \texttt{save\_in} functionality in MOOSE.  The post-processor for $\bar{\beta}_i$ must be re-written for transport, but takes a very similar form as diffusion.
\end{enumerate}

IQS will be made available to any action system capable of transient simulation, which includes:
\begin{itemize}
\item CFEM-Diffusion
\item DFEM-Diffusion
\item SAAF-CFEM-SN
\item SAAF-CFEM-PN
\item LS-CFEM-SN
\item DFEM-SN
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Predictor-Corrector Modification}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In order to preserve the already implemented standard version of IQS, an option in the IQS executioner was created to specify which method is desired.  Because the diffusion solve is flux instead of shape, when predictor-corrector option is specified, the IQS removal kernel ($\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$) and the modified precursor kernel are bypassed, while all the postprocessors are still executed.  However, it is difficult to rescale the flux to shape before the PRKE parameter postprocessor are executed.  So the parameters are computed using the full flux, but amplitude is space independent and comes out of the integrals.  As seen in parameter definitions, when shape is replaced with flux, the amplitude comes out of the integral and cancels out.  So the conversion of the predicted flux in  \eqt{eq:rescale} to shape is unnecessary if the corrected flux is solved with \eqt{eq:phicorr2}.  After obtaining the corrected flux, the precursors are re-evaluated using a \texttt{EXEC\_LINEAR} statement.
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}} p_{n+1}
\label{eq:phicorr2}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Input}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The input file for IQS is very similar to the current transient diffusion input file.  The IQS input has a different executioner type and parameters.  The executioner type is simply IQS and input parameters include the number of micro time steps per macro step, the IQS error tolerance, and the initial power.  The Rattlesnake transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $\keff$, which is present in the transient input deck.  However, IQS also requires an initial evaluation of the adjoint flux, for the weighting function.  So another input file and multi-app transfer was made for the adjoint calculation.  \vspace{3mm}

Below is the syntax for the executioner block for an IQS input file.  The \texttt{predictor\_corrector} logical determines whether to do IQS P-C or regular IQS.  The \texttt{IQS\_error\_tol} is the tolerance for the IQS error represented by \eqt{eq:eiqs} and will at most \texttt{picard\_max\_its} iterations until convergence.  The \texttt{prke\_scheme} defines the time discretization used for the PRKE solution. Where \texttt{RK} is SDIRK33, \texttt{CN} is Crank-Nicholson, and \texttt{IE} is implicit-Euler.
\begin{lstlisting}
[Executioner]
  type = IQS
  predictor_corrector = true/false
  picard_max_its = 5
  ...
  n_micro = 10000
  IQS_error_tol = 1e-7
  prke_scheme = 'RK'
[]
\end{lstlisting}

Since IQS needs to use the adjoint solution for the PRKE parameter evaluation, auxiliary variables need to be created for each group in the input file.  Below is an example of their definition: 
\begin{lstlisting}
[AuxVariables]
  [./adjoint_flux_g0] 
    family = LAGRANGE
    order = FIRST
  [../]
  [./adjoint_flux_g1]
    family = LAGRANGE
    order = FIRST
  [../]
  ...
[]
\end{lstlisting}

Below is the syntax for the multi-app block to perform forward and adjoint steady-state evaluations:
\begin{lstlisting}
[MultiApps]
 [./initial_solve]
   type = FullSolveMultiApp
   execute_on = initial
   input_files = initial.i
 [../]
 [./adjoint_solve]
   type = FullSolveMultiApp
   execute_on = initial
   input_files = adjoint.i
 [../]
[]
\end{lstlisting}

Below is the syntax for the Transfer block to copy the initial and adjoint solutions from the multi-apps.
\begin{lstlisting}
[Transfers]
 [./copy_solution]
   type = TransportSystemVariableTransfer
   direction = from_multiapp
   multi_app = initial_solve
   execute_on = initial
   from_transport_system = diff
   to_transport_system = diff
 [../]
 [./copy_adjoint]
   type = MultiAppVariableTransfer
   execute_on = initial
   direction = from_multiapp
   multi_app = adjoint_solve
   from_variables = 'sflux_g0 sflux_g1 ...'
   to_variables = 'adjoint_flux_g0 adjoint_flux_g1 ...'
 [../]
 [./copy_eigenvalue]
   type = EigenvalueTransfer
   execute_on = initial
   direction = from_multiapp
   multi_app = initial_solve
 [../]
[]
\end{lstlisting}

### 2017-02-04 doc/Thesis/data/section3.tex
Initial thesis draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2016.
%
%  Author: Sean Zachary Roberson
%	 Version 3.16.09
%  Last updated 9/12/2016
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           SECTION III
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{KINETICS EXAMPLES}

Neutron dynamics is the study of the time-dependent nature of neutrons in a reactor.  More specifically, there is no coupling of neutron behavior with other physics, like thermal hydraulic feedback.  This section describes kinetics examples that IQS is tested with and analysis of its performance.  The examples range in complexity and application.  The first is a one-dimensional problem, designed for the prototype code in MATLAB.  The next four are from the Argonne National Lab (ANL) Benchmark Problem Book (BPB), and are common problems for testing codes and developing methods \cite{ANL_BPB}.

\section{One Dimensional Prototype Problem}

This example is very simple and computes quickly; it entails a one dimensional, homogeneous 400 cm slab with a heterogenous perturbation in absorption cross section.  \fig{fig:slab} shows how the regions of the slab are divided and \tbl{tab:1Dmat} shows the initial material properties.  Regions 2, 3, and 4 have slope perturbations at different points in time, \tbl{tab:1Dslope} shows the values of the absorption cross-section in each region at the times of interest.  The values of $\Sigma_a$ between these times of interest are linear interpolations between the given values.

\begin{figure}[!htbp]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l |}
\hline \hline \hline
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 4 & 4 & 1 & 1 & 1 & 1 \\
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
\hline \hline \hline
\end{tabular}
\caption{1-D slab region identification}
\label{fig:slab}
\end{center}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{1-D slab material properties and problem parameters}
\label{tab:1Dmat}
\begin{tabular}{llllll}
\hline
$D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu \Sigma_f (cm^{-1})$ & $v (cm/s)$ & $\beta$ & $\lambda (s^{-1})$ \\
\hline
1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\

\hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[!htbp]
\begin{center}
\caption{1-D slab absorption cross-section at times of interest}
\label{tab:1Dslope}
\begin{tabular}{llllllll}
\hline
Region & Material Property & 0.0 s & 0.1 s & 0.6 s & 1.0 s & 1.7 s \\
\hline
2 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.095 & 1.095 & 1.095 \\
3 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.09 & 1.09 & 1.1 \\
4 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.105 & 1.105 & 1.105 \\
\hline
\end{tabular}
\end{center}
\end{table}

\subsection{IQS Iteration Convergence}

\subsection{Time Step Convergence}

\pagebreak
\section{TWIGL Benchmark}

This benchmark problem originates from the Argonne National Lab Benchmark Problem Book \cite{ANL_BPB}.  It is a 2D, 2-group reactor core model with no reflector region shown in \fig{fig:TWIGL_reg} \cite{TWIGL_benchmark}.  This example is meant to be of progressive complexity from the previous example.  The transient of this reactor is very geometrically symmetrical with very little temporal shape change.  Therefore, IQS is expected to perform significantly better than the implicit discretization method.  \tbl{tab:TWIGL_mat} shows the material properties of each fuel region and the ramp perturbation of Material 1.

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.75\textwidth]{\FiguresDir/TWIGL_regions.jpg}
\caption{TWIGL benchmark problem description}
\label{fig:TWIGL_reg}
\end{center}
\end{figure}
\begin{table}[!htbp]
\begin{center}
\caption{TWIGL benchmark material properties and slope perturbation}
\label{tab:TWIGL_mat}
\begin{tabular}{llllllll}
\hline
  &  &  &  &  &  &  \multicolumn{2}{c}{$\underline{\Sigma_s (cm^{-1})} $} \\
Material & Group & $D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu\Sigma_f (cm^{-1})$ & $\chi$ & $g \rightarrow 1$ & $g \rightarrow 2$ \\
\hline
1 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01 \\
  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
2 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01  \\
  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
3 & 1 & 1.3 & 0.008 & 0.003 & 1.0 & 0.0 & 0.01  \\
  & 2 & 0.5 & 0.050 & 0.060 & 0.0 & 0.0 & 0.00  \\
\hline
  & $\nu$ & $v_1 (cm/s)$ & $v_2 (cm/s)$ & $\beta$ & $\lambda (1/s)$ &   &   \\
\hline
  & 2.43 & 1.0E7 & 2.0E5 & 0.0075 & 0.08 &   &   \\
\hline
 \multicolumn{8}{l}{\footnotesize Material 1 ramp perturbation:} \\
\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (1-0.11667t) \quad t \leq 0.2 s$} \\
\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (0.97666t) \quad t > 0.2 s$} \normalsize
\end{tabular}
\end{center}
\end{table}

\subsection{TWIGL Convergence Analysis}

Figs. \ref{fig:TWIGL_power} and \ref{fig:TWIGL_plots} show the IQS  solution as compared with the implicit discretization solution.  It is important to note the IQS shape plot is scaled differently than the Brute Force flux plot (\fig{fig:TWIGL_plots}) because the amplitude term is not included, but the gradients of colors is comparable. These plots show that IQS is consistent in more complex, higher dimensional problems in Rattlesnake. These plots also serve to illustrate that IQS has a much more accurate solution, even at a significantly larger time step than the implicit discretization. In order to demonstrate asymptotic convergence of IQS, implicit Euler (IE) and second order BDF (BDF2) were applied to the TWIGL simulation. \fig{fig:TWIGL_conv} plots the error convergence of IQS and the implicit discretization methods.  The curves show the impressive convergence of IQS for the highly transient TWIGL example. The slope indicated in the legend are the linear slope of curves on the log plot, these slopes should be similar to the order of the method (1 for IE and 2 for BDF2).  IQS shows a increased order because the PRKE is performing much of accuracy convergence and it is computed using SDIRK33, a third order method.

\begin{figure}[!htbp]
\centering
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/TWIGL_power_plot.png}
\caption{Power profile for entire transient}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/TWIGL_power_plot2.png}
\caption{Power at cusp of profile}
\end{subfigure}
\caption{Power level comparison of TWIGL Benchmark}
\label{fig:TWIGL_power}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\begin{subfigure}[!htbp]{0.4\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/ndiff_ramp_flux.jpg}
\caption{Implicit Discretization flux}
\end{subfigure}
\quad
\begin{subfigure}[!htbp]{0.4\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/iqs_ramp_shape.jpg}
\caption{IQS Shape}
\end{subfigure}
\caption{TWIGL Benchmark flux/shape comparison at $t=0.2$}
\label{fig:TWIGL_plots}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[height=3in]{\FiguresDir/TWIGL_convergence.jpg}
\caption{Error convergence comparison of TWIGL Benchmark}
\label{fig:TWIGL_conv}
\end{figure}

\subsection{TWIGL with Step Doubling Time Adaptation}

\tbl{tab:TWIGLdt2} and \fig{fig:TWIGL_power_dt2} show the results for TWIGL with time adaptation.  The results show that both IQS methods perform exceptionally well compared to implicit discretization.  It also shows that traditional IQS performed better with large $e_{tol}$, while IQS P-C was better with smaller $e_{tol}$.

\begin{table}[!htbp]
\begin{center}
\caption{TWIGL step doubling results}
\label{tab:TWIGLdt2}
\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
\hline
  &  & \multicolumn{3}{|c|}{Implicit Discretization} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
\hline
1 &	0.05    &	0.00012677 &	9   &	29  &	0.03380433 &	4   &	20   &	0.00323100 &	4  &	9   \\
2 &	0.01    &	3.5555e-05 &	11  &	35  &	0.00166991 &	5   &	40   &	0.00263068 &	5  &	12  \\
3 &	0.005   &	4.0364e-05 &	11  &	31  &	0.00886584 &	5   &	40   &	0.00160486 &	6  &	21  \\
4 &	0.001   &	0.00294822 &	33  &	122 &	0.02976305 &	5   &	36   &	1.7527e-05 &	10 &	35  \\
5 &	0.0005  &	0.00099778 &	39  &	131 &	0.00143781 &	6   &	55   &	1.4185e-05 &	16 &	74  \\
6 &	0.0001  &	0.00019510 &	78  &	236 &	0.00016175 &	8   &	65   &	6.2903e-06 &	19 &	78  \\
7 &	5.0e-05 &	0.00018372 &	112 &	342 &	6.0328e-05 &	12  &	163  &	1.5247e-06 &	24 &	92  \\
8 &	1.0e-05 &	8.0564e-05 &	263 &	794 &	7.7103e-05 &	379 &	5729 &	9.8321e-07 &	48 &	210 \\
\hline

\end{tabular}}
\end{center}
\end{table}

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\linewidth]{\FiguresDir/TWIGL_power_plot_dt2.png}
\caption{Power level comparison of TWIGL Benchmark with time adaptation}
\label{fig:TWIGL_power_dt2}
\end{figure}

\pagebreak
\section{C5G7-TD Benchmark}

The C5G7 benchmark is a MOX fueled, pressurized water reactor (PWR) minicore configuration. The C5G7 geometry depicted in Fig. \ref{fig:c5g7-2D} comprises a 2-by-2 array of UO$_2$ and MOX assemblies surrounded by moderator. In Fig. \ref{fig:c5g7-2D} vacuum and reflective boundary conditions are denoted by V and R, respectively. The fuel pins are not homogenized, each pin cell is comprised of a fuel pin, fission chamber, or guide tube surrounded by moderator as depicted in Fig. \ref{fig:c5g7-pincell}. The cladding and gap are homogenized into the fuel and are not explicitly modeled. Each assembly is made up of a regular 17-by-17 grid of pin cells.
Seven energy-group cross sections for the seven material regions are given in \cite{c5g7}. The energy group boundaries are provided in Table \ref{tab:c5g7-ebounds}; three of the seven energy groups are fast, four are thermal.

\begin{figure}[!htpb]
\begin{center}
\includegraphics[width=\linewidth]{\FiguresDir/C5G7-2D.png}
\end{center}
\caption{Geometry of the two-dimensional C5G7 benchmark problem. Boundary conditions are vacuum (V) or reflective (R).}
\label{fig:c5g7-2D}
\end{figure}

\begin{figure}[!htpb]
\begin{center}
\includegraphics[width=0.5\linewidth]{\FiguresDir/C5G7-pincell.png}
\end{center}
\caption{Geometry of a C5G7 pin-cell.}
\label{fig:c5g7-pincell}
\end{figure}

\begin{table}[!htpb]
\centering
\caption{Energy group boundaries for C5G7 MOX benchmark.}
\label{tab:c5g7-ebounds}
\begin{tabular}{c c}
\hline
Group & Upper Energy\\
\hline
 1&20 MeV  \\
 2&1 MeV  \\
 3&500 keV  \\
 4&3 eV  \\
 5&0.625 MeV  \\
 6&0.1 MeV  \\
71&0.02 MeV  \\
\hline
\end{tabular}
\end{table}

\pagebreak
\section{LMW Benchmark?}

### 2017-02-04 doc/Thesis/data/section4.tex
Initial thesis draft

--- 2016-06-15 doc/NEAMS_Report/NEAMS_Report.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[12pt]{scrartcl} 
-%########################### Preferences #################################
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%  New template code for TAMU Theses and Dissertations starting Fall 2016.
+%
+%  Author: Sean Zachary Roberson
+%	 Version 3.16.09
+%  Last updated 9/12/2016
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                           SECTION IV
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\usepackage{vmargin}
-\usepackage{color}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amstext}
-\usepackage{amsbsy}
-%\usepackage{mathbbol}
-\usepackage{graphicx} 
-\usepackage{verbatim}
-\usepackage{csvsimple} 
-\usepackage{subcaption}
-\usepackage{hyperref}
-\usepackage{fancyhdr}
-\usepackage{multirow}
-\usepackage{tikz}
-\usepackage{listings}
-\usepackage[numbers,sort&compress]{natbib}[2010/09/13]
+\chapter{DYNAMICS EXAMPLES \label{sect:dyn}}
 
-%  new definitions
-\renewcommand{\div}{\bs{\nabla}\! \cdot \!}
-\newcommand{\grad}{\bs{\nabla}}
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert_{L^2}}
-% extra space
-\newcommand{\qq}{\quad\quad}
-% common reference commands
-\newcommand{\eqt}[1]{Equation~\ref{#1}}                     % equation
-\newcommand{\fig}[1]{Figure~\ref{#1}}                      % figure
-\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table
-\newcommand{\sct}[1]{Section~\ref{#1}}                   % section
-\newcommand{\app}[1]{Appendix~\ref{#1}}                   % appendix
+Reactor dynamics is the study of the time-dependent behavior of reactors as an entire system.
+This study includes the physical nature of neutrons and their feedback with power generation.
+This feedback includes coupling with other physical properties such as temperature, fluids, material dynamics, etc.
+This section describes several dynamics examples, including the LRA benchmark and two TREAT experiments.
+These examples are of increased complexity from the kinetics examples of \sct{sect:kin}.
+This section also analyzes IQS's performance with these, which is vital for verification of IQS in real-world problems.
 
-\newcommand{\bs}[1]{\mathbf{#1}}
-\newcommand{\dd}{\mathrm{d}}
-\newcommand{\keff}{k_\textit{eff}}
-
-\newcommand{\be}{\begin{equation}}
-\newcommand{\ee}{\end{equation}}
-\newcommand{\vn}{\vec{n}}
-\newcommand{\vel}{\vec{\mathrm{v}}}
-\newcommand{\adj}{\Phi^\dagger_0}
-\newcommand{\tcr}[1]{\textcolor{red}{#1}}
-\newcommand{\tcb}[1]{\textcolor{blue}{#1}}
-
-% tikz stuff
-\usetikzlibrary{shapes,arrows,positioning}
-\pgfdeclarelayer{background}
-\pgfdeclarelayer{foreground}
-\pgfsetlayers{background,main,foreground}
-\tikzstyle{greenblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=green!20, ultra thick, minimum width=60mm, minimum height=15mm,]
-\tikzstyle{blueblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=blue!20, ultra thick, minimum width=60mm, minimum height=15mm]
-\tikzstyle{reddiamond}=[diamond, draw, align=center, top color=white, bottom color=red!20, ultra thick, minimum width=60mm, aspect=2]
-\newcommand{\tikzback}[5]{
- \begin{pgfonlayer}{background}
-  \path (#1.west |- #2.north)+(-0.5,0.75) node (a1) {};
-  \path (#3.east |- #4.south)+(+0.5,-0.25) node (a2) {};
-  \path[fill=yellow!10,rounded corners, draw=black!100, dashed] (a1) rectangle (a2);
-   \path (#3.east |- #2.north)+(0,1.25)--(#1.west |- #2.north) node[midway] (#5-n) {};
-   \path (#3.east |- #2.south)+(0,-0.35)--(#1.west |- #2.south) node[midway] (#5-s) {};
-   \path (#1.west |- #2.north)+(-0.75,0)--(#1.west |- #4.south) node[midway] (#5-w) {};
- \end{pgfonlayer}}
-
-% input stuff
-\lstset{%
-%  language=C++,
-  showstringspaces=false,
-% basicstyle=\scriptsize\ttfamily,
-  basicstyle=\small\ttfamily,
-  commentstyle=\color{inl@green},
-%  keywordstyle=\bfseries,
-%  escapeinside=$$
-%  framesep=0pt,
-%  rulesep=0pt,
-  frame=single
-}
-% Make nice big tildes inside lstlistings instead of
-% tiny ones.
-\lstset{
-    literate={~} {$\sim$}{1}
-}
-
-% ********* Caption Layout ************
-%\usepackage{ccaption} % allows special formating of the captions
-%\captionnamefont{\bf\footnotesize\sffamily} % defines the font of the caption name (e.g. Figure: or Table:)
-%\captiontitlefont{\footnotesize\sffamily} % defines the font of the caption text (same as above, but not bold)
-%\setlength{\abovecaptionskip}{0mm} %lowers the distance of captions to the figure
-\setlength\parindent{0pt}
-\setlength{\oddsidemargin}{1in}
-\setlength{\evensidemargin}{1in}
-\setlength{\textwidth}{6.5in}
-% ********* Header and Footer **********
-% This is something to play with forever. I use here the advanced settings of the KOMA script
-
-\title{Improved Quasi-Static Methods in Rattlesnake}
-\subtitle{Development and Implementation of Improved Quasi-Static (IQS) methods for time-dependent neutron diffusion and neutron transport solvers in Rattlesnake}
-\author{ \normalsize
-  \textbf{Zachary M. Prince$^\dagger$, Jean C. Ragusa$^\dagger$, Yaqi Wang$^\star$} \\
- \normalsize $^\dagger$Department of Nuclear Engineering \\
- \normalsize Texas A\&M University, College Station, TX, USA\\
- \normalsize $^\star$Idaho National Laboratory\\
- \normalsize \href{mailto:zachmprince@tamu.edu}{zachmprince@tamu.edu}, \href{jean.ragusa@tamu.edu}{jean.ragusa@tamu.edu}, \href{yaqi.wang@inl.gov}{yaqi.wang@inl.gov} 
-}
-
-\hypersetup{
-  colorlinks=true,
-  urlcolor=blue,
-}
-
-\pagestyle{fancy}
-\fancyhf{}
-\lhead{INL/EXT-16-38059}
-\rfoot{DRAFT}
-\cfoot{\thepage}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% boxes
-\usepackage{color}
-\definecolor{myblue}{rgb}{.8, .8, 1}
-\usepackage{empheq}
-
-\newlength\mytemplen
-\newsavebox\mytempbox
-
-\makeatletter
-\newcommand\mybluebox{%
-    \@ifnextchar[%]
-       {\@mybluebox}%
-       {\@mybluebox[0pt]}}
-
-\def\@mybluebox[#1]{%
-    \@ifnextchar[%]
-       {\@@mybluebox[#1]}%
-       {\@@mybluebox[#1][0pt]}}
-
-\def\@@mybluebox[#1][#2]#3{
-    \sbox\mytempbox{#3}%
-    \mytemplen\ht\mytempbox
-    \advance\mytemplen #1\relax
-    \ht\mytempbox\mytemplen
-    \mytemplen\dp\mytempbox
-    \advance\mytemplen #2\relax
-    \dp\mytempbox\mytemplen
-    \colorbox{myblue}{\hspace{1em}\usebox{\mytempbox}\hspace{1em}}}
-
-\makeatother
-
-%################ End Preferences, Begin Document #####################
-
-\pagestyle{plain} % on headers or footers on the first page
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% set relative directory for figures. can be changed at any time later with \renewcommand
-\newcommand{\FiguresDir}{./figs}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\maketitle
-\pagenumbering{arabic}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph*{Abstract}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This report presents  the implementation of the Improved Quasi-Static method (IQS) in the Yak/Rattlesnake applications of the MOOSE framework.
-In the IQS method, the neutron flux is factored into a time-dependent amplitude and a spatial- and weakly time-dependent shape.
-The shape evaluation is very similar to a standard flux solve and is computed at large (macro) time steps.
-The amplitude is obtained by solving the PRKEs on a micro time scale.
-The PRKE parameters are regularly updated based on the shape changes.
-Two versions of the IQS methodology have been implemented: (i) the standard approach, which leads to a system of nonlinear equations between the shape and the amplitude, and (ii) a predictor corrector approach that linearizes the system of equations.
-A time-adaptation technique has been developed to automatically select an appropriate macro time step size based on user-specified tolerances.
-The implementation has been verified and tested using manufactured solutions, a custom one-dimensional example, the TWIGL ramp benchmark, and the LRA benchmark.
-The numerical results include error convergence plots and performance with time adaptation.
-These examples prove it to be a viable and effective method for  transient neutronic simulations.
-More complex simulations (e.g., TREAT transients) are planned to perform coupled neutronics/heat conduction transient simulations within the MAMMOTH application.
-
-\pagebreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Theory}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\par
-The improved quasi-static (IQS) method is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (the shape) and a time-dependent component (the amplitude) \cite{Ott_1966,Dulla2008}.
-The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates.
-The IQS method has primarily been developed in the context of the neutron diffusion approximation \cite{Ban_2012,Ferguson_1973}, but reference \cite{Goluoglu_2001} provides an extension of the technique to transport, implemented in TDTORT.
-The method has mostly being applied to neutron kinetics, without thermal-hydraulic feedback.
-In this report, we extend the IQS method in the context of transient multiphysics simulations.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{IQS for Multigroup Diffusion Equations}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In this Section, we recall the equations for the IQS method, using the standard multigroup diffusion equations written below:
-\begin{subequations}
-\begin{align}
-\frac{\partial}{\partial t}\left(\frac{\phi^g}{v^g}\right)  =& \frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \phi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g  \nonumber \\
-&  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \phi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
-\label{eq:flux}
-\end{align}
-\be
-\frac{dC_i}{dt} = \frac{\beta_i}{\keff}\sum_{g=1}^G\nu^{g} \Sigma_f^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
-\label{eq:precursor}
-\ee
-\end{subequations}
-%
-Extension to the transport equations will be provided subsequently, once a concise operator notation has been introduced. 
-In the above equations,  \\
-
-\begin{tabular}{lll}
-$\phi^g$   			&	$=$	&	Scalar flux in energy group $g$ \\
-$C_i$					  &	$=$	&	Concentration of delayed neutron precursor $i$ \\
-$\Sigma_f^{g}$	&	$=$	&	Fission cross section in energy group $g$ \\
-$\Sigma_r^{g}$	&	$=$	&	Removal cross section in energy group $g$ \\
-$\Sigma_s^{g' \to g}$	&	$=$	&	Scattering cross section from energy group $g'$ to $g$ \\
-$v^g$					  &	$=$	&	Neutron velocity in energy group $g$ \\
-$\chi_p^g$			&	$=$	&	Fission spectrum of prompt neutrons \\
-$\chi_{d,i}^g$	&	$=$	&	Fission spectrum of delayed neutrons from precursor $i$ \\
-$\nu^g$					&	$=$	&	Total number of neutrons per fission \\
-$D^g$					  &	$=$	&	Diffusion coefficient in energy group $g$\\
-$\lambda_i$			&	$=$	&	Decay constant of precursor $i$ \\
-$\beta_i$				&	$=$	&	Delayed neutron fraction  from precursor $i$ \\
-$\beta$			 	  &	$=$	&	Total delayed neutron fraction ($\beta = \sum_{i=1}^I \beta_{i}$) \\
-  & & 
-\end{tabular}
-
-Note that the prompt fission terms have been normalized by $\keff$, the eigenvalue of the initial state, as transient simulations are often carried out for initially critical reactors.
-The expressions for subcritical and source-driven problems are straightforward to derive but omitted in this report for brevity.
-\\
-
-The factorization approach in IQS leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent multigroup shape ($\varphi^g$):
-\be
-\phi^g(\vec{r},t)=p(t)\varphi^g(\vec{r},t)
-\ee
-We stress that the factorization is not an approximation. Reporting it in the flux and precursor equations yields, after some simple algebra, the following equations for the shape and the precursors:
-\begin{subequations}
-\begin{align}
-\frac{\partial}{\partial t}\left(\frac{\varphi^g}{v^g}\right) = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\
-& -  \left( -\div D^g \grad  + \Sigma_r^g + \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right) \varphi^g + \boxed{\frac{1}{p}}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
-\label{eq:shape2}
-\end{align}
-\be
-\frac{dC_i}{dt} = \frac{\beta_i}{\keff}\boxed{p} \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
-\label{eq:preq2}
-\ee
-\end{subequations}
-
-We note that the time-dependent shape equation is similar to the time-dependent flux, with the following modifications:
-\begin{enumerate}
-\item The shape equation contains an additional term equivalent to a removal cross section,  $ \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$.
-\item The delayed neutron source term is divided by $p$.
-\item The system of equations is now nonlinear due to the factorization.
-\item An equation is needed to obtain the amplitude $p$.
-\end{enumerate}
-
-To derive the amplitude equation (also known as the PRKE equations), the shape/precursors equations are weighed by a space-dependent function (typically the initial adjoint flux, $\phi^{*g}$) and integrated over the phase-space.
-The final expressions are given below:
-\begin{subequations}
-\be
-\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
-\label{eq:p}
-\ee
-\be
-\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
-\label{eq:c}
-\ee
-\end{subequations}
-where we have defined the reactivity, effective delayed-neutron fraction, and delayed-neutron precursor decay constant as follows:
-\begin{subequations}
-\be
-\frac{\rho-\bar{\beta}}{\Lambda} = 
-\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}
-{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-\label{eq:rmb}
-\ee
-\be
-\frac{\bar{\beta}}{\Lambda} = \sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda} = 
-\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}
-{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-\label{eq:b}
-\ee
-\be
-\bar{\lambda}_i = \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
-\label{eq:l}
-\ee 
-\end{subequations}
-The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.\\  
-Additionally, in order to impose uniqueness on the factorization and to derive the PRKE, the following normalization condition is imposed: $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}$.
-This condition is later used  as the convergence criteria when solving the nonlinear system of equations resulting from the IQS method.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Operator Notation and Extension of IQS to Multigroup Transport Equations}
-\label{sect:transport}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-For simplicity, the previous section only described the IQS implementation using the neutron diffusion equation.
-However, deriving the IQS form for other neutron balance equations (e.g., transport, simplified transport, etc.) is very similar.
-To this end, we re-write the neutron conservation equations in operator form, shown in \eqt{eq:transport}.
-\be
-\frac{\partial}{\partial t}\left(\frac{\Psi^g}{v^g}\right) = \sum_{g'} \left(H^{g'\to g} + P_p^{g' \to g} \right) \Psi^{g'} - L^g\Psi^g + S_d^g
-\label{eq:transport}
-\ee
-where $\Psi^g$ is the multigroup neutron flux (angular flux in the case of transport), $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt neutron production operator, $L^g$ is the loss operator, and $S_d^g$ is the delayed neutron source operator.
-Using \eqt{eq:flux}, the reader may easily obtain the functional form for these operators in the case of a diffusion approximation.
-%
-%Factorization is defined as $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ and the resulting shape equation is defined by \eqt{eq:tshape}.  
-Next, the factorization $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ is introduced, where the shape is denoted by $\psi^g$, leading to the following shape equations, \eqt{eq:tshape}:
-\be
-\frac{\partial}{\partial t}\left(\frac{\psi^g}{v^g}\right) = \sum_{g'} \left(H^{g' \to g} + P_p^{g' \to g}\right) \psi^{g'} - \left( L^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \psi^g + \frac{1}{p} S_d^g
-\label{eq:tshape}
-\ee
-Finally, the PRKE parameters are defined by Equations~\ref{eq:trmb}~and~\ref{eq:tb}, where 
-$\left(\Psi^{*g},f^g\right) =$\\ $ \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d^2\Omega$.
-\be
-\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},\sum_{g'}(H^{g' \to g}g+P_p^{g' \to g}-L^{g'}\delta_{g'g})\psi^{g'}\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
-\label{eq:trmb}
-\ee
-\be
-\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{\sum_{g=1}^G(\Psi^{*g}, \sum_{g'} P_{d,i}^{g' \to g} \psi^{g'})}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
-\label{eq:tb}
-\ee
-where $P_{d,i}^{g' \to g}$ is the delayed-neutron operator for precursor group $i$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Solving the IQS System of Equations}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Solving for the shape can become expensive, especially in two or three dimensions, and even more so when using the transport equations in lieu of the diffusion equations.
-Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself,  thus allowing for larger time step sizes in updating the shape.
-The PRKE equations form a small ODE system and can be solved using a much smaller time step size.
-In transients where the shape varies much less than the flux, IQS can thus be very computationally effective.
-The two time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.
-
-\begin{figure}[!htbp]
-\includegraphics[width=\linewidth]{\FiguresDir/IQS_visualization.jpg}
-\caption{IQS method visualization}
-\label{fig:iqsviz}
-\end{figure}
-
-By this design, the benefits of IQS is most realized when using time adaptive techniques, one of these is described in  \sct{sect:dt2}.
-As we noted earlier, shape-PRKE equations are a nonlinear system and thus need to be solved in a iterative manner.
-Each macro time step can be iterated so the best shape is used to compute power at the micro time steps.
-This iteration process must converge the shape such that the uniqueness condition $(\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=0)$ is preserved.
-
-The most convenient method to implement IQS in RATTLESNAKE is through Picard or fixed-point iterations.
-Over a given Picard iteration, the shape is evaluated first with the latest available amplitude.
-The updated shape is employed to compute the PRKE parameters (using a linear interpolation in time) and then the PRKEs are solved using micro time steps.
-This process is repeated until convergence of the normalization condition.
-Picard iterations can also be utilized for multiphysics simulations, where neutronics (e.g., IQS) is nonlinearly coupled to thermal-hydraulics, for instance.
-Thus, inlcuding the IQS nonlinear system solve into the Picard iteration of MOOSE is a straightforward process and is described in \fig{fig:picard}.
-
-\begin{figure}[!htpb]
-\centering
-\begin{tikzpicture}[every node/.style = {font=\normalsize}]
-
-\node[greenblock](p1) at (0,0) {Solve shape equations \\ over macro time step};
-\node[greenblock](p2) at (0,-2.5) {Compute PRKE parameters \\ over macro time step};
-\node[greenblock](p3) at (0,-5) {Solve PRKE using micro \\ steps};
-\node[greenblock](p4) at (0,-7.5) {Update $p$ and $\frac{dp}{dt}$};
-\node[blueblock] (p5) at (0,-11) {Solve other physics \\ components};
-\node[reddiamond] (check) at (0,-15) {Check for \\ convergence};
-\node [above =0.1mm of p1] {\large{IQS Solve:}};
-\node [above =0.1mm of p5] {\large{Multi-physics:}};
-
-\tikzback{p1}{p1}{p4}{p4}{bk1}
-\tikzback{p5}{p5}{p5}{p5}{bk2}
-
-\draw[->,ultra thick](p1.south) -- (p2.north);
-\draw[->,ultra thick](p2.south) -- (p3.north);
-\draw[->,ultra thick](p3.south) -- (p4.north);
-\draw [->,ultra thick] (p4.south)+(0,-0.25) -- node [above] {} (bk2-n);
-\draw[->,ultra thick](p5.south)+(0,-0.25) -- (check.north);
-%\draw[->,ultra thick](check.west) -- (-5,-15) -- node[above,sloped] {\large{no}}(-5,0) -- node[above] {}(bk1-w);
-\draw[->,ultra thick](check.west) -- node[above,sloped] {\large{no}} (-5,-15) |-  node[above] {}(bk1-w);
-\draw[->,ultra thick](check.south) -- node[right] {\large{yes}} ++(0,-2);
-
-\end{tikzpicture}
-\caption{Visualization of Picard iteration process}
-\label{fig:picard}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{A Predictor-Corrector version of IQS (IQS P-C)}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The Predictor-Corrector (P-C) version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the solution of the coupled system of equations is different.
-In the IQS P-C version, the flux equations (not the shape equations) are solved (represented by Equations~\ref{eq:flux}~and~\ref{eq:precursor}) in order to obtain a predicted flux solution.
-This predicted flux is then converted to shape by rescaling as follows:
-\be
-\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
-\label{eq:rescale}
-\ee
-where the scaling factors are given by
-\be
-K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
-\ee
-\be
-K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0} \,.\right)
-\ee
-
-The PRKE parameters are then computed with this shape using Equations~\ref{eq:rmb}~-~\ref{eq:l} and interpolated over the macro step, then the PRKE ODE system is solved on the micro time scale.
-With the newly computed amplitude, the shape is rescaled into a flux and the final corrected flux is given by:
-\be
-\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1} \,.
-\ee
-
-The advantage to the predictor-corrector method is there is no iteration necessary for this method and, in turn, is much simpler and faster than the standard IQS.  
-%The disadvantage is that the method assumes, by construction, that $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)$ is inherently constant.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Precursor Integration}
-\label{sect:prec}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This section presents the time-integration method used to solve coupled IQS shape and precursor equations, represented by Equations~\ref{eq:shape2}~and~\ref{eq:preq2}.
-First, we note that we could keep this system of two time-dependent equations and solve it as a coupled system.
-However, this is unnecessary and a memory expensive endeavor because the precursor equation is only an ODE and not a PDE.
-Instead, one may discretize in time the shape equation, which typically requires the knowledge of the precursor concentrations at the end of the time step.
-This precursor value is taken from the solution, numerical or analytical, of the precursors ODE.
-\\
-
-Through prototyping, it has been found that typical implicit nor Crank-Nicholson time discretizations of precursors are not preferable methods for solving the shape equation in IQS.
-It has been found that these discretizations result in a lack of convergence of the shape over the IQS iteration process.
-In order to remedy the error, a analytical representation of the precursors was implemented in the prototype and the shape solution was able to converge.
-The method entails linearly interpolating the fission source ($S_f = \nu \Sigma_f \varphi$) and integrating the resulting continuous function analytically.
-\eqt{eq:c_dis} is the equation to solve for the precursors and Equations~\ref{eq:a1}~and~\ref{eq:a2} are integrating factors.
-
-\be
-C^{n+1} = C^n e^{-\lambda \Delta t} 
-+ \beta_i \left(a_2 S_f^{n+1} + a_1 S_f^{n}\right)
-\label{eq:c_dis}
-\ee
-
-\be
-a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)p(t')e^{-\lambda(t_{n+1}-t')}dt'
-\label{eq:a1}
-\ee
-\be
-a_2 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)p(t')e^{-\lambda(t_{n+1}-t')}dt'
-\label{eq:a2}
-\ee
-
-The amplitude $(p)$ is included in the integration coefficient because it has been highly accurately calculated in the micro step scheme, so a piecewise interpolation between those points can be done to maximize accuracy.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Time Adaptation}
-\label{sect:dt2}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-IQS aims at reducing the time discretization error in the flux solution by splitting the flux into an amplitude (that is highly resolved with at a micro time scale) and a shape (whose time-dependence is weaker than that of the flux itself).
-Thus, by construction, the IQS approach may employ larger time-step sizes for comparable temporal error.
-Further enhancements can be gained by using time adaptation (or time step control) in order to increase or reduce the macro time step size for the
-shape evaluation, depending on error estimates. Here, we employ a step-doubling technique to carry out the time adaptation.
-The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps ($\varphi^g_{\Delta t/2}$).
-Note:
-$\varphi$ is changed to $\phi$ for regular flux evaluation (brute-force time discretization of the governing equations) and IQS P-C.
-
-The relative error is computed as follows:
-\be
-e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2} - \sum_{g=1}^G\varphi^g_{\Delta t}}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2}},\norm{\sum_{g=1}^G\varphi^g_{\Delta t}}\right)}
-\label{eq:edt2}
-\ee
-If the error is smaller than the user-specified tolerance, $e_{tol}$, the time step is accepted. In addition, a new time step size is estimated as follows:
-\be
-\Delta t_{new} = S \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}} \,,
-\label{eq:dt2}
-\ee
-where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor.\\
-If the error is larger than the user-specified tolerance, the time step is rejected. A new time step size is estimated using  \eqt{eq:edt2} as well.
-
-To sum up, each solution undergoes Picard iterations until convergence before the error is calculated; for IQS P-C, each solution is re-scaled by its corresponding amplitude $p$.
-Additionally, if the nonlinear iteration does not converge, the entire step doubling process is repeated with half the time original time step.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{RATTLESNAKE Implementation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-RATTLESNAKE is a MOOSE-based application developed INL specific to solving radiation transport problems with multiphysics capabilities.
-MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like RATTLESNAKE.
-At the heart of RATTLESNAKE is the action system, which provides a means to consolidate the MOOSE input syntax (which can be quite larger for multigroup transport simulations), so that a user does not have to define every kernel, variable, etc., used in the problem.
-Rather, the user inputs an equation description
-(e.g., Diffusion, S$_n$, P$_n$, etc.) and a solution method (e.g., SAAF, LS, CFEM, DFEM, etc.), and the action system will incorporate all the necessary physics involved (kernels, boundary conditions, postprocessor, etc.).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Executioner}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-An IQS excecutioner was created to implement the IQS method.
-The IQS executioner derives from the Transient executioner in MOOSE.
-The IQS executioner contains a loop over micro time steps that computes the PRKE solution and then passes $p$ and $\frac{dp}{dt}$ for the Transient executioner to evaluate the shape equation at each macro step.
-The PRKE solve is performed with a user specified option of backward-Euler, Crank-Nicholson, or SDIRK33.
-The IQS executioner also supplements Transient Picard iteration process by adding its own error criteria:
-\be
-Error_{IQS}=\left|\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,n}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,0}\right)}-1\right|
-\label{eq:eiqs}
-\ee
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Action System}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The IQS implementation mostly require the specific IQS executioner, described above.
-However, IQS additional changes are in the RATTLESNAKE action system in order to support the IQS execution.
-First, changes needed to be made in order to evaluate the shape equation.
-The shape equation, after some manipulation, is very similar to the time-dependent governing laws that RATTLESNAKE already solves.
-Using multigroup diffusion as an example again, we show the various kernels employed and highlight the new or modified kernels.
-
-\begin{align}
-\frac{\partial}{\partial t}\left(\frac{\varphi^g}{v^g}\right)=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
-& - \underbrace{\frac{1}{v^g} \boxed{\overbrace{\frac{1}{p}\frac{dp}{dt}}^{From Executioner}}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
-\end{align}
-
-To enable RATTLESNAKE to solve this equation, an IQS removal kernel was created to evaluate $\sum_{g=1}^G\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.
-Also, the precursor kernel was modified to include the $\frac{1}{p}$ term.
-Finally, the precursor auxkernel that evaluates \eqt{eq:preq2} using the analytical integration method described in \sct{sect:prec}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{PRKE coefficients}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In order to evaluate the PRKE coefficients, defined by Equations~\ref{eq:rmb}~-~\ref{eq:l}, four postprocessors were created.
-The parameter calculations were separated by $\frac{\bar{\beta}_i}{\Lambda}$ numerator, $\bar{\lambda}_i$ numerator/denominator, $\frac{\rho-\bar{\beta}}{\Lambda}/\frac{\bar{\beta}}{\Lambda}$ denominator, and $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator.
-The first three are relatively simple, only relying on material properties and solution quantities, then computing the elemental integral.
-The $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator requires the use of the MOOSE \texttt{save\_in} feature.
-This feature saves the residual from a calculated kernel or boundary contribution in the shape evaluation to an auxiliary variable.
-The postprocessor then computes the inner product of this variable and the initial adjoint solution.
-After each of these postprocessors are evaluated, a user object pulls together all the values and performs the numerator/denominator divisions.
-The resulting values are then passed to the executioner for the PRKE solve.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Other Action Systems}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-For simplicity, IQS implementation has only been described for CFEM diffusion.
-However, RATTLESNAKE has other action systems capable of transient simulation, where IQS can be implemented and be effective.
-One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equations~\ref{eq:flux}~and~\ref{eq:shape2}.
-However, in the derivation for IQS, this term is unaffected between shape and flux evaluation.
-So saving the residual for this diffusion kernel in the \texttt{save\_in} variable is the only alteration to this action for IQS to function.\\
-
-Additional action systems involve transport, but it is evident from \sct{sect:transport} that IQS implementation in these is straightforward as well.
-The main differences between a diffusion implementation and a transport implementation are outlined below:
-\begin{enumerate}
-\item The form of the operators in the shape equations is different, but RATTLESNAKE has already implemented all the kernels necessary to represent these operators.
-So no change is necessary to these kernels is necessary for IQS.
-Additionally, the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ is the same, so both action systems can use the same kernel.
-\item The PRKE parameters also change because of the operators.
-For the $(\rho-\bar{\beta})$ parameter, the same post-processor can be used with the \texttt{save\_in} functionality in MOOSE.
-The post-processor for $\bar{\beta}_i$ must be re-written for transport, but takes a very similar form as diffusion.
-\end{enumerate}
-
-IQS will be made available to any action system capable of transient simulation, which includes:
-\begin{itemize}
-\item CFEM-Diffusion
-\item DFEM-Diffusion
-\item SAAF-CFEM-SN
-\item SAAF-CFEM-PN
-\item LS-CFEM-SN
-\item DFEM-SN
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Predictor-Corrector Modification}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In order to preserve the already implemented standard version of IQS, an option in the IQS executioner was created to specify which method is desired.
-Because the diffusion solve is flux instead of shape, when predictor-corrector option is specified, the IQS removal kernel ($\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$) and the modified precursor kernel are bypassed, while all the postprocessors are still executed.
-However, it is difficult to rescale the flux to shape before the PRKE parameter postprocessor are executed.
-So the parameters are computed using the full flux, but amplitude is space independent and comes out of the integrals.
-As seen in parameter definitions, when shape is replaced with flux, the amplitude comes out of the integral and cancels out.
-So the conversion of the predicted flux in  \eqt{eq:rescale} to shape is unnecessary if the corrected flux is solved with \eqt{eq:phicorr2}.
-After obtaining the corrected flux, the precursors are re-evaluated using a \texttt{EXEC\_LINEAR} statement.
-\be
-\underbrace{\phi^g_{n+1}}_{\text{corrected}} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}} p_{n+1}
-\label{eq:phicorr2}
-\ee
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Input}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The input file for IQS is very similar to the current transient diffusion input file.
-The IQS input has a different executioner type and parameters.
-The executioner type is simply IQS and input parameters include the number of micro time steps per macro step, the IQS error tolerance, and the initial power.
-The RATTLESNAKE transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $\keff$, which is present in the transient input deck.
-However, IQS also requires an initial evaluation of the adjoint flux, for the weighting function.
-So another input file and multi-app transfer was made for the adjoint calculation.
-\\
-
-Below is the syntax for the executioner block for an IQS input file.
-The \texttt{predictor\_corrector} logical determines whether to do IQS P-C or regular IQS.
-The \texttt{IQS\_error\_tol} is the tolerance for the IQS error represented by \eqt{eq:eiqs} and will at most \texttt{picard\_max\_its} iterations until convergence.
-The \texttt{prke\_scheme} defines the time discretization used for the PRKE solution.
-Where \texttt{RK} is SDIRK33, \texttt{CN} is Crank-Nicholson, and \texttt{IE} is implicit-Euler.
-\begin{lstlisting}
-[Executioner]
-  type = IQS
-  predictor_corrector = true/false
-  picard_max_its = 5
-  ...
-  n_micro = 10000
-  IQS_error_tol = 1e-7
-  prke_scheme = 'RK'
-[]
-\end{lstlisting}
-
-Since IQS needs to use the adjoint solution for the PRKE parameter evaluation, auxiliary variables need to be created for each group in the input file.  Below is an example of their definition: 
-\begin{lstlisting}
-[AuxVariables]
-  [./adjoint_flux_g0] 
-    family = LAGRANGE
-    order = FIRST
-  [../]
-  [./adjoint_flux_g1]
-    family = LAGRANGE
-    order = FIRST
-  [../]
-  ...
-[]
-\end{lstlisting}
-
-Below is the syntax for the multi-app block to perform forward and adjoint steady-state evaluations:
-\begin{lstlisting}
-[MultiApps]
- [./initial_solve]
-   type = FullSolveMultiApp
-   execute_on = initial
-   input_files = initial.i
- [../]
- [./adjoint_solve]
-   type = FullSolveMultiApp
-   execute_on = initial
-   input_files = adjoint.i
- [../]
-[]
-\end{lstlisting}
-
-Below is the syntax for the Transfer block to copy the initial and adjoint solutions from the multi-apps.
-\begin{lstlisting}
-[Transfers]
- [./copy_solution]
-   type = TransportSystemVariableTransfer
-   direction = from_multiapp
-   multi_app = initial_solve
-   execute_on = initial
-   from_transport_system = diff
-   to_transport_system = diff
- [../]
- [./copy_adjoint]
-   type = MultiAppVariableTransfer
-   execute_on = initial
-   direction = from_multiapp
-   multi_app = adjoint_solve
-   from_variables = 'sflux_g0 sflux_g1 ...'
-   to_variables = 'adjoint_flux_g0 adjoint_flux_g1 ...'
- [../]
- [./copy_eigenvalue]
-   type = EigenvalueTransfer
-   execute_on = initial
-   direction = from_multiapp
-   multi_app = initial_solve
- [../]
-[]
-\end{lstlisting}
-
-\newpage
-%------------------------------------------------------------------------------
-%
-%------------------------------------------------------------------------------
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{RESULTS} 
-\label{sect::results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This section describes results of examples that test the IQS implementation and shows its effectiveness on computation speed and accuracy.
-Three examples were selected for this purpose.
-The first is a homogeneous one-group problem, subjected to a heterogeneous material change (absorption cross-section change as a ramp in time for a subset of the geometry).
-The second is the two-dimensional TWIGL ramp transient benchmark, described further.
-The final is the LRA benchmark, which a two-dimensional, temperature feedback example.
-Each example was solved with regular diffusion (brute force), IQS, and IQS P-C.
-Each method was tested with backward-Euler and BDF2 for the first two examples; just Crank-Nicholson was used for third example.
-The performance for each method are represented by convergence plots with constant $\Delta t$ and by the number of time steps with step doubling time adaptation.
-For time adaptation, backward-Euler time discretization was used.
-% and, for a safety factor, $e_{tol} = 0.8e_{max}$. I found tol and max to be confusing to the reader. i have simplified a bit
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{One-Dimensional Custom Example}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The example is very simple and computes quickly, it entails a one dimensional, homogeneous 400 cm slab with a heterogenous perturbation in absorption cross section.
-\fig{fig:slab} shows how the regions of the slab are divided and \tbl{tab:1Dmat} shows the initial material properties.
-Regions 2, 3, and 4 have slope perturbations at different points in time, \tbl{tab:1Dslope} shows the values of the absorption cross-section in each region at the times of interest.
-The values of $\Sigma_a$ between these times of interest are linear interpolations between the given values.
-
-\begin{figure}[!htbp]
-\begin{center}
-\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l |}
-\hline \hline \hline
-  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
-1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 4 & 4 & 1 & 1 & 1 & 1 \\
-  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
-\hline \hline \hline
-\end{tabular}
-\caption{1-D slab region identification}
-\label{fig:slab}
-\end{center}
-\end{figure}
-
-\begin{table}[!htbp]
-\begin{center}
-\caption{1-D slab material properties and problem parameters}
-\label{tab:1Dmat}
-\begin{tabular}{llllll}
-\hline
-$D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu \Sigma_f (cm^{-1})$ & $v (cm/s)$ & $\beta$ & $\lambda (s^{-1})$ \\
-\hline
-1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\
-
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[!htbp]
-\begin{center}
-\caption{1-D slab absorption cross-section at times of interest}
-\label{tab:1Dslope}
-\begin{tabular}{llllllll}
-\hline
-Region & Material Property & 0.0 s & 0.1 s & 0.6 s & 1.0 s & 1.7 s \\
-\hline
-2 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.095 & 1.095 & 1.095 \\
-3 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.09 & 1.09 & 1.1 \\
-4 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.105 & 1.105 & 1.105 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\fig{fig:power} shows the power at each macro time step as compared to the traditional brute force (full flux time discretization) method.
-The strong correlation between the two curves shows that IQS is consistent with a proven method for a highly transient example.
-\fig{fig:conv} shows that IQS is not only consistent for this example, but also has a better error constant in the convergence study.
-\fig{fig:conv} shows the error convergence for brute force and IQS methods.
-This plot shows that IQS has consistent convergence slope for first and second order time schemes and that IQS has a better error constant for the convergence.
-\tbl{tab:1Ddt2} shows the results for time adaptation.
-These results show that traditional IQS does not perform well with step doubling time adaptation and IQS P-C performs moderately better than brute force.
-The IQS performance in this example are not exceptional because it is not an easy problem.
-As seen in \fig{eq:D_shape}, the shape undergoes significant change through the transient, which requires many shape evaluations to retain accuracy.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.49\textwidth]{\FiguresDir/1D_het_power_plot.png}
-\includegraphics[width=0.5\textwidth]{\FiguresDir/1D_het_power_plot2.png}
-\caption{Power level of 1D example}
-\label{fig:power}
-\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=5in]{\FiguresDir/1D_het_convergence.jpg}
-\caption{Error convergence comparison of 1D example}
-\label{fig:conv}
-\end{figure}
-
-\begin{table}[!htbp]
-\begin{center}
-\caption{1-D slab step doubling results}
-\label{tab:1Ddt2}
-\resizebox{\textwidth}{!}{
-\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
-\hline
-  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
-\hline
-Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
-\hline
-1 &	0.05 	&	0.021596 &	10  &	32	 &	0.10084 &	12	  &	175		&	0.028019 &	10	&	33 \\
-2 &	0.01 	&	0.032864 &	21  &	80	 &	0.08030 &	41	  &	850	 	&	0.044721 &	20	&	74 \\
-3 &	0.005 	&	0.053159 &	27  &	96	 &	0.01215 &	115	  &	2220 	&	0.052095 &	25	&	96 \\
-4 & 0.001 	&	0.056546 &	56  &	188	 &	0.03242 &	848	  &	12511 	&	0.061274 &	50	&	162 \\
-5 &	0.0005 	&	0.062882 &	77  &	245	 &	0.03491 &	1718  &	25841 	&	0.061240 &	69	&	225 \\
-6 &	0.0001 	&	0.060089 &	177 &	537	 &	0.03554 &	8702  &	129985 	&	0.060824 &	159	&	480 \\
-7 &	5.0e-05	&	0.059513 &	252 &	767	 &	0.03622 &	17282 &	256463 	&	0.061078 &	224	&	680 \\
-8 &	1.0e-05	&	0.061063 &	561 &	1691 &	0.03142 &	79988 &	1104227 &	0.061901 &	501	&	1509 \\
-\hline
-\end{tabular}}
-\end{center}
-\end{table}
-
-\begin{figure}[!htbp]
-\begin{center}
-\includegraphics[width=\linewidth]{\FiguresDir/1D_shape.png}
-\caption{Flux profile of 1D slab at various times}
-\label{fig:1D_shape}
-\end{center}
-\end{figure}
-
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TWIGL Benchmark}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This benchmark problem originates from the Argonne National Lab Benchmark Problem Book \cite{ANL_BPB}.
-It is a 2D, 2-group reactor core model with no reflector region shown in \fig{fig:TWIGL_reg}.
-\tbl{tab:TWIGL_mat} shows the material properties of each fuel region and the ramp perturbation of Material 1.
-
-\begin{figure}[!htbp]
-\begin{center}
-\includegraphics[width=0.75\textwidth]{\FiguresDir/TWIGL_regions.jpg}
-\caption{TWIGL benchmark problem description}
-\label{fig:TWIGL_reg}
-\end{center}
-\end{figure}
-
-\begin{table}[!htbp]
-\begin{center}
-\caption{TWIGL benchmark material properties and slope perturbation}
-\label{tab:TWIGL_mat}
-\begin{tabular}{llllllll}
-\hline
-  &  &  &  &  &  &  \multicolumn{2}{c}{$\underline{\Sigma_s (cm^{-1})} $} \\
-Material & Group & $D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu\Sigma_f (cm^{-1})$ & $\chi$ & $g \rightarrow 1$ & $g \rightarrow 2$ \\
-\hline
-1 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01 \\
-  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
-2 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01  \\
-  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
-3 & 1 & 1.3 & 0.008 & 0.003 & 1.0 & 0.0 & 0.01  \\
-  & 2 & 0.5 & 0.050 & 0.060 & 0.0 & 0.0 & 0.00  \\
-\hline
-  & $\nu$ & $v_1 (cm/s)$ & $v_2 (cm/s)$ & $\beta$ & $\lambda (1/s)$ &   &   \\
-\hline
-  & 2.43 & 1.0E7 & 2.0E5 & 0.0075 & 0.08 &   &   \\
-\hline
- \multicolumn{8}{l}{\footnotesize Material 1 ramp perturbation:} \\
-\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (1-0.11667t) \quad t \leq 0.2 s$} \\
-\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (0.97666t) \quad t > 0.2 s$} \normalsize
-\end{tabular}
-\end{center}
-\end{table}
-
-\fig{fig:TWIGL_power} shows the IQS  solution as compared with the Brute Force solution.
-These plots show that IQS is consistent in more complex, higher dimensional problems in Rattlesnake.
-Finally, \fig{fig:TWIGL_conv} plots the error convergence of IQS and the Brute Force methods.
-The curves show the impressive convergence of IQS for the highly transient TWIGL example.
-\tbl{tab:TWIGLdt2} and \fig{fig:TWIGL_power_dt2} show the results for time adaptation.
-The results show that both IQS methods perform exceptionally well compared to brute force.
-It also shows that traditional IQS performed better with large $e_{tol}$, while IQS P-C was better with smaller $e_{tol}$.
-
-%It is important to note the IQS shape plot is scaled differently than the Brute Force flux plot because the amplitude term is not included, but the gradients of colors is comparable.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.49\linewidth]{\FiguresDir/TWIGL_power_plot.png}
-\includegraphics[width=0.5\linewidth]{\FiguresDir/TWIGL_power_plot2.png}
-\caption{Power level comparison of TWIGL Benchmark}
-\label{fig:TWIGL_power}
-\end{figure}
-
-%\begin{figure}[!htbp]
-%\begin{center}
-%\begin{subfigure}[!htbp]{0.4\textwidth}
-%\includegraphics[width=\textwidth]{\FiguresDir/ndiff_ramp_flux.jpg}
-%\caption{Brute force flux}
-%\end{subfigure}
-%\quad
-%\begin{subfigure}[!htbp]{0.4\textwidth}
-%\includegraphics[width=\textwidth]{\FiguresDir/iqs_ramp_shape.jpg}
-%\caption{IQS Shape}
-%\end{subfigure}
-%\caption{TWIGL Benchmark flux/shape comparison at $t=0.2$}
-%\label{fig:TWIGL_plots}
-%\end{center}
-%\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=5in]{\FiguresDir/TWIGL_convergence.jpg}
-\caption{Error convergence comparison of TWIGL Benchmark}
-\label{fig:TWIGL_conv}
-\end{figure}
-
-\begin{table}[!htbp]
-\begin{center}
-\caption{TWIGL step doubling results}
-\label{tab:TWIGLdt2}
-\resizebox{\textwidth}{!}{
-\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
-\hline
-  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
-\hline
-Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
-\hline
-1 &	0.05    &	0.00012677 &	9   &	29  &	0.03380433 &	4   &	20   &	0.00323100 &	4  &	9   \\
-2 &	0.01    &	3.5555e-05 &	11  &	35  &	0.00166991 &	5   &	40   &	0.00263068 &	5  &	12  \\
-3 &	0.005   &	4.0364e-05 &	11  &	31  &	0.00886584 &	5   &	40   &	0.00160486 &	6  &	21  \\
-4 &	0.001   &	0.00294822 &	33  &	122 &	0.02976305 &	5   &	36   &	1.7527e-05 &	10 &	35  \\
-5 &	0.0005  &	0.00099778 &	39  &	131 &	0.00143781 &	6   &	55   &	1.4185e-05 &	16 &	74  \\
-6 &	0.0001  &	0.00019510 &	78  &	236 &	0.00016175 &	7   &	65   &	6.2903e-06 &	19 &	78  \\
-7 &	5.0e-05 &	0.00018372 &	112 &	342 &	6.0328e-05 &	12  &	163  &	1.5247e-06 &	24 &	92  \\
-8 &	1.0e-05 &	8.0564e-05 &	263 &	794 &	7.7103e-05 &	379 &	5729 &	9.8321e-07 &	48 &	210 \\
-\hline
-
-\end{tabular}}
-\end{center}
-\end{table}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.8\linewidth]{\FiguresDir/TWIGL_power_plot_dt2.png}
-\caption{Power level comparison of TWIGL Benchmark with time adaptation}
-\label{fig:TWIGL_power_dt2}
-\end{figure}
-
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{LRA Benchmark}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{LRA Benchmark}
 
@@
 -\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
--\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1. \\
+-\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1.
 \end{align}
@@
 \end{table}
-
 The transient is initiated by changing the thermal absorption cross section as the following:
@@
 \end{align}
-where $t$ is time in seconds.\\
+where $t$ is time in seconds.
 
-The \fig{fig:LRA_plots} plots show the IQS  solution as compared with the Brute Force solution.
-These plots show that IQS is consistent with Doppler feedback problems in RATTLESNAKE.
-\fig{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods.
-\tbl{tab:LRAdt2} shows the time adaptation performance of brute force and IQS P-C.
-These results show that IQS, again, performs much better than the brute force method by taking significantly fewer diffusion solves while retaining similar accuracy.
+\section{TREAT Transient-15 Problem}
 
-\begin{figure}[!htbp]
-\begin{center}
-\begin{subfigure}[!htbp]{0.49\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/lra_power_profile.png}
-\caption{Power Profile}
-\end{subfigure}
-\begin{subfigure}[!htbp]{0.49\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/lra_power_profile2.png}
-\caption{Power Profile at Peak}
-\end{subfigure}
-\begin{subfigure}[!htbp]{0.49\textwidth}
-\includegraphics[width=\textwidth]{\FiguresDir/lra_temp_profile.png}
-\caption{Temperature Profile}
-\end{subfigure}
-\caption{LRA Benchmark power and temperature comparison}
-\label{fig:LRA_plots}
-\end{center}
-\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.6\linewidth]{\FiguresDir/lra_convergence.jpg}
-\caption{Error convergence comparison of LRA Benchmark}
-\label{fig:lra_conv}
-\end{figure}
-
-\begin{table}[!htbp]
-\caption{LRA step doubling results}
-\label{tab:LRAdt2}
-\begin{center}
-\begin{tabular}{|l|l|l|l|l|l|l|l|}
-\hline
-  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS P-C} \\
-\hline
-Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves \\
-\hline
-1 &	 0.01 &	 0.0437566 &	 145 &	 735 &	 0.143855 &	 105 &	 616 \\ 
-2 &	 0.005 &	 0.0309732 &	 266 &	 948 &	 0.104782 &	 157 &	 814 \\ 
-3 &	 0.001 &	 0.158242 &	 585 &	 2046 &	 0.155676 &	 218 &	 1062 \\ 
-4 &	 0.0005 &	 0.270949 &	 833 &	 2908 &	 0.211994 &	 480 &	 2038 \\ 
-5 &	 0.0001 &	 0.280468 &	 1793 &	 6155 &	 0.0437761 &	 434 &	 1458 \\ 
-6 &	 5e-05 &	 0.212274 &	 2485 &	 8511 &	 0.0385312 &	 542 &	 1719 \\ 
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-The purpose of this report is to briefly review the improved quasi-static method for neutronic transient simulations, to describe its implementation in RATTLESNAKE, the radiation transport application based on MOOSE, and to show its effectiveness using several test cases.
-%The derivation of IQS was done with multigroup diffusion, but it was seen in \sct{sect:transport} that application to transport is relatively simple.  
-Two IQS implementations are available in RATTLESNAKE.
-The first one is the traditional IQS approach, where the shape equation and PRKE are solved as a coupled nonlinear system with Picard iterations.
-The second approach, IQS P-C, is a predictor-corrector linearization, where the flux is solved and then rescaled using the amplitude obtained through the PRKE
-(Point Reactor Kinetic Equations).  IQS P-C is a simpler method and is computationally much less expensive to perform. \\
-
-IQS's implementation in RATTLESNAKE involves a new kernel, an auxiliary kernel, a new executioner derived from the current transient, four postprocessors, as well as alterations to the action system.
-The new kernel implements the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term in the shape equation.
-The auxiliary kernel alters how the precursors are solved, the method is outlined by \sct{sect:prec}.
-The executioner is where the PRKE are evaluated after shape solves are performed.
-The postprocessors are used to evaluate the parameters for the PRKE and a user object is used to pass them to the executioner.
-IQS's implementation the action system made sure that all the correct kernels and postprocessors are executed when the IQS executioner is specified in the input.
-\\
-
-The IQS implementations have been verified and comparing against the traditional ``brute force'' temporal discretization of the governing laws.
-Error convergence plots and time adaptation performance are provided.
-Three examples were used including a custom one-dimensional example, the TWIGL benchmark, and LRA benchmark.
-The one-dimensional example shows large shape changes in the flux and proves consistency of the shape evaluation.
-IQS shows proper error convergence for first and second order time discretization.
-IQS P-C shows marginal improvement with time adaptation, but traditional IQS shows poor performance.
-The TWIGL benchmark shows the effectiveness of the PRKE in IQS because there is very little shape change through the transient.
-IQS shows very satisfactory performance with both constant time steps and time adaptation.
-The LRA benchmark tests IQS when utilized in the context of a multiphysics simulation.
-Again, IQS shows good performance by requiring half the time steps as brute force for a given error.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\clearpage
-\pagebreak
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{abbrvnat}
-\bibliography{references_IQS}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Appendix: aaaa}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\tcr{putting stuff here for now}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%It can be shown that the most appropriate weighting function $(w^g)$ is the initial adjoint flux $(\phi^{*g})$.
-For brevity, the following definition will be applied: $\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r=\left(\phi^{*g},f\right)$
-%\begin{align}
-%& \quad \frac{dp}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)+ p\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)= \nonumber \\
-%& p\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)  \\
-%& + \sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i) \nonumber
-%\end{align}
-%\be
-%\frac{d}{dt}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)=\frac{1}{\keff}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})p - \sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \quad 1 \le i \le I
-%\ee
-%
-%In order to impose uniqueness of the factorization, one imposes:
-%\be
-%\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=constant
-%\ee
-%
-%Therefore the PRKE formulation reduces to:
-%\begin{align}
-%\frac{dp}{dt}= &\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g}  \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p \nonumber \\
-%& +\sum_{i=1}^I\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i
-%\end{align}
-%\be
-%\frac{d\xi_i}{dt}=\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p - \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i \quad 1 \le i \le I
-%\ee
-%Where:
-%\be
-%\xi_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g C_i)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-%\ee
-%
-%It is convenient to define the effective reactivity, delay-neutron fraction, and delayed-neutron precursor decay constant:
-%\be
-%\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-%%\label{eq:rmb}
-%\ee
-%\be
-%\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
-%%\label{eq:b}
-%\ee
-%\be
-%\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
-%%\label{eq:l}
-%\ee
-%
-%So:
-%\be
-%\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
-%%\label{eq:p}
-%\ee
-%\be
-%\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
-%%\label{eq:c}
-%\ee
-%
-%Equations (\ref{eq:p}) and (\ref{eq:c}) are a formulation of the point reactor kinetics equation (PRKE), but the parameters (Equations (\ref{eq:rmb})-(\ref{eq:l})) are dependent on the shape.
-If the assumption is made that the shape is time-independent, the shape is computed once at the first time step and used for the PRKE parameter evaluation at all other steps.
-However, if the shape is dependent on time, the shape needs to be computed in transient using equation (\ref{eq:shape}) and (\ref{eq:prec}) in order retain accuracy.
-Equations (\ref{eq:shape2}) and (\ref{eq:preq2}) shows the usual form of the shape and precursor equations with amplitude put on the right hand side.
-Equation (\ref{eq:shape2}) is very similar to the multigroup flux equation (\ref{eq:flux}), except the removal cross-section term is augmented by a $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term and the precursor contribution has a $\frac{1}{p}$ multiplier.
-Equation (\ref{eq:preq2}) is very similar to the normal precursor equation (\ref{eq:precursor}), except the fission source term is multiplied by $p$.
-These differences are crucial for IQS implementation in Rattlesnake.
-%
-%\begin{align}
-%\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\
-%& -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
-%%\label{eq:shape2}
-%\end{align}
-%\be
-%\frac{dC_i}{dt} = \frac{\beta_i}{\keff}p \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
-%%\label{eq:preq2}
-%\ee
-
-%\begin{align}
-%& \quad \sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}\right)= \nonumber \\ 
-%&\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g\right) \\
-%&+\frac{1}{p}\sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \nonumber
-%\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}
+\section{TREAT M8-CAL Problem}

### 2017-02-07 doc/M&C_2017/FullPaper/MC2017_FullPaper.tex
Completed draft

@@
 \centering{\parbox{153mm}{{\bf Abstract} \itshape - 
-This template provides the standard format and styles of the full papers to be submitted to M\&C 2017.
-It is basically the same as the extended summary template except that the abstract is given and the section numbers are specified.
+The purpose of this paper is to present a new technique in dealing with multi-physics feedback in transient reactor simulations, specifically utilizing the improved quasi-static method (IQS).
+the technique involves the addition of an intermediate time scale to evaluate temperature and PRKE parameters.
+The theory is that temperature and the parameters are more quickly varying in time than the shape, thus further reducing the number of multi-group diffusion evaluations and execution time for a given accuracy.
+The technique was implemented into the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) and tested with the LRA benchmark and the Transient-15 TREAT problem.
+The results proved that the addition of this time scale was effective in reducing error and execution time for these problems.
 }\par}
@@
 The purpose of this paper is to introduce several techniques in dealing with multi-physics feedback for transient neutron diffusion calculations.
-In dynamic simulations, the neutron temporal distribution in a nuclear core can be strongly influenced by non-neutronics variables, e.g., temperature.The improved quasi-static method (IQS) is an effective technique for simulating the temporal behavior of the neutron flux in reactors.
+In dynamic simulations, the neutron temporal distribution in a nuclear core can be strongly influenced by non-neutronics variables, e.g., temperature.
+The improved quasi-static method (IQS) is an effective technique for simulating the temporal behavior of the neutron flux in reactors.
 Here, we present a study combining the IQS method with multi-physics solution techniques for coupled transient calculations.
@@
 The PRKE equations form a small ODE system and can be solved using a much smaller time step size.
-In transients where the shape varies much less than the flux, IQS can thus be very computationally effective.
+In transients where the shape varies much less than the flux, IQS can be very computationally effective.
 The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.
@@
 It is a super prompt-critical transient.
-The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL).
+The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.
 The spacial discretization was performed using continuous finite element method with first order Lagrangian basis functions.
@@
 Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.
-Third order Runge–Kutta discretization with step doubling adaptation was used for the PRKE evaluation.
+Third order Runge-Kutta discretization with step doubling adaptation was used for the PRKE evaluation.
 The performance of IQS and the temperature updates were measured by its improvement in accuracy at peak power over the Brute force technique.
@@
 
-In the temperature evaluation, a typical implicit solver would simply use the interpolated flux at end of the temperature time step for the right hand side of the equation.
+In the temperature evaluation, a typical implicit solver would simply use the interpolated flux at end of the temperature time step for the right-hand-side of the equation.
 However, IQS has much more information about the profile of the flux along the time step because of the micro-step amplitude evaluation.
@@
 \fig{fig:conv} is an error convergence plot comparing the three techniques where temperature is evaluated 5 times within a macro step (5 temperature updates).
-Finally, \fig{fig:mp} shows the affect of various temperature updates.
+Finally, \fig{fig:mp} shows the effect of various temperature updates.
 The dashed lines correspond to brute force at different flux step sizes, while the IQS macro step size is kept constant.
@@
 
-The convergence plots show that updating temperature and the PRKE parameters within a macro step has a significant affect on the performance of IQS.
+The convergence plots show that updating temperature and the PRKE parameters within a macro step has a significant effect on the performance of IQS.
 With only one update, IQS was only slightly better than brute force, brute force required about 150\% more time steps than IQS for the same error.
@@
 These run-times are based on total alive time of the execution where the diffusion evaluation is distributed over 24 processors.
+These run-times show a marginal performance for IQS and impressive performance for IQS P-C.
+Some of the execution times were able to decrease from implicit discretization with the same number of macro steps because IQS is better equipped to resolve the nonlinearity between temperature and amplitude.
+Furthermore, there does seem to be an ideal number of temperature updates to optimize execution time: IQS only needs one and IQS P-C seems to be ideal at 4 updates.
+This discrepancy in the number of updates shows that a adaptive type implementation of the updates would be ideal, and could enforce a constant error over the transient.
 \\
@@
 \hline
-1	& 1		& 2.612e-3 	&  `&	\\
-2	& 2		& 9.893e-4 	& 	& 	\\
-3 	& 4 	& 5.796e-4 	& 	&	\\
-4 	& 8 	& 4.772e-4 	& 	&	\\
-5 	& 16	& 4.516e-4 	& 	&	\\
+1	& 1		& 2.612e-3 	& 3.96 `& -3.18\%	\\
+2	& 2		& 9.893e-4 	& 6.02	&  47.1\%	\\
+3 	& 4 	& 5.796e-4 	& 7.87	&  92.3\%	\\
+4 	& 8 	& 4.772e-4 	& 12.61	& 207.9\% 	\\
+5 	& 16	& 4.516e-4 	& 22.14	& 440.7\%	\\
 \hline
@@
 \end{center}
-\caption{Traditional IQS run time results with $\Delta t = 0.004$}
+\vspace{-3mm}
+$^*$ difference in runtime from $\Delta t = 0.004$ implicit discretization 
+\caption{IQS run time results with $\Delta t = 0.004$}
 \label{tab:iqs_lra}
@@
 \hline
-1	& 1		& 3.488e-3 	& 6.24 `&	\\
-2	& 2		& 1.349e-3 	& 6.18	& 	\\
-3 	& 4 	& 9.161e-4 	& 7.05	&	\\
-4 	& 8 	& 8.052e-4 	& 7.72	&	\\
-5 	& 16	& 7.905e-4 	& 10.82	&	\\
+1	& 1		& 3.488e-3 	& 2.91 `& -28.9\%	\\
+2	& 2		& 1.349e-3 	& 3.73	& -9.00\%	\\
+3 	& 4 	& 9.161e-4 	& 3.97	& -3.04\%	\\
+4 	& 8 	& 8.052e-4 	& 5.39	&  31.7\%	\\
+5 	& 16	& 7.905e-4 	& 8.19	&  100\%	\\
 \hline
@@
 \end{center}
+\vspace{-3mm}
+$^*$ difference in runtime from $\Delta t = 0.004$ implicit discretization 
 \caption{IQS PC run time results with $\Delta t = 0.004$}
@@
 
-Transient 15 is a small core TREAT experiment that was modeled by the Rattlesnake group at INL for demonstrating their model's fidelity to TREAT’s feedback mechanisms.
+Transient 15 is a small core TREAT experiment that was modeled by the Rattlesnake group at INL for demonstrating their model's fidelity to TREAT's feedback mechanisms.
 The goal of the simulations is to obtain a power peak whose position and magnitude that sufficiently matches experimental results.
-The model involves an 11 energy group diffusion approximation and is discretized into 355,712 hexahedral continuous finite elements totaling 4,109,523 degrees of freedom.
+The model involves an 11-energy group diffusion approximation and is discretized into 355,712 hexahedral continuous finite elements totaling 4,109,523 degrees of freedom.
 The three second transient involves a linear ramp decrease in boron content throughout the fuel region.
@@
 
+In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time step, implicit discretization, IQS with one and 5 temperature updates per macro step, and IQS P-C with one and 5 updates.
 \fig{fig:Tran15_profile} shows the baseline power and temperature profile for the Transient-15 example.
+\tbl{tab:tran15} shows the error and runtime results.
 
@@
 
+\begin{table*}[htb]
+\begin{center}
+%\resizebox{\linewidth}{!}{
+\begin{tabular}{|l|cccccc|}
+\hline
+ & No.	& Max &	Time at Max &	Max Average &	\% Increase &	Max Power \\
+Method & of Steps &	Power (W) &	Power (s) &	Temperature (K) & Runtime$^*$ & Error \\
+\hline
+Baseline 			& 3000 	& 3.504e+08 & 1.9 	& 370 &	---		&	---		\\
+Implicit Dis. 		& 65 	& 3.424e+08 & 1.875 & 370 &	---		&	0.0230 	\\
+IQS 				& 61 	& 3.527e+08 & 1.9 	& 371 &	 17.9\%	&	0.00643 \\
+IQS (5 updates) 	& 61 	& 3.514e+08 & 1.9 	& 371 &	-19.2\%	&	0.00339	\\
+IQS P-C 			& 66 	& 4.340e+08 & 1.9 	& 374 &	 20.3\%	&	0.238	\\
+IQS P-C (5 updates) & 66 	& 3.537e+08 & 1.875 & 372 &  81.9\%	&	0.00935	\\
+\hline
+\end{tabular}
+%}
+\end{center}
+\vspace{-3mm}
+$^*$ difference in runtime from implicit discretization 
+\caption{Transient-15 Error and Runtime Results}
+\label{tab:tran15}
+\end{table*}
+
+The results from \tbl{tab:tran15} prove the performance of IQS with the temperature updates is slightly more ambiguous than the results from the LRA.
+IQS with 5 temperature updates shows a performance that reduces the error to approximately a tenth of the implicit discretization error, and reduces the execution time by about 20\%.
+This shows that the 5 updates were able to resolve the nonlinearity between flux and temperature with significantly less diffusion evaluations.
+However, IQS P-C did not perform nearly as well and proved almost detrimental to the problem.
+With only one update, IQS P-C achieved a less accurate solution and increased the execution time.
+Only with 5 updates was the error reduced from the implicit discretization error, but the increase in run-time was significant.
+\\
+
+%\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 IQS is a powerful tool for reactor simulation, it has the capability of greatly reducing computation time while retaining accuracy of relevant quantities.
-The goal of this summary is to demonstrate IQS's performance with multi-physics simulations, as well as techniques to improve its performance.
+The goal of this paper is to demonstrate IQS's performance with multi-physics simulations, as well as techniques to improve its performance.
 The technique focused on incorporating a new time scale in IQS for feedback quantities and PRKE parameters.
-The results when IQS is applied to the temperature feedback model of the LRA benchmark showed that incorporating a intermediate time scale for temperature has significant impact on IQS performance.
+The methodology is that for problems involving prompt-critical transients, temperature and its feedback to amplitude is more quickly varying than shape, but less time-dependent than amplitude itself.
+Therefore, an intermediate time scale was implemented, as well as a fixed-point iteration between amplitude and temperature.
+The attempt is to reduce the number of multi-group diffusion evaluations in a transient by reducing the number of time steps required for a given accuracy, and reduce the number of nonlinear iterations between flux/shape and temperature.
+Since multi-group diffusion evaluations are significantly more computationally expensive than temperature and amplitude evaluations, this result will show a decrease in execution time for a given accuracy.
 
-The full paper version will include time adaptive simulations for multiphysics dynamic transients: LRA benchmark, a full core TREAT model, and the Lady Godiva benchmark with structural feedback \cite{CritSafetyHandbook}.
-Performance of the IQS method with intermediate time scale for physics feedback will be compared to that of an implicit discretization of the flux equations with feedback.
+Two examples were chosen to test the implementation of the temperature feedback treatment: the LRA benchmark and the Transient-15 TREAT problem.
+Both problems involve adiabatic heat up of the core, the LRA has a constant specific heat, while Transient-15's is dependent on temperature.
+The LRA showed clear results concluding a marginal improvement in performance for IQS and a impressive performance for IQS P-C.
+The results proved that the updates improved the performance for both IQS and IQS P-C, but optimization of the number of updates is important, possibly requiring adaptation.
+The implementation to the Transient-15 problem gave more ambiguous results.
+IQS proved to significantly benefit from the updates, increasing accuracy and decreasing execution time.
+IQS P-C was able to reduce the error with the updates, but the execution time was significantly greater.
+
+In conclusion, the multi-physics treatment with IQS presented in this paper is worth further investigation, optimization, and implementation into other transient reactor problems with other feedback mechanisms.
 

### 2017-02-08 PetSc_mat_view/cross_section_mat.m
Trying two-grid

@@ -1,0 +1,35 @@
+function[g,f] = cross_section_mat()
+
+% Graphite
+g.v = [2957766058.45137 943013683.128542 173558681.925946 32455098.3714032 7324987.73064555 1923894.57827269 805568.086614681 458909.264460231 338055.975308392 247182.734780341 139673.973012195];
+profile = [1 1 1 2 1 3 1 4 4 6 5 7 6 10 6 11 7 11 7 11 7 11];
+value = [0.0669562 0.0337001 0.17454 6.05895e-05 0.0115368 0.256327 1.48423e-07 2.02861e-08 0.0120009 0.257911 0.0129102 0.256334 3.40804e-06 0.015042 0.255465 0.0010264 0.0159717 0.234959 0.00267275 3.35983e-05 2.38683e-06 2.84804e-05 0.0352161 0.236534 0.0269672 0.00404431 0.000820581 0.00077844 0.0284825 0.213021 0.025675 0.00427907 0.000302635 0.00646945 0.0372431 0.232662 0.0420183 3.68845e-05 0.000639437 0.0031687 0.020822 0.227415];
+g.ss = zeros(11);
+k=1;
+for i=2:2:length(profile)
+    g.ss(i/2,profile(i-1):profile(i)) = value(k:k+profile(i)-profile(i-1));
+    k = k+profile(i)-profile(i-1)+1;
+end
+g.str = [0.07648 0.152442 0.252388 0.25574 0.256325 0.256506 0.257873 0.263153 0.274103 0.275375 0.278619];
+g.sr = [0.034094 0.0115407 0.0120129 0.0129178 0.0150733 0.0161241 0.0376442 0.0387646 0.0680895 0.0514684 0.0487483];
+g.st = [0.10105 0.186081 0.26834 0.270829 0.271407 0.271589 0.272604 0.275299 0.28111 0.284131 0.276164];
+g.sa = [0.000327327 3.9339e-06 1.20226e-05 8.24054e-06 3.11858e-05 0.000119359 0.000284632 0.000499244 0.00067722 0.000924879 0.00163082];
+g.ngamma = [0.000327396 3.9339e-06 1.20226e-05 8.24054e-06 3.11858e-05 0.000119359 0.000284632 0.000499244 0.00067722 0.000924879 0.00163082];
+
+% Fuel
[…]

### 2017-02-08 PetSc_mat_view/two_step.m

@@ -1,0 +1,105 @@
+close all; clear;clc
+
+fd = PetscOpenFile('A');
+A = PetscBinaryRead(fd);
+
+fd = PetscOpenFile('D');
+D = PetscBinaryRead(fd);
+
+fd = PetscOpenFile('S');
+S = PetscBinaryRead(fd);
+
+fd = PetscOpenFile('F');
+F = PetscBinaryRead(fd);
+
+fd = PetscOpenFile('Dd');
+Dd = PetscBinaryRead(fd);
+
+G=11;
+[N,~]=size(A);
+n=N/G;
+
+k=1:G:N;
+if length(k)~=n
+    error('lengths');
+end
+
+AA = spalloc(N,N,nnz(A));
+row_id=0;
+for g=1:G
+    for i=1:n
+        ia=k(i)+(g-1);
+        row=[];
+        for gg=1:G
+            ja=k+(gg-1);
+            row = [row A(ia,ja)];
+        end
+        row_id=row_id+1;
+        AA(row_id,:)=row;
+    end
+end
+
+DD = spalloc(N,N,nnz(D));
+row_id=0;
+for g=1:G
+    for i=1:n
+        ia=k(i)+(g-1);
+        row=[];
+        for gg=1:G
+            ja=k+(gg-1);
+            row = [row D(ia,ja)];
+        end
+        row_id=row_id+1;
+        DD(row_id,:)=row;
+    end
+end
+
+SS = spalloc(N,N,nnz(S));
+row_id=0;
+for g=1:G
+    for i=1:n
+        ia=k(i)+(g-1);
+        row=[];
+        for gg=1:G
+            ja=k+(gg-1);
+            row = [row S(ia,ja)];
+        end
+        row_id=row_id+1;
+        SS(row_id,:)=row;
+    end
+end
+
+FF = spalloc(N,N,nnz(F));
+row_id=0;
+for g=1:G
+    for i=1:n
+        ia=k(i)+(g-1);
+        row=[];
+        for gg=1:G
+            ja=k+(gg-1);
+            row = [row F(ia,ja)];
+        end
+        row_id=row_id+1;
+        FF(row_id,:)=row;
+    end
+end
+
+DDd = spalloc(N,N,nnz(Dd));
+row_id=0;
+for g=1:G
+    for i=1:n
+        ia=k(i)+(g-1);
+        row=[];
+        for gg=1:G
+            ja=k+(gg-1);
[…]

### 2017-02-10 doc/Thesis/data/section3.tex
Thesis: Adding 1D Results

@@
 
+\fig{fig:1D_power} shows the resulting baseline relative power profile of the one-dimensional problem.
+The baseline was computed using MATLAB's ode15s function which is a embedded Runge-Kutta time adaptive method for stiff problems, the error tolerance was set very tightly ($10^{-12}$).
+This baseline computation is used to compute the error of the other time discretization methods.
+\fig{fig:1D_shape} shows the shape profile at various times during the transient.
+It is apparent that the shape is very time-dependent, so it is expected that IQS has marginal accuracy gain for a given time step.
+
+\begin{figure}[!htbp]
+\begin{center}
+\includegraphics[width=\linewidth]{\FiguresDir/1D_power_base.png}
+\caption{Baseline power profile of 1D slab}
+\label{fig:1D_power}
+\end{center}
+\end{figure}
+
+\begin{figure}[!htbp]
+\begin{center}
+\includegraphics[width=\linewidth]{\FiguresDir/1D_shape.png}
+\caption{Flux profile of 1D slab at various times}
+\label{fig:1D_shape}
+\end{center}
+\end{figure}
+
 \subsection{IQS Iteration Convergence}
@@
 \subsection{Time Step Convergence}
+
+In order to evaluate the performance and error convergence of IQS, the slab was simulated with varying time discretization methods and time step sizes.
+\fig{1D_conv} shows these convergence plots of five different discretization methods for implicit discretization, IQS, and IQS P-C.
+These plots were generated from the results using the MATLAB prototype program.
+The plots show that IQS and IQS P-C are convergent through fourth order BDF.
+Third order SDIRK did not show third order convergence, but, through extensive testing, SDIRK shows non-convergent behavior for too stiff of problems.
+
+There are higher discretization order that can be tested, but most practical application do not go beyond second order.
+This paper shows an analysis of the first publicized application of IQS with higher than second order discretization, which exposed unforeseen properties of IQS.
+When using higher order techniques, the interpolation of PRKE parameters and shape for precursor integration become important to consider.
+Every other application that was investigated linearly interpolates parameters for the PRKE evaluation.
+Similarly, the shape used for the integration of the ODE for the precursors needs to have higher order interpolation to preserve high order error convergence.
+This interpolation was done using Lagragian and Hermite methods, both leading to successful convergence.
+
+The 1D slab problem was also applied to the Rattlesnake implementation of IQS.
+\fig{fig:1D_conv_Rat} shows the error convergence for implicit Euler and BDF2 discretization of the three methods.
+The results differ slightly from the prototype, especially in the fact that IQS P-C performs better than IQS.
+Since Rattlesnake uses a PJFNK solver for the FEM model, number of time steps isn't strictly proportional to the execution time.
+\fig{fig:1D_conv_lin} shows the error of the three methods as compared to the number of linear GMRES iterations, which is a better mark for comparison in computation time.
+This figure shows that IQS performs worse than implicit discretization for most time-step sizes, but IQS P-C performs significantly better.
+This is due to the fact that IQS needs to iterate between amplitude and shape to resolve its nonlinearity.
+
+\begin{figure}[!htbp]
+\centering
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/1D_conv_IE.png}
+\caption{Implicit Euler}
+\end{subfigure}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/1D_conv_BDF2.png}
+\caption{BDF2}
+\end{subfigure}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/1D_conv_BDF3.png}
+\caption{BDF3}
+\end{subfigure}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/1D_conv_BDF4.png}
+\caption{BDF4}
+\end{subfigure}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/1D_conv_SDIRK33.png}
+\caption{SDIRK33}
+\end{subfigure}
+\caption{Error convergence plots of implicit discretization, IQS, and IQS P-C with various time discretization schemes}
+\label{fig:1D_conv}
+\end{figure}
+
+\begin{figure}[!htbp]
+\begin{center}
+\includegraphics[width=0.75\linewidth]{\FiguresDir/1D_conv_Rat.png}
+\caption{Error convergence plots of implicit discretization, IQS, and IQS P-C from Rattlesnake implementation}
+\label{fig:1D_conv_Rat}
+\end{center}
+\end{figure}
+
+\begin{figure}[!htbp]
+\begin{center}
+\includegraphics[width=0.75\linewidth]{\FiguresDir/1D_conv_lin.png}
+\caption{Error convergence plots of implicit discretization, IQS, and IQS P-C vs. number of GMRES iterations}
+\label{fig:1D_conv_lin}
+\end{center}
+\end{figure}
 

### 2017-02-16 doc/M&C_2017/FullPaper/MC2017_FullPaper.tex
Changes to language and addition of time scale theory and plots

@@
 \begin{figure}[!htbp]
+\centering
 \includegraphics[width=\linewidth]{IQS_visualization.jpg}
@@
 \centering
+\resizebox{\linewidth}{!}{%
 \begin{tikzpicture}[scale=1]
@@
 
-\end{tikzpicture}
+\end{tikzpicture}}
 \caption{Time scales and process of IQS}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Dynamical Time Scale Analysis}
+\label{sect:tau}
+
+The temporal variance of flux, shape, temperature, and amplitude can be quantified by defining a dynamical time scale ($\tau$) for each physics.
+A small value $\tau$ means the variable is quickly varying in time and consequently needs relatively small time steps for accuracy, vice-versa for large $\tau$.
+The general definition of $tau$ is defined by \eqt{eq:tau1}, where $\theta$ is the variable of interest.
+Since each variable in discretized in time, a finite difference approximation will be made for the $\frac{d\theta}{dt}$ term and the average between the two corresponding time steps will be made for the $\frac{1}{\theta}$ term.
+Additionally, $\tau$ is spatially dependent for flux and temperature,  but only the time dependent behaivior of this quantity is of interest.
+Therefore, the $L^2$ norm of each term will be used to compute the approximate time scale ($\tilde{\tau}$), formally defined by \eqt{eq:tau2}.
+$\theta$ represents a summation over groups for flux and shape.
+
+\be
+\tau = \frac{1}{\left|\frac{1}{\theta}\frac{d\theta}{dt}\right|}
+\label{eq:tau1}
+\ee
+
+\be
+\tilde{\tau}_{n+1} = \frac{\norm{\theta_{n+1} + \theta_{n}}/2}{\norm{\theta_{n+1} - \theta_{n}}/\Delta t}
+\label{eq:tau2}
+\ee
+
+According to the a priori hypothesis from previous sections, $\tau$ is large for shape, somewhat smaller for temperature, and very small for amplitude and flux.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results and Analysis}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+.
 To test the multiple time scale implementation of IQS, the LRA benchmark was chosen for reference and a TREAT example was employed.
@@
 
-Any interpolation of the amplitude along the micro steps is possible for the integration, this application uses piece-wise linear.
+Any interpolation of the amplitude function $p(t)$ along the micro steps is possible to carry out the above macro-step integrals. Here, we use piece-wise linear interpolation.
 
@@
 \begin{figure}[htbp!]
+\centering
 \includegraphics[width=\linewidth]{lra_profile.png}
@@
 \hline
-No. of Mesh 			& 3872 		& 1936 \\
+No. of Spatial Nodes	& 3872 		& 1936 \\
 Eigenvalue 				& 0.99637	& 0.99637 \\
@@
 \begin{figure}[!htpb]
+\centering
 \includegraphics[width=\linewidth]{lra_bad.jpg}
@@
 This convergence makes sense because temperature can only be so accurate before the error in shape is dominating.
-Table \ref{tab:ndiff_lra} shows the run time results for the implicit discretization calculations.
+Table \ref{tab:ndiff_lra} shows the run time results for the brute force calculations.
 Tables \ref{tab:iqs_lra} and \ref{tab:iqspc_lra} present the IQS run-times with various numbers of temperature updates.
@@
 These run-times show a marginal performance for IQS and impressive performance for IQS P-C.
-Some of the execution times were able to decrease from implicit discretization with the same number of macro steps because IQS is better equipped to resolve the nonlinearity between temperature and amplitude.
+Some of the execution times were able to decrease from brute force with the same number of macro steps because IQS is better equipped to resolve the nonlinearity between temperature and amplitude.
 Furthermore, there does seem to be an ideal number of temperature updates to optimize execution time: IQS only needs one and IQS P-C seems to be ideal at 4 updates.
 This discrepancy in the number of updates shows that a adaptive type implementation of the updates would be ideal, and could enforce a constant error over the transient.
+It is also important to compare the error of brute force with IQS at one update and IQS P-C at 4 updates.
+IQS shows an error comparable to brute force at $\Delta t = 0.002$, signifying an actual increase in runtime by -34.1\%.
+IQS P-C shows an error less than brute force at $\Delta t = 0.002$, signifying an actual increase in runtime by <-34.9\%.
 \\
@@
 %\resizebox{\linewidth}{!}{
-\begin{tabular}{|l|l|cc|}
+\begin{tabular}{|l|l|ccc|}
 \hline
-Run  &  $\Delta t$ & Error & Runtime (hr) \\
+Run  &  $\Delta t$ & Error & Runtime (hr) & Linear Iter.\\
 \hline
-1	& 4.0e-3	& 1.407e-2 	& 4.11	\\
-2	& 2.0e-3	& 3.174e-3 	& 5.68 	\\
-3 	& 1.0e-3 	& 7.690e-4 	& 5.55	\\
-4 	& 5.0e-4 	& 1.892e-4 	& 13.57	\\
-5 	& 2.5e-4	& 4.590e-5 	& 23.81	\\
+1	& 4.0e-3	& 1.407e-2 	& 4.11	& 7.13e4	\\
+2	& 2.0e-3	& 3.174e-3 	& 6.01	& 9.49e4 	\\
+3 	& 1.0e-3 	& 7.690e-4 	& 10.38	& 1.45e5	\\
+4 	& 5.0e-4 	& 1.892e-4 	& 21.91	& 2.08e5	\\
+5 	& 2.5e-4	& 4.590e-5 	& 25.23	& 3.16e5	\\
 \hline
@@
 \end{center}
-\caption{Implicit discretization run time results}
+\caption{Brute force run time results}
 \label{tab:ndiff_lra}
@@
 \hline
-1	& 1		& 2.612e-3 	& 3.96 `& -3.18\%	\\
+1	& 1		& 2.612e-3 	& 3.96 	& -3.18\%	\\
 2	& 2		& 9.893e-4 	& 6.02	&  47.1\%	\\
@@
 \vspace{-3mm}
-$^*$ difference in runtime from $\Delta t = 0.004$ implicit discretization 
+$^*$ difference in runtime from $\Delta t = 0.004$ brute force 
 \caption{IQS run time results with $\Delta t = 0.004$}
@@
 \hline
-1	& 1		& 3.488e-3 	& 2.91 `& -28.9\%	\\
+1	& 1		& 3.488e-3 	& 2.91 	& -28.9\%	\\
 2	& 2		& 1.349e-3 	& 3.73	& -9.00\%	\\
@@
 \vspace{-3mm}
-$^*$ difference in runtime from $\Delta t = 0.004$ implicit discretization 
+$^*$ difference in runtime from $\Delta t = 0.004$ brute force 
 \caption{IQS PC run time results with $\Delta t = 0.004$}
@@
 
+The performance of the quasi-statics can also be explained by the computation of the dynamical time scale described by \sct{sect:tau}.
+\fig{fig:LRAtc} shows the time scale profile over the transient, computed using \eqt{eq:tau2}.
+This plot shows that in a brute force simulation, the flux dominates the time dependent behaivior, while temperature lags in its variance for the majority of the transient.
+In an IQS simulation, the time scale behavior of amplitude almost exactly matches the flux, while shape is more varying than temperature throughout most of the transient.
+The large $\tau$ for temperature during the beginning of the transient shows that adaptation of the number of updates is important; computational expense on temperature evaluations is being wasted during this time.
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=\linewidth]{time_constant_lra.png}
+\caption{Dynamical time scale for LRA benchmark}
+\label{fig:LRAtc}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Transient 15 is a small core TREAT experiment that was modeled by the Rattlesnake group at INL for demonstrating their model's fidelity to TREAT's feedback mechanisms.
-The goal of the simulations is to obtain a power peak whose position and magnitude that sufficiently matches experimental results.
-The model involves an 11-energy group diffusion approximation and is discretized into 355,712 hexahedral continuous finite elements totaling 4,109,523 degrees of freedom.
-The three second transient involves a linear ramp decrease in boron content throughout the fuel region.
+Transient 15 is a test case based on the TREAT core.
+The purpose of the original creation of this simulation in Rattlesnake is to test the model's fidelity with the thermal feedback of TREAT, but it is not meant to exactly match any previous experiments.
+Nevertheless, the goal of the following simulations is to test IQS and its time scale based treatment of temperature with a more complex model.
+Transient 15 involves an 11-energy group diffusion approximation and is discretized into $355,712$ hexahedral continuous finite elements totaling $4,109,523$ degrees of freedom.
+The three-second transient involves a linear ramp decrease in the absorption cross section throughout the control rod region.
 \fig{fig:Tran15} shows a visualization of the flux profile within the core, hidden is the massive amount of graphite surrounding the core.
@@
 
-In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time step, implicit discretization, IQS with one and 5 temperature updates per macro step, and IQS P-C with one and 5 updates.
+In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time step, brute force, IQS with one and 5 temperature updates per macro step, and IQS P-C with one and 5 updates.
 \fig{fig:Tran15_profile} shows the baseline power and temperature profile for the Transient-15 example.
@@
 %\resizebox{\linewidth}{!}{
-\begin{tabular}{|l|cccccc|}
+\begin{tabular}{|l|ccccccc|}
 \hline
- & No.	& Max &	Time at Max &	Max Average &	\% Increase &	Max Power \\
-Method & of Steps &	Power (W) &	Power (s) &	Temperature (K) & Runtime$^*$ & Error \\
+ & No.	& Max &	Time at Max &	Max Average &	\% Increase &	Max Power & Linear	\\
+Method & of Steps &	Power (W) &	Power (s) &	Temperature (K) & Runtime$^*$ & Error & Iterations \\
 \hline
-Baseline 			& 3000 	& 3.504e+08 & 1.9 	& 370 &	---		&	---		\\
-Implicit Dis. 		& 65 	& 3.424e+08 & 1.875 & 370 &	---		&	0.0230 	\\
-IQS 				& 61 	& 3.527e+08 & 1.9 	& 371 &	 17.9\%	&	0.00643 \\
-IQS (5 updates) 	& 61 	& 3.514e+08 & 1.9 	& 371 &	-19.2\%	&	0.00339	\\
-IQS P-C 			& 66 	& 4.340e+08 & 1.9 	& 374 &	 20.3\%	&	0.238	\\
-IQS P-C (5 updates) & 66 	& 3.537e+08 & 1.875 & 372 &  81.9\%	&	0.00935	\\
+Baseline 			& 3000 	& 3.5039e+08 & 1.901 & 371 &	---		&	---		 & ---		\\
+Brute Force 		& 300 	& 3.5011e+08 & 1.90  & 371 &	---		&	7.875e-4 & 41020	\\
+IQS 				& 300	& 3.5036e+08 & 1.90  & 371 & -11.9\%	&	8.385e-5 & 23949	\\
+IQS (5 updates) 	& 300 	& 3.5040e+08 & 1.90  & 371 &  49.7\%	&	3.687e-5 & 24035	\\
+IQS P-C 			& 66 	& 4.3400e+08 & 1.90  & 374 &  20.3\%	&	0.238	 & 12318	\\
+IQS P-C (5 updates) & 66 	& 3.5370e+08 & 1.90  & 372 &  81.9\%	&	0.00935	 & 15598	\\
 \hline
@@
 \vspace{-3mm}
-$^*$ difference in runtime from implicit discretization 
+$^*$ difference in runtime from brute force 
 \caption{Transient-15 Error and Runtime Results}
@@
 
-The results from \tbl{tab:tran15} prove the performance of IQS with the temperature updates is slightly more ambiguous than the results from the LRA.
-IQS with 5 temperature updates shows a performance that reduces the error to approximately a tenth of the implicit discretization error, and reduces the execution time by about 20\%.
-This shows that the 5 updates were able to resolve the nonlinearity between flux and temperature with significantly less diffusion evaluations.
+The results from \tbl{tab:tran15} show similar performance of IQS with the temperature updates as the LRA.
+IQS with 1 temperature update shows a performance that reduces the error to approximately a tenth of the brute force error, and reduces the execution time by about 12\%.
+This shows that IQS was able to resolve the nonlinearity between flux and temperature with significantly less diffusion evaluations.
+Having IQS with 5 updates significantly increased the execution time for the same time step, but the error was reduced.
+Comparing this error to a similar brute force error at a smaller time step could show that the runtime was reduced.
 However, IQS P-C did not perform nearly as well and proved almost detrimental to the problem.
 With only one update, IQS P-C achieved a less accurate solution and increased the execution time.
-Only with 5 updates was the error reduced from the implicit discretization error, but the increase in run-time was significant.
+Only with 5 updates was the error reduced from the brute force error, but the increase in run-time was significant.
+The transient profile of the varibles' dynamical time scales is shown in \fig{fig:tran15tc}.
+This plot exhibits a similar response to that of the LRA.
+The response of temperature shows that the updates are a computational frugal treatment of the feedback and adaptation of the number of updates is vital for optimization.
 \\
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=\linewidth]{time_constant_tran15.png}
+\caption{Dynamical time scale for the Transient 15 example}
+\label{fig:tran15tc}
+\end{figure}
 

### 2017-02-16 doc/M&C_2017/FullPaper/MC2017_FullPaper.tex
Adding more results

@@
 The technique was implemented into the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) and tested with the LRA benchmark and the Transient-15 TREAT problem.
-The results proved that the addition of this time scale was effective in reducing error and execution time for these problems.
+The results proved that the addition of this time scale was effective in reducing execution time for these problems while maintaining accuracy.
 }\par}
@@
 IQS (5 updates) 	& 300 	& 3.5040e+08 & 1.90  & 371 &  49.7\%	&	3.687e-5 & 24035	\\
-IQS P-C 			& 66 	& 4.3400e+08 & 1.90  & 374 &  20.3\%	&	0.238	 & 12318	\\
-IQS P-C (5 updates) & 66 	& 3.5370e+08 & 1.90  & 372 &  81.9\%	&	0.00935	 & 15598	\\
+IQS P-C 			& 300 	& 3.5065e+08 & 1.90  & 371 &  -2.1\%	&	7.527e-4 & 39020	\\
+IQS P-C (5 updates) & 300 	& 3.5043e+08 & 1.90  & 371 &  26.5\%	&	1.227e-4 & 37866	\\
 \hline
@@
 Comparing this error to a similar brute force error at a smaller time step could show that the runtime was reduced.
-However, IQS P-C did not perform nearly as well and proved almost detrimental to the problem.
-With only one update, IQS P-C achieved a less accurate solution and increased the execution time.
-Only with 5 updates was the error reduced from the brute force error, but the increase in run-time was significant.
+IQS P-C performed not nearly as well as it did with the LRA benchmark, but still proved to be effective.
+Having 5 updates for IQS P-C increased the runtime marginally, but decreased the error significantly.
 The transient profile of the varibles' dynamical time scales is shown in \fig{fig:tran15tc}.
@@
 The results proved that the updates improved the performance for both IQS and IQS P-C, but optimization of the number of updates is important, possibly requiring adaptation.
-The implementation to the Transient-15 problem gave more ambiguous results.
-IQS proved to significantly benefit from the updates, increasing accuracy and decreasing execution time.
-IQS P-C was able to reduce the error with the updates, but the execution time was significantly greater.
+The implementation to the Transient-15 problem gave similar results.
+IQS proved to marginally benefit from the updates, decreasing error but increasing execution time.
+IQS P-C was benefited more by the updates in this example.
+The updates were able to reduce the error with a marginal increase in execution time.
 
-In conclusion, the multi-physics treatment with IQS presented in this paper is worth further investigation, optimization, and implementation into other transient reactor problems with other feedback mechanisms.
+In conclusion, the multi-physics treatment with IQS presented in this paper is worth further investigation, optimization, and implementation into other transient reactor problems with more feedback mechanisms.
 

### 2017-02-17 doc/M&C_2017/FullPaper/MC2017_FullPaper.tex
Final changes before Ragusa proof-read

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-.
+
 To test the multiple time scale implementation of IQS, the LRA benchmark was chosen for reference and a TREAT example was employed.
+The performance of the implementation is evaluated comparing errors in power at peak points with brute force and IQS without quasi-static temperature treatment.
+Execution times were also compared at given error values to judge the computational efficiency of adding more temperature evaluations.
+These examples were executed using the MOOSE/Rattlesnake framework at INL, which uses a PJFNK (preconditioned Jacobian-free Newton Krylov) method to evaluate the nonlinear problems.
+Each time step evaluation consists of three iteration loops: GMRES iterations for the linear system, Newton iterations for the temperature-flux nonlinearity, and Picard iterations for the shape-amplitude nonlinearity.
+Another judge of computational expense is the total number of linear iterations during the execution, these are shown specifically for the brute force LRA simulations and all the simulations for the TREAT example.
 
@@
 
-\fig{fig:lra_profile} shows the baseline power and temperature profile for the LRA benchmark.
+\fig{fig:lra_profile} shows the baseline power and temperature transient profile for the LRA benchmark, and \fig{fig:lra_profile2D} shows the spacial power distribution at the peak power.
 The baseline results are compared to the results achieved by Sutton and Aviles in \cite{Sutton_1996} and presented in \tbl{tab:base}.
@@
 \label{fig:lra_profile}
+\end{figure}
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[width=\linewidth]{lra_profile2D.png}
+\caption{LRA baseline spacial power profile at $t=1.44 s$}
+\label{fig:lra_profile2D}
 \end{figure}

### 2017-02-18 doc/M&C_2017/FullPaper/MC2017_FullPaper.tex
updated abstract

@@
 \centering{\parbox{153mm}{{\bf Abstract} \itshape - 
-The purpose of this paper is to present a new technique in dealing with multi-physics feedback in transient reactor simulations, specifically utilizing the improved quasi-static method (IQS).
-the technique involves the addition of an intermediate time scale to evaluate temperature and PRKE parameters.
-The theory is that temperature and the parameters are more quickly varying in time than the shape, thus further reducing the number of multi-group diffusion evaluations and execution time for a given accuracy.
-The technique was implemented into the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) and tested with the LRA benchmark and the Transient-15 TREAT problem.
-The results proved that the addition of this time scale was effective in reducing execution time for these problems while maintaining accuracy.
+In this paper, we  present a tightly coupled technique to solve multi-physics transient reactor simulations, whereby the neutronic solution  is based on the improved quasi-static method (IQS).
+In IQS, the flux is typically factored into components, shape and amplitude, have greatly varying time scales.
+When multiphysic simulations are considered,
+an additional intermediate time scale to evaluate feedback (temperature and PRKE parameters) is used.
+The rationale for the intermediate time scale is based on the observation that temperature and point-reactor parameters may vary quickly in time than the shape itself.
+The method was implemented in the Rattlesnake/MOOSE framework of Idaho National Laboratory (INL) and tested with the LRA benchmark and the Transient-15 TREAT problem.
+The results show that the addition of this time scale was effective in reducing execution time for these problems while yielding accurate answers.
 }\par}

### 2017-02-18 doc/M&C_2017/FullPaper/MC2017_FullPaper.tex
final version

@@
 In dynamic simulations, the neutron temporal distribution in a nuclear core can be strongly influenced by non-neutronics variables, e.g., temperature.
-The improved quasi-static method (IQS) is an effective technique for simulating the temporal behavior of the neutron flux in reactors.
+The improved quasi-static method (IQS) is an effective technique for simulating the kinetic behavior of the neutron flux in reactors.
 Here, we present a study combining the IQS method with multi-physics solution techniques for coupled transient calculations.
 
-IQS is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (shape) and a time-dependent component (amplitude) \cite{Ott_1966,Dulla2008}.
+IQS is a transient spatial kinetic method that involves factorizing the neutron flux into a space- and time-dependent component (shape) and a time-dependent component (amplitude) \cite{Ott_1966,Dulla2008}.
 The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates.
 IQS has largely been applied to neutron kinetics, without other feedback mechanisms.
-This paper presents the application of multi-physics feedback with IQS and analyzes performance with temperature feedback benchmark problems.
+This paper presents the application of multi-physics feedback with IQS and analyzes performance with temperature feedback problems.
 
@@
 The intention of this implementation is to further optimize solution accuracy with computational effort.
-In order to evaluate the performance of the methodology, it will be tested with two temperature feedback problems: the LRA benchmark and an example from Transient Reactor Testing Facility (TREAT) reactor.
+In order to evaluate the performance of the methodology, it will be tested with two temperature feedback problems: the LRA benchmark and an example from the Transient Reactor Testing Facility (TREAT) reactor.
 The performance is quantified by comparing accuracies with traditional implementations of neutron dynamics and IQS.
@@
 \ee
-
-Where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the streaming and interaction operator, 
+%
+where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the diffusion operator, 
 $S_{d}^g$ is the delay source, and $P_{d,i}^g$ is the delayed-neutron fission operator.
@@
 Note that the time-dependent shape equation is similar to the time-dependent flux equation. 
-However, the shape/amplitude equations are now non-linearly coupled (boxed terms).
+However, the shape equations are now non-linearly coupled (boxed terms) to the amplitude equations.
 
@@
 In order to impose uniqueness of the factorization, one requires $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)$ to be constant.
-And after some manipulation, the standard point reactor kinetics equations (PRKE) for the amplitude solution are obtained:
+After some manipulation, the point reactor kinetics equations (PRKE) for the amplitude solution are obtained:
 \be
@@
 \ee
-Where the functional coefficients are calculated using the space-/time-dependent shape function as follows:
+where the functional coefficients are calculated using the space-/time-dependent shape function as follows:
 \be
@@
 
-It is important to note that the PRKE parameters are evaluated at each macro step and linearly interpolated for the PRKE evaluation.
+It is important to note that the PRKE parameters are evaluated at each macro step and (linearly) interpolated for the PRKE evaluation.
 In order to preserve the error convergence of high order discretization schemes for shape, higher order interpolation of the parameters is required.
@@
 For IQS, this feedback affects both the shape equation and the parameters of the PRKE; thus, it is an additional nonlinear component to the 
-already non-linear shape-amplitude equations.  Each of these components have different temporal dependencies; so it may be beneficial for 
+already non-linear shape-amplitude equations.  Each of these components have different temporal behaviors; so it may be beneficial for 
 efficiency to evaluate them on different time scales.  The amplitude is more rapidly varying than the shape which is computationally expensive to evaluate
@@
 
-The top time scale represents a shape diffusion evaluation on a macro step, the middle has an arbitrary three steps within the macro step where temperature and the PRKE parameters are evaluated, and the bottom one represents the PRKE evaluation on micro steps.
+The top time scale represents a shape diffusion evaluation on a macro step, the middle has an arbitrary three steps 
+(for illustration) within the macro step where temperature and the PRKE parameters are evaluated, and the bottom one represents the PRKE evaluation on micro steps.
 The shape is linearly interpolated within the macro step for the temperature and PRKE parameter evaluation, and the parameters are interpolated within the temperature step for the PRKE evaluation.
@@
 A small value $\tau$ means the variable is quickly varying in time and consequently needs relatively small time steps for accuracy, vice-versa for large $\tau$.
-The general definition of $tau$ is defined by \eqt{eq:tau1}, where $\theta$ is the variable of interest.
-Since each variable in discretized in time, a finite difference approximation will be made for the $\frac{d\theta}{dt}$ term and the average between the two corresponding time steps will be made for the $\frac{1}{\theta}$ term.
-Additionally, $\tau$ is spatially dependent for flux and temperature,  but only the time dependent behaivior of this quantity is of interest.
-Therefore, the $L^2$ norm of each term will be used to compute the approximate time scale ($\tilde{\tau}$), formally defined by \eqt{eq:tau2}.
-$\theta$ represents a summation over groups for flux and shape.
-
+The general definition of $\tau$ is defined by \eqt{eq:tau1}, where $\theta$ is the physic component of interest.
 \be
@@
 \ee
-
+Since each variable in discretized in time, a finite difference approximation will be made for the $\frac{d\theta}{dt}$ term and the average between the two corresponding time steps will be made for the $\frac{1}{\theta}$ term.
+Additionally, $\tau$ is spatially dependent for flux and temperature,  but only the time dependent behavior of this quantity is of interest.
+Therefore, the $L^2$ norm of each term will be used to compute the approximate time scale ($\tilde{\tau}$), formally defined by \eqt{eq:tau2}.
+$\theta$ represents a summation over groups for flux and shape.
 \be
-\tilde{\tau}_{n+1} = \frac{\norm{\theta_{n+1} + \theta_{n}}/2}{\norm{\theta_{n+1} - \theta_{n}}/\Delta t}
+\tilde{\tau}_{n+1} = \frac{\norm{\theta_{n+1} + \theta_{n}}}{2}\frac{\Delta t}{\norm{\theta_{n+1} - \theta_{n}}}
 \label{eq:tau2}
 \ee
-
-According to the a priori hypothesis from previous sections, $\tau$ is large for shape, somewhat smaller for temperature, and very small for amplitude and flux.
+%
+According to the a priori hypothesis from previous sections, $\tau$ is large for shape, somewhat smaller for temperature, and much smaller for amplitude and flux.
 
@@
 
-To test the multiple time scale implementation of IQS, the LRA benchmark was chosen for reference and a TREAT example was employed.
-The performance of the implementation is evaluated comparing errors in power at peak points with brute force and IQS without quasi-static temperature treatment.
+To test the multiple time scale implementation of IQS, the LRA benchmark
+\cite{ANL_BPB} and a TREAT reactor example have been selected.
+The performance of the implementation is evaluated comparing errors in power at peak time with standard flux solution techniques (``brute force'') and IQS without quasi-static temperature treatment.
 Execution times were also compared at given error values to judge the computational efficiency of adding more temperature evaluations.
-These examples were executed using the MOOSE/Rattlesnake framework at INL, which uses a PJFNK (preconditioned Jacobian-free Newton Krylov) method to evaluate the nonlinear problems.
+These examples were executed using INL's MOOSE/Rattlesnake framework, which uses a PJFNK (preconditioned Jacobian-free Newton Krylov) method to evaluate the nonlinear problems.
 Each time step evaluation consists of three iteration loops: GMRES iterations for the linear system, Newton iterations for the temperature-flux nonlinearity, and Picard iterations for the shape-amplitude nonlinearity.
@@
 
-The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.
+The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor.
 It is a super prompt-critical transient.
@@
 The spacial discretization was performed using continuous finite element method with first order Lagrangian basis functions.
-The mesh consisted of blocks 11X11 with five uniform refinements, totaling 165,165 elements and 124,609 nodes.
-Three different temporal techniques were applied: implicit discretization of the flux equation (Brute force), IQS, and IQS-PC.
+The mesh consisted of blocks $11\times 11$ with five uniform refinements, totaling $165,165$ elements and $124,609$ nodes.
+Three different temporal techniques were applied: implicit discretization of the flux equation (``brute force''), IQS, and IQS-PC.
 Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.
@@
 \ee
-Where $n$ corresponds to the beginning of the temperature step.  $a_1$ and $a_2$ are integration coefficients defined by \eqt{eq:a1} and \eqt{eq:a2}.
+where $n$ corresponds to the beginning of the temperature step.  $a_1$ and $a_2$ are integration coefficients defined by \eqt{eq:a1} and \eqt{eq:a2}.
 \be
@@
 
-\fig{fig:lra_profile} shows the baseline power and temperature transient profile for the LRA benchmark, and \fig{fig:lra_profile2D} shows the spacial power distribution at the peak power.
+\fig{fig:lra_profile} shows the baseline power and temperature transient profile for the LRA benchmark.
+\fig{fig:lra_profile2D} shows the spacial power distribution at the peak power.
 The baseline results are compared to the results achieved by Sutton and Aviles in \cite{Sutton_1996} and presented in \tbl{tab:base}.
@@
 \includegraphics[width=\linewidth]{lra_bad.jpg}
-\caption{LRA convergence plots with only one temperature update each macro step}
+\caption{LRA convergence plots with only one temperature update per macro step}
 \label{fig:lra_bad}
@@
 IQS with 1 temperature update shows a performance that reduces the error to approximately a tenth of the brute force error, and reduces the execution time by about 12\%.
-This shows that IQS was able to resolve the nonlinearity between flux and temperature with significantly less diffusion evaluations.
+This shows that IQS was able to resolve the nonlinearity between flux and temperature with significantly fewer diffusion evaluations.
 Having IQS with 5 updates significantly increased the execution time for the same time step, but the error was reduced.
@@
 The technique focused on incorporating a new time scale in IQS for feedback quantities and PRKE parameters.
-The methodology is that for problems involving prompt-critical transients, temperature and its feedback to amplitude is more quickly varying than shape, but less time-dependent than amplitude itself.
+The rationale is that for problems involving prompt-critical transients, temperature is more quickly varying than shape, but still slower than the amplitude itself.
 Therefore, an intermediate time scale was implemented, as well as a fixed-point iteration between amplitude and temperature.
@@
 
-In conclusion, the multi-physics treatment with IQS presented in this paper is worth further investigation, optimization, and implementation into other transient reactor problems with more feedback mechanisms.
+In conclusion, the multi-physics treatment with IQS presented in this paper is worth further investigation, optimization, and testing with additional transient reactor problems with other feedback mechanisms.
 

### 2017-02-18 doc/M&C_2017/FullPaper/MC2017_FullPaper.tex
final version

@@
 In IQS, the flux is typically factored into components, shape and amplitude, have greatly varying time scales.
-When multiphysic simulations are considered,
-an additional intermediate time scale to evaluate feedback (temperature and PRKE parameters) is used.
+When multiphysic simulations are considered, an additional intermediate time scale to evaluate feedback (temperature and PRKE parameters) is used.
 The rationale for the intermediate time scale is based on the observation that temperature and point-reactor parameters may vary quickly in time than the shape itself.
@@
 Other physical quantities, such as temperature, are affected by reaction rates and subsequently affect the operators of the flux equations.
-For IQS, this feedback affects both the shape equation and the parameters of the PRKE; thus, it is an additional nonlinear component to the 
-already non-linear shape-amplitude equations.  Each of these components have different temporal behaviors; so it may be beneficial for 
-efficiency to evaluate them on different time scales.  The amplitude is more rapidly varying than the shape which is computationally expensive to evaluate
-and is evaluated only on macro-time steps. In multiphysics simulations, one may take advantage of a fine-scale power distribution in the coupled
-physics components (temperature). Figure~\ref{fig:time} shows a such a solution process for temperature feedback.
+For IQS, this feedback affects both the shape equation and the parameters of the PRKE; thus, it is an additional nonlinear component to the already non-linear shape-amplitude equations.
+Each of these components have different temporal behaviors; so, it may be beneficial for efficiency to evaluate them on different time scales.
+The amplitude is more rapidly varying than the shape which is computationally expensive to evaluate and is evaluated only on macro-time steps.
+In multiphysics simulations, one may take advantage of a fine-scale power distribution in the coupled physics components (temperature).
+Figure~\ref{fig:time} shows a such a solution process for temperature feedback.
 
@@
 
-The top time scale represents a shape diffusion evaluation on a macro step, the middle has an arbitrary three steps 
-(for illustration) within the macro step where temperature and the PRKE parameters are evaluated, and the bottom one represents the PRKE evaluation on micro steps.
+The top time scale represents a shape diffusion evaluation on a macro step, the middle has an arbitrary three steps (for illustration) within the macro step where temperature and the PRKE parameters are evaluated, and the bottom one represents the PRKE evaluation on micro steps.
 The shape is linearly interpolated within the macro step for the temperature and PRKE parameter evaluation, and the parameters are interpolated within the temperature step for the PRKE evaluation.
@@
 
-To test the multiple time scale implementation of IQS, the LRA benchmark
-\cite{ANL_BPB} and a TREAT reactor example have been selected.
+To test the multiple time scale implementation of IQS, the LRA benchmark \cite{ANL_BPB} and a TREAT reactor example have been selected.
 The performance of the implementation is evaluated comparing errors in power at peak time with standard flux solution techniques (``brute force'') and IQS without quasi-static temperature treatment.
@@
 \fig{fig:LRAtc} shows the time scale profile over the transient, computed using \eqt{eq:tau2}.
-This plot shows that in a brute force simulation, the flux dominates the time dependent behaivior, while temperature lags in its variance for the majority of the transient.
+This plot shows that in a brute force simulation, the flux dominates the time dependent behavior, while temperature lags in its variance for the majority of the transient.
 In an IQS simulation, the time scale behavior of amplitude almost exactly matches the flux, while shape is more varying than temperature throughout most of the transient.
@@
 Having 5 updates for IQS P-C increased the runtime marginally, but decreased the error significantly.
-The transient profile of the varibles' dynamical time scales is shown in \fig{fig:tran15tc}.
+The transient profile of the variables' dynamical time scales is shown in \fig{fig:tran15tc}.
 This plot exhibits a similar response to that of the LRA.
@@
 This work was supported by the Department of Energy, Idaho National Laboratory, and the Integrated University Program Fellowship.
-We thank Mark Dehart, Yaqi Wang, the NEAMS program, and INL's Moose/Rattlesnake team for their support.
+We thank Mark Dehart, Yaqi Wang, the NEAMS program, and INL's MOOSE/Rattlesnake team for their support.
 

### 2017-02-21 doc/Thesis/data/section1.tex
some initial comments to the draft

@@
 
-Transient modeling of nuclear reactors has been a chronically foreboding task due to its computationally expensive nature.
-However, recent developments in nuclear testing, including the revitalization of the Transient Reactor Testing (TREAT) Facility at Idaho National Laboratory (INL), have brought significant attention to the development of transient reactor modeling.
+\tcr{Advice: rename through git the files section1 ... with more meaningful filenames, e.g., introduction, theory, ...}
+
+\tcr{Advice: use bibtex for bibliography in the future, rather than bibitem, so that you can change the bibliography
+style super quickly. Also, most journals provide the bibtex entry for the article, so you just save that in your bib file - yes, no need to type much if you use that feature-}
+
+Transient modeling of nuclear reactors has been a chronically foreboding task due to its computationally expensive nature.  However, recent developments in nuclear testing
+\tcr{this is not the introduction of a short summary, you can spend a bit more time in the topic of transient testing,
+what is it for, what facilities, how what it simulated before, why a regain of interest in simulation for this today, ...
+build the storyline...}, including the revitalization of the Transient Reactor Testing (TREAT) Facility at Idaho National Laboratory (INL), have brought significant attention to the development of transient reactor modeling.
 3-D transient computational methods currently in use at national laboratories, including INL, have proven to be overbearingly computationally expensive for complex, real-world scenarios.
-Therefore, implementation of more efficient and faster methods for 3-D transient simulations is highly desired.
+Therefore, implementation of more efficient and faster methods for 3-D transient simulations is highly desired \tcr{this was already a problem when Karl Ott introduced IQS in 1977(?) so to rationale sounds a bit hollow the way you present it and yet we are doing it and Downar was awarded last year a grant to do this too.
+I think the real reason is that most modern reactor physics tools have not being used much for fast transients, so the only thing we are doing is bringing back the method into modern tools.
+However, in doing so, }.
 The Department of Energy (DOE), through its NEAMS (Nuclear Energy Advanced Modeling and Simulation) program, has especially sought the development of transient multiphysics capability in INL's MOOSE (Multiphysics Object-Oriented Simulation Environment).
@@
 
+\tcr{
+I think an outline for the introduction should be 
+\begin{enumerate}
+\item reactor transient testing.
+Background, etc...
+this subsection should conclude on the overarching need of transient testing: computational methods for multiphysics reactor transients.
+It should also introduce the next 2 subsections, which are both about lit review, one on time-dependent  neutronics, one on multiphysics solution techniques for reactor physics.
+\item literature review on mostly neutronics time dependent: challenges in computational methods for time-dependent neutron balance equations.
+the size of the phase-space, the need for implicitness in the temporal treatment due to the stiffness, ....
+this should open up to a discussion on alternate methods, such as IQS, ...
+it is a lit review, so no need for equations, but rather a demonstration that you know the background and have a solid understand of that foundation to lay the ground from your OWN work.
+As part of this lit, some discussion of Mund's, Dulla's, Ott's, Monier,'s the Canadian's papers is welcome.
+\item now, reactor transients also means coupled to other physics.
+this is were you discuss some aspect related to multiphysics.
+you can start with the general discussion that this is typically a nonlinear problem, that it can be solved via picard or newton.
+this is maybe a segway into a discussion/introduction about MOOSE.
+Then you go back on IQS and add the multiphysics aspects.
+Ott did some work with multiphysics, I gave you the code manual of his own software where he discusses when he decided to do shape updates.
+\end{enumerate}
+Generally speaking, the rest is real good but you need to spend some time on these 5-8 pages of introduction that set the tone and the background.
+Everybody needs to understand that you know a lot; that your thoughts are well organized and clear; the rest will come easy (except some results ...).\\
+%
+side note: we need to put the small results about sdirk being poor for stiffness problem. maybe as an appendix. I want to keep that knowledge somewhere. 
+}
 MOOSE is a multiphysics framework being developed at INL that presents the architecture for physics-based applications \cite{moose}.

### 2017-02-23 doc/Thesis/data/introduction.tex
Thesis: Addressing initial comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2016.
%
%  Author: Sean Zachary Roberson 
%	 Version 3.08.16
%  Last updated 8/19/2016
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           SECTION I
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagestyle{plain} % No headers, just page numbers
\pagenumbering{arabic} % Arabic numerals
\setcounter{page}{1}

\chapter{\uppercase{Introduction}}

\section{Background on Transient Reactor Testing}

The primary purpose of transient reactor testing is the safety and performance analysis of fuel and other reactor components. Since the March 2011 events at the Fukushima Daiichi Nuclear Power Plant in Japan, there has been a significant governmental demand for the development of accident tolerant fuels in light water reactors (LWR) \cite{ConRepTREAT}.  Currently, two facilities are capable of testing fuels under these extreme conditions: Annular Core Research Reactor (ACRR) at Sandia National Laboratory (SNL) and the Transient Reactor Testing Facitlity (TREAT) at Idaho National Laboratory (INL) \cite{SandiaTREAT}. TREAT had been put in stand-by status in 1994, but in 2014 a Final Environmental Assessment by the Department of Energy (DOE) approved the restart of the facility \cite{FONSI}.

TREAT is an graphite-moderated, air-cooled, thermal reactor built and operated starting in 1959.  The reactor was designed to subject fuels and experimental apparatus to extreme power pulses, for fast reactor type components \cite{TREATSummary}. The primary purpose for the restart of the facility is to use these pulses to simulate accident scenarios and test the fidelity of accident tolerant LWR fuels. TREAT is expected to resume experimentation and testing by 2020.  In addition to the substantial upgrades of TREAT's electronic and mechanical systems during the restart, advance computer models of the reactor are in development.  These models have a mutualistic relationship with TREAT: validating fuel models with experimentation and using reactor models to streamline experimental procedures. Development of these models have renewed a significant interest in implementing transient reactor simulation methods in modern computational tools.

Transient reactor simulation methods have been developed for several decades. The revitalization of TREAT and other transient reactor testing efforts has brought special attention to the methods' implementation in modern computational tools.  The DOE, through its NEAMS (Nuclear Energy Advanced Modeling and Simulation) program, has especially sought the development of transient multiphysics capability in INL's MOOSE (Multiphysics Object-Oriented Simulation Environment).  The following two sections overview time-dependent neutronics and multiphysics solution methods.

\section{Time-Dependent Neutron Transport}

Transient modeling of nuclear reactors has been a chronically foreboding task due to its computationally expensive nature. The difficulty in transient reactor modeling is due to:
\begin{enumerate}
\item
the high-dimensionality of the phase-space for the governing equations that describe the flux of neutrons (6-D+time for multigroup neutron transport and 4-D+time for multigroup neutron diffusion), and 
\item
the fact that the time discretization has to be implicit, which  leads to stiff system of equations.  
\end{enumerate}
The transport equation has seven independent variables: space ($\vec{r}$), energy ($E$), direction ($\vec{\Omega}$), and time ($t$) \cite{duderstadt1976nuclear}.  Nuclear reactor simulations often utilize the neutron diffusion approximation, which carefully eliminates the dependence on direction.  For the purpose of this research, neutron diffusion will be discussed exclusively.

For computational purposes, the neutron diffusion equation is discretized in space, energy, and time.  There are several viable discretization schemes for each of these variables; FEM for space and multigroup in energy are used for the development of this research \cite{zeinkiewicz2005finite, duderstadt1976nuclear}. Since the purpose of this research involves temporal dependence, discretization in time is kept general. Most reactor computation frameworks discretize the time variable of flux directly using a multitude of schemes (Euler, Crank-Nicholson, Runge-Kutta, etc.).  However, the solutions of flux can be incredibly stiff due to the speed of neutrons, especially in a pulsing reactor \cite{TWIGL_benchmark}. Therefore, computationally expensive implicit schemes are necessary with many points of discretization (time steps). The following subsections describe two alternate techniques: point reactor kinetics and the improved quasi-static method (IQS). These methods are meant to reduce computational expense of evaluating transient flux while maintaining accuracy.

\subsection{Point Reactor Kinetics}

A common solution process for neutron kinetics is the evaluation of the point reactor kinetics equation (PRKE) \cite{Planchard_1991}.  The derivation of this equation involves factorization flux into a space-dependent shape and time-dependent amplitude.  This factorization immediately creates a large assumption that the spacial variance of the flux is time-independent.  The amplitude maintains the temporal stiffness as flux; however, the PRKE takes very little computational effort to evaluate because the multitudinous variables from the spacial discretization only needs to be evaluated on the initial time step.  The rest of the time steps only evaluated one variable: amplitude.

\subsection{Improved Quasi-Static Method}

IQS is a spatial kinetics method that involves factorizing the flux solution into space- and time-dependent components \cite{Ott_1966,Dulla2008,Devooght_1984,Monier_diss,Sissaoui_1995}.  These components are the flux's amplitude and its shape. Amplitude is only time-dependent, while the shape is both space- and time-dependent.  However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the variable for the diffusion evaluation is less stiff without making the assumption that spacial variance is time-independent, as the PRKE does.  As a result, the shape may not require an update at the same frequency as the amplitude function, but only on larger macro-time steps.

Since shape and amplitude share a independent variable, they are temporally coupled. Ott in \cite{Ott_1966} was the first publication that investigated the coupling of shape and amplitude in a quasi-static nature, but kept steady-state shape equation to evaluate. Ott later in \cite{Ott_1969} incorporated the time derivative of shape in the equation, yielding better results.  The resulting system of equations is heavily nonlinear \cite{Dulla2008}. Nonlinear problems require an iterative process to evaluate, either fixed-point (Picard) or Newton iteration. Sissaoui et al. \cite{Sissaoui_1995}, Koclas et al. \cite{Koclas_1996}, Devooght et al. \cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their quasi-static simulations, the main difference is their criteria for convergence.  Devooght et al. in \cite{Devooght_1984} also utilizes a Newton iteration technique.

\section{Time-Dependent Reactor Dynamics}

Transient reactor analysis also involves coupling neutronics with other physics like temperature feedback, thermal hydraulics, fuel structure and composition, etc., this coupling is known as multiphysics simulation. Multiphysics simulation is a daunting task because it requires communication of coupled variables that are being evaluated with a myriad of techniques. The coupling is also usually nonlinear requiring Picard or Newton iteration techniques. Programs, such as MOOSE, provide the capability of evaluating, communicating, and iterating different physics.

MOOSE is a multiphysics framework being developed at INL that presents the architecture for physics-based applications \cite{moose}.  Rattlesnake is a large application in MOOSE that involves deterministic radiation transport physics.  Currently, Rattlesnake is able to solve steady-state, transient, and k-eigenvalue neutron transport and diffusion problems using finite element methods (FEM) \cite{wang2013}. BISON is the application for evaluating fuel temperature and structural properties \cite{bison}.  RELAP-7 is the main safety analysis application, but focuses on variables involved with hydrodynamic analysis \cite{relap7}.  MAMMOTH is the over-arching application that couples all these physics by recomputing cross-sections for Rattlesnake and source terms for BISON and RELAP-7 \cite{mammoth}.

IQS is a manipulation of neutronics variables, so it is solely developed within the Rattlesnake application. However, proper treatment of multiphysics with IQS is vital for computational optimization.  The quasi-static nature of IQS can be extended to multiphysics simulation by determining the proper time scales for each physics.  Considering TREAT, shape is the slowest varying variable and amplitude is fastest, while fuel temperature is in between.  Varying time step sizes for each variable is optimal for minimizing computation expense, while maintaining accuracy.

\section{Objective}

The goal of this research is to continue the investigation and development of the improved quasi-static method.  This goal requires the development of IQS in a easily modifiable, prototype-like code, as well as a robust FEM framework. \\
\indent The prototyping code is used to investigate time discretization and nonlinear iteration schemes. The prototype will also guide the development of IQS in the large FEM framework of Rattlesnake/MOOSE, where more complex examples and benchmarks can be tested.  Development in Rattlesnake also spawns the opportunity to test IQS with multiphysics, namely temperature feedback, and time adaptation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tcr{
I think an outline for the introduction should be 
\begin{enumerate}
\item reactor transient testing. Background, etc... this subsection should conclude on the overarching need of transient testing: computational methods for multiphysics reactor transients. It should also introduce the next 2 subsections, which are both about lit review, one on time-dependent  neutronics, one on multiphysics solution techniques for reactor physics.
\item literature review on mostly neutronics time dependent: challenges in computational methods for time-dependent neutron balance equations. the size of the phase-space, the need for implicitness in the temporal treatment due to the stiffness, .... this should open up to a discussion on alternate methods, such as IQS, ... it is a lit review, so no need for equations, but rather a demonstration that you know the background and have a solid understand of that foundation to lay the ground from your OWN work. As part of this lit, some discussion of Mund's, Dulla's, Ott's, Monier,'s the Canadian's papers is welcome. 
\item now, reactor transients also means coupled to other physics. this is were you discuss some aspect related to multiphysics. you can start with the general discussion that this is typically a nonlinear problem, that it can be solved via picard or newton. this is maybe a segway into a discussion/introduction about MOOSE. Then you go back on IQS and add the multiphysics aspects. Ott did some work with multiphysics, I gave you the code manual of his own software where he discusses when he decided to do shape updates.
\end{enumerate}
Generally speaking, the rest is real good but you need to spend some time on these 5-8 pages of introduction that set the tone and the background. Everybody needs to understand that you know a lot; that your thoughts are well organized and clear; the rest will come easy (except some results ...).\\
side note: we need to put the small results about sdirk being poor for stiffness problem. maybe as an appendix. I want to keep that knowledge somewhere. 
}

### 2017-02-23 doc/Thesis/data/theory.tex
Thesis: Addressing initial comments

--- 2017-02-04 doc/Thesis/data/section2.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To derive the IQS equations, the time dependent neutron diffusion equation with delayed neutron precursors are presented by Equations~\ref{eq:flux}~and~\ref{eq:precursor}.
+
+\begin{subequations}
+\begin{multline}
+\frac{1}{v^g}\frac{\partial \phi^g}{\partial t}  = 
+\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \phi^{g'} 
++ \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \phi^{g'}  \\ + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i 
+-  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g   
+\ , \quad 1 \le g \le G 
+\label{eq:flux}
+\end{multline}
+\be
+\frac{dC_i}{dt} = \frac{\beta_i}{\keff}\sum_{g=1}^G\nu^{g} \Sigma_f^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
+\label{eq:precursor}
+\ee
+\end{subequations}
+
+where,
+
+\begin{longtable}{lll}
+$\phi^g$   			&	$=$	&	Scalar flux in energy group $g$ \\
+$C_i$					  &	$=$	&	Concentration of delayed neutron precursor $i$ \\
+$\Sigma_f^{g}$	&	$=$	&	Fission cross section in energy group $g$ \\
+$\Sigma_r^{g}$	&	$=$	&	Removal cross section in energy group $g$ \\
+$\Sigma_s^{g' \to g}$	&	$=$	&	Scattering cross section from energy group $g'$ to $g$ \\
+$v^g$					  &	$=$	&	Neutron velocity in energy group $g$ \\
+$\chi_p^g$			&	$=$	&	Fission spectrum of prompt neutrons \\
+$\chi_{d,i}^g$	&	$=$	&	Fission spectrum of delayed neutrons from precursor $i$ \\
+$\nu^g$					&	$=$	&	Total number of neutrons per fission \\
+$D^g$					  &	$=$	&	Diffusion coefficient in energy group $g$\\
+$\lambda_i$			&	$=$	&	Decay constant of precursor $i$ \\
+$\beta_i$				&	$=$	&	Delayed neutron fraction  from precursor $i$ \\
+$\beta$			 	  &	$=$	&	Total delayed neutron fraction ($\beta = \sum_{i=1}^I \beta_{i}$) \\
+  & & 
+\end{longtable}
+
+Most reactor computation frameworks, including Rattlesnake, discretize the time variable directly with these equations using a multitude of schemes (Implicit Euler, Crank-Nicholson, implicit Runge-Kutta, etc.).
+In this paper, the method of discretizing Equations~\ref{eq:flux}~and~\ref{eq:precursor} is generally referred to as ``implicit discretization''.
+This research intends to improve upon this method by instead implementing the improved quasi-static method (IQS) for neutron kinetics and
+implement it within a multiphysics setting.
 

### 2017-02-23 doc/Thesis/data/dynamics.tex
Thesis: Adding LRA and Tran15 results

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2016.
%
%  Author: Sean Zachary Roberson
%	 Version 3.16.09
%  Last updated 9/12/2016
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           SECTION IV
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{DYNAMICS EXAMPLES \label{sect:dyn}}

Reactor dynamics is the study of the time-dependent behavior of reactors as an entire system. This study includes the physical nature of neutrons and their feedback with power generation.  This feedback includes coupling with other physical properties such as temperature, fluids, material dynamics, etc.  This section describes several dynamics examples, including the LRA benchmark and two TREAT experiments.  These examples are of increased complexity from the kinetics examples of \sct{sect:kin}.  This section also analyzes IQS's performance with these, which is vital for verification of IQS in real-world problems.

\section{LRA Benchmark}

The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.  It is a super prompt-critical transient.  To have better understanding on the cross sections given later, we present the equations here:
\begin{subequations}
\begin{align}
-\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 - \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
-\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
S_f &= \sum_{g=1}^2 \Sigma_{f,g} \phi_g, \\
\frac{\partial C_i}{\partial t} &= \nu\beta_i f - \lambda_i C_i, \quad i=1,2, \\
\frac{\partial T}{\partial t} &= \alpha f, \label{eq:lra-temp} \\
\Sigma_{a,1} &= \Sigma_{a,1}(\vec{r}, t=0) \left[1+\gamma\left(\sqrt{T} - \sqrt{T_0}\right)\right], \\
P &= \kappa S_f,
\end{align}
\end{subequations}
where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0=T(\vec{r}, t=0)$.
The two-group diffusion equation are solved with zero flux boundary conditions on external surfaces, reflecting conditions at symmetry boundaries and steady state initial conditions which are obtained by solving
\begin{align}
-\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
-\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1.
\end{align}
The eigenvalue $k$ is used to modify the fission cross section for the transient simulations with $\frac{1}{k}\Sigma_{f,g}, g=1,2$.  The initial flux distribution shall be normalized such that the averaged power density
\begin{align}
\bar{P} \equiv \frac{\int_{V_{core}} P(\vec{r}, t=0) d\vec{r}}{\int_{V_{core}} d\vec{r}},
\end{align}
where $V_{core}$ is the core region with fuels, is equal to $10^{-6} W\cdot cm^{-3}$.
The initial precursor concentrations are in equilibrium with the initial critical flux distribution.\\

The geometry is illustrated in~\fig{fig:lra-geometry}.\\
\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\linewidth]{\FiguresDir/lra.png}
\caption{LRA benchmark geometry with region assignment.}
\label{fig:lra-geometry}
\end{figure}

Initial two-group constants are presented in~\tbl{tab:lra-xs}.
$\nu$ is equal to 2.43.
Axial bulking $B^2 = 10^{-4}$ is applied for both energy groups.
Delayed neutron data are presented in~\tbl{tab:lra-dnp}.
All fuel materials have the same delayed neutron data.
Some scalar data are listed in~\tbl{tab:lra-scalar}.\\
\begin{table}[!htbp]
    \centering
    \caption{LRA benchmark initial two-group constants.\label{tab:lra-xs}}
    \resizebox{\textwidth}{!}{
      \begin{tabular}{|c|c|c|c|l|l|l|l|l|}
      \hline
             &                 & Group   & $D_g$      & $\Sigma_{a,g}$ & $\nu\Sigma_{f,g}$ & $\Sigma_{s,1\rightarrow 2}$ & $\chi_g$ & $v_g$              \\
      Region & Material        &       g & (cm)       &  ($cm^{-1}$)   &  ($cm^{-1}$)      &  ($cm^{-1}$)                &          & ($cm\cdot s^{-1}$) \\
      \hline
      1      & Fuel 1 with rod & 1       & 1.255      & 0.008252       & 0.004602          &                             & 1        & $3.0\times10^7$    \\
             &                 & 2       & 0.211      & 0.1003         & 0.1091            & 0.02533                     & 0        & $3.0\times10^5$    \\
      \hline
      2      & Fuel 1 without rod & 1    & 1.268      & 0.007181       & 0.004609          &                             & 1        & $3.0\times10^7$    \\
             &                    & 2    & 0.1902     & 0.07047        & 0.08675           & 0.02767                     & 0        & $3.0\times10^5$    \\
      \hline
      3      & Fuel 2 with rod & 1       & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
             &                 & 2       & 0.2091     & 0.08344        & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
      \hline
      4      & Fuel 2 without rod & 1    & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
             &                    & 2    & 0.2091     & 0.073324       & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
      \hline
      5      & Reflector        & 1      & 1.257      & 0.0006034      & -                 &                             & -        & $3.0\times10^7$    \\
             &                  & 2      & 0.1592     & 0.01911        & -                 & 0.04754                     & -        & $3.0\times10^5$    \\
      \hline
      \end{tabular}}
\end{table}

\begin{table}[!htbp]
    \centering
    \caption{LRA benchmark delayed neutron data.\label{tab:lra-dnp}}
      \begin{tabular}{|c|l|l|l|l|}
      \hline
      Group i & $\beta_i$ & $\lambda_i$ ($s^{-1}$) & $\chi_{d,i,1}$ & $\chi_{d,i,2}$ \\
      \hline
      1       & 0.0054    & 0.0654  & 1 & 0 \\
      2       & 0.001087  & 1.35    & 1 & 0 \\
      \hline
      \end{tabular}
\end{table}

\begin{table}[!htbp]
    \centering
    \caption{LRA benchmark scalar values.\label{tab:lra-scalar}}
      \begin{tabular}{|l|c|l|}
      \hline
      Meaning & Notation & value \\
      \hline
      Axial buckling for both energy groups & $B_g^2$  & $10^{-4}$ ($cm^{-2}$)\\
      Mean number of neutrons per fission   & $\nu$    & 2.43 \\
      Conversion factor                     & $\alpha$ & $3.83\times 10^{-11}$ ($K\cdot cm^{3}$) \\
      Feedback constant                     & $\gamma$ & $3.034\times 10^{-3}$ ($K^{1/2}$) \\
      Energy released per fission           & $\kappa$ & $3.204\times 10^{-11} $ ($J/fission$) \\
      Initial and reference temperature     & $T_0$    & 300 (K) \\
      Active core volume                    & $V_{core}$ & 17550 ($cm^2$)\\
      \hline
      \end{tabular}
\end{table}
The transient is initiated by changing the thermal absorption cross section as the following:
\begin{align}
\Sigma_{a,2}(t) = \Sigma_{a,2}(t=0) \left\{\begin{array}{lr} 1-0.0606184t, & t\leq 2 \\
                                                             0.8787631, & t>2
                                           \end{array}\right.
\end{align}
where $t$ is time in seconds.

\subsection{LRA Multiphysics Time Scale Results}

\fig{fig:lra_profile} shows the baseline power and temperature transient profile for the LRA benchmark. \fig{fig:lra_profile2D} shows the spacial power distribution at the peak power.  The baseline results are compared to the results achieved by Sutton and Aviles in \cite{Sutton_1996} and presented in \tbl{tab:base}.  The relative difference in the magnitude of the peak power ($t\approx1.44 s$) from the baseline was used for error comparison.  \fig{fig:lra_bad} is an error convergence plot comparing the three techniques where temperature is evaluated only on the macro step (1 temperature update).  \fig{fig:lra_mpconv} is an error convergence plot comparing the three techniques where temperature is evaluated 5 times within a macro step (5 temperature updates).  Finally, \fig{fig:mp} shows the effect of various temperature updates. The dashed lines correspond to implicit discretization at different flux step sizes, while the IQS macro step size is kept constant.

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{\FiguresDir/lra_profile.png}
\caption{LRA baseline temperature and power profile}
\label{fig:lra_profile}
\end{figure}

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{\FiguresDir/lra_profile2D.png}
\caption{LRA baseline spacial power profile at $t=1.44 s$}
\label{fig:lra_profile2D}
\end{figure}

\begin{table}[!htbp]
\begin{center}
%\resizebox{\linewidth}{!}{
\begin{tabular}{|l|cc|}
\hline
Calculation  &  Baseline & Sutton (Spandex 1936) \\
\hline
No. of Spatial Nodes	& 3872 		& 1936 \\
Eigenvalue 				& 0.99637	& 0.99637 \\
No. of Time Steps 		& 6000 		& 23,890 \\
Time to Peak Power (s) 	& 1.441 	& 1.441 \\
Peak Power (W/cm$^3$) 	& 5456 		& 5461 \\
\hline
\end{tabular}
%}
\end{center}
\caption{LRA baseline verification}
\label{tab:base}
\end{table}

\begin{figure}[!htbp]
\centering
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/lra_bad.png}
\caption{Only one temperature update per macro step}
\label{fig:lra_bad}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/lra_mp_convergence.png}
\caption{Five temperature updates per macro step}
\label{fig:lra_mpconv}
\end{subfigure}
\caption{LRA error convergence plots}
\end{figure}

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{\FiguresDir/lra_mp.png}
\caption{Error plot with various temperature updates per macro step}
\label{fig:mp}
\end{figure}

The convergence plots show that updating temperature and the PRKE parameters within a macro step has a significant effect on the performance of IQS.  With only one update, IQS was only slightly better than implicit discretization, implicit discretization required about 150\% more time steps than IQS for the same error.  While 5 temperature updates showed a much more significant IQS performance, implicit discretization required about 400\% more time steps than IQS for the same error.  \fig{fig:mp} shows that error has a convergent behavior for the number of temperature updates.  This convergence makes sense because temperature can only be so accurate before the error in shape is dominating. Table \ref{tab:ndiff_lra} shows the run time results for the implicit discretization calculations. The number of GMRES linear iterations is included because it is proportional measure of the computational effort. Tables \ref{tab:iqs_lra} and \ref{tab:iqspc_lra} present the IQS run-times with various numbers of temperature updates.  These run-times are based on total alive time of the execution where the diffusion evaluation is distributed over 24 processors. These run-times show a marginal performance for IQS and impressive performance for IQS P-C.  Some of the execution times were able to decrease from implicit discretization with the same number of macro steps because IQS is better equipped to resolve the nonlinearity between temperature and amplitude. Furthermore, there does seem to be an ideal number of temperature updates to optimize execution time: IQS only needs one and IQS P-C seems to be ideal at 4 updates. This discrepancy in the number of updates shows that a adaptive type implementation of the updates would be ideal, and could enforce a constant error over the transient. It is also important to compare the error of implicit discretization with IQS at one update and IQS P-C at 4 updates.  IQS shows an error comparable to implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by -34.1\%.  IQS P-C shows an error less than implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by <-34.9\%.

\begin{table}[!htbp]
\begin{center}
%\resizebox{\linewidth}{!}{
\begin{tabular}{|l|l|ccc|}
\hline
Run  &  $\Delta t$ & Error & Runtime (hr) & Linear Iter.\\
\hline
1	& 4.0e-3	& 1.407e-2 	& 4.11	& 7.13e4	\\
2	& 2.0e-3	& 3.174e-3 	& 6.01	& 9.49e4 	\\
3 	& 1.0e-3 	& 7.690e-4 	& 10.38	& 1.45e5	\\
4 	& 5.0e-4 	& 1.892e-4 	& 21.91	& 2.08e5	\\
5 	& 2.5e-4	& 4.590e-5 	& 25.23	& 3.16e5	\\
\hline
\end{tabular}
%}
\end{center}
\caption{Implicit discretization run time results}
\label{tab:ndiff_lra}
\end{table}

\begin{table}[!htbp]
\begin{center}
%\resizebox{\linewidth}{!}{
\begin{tabular}{|l|l|ccc|}
\hline
	&  Temperature 	&  		& Runtime 	& \% Increase	\\
Run	&  Updates 	& Error & (hr)		& in Runtime$^*$\\
\hline
1	& 1		& 2.612e-3 	& 3.96 	& -3.18\%	\\
2	& 2		& 9.893e-4 	& 6.02	&  47.1\%	\\
3 	& 4 	& 5.796e-4 	& 7.87	&  92.3\%	\\
4 	& 8 	& 4.772e-4 	& 12.61	& 207.9\% 	\\
5 	& 16	& 4.516e-4 	& 22.14	& 440.7\%	\\
\hline
\end{tabular}
%}
\\
$^*$ difference in runtime from $\Delta t = 0.004$ implicit discretization 
\caption{IQS run time results with $\Delta t = 0.004$}
\label{tab:iqs_lra}
\end{center}
\end{table}

\begin{table}[!htbp]
\begin{center}
%\resizebox{\linewidth}{!}{
\begin{tabular}{|l|l|ccc|}
\hline
	&  Temperature 	&  		& Runtime 	& \% Increase	\\
Run	&  Updates 	& Error & (hr)		& in Runtime$^*$\\
\hline
1	& 1		& 3.488e-3 	& 2.91 	& -28.9\%	\\
2	& 2		& 1.349e-3 	& 3.73	& -9.00\%	\\
3 	& 4 	& 9.161e-4 	& 3.97	& -3.04\%	\\
4 	& 8 	& 8.052e-4 	& 5.39	&  31.7\%	\\
5 	& 16	& 7.905e-4 	& 8.19	&  100\%	\\
\hline
\end{tabular}
%}
\\
$^*$ difference in runtime from $\Delta t = 0.004$ implicit discretization 
\caption{IQS PC run time results with $\Delta t = 0.004$}
\label{tab:iqspc_lra}
\end{center}
\end{table}

The performance of the quasi-statics can also be explained by the computation of the dynamical time scale described by \sct{sect:tau}.  \fig{fig:LRAtc} shows the time scale profile over the transient, computed using \eqt{eq:tau2}.  This plot shows that in a implicit discretization simulation, the flux dominates the time dependent behavior, while temperature lags in its variance for the majority of the transient.  In an IQS simulation, the time scale behavior of amplitude almost exactly matches the flux, while shape is more varying than temperature throughout most of the transient.  The large $\tau$ for temperature during the beginning of the transient shows that adaptation of the number of updates is important; computational expense on temperature evaluations is being wasted during this time.

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{\FiguresDir/time_constant_lra.png}
\caption{Dynamical time scale for LRA benchmark}
\label{fig:LRAtc}
\end{figure}

\subsection{LRA with Time Adaptation}

\fig{fig:lra_dt2} shows the power profile of the LRA with time adaptation of implicit discretization and IQS P-C, and \tbl{tab:lra_dt2} compiles the results. These time adaptation results show the significant decrease in macro time steps required for IQS P-C. These profiles were obtaining with only one temperature update per macro step; so based on previous results, the IQS P-C performance would improve even more with more updates.

\begin{figure}[!htbp]
\centering
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/LRA_DT2.png}
\caption{Full power profile}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/LRA_DT2_peak.png}
\caption{Peak power profile at peak}
\end{subfigure}
\caption{LRA power profile with time adaptation of implicit discretization and IQS P-C}
\label{fig:lra_dt2}
\end{figure}

\begin{table}
\begin{center}
\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
 & \multicolumn{3}{|c|}{Implicit Dis.} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Event & Power (W/cm$^3$) & Error & Steps & Power (W/cm$^3$) & Error & Steps \\
\hline
Max Power & 5567.3 & 0.019454 & 423 & 5568.3 & 0.019274 & 47 \\
End (3 s) & 109.66 & 2.3650e-4 & 603 & 109.65 & 3.0622e-4 & 97 \\
\hline
\end{tabular}}
\end{center}
\caption{LRA step doubling adaptation results with implicit discretization and IQS P-C}
\label{tab:lra_dt2}
\end{table}

\section{TREAT Transient-15 Problem}

Transient 15 is a test case based on the TREAT core. The preliminary purpose of this model was to match an early test of TREAT; due to lack of experimental data and procedures, model validation was impossible \cite{mammoth}. Therefore, ultimate purpose of this simulation in Rattlesnake is to test the model's fidelity with the thermal feedback of TREAT, but it is not meant to exactly match any previous experiments.  Nevertheless, the goal of the following simulations is to test IQS and its time scale based treatment of temperature with a more complex model. The model involves a 159-element "small core" configuration of TREAT, shown in \fig{fig:Tran15_config} \cite{Tran15}. \fig{fig:Tran15_mesh} shows the meshing techniques of the model \cite{mammoth}. To ease computation, the fully homogenized version (without air gaps) was used  for the following simulations. Transient 15 involves an 11-energy group diffusion approximation and is discretized into $355,712$ hexahedral continuous finite elements totaling $4,109,523$ degrees of freedom.  The three-second transient involves a linear ramp decrease in the absorption cross section throughout the control rod region. \fig{fig:Tran15} shows a visualization of the flux profile within the core, hidden is the massive amount of graphite surrounding the core.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\linewidth]{\FiguresDir/Tran15_config.png}
\caption{Transient 15 159-element small core configuration}
\label{fig:Tran15_config}
\end{figure}

\begin{figure}[!htbp]
\centering
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/Tran15_mesh_fine.png}
\caption{Explicit meshing of air channels}
\label{fig:Tran15_mesh_fine}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/Tran15_mesh_homo.png}
\caption{Fully homogenized fuel elements}
\label{fig:Tran15_mesh_homo}
\end{subfigure}
\caption{Top quarter view of Transient 15 mesh}
\label{fig:Tran15_mesh}
\end{figure}

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.6\linewidth]{\FiguresDir/Tran15_core2.png}
\caption{Transient 15 core power profile at peak power}
\label{fig:Tran15}
\end{figure}

\subsection{Transient-15 Temperature Feedback}

The Transient-15 model uses a adiabatic temperature feedback mechanism, similar to the one explored by the LRA. \eqt{eq:temp2} describes the heat up of the fuel.  It is very similar, except the specific heat is now dependent on temperature is described by \eqt{eq:cp}.  The temperature evaluation is identical to the one described in LRA section, except a Newton iteration process is employed to resolve the nonlinearity from the specific heat term.  The feedback to the cross-sections are applied using linear interpolation of tabular data provided by INL.

\be
\rho c_p(T) \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
\label{eq:temp2}
\ee

\be
c_p = -5.8219\text{e-}10T^3 - 4.3694\text{e-}7T^2 + 2.8369\text{e-}3T -1.009\text{e-}2 \ (J/g/K)
\label{eq:cp}
\ee

\subsection{Transient-15 Multiphysics Time Scale Results}

In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time step, implicit discretization, IQS with one and 5 temperature updates per macro step, and IQS P-C with one and 5 updates.  \fig{fig:Tran15_profile} shows the baseline power and temperature profile for the Transient-15 example.  \tbl{tab:tran15} shows the error and runtime results.

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{\FiguresDir/Tran15_profile.png}
\caption{Transient 15 total power and average temperature profile during transient}
\label{fig:Tran15_profile}
\end{figure}

\begin{table}[htb]
\begin{center}
\resizebox{\linewidth}{!}{
\begin{tabular}{|l|ccccccc|}
\hline
 & No.	& Max &	Time at Max &	Max Average &	\% Increase &	Max Power & Linear	\\
Method & of Steps &	Power (W) &	Power (s) &	Temperature (K) & Runtime$^*$ & Error & Iterations \\
\hline
Baseline 			& 3000 	& 3.5039e+08 & 1.901 & 371 &	---		&	---		 & ---		\\
Implicit Dis. 		& 300 	& 3.5011e+08 & 1.90  & 371 &	---		&	7.875e-4 & 41020	\\
IQS 				& 300	& 3.5036e+08 & 1.90  & 371 & -11.9\%	&	8.385e-5 & 23949	\\
IQS (5 updates) 	& 300 	& 3.5040e+08 & 1.90  & 371 &  49.7\%	&	3.687e-5 & 24035	\\
IQS P-C 			& 300 	& 3.5065e+08 & 1.90  & 371 &  -2.1\%	&	7.527e-4 & 39020	\\
IQS P-C (5 updates) & 300 	& 3.5043e+08 & 1.90  & 371 &  26.5\%	&	1.227e-4 & 37866	\\
\hline
\end{tabular}
}
\end{center}
\vspace{-3mm}
$^*$ difference in runtime from implicit discretization 
\caption{Transient-15 Error and Runtime Results}
\label{tab:tran15}
\end{table}

The results from \tbl{tab:tran15} show similar performance of IQS with the temperature updates as the LRA. Again, the number of linear GMRES iterations is shown as a measure of computational expense. However, these iterations do not consider the temperature updates, so the iterations of the simulations with multiple updates should be taken with a grain of salt. IQS with 1 temperature update shows a performance that reduces the error to approximately a tenth of the implicit discretization error, and reduces the execution time by about 12\%.  This shows that IQS was able to resolve the nonlinearity between flux and temperature with significantly fewer diffusion evaluations.  Having IQS with 5 updates significantly increased the execution time for the same time step, but the error was reduced.  Comparing this error to a similar implicit discretization error at a smaller time step could show that the runtime was reduced. IQS P-C performed not nearly as well as it did with the LRA benchmark, but still proved to be effective.  Having 5 updates for IQS P-C increased the runtime marginally, but decreased the error significantly.  The transient profile of the variables' dynamical time scales is shown in \fig{fig:tran15tc}.  This plot exhibits a similar response to that of the LRA. The response of temperature shows that the updates are a computational frugal treatment of the feedback and adaptation of the number of updates is vital for optimization. \\

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\linewidth]{\FiguresDir/time_constant_tran15.png}
\caption{Dynamical time scale for the Transient 15 example}
\label{fig:tran15tc}
\end{figure}

\subsection{Transient-15 with Time Adaptation}

\fig{fig:Tran15_dt2} shows the power profile of the LRA with time adaptation of implicit discretization and IQS P-C.  These plots show that IQS P-C requires marginally fewer macro time steps than implicit discretization, but is qualitatively much closer to the baseline profile. Like the LRA step doubling results, IQS P-C only performs one temperature update per macro step. Adding more updates would most likely improve the error, but increase the computation time significantly.

\begin{figure}[!htbp]
\centering
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/Tran15_DT2.png}
\caption{Full power profile}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/Tran15_DT2_peak.png}
\caption{Peak power profile}
\end{subfigure}
\caption{Transient-15 step doubling adaptation results with implicit discretization and IQS P-C}
\label{fig:Tran15_dt2}
\end{figure}

\section{TREAT M8-CAL Problem}

### 2017-02-23 doc/Thesis/data/theory.tex
Thesis: Adding LRA and Tran15 results

@@
 
-To derive the IQS equations, the time dependent neutron diffusion equation with delayed neutron precursors are presented by Equations~\ref{eq:flux}~and~\ref{eq:precursor}.
+To derive the IQS equations, the time dependent neutron diffusion equation with delayed neutron precursors are presented by \eqts{eq:flux}{eq:precursor}.
 
@@
 Most reactor computation frameworks, including Rattlesnake, discretize the time variable directly with these equations using a multitude of schemes (Implicit Euler, Crank-Nicholson, implicit Runge-Kutta, etc.).
-In this paper, the method of discretizing Equations~\ref{eq:flux}~and~\ref{eq:precursor} is generally referred to as ``implicit discretization''.
+In this paper, the method of discretizing \eqts{eq:flux}{eq:precursor} is generally referred to as ``implicit discretization''.
 This research intends to improve upon this method by instead implementing the improved quasi-static method (IQS) for neutron kinetics and
@@
 IQS involves factorizing the flux from \eqt{eq:flux} into time-dependent amplitude ($p$) and space- and time- dependent shape ($\varphi$).
-The resulting equation is the shape-diffusion equation with precursors represented by Equations~\ref{eq:shape}~and~\ref{eq:prec}.
+The resulting equation is the shape-diffusion equation with precursors represented by \eqts{eq:shape}{eq:prec}.
 
@@
 \ee
-Finally, the PRKE parameters are defined by Equations~\ref{eq:trmb}~and~\ref{eq:tb}, where 
+Finally, the PRKE parameters are defined by \eqts{eq:trmb}{eq:tb}, where 
 $\left(\Psi^{*g},f^g\right) =$\\ $ \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d^2\Omega$.
@@
 \ee
-This resulting system can be split into two systems, represented by Equations~\eqref{eq:PJFNK_1}~and~\eqref{eq:PJFNK_2}.
+This resulting system can be split into two systems, represented by \eqts{eq:PJFNK_1}{eq:PJFNK_2}.
 \be 
@@
 The Predictor-Corrector (P-C) version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the solution of the coupled system of equations is different.
-In the IQS P-C version, the flux equations (not the shape equations) are solved (represented by Equations~\ref{eq:flux}~and~\ref{eq:precursor}) in order to obtain a predicted flux solution.
+In the IQS P-C version, the flux equations (not the shape equations) are solved (represented by \eqts{eq:flux}{eq:precursor}) in order to obtain a predicted flux solution.
 This predicted flux is then converted to a shape by rescaling it as follows:
@@
 
-The PRKE parameters are then computed with this shape using Equations~\ref{eq:rmb}~-~\ref{eq:l} and interpolated over the macro step, then the PRKE ODE system is solved on the micro time scale.
+The PRKE parameters are then computed with this shape using \eqtss{eq:rmb}{eq:l} and interpolated over the macro step, then the PRKE ODE system is solved on the micro time scale.
 With the newly computed amplitude, the shape is rescaled into a flux and the final corrected flux is given by:
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Dynamical Time Scale Analysis}
+\label{sect:tau}
+
+The temporal variance of flux, shape, temperature, and amplitude can be quantified by defining a dynamical time scale ($\tau$) for each physics.
+A small value $\tau$ means the variable is quickly varying in time and consequently needs relatively small time steps for accuracy, vice-versa for large $\tau$.
+The general definition of $\tau$ is defined by \eqt{eq:tau1}, where $\theta$ is the physic component of interest.
+\be
+\tau = \frac{1}{\left|\frac{1}{\theta}\frac{d\theta}{dt}\right|}
+\label{eq:tau1}
+\ee
+Since each variable in discretized in time, a finite difference approximation will be made for the $\frac{d\theta}{dt}$ term and the average between the two corresponding time steps will be made for the $\frac{1}{\theta}$ term.
+Additionally, $\tau$ is spatially dependent for flux and temperature, but only the time dependent behavior of this quantity is of interest.
+Therefore, the $L^2$ norm of each term will be used to compute the approximate time scale ($\tilde{\tau}$), formally defined by \eqt{eq:tau2}.
+$\theta$ represents a summation over groups for flux and shape.
+\be
+\tilde{\tau}_{n+1} = \frac{\norm{\theta_{n+1} + \theta_{n}}}{2}\frac{\Delta t}{\norm{\theta_{n+1} - \theta_{n}}}
+\label{eq:tau2}
+\ee
+%
+According to the a priori hypothesis from previous sections, $\tau$ is large for shape, somewhat smaller for temperature, and much smaller for amplitude and flux.
 
@@
 A new time step size is estimated using  \eqt{eq:edt2} as well.
-This process can be visualized by Figs.~\ref{fig:dt2_1}~and~\ref{fig:dt2_2}.
+This process can be visualized by \figs{fig:dt2_1}{fig:dt2_2}.
 Where a step involves a full convergence of shape, amplitude, and any multiphysics on the respective time step.
@@
 
-In order to evaluate the PRKE coefficients, defined by Equations~\ref{eq:rmb}~-~\ref{eq:l}, four postprocessors were created.
+In order to evaluate the PRKE coefficients, defined by \eqtss{eq:rmb}{eq:l}, four postprocessors were created.
 The parameter calculations were separated by $\frac{\bar{\beta}_i}{\Lambda}$ numerator, $\bar{\lambda}_i$ numerator/denominator, $\frac{\rho-\bar{\beta}}{\Lambda}/\frac{\bar{\beta}}{\Lambda}$ denominator, and $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator.
@@
 However, Rattlesnake has other action systems capable of transient simulation, where IQS can be implemented and be effective.
-One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equations~\eqref{eq:flux}~and~\eqref{eq:shape}.
+One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in \eqts{eq:flux}{eq:shape}.
 However, in the derivation for IQS, this term is unaffected between shape and flux evaluation.

### 2017-02-26 doc/Thesis/data/kinetics.tex
Thesis: Adding IQS iteration analysis

--- 2017-02-10 doc/Thesis/data/section3.tex
@@
 
+IQS, as previously stated, is a system of nonlinear equations between shape and amplitude.
+These equations needed to be iterated to numerically converge to an accurate solution.
+\sct{sect:iter} describes various iteration techniques for fixed-point and Newton schemes.
+\fig{fig:iter} shows the number of fixed-point iterations required for a $10^{-11}$ tolerance over the transient.
+The criteria listed in the legend are described as:
+\begin{enumerate}
+\item $L^{\infty} \rightarrow$ \eqt{eq:shape_Linf}
+\item $L^{2} \rightarrow$ \eqt{eq:shape_L2}
+\item Reactivity $\rightarrow$ \eqt{eq:rho_conv}
+\item Amplitude $\rightarrow$ \eqt{eq:p_conv}
+\item K criteria $\rightarrow$ \eqt{eq:K_conv}
+\item All properties $\rightarrow$ \eqtss{eq:rho_conv}{eq:K_conv}
+\end{enumerate}
+This plot shows that 1-4 have approximately the same convergence behavior, but the K criteria converges to a certain error.
+\fig{fig:iter_err} shows the resulting error the K criteria converges to for different points of rescaling the shape.
+The rescaling is described by \eqt{eq:shape_scale}.
+Rescaling shape more frequently helps the error.
+However, rescaling every iteration is somewhat artificial because it does not consider changes spatially.
+Regardless, an error of $~10^{-5}$ is quite large and it is expected that the magnitude is due to the explicit treatment of precursors (\eqt{eq:dnp_theta}).
+Switching to an analytical elimination (\eqt{eq:dnp_an}) does not converge as well, but the error is much smaller, seen in \fig{fig:iter_err_an}.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.75\linewidth]{\FiguresDir/iter_renorm.png}
+\caption{\# of iterations for various convergence criteria, tolerance$=10^{-11}$, max iterations$=20$}
+\label{fig:iter}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.75\linewidth]{\FiguresDir/iter_error.png}
+\caption{Final iteration error for K convergence criteria}
+\label{fig:iter_err}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.75\linewidth]{\FiguresDir/iter_error_an.png}
+\caption{Final iteration error for K convergence criteria with analytical precursor elimination}
+\label{fig:iter_err_an}
+\end{figure}
+
 \subsection{Time Step Convergence}
@@
 In order to evaluate the performance and error convergence of IQS, the slab was simulated with varying time discretization methods and time step sizes.
-\fig{1D_conv} shows these convergence plots of five different discretization methods for implicit discretization, IQS, and IQS P-C.
+\fig{fig:1D_conv} shows these convergence plots of five different discretization methods for implicit discretization, IQS, and IQS P-C.
 These plots were generated from the results using the MATLAB prototype program.

### 2017-02-26 matlab/PRKE_testing/assemble_source.m

@@ -1,0 +1,36 @@
+function [S] = assemble_source(sfun,curr_time)
+
+porder= 3;
+gn    = [1,2,3,4];
+n=4;
+% allocate memory
+S=zeros(n,1);
+% initialize local matrices
+v=zeros(porder+1,1);
+% shape set
+b=[0.660005665072803,0.520937687711704,-0.230187903250739,0.0492445504662318;...
+   0.00337373643277257,1.00488585482565,-0.00992135357232475,0.00166176231390641;...
+   0.00166176231390640,-0.00992135357232467,1.00488585482565,0.00337373643277254;...
+   0.0492445504662318,-0.230187903250739,0.520937687711703,0.660005665072804];
+% quadrature stuff
+xq=[-0.861136311594053;-0.339981043584856;0.339981043584856;0.861136311594053];
+wq=[0.347854845137454;0.652145154862546;0.652145154862546;0.347854845137454];
+
+% element extremities
+x0=0;
+x1=1;
+% jacobian of the transformation to the ref. element
+Jac=(x1-x0)/2;
+% get x values in the interval
+x=(x1+x0)/2+xq*(x1-x0)/2;
+% b and dbdx(:,:) are of size (nbr of xq values) x (porder+1),
+% 2/dx is the 1d jacobian
+% old: d=fct_ptr(curr_time{imat},x)/Jac;
+s = sfun(x,curr_time)*Jac;%/npar.keff;
+
+% assemble
+for i=1:porder+1
+    v(i) = dot(s.*wq , b(:,i));
+end
+S(gn(1,:)) = S(gn(1,:)) + v;
+

### 2017-02-26 matlab/PRKE_testing/rho_test.m

@@ -1,0 +1,298 @@
+function rho_test
+
+clear;clc;close all;
+
+syms x t
+syms S_p(x,t) phi(x,t) C(x,t) S_c(x,t)
+syms varphi(x,t) amp(t) dvarphi(x,t)
+
+syms f(t) a(x,t) A(t)
+f(t) = (t+1);
+a(x,t) = x*(1-x)*(x+t);
+A(t) = int(a(x,t)*a(x,0),x,0,1);
+phi(x,t) = f(t)*a(x,t)/A(t);
+varphi(x,t) = a(x,t)/A(t);
+amp(t) = f(t);
+
+iv = 1e-3; Sa = 1.0; nfSf = 1.1; cdiff = 1.0; b = 0.0;
+S_p(x,t) = iv*diff(phi,t) + (Sa-nfSf*(1-b))*phi - cdiff*diff(diff(phi,x),x);
+
+phi_exact = matlabFunction(phi);
+shape_exact = matlabFunction(varphi);
+dshape_exact = matlabFunction(diff(varphi,t));
+p_exact = matlabFunction(amp);
+source_phi = matlabFunction(S_p);
+clear x t S_p phi C S_c 
+clear varphi(x,t) amp(t) dvarphi(x,t)
+clear f(t) a(x,t) A(t)
+
+x = [0 1/3 2/3 1]';
+tend = 1.0;
+
+pinit = 1.0;
+phi_adj = phi_exact(x,0);
+K0 = 0.105;
+
+PL = zeros(4);
+PL(2,2) = -10.7614285714286;
+PL(3,2) =  7.42017857142857;
+PL(2,3) =  7.42017857142857;
+PL(3,3) = -10.7614285714286;
+
+rho_exact = @(t) (phi_adj' * PL * shape_exact(x,t)) / K0;
+q = @(t) (phi_adj' * assemble_source(source_phi,t))/K0;
+
+n_micro = 10;
+n_macro = [10 20 40 80 500 1000];% 5000 1e4];
+ntimes=n_macro*n_micro; % 5e4 1e5];% 5e5];% 5e5];
+
+% rho_fun = @(t,told,tnew) rho_exact(t); type = 'Exact \rho(t)';
+% rho_fun = @(t,told,tnew) lagrange_fun(rho_exact,told,tnew,1,t); type = 'Linear interpolation of \rho(t) over macro step';
[…]

### 2017-02-26 matlab/compute_conv_error.m

@@ -1,0 +1,47 @@
+function [err] = compute_conv_error(phi,phiprev,fac)
+% make the problem-data a global variable
+global npar
+
+% shortcuts
+porder= npar.porder;
+gn    = npar.gn;
+nel   = npar.nel;
+% n: linear system size
+n=npar.n;
+% shape set
+b=npar.b;
+% quadrature stuff
+xq=npar.xq;
+wq=npar.wq;
+
+num=0;
+deno=0;
+
+if fac < Inf
+    % loop over elements
+    for iel=1:npar.nel
+        % element extremities
+        x0=npar.x(iel);
+        x1=npar.x(iel+1);
+        % jacobian of the transformation to the ref. element
+        Jac=(x1-x0)/2;
+        % get x values in the interval
+        x=(x1+x0)/2+xq*(x1-x0)/2;
+        % b and dbdx(:,:) are of size (nbr of xq values) x (porder+1),
+        fem_phi = phi(gn(iel,:));
+        fem_phi_xq = b*fem_phi;
+        fem_phiprev = phiprev(gn(iel,:));
+        fem_phiprev_xq = b*fem_phiprev;
+        num = num + Jac*dot(wq,(fem_phi_xq-fem_phiprev_xq).^fac);
+        deno = deno + Jac*dot(wq,(fem_phi_xq).^fac);    
+    end
+    num=num^(1/fac);
+    deno=deno^(1/fac);
+else
+    num = max(abs(phi - phiprev));
+    deno = max(abs(phi));
+end
+
+err = num/deno;
+return
+end

### 2017-02-26 matlab/problem_init.m

@@ -9,10 +9,10 @@ dat.PbID = problem_ID;
 
 % rod mov times
 dat.rod_mov.t_beg_1=0.1; dat.rod_mov.t_end_1=0.6;
-dat.rod_mov.t_beg_2=1.0; dat.rod_mov.t_end_2=2.7;
+dat.rod_mov.t_beg_2=1.0; dat.rod_mov.t_end_2=1.7;
 
 % kinetic parameters
-dat.beta_tot=600e-5*0;
+dat.beta_tot=600e-5;
 dat.lambda=0.1;
 dat.invvel=1e-3;
 
@@ -309,9 +309,12 @@ npar.solve_prke_compute_rho_each_time = false;
 npar.prke_solve = 'matlab' ;
 npar.int_order = 1;
 % npar.prke_solve = 'no' ;
-npar.max_iter_iqs = 6;
+npar.max_iter_iqs = 20;
 npar.tol_iqs      = 1e-11;
 npar.iqs_prke_interpolation_method=1;
+npar.conv_criteria = 'none';
+npar.scale_IQS = false;
+npar.scale_IQS_each = false;
 
 if ~strcmpi(npar.prke_solve,'matlab')
     npar.n_micro=1000;
@@ -322,10 +325,10 @@ end
 dat.max_y_val_movie = 2.;
 
 % time integration
-t_order=3;
+t_order=1;
 
-% npar.method = 'SDIRK';
-npar.method = 'BDF';
+npar.method = 'SDIRK';
+% npar.method = 'BDF';
 
 if strcmp(npar.method,'SDIRK')
     time_integration = t_order;

### 2017-02-26 matlab/solve_IQS_diffusion_an_prec.m

@@ -10,7 +10,7 @@ rk=npar.rk;
 if strcmp(npar.method,'BDF')
     bdf = npar.bdf;
 end
-order = length(tn);
+order = length(tn)-1;
 
 max_iter_iqs = npar.max_iter_iqs;
 tol_iqs      = npar.tol_iqs;
@@ -23,9 +23,8 @@ X_beg=X;
 shape_beg=u_shape(1:npar.n,:);
 shape_end=shape_beg(:,end);
 
-tn = tn - dt;
-time_beg = tn(end);
-time_end = time_beg+dt;
+time_end = tn(end);
+time_beg = time_end-dt;
 
 % storage for temp SDIRK quantities
 K=zeros(length(shape_beg(:,1)),rk.s);
@@ -33,10 +32,11 @@ K=zeros(length(shape_beg(:,1)),rk.s);
 NFId_old = assemble_mass(dat.nusigf_d,time_beg) / npar.keff;
 
 for iter = 1: max_iter_iqs
-    
+    shape_prev = shape_end;
+    X_prev = X;
     % solve for amplitude function
     if strcmpi(npar.prke_solve,'matlab')
-        [X,dXdt,w,t,y]    = solve_prke_ode(X_beg(:,end),dt_macro,time_end,shape_beg(:,end),shape_end);
+        [X,dXdt,w,t,y] =  solve_prke_ode(X_beg(:,end),dt_macro,tn,shape_beg,shape_end);
     else
         [X,dpdt,t,y] = solve_prke_iqs(X_beg(:,end),dt_macro,time_end,shape_beg(:,end),shape_end,npar.n_micro,npar.freq_react);
     end
@@ -91,12 +91,12 @@ for iter = 1: max_iter_iqs
         ti = time_beg + rk.c(i)*dt;
         t2=ti;
         t1=time_beg;
-        t_all = [tn t2];
+        t_all = tn;
         dti=rk.c(i)*dt;
         
         % OLD shortcut
         % OLD p=X(1);
-        pi = ppval(pp,ti);
+        pi = X(1);
 %         dpdti = ppval(pp_der,ti);
         dpdti = dXdt(1);
         
@@ -208,17 +208,40 @@ for iter = 1: max_iter_iqs
     u_shape = [ shape_end ; C_new];
     
     % check for tolerance
-    err = abs( (npar.phi_adj)'*IV*shape_end/npar.K0  - 1);
+    err = 0;
[…]

### 2017-02-26 matlab/solve_IQS_diffusion_elim_prec.m

@@ -37,7 +37,8 @@ NFId_old = assemble_mass(dat.nusigf_d,time_beg) / npar.keff;
 
 
 for iter = 1: max_iter_iqs
-    
+    shape_prev = shape_end;
+    X_prev = X;
     % solve for amplitude function over the entire macro-step
     if strcmpi(npar.prke_solve,'matlab')
         [X,dXdt,w,t,y] =  solve_prke_ode(X_beg(:,end),dt_macro,tn,shape_beg(:,1:end),shape_end);
@@ -165,17 +166,40 @@ for iter = 1: max_iter_iqs
     u_shape = [ shape_end ; Ci];
     
     % check for tolerance
-    err = abs( (npar.phi_adj)'*IV*shape_end/npar.K0  - 1);
+    err = 0;
+    if strcmp(npar.conv_criteria,'Linf')
+        err = compute_conv_error(shape_end,shape_prev,Inf);
+    elseif strcmp(npar.conv_criteria,'L2')
+        err = compute_conv_error(shape_end,shape_prev,2);
+    elseif max(strcmp(npar.conv_criteria,{'amp', 'rho', 'param_all'}))
+        if max(strcmp(npar.conv_criteria,{'rho', 'param_all'}))
+            [rho_end,~,~]=compute_prke_parameters(time_end,shape_end);
+            [rho_prev,~,~]=compute_prke_parameters(time_end,shape_prev);
+            err = max(err,abs(rho_end - rho_prev));
+        end
+        if max(strcmp(npar.conv_criteria,{'amp', 'param_all'}))
+            err = max(err,abs(X(1) - X_prev(1)));
+        end
+        if strcmp(npar.conv_criteria,'param_all')
+            err = max(err,abs( (npar.phi_adj)'*IV*shape_end/npar.K0  - 1));
+        end
+    else
+        err = abs( (npar.phi_adj)'*IV*shape_end/npar.K0  - 1);
+    end
+    
     if io.console_print
         fprintf('  IQS iter %d, err %g \n',iter,err);
     end
     if err<tol_iqs
         break
-    else
-%                 u_shape = u_shape / ((npar.phi_adj)'*IV*shape_end/npar.K0);
-%                 shape_end=u_shape(1:npar.n);
+    elseif npar.scale_IQS_each
+        u_shape = u_shape / ((npar.phi_adj)'*IV*shape_end/npar.K0);
[…]

### 2017-02-26 matlab/time_marching_IQS.m

@@ -18,7 +18,7 @@ end
 order=1;
 if strcmp(npar.method,'BDF')
     order = npar.bdf_order;
-elseif strcmp(func2str(FUNHANDLE),'solve_TD_diffusion_an_prec') && strcmp(npar.an_interp_type,'lagrange')
+elseif strcmp(func2str(FUNHANDLE),'solve_IQS_diffusion_an_prec') && strcmp(npar.an_interp_type,'lagrange')
     order=npar.interpolation_order;
 end
 u_shape=zeros(length(u0),order+1);
@@ -32,9 +32,11 @@ Ptot = dat.Ptot*ones(ntimes+1,1);
 [~,beff_MGT]=compute_prke_parameters(0.,u0(1:npar.n));
 X(:,end)=[1;beff_MGT/dat.lambda];
 
-    dat.ode.f_beg=zeros(npar.n,1);
-    dat.ode.f_end=zeros(npar.n,1);
+dat.ode.f_beg=zeros(npar.n,1);
+dat.ode.f_end=zeros(npar.n,1);
 
+dat.iter_IQS = [0];
+dat.iter_err = [0];
 
 time_prke_iqs=[];
 power_prke_iqs=[];
@@ -110,6 +112,33 @@ if io.plot_power_figure
     plot_power_level( func2str(FUNHANDLE),...
         linspace(0,dt*ntimes,ntimes+1), amplitude_norm,...
         time_prke_iqs, power_prke_iqs );
+end
+
+if io.plot_iqs_conv
+    figure(23)
+%     plot(tn,(amplitude_norm-min(amplitude_norm))/(max(amplitude_norm)-min(amplitude_norm))*npar.max_iter_iqs,'k--')
+%     semilogy(tn,10.^((amplitude_norm-min(amplitude_norm))/(max(amplitude_norm)-min(amplitude_norm))*max(dat.iter_err)),'k--')
+    hold on
+%     plot(tn,dat.iter_IQS)
+    semilogy(tn,dat.iter_err)
+    xlabel('time (s)')
+    ylabel('# of Iterations')
+    grid on
+%     figid = figure(23);
+%     Pline = line(tn,amplitude_norm,'Color','b');
+%     ax1 = gca;
+%     ax1.XColor = 'k';
+%     ax1.YColor = 'b';
+%     ax1.YLabel.String = 'log_{10}(Total Power (W))';
+%     ax1_xlim = ax1.XLim;
[…]

### 2017-02-26 matlab/IQSfun.m
Attempting IQS with JFNK (not working right yet)

@@ -1,0 +1,10 @@
+function y = IQSfun(y,shape_end,shape_beg,X_beg,tn,TR,NFId_new,NFId_old,IV,Re,eps,C_old)
+
+Rep = residual_IQS(shape_end + eps*y,shape_beg,X_beg,tn,TR,NFId_new,NFId_old,IV,C_old);
+
+y = (Rep - Re) / eps;
+
+return
+end
+
+

### 2017-02-26 matlab/kinetics_driver_conv.m

@@ -2,7 +2,7 @@ function kinetics_driver_conv
 
 clc;
 % clear all;
-close all;
+% close all;
 clear global;
 % turn off warning due to interp1
 warning('OFF','MATLAB:interp1:ppGriddedInterpolant')
@@ -10,12 +10,13 @@ warning('OFF','MATLAB:interp1:ppGriddedInterpolant')
 global npar io
 
 % verbose/output parameters
-io.console_print         = false;
+io.console_print         = true;
 io.plot_transient_figure = false;
-io.plot_power_figure     = false;
+io.plot_power_figure     = true;
 io.make_movie            = false;
 io.save_flux             = false;
 io.print_progress        = false;
+io.plot_iqs_conv         = false;
 io.figID = 99;
 % one of the two choices for applying BC
 npar.set_bc_last=true;
@@ -44,7 +45,7 @@ u=[phi0;C0];
 u0=u;
 
 % time steping data
-t_end = 1.1;
+t_end = 2.0;
 % t_end = 1.3;
 % t_end = 0.1;
 
@@ -58,8 +59,8 @@ else
     amplitude_norm_ref = reference_solution( t_end, u0);
 end
 
-nn=5;
-ntimes = 2.^(0:nn-1)*10;
+nn=1;
+ntimes = 2.^(0:nn-1)*100;
 dt = t_end./ntimes;
 
 % Interpolation type of shape for IQS prke parameters
@@ -69,7 +70,7 @@ dt = t_end./ntimes;
 %   2      = linear interpolation of shape
 %   3      = cubic interpolation of shape
 %   4      = quadratic interpolation of shape
-npar.iqs_prke_interpolation_method=5;
+npar.iqs_prke_interpolation_method=1;
 npar.solve_prke_compute_rho_each_time = false;
 
 % Type of precursor interpolation for analytical elimination
@@ -98,12 +99,13 @@ npar.prec_solve_type = 'linear';
 i=0;
 % not to be used for conv. studies % i=i+1; list_runs{i}= 'brute_force_matlab';
 % i=i+1; list_runs{i}= 'brute_force';
-i=i+1; list_runs{i}= 'brute_force_elim_prec';
[…]

### 2017-02-26 matlab/plot_power_level.m

@@ -22,6 +22,9 @@ switch method_name
     case 'solve_IQS_diffusion_elim_prec'
         curr_leg = 'IQS-elim';
     
+    case 'solve_IQS_diffusion_JFNK'
+        curr_leg = 'IQS-JFNK';
+    
     case 'solve_IQS_PC_diffusion_an_prec'
         curr_leg = 'IQS-PC-an';
     

### 2017-02-26 matlab/residual_IQS.m

@@ -1,0 +1,56 @@
+function Re = residual_IQS(shape,shape_beg,X_beg,tn,TR,NFId_new,NFId_old,IV,C_old)
+
+global dat npar
+
+lambda = dat.lambda;
+
+order = length(tn)-1;
+dt = tn(end) - tn(end-1);
+time_end = tn(end);
+time_beg = time_end-dt;
+
+[X,dXdt,~,t,y] =  solve_prke_ode(X_beg,dt,tn,shape_beg,shape);
+
+% interpolating polynomial
+pp = interp1(t,y,'spline','pp');
+
+p = X(1);
+dpdt = dXdt(1);
+
+% build rhs
+a = zeros(order+1,2);
+for j=1:(order+1)
+    Aj = @(t) LagrangeInterp_section(t,tn,j).*exp(-lambda*(time_end-t)).*ppval(pp,t);
+    a(j,1) = integral(@(t)Aj(t).*(time_end-t)/dt,time_beg,time_end);
+    a(j,2) = integral(@(t)Aj(t).*(t-time_beg)/dt,time_beg,time_end);
+end
+
+zi =  lambda *( C_old*exp(-lambda*dt) )/p;
+TR = TR + IV*dpdt/p;
+TR = TR + lambda * ( a(end,1)*NFId_old + a(end,2)*NFId_new )/p ;
+
+for n=1:order
+    zi = zi + lambda *( a(n,1)*NFId_old*shape_beg(:,n) + a(n,2)*NFId_new*shape_beg(:,n) )/p;
+end
+A = IV-dt*TR;
+rhs = IV*shape_beg(:,end) + dt*zi;
+
+% apply BC
+if npar.set_bc_last
+    [A,rhs]=apply_BC(A,rhs,npar.add_ones_on_diagonal);
+else
+    rhs=apply_BC_vec_only(rhs);
+end
+
+Re = A*shape - rhs;
+    
+if dat.solve_prec 
+    C_new =  C_old*exp(-lambda*dt) + ( a(end,1)*NFId_old + a(end,2)*NFId_new )*shape ;
+    for n=1:order
+        C_new = C_new + ( a(n,1)*NFId_old + a(n,2)*NFId_new )*shape_beg(:,n);
+    end
+    dat.C_new = C_new;
+    dat.t = t;
+    dat.y = y;
+    dat.X = X;
+end

### 2017-02-26 matlab/solve_IQS_diffusion_JFNK.m

@@ -1,0 +1,60 @@
+function [u_shape, X,t,y] = solve_IQS_diffusion_JFNK(u_shape,X,dt_macro,tn)
+
+global io dat npar
+
+% shortcuts
+dt = dt_macro;
+C_old = u_shape(npar.n+1:end,end);
+
+% save values at beginning of macro time step: they are needed in the IQS iteration
+X_beg=X;
+% for clarity, I also single out the shape function at the beginning/end of the macro time step
+shape_beg=u_shape(1:npar.n,:);
+shape_end=shape_beg(:,end);
+
+time_end = tn(end);
+time_beg = time_end-dt;
+
+% assemble matrices
+D    = assemble_stiffness(dat.cdiff   ,time_end);
+A    = assemble_mass(     dat.siga    ,time_end);
+NFIp = assemble_mass(     dat.nusigf_p,time_end) / npar.keff;
+IV   = assemble_mass(     dat.inv_vel ,time_end);
+NFId_new = assemble_mass(dat.nusigf_d ,time_end) / npar.keff;
+NFId_old = assemble_mass(dat.nusigf_d ,time_beg) / npar.keff;
+TR=NFIp -(D+A);
+
+eps = 1e-6;
+
+for iter = 1: npar.max_iter_iqs   
+
+    % Compute residual and precursors
+    dat.solve_prec = true;
+    Re = residual_IQS(shape_end,shape_beg,X_beg,tn,TR,NFId_new,NFId_old,IV,C_old);
+    residual = norm(Re,2);
+    if io.console_print
+        fprintf('  IQS iter %d, Nonlinear|R| %g \n',iter,residual);
+    end
+    if residual < npar.tol_iqs
+        break;
+    end
+    
+    % Solve linear system
+    dat.solve_prec = false;
+    dshape = gmres(@(y) IQSfun(y,shape_end,shape_beg,X_beg,tn,TR,NFId_new,NFId_old,IV,Re,eps,C_old), -Re,[],1e-4,npar.n);
+    shape_end = shape_end + dshape;
+    
+end
+
+u_shape = [ shape_end ; dat.C_new];
+t = dat.t;
+y = dat.y;
+X = dat.X;
+
+% dat.iter_IQS = [dat.iter_IQS iter];
+% dat.iter_err = [dat.iter_err err];
+
+% renormalize anyway
+if npar.scale_IQS
+    u_shape = u_shape / ( ((npar.phi_adj)'*npar.IV*shape_end) / npar.K0 );
+end

### 2017-02-28 doc/Thesis/data/introduction.tex
Thesis: Making some changes to Introduction

@@
 The primary purpose of transient reactor testing is the safety and performance analysis of fuel and other reactor components.
-Since the March 2011 events at the Fukushima Daiichi Nuclear Power Plant in Japan, there has been a significant governmental demand for the development of accident tolerant fuels in light water reactors (LWR) \cite{ConRepTREAT}.
-Currently, two facilities are capable of testing fuels under these extreme conditions: Annular Core Research Reactor (ACRR) at Sandia National Laboratory (SNL) and the Transient Reactor Testing Facitlity (TREAT) at Idaho National Laboratory (INL) \cite{SandiaTREAT}.
+Since the March 2011 events at the Fukushima Daiichi Nuclear Power Plant in Japan, there has been significant governmental demand for the development of accident tolerant fuels in light water reactors (LWR) \cite{ConRepTREAT}.
+Currently, two facilities are capable of testing fuels under these extreme conditions: Annular Core Research Reactor (ACRR) at Sandia National Laboratory (SNL) and the Transient Reactor Testing Facility (TREAT) at Idaho National Laboratory (INL) \cite{SandiaTREAT}.
 TREAT had been put in stand-by status in 1994, but in 2014 a Final Environmental Assessment by the Department of Energy (DOE) approved the restart of the facility \cite{FONSI}.
 
-TREAT is an graphite-moderated, air-cooled, thermal reactor built and operated starting in 1959.
+TREAT is an graphite-moderated, air-cooled, thermal reactor which began operation in 1959.
 The reactor was designed to subject fuels and experimental apparatus to extreme power pulses, for fast reactor type components \cite{TREATSummary}.
-The primary purpose for the restart of the facility is to use these pulses to simulate accident scenarios and test the fidelity of accident tolerant LWR fuels.
+The primary purpose for the restart of the facility is to use these pulses to simulate accident scenarios and test the integrity of accident tolerant LWR fuels.
 TREAT is expected to resume experimentation and testing by 2020.
@@
 These models have a mutualistic relationship with TREAT: validating fuel models with experimentation and using reactor models to streamline experimental procedures.
-Development of these models have renewed a significant interest in implementing transient reactor simulation methods in modern computational tools.
+Development of these models has renewed a significant interest in implementing transient reactor simulation methods in modern computational tools.
 
 Transient reactor simulation methods have been developed for several decades.
-The revitalization of TREAT and other transient reactor testing efforts has brought special attention to the methods' implementation in modern computational tools.
+The revitalization of TREAT and other transient reactor testing efforts have brought special attention to the methods' implementation in modern computational tools.
 The DOE, through its NEAMS (Nuclear Energy Advanced Modeling and Simulation) program, has especially sought the development of transient multiphysics capability in INL's MOOSE (Multiphysics Object-Oriented Simulation Environment).
-The following two sections overview time-dependent neutronics and multiphysics solution methods.
+This research intends to investigate and develop the improved quasi-static method (IQS) with these modern tools.
+IQS is a transient neutronics method intended to improve computational efficiency.
+Improving this efficiency is vital for reactor simulations that would otherwise be overbearing to evaluate.
+In order to apply IQS to full transient reactor models, implementation of multiphysics is required.
+The following two sections provide an overview of time-dependent neutronics and multiphysics solution methods.
 
@@
 
-Transient modeling of nuclear reactors has been a chronically foreboding task due to its computationally expensive nature. The difficulty in transient reactor modeling is due to:
+Transient modeling of nuclear reactors has been a chronically formidable task due to its computationally expensive nature. The difficulty in transient reactor modeling is due to:
 \begin{enumerate}
@@
 \item
-the fact that the time discretization has to be implicit, which  leads to stiff system of equations.  
+the fact that the time discretization has to be implicit, which  leads to a stiff system of equations.  
 \end{enumerate}
@@
 Nuclear reactor simulations often utilize the neutron diffusion approximation, which carefully eliminates the dependence on direction.
-For the purpose of this research, neutron diffusion will be discussed exclusively.
+For the purpose of this research, neutron diffusion is discussed exclusively.
 
 For computational purposes, the neutron diffusion equation is discretized in space, energy, and time.
-There are several viable discretization schemes for each of these variables; FEM for space and multigroup in energy are used for the development of this research \cite{zeinkiewicz2005finite, duderstadt1976nuclear}.
+There are several viable discretization schemes for each of these variables.
+In this research effort, FEM is used for space discretization and multigroup for energy discretization \cite{zeinkiewicz2005finite, duderstadt1976nuclear}.
 Since the purpose of this research involves temporal dependence, discretization in time is kept general.
 Most reactor computation frameworks discretize the time variable of flux directly using a multitude of schemes (Euler, Crank-Nicholson, Runge-Kutta, etc.).
-However, the solutions of flux can be incredibly stiff due to the speed of neutrons, especially in a pulsing reactor \cite{TWIGL_benchmark}.
+However, the solutions of flux can be particullary stiff due to the speed of neutrons, especially in a pulsing reactor \cite{TWIGL_benchmark}.
 Therefore, computationally expensive implicit schemes are necessary with many points of discretization (time steps).
-The following subsections describe two alternate techniques: point reactor kinetics and the improved quasi-static method (IQS).
+The following subsections describe two alternate techniques for transient simulation: point reactor kinetics and the improved quasi-static method (IQS).
 These methods are meant to reduce computational expense of evaluating transient flux while maintaining accuracy.
@@
 A common solution process for neutron kinetics is the evaluation of the point reactor kinetics equation (PRKE) \cite{Planchard_1991}.
-The derivation of this equation involves factorization flux into a space-dependent shape and time-dependent amplitude.
+The derivation of this equation involves factorizing flux into a space-dependent shape and time-dependent amplitude.
 This factorization immediately creates a large assumption that the spacial variance of the flux is time-independent.
-The amplitude maintains the temporal stiffness as flux; however, the PRKE takes very little computational effort to evaluate because the multitudinous variables from the spacial discretization only needs to be evaluated on the initial time step.
-The rest of the time steps only evaluated one variable: amplitude.
+The amplitude has the same temporal stiffness as flux; however, the PRKE takes very little computational effort to evaluate because the many variables from the spatial discretization only need to be evaluated on the initial time step.
+The rest of the time steps just evaluate one variable: amplitude.
 
@@
 Amplitude is only time-dependent, while the shape is both space- and time-dependent.
-However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the variable for the diffusion evaluation is less stiff without making the assumption that spacial variance is time-independent, as the PRKE does.
+However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the variable for the diffusion evaluation is less stiff without making the assumption that spatial variance is time-independent, as the PRKE does.
 As a result, the shape may not require an update at the same frequency as the amplitude function, but only on larger macro-time steps.
 
-Since shape and amplitude share a independent variable, they are temporally coupled.
-Ott in \cite{Ott_1966} was the first publication that investigated the coupling of shape and amplitude in a quasi-static nature, but kept steady-state shape equation to evaluate.
+Since shape and amplitude share an independent variable, they are temporally coupled.
+Ott in \cite{Ott_1966} was the first publication that investigated the coupling of shape and amplitude in a quasi-static nature, but did not include the time derivative in the shape equation.
 Ott later in \cite{Ott_1969} incorporated the time derivative of shape in the equation, yielding better results.
@@
 \cite{Koclas_1996}, Devooght et al.
-\cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their quasi-static simulations, the main difference is their criteria for convergence.
+\cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their quasi-static simulations, the main difference among them is their criteria for convergence.
 Devooght et al.
@@
 
-Transient reactor analysis also involves coupling neutronics with other physics like temperature feedback, thermal hydraulics, fuel structure and composition, etc., this coupling is known as multiphysics simulation.
-Multiphysics simulation is a daunting task because it requires communication of coupled variables that are being evaluated with a myriad of techniques.
+Transient reactor analysis also involves coupling neutronics with other physics like temperature feedback, thermal hydraulics, fuel structure and composition, etc.
+This coupling is known as multiphysics simulation.
+Multiphysics simulation is often a daunting task because it requires communication of coupled variables that are being evaluated with a myriad of techniques.
 The coupling is also usually nonlinear requiring Picard or Newton iteration techniques.
@@
 MOOSE is a multiphysics framework being developed at INL that presents the architecture for physics-based applications \cite{moose}.
+In order to robustly implement IQS in the MOOSE framework for multiphysics simulation, each coupled physics application requires consideration.
 Rattlesnake is a large application in MOOSE that involves deterministic radiation transport physics.
@@
 Considering TREAT, shape is the slowest varying variable and amplitude is fastest, while fuel temperature is in between.
-Varying time step sizes for each variable is optimal for minimizing computation expense, while maintaining accuracy.
+Varying time step sizes for each variable can considerably improve computational efficiency, which is the nature and purpose of IQS application.
 
@@
 
-The goal of this research is to continue the investigation and development of the improved quasi-static method.
-This goal requires the development of IQS in a easily modifiable, prototype-like code, as well as a robust FEM framework.
-\\
-\indent The prototyping code is used to investigate time discretization and nonlinear iteration schemes.
-The prototype will also guide the development of IQS in the large FEM framework of Rattlesnake/MOOSE, where more complex examples and benchmarks can be tested.
-Development in Rattlesnake also spawns the opportunity to test IQS with multiphysics, namely temperature feedback, and time adaptation.
+The goal of this research is to continue the investigation and development of the improved quasi-static method for minimizing computation expense for transient reactor simulations, while maintaining accuracy.
+This goal is pursued with three objectives:
+
+\begin{enumerate}
+\item Establish IQS performance for various iteration techniques
+\item Validate time step convergence for IQS with high order time discretization schemes
+\item Apply IQS to multiphysics simulations
+\end{enumerate}
+
+Iteration convergence analysis involves investigating iteration techniques for nonlinear problems and convergence criteria.
+For this objective, IQS is implemented into an easily modifiable, prototype-like code in MATLAB.
+High order time discretizing schemes were also implemented in the prototyping code to investigate time step convergence.
+The prototype also guides the development of IQS in the large FEM framework of Rattlesnake/MOOSE, where more complex examples and benchmarks can be tested.
+Time step convergence is also analyzed for these examples.
+Development in Rattlesnake also spawns the opportunity to test IQS with multiphysics, namely temperature feedback and time adaptation.
+Applying IQS to multiphysics simulations allows insight to the performance of IQS for full transient reactor simulations, particularly for TREAT modeling.
 
@@
 this is were you discuss some aspect related to multiphysics.
-you can start with the general discussion that this is typically a nonlinear problem, that it can be solved via picard or newton.
+you can start with the general discussion that this is typically a nonlinear problem, that it can be solved via Picard or newton.
 this is maybe a segway into a discussion/introduction about MOOSE.

### 2017-02-28 doc/Thesis/data/abstract.tex
Thesis: Adding abstract

@@
 
-\indent Going to finish this when I finish the rest of the paper.
+\indent Transient reactor simulation is a chronically formidable challenge due to the computational rigor of evaluating the neutron transport equation, as well as its coupling with other physical phenomena.
+The goal of this thesis research is to mitigate the computational expense of these simulations by investigating and developing the improved quasi-static method (IQS).
+IQS involves factorizing neutron flux within a reactor into a time-dependent amplitude and time-space-energy dependent shape.
+The purpose is to evaluate amplitude and shape on different time scales in order to reduce computational expense, while maintaining solution accuracy.
+IQS factorization leads to a nonlinear system of equations that requires iteration of shape and amplitude.
+Additionally, reactor simulation requires multiphysics simulation, where the quasi-static methodology can also be applied.
+
+The objectives of this research is to establish IQS performance with various iteration techniques, validate time step convergence of IQS, and apply IQS to multiphysics simulation.
+IQS iteration techniques involve fixed-point (Picard) iteration with various convergence criteria and Newton iteration, namely preconditioned Jacobian-free Newton Krylov (PJFNK) method.
+Nonlinear convergence of each of these techniques is investigated.
+Validation of IQS with analysis of time step convergence is vital for implementation of time adaptive methods and error prediction.
+The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicolson, backward difference formula (BDF), and singly-diagonally-implicit Runge-Kutta (SDIRK) methods.
+IQS application to multiphysics simulations involves its implementation into the Rattlesnake/MOOSE framework.
+These simulations allow insight to the performance of IQS for full transient reactor simulations.
+
+The results of the iteration convergence analysis show that the most rigorous and comprehensive iteration technique is fixed-point iteration with consistency in the IQS uniqueness specification as the convergence criteria.
+This iteration technique revealed the need for analytical treatment of the precursor equation for proper convergence.
+For time step convergence analysis, IQS was applied to a one-dimensional prototype example, as well as the TWIGL and LRA benchmark.
+The prototype results show that IQS has proper error convergence through fourth order discretization schemes.
+The TWIGL and LRA benchmark results show that IQS has proper convergence for implicit Euler, second-order BDF, and Crank-Nicolson schemes, validating IQS for more complex problems.
+For multiphysics simulation, IQS was applied to the LRA benchmark and a full core TREAT model.
+The results show that integration of temperature into the quasi-static process made considerable improvement to IQS performance.
+These results and conclusions helped gain insight into the behavior of IQS and furthered its development in complete transient reactor models.
 

### 2017-03-01 doc/Thesis/data/conclusion.tex
Thesis: adding conclusion

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2016.
%
%  Author: Sean Zachary Roberson 
%	 Version 3.08.16
%  Last updated 8/19/2016
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           SECTION IV
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{\uppercase{Conclusion}}

The goal of this thesis research is to continue the investigation and development of the improved quasi-static method for the optimization of transient reactor simulations. In pursuit of this goal, three objectives are formulated: establish IQS behavior for various iteration techniques, validate time step convergence for IQS, and apply IQS to multiphysics simulations. The following three sections describe the purpose of each objective, as well as conclusions on their results.

\section{Iteration Convergence Analysis}

IQS is a nonlinear system of equations; therefore, either fixed-point or Newton iteration is needed to evaluate shape and amplitude. Through literary review, most IQS applications used fixed-point iteration with various convergence criteria. Previous application of the Newton method is not discussed in detail, nor is any analysis of its convergence presented. Investigating these iteration techniques is important for understanding the behavior of IQS, as well as determining the most appropriate technique for optimal performance. The iteration techniques were applied to a one-dimensional prototype problem for testing. For fixed-point iteration, five different criteria were tested:

\begin{enumerate}
\item $L^\infty$ norm of the change in shape between iterations
\item $L^2$ norm of the change in shape between iterations
\item Difference in reactivity between iterations
\item Difference in amplitude between iterations
\item IQS uniqueness consistency criteria
\end{enumerate}

The results showed that criteria 1-4 had relatively equivalent convergence behavior. However, iteration with the criteria 5 could not converge without a highly-accurate analytical treatment of the precursor equation. This criteria proved to be the most rigorous and thorough convergence criteria and is recommended for any application of IQS.

\section{Time Step Convergence Analysis}

Time step convergence analysis involves evaluating a problem with various refinements in step size and comparing the resulting errors with the time step size. Plotting error versus $\Delta t$ on a log-scale should produce a relatively strait line with a slope equal to the order of the time discretization method. Validating IQS by demonstrating proper error convergence is essential for predicting model error and using time adaptation techniques. Implicit Euler and BDF discretization schemes were applied to the one-dimensional prototype problem. IQS showed proper error convergence up through fourth order BDF schemes. Implicit Euler and second order BDF schemes were applied to the TWIGL benchmark, where IQS showed proper error convergence. Step doubling adaptation was also applied to TWIGL, where IQS performed impressively, reducing the number of diffusion evaluations considerably. The Crank-Nicolson scheme was applied to the LRA benchmark with IQS. IQS again showed proper second order convergence and was able to reduce the number of diffusion evaluations with time adaptation by more than a factor of 6. The results of the TWIGL and LRA benchmarks proved that IQS is capable of proper convergence and time adaptation behavior for more complex problems.

\section{IQS Application to Multiphysics}

Full transient reactor simulation requires the implementation of multiphysics solution methods. In order to test IQS with these types of simulations, it was implemented into the MOOSE/Rattlesnake framework and executed using MAMMOTH. The LRA benchmark and a full core TREAT model were evaluated as test cases for multiphysics treatment. These examples involve a adiabatic heat up of the fuel with Doppler broadening feedback of cross-sections. The evaluation of temperature was integrated in the quasi-static process by introducing an intermediate time scale for temperature and PRKE parameter evaluation. The results of the LRA benchmark showed that quasi-static approach to temperature greatly improved the accuracy for a given time step size. However, the increase in computation time due to the extra temperature evaluations was significant. Therefore, for this benchmark, IQS with one update and IQS P-C with four updates is optimal. The results of the TREAT example showed similar behavior, except the updates only produced a marginal increase in accuracy. These examples' results and the dynamical time scale analysis show that performing a variant number of updates during the transient could further improve the performance of this quasi-static process. In conclusion, IQS and the quasi-static treatment of temperature can drastically improve the computational efficiency for burdensome full transient reactor simulations.

\section{Recommendations for Further Research}

This research intended to investigate IQS performance for transient reactor simulation and develop means for maximizing computational efficiency for these simulations. Further improvement and testing of IQS can be done in the following areas:

\begin{enumerate}

\item This thesis only applied IQS to fuel temperature feedback models. It is recommended that IQS be applied and tested with other feedback mechanisms such as thermal hydraulic and structural feedback, especially within the quasi-static process.

\item Application of IQS to neutron transport was hinted at in \sct{sect:transport}, but diffusion is typically sufficient for reactor simulation. It is recommended that IQS be applied to neutron transport with multiphysics and tested to see if similar performance is observed.

\item The large systems of full reactor models require a linear iteration process, such as GMRES, to evaluate efficiently. These processes usually implement a preconditioner to minimize the number of linear iterations and improve computation time. Investigating of an optimal preconditioner would further improve IQS performance in terms of computational efficiency.

\end{enumerate}

### 2017-03-01 doc/Thesis/data/kinetics.tex
Thesis: adding conclusion

@@
 
-\pagebreak
-\section{C5G7-TD Benchmark}
-
-The C5G7 benchmark is a MOX fueled, pressurized water reactor (PWR) minicore configuration.
+%\pagebreak
+%\section{C5G7-TD Benchmark}
+%
+%The C5G7 benchmark is a MOX fueled, pressurized water reactor (PWR) minicore configuration.
 The C5G7 geometry depicted in Fig.
@@
 Each assembly is made up of a regular 17-by-17 grid of pin cells.
-Seven energy-group cross sections for the seven material regions are given in \cite{c5g7}.
+%Seven energy-group cross sections for the seven material regions are given in \cite{c5g7}.
 The energy group boundaries are provided in Table \ref{tab:c5g7-ebounds}; three of the seven energy groups are fast, four are thermal.
-
-\begin{figure}[!htpb]
-\begin{center}
-\includegraphics[width=\linewidth]{\FiguresDir/C5G7-2D.png}
-\end{center}
-\caption{Geometry of the two-dimensional C5G7 benchmark problem. Boundary conditions are vacuum (V) or reflective (R).}
-\label{fig:c5g7-2D}
-\end{figure}
-
-\begin{figure}[!htpb]
-\begin{center}
-\includegraphics[width=0.5\linewidth]{\FiguresDir/C5G7-pincell.png}
-\end{center}
-\caption{Geometry of a C5G7 pin-cell.}
-\label{fig:c5g7-pincell}
-\end{figure}
-
-\begin{table}[!htpb]
-\centering
-\caption{Energy group boundaries for C5G7 MOX benchmark.}
-\label{tab:c5g7-ebounds}
-\begin{tabular}{c c}
-\hline
-Group & Upper Energy\\
-\hline
- 1&20 MeV  \\
- 2&1 MeV  \\
- 3&500 keV  \\
- 4&3 eV  \\
- 5&0.625 MeV  \\
- 6&0.1 MeV  \\
-71&0.02 MeV  \\
-\hline
-\end{tabular}
-\end{table}
-
-\pagebreak
-\section{LMW Benchmark?}
+%
+%\begin{figure}[!htpb]
+%\begin{center}
+%\includegraphics[width=\linewidth]{\FiguresDir/C5G7-2D.png}
+%\end{center}
+%\caption{Geometry of the two-dimensional C5G7 benchmark problem. Boundary conditions are vacuum (V) or reflective (R).}
+%\label{fig:c5g7-2D}
+%\end{figure}
+%
+%\begin{figure}[!htpb]
+%\begin{center}
+%\includegraphics[width=0.5\linewidth]{\FiguresDir/C5G7-pincell.png}
+%\end{center}
+%\caption{Geometry of a C5G7 pin-cell.}
+%\label{fig:c5g7-pincell}
+%\end{figure}
+%
+%\begin{table}[!htpb]
+%\centering
+%\caption{Energy group boundaries for C5G7 MOX benchmark.}
+%\label{tab:c5g7-ebounds}
+%\begin{tabular}{c c}
+%\hline
+%Group & Upper Energy\\
+%\hline
+% 1&20 MeV  \\
+% 2&1 MeV  \\
+% 3&500 keV  \\
+% 4&3 eV  \\
+% 5&0.625 MeV  \\
+% 6&0.1 MeV  \\
+%71&0.02 MeV  \\
+%\hline
+%\end{tabular}
+%\end{table}
+%
+%\pagebreak
+%\section{LMW Benchmark?}

### 2017-03-01 doc/Thesis/data/abstract.tex
updated abstract

@@
 
-\indent Transient reactor simulation is a chronically formidable challenge due to the computational rigor of evaluating the neutron transport equation, as well as its coupling with other physical phenomena.
+\indent Transient reactor simulations are a chronically formidable challenge due to the computational rigor of evaluating the neutron transport equation, as well as its coupling with other physical phenomena.
 The goal of this thesis research is to mitigate the computational expense of these simulations by investigating and developing the improved quasi-static method (IQS).
-IQS involves factorizing neutron flux within a reactor into a time-dependent amplitude and time-space-energy dependent shape.
-The purpose is to evaluate amplitude and shape on different time scales in order to reduce computational expense, while maintaining solution accuracy.
+IQS is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent amplitude and a time-space-energy dependent shape.
+The objective of the IQS factorization is to evaluate amplitude and shape on different time scales in order to reduce computational burden associated with solving the multi-dimensional flux equations, while maintaining solution accuracy.
 IQS factorization leads to a nonlinear system of equations that requires iteration of shape and amplitude.
-Additionally, reactor simulation requires multiphysics simulation, where the quasi-static methodology can also be applied.
+Furthermore, reactor simulations often require coupling to additional multiphysics components for temperature feedback.
+These additional physics may evolve on different time scales and the applicability of IQS in multiphysic simulations needs to be evaluated from this perspective as well.
 
@@
 Validation of IQS with analysis of time step convergence is vital for implementation of time adaptive methods and error prediction.
-The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicolson, backward difference formula (BDF), and singly-diagonally-implicit Runge-Kutta (SDIRK) methods.
+The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicolson, backward difference formulae (BDF), and singly-diagonally-implicit Runge-Kutta (SDIRK) methods.
 IQS application to multiphysics simulations involves its implementation into the Rattlesnake/MOOSE framework.

### 2017-03-01 doc/Thesis/data/introduction.tex
reviewed introduction

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter{\uppercase{Introduction}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Background on Transient Reactor Testing}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Since the March 2011 events at the Fukushima Daiichi Nuclear Power Plant in Japan, there has been significant governmental demand for the development of accident tolerant fuels in light water reactors (LWR) \cite{ConRepTREAT}.
-Currently, two facilities are capable of testing fuels under these extreme conditions: Annular Core Research Reactor (ACRR) at Sandia National Laboratory (SNL) and the Transient Reactor Testing Facility (TREAT) at Idaho National Laboratory (INL) \cite{SandiaTREAT}.
-TREAT had been put in stand-by status in 1994, but in 2014 a Final Environmental Assessment by the Department of Energy (DOE) approved the restart of the facility \cite{FONSI}.
+Currently, two facilities are capable of testing fuels under these extreme conditions: the Annular Core Research Reactor (ACRR) at Sandia National Laboratory (SNL) and the Transient Reactor Testing Facility (TREAT) at Idaho National Laboratory (INL) \cite{SandiaTREAT}.
+TREAT had been put in stand-by status in 1994, but in 2014 a Final Environmental Assessment by the Department of Energy (DOE) approved the restart of the facility in order to resume nuclear fuel testing\cite{FONSI}.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Time-Dependent Neutron Transport}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 In this research effort, FEM is used for space discretization and multigroup for energy discretization \cite{zeinkiewicz2005finite, duderstadt1976nuclear}.
-Since the purpose of this research involves temporal dependence, discretization in time is kept general.
-Most reactor computation frameworks discretize the time variable of flux directly using a multitude of schemes (Euler, Crank-Nicholson, Runge-Kutta, etc.).
+Since the purpose of this research involves temporal dependence, discussion of tempoeral discretization is kept general at this stage.
+\tcr{ok with you?} Most reactor computation frameworks discretize the time variable of flux directly using a multitude of schemes (Euler, Crank-Nicholson, Runge-Kutta, etc.).
 However, the solutions of flux can be particullary stiff due to the speed of neutrons, especially in a pulsing reactor \cite{TWIGL_benchmark}.
@@
 The following subsections describe two alternate techniques for transient simulation: point reactor kinetics and the improved quasi-static method (IQS).
-These methods are meant to reduce computational expense of evaluating transient flux while maintaining accuracy.
+These methods are meant to reduce computational expense of evaluating time-dependent neutron population distributions while maintaining accuracy.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Point Reactor Kinetics}
 
-A common solution process for neutron kinetics is the evaluation of the point reactor kinetics equation (PRKE) \cite{Planchard_1991}.
+A common solution process for neutron kinetics is the evaluation of the point reactor kinetics equations (PRKE) \cite{Planchard_1991}.
 The derivation of this equation involves factorizing flux into a space-dependent shape and time-dependent amplitude.
-This factorization immediately creates a large assumption that the spacial variance of the flux is time-independent.
-The amplitude has the same temporal stiffness as flux; however, the PRKE takes very little computational effort to evaluate because the many variables from the spatial discretization only need to be evaluated on the initial time step.
-The rest of the time steps just evaluate one variable: amplitude.
+This factorization immediately creates a large assumption that the spatial variance of the flux is time-independent.
+The amplitude has the same temporal stiffness as flux; however, the PRKE takes very little computational effort to evaluate because the many variables from the spatial discretization only need to be evaluated for the initital (steady state) distribution.
+For all times into the transient, the spatial shape is assumed fixed and one just need to evaluate the PRKE, a small system of ODEs,for the time-dependent amplitude.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Improved Quasi-Static Method}
@@
 These components are the flux's amplitude and its shape.
-Amplitude is only time-dependent, while the shape is both space- and time-dependent.
-However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the variable for the diffusion evaluation is less stiff without making the assumption that spatial variance is time-independent, as the PRKE does.
+Amplitude is only time-dependent, while the shape is both space- and time-dependent (as opposed to the PRKE approach where the shape was constant in time).
+However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the variable for the multi-D diffusion solves is expected to be less stiff than the flux itself, without making the assumption that spatial variance is time-independent.
 As a result, the shape may not require an update at the same frequency as the amplitude function, but only on larger macro-time steps.
 
-Since shape and amplitude share an independent variable, they are temporally coupled.
-Ott in \cite{Ott_1966} was the first publication that investigated the coupling of shape and amplitude in a quasi-static nature, but did not include the time derivative in the shape equation.
-Ott later in \cite{Ott_1969} incorporated the time derivative of shape in the equation, yielding better results.
-The resulting system of equations is heavily nonlinear \cite{Dulla2008}.
+Due to the factorization of the flux into a shape and an amplitude, the latter two variables are coupled.
+Ott in \cite{Ott_1966} first investigated the coupling of shape and amplitude in a quasi-static nature, but did not include the time derivative in the shape equation.
+Later, in \cite{Ott_1969}, Ott incorporated the time derivative of shape in the equation, yielding better results; this also led to the technique's name: the Improved Quasi-Static method.
+The resulting system of equations is  nonlinear \cite{Dulla2008}.
 Nonlinear problems require an iterative process to evaluate, either fixed-point (Picard) or Newton iteration.
@@
 \cite{Koclas_1996}, Devooght et al.
-\cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their quasi-static simulations, the main difference among them is their criteria for convergence.
+\cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their IQS simulations, the main difference among them is their criteria for convergence.
 Devooght et al.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Time-Dependent Reactor Dynamics}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Transient reactor analysis also involves coupling neutronics with other physics like temperature feedback, thermal hydraulics, fuel structure and composition, etc.
+Transient reactor analysis  involves coupling neutronics with other physics such as thermal hydraulics, fuel mechanics, etc.
+to to enable temperature feedback in the neutronics simulations.
 This coupling is known as multiphysics simulation.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Objective}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\tcr{
-I think an outline for the introduction should be 
-\begin{enumerate}
-\item reactor transient testing.
-Background, etc...
-this subsection should conclude on the overarching need of transient testing: computational methods for multiphysics reactor transients.
-It should also introduce the next 2 subsections, which are both about lit review, one on time-dependent  neutronics, one on multiphysics solution techniques for reactor physics.
-\item literature review on mostly neutronics time dependent: challenges in computational methods for time-dependent neutron balance equations.
-the size of the phase-space, the need for implicitness in the temporal treatment due to the stiffness, ....
-this should open up to a discussion on alternate methods, such as IQS, ...
-it is a lit review, so no need for equations, but rather a demonstration that you know the background and have a solid understand of that foundation to lay the ground from your OWN work.
-As part of this lit, some discussion of Mund's, Dulla's, Ott's, Monier,'s the Canadian's papers is welcome.
-\item now, reactor transients also means coupled to other physics.
-this is were you discuss some aspect related to multiphysics.
-you can start with the general discussion that this is typically a nonlinear problem, that it can be solved via Picard or newton.
-this is maybe a segway into a discussion/introduction about MOOSE.
-Then you go back on IQS and add the multiphysics aspects.
-Ott did some work with multiphysics, I gave you the code manual of his own software where he discusses when he decided to do shape updates.
-\end{enumerate}
-Generally speaking, the rest is real good but you need to spend some time on these 5-8 pages of introduction that set the tone and the background.
-Everybody needs to understand that you know a lot; that your thoughts are well organized and clear; the rest will come easy (except some results ...).\\
-side note: we need to put the small results about sdirk being poor for stiffness problem. maybe as an appendix. I want to keep that knowledge somewhere. 
-}

### 2017-03-01 doc/Thesis/data/theory.tex
Thesis: Adding IQS iteration analysis
reviewed theory chapter

@@
 The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.\\  
+\tcr{I vaguely remind something about $\bar{\lambda}$ not being so easy to obtain/derive/or conceptualize. What was that story?}
 
@@
 \be
-\frac{\partial}{\partial t}\left(\frac{\Psi^g}{v^g}\right) = \sum_{g'} \left(H^{g'\to g} + P_p^{g' \to g} \right) \Psi^{g'} - L^g\Psi^g + S_d^g
+\frac{1}{v^g}\frac{\partial \Psi^g}{\partial t} = \sum_{g'} \left(H^{g'\to g} + P_p^{g' \to g} \right) \Psi^{g'} - L^g\Psi^g + S_d^g
 \label{eq:transport}
@@
 \be
-\frac{\partial}{\partial t}\left(\frac{\psi^g}{v^g}\right) = \sum_{g'} \left(H^{g' \to g} + P_p^{g' \to g}\right) \psi^{g'} - \left( L^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \psi^g + \frac{1}{p} S_d^g
+\frac{1}{v^g}\frac{\partial \psi^g}{\partial t}  = \sum_{g'} \left(H^{g' \to g} + P_p^{g' \to g}\right) \psi^{g'} - \left( L^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \psi^g + \frac{1}{p} S_d^g
 \label{eq:tshape}
@@
 There is no guarantee that the shape converges, so max number of iterations is typically enforced.
+\tcr{rephrase last bit}
 
@@
 \ee
+\tcr{here, and before, $n$ seems to indicate the current time at which the solution is sought. I believe that later, the unknowns
+will be described as being at $n+1$. this is a slight inconsistency.}
 
@@
 In nuclear reactors, multiple physics affect the profile of the neutron flux.
-The most simple example of mulitphysics reactor simulations is adiabatic heat up with Doppler feedback.
+One of the simplest examples of mulitphysics reactor simulations is adiabatic heat up with Doppler feedback.
 The principle of Doppler feedback is that fission in a fuel causes the material to increase temperature and induces a change in the neutronics properties.
@@
 \fig{fig:proc} shows the programming diagram implement to execute this process.
-The time increment of $\Delta t/3$ for the temperature solve is arbitrary and is meant only to match \fig{fig:time}.
+The time increment of $\Delta t/3$ for the temperature solve is arbitrary and is meant only to match \fig{fig:time} where three temperature updates have been used as illustration.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Backward Difference Formulas}
+\subsection{Backward Difference Formulae}
 
-An extension of implicit discretization is the backward difference formulas (BDF) \cite{Gear:2007}.
+An extension of implicit discretization is the backward difference formulae (BDF) \cite{Gear:2007}.
 BDF's can increase the order of error convergence by interpolating solutions from previous time time steps.
-The general formula for BDF is discribed by \eqt{eq:bdf}.
+The general formula for BDF is described by \eqt{eq:bdf}.
 \be 
@@
 Where $k$ is the order of the error convergence and $\alpha_{ij}$ are coefficients chosen such that the temporal truncation error is minimized.
-The benifit of using BDF is that any order of convergence can be applied and are all unconditionally stable, so the IQS approximation can easily be validated for high order discretization.
+The benefit of using BDF is that any order of convergence can be applied and are all unconditionally stable, so the IQS approximation can easily be validated for high order discretization.
+\tcr{The $k$ subscript is confusing. It reads like $\alpha_{jk}$ is a matrix entry but the rest of the equation 2.35 is independent of $k$. Something is fishy.}
 
@@
 C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta}{1+\theta\Delta t\lambda}S_f^n p^n +  \frac{\theta\Delta t \beta}{1+\theta\Delta t\lambda}S_f^{n+1} p^{n+1}
+\label{eq:dnp_theta}
 \ee
@@
 C^{n+1} = C^n e^{-\lambda \Delta t} + \left(\hat{a}_2 S_f^{n+1}+\hat{a}_1 S_f^n\right)\beta
+\label{eq:dnp_an}
 \ee
@@
 
-IQS will be made available to any action system capable of transient simulation, which includes:
+IQS will be made available \tcr{this is a copy paste from an old report but it is no longer fully accurate, right?}
+to any action system capable of transient simulation, which includes:
 \begin{itemize}

### 2017-03-01 doc/Thesis/data/conclusion.tex
updated conclusions

@@
 
-\chapter{\uppercase{Conclusion}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{\uppercase{Conclusions}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\tcr{I am not sure but the present tense sounds weird when you conclude. How about a reformation as follows:
+In this work, we analyzed/investigated ....}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Iteration Convergence Analysis}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Time Step Convergence Analysis}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Plotting error versus $\Delta t$ on a log-scale should produce a relatively strait line with a slope equal to the order of the time discretization method.
+\tcr{you are describing how to visualize convergence order.
+This should not be said here, in the conclusions, but before.}
 Validating IQS by demonstrating proper error convergence is essential for predicting model error and using time adaptation techniques.
@@
 Step doubling adaptation was also applied to TWIGL, where IQS performed impressively, reducing the number of diffusion evaluations considerably.
-The Crank-Nicolson scheme was applied to the LRA benchmark with IQS.
+The Crank-Nicholson scheme was applied to the LRA benchmark with IQS.
 IQS again showed proper second order convergence and was able to reduce the number of diffusion evaluations with time adaptation by more than a factor of 6.
 The results of the TWIGL and LRA benchmarks proved that IQS is capable of proper convergence and time adaptation behavior for more complex problems.
+\tcr{Popov may balk at the repeated use of proper, but that's ok. If he says nothing, we let it go.}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{IQS Application to Multiphysics}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Recommendations for Further Research}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 These processes usually implement a preconditioner to minimize the number of linear iterations and improve computation time.
-Investigating of an optimal preconditioner would further improve IQS performance in terms of computational efficiency.
+Investigating of an optimal preconditioner would \tcr{further} improve IQS performance in terms of computational efficiency.
+\tcr{this reminds me that you may want to describe a bit the types of linear solution techniques (and their associated preconditioners) used when employing Rattlesnake.
+This is missing from the document I believe}
 

### 2017-03-01 doc/Thesis/data/conclusion.tex
Thesis: addressing latest comments

--- 2017-03-01 doc/Thesis/data/conclusion.tex
@@
 
-\chapter{\uppercase{Conclusion}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{\uppercase{Conclusions}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-The goal of this thesis research is to continue the investigation and development of the improved quasi-static method for the optimization of transient reactor simulations.
-In pursuit of this goal, three objectives are formulated: establish IQS behavior for various iteration techniques, validate time step convergence for IQS, and apply IQS to multiphysics simulations.
+The goal of this thesis research was to continue the investigation and development of the improved quasi-static method for the optimization of transient reactor simulations.
+In pursuit of this goal, three objectives were formulated: establish IQS behavior for various iteration techniques, validate time step convergence for IQS, and apply IQS to multiphysics simulations.
 The following three sections describe the purpose of each objective, as well as conclusions on their results.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Iteration Convergence Analysis}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-IQS is a nonlinear system of equations; therefore, either fixed-point or Newton iteration is needed to evaluate shape and amplitude.
+IQS is a nonlinear system of equations; therefore, either fixed-point or Newton iteration were needed to evaluate shape and amplitude.
 Through literary review, most IQS applications used fixed-point iteration with various convergence criteria.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Time Step Convergence Analysis}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Time step convergence analysis involves evaluating a problem with various refinements in step size and comparing the resulting errors with the time step size.
-Plotting error versus $\Delta t$ on a log-scale should produce a relatively strait line with a slope equal to the order of the time discretization method.
 Validating IQS by demonstrating proper error convergence is essential for predicting model error and using time adaptation techniques.
@@
 IQS showed proper error convergence up through fourth order BDF schemes.
-Implicit Euler and second order BDF schemes were applied to the TWIGL benchmark, where IQS showed proper error convergence.
+Implicit Euler and second order BDF schemes were applied to the TWIGL benchmark, where IQS showed normal error convergence.
 Step doubling adaptation was also applied to TWIGL, where IQS performed impressively, reducing the number of diffusion evaluations considerably.
-The Crank-Nicolson scheme was applied to the LRA benchmark with IQS.
-IQS again showed proper second order convergence and was able to reduce the number of diffusion evaluations with time adaptation by more than a factor of 6.
+The Crank-Nicholson scheme was applied to the LRA benchmark with IQS.
+IQS again showed expected second order convergence and was able to reduce the number of diffusion evaluations with time adaptation by more than a factor of 6.
 The results of the TWIGL and LRA benchmarks proved that IQS is capable of proper convergence and time adaptation behavior for more complex problems.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{IQS Application to Multiphysics}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2017-03-01 doc/Thesis/data/introduction.tex
Thesis: addressing latest comments

@@
 Since the purpose of this research involves temporal dependence, discussion of tempoeral discretization is kept general at this stage.
-\tcr{ok with you?} Most reactor computation frameworks discretize the time variable of flux directly using a multitude of schemes (Euler, Crank-Nicholson, Runge-Kutta, etc.).
+Most reactor computation frameworks discretize the time variable of flux directly using a multitude of schemes (Euler, Crank-Nicholson, Runge-Kutta, etc.).
 However, the solutions of flux can be particullary stiff due to the speed of neutrons, especially in a pulsing reactor \cite{TWIGL_benchmark}.

### 2017-03-01 doc/Thesis/data/kinetics.tex
Thesis: addressing latest comments

@@
 
+Time step convergence analysis involves evaluating a problem with various refinements in step size and comparing the resulting errors with the time step size.
+Plotting error versus $\Delta t$ on a log-scale should produce a relatively strait line with a slope equal to the order of the time discretization method.
 In order to evaluate the performance and error convergence of IQS, the slab was simulated with varying time discretization methods and time step sizes.

### 2017-03-01 doc/Thesis/data/theory.tex
Thesis: addressing latest comments

@@
 \end{subequations}
-The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.\\  
-\tcr{I vaguely remind something about $\bar{\lambda}$ not being so easy to obtain/derive/or conceptualize. What was that story?}
+The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.
 
 Additionally, in order to impose uniqueness on the factorization and to derive the PRKE, the following normalization condition is imposed: $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}$ \cite{Dulla2008}.
-\\
 
@@
 Where $n$ is the time step, $k$ is the iteration number, and $\epsilon_{\varphi}$ is the numerical criteria provided by a user.
-There is no guarantee that the shape converges, so max number of iterations is typically enforced.
-\tcr{rephrase last bit}
+Through testing, iterations do not always converge this criteria; in this circumstance, max number of iterations is enforced.
 
@@
 \ee
-\tcr{here, and before, $n$ seems to indicate the current time at which the solution is sought. I believe that later, the unknowns
-will be described as being at $n+1$. this is a slight inconsistency.}
 
@@
 \subsection{Preconditioned Jacobian-Free Newton-Krylov}
+\label{sect:pjfnk}
 
@@
 Where $k$ is the order of the error convergence and $\alpha_{ij}$ are coefficients chosen such that the temporal truncation error is minimized.
-The benefit of using BDF is that any order of convergence can be applied and are all unconditionally stable, so the IQS approximation can easily be validated for high order discretization.
-\tcr{The $k$ subscript is confusing. It reads like $\alpha_{jk}$ is a matrix entry but the rest of the equation 2.35 is independent of $k$. Something is fishy.}
+The benefit of using BDF is that any order of convergence can be applied and are all unconditionally stable, so the IQS approximation can easily be validated for high order discretization.S
 
@@
 However, in the derivation for IQS, this term is unaffected between shape and flux evaluation.
-So saving the residual for this diffusion kernel in the \texttt{save\_in} variable is the only alteration to this action for IQS to function.\\
+So saving the residual for this diffusion kernel in the \texttt{save\_in} variable is the only alteration to this action for IQS to function.
+
 Additional action systems involve transport, but it is evident from \sct{sect:transport} that IQS implementation in these is straightforward as well.
 The main differences between a diffusion implementation and a transport implementation are outlined below:
+
 \begin{enumerate}
@@
 
-IQS will be made available \tcr{this is a copy paste from an old report but it is no longer fully accurate, right?}
-to any action system capable of transient simulation, which includes:
+It is possible to implement IQS to any action system capable of transient simulation, which includes:
 \begin{itemize}
@@
 \end{itemize}
+However, for TREAT models and examples simulated in this research, CFEM-Diffusion is a sufficient approximation. Therefore, IQS was only implemented and tested in this action system.
 
@@
 \end{lstlisting}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Linear Solution Techniques}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+MOOSE contains a suite of solution techniques to evaluate a linear system generated by the Rattlesnake action system.
+For most neutronics and reactor applications, direct evaluation of the system is overbearingly expensive, both computationally and by memory usage.
+Therefore, the most common technique (default in MOOSE) is PJFNK.
+This technique is briefly described in \sct{sect:pjfnk}.
+PJFNK entails approximately solving the Jacobian action on the solution with a finite difference approach using GMRES iteration.
+Each GMRES iteration requires a residual evaluation, which becomes globally expensive with many iterations.
+Employing a preconditioner on the solution can decrease the number of iterations and save execution time.
+The purpose of the preconditioner is to resemble the Jacobian (reduce number of iterations), while being easily invertable (preconditioner overhead).
+MOOSE has four useful assembly methods for building an effective preconditioner:
+
+\begin{enumerate}
+\item Block Diagonal Preconditioning (default)
+\item Single Matrix Preconditioner (SMP)
+\item Finite Difference Preconditioner (FDP)
+\item Physics Based Preconditioner (PBP)
+\end{enumerate}
+
+When developing a kernel in MOOSE, a function is implement that evaluates the Jacobian for each variable it depends on.
+In a multi-variable system, like multigroup diffusion, the Jacobian contains "blocks" for each variable and the coupling between them.
+Diagonal blocks contain the variable contribution to itself, i.e.
+diffusion and removal.
+The off-diagonal blocks contain variable-to-variable contributions, i.e.
+scattering and fission.
+The default method assembles only the diagonal blocks; this preconditioner is relatively easy to invert, but is not optimal for reducing linear iterations.
+SMP assembles the diagonal blocks and any off-diagonal blocks specified by the user.
+Assembling more off-diagonal blocks reduces linear iterations (inputting all blocks requires only one GMRES iteration for a linear system), but increases the expense of inverting the preconditioner.
+Therefore, SMP requires extensive analysis for optimization.
+FDP simply does a finite difference approximation for the Jacobian based on residual evaluations.
+This method only requires one GMRES iteration for a linear system, but is impractically expensive.
+PBP is an advanced adaptation to SMP.
+This method assembles the preconditioner similarly to SMP, but each block is inverted as a independent system and iteratively communicates with the rest of the blocks.
+PBP can be effective for a lower block diagonal system, i.e.
+including only downscattering blocks, because no communication iterations are required.
+Selecting an effective preconditioner and assembly method is essential for maximizing computational efficiency of large reactor simulations.

### 2017-03-01 doc/Thesis/data/theory.tex
some last minute update to the theory section

@@
 Where $n$ is the time step, $k$ is the iteration number, and $\epsilon_{\varphi}$ is the numerical criteria provided by a user.
-Through testing, iterations do not always converge this criteria; in this circumstance, max number of iterations is enforced.
+Through testing, iterations do not always converge this criteria; in this circumstance, the maximum number of iterations is reached.
 
@@
 \be 
-\varphi_{n+1} = (IV-\Delta t A_{n+1})^{-1}\left[IV\sum_{j=0}^{k-1}\alpha_{jk}\varphi_{n-j} +
-\Delta t b_{n+1}\right]
+\varphi_{n+1} = (IV-\Delta t A_{n+1})^{-1}\left[IV\sum_{j=0}^{k-1}\alpha_{j}^{k}\varphi_{n-j} +
+\Delta t \alpha_{k+1}^{k} b_{n+1}\right]
 \label{eq:bdf}
 \ee
-Where $k$ is the order of the error convergence and $\alpha_{ij}$ are coefficients chosen such that the temporal truncation error is minimized.
-The benefit of using BDF is that any order of convergence can be applied and are all unconditionally stable, so the IQS approximation can easily be validated for high order discretization.S
-
+where $k$ is the order of the error convergence and $\alpha_{j}^{k}$ (vector of length $k+1$) are coefficients chosen such that the temporal truncation error is minimized.
+The benefit of using BDF is that any order of convergence can be applied and are all unconditionally stable, so the IQS approximation can easily be validated for high order discretization.
+\tcr{I think one term was missing in your eq 2.35. also, it would be good to given the alpha values for the methods you implemented}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\tcr{JFNK is really for nonlinear systems. I would avoid confusing the reader here. I would start with PJFNK and Picard are
+the nonlinear solvers in MOOSE. Each of the nonlinear solvers requires linear solves ... and then you talk about linear solves ...}
 
-MOOSE contains a suite of solution techniques to evaluate a linear system generated by the Rattlesnake action system.
-For most neutronics and reactor applications, direct evaluation of the system is overbearingly expensive, both computationally and by memory usage.
+MOOSE contains a suite of solution techniques to solve the linear system generated by the Rattlesnake action system.
+For most neutronics and reactor applications, direct inversion (e.g., LU solve) of the system is overbearingly expensive, both computationally and by memory usage.
 Therefore, the most common technique (default in MOOSE) is PJFNK.
@@
 In a multi-variable system, like multigroup diffusion, the Jacobian contains "blocks" for each variable and the coupling between them.
-Diagonal blocks contain the variable contribution to itself, i.e.
-diffusion and removal.
+Diagonal blocks contain the variable contribution to itself, i.e., diffusion and removal.
 The off-diagonal blocks contain variable-to-variable contributions, i.e.
@@
 This method assembles the preconditioner similarly to SMP, but each block is inverted as a independent system and iteratively communicates with the rest of the blocks.
-PBP can be effective for a lower block diagonal system, i.e.
-including only downscattering blocks, because no communication iterations are required.
+PBP can be effective for a lower block diagonal system, i.e., including only downscattering blocks, because no communication iterations are required.
 Selecting an effective preconditioner and assembly method is essential for maximizing computational efficiency of large reactor simulations.

### 2017-03-01 doc/Thesis/data/dynamics.tex
Thesis: adding conclusion
more comments pushed. almost there

@@
 This section describes several dynamics examples, including the LRA benchmark and two TREAT experiments.
-These examples are of increased complexity from the kinetics examples of \sct{sect:kin}.
-This section also analyzes IQS's performance with these, which is vital for verification of IQS in real-world problems.
+These examples are of increased complexity from the kinetics examples of \chpt{sect:kin}.
+This Chapter also analyzes IQS's performance with these, which is vital for verification of IQS in real-world problems.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{LRA Benchmark}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 To have better understanding on the cross sections given later, we present the equations here:
+\tcr{ooops, wrong sign on the RHS for neutronics!}
 \begin{subequations}
@@
 -\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 - \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
--\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
+-\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\div D_2 \grad\phi_2 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
 S_f &= \sum_{g=1}^2 \Sigma_{f,g} \phi_g, \\
@@
 \end{subequations}
-where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0=T(\vec{r}, t=0)$.
+where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0(\vec{r})=T(\vec{r}, t=0)$.
 The two-group diffusion equation are solved with zero flux boundary conditions on external surfaces, reflecting conditions at symmetry boundaries and steady state initial conditions which are obtained by solving
@@
 -\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
--\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1.
+-\div D_2 \grad\phi_2 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1.
 \end{align}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{LRA Multiphysics Time Scale Results}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{LRA with Time Adaptation}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{TREAT Transient-15 Problem}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \fig{fig:Tran15} shows a visualization of the flux profile within the core, hidden is the massive amount of graphite surrounding the core.
-
+\tcr{I do not see/understand the air channel meshing}
 \begin{figure}[htbp!]
@@
 \includegraphics[width=0.7\linewidth]{\FiguresDir/Tran15_config.png}
-\caption{Transient 15 159-element small core configuration}
+\caption{Transient-15 159-element small core configuration}
 \label{fig:Tran15_config}
@@
 \end{subfigure}
-\caption{Top quarter view of Transient 15 mesh}
+\caption{Top quarter view of Transient-15 mesh}
 \label{fig:Tran15_mesh}
@@
 \includegraphics[width=0.6\linewidth]{\FiguresDir/Tran15_core2.png}
-\caption{Transient 15 core power profile at peak power}
+\caption{Transient-15 core power profile at peak power}
 \label{fig:Tran15}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Transient-15 Temperature Feedback}
@@
 \be
-c_p = -5.8219\text{e-}10T^3 - 4.3694\text{e-}7T^2 + 2.8369\text{e-}3T -1.009\text{e-}2 \ (J/g/K)
+c_p = -5.8219\times 10^{-10}T^3 - 4.3694\times 10^{-7}T^2 + 2.8369\times 10^{-3}T -1.009\times 10^{-2} \ (J/g/K)
 \label{eq:cp}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Transient-15 Multiphysics Time Scale Results}
@@
 \includegraphics[width=\linewidth]{\FiguresDir/Tran15_profile.png}
-\caption{Transient 15 total power and average temperature profile during transient}
+\caption{Transient-15 total power and average temperature profile during transient}
 \label{fig:Tran15_profile}
@@
 \includegraphics[width=0.7\linewidth]{\FiguresDir/time_constant_tran15.png}
-\caption{Dynamical time scale for the Transient 15 example}
+\caption{Dynamical time scale for the Transient-15 example}
 \label{fig:tran15tc}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Transient-15 with Time Adaptation}
@@
 
-\section{TREAT M8-CAL Problem}
+%\section{TREAT M8-CAL Problem}

### 2017-03-02 doc/Thesis/data/dynamics.tex
Thesis: Final changes before sending to committee

@@
 This feedback includes coupling with other physical properties such as temperature, fluids, material dynamics, etc.
-This section describes several dynamics examples, including the LRA benchmark and two TREAT experiments.
+This section describes two dynamics examples, including the LRA benchmark and a TREAT experiment.
 These examples are of increased complexity from the kinetics examples of \chpt{sect:kin}.
@@
 To have better understanding on the cross sections given later, we present the equations here:
-\tcr{ooops, wrong sign on the RHS for neutronics!}
+
 \begin{subequations}
@@
 where $V_{core}$ is the core region with fuels, is equal to $10^{-6} W\cdot cm^{-3}$.
-The initial precursor concentrations are in equilibrium with the initial critical flux distribution.\\
+The initial precursor concentrations are in equilibrium with the initial critical flux distribution.
 
-The geometry is illustrated in~\fig{fig:lra-geometry}.\\
+The geometry is illustrated in~\fig{fig:lra-geometry}.
 \begin{figure}[!htbp]
@@
 All fuel materials have the same delayed neutron data.
-Some scalar data are listed in~\tbl{tab:lra-scalar}.\\
+Some scalar data are listed in~\tbl{tab:lra-scalar}.
 \begin{table}[!htbp]
@@
 \fig{fig:Tran15_mesh} shows the meshing techniques of the model \cite{mammoth}.
+\fig{fig:Tran15_mesh_zoom} shows the single block mesh for each meshing technique, the red area of \fig{fig:Tran15_mesh_fine_zoom} represents the air gap.
 To ease computation, the fully homogenized version (without air gaps) was used  for the following simulations.
@@
 \fig{fig:Tran15} shows a visualization of the flux profile within the core, hidden is the massive amount of graphite surrounding the core.
-\tcr{I do not see/understand the air channel meshing}
+
 \begin{figure}[htbp!]
@@
 
+\begin{figure}[!htbp]
+\centering
+\begin{subfigure}[!htbp]{0.49\textwidth}
+\includegraphics[width=\textwidth]{\FiguresDir/Tran15_mesh_fine_zoom.png}
+\caption{Explicit meshing of air channels}
+\label{fig:Tran15_mesh_fine_zoom}
+\end{subfigure}
+\begin{subfigure}[!htbp]{0.49\textwidth}
+\includegraphics[width=\textwidth]{\FiguresDir/Tran15_mesh_homo_zoom.png}
+\caption{Fully homogenized fuel elements}
+\label{fig:Tran15_mesh_homo_zoom}
+\end{subfigure}
+\caption{Top single block view of Transient-15 mesh}
+\label{fig:Tran15_mesh_zoom}
+\end{figure}
+
 \begin{figure}[htbp!]
@@
 \eqt{eq:temp2} describes the heat up of the fuel.
-It is very similar, except the specific heat is now dependent on temperature is described by \eqt{eq:cp}.
+It is very similar, except the specific heat is now dependent on temperature, which is described by \eqt{eq:cp}.
 The temperature evaluation is identical to the one described in LRA section, except a Newton iteration process is employed to resolve the nonlinearity from the specific heat term.
@@
 \centering
-\begin{subfigure}[!htbp]{0.49\textwidth}
+\begin{subfigure}[!htbp]{0.75\textwidth}
 \includegraphics[width=\textwidth]{\FiguresDir/Tran15_DT2.png}
@@
 \end{subfigure}
-\begin{subfigure}[!htbp]{0.49\textwidth}
+\begin{subfigure}[!htbp]{0.75\textwidth}
 \includegraphics[width=\textwidth]{\FiguresDir/Tran15_DT2_peak.png}

### 2017-03-02 doc/Thesis/data/introduction.tex
Thesis: Final changes before sending to committee

@@
 The amplitude has the same temporal stiffness as flux; however, the PRKE takes very little computational effort to evaluate because the many variables from the spatial discretization only need to be evaluated for the initital (steady state) distribution.
-For all times into the transient, the spatial shape is assumed fixed and one just need to evaluate the PRKE, a small system of ODEs,for the time-dependent amplitude.
+For all time steps in the transient, the spatial shape is assumed fixed and only the PRKE, a small system of ODEs, needs to be evaluated for the time-dependent amplitude.
 
@@
 Amplitude is only time-dependent, while the shape is both space- and time-dependent (as opposed to the PRKE approach where the shape was constant in time).
-However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the variable for the multi-D diffusion solves is expected to be less stiff than the flux itself, without making the assumption that spatial variance is time-independent.
+However, the impetus of the method is the assumption that the shape is only weakly dependent on time.
+Therefore, the variable for the multi-dimensional diffusion evaluation is expected to be less stiff than the flux itself, without making the assumption that spatial variance is time-independent.
 As a result, the shape may not require an update at the same frequency as the amplitude function, but only on larger macro-time steps.
@@
 Transient reactor analysis  involves coupling neutronics with other physics such as thermal hydraulics, fuel mechanics, etc.
-to to enable temperature feedback in the neutronics simulations.
+to enable temperature feedback in the neutronics simulations.
 This coupling is known as multiphysics simulation.

### 2017-03-02 doc/Thesis/data/kinetics.tex
more comments pushed. almost there
Thesis: Final changes before sending to committee

@@
 
-\chapter{KINETICS EXAMPLES}
+\chapter{KINETICS EXAMPLES}\label{sect:kin}
 
@@
 The first is a one-dimensional problem, designed for the prototype code in MATLAB.
-The next four are from the Argonne National Lab (ANL) Benchmark Problem Book (BPB), and are common problems for testing codes and developing methods \cite{ANL_BPB}.
+The next one is from the Argonne National Lab (ANL) Benchmark Problem Book (BPB), and is a common problem for testing codes and developing methods \cite{ANL_BPB}.
 
@@
 \fig{fig:1D_shape} shows the shape profile at various times during the transient.
-It is apparent that the shape is very time-dependent, so it is expected that IQS has marginal accuracy gain for a given time step.
+It is apparent that the shape is very time-dependent, so it is expected that IQS has marginal accuracy gain for a given time step size.
 
@@
 \includegraphics[width=\linewidth]{\FiguresDir/1D_power_base.png}
-\caption{Baseline power profile of 1D slab}
+\caption{Baseline total power level for the 1D slab test case}
 \label{fig:1D_power}
@@
 The results differ slightly from the prototype, especially in the fact that IQS P-C performs better than IQS.
-Since Rattlesnake uses a PJFNK solver for the FEM model, number of time steps isn't strictly proportional to the execution time.
+Since Rattlesnake uses a PJFNK solver for the FEM model, the number of time steps isn't strictly proportional to the execution time.
 \fig{fig:1D_conv_lin} shows the error of the three methods as compared to the number of linear GMRES iterations, which is a better mark for comparison in computation time.

### 2017-03-02 doc/Thesis/data/theory.tex
Thesis: Final changes before sending to committee

@@
 $\Sigma_f^{g}$	&	$=$	&	Fission cross section in energy group $g$ \\
+$k_{eff}$	&	$=$	&	k-eigenvalue of initial flux evaluation \\
 $\Sigma_r^{g}$	&	$=$	&	Removal cross section in energy group $g$ \\
@@
 To derive the amplitude equation, the shape/precursors equations are weighted by a space-dependent function and integrated over the phase-space.
-The weight function is typically the adjoint flux $\phi^{*g}$, which can be proven to minimize truncation error \cite{duderstadt1976nuclear}.
+The weight function is typically the adjoint flux $\phi^{*g}$ at the initial time, which can be proven to minimize truncation error \cite{duderstadt1976nuclear}.
 The final expressions are given below:
@@
 Monier in \cite{Monier_diss} also does fixed point iterations with the same criteria, except the solution is scaled by $\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n})\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n+1})\right)}$ after each iteration.
-Devooght in \cite{Devooght_1984} does a Newton-SOR iteration where the residual of the shape function evaluation is the convergence criteria and next iteration's solution is computed using Newton-Raphson method.\\
+Devooght in \cite{Devooght_1984} does a Newton-SOR iteration where the residual of the shape function evaluation is the convergence criteria and next iteration's solution is computed using Newton-Raphson method.
 
@@
 Where $n$ is the time step, $k$ is the iteration number, and $\epsilon_{\varphi}$ is the numerical criteria provided by a user.
-Through testing, iterations do not always converge this criteria; in this circumstance, the maximum number of iterations is reached.
+Through testing, iterations do not always converge; in this circumstance, the maximum number of iterations is reached.
 
@@
 An extension of implicit discretization is the backward difference formulae (BDF) \cite{Gear:2007}.
-BDF's can increase the order of error convergence by interpolating solutions from previous time time steps.
+BDF's can increase the order of error convergence by interpolating solutions from previous time steps.
 The general formula for BDF is described by \eqt{eq:bdf}.
 \be 
-\varphi_{n+1} = (IV-\Delta t A_{n+1})^{-1}\left[IV\sum_{j=0}^{k-1}\alpha_{j}^{k}\varphi_{n-j} +
+\varphi_{n+1} = (IV-\Delta t A_{n+1})^{-1}\left[IV\sum_{j=1}^{k}\alpha_{j}^{k}\varphi_{n-(k-j)} +
 \Delta t \alpha_{k+1}^{k} b_{n+1}\right]
@@
 where $k$ is the order of the error convergence and $\alpha_{j}^{k}$ (vector of length $k+1$) are coefficients chosen such that the temporal truncation error is minimized.
-The benefit of using BDF is that any order of convergence can be applied and are all unconditionally stable, so the IQS approximation can easily be validated for high order discretization.
-\tcr{I think one term was missing in your eq 2.35. also, it would be good to given the alpha values for the methods you implemented}
+\tbl{tab:bdfa} shows the values of $\alpha$ for every order implemented in this thesis.
+The benefit of using BDF is that any order of convergence can be applied and are unconditionally stable up through fourth order, so the IQS approximation can easily be validated for high order discretization.
+
+\begin{table}
+\centering
+\begin{tabular}{c|lllll}
+\hline
+Order ($k$) & $\alpha_{1}^{k}$ & $\alpha_{2}^{k}$ & $\alpha_{3}^{k}$ & $\alpha_{4}^{k}$ & $\alpha_{5}^{k}$ \\
+\hline
+1 & 1 & 1 & & & \\
+2 & -1/3 & 4/3 & 2/3 & & \\
+3 & 2/11 & -9/11 & 18/11 & 6/11 & \\
+4 & -3/25 & 16/25 & -36/25 & 48/25 & 12/25 \\
+\hline
+\end{tabular}
+\caption{$\alpha$ values for relevant BDF methods}
+\label{tab:bdfa}
+\end{table}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Once $\varphi^{n+1}$ has been determined, $C^{n+1}$ is updated.
-Applying this technique to implicit discretization is done by changing the definition the fission source ($S_f^n=(\nu\Sigma_f)^n\varphi^n$) and eliminating all $p$ terms.
+Applying this technique to implicit discretization is done by changing the definition the fission source ($S_f^n=(\nu\Sigma_f)^n\phi^n$) and eliminating all $p$ terms.
 
@@
 
-The IQS implementation mostly require the specific IQS executioner, described above.
+The IQS implementation mostly requires the specific IQS executioner, described above.
 However, IQS additional changes are in the Rattlesnake action system in order to support the IQS execution.
@@
 
-It is possible to implement IQS to any action system capable of transient simulation, which includes:
+It is possible to implement IQS to any action system capable of transient simulation, which include:
 \begin{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\tcr{JFNK is really for nonlinear systems. I would avoid confusing the reader here. I would start with PJFNK and Picard are
-the nonlinear solvers in MOOSE. Each of the nonlinear solvers requires linear solves ... and then you talk about linear solves ...}
 
@@
 For most neutronics and reactor applications, direct inversion (e.g., LU solve) of the system is overbearingly expensive, both computationally and by memory usage.
-Therefore, the most common technique (default in MOOSE) is PJFNK.
+Therefore, the most common technique (default in MOOSE) is GMRES linear iteration.
+For nonlinear systems, MOOSE employs Picard or PJFNK nonlinear iterations, PJFNK is the default technique in MOOSE and will be the basis of the following discussion.
 This technique is briefly described in \sct{sect:pjfnk}.

### 2017-03-07 matlab/kinetics_driver_conv.m
Adding PJFNK for brute force

@@ -2,7 +2,7 @@ function kinetics_driver_conv
 
 clc;
 % clear all;
-% close all;
+close all;
 clear global;
 % turn off warning due to interp1
 warning('OFF','MATLAB:interp1:ppGriddedInterpolant')
@@ -56,7 +56,7 @@ t_end = 2.0;
 if pbID==12
     amplitude_norm_ref = 0;
 else
-    amplitude_norm_ref = reference_solution( t_end, u0);
+%     amplitude_norm_ref = reference_solution( t_end, u0);
 end
 
 nn=1;
@@ -100,10 +100,11 @@ i=0;
 % not to be used for conv. studies % i=i+1; list_runs{i}= 'brute_force_matlab';
 % i=i+1; list_runs{i}= 'brute_force';
 % i=i+1; list_runs{i}= 'brute_force_elim_prec';
+% i=i+1; list_runs{i}= 'brute_force_jfnk';
 % i=i+1; list_runs{i}= 'brute_force_an_prec';
-% i=i+1; list_runs{i}= 'iqs_an_prec';
+i=i+1; list_runs{i}= 'iqs_an_prec';
 % i=i+1; list_runs{i}= 'iqs_elim_prec';
-i=i+1; list_runs{i}= 'iqs_jfnk';
+% i=i+1; list_runs{i}= 'iqs_jfnk';
 % i=i+1; list_runs{i}= 'iqsPC_an_prec';
 % i=i+1; list_runs{i}= 'iqsPC_elim_prec';
 % i=i+1; list_runs{i}= 'iqs_theta_prec';
@@ -156,6 +157,15 @@ for iconv=1:length(ntimes)
     if should_I_run_this(list_runs,'brute_force_elim_prec')
         display('brute_force_elim_prec')
         FUNHANDLE = @solve_TD_diffusion_elim_prec;
+        [brute_force_elim_prec.ampl(iconv)]=time_marching_BF( dt(iconv), ntimes(iconv), u0, FUNHANDLE);
+    end
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    %%% brute force discretization of
+    %%%   the TD neutron diffusion eq with JFNK
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    if should_I_run_this(list_runs,'brute_force_JFNK')
+        display('brute_force_JFNK')
[…]

### 2017-03-07 matlab/plot_power_level.m

@@ -16,6 +16,9 @@ switch method_name
     case 'solve_TD_diffusion_elim_prec'
         curr_leg = 'space-time-elim';
     
+    case 'solve_TD_diffusion_JFNK'
+        curr_leg = 'space-time-JFNK';
+    
     case 'solve_IQS_diffusion_an_prec'
         curr_leg = 'IQS-an';
     

### 2017-03-07 matlab/solve_TD_diffusion_JFNK.m

@@ -1,0 +1,74 @@
+function [ u ] = solve_TD_diffusion_JFNK(u,dt,tn)
+
+global io dat npar
+
+lambda = dat.lambda;
+
+% split initial solution vector into flux and precursors:
+Phi_old = u(1:npar.n,:);
+Phi_new = u(1:npar.n,end);
+C_old   = u(npar.n+1:end,:);
+
+order = length(tn) - 1;
+time_end = tn(end);
+time_beg = time_end-dt;
+
+% assemble matrices
+D    = assemble_stiffness(dat.cdiff   ,time_end);
+A    = assemble_mass(     dat.siga    ,time_end);
+NFIp = assemble_mass(     dat.nusigf_p,time_end) / npar.keff;
+IV   = assemble_mass(     dat.inv_vel ,time_end);
+NFId_new = assemble_mass(dat.nusigf_d ,time_end) / npar.keff;
+NFId_old = assemble_mass(dat.nusigf_d ,time_beg) / npar.keff;
+
+TR=NFIp -(D+A);
+
+% Ci = C_old; 
+% deno = (1+lambda*dt);
+% Ci = Ci/deno;
+% 
+% TR = TR + lambda*dt/deno*NFId_new;
+% zi = lambda*Ci;
+
+zi =  lambda*(C_old*exp(-lambda*dt));
+
+for j=1:(order+1)
+    Aj = @(t) LagrangeInterp_section(t,tn,j).*exp(-lambda*(time_end-t));
+    a(j,1) = integral(@(t)Aj(t).*(time_end-t)/dt,time_beg,time_end);
+    a(j,2) = integral(@(t)Aj(t).*(t-time_beg)/dt,time_beg,time_end);
+end
+
+TR = TR + lambda * ( a(end,1)*NFId_old + a(end,2)*NFId_new ) ;
+
+for n=1:order
+    zi = zi + lambda *( a(n,1)*NFId_old*Phi_old(:,n) + a(n,2)*NFId_new*Phi_old(:,n) );
+end
+
+A = IV - dt*TR;
+rhs = IV*Phi_old(:,end) + dt*zi;
+
+[A,rhs]=apply_BC(A,rhs,npar.add_ones_on_diagonal);
+
+Re_fun = @(phi) A*phi - rhs;
+
+eps = 1e-6;
+Jac_fun = @(phi,y) (Re_fun(phi+eps*y) - Re_fun(phi)) / eps; 
+
+% Compute residual
[…]

### 2017-03-07 doc/Thesis/data/appendix2.tex
Thesis: Adding SDIRK convergence appendix

@@
 
-\chapter{\uppercase {A Second Appendix Whose Title Is Much Longer Than The First}}
+\chapter{\uppercase {Other Testing}}
 
-Text for the Appendix follows.
+\section{SDIRK Convergence Problems}
 
-\section{Appendix Section}
+In order to figure what is causing convergence problems for SDIRK, a manufactured solution was created in the one-dimensional prototype code, described by \eqt{eq:mms}.
+A second order finite element basis function was used, so the spatial profile was known exactly with only one degree of freedom.
+The ODE describing this degree of freedom is shown by \eqt{eq:mms_ode}.
+The coefficients represented in this equation are defined by \eqt{eq:mms_coef}.
 
-\section{Second Appendix Section}
+\be 
+\phi(x,t) = x(1-x)(1+t)^4 \qq 0 \leq x \leq 1
+\label{eq:mms}
+\ee
+\be 
+\frac{1}{v}\frac{\partial\phi}{\partial t} = \alpha\phi + \frac{\beta}{v}4(1+t)^3 + \gamma(1+t)^4
+\label{eq:mms_ode}
+\ee
+
+\begin{subequations}
+\be 
+\alpha = -D\int_0^1 (4(1-2x))^2dx + (\nu\Sigma_f-\Sigma_a)\int_0^1 (4x(1-x))^2dx = -5.28
+\ee
+\be 
+\beta = \frac{1}{4}\int_0^1(4x(1-x))^2dx = 0.1333
+\ee
+\be 
+\gamma = (\Sigma_a-\nu\Sigma_f)\beta+2D\int_0^1 4x(1-x)dx = 1.32
+\ee
+\label{eq:mms_coef}
+\end{subequations}
+
+The ODE was then evaluated using SDIRK and BDF schemes with varying velocities.
+\fig{fig:mms1} compares SDIRK and BDF convergence with a relatively large $v$.
+This plot shows the SDIRK is unable to establish a proper error convergence for such a stiff problem.
+\fig{fig:mms2} compares SDIRK convergence lines with different velocities.
+This plot shows that only at relatively small values of $v$, is SDIRK able to show proper convergence.
+
+\begin{figure}[htpb!]
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/sdirk_mms1.png}
+\caption{SDIRK33 vs. BDF convergence with $v=1e5$}
+\label{fig:mms1}
+\end{figure}
+
+\begin{figure}[htpb!]
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/sdirk_mms2.png}
+\caption{SDIRK33 convergence with different velocities}
+\label{fig:mms2}
+\end{figure}
 

### 2017-03-07 matlab/sdirk_test.m

@@ -1,11 +1,13 @@
-clear all; close all; clc;
+clear all; 
+% close all; 
+clc;
 
 bf=@(x)4*x.*(1-x);
 a=integral(@(x)bf(x).*bf(x),0,1);
 b=integral(@(x)bf(x),0,1);
 g=integral(@(x)(4*(1-2*x)).^2,0,1);
 D=1;
-v=1e4;
+v=1e-1;
 nsf=1.1;sa=1;
 p=4;
 
@@ -15,14 +17,14 @@ S =@(t) a/4/v*p*(1+t).^(p-1) + (1+t).^p*((sa-nsf)*a/4+2*D*b);
 ex=@(t)(1+t).^p/4;
 
 yinit=0.25;
-tend=55;
+tend=10;
 
-do_bdf2=true;
-do_bdf3=true;
+do_bdf2=false;
+do_bdf3=false;
 
-ntimes=[10 20 40 80 500 1000 5000 1e4]; % 5e4 1e5];% 5e5];% 5e5];
+ntimes=[10 20 40 80 160 320 640];% 5000 1e4]; % 5e4 1e5];% 5e5];% 5e5];
 
-time_integration='sdirk4a';
+time_integration='sdirk3';
 switch time_integration
     case 'sdirk3'
         rk.nstages=3;
@@ -106,7 +108,8 @@ for iconv=1:length(ntimes)
     err(iconv)=abs(yn-ex(tend));
 end
 
-plot( log10(tend./ntimes), log10(err), '+-' )
+% plot( log10(tend./ntimes), log10(err), '+-' )
+loglog( (tend./ntimes), (err), '+-' )
 % slope_err = polyfit(log10(tend./ntimes), log10(err),1)
 % reference line y=Cx^s log(y) = log(C) + s log(x)
 % we pick one value of (x,y) to get C
@@ -115,9 +118,11 @@ nn=1; %
 nn=length(err);
 s=rk.order; C=0.5*err(nn)/(tend/ntimes(nn))^s;
 y=C*(tend./ntimes).^s;
-hold all;plot(log10(tend./ntimes), log10(y), 'r-')
+% hold all;plot(log10(tend./ntimes), log10(y), 'r-')
+hold all;loglog((tend./ntimes), (y), 'r-')
 leg=char(time_integration); 
-leg=char(leg,sprintf('slope %d',s)); 
+% leg=char(leg,sprintf('slope %d',s)); 
[…]

### 2017-03-09 doc/ANS_winter_2016/example.tex

\documentclass{anstrans}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{An Improved ANS Transaction Template}
\author{Seth R.~Johnson,$^{*}$ Abraham Lincoln$^{\dagger}$}

\institute{
$^{*}$Radiation Transport Group, Oak Ridge National Laboratory, P.O.\ Box 2008,
Oak Ridge, TN
\and
$^{\dagger}$State Capitol Building, Springfield, IL
}

\email{sethrj@umich.edu \and honestabe@example.com}

% Optional disclaimer: remove this command to hide
\disclaimer{Notice: this manuscript is a work of fiction. Any resemblance to
actual articles, living or dead, is purely coincidental.}

%%%% packages and definitions (optional)
\usepackage{graphicx} % allows inclusion of graphics
\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{microtype} % improves typography for PDF

\newcommand{\SN}{S$_N$}
\renewcommand{\vec}[1]{\bm{#1}} %vector is bold italic
\newcommand{\vd}{\bm{\cdot}} % slightly bold vector dot
\newcommand{\grad}{\vec{\nabla}} % gradient
\newcommand{\ud}{\mathop{}\!\mathrm{d}} % upright derivative symbol

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Microsoft Word can be a finicky beastie. \LaTeX\ abstracts content from
formatting, allowing the user to let a style file such as this take care of
uppercasing the section headings, spacing the paragraphs, and shuffle around
the figures.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu
fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

The \SN\ equations were developed by Carlson \cite{Car1953}. Another
paper is cited here \cite{Lar2008}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur faucibus
erat sed nisi aliquet molestie. Etiam malesuada, sapien at lobortis lacinia,
justo ante volutpat nunc, gravida commodo justo purus ut quam. Proin tincidunt
sem quis dui condimentum rhoncus. Nulla ut libero est, ut sollicitudin ligula.
Curabitur quam orci, aliquet dignissim feugiat eu, porta ac leo. Aenean in
ipsum arcu. Duis tempus porttitor turpis, eu volutpat odio fringilla sit amet.
Donec malesuada, arcu id porttitor mattis, arcu est molestie arcu, quis
dignissim tellus justo nec sapien. Praesent pretium interdum odio ac varius.
Suspendisse dui mauris, posuere in varius a, semper id dui. Suspendisse
placerat, quam quis luctus aliquam, metus justo hendrerit massa, vel tempus
ligula sem et purus.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur faucibus
erat sed nisi aliquet molestie.

Equations look exceedingly pretty. Here is a 3-D, monoenergetic, steady-state
transport equation with isotropic scattering and an isotropic extraneous source:
\begin{subequations} \label{eqs:fullTransport}
\begin{multline} \label{eq:fullTransportVol}
  \vec{\Omega}\vd \grad \psi(\vec{x}, \vec{\Omega})
  + \sigma(\vec{x}) \psi (\vec{x}, \vec{\Omega})
\\ =
  \frac{\sigma_s(\vec{x})}{4\pi} \int_{4\pi} \psi(\vec{x},\vec{\Omega}')
  \ud\Omega' + \frac{q(\vec{x})}{4\pi}
  \equiv \frac{1}{4\pi} Q(\vec{x}) \,,
\end{multline}
inside $\vec{x} \in V$, $\vec{\Omega} \in 4\pi$, with an incident boundary
condition
\begin{equation} \label{eq:fullTransportBndy}
  \psi(\vec{x}, \vec{\Omega}) = \psi^b(\vec{x}, \vec{\Omega}) \,,
 \quad \vec{x} \in \partial V, \ \vec{\Omega} \vd \vec{n} < 0\,.
\end{equation}
\end{subequations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Analysis}
The results were interesting, so interesting in fact that we have decided to
present them here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Subsection Goes Here}
The user must manually capitalize initial letters of a subsection heading.

For those who like equations in their papers, \LaTeX\ is a good choice. Here is
an equation for the Marshak diffusion boundary condition:
\begin{equation} \label{eq:marshak}
  4 J^- = \phi + 2 D \vec{n} \vd \grad \phi \,.
\end{equation}
If we so choose, we can effortlessly reference the equation later.

Another paragraph starts with Eq.~\eqref{eq:marshak} and sets $J^-$ to zero, a
vacuum boundary condition:
\begin{equation*}
  0 = \phi + \frac{2}{3} \frac{1}{\sigma} \vec{n} \vd \grad \phi \,.
\end{equation*}
The extrapolation distance is $2/3$. A more detailed asymptotic analysis yields
an extrapolation distance of about $0.71045$.

Figure~\ref{fig:voltage} shows how a plot might conceivably look in your
document. Always place figures after they are referenced so as not to throw
off the reader. You can use symbols and different line styles to help
differentiate your results, especially if they are printed in black and white.
Note how Fig.~\ref{fig:voltage} uses dashed lines \verb|--| for the exact
solution, solid lines \verb|-| for the new method's solutions, and dotted lines
\verb|:| for existing inaccurate methods.
\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics{example_figure}
  \caption{Captions are flush with the left.}
  \label{fig:voltage}
\end{figure}

Later on, we can include a table, even one that spans two columns such as
Table~\ref{tab:widetable}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[htb]
  \centering
\begin{tabular}{llllllllll}\toprule
      & $\phi_T(0)$      & $\phi_T(10)$      & $\phi_T(20)$      &
      $\phi_D(0)$      & $\phi_D(10)$      & $\phi_D(20)$      & $\rho$      &
      $\varepsilon$      & $N_\text{it}$
\\ \midrule
$c=0.999$  & 0.9038 & 20.63 & 31.24 & 0.9087 & 20.63 & 31.23 & 0.2192 & $10^{-7}$ & 15
\\
$c=0.990$  & 0.3675 & 13.04 & 24.7 & 0.3696 & 13.04 & 24.69 & 0.2184 & $10^{-7}$ & 15
\\
$c=0.900$  & 0.009909 & 4.776 & 17.64 & 0.009984 & 4.786 & 17.63 & 0.2118 & $10^{-7}$ & 14
\\
$c=0.500$  & $6.069\times 10^{-5}$ & 2.212 & 15.53 & 6.213$\times 10^{-5}$ & 2.239 & 15.53 & 0.2068 & $10^{-7}$ & 13
\\
\bottomrule
\end{tabular}
  \caption{This is an example of a really wide table which might not normally
  fit in the document.}
  \label{tab:widetable}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Notice how the table reference uses a Roman numeral
for its numbering scheme, whereas the figure reference uses an Arabic numeral.
For one-column tables, use the \verb|table| environment; two-column tables use
\verb|table*|. The same applies to figures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Another Subsection}
Excessive sectioning in a three-page document is discouraged, but here are more
subsections to demonstrate compliance with the ANS formatting guidelines.

\subsubsection{Third-level Heading}
This subsubsection shows compliance with the ANS-specified standard. This level
of heading should be used rarely.

\subsubsection{Another Such Heading}
And, if you really think you need a third-level heading, you should make sure
that your subsection needs at least two of them.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

The included ANS style file and this clear example file are a panacea for
the hours of headache that invariably results from formatting a document in
Microsoft Word.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Appendix}

Numbering in the appendix is different:
\begin{equation} \label{eq:appendix}
  2 + 2 = 5\,.
\end{equation}
and another equation:
\begin{equation} \label{eq:appendix2}
  a + b = c\,.
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
This material is based upon work supported a Department of Energy Nuclear
Energy University Programs Graduate Fellowship.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{ans}
\bibliography{references_IQS}
\end{document}

### 2017-03-09 doc/NEAMS_Report/NEAMS_Report.tex

@@
 % basicstyle=\scriptsize\ttfamily,
-  basicstyle=\small\ttfamily,
+  basicstyle=\footnotesize\ttfamily,
   commentstyle=\color{inl@green},
@@
 \rattlesnake is a MOOSE-based application developed INL specific to solving radiation transport problems with multiphysics capabilities.
-MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like \rattlesnake.
+MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like Rattlesnake.
 At the heart of \rattlesnake is the action system, which provides a means to consolidate the MOOSE input syntax (which can be quite larger for multigroup transport simulations), so that a user does not have to define every kernel, variable, etc., used in the problem.
@@
 So another input file and multi-app transfer was made for the adjoint calculation.
-\\
+\vspace{3mm}
 
@@
 \fig{fig:TWIGL_power} shows the IQS  solution as compared with the Brute Force solution.
-These plots show that IQS is consistent in more complex, higher dimensional problems in \rattlesnake.
+These plots show that IQS is consistent in more complex, higher dimensional problems in Rattlesnake.
 Finally, \fig{fig:TWIGL_conv} plots the error convergence of IQS and the Brute Force methods.
@@
 The \fig{fig:LRA_plots} plots show the IQS  solution as compared with the Brute Force solution.
-These plots show that IQS is consistent with Doppler feedback problems in \rattlesnake.
+These plots show that IQS is consistent with Doppler feedback problems in Rattlesnake.
 \fig{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods.
@@
 
-The purpose of this report is to briefly review the improved quasi-static method for neutronic transient simulations, to describe its implementation in \rattlesnake, the radiation transport application based on MOOSE, and to show its effectiveness using several test cases.
+The purpose of this report is to briefly review the improved quasi-static method for neutronic transient simulations, to describe its implementation in Rattlesnake, the radiation transport application based on MOOSE, and to show its effectiveness using several test cases.
 %The derivation of IQS was done with multigroup diffusion, but it was seen in \sct{sect:transport} that application to transport is relatively simple.  
-Two IQS implementations are available in \rattlesnake.
+Two IQS implementations are available in Rattlesnake.
 The first one is the traditional IQS approach, where the shape equation and PRKE are solved as a coupled nonlinear system with Picard iterations.
 The second approach, IQS P-C, is a predictor-corrector linearization, where the flux is solved and then rescaled using the amplitude obtained through the PRKE
-(Point Reactor Kinetic Equations).  IQS P-C is a simpler method and is computationally much less expensive to perform. \\
+(Point Reactor Kinetic Equations). \vspace{2mm}
 
@@
 IQS's implementation the action system made sure that all the correct kernels and postprocessors are executed when the IQS executioner is specified in the input.
-\\
+\vspace{2mm}
 
@@
 Again, IQS shows good performance by requiring half the time steps as brute force for a given error.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Future Work}
+
+Now that we are confident in the transient modeling capabilities with IQS, our next step in this research will be to compare results with a direct Rattlesnake solution for TREAT.
+INL has performed a number of transient tests for various TREAT configurations with thermal feedback.
+Use of the MAMMOTH reactor physics app will allow coupling of Rattlesnake (including the new IQS routines) with BISON.
+In this application, we will need to investigate the temporal coupling of BISON with the IQS time stepping process.
+It may or may not be necessary to couple each micro time step to BISON updates, but further research in this area is needed.
+Another interest to improve IQS's capability is to develop a smarter time adaptation scheme.
+The adaptation would entail predicting the perturbation of the shape based on transient changes in material properties.
+The ability to update PRKE parameters within macro steps may also be a valuable improvement.
 
@@
 Equation (\ref{eq:preq2}) is very similar to the normal precursor equation (\ref{eq:precursor}), except the fission source term is multiplied by $p$.
-These differences are crucial for IQS implementation in \rattlesnake.
+These differences are crucial for IQS implementation in Rattlesnake.
 %

### 2017-03-09 dr_sdirk4.f
Revert "M&c 2017"

@@ -1,103 +1,0 @@
-C * * * * * * * * * * * * * * * * * * * * * * * * *
-C --- DRIVER FOR SDIRK4 AT VAN DER POL'S EQUATION
-C * * * * * * * * * * * * * * * * * * * * * * * * *
-clink dr_sdirk4 sdirk4 decsol
-        IMPLICIT REAL*8 (A-H,O-Z)
-C --- PARAMETERS FOR SDIRK4 (FULL JACOBIAN)
-        PARAMETER (ND=2,LWORK=2*ND*ND+12*ND+7,LIWORK=2*ND+7)
-        PARAMETER (LRCONT=5*ND+2)
-        COMMON /CONT/ICONT(4),RCONT(LRCONT)
-C --- DECLARATIONS
-        DIMENSION Y(ND),WORK(LWORK),IWORK(LIWORK)
-        COMMON/STAT/NFCN,NJAC,NSTEP,NACCPT,NREJCT,NDEC,NSOL
-        EXTERNAL FVPOL,JVPOL,SOLOUT
-C --- DIMENSION OF THE SYSTEM
-        N=1
-C --- COMPUTE THE JACOBIAN ANALYTICALLY
-        IJAC=1
-C --- JACOBIAN IS A FULL MATRIX
-        MLJAC=N
-C --- DIFFERENTIAL EQUATION IS IN EXPLICIT FORM
-        IMAS=0
-C --- OUTPUT ROUTINE IS USED DURING INTEGRATION
-        IOUT=0
-C --- INITIAL VALUES
-        X=0.0D0
-        Y(1)=0.0D0
-C --- ENDPOINT OF INTEGRATION
-        XEND=2.0D0
-C --- REQUIRED TOLERANCE
-        RTOL=1.0D-5
-        ATOL=1.0D0*RTOL
-        ITOL=0
-C --- INITIAL STEP SIZE
-        H=1.0D-0
-C --- SET DEFAULT VALUES 
-        DO 10 I=1,7
-          IWORK(I)=0
-  10    WORK(I)=0.D0
-C      IWORK(4)=1 uses gamma=0.25
-       IWORK(4)=1 
-C --- CALL OF THE SUBROUTINE SDIRK4
-        CALL SDIRK4(N,FVPOL,X,Y,XEND,H,
-     &                  RTOL,ATOL,ITOL,
-     &                  JVPOL,IJAC,MLJAC,MUJAC,
-     &                  FVPOL,IMAS,MLMAS,MUMAS,
[…]

### 2017-03-09 matlab/SDIRK_testing/SDIRK.m

@@ -1,19 +1,0 @@
-function [out] = SDIRK(yn,tn,A,a,c,dt)
-
-stages = length(c);
-var = length(yn);
-k=zeros(var,stages);
-Y=zeros(var,stages);
-t = tn+c*dt;
-
-for i=1:stages
-    M = eye(var)-a(i,i)*dt*A(t(i));
-    rhs = yn;
-    for j=1:i-1,
-        rhs = rhs + dt*a(i,j)*k(:,j);
-    end
-    Y(:,i) = M\rhs;
-    X = Y(:,i);
-    k(:,i) = A(t(i))*Y(:,i);
-end
-out=Y(:,end);

### 2017-03-09 matlab/SDIRK_testing/SDIRK_an_lag.m

@@ -1,42 +1,0 @@
-function [out] = SDIRK_an_lag(yn,tn,A,a_rk,c_rk,dt)
-
-[var,order] = size(yn);
-if length(tn) ~= order
-    error('Number of solutions must match number of times')
-end
-a = zeros(order+1,1);
-stages = length(c_rk);
-k=zeros(1,stages);
-Y=zeros(var,stages);
-tt = tn(end)+c_rk*dt;
-t_all = [tn 0];
-
-for i=1:stages
-    t_all(end) = tt(i);
-    Ai  = A(tt(i));
-    dti = c_rk(i)*dt;
-    
-    for j=1:(order+1)
-        a(j) = integral(@(t) LagrangeInterp_section(t,t_all,j).*exp(Ai(2,2)*(tt(i)-t)), tn(end),tt(i));
-    end
-    
-    Ki = Ai(1,1) + Ai(1,2)*Ai(2,1)*a(end);
-    
-    zi = Ai(1,2)*yn(2,end)*exp(Ai(2,2)*(tt(i)-tn(end)));
-    for n=1:order
-        zi = zi + Ai(1,2)*Ai(2,1)*yn(1,n)*a(n);
-    end
-    
-    M = 1-a_rk(i,i)*dt*Ki;
-    rhs = yn(1,end) + a_rk(i,i)*dt*zi;
-    for j=1:i-1,
-        rhs = rhs + dt*a_rk(i,j)*k(:,j);
-    end
-    Y(1,i) = M\rhs;
-    k(i) = Ki*Y(1,i) + zi;
-end
-Y(2,end) = yn(2,end)*exp(Ai(2,2)*(tt(i)-tn(end))) + Ai(2,1)*a(end)*Y(1,end);
-for n=1:order
-    Y(2,end) = Y(2,end) + Ai(2,1)*yn(1,n)*a(n);
-end
-out=Y(:,end);

### 2017-03-09 matlab/assemble_transient_operator_iqs.m

@@ -1,44 +1,0 @@
-function M = assemble_transient_operator_iqs(curr_time,p,dpdt)
-
-global dat npar
-
-n   = npar.n;
-nnz = npar.nnz;
-M   = sparse(2*n,2*n,2*(nnz+n));
-
-D    = assemble_stiffness(dat.cdiff   ,curr_time);
-A    = assemble_mass(     dat.siga    ,curr_time);
-NFIp = assemble_mass(     dat.nusigf_p,curr_time) / npar.keff;
-NFId = assemble_mass(     dat.nusigf_d,curr_time) / npar.keff * p;
-Aiqs = assemble_mass(     dat.inv_vel ,curr_time) * dpdt/p; 
-
-% flux-flux matrix
-tmp=NFIp-(D+A+Aiqs);
-% prec-flux matrix
-if ~npar.set_bc_last
-    NFId=apply_BC_mat_only(NFId,npar.add_zero_on_diagonal);
-    tmp =apply_BC_mat_only(tmp ,npar.add_zero_on_diagonal);
-end
-
-% generic matrix for precursors (it is diagonal because the unknown
-% precursor concentrations are not the FEM expansion values but int b_i C
-L=dat.lambda*speye(n);
-
-% prec-prec matrix
-Ldiag=L;
-% flux-prec matrix
-Loffd=L/p;
-
-if ~npar.set_bc_last
-    Ldiag=apply_BC_mat_only(L,npar.add_zero_on_diagonal);
-    Loffd=apply_BC_mat_only(L,npar.add_zero_on_diagonal);
-end
-
-% finally, build the rhs operator for the transient problem
-M(1:n    ,1:n    ) = tmp;
-M(1:n    ,n+1:2*n) =  Loffd;
-M(n+1:2*n,1:n    ) = NFId;
-M(n+1:2*n,n+1:2*n) = -Ldiag;
-
-return
-end

### 2017-03-09 matlab/compute_prke_parameters.m

@@ -1,4 +1,4 @@
-function [rho_MGT,beff_MGT]=compute_prke_parameters(curr_time,shape)
+function [rho_MGT,beff_MGT,prec_MGT]=compute_prke_parameters(curr_time,shape,C)
 
 global dat npar
 
@@ -17,19 +17,12 @@ PmM = apply_BC_mat_only(NFI-D-A,npar.add_zero_on_diagonal);
 % rho  = phi_adjoint' * (NFI-D-A) * shape;
 rho  = phi_adjoint' * (PmM) * shape;
 beff = phi_adjoint' * NFId * shape;
-% MGT  = phi_adjoint' * IV * shape; % obviously, MGT = npar.K0 as well
+MGT  = phi_adjoint' * IV * shape;
 
+rho_MGT  = rho/MGT;
+beff_MGT = beff/MGT;
 
-% rho_MGT  = rho/MGT;
-% beff_MGT = beff/MGT;
-rho_MGT  = rho/npar.K0;
-beff_MGT = beff/npar.K0;
-
-% prec_MGT = phi_adjoint' * C / MGT; 
-% % obviously, this is the same as beff_MGT/dat.lambda because, at steady
-% % state, C = NFid * phi0 / lambda, thus: 
-% % phi_adjoint' * C = beff / lambda, and finally
-% % phi_adjoint' * C / MGT = beff_MGT /lambda 
+prec_MGT = phi_adjoint' * C / MGT;
 
 
 return

### 2017-03-09 matlab/funprke_lin_interp.m

@@ -1,27 +1,0 @@
-function dydt=funprke_lin_interp(time,y)
-
-% global
-global dat
-
-time_beg = dat.ode.time_beg;
-time_end = dat.ode.time_end;
-dt_macro = time_end - time_beg;
-
-w1 = (time_end-time)/dt_macro;
-w2 = (time-time_beg)/dt_macro;
-
-% reactivity coef.  interpolation
-rho_MGT  = dat.ode.rho_MGT_beg  * w1 + dat.ode.rho_MGT_end  * w2;
-beff_MGT = dat.ode.beff_MGT_beg * w1 + dat.ode.beff_MGT_end * w2;
-
-
-% compute PRKE matrix at current time
-J=[(rho_MGT-beff_MGT)   dat.lambda ; ...
-    beff_MGT           -dat.lambda ];
-
-% provide dydt to ode solver
-dydt=J*y;
-
-return
-end
-

### 2017-03-09 matlab/funprke_linhermite_interp.m

@@ -1,32 +1,0 @@
-function dydt=funprke_linhermite_interp(time,y)
-
-% global
-global dat
-
-nw = dat.ode.nw;
-
-time_beg = dat.ode.time_beg;
-time_end = dat.ode.time_end;
-dt_macro = time_end - time_beg;
-
-w1 = (time_end-time)/dt_macro;
-w2 = (time-time_beg)/dt_macro;
-
-% reactivity coef.  interpolation
-rho_MGT  = 0;
-beff_MGT = 0;
-for k=1:nw
-    rho_MGT  = rho_MGT  + ( dat.ode.rho_MGT_beg(k)  * w1 + dat.ode.rho_MGT_end(k)  * w2 ) * time^(nw-k);
-    beff_MGT = beff_MGT + ( dat.ode.beff_MGT_beg(k) * w1 + dat.ode.beff_MGT_beg(k) * w2 ) * time^(nw-k);
-end
-
-% compute PRKE matrix at current time
-J=[(rho_MGT-beff_MGT)   dat.lambda ; ...
-    beff_MGT           -dat.lambda ];
-
-% provide dydt to ode solver
-dydt=J*y;
-
-return
-end
-

### 2017-03-09 matlab/funprke_linlin_interp.m

@@ -1,28 +1,0 @@
-function dydt=funprke_linlin_interp(time,y)
-
-% global
-global dat
-
-time_beg = dat.ode.time_beg;
-time_end = dat.ode.time_end;
-dt_macro = time_end - time_beg;
-
-w1 = (time_end-time)/dt_macro;
-w2 = (time-time_beg)/dt_macro;
-
-% reactivity coef.  interpolation
-rho_MGT  = ( dat.ode.rho_MGT_beg(1)  * w1 + dat.ode.rho_MGT_end(1)  * w2 ) * w1 + ...
-          +( dat.ode.rho_MGT_beg(2)  * w1 + dat.ode.rho_MGT_end(2)  * w2 ) * w2;
-beff_MGT = ( dat.ode.beff_MGT_beg(1) * w1 + dat.ode.beff_MGT_beg(1) * w2 ) * w1 + ...
-          +( dat.ode.beff_MGT_beg(2) * w1 + dat.ode.beff_MGT_beg(2) * w2 ) * w2;
-
-% compute PRKE matrix at current time
-J=[(rho_MGT-beff_MGT)   dat.lambda ; ...
-    beff_MGT           -dat.lambda ];
-
-% provide dydt to ode solver
-dydt=J*y;
-
-return
-end
-

### 2017-03-09 matlab/integ.m

@@ -1,17 +1,0 @@
-syms t1 t2 L t
-
-int(exp(L*t), t1, t2)
-
-
-A=int((t2-t)/(t2-t1)*exp(L*t), t1, t2)
-B=expand(exp(-L*t2)*A);
-simplify(B, 'IgnoreAnalyticConstraints', true)
-% (exp(-L*t2)*(exp(L*t1) - exp(L*t2) - L*t1*exp(L*t1) + L*t2*exp(L*t1)))/(L^2*(t1 - t2))
-
-C=int((t-t1)/(t2-t1)*exp(L*t), t1, t2)
-D=expand(exp(-L*t2)*C);
-simplify(D, 'IgnoreAnalyticConstraints', true)
-% -(exp(L*(t1 - t2)) - L*t1 + L*t2 - 1)/(L^2*(t1 - t2))
-
-simplify(B+D)
-% -(exp(L*(t1 - t2)) - 1)/L

### 2017-03-09 matlab/kinetics_driver_conv.m

@@ -1,386 +1,0 @@
-function kinetics_driver_conv
-
-clc;
-% clear all;
-close all;
-% turn off warning due to interp1
-warning('OFF','MATLAB:interp1:ppGriddedInterpolant')
-
-global npar io
-
-% verbose/output parameters
-io.console_print         = false;
-io.plot_transient_figure = false;
-io.plot_power_figure     = false;
-io.make_movie            = false;
-io.save_flux             = false;
-io.print_progress        = true;
-io.figID = 99;
-% one of the two choices for applying BC
-npar.set_bc_last=true;
-
-% select problem
-pbID=11; refinements=1;
-problem_init(pbID,refinements);
-
-% compute fundamental eigenmode
-curr_time=0;
-[phi0]=steady_state_eigenproblem(curr_time);
-
-% initialize kinetic values
-C0 = kinetics_init(phi0,curr_time);
-
-% initial solution vector
-u=[phi0;C0];
-% save a copy of it
-u0=u;
-
-% time steping data
-t_end = 1.28;
-% t_end = 1.3;
-% t_end = 0.1;
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% MATLAB time discretization of
-%%%   the TD neutron diffusion eq and precursors eq
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-amplitude_norm_ref = reference_solution( t_end, u0);
-
-nn=5;
-ntimes = 2.^(0:nn-1)*4;
-dt = t_end./ntimes;
-
-% Interpolation type of shape for IQS prke parameters
-% (see solve_prke)
-%   'none' = parameters are evaluated every micro step
-%   1      = linear interpolation of parameters evauated at macro steps
-%   2      = linear interpolation of shape
-%   3      = cubic interpolation of shape
-%   4      = quadratic interpolation of shape
-npar.iqs_prke_interpolation_method=2;
-
-% Type of precursor interpolation for analytical elimination
[…]

### 2017-03-09 matlab/kinetics_init.m

@@ -16,17 +16,16 @@ C = NFId*phi/dat.lambda;
 % adjoint flux
 npar.phi_adj=phi;
 
-% store constant IV matrix
-npar.IV = assemble_mass(dat.inv_vel,curr_time);
-% npar.IV0=IV;
 % normalization constant
-K0 = npar.phi_adj' * npar.IV * phi;
+IV = assemble_mass(dat.inv_vel,curr_time);
+% npar.IV0=IV;
+K0 = npar.phi_adj' * IV * phi;
 npar.K0 = K0;
 
 % total power
 dat.Ptot = compute_power(dat.nusigf,curr_time,phi);
 
-if ~npar.set_bc_last % 0=neumann, 1=robin, 2=dirichlet
+if ~npar.set_bc_last
     % add precursors BC values
     if dat.bc.left.type==2
         dat.bc.left.C(2) = C(1);

### 2017-03-09 matlab/plot_power_level.m

@@ -1,77 +1,0 @@
-function plot_power_level( method_name, t, p, varargin )
-
-global io
-
-% get the legend name
-switch method_name
-    case 'brute_force_matlab'
-        curr_leg = 'space-time-matlab';
-
-    case 'solve_TD_diffusion'
-        curr_leg = 'space-time';
-
-    case 'solve_TD_diffusion_an_prec'
-        curr_leg = 'space-time-ANALY';
-    
-    case 'solve_TD_diffusion_elim_prec'
-        curr_leg = 'space-time-elim';
-    
-    case 'solve_IQS_diffusion_an_prec'
-        curr_leg = 'IQS-an';
-    
-    case 'solve_IQS_diffusion_elim_prec'
-        curr_leg = 'IQS-elim';
-    
-    case 'solve_IQS_PC_diffusion_an_prec'
-        curr_leg = 'IQS-PC-an';
-    
-    case 'solve_IQS_PC_diffusion_elim_prec'
-        curr_leg = 'IQS-PC-elim';
-    
-    case 'solve_IQS_diffusion_td_prec'
-        curr_leg = 'IQS-theta-prec';
-    
-    case 'solve_IQS_diffusion'
-        curr_leg = 'IQS';
-    
-    case 'solve_PRKE'
-        curr_leg = 'PRKE';
-    
-    case 'solve_PRKE_exact'
-        curr_leg = 'PRKE exact';
-    
-    case 'solve_PRKE_QS'
-        curr_leg = 'PRKE QS';
-        
-    otherwise
-        whos method_name
-        error('Method %s is UNKNOWN',method_name);
-end
-
- 
-figure(io.figID); hold all;
-
-% findobj(gcf,'Type','axes','Tag','legend')
-a=get(legend(gca),'String');
-if isempty(a)
-    leg=char(curr_leg);
-else
-    leg=char(char(a),curr_leg);
-end
-plot(t,p,'o-');
-
-% if IQS
-nVarargs = length(varargin);
-if nVarargs ==2
-    t_iqs_fine = varargin{1};
-    p_iqs_fine = varargin{2};
-    plot(t_iqs_fine,p_iqs_fine,'c.-');
-    leg = char(leg, char(strcat(curr_leg, ' fine')) );
-end
-
-legend(leg,'Location','Best');
-
-hold off;
-
-
-end

### 2017-03-09 matlab/problem_init.m

@@ -2,16 +2,15 @@ function problem_init(problem_ID,nbr_refinements_per_region)
 % load the data structure with info pertaining to the physical problem
 
 % make the problem data a global variable
-global dat npar
-
-% save pb ID for later use
-dat.PbID = problem_ID;
+global dat npar 
 
 % rod mov times
-dat.rod_mov.t_beg_1=0.1; dat.rod_mov.t_end_1=0.6;
-dat.rod_mov.t_beg_2=1.0; dat.rod_mov.t_end_2=2.7;
+dat.rod_mov.t_beg_1=0.1; 
+dat.rod_mov.t_end_1=0.6;
+dat.rod_mov.t_beg_2=1.; 
+dat.rod_mov.t_end_2=2.7;
 
-% kinetic parameters
+% kinetic parameters 
 dat.beta_tot=600e-5;
 dat.lambda=0.1;
 dat.invvel=1e-3;
@@ -20,7 +19,7 @@ dat.invvel=1e-3;
 switch problem_ID
     
     case 1
-        % one material, constant in space and time
+        % one material, constant in space and time   
         
         b=dat.beta_tot;
         iv=dat.invvel;
@@ -31,15 +30,15 @@ switch problem_ID
         dat.nusigf_d{1}= create_material_prop('constant_in_time',1.1*b    ,[],'constant_in_space',0);
         dat.inv_vel{1} = create_material_prop('constant_in_time',iv       ,[],'constant_in_space',0);
         dat.ext_src{1} = create_material_prop('constant_in_time',0        ,[],'constant_in_space',0);
-        
+
         n_regions=1;
         region_width=400;
         dat.width = region_width * n_regions;
-        
+
         imat = ones(n_regions,1);
-        
+
     case 2
-        % one material, constant in space, ramp in time 
+        % one material, constant in space      
         
         b=dat.beta_tot;
         iv=dat.invvel;
@@ -51,36 +50,36 @@ switch problem_ID
         dat.nusigf_d{1}= create_material_prop('constant_in_time',1.1*b    ,[],'constant_in_space',0);
         dat.inv_vel{1} = create_material_prop('constant_in_time',iv       ,[],'constant_in_space',0);
         dat.ext_src{1} = create_material_prop('constant_in_time',0        ,[],'constant_in_space',0);
-        
+
         n_regions=1;
         region_width=400;
         dat.width = region_width * n_regions;
[…]

### 2017-03-09 matlab/reference_solution.m

@@ -1,73 +1,0 @@
-function  [varargout] = reference_solution( t_end, u0)
-
-global dat npar io
-
-% options = odeset('RelTol',rtol,'AbsTol',atol,'InitialStep',1e-10,'OutputFcn',@odephas2,'OutputSel',[1:npar.ndofs]);
-% options = odeset('RelTol',rtol,'AbsTol',atol,'InitialStep',1e-10,'OutputFcn',@odeplot,'OutputSel',[1:npar.ndofs]);
-% options = odeset('RelTol',rtol,'AbsTol',atol);
-options = odeset('RelTol',1e-12,'AbsTol',1e-12,'InitialStep',1e-8,'Stats','on','MaxStep',1e-1,'Stats','off');
-% options = odeset('RelTol',1e-4,'AbsTol',1e-4);
-[t_ref,u_arr]=ode15s(@residual_TD_diffusion,[0 t_end],u0,options);
-u_arr=u_arr';
-
-
-% prepare for movie
-if io.make_movie
-    %# figure
-    figure, set(gcf, 'Color','white')
-    axis([0 dat.width 0 dat.max_y_val_movie]);
-    set(gca, 'nextplot','replacechildren', 'Visible','off');
-    %# preallocate
-    mov(1:ntimes) = struct('cdata',[], 'colormap',[]);
-    % movie name
-    movie_name = sprintf('PbID%d_method%s_nx%d.avi',dat.PbID,'ref_matlab',npar.nel);
-end
-
-% initial power level
-amplitude_norm=ones(length(t_ref),1);
-Ptot = dat.Ptot*ones(length(t_ref),1);
-
-%%% post-process solution %%%
-if io.plot_transient_figure
-    figure;
-end
-
-for it=1:length(t_ref)
-    % plot/movie
-    if io.plot_transient_figure
-        plot(npar.x_dofs,u_arr(1:npar.n,it));drawnow;
-        if io.make_movie && it>1, mov(it) = getframe(gca); end
-    end
-    % compute end time power for plotting
-    Ptot(it) = compute_power(dat.nusigf,t_ref(it),u_arr(1:npar.n,it));
-    % ratio of <u*,IVu> to its initial value
-    amplitude_norm(it) = (npar.phi_adj'*npar.IV*u_arr(1:npar.n,it))/npar.K0;
-end
-
-% make movie
-if io.plot_transient_figure && io.make_movie
-    close(gcf)
-    % save as AVI file
[…]

### 2017-03-09 matlab/solve_IQS_PC_diffusion_an_prec.m

@@ -1,99 +1,0 @@
-function [u_end, X,t,y] = solve_IQS_PC_diffusion_an_prec(u,X,dt_macro,time_end)
-
-global io dat npar
-
-% shortcuts
-lambda = dat.lambda;
-C_old = u(npar.n+1:end);
-
-max_iter_iqs = npar.max_iter_iqs;
-tol_iqs      = npar.tol_iqs;
-
-npar.theta_old=[];
-
-% save values at beginning of macro time step: they are needed in the IQS iteration
-X_beg=X;
-u_beg=u;
-% for clarity, I also single out the shape function at the beginning/end of the macro time step
-% IV   = assemble_mass(     dat.inv_vel ,time_end);
-z = (npar.phi_adj)'*npar.IV*u(1:npar.n)/npar.K0;
-shape_beg=u(1:npar.n)/z;
-
-for iter = 1: max_iter_iqs
-    
-    % solve time-dependent diffusion for flux
-    u_end = solve_TD_diffusion_elim_prec(u_beg,dt_macro,time_end)
-
-    % get a predicted value of the end flux
-    flux_end = u_end(1:npar.n);
-    
-    % get a shape
-    z = (npar.phi_adj)'*npar.IV*flux_end/npar.K0;
-    shape_end = u_end(1:npar.n) / z;
-    
-    % solve for amplitude function
-    if strcmpi(npar.prke_solve,'matlab')
-        [X,dpdt,t,y] =  solve_prke_ode(X_beg,dt_macro,time_end,shape_beg,shape_end);
-    else
-        [X,dpdt,t,y] =  solve_prke_iqs(X_beg,dt_macro,time_end,shape_beg,shape_end);
-    end
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % assemble IQS
-    
-    % interpolating polynomial
-    if strcmpi(npar.prke_solve,'matlab')
-        if npar.int_order==3
-            pp = interp1(t,y,'spline','pp');
-        elseif npar.int_order==2
-            pp = interp1(t,y,'pchip','pp');
-        elseif npar.int_order==1
-            pp = interp1(t,y,'linear','pp');
-        end
-    else
-        pp = interp1(t,y,'linear','pp');
-    end
-%     % extract details from piece-wise polynomial by breaking it apart
[…]

### 2017-03-09 matlab/solve_IQS_diffusion_an_prec.m

@@ -1,211 +1,0 @@
-function [u_shape, X,t,y] = solve_IQS_diffusion_an_prec(u_shape,X,dt_macro,tn)
-
-global io dat npar
-
-% shortcuts
-lambda = dat.lambda;
-dt = dt_macro;
-C_old = u_shape(npar.n+1:end,end);
-rk=npar.rk;
-
-max_iter_iqs = npar.max_iter_iqs;
-tol_iqs      = npar.tol_iqs;
-
-npar.theta_old=[];
-
-% save values at beginning of macro time step: they are needed in the IQS iteration
-X_beg=X;
-% for clarity, I also single out the shape function at the beginning/end of the macro time step
-shape_beg=u_shape(1:npar.n,:);
-shape_end=shape_beg(:,end);
-
-tn = tn-dt;
-time_beg = tn(end);
-time_end = time_beg+dt;
-
-% storage for temp SDIRK quantities
-K=zeros(length(shape_beg(:,1)),rk.s);
-
-NFId_old = assemble_mass(dat.nusigf_d,time_beg) / npar.keff;
-
-for iter = 1: max_iter_iqs
-    
-    % solve for amplitude function
-    if strcmpi(npar.prke_solve,'matlab')
-        [X,w,t,y] =  solve_prke_ode(X_beg(:,end),dt_macro,time_end,shape_beg(:,end),shape_end);
-    else
-        [X,dpdt,t,y] =  solve_prke_iqs(X_beg(:,end),dt_macro,time_end,shape_beg,shape_end,n_micro,freq_react);
-    end
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % assemble IQS
-    
-    % shortcut
-    p=X(1);
-    % interpolating polynomial
-    if strcmpi(npar.prke_solve,'matlab')
-        if npar.int_order==3
-            pp = interp1(t,y,'spline','pp');
-        elseif npar.int_order==2
-            pp = interp1(t,y,'pchip','pp');
-        elseif npar.int_order==1
-            pp = interp1(t,y,'linear','pp');
-        end        
-    else
-        pp = interp1(t,y,'linear','pp');
-    end
-    
-    % extract details from piece-wise polynomial by breaking it apart
-    [breaks,coefs,l,k,d] = unmkpp(pp);
[…]

### 2017-03-09 matlab/solve_IQS_diffusion_elim_prec.m

@@ -1,145 +1,0 @@
-function [u_shape, X,t,y] = solve_IQS_diffusion_elim_prec(u_shape,X,dt_macro,time_end)
-
-global io dat npar
-
-% shortcuts
-lambda = dat.lambda;
-dt = dt_macro;
-C_old = u_shape(npar.n+1:end);
-
-max_iter_iqs = npar.max_iter_iqs;
-tol_iqs      = npar.tol_iqs;
-
-npar.theta_old=[];
-% save values at beginning of macro time step: they are needed in the IQS iteration
-X_beg=X;
-% for clarity, I also single out the shape function at the beginning/end of the macro time step
-shape_beg=u_shape(1:npar.n);
-shape_end=shape_beg;
-
-% SDIRK solve
-% Yi = yn + h sum_j aij f(tj,Yj)
-% gives (f=Qy+z)
-%  [I-aii.h.Q(ti)]Yi = yn + h.aii zi + h sum(j=1..i-1) ( aij {Q(tj)Yj + zj} )
-rk=npar.rk;
-% beginning of the time interval
-time_beg = time_end - dt;
-% storage for temp SDIRK quantities (-lambda Ci + Bi Phii)
-f=zeros(length(shape_beg),rk.s);
-fC=f;
-NFId_old = assemble_mass(dat.nusigf_d,time_beg) / npar.keff;
-
-
-for iter = 1: max_iter_iqs
-    
-    % solve for amplitude function over the entire macro-step
-    if strcmpi(npar.prke_solve,'matlab')
-        [X,w,t,y] =  solve_prke_ode(X_beg,dt_macro,time_end,shape_beg,shape_end);
-    else
-        [X,dpdt,t,y] =  solve_prke_iqs(X_beg,dt_macro,time_end,shape_beg,shape_end,n_micro,freq_react);
-    end
-    
-    % interpolating polynomial
-    if strcmpi(npar.prke_solve,'matlab')
-        if npar.int_order==3
-            pp = interp1(t,y,'spline','pp');
-        elseif npar.int_order==2
-            pp = interp1(t,y,'pchip','pp');
-        elseif npar.int_order==1
-            pp = interp1(t,y,'linear','pp');
-        end
-    else
-        pp = interp1(t,y,'linear','pp');
-    end
-    
-    % extract details from piece-wise polynomial by breaking it apart
-    [breaks,coefs,l,k,d] = unmkpp(pp);
[…]

### 2017-03-09 matlab/solve_IQS_diffusion_td_prec.m

@@ -1,95 +1,0 @@
-function [u_shape, X, err] = solve_IQS_diffusion_td_prec(u_shape,X,dt_macro,time_end,theta)
-
-global io dat npar
-
-% shortcuts
-lambda = dat.lambda;
-dt = dt_macro;
-C_old = u_shape(npar.n+1:end);
-
-max_iter_iqs = npar.max_iter_iqs;
-tol_iqs      = npar.tol_iqs;
-
-npar.theta_old=[];
-
-% save values at beginning of macro time step: they are needed in the IQS iteration
-X_beg=X;
-u_shape_beg=u_shape;
-% for clarity, I also single out the shape function at the beginning/end of the macro time step
-shape_beg=u_shape(1:npar.n);
-shape_end=shape_beg;
-
-for iter = 1: max_iter_iqs
-    
-    % solve for amplitude function
-    %     [X,dpdt] =  solve_prke_iqs(X_beg,dt_macro,time_end,shape_beg,shape_end,n_micro,freq_react);
-    [X,dpdt,t,y] =  solve_prke_ode(X_beg,dt_macro,time_end,shape_beg,shape_end);
-    
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % assemble IQS
-    
-    % shortcut
-    p_old=X_beg(1);
-    p=X(1);
-
-    TR  = sparse(npar.n,npar.n,npar.nnz);
-    
-    D    = assemble_stiffness(dat.cdiff   ,time_end);
-    A    = assemble_mass(     dat.siga    ,time_end);
-    NFIp = assemble_mass(     dat.nusigf_p,time_end) / npar.keff;
-    IV   = assemble_mass(     dat.inv_vel ,time_end);
-    Aiqs = assemble_mass(     dat.inv_vel ,time_end) * dpdt/p;
-    
-    NFId_old = assemble_mass(dat.nusigf_d,time_end-dt) / npar.keff ;
-    NFId_new = assemble_mass(dat.nusigf_d,time_end)    / npar.keff ;
-    
-    % flux-flux matrix
-    TR=NFIp-(D+A+Aiqs);
-    
-    a1= (1-(1-theta)*dt*lambda)/(1+theta*dt*lambda);
-    a2= (1-theta)*dt/(1+theta*dt*lambda)*p_old;
-    a3= theta*dt/(1+theta*dt*lambda)*p;
-    
-    
-    % build transient matrix (divide lambda by p)
[…]

### 2017-03-09 matlab/solve_prke.m

@@ -1,14 +1,0 @@
-function X =  solve_prke(X,dt,time_end,shape)
-
-% make the problem-data a global variable
-global dat 
-
-% recompute prke parameters if XS change in time 
-[rho_MGT,beff_MGT]=compute_prke_parameters(time_end,shape);
-
-% build prke matrix
-A=[(rho_MGT-beff_MGT) dat.lambda ; ...
-    beff_MGT         -dat.lambda];
-
-% solve
-X=(eye(2)-dt*A)\X;

### 2017-03-09 matlab/solve_prke_ode.m

@@ -1,100 +1,0 @@
-function [X,w,t,pow] =  solve_prke_ode(X,dt_macro,time_end,shape_beg,shape_end)
-
-% make the problem-data a global variable
-global dat npar
-
-% create a list of times when reactivity will be recomputed for this macro
-% time step
-time_beg = time_end - dt_macro;
-
-dat.ode.time_beg = time_beg;
-dat.ode.time_end = time_end;
-
-% tolerances for odesolvers
-rtol = 3e-14; abso = 3e-14;
-atol  = abso*ones(length(X),1);
-options = odeset('RelTol',rtol,'AbsTol',atol,'InitialStep',1e-10);
-
-% ODE solve for PRKEs
-if npar.solve_prke_compute_rho_each_time
-    % save shape to retrieve them in funprke
-    dat.ode.shape_beg = shape_beg;
-    dat.ode.shape_end = shape_end;
-    % call ode solver with prke function that re-compute rho/beta at each time (expensive)
-    [t,y]=ode15s(@funprke,[time_beg time_end],X,options);
-    w = 0;
-else
-    switch npar.iqs_prke_interpolation_method
-        case 1% compute prke parameters at beg/end time and linearly interpolate
-            w = [ shape_beg shape_end]; 
-            [dat.ode.rho_MGT_beg,dat.ode.beff_MGT_beg]=compute_prke_parameters(time_beg,shape_beg);
-            [dat.ode.rho_MGT_end,dat.ode.beff_MGT_end]=compute_prke_parameters(time_end,shape_end);
-            % call ode solver with prke function that linearly interpolates
-            [t,y]=ode15s(@funprke_lin_interp,[time_beg time_end],X,options);
-        case 2 % linear variation for XS, linear for shape
-            w = [ shape_beg shape_end]; 
-            for k=1:size(w,2)
-                [dat.ode.rho_MGT_beg(k),dat.ode.beff_MGT_beg(k)]=compute_prke_parameters(time_beg,w(:,k));
-                [dat.ode.rho_MGT_end(k),dat.ode.beff_MGT_end(k)]=compute_prke_parameters(time_end,w(:,k));
-            end
-            % call ode solver with prke function that linearly interpolates
-            [t,y]=ode15s(@funprke_linlin_interp,[time_beg time_end],X,options);
-        case 3 % linear variation for XS, cubic hermite for shape
-            % hermite interpolant
-            t1=time_beg;
[…]

### 2017-03-09 matlab/steady_state_eigenproblem.m

@@ -1,11 +1,16 @@
-function [u]=steady_state_eigenproblem(curr_time)
+function [u,keff]=steady_state_eigenproblem(curr_time)
 
-global npar io
-
+global npar
 [M,P]=assemble_steady_state_operator(curr_time);
+M_=M;
+P_=P;
+save  beforebc.mat M_ P_
 
 M=apply_BC_mat_only(M,npar.add_ones_on_diagonal);
 P=apply_BC_mat_only(P,npar.add_zero_on_diagonal);
+M_=M;
+P_=P;
+save  afterbc.mat M_ P_
 
 % % kold=1;uold=rand(length(M),1);
 % % uold=uold/norm(uold,2);
@@ -38,16 +43,8 @@ if curr_time < eps
     npar.keff=keff;
 end
 
-% make sure the fundamental mode has the right sign
 if sum(u)<0
     u=-u;
-end
-
-if io.plot_transient_figure
-    plot(npar.x_dofs,u);
-end
-if io.console_print
-    fprintf('Initial SS eigenvalue = %10.8g \n--------------------------\n\n',keff);
 end
 
 return

### 2017-03-09 matlab/time_marching_BF.m

@@ -1,104 +1,0 @@
-function  [varargout] = time_marching_BF( dt, ntimes, u0, FUNHANDLE)
-
-global dat npar io
-
-% prepare for movie
-if io.make_movie
-    %# figure
-    figure, set(gcf, 'Color','white')
-    axis([0 dat.width 0 dat.max_y_val_movie]);
-    set(gca, 'nextplot','replacechildren', 'Visible','off');
-    %# preallocate
-    mov(1:ntimes) = struct('cdata',[], 'colormap',[]);
-    % movie name
-    movie_name = sprintf('PbID%d_method%s_nx%d_dt%g.avi',dat.PbID,func2str(FUNHANDLE),npar.nel,dt);
-end
-
-% initial stuff
-if strcmp(func2str(FUNHANDLE),'solve_TD_diffusion_an_prec') && strcmp(npar.an_interp_type,'lagrange')
-    order=npar.interpolation_order;
-else
-    order=1;
-end
-u=zeros(length(u0),order+1); u(:,end)=u0;
-t = 0:dt:ntimes*dt;
-amplitude_norm=ones(ntimes+1,1);
-Ptot = dat.Ptot*ones(ntimes+1,1);
-
-dat.ode.f_beg=zeros(npar.n,1);
-dat.ode.f_end=zeros(npar.n,1);
-
-
-%%% loop on time steps %%%
-for it=1:ntimes
-    
-    time_end=it*dt;
-    if io.console_print, fprintf('time end = %g \n',time_end); end
-    
-    u(:,1:end-1) = u(:,2:end);
-    
-    % solve time-dependent diffusion for flux
-    if it<order;
-        od = it;
-    else
-        od = order;
-    end
-    u(:,end) = FUNHANDLE(u(:,end-od:end-1),dt,t(it-od+2:it+1));
-    
-    % update data for hermite interpolation
-    dat.ode.f_beg=dat.ode.f_end;
-    
-    % plot/movie
-    if io.plot_transient_figure
-        figure(1);
-        plot(npar.x_dofs,u(1:npar.n,end));drawnow;
-        if io.make_movie, mov(it) = getframe(gca); end
-    end
-    % compute end time power for plotting
-    Ptot(it+1) = compute_power(dat.nusigf,time_end,u(1:npar.n,end));
-    
-    % ratio of <u*,IVu> to its initial value
-    amplitude_norm(it+1) = (npar.phi_adj'*npar.IV*u(1:npar.n,end))/npar.K0;
-    
[…]

### 2017-03-09 matlab/time_marching_PRKE.m

@@ -1,43 +1,0 @@
-function  [Pnorm_prke] = time_marching_PRKE( dt, ntimes, u, FUNHANDLE)
-
-global dat npar io
-
-shape0=u(npar.n);
-[~,beff_MGT]=compute_prke_parameters(0.,shape0);
-X=[1;beff_MGT/dat.lambda];
-Pnorm_prke=X(1)*ones(ntimes+1);
-
-% loop over time steps
-for it=1:ntimes
-    time_end=it*dt;
-    if io.console_print, fprintf('time end = %g \n',time_end); end
-    switch FUNHANDLE
-        case strcmp('solve_PRKE',FUNHANDLE)
-            shape_curr = shape0;
-            
-        case strcmp('solve_PRKE_exact',FUNHANDLE)
-            if io.save_flux
-                % shape from saved flux
-                shape_curr = io.phi_save(:,it) / (npar.phi_adj'*npar.IV*io.phi_save(:,it)) * npar.K0;
-            else
-                warning('Cannot retrieve saved flux, using initial shape instead');
-                shape_curr = shape0;
-            end
-            
-        case strcmp('solve_PRKE_QS',FUNHANDLE)
-            % compute shape for steady state
-            shape_curr = steady_state_eigenproblem(time_end);
-    end
-    % solve prke
-    X =  solve_prke(X,dt,time_end,shashape_currpe0);
-    % store power level for plotting
-    Pnorm_prke(it+1)=X(1);
-end
-
-if io.plot_power_figure
-    plot_power_level( func2str(FUNHANDLE), ...
-        linspace(0,dt*ntimes,ntimes+1), Pnorm_prke);
-end
-
-end
-

### 2017-03-09 original_files/decsol.f

@@ -1,1006 +1,0 @@
-      SUBROUTINE DEC (N, NDIM, A, IP, IER)
-C VERSION REAL DOUBLE PRECISION
-      INTEGER N,NDIM,IP,IER,NM1,K,KP1,M,I,J
-      DOUBLE PRECISION A,T
-      DIMENSION A(NDIM,N), IP(N)
-C-----------------------------------------------------------------------
-C  MATRIX TRIANGULARIZATION BY GAUSSIAN ELIMINATION.
-C  INPUT..
-C     N = ORDER OF MATRIX.
-C     NDIM = DECLARED DIMENSION OF ARRAY  A .
-C     A = MATRIX TO BE TRIANGULARIZED.
-C  OUTPUT..
-C     A(I,J), I.LE.J = UPPER TRIANGULAR FACTOR, U .
-C     A(I,J), I.GT.J = MULTIPLIERS = LOWER TRIANGULAR FACTOR, I - L.
-C     IP(K), K.LT.N = INDEX OF K-TH PIVOT ROW.
-C     IP(N) = (-1)**(NUMBER OF INTERCHANGES) OR O .
-C     IER = 0 IF MATRIX A IS NONSINGULAR, OR K IF FOUND TO BE
-C           SINGULAR AT STAGE K.
-C  USE  SOL  TO OBTAIN SOLUTION OF LINEAR SYSTEM.
-C  DETERM(A) = IP(N)*A(1,1)*A(2,2)*...*A(N,N).
-C  IF IP(N)=O, A IS SINGULAR, SOL WILL DIVIDE BY ZERO.
-C
-C  REFERENCE..
-C     C. B. MOLER, ALGORITHM 423, LINEAR EQUATION SOLVER,
-C     C.A.C.M. 15 (1972), P. 274.
-C-----------------------------------------------------------------------
-      IER = 0
-      IP(N) = 1
-      IF (N .EQ. 1) GO TO 70
-      NM1 = N - 1
-      DO 60 K = 1,NM1
-        KP1 = K + 1
-        M = K
-        DO 10 I = KP1,N
-          IF (DABS(A(I,K)) .GT. DABS(A(M,K))) M = I  
- 10     CONTINUE
-        IP(K) = M
-        T = A(M,K)
-        IF (M .EQ. K) GO TO 20
-        IP(N) = -IP(N)
-        A(M,K) = A(K,K)
-        A(K,K) = T
- 20     CONTINUE
-        IF (T .EQ. 0.D0) GO TO 80
[…]

### 2017-03-09 original_files/dr_sdirk4.f

@@ -1,107 +1,0 @@
-C * * * * * * * * * * * * * * * * * * * * * * * * *
-C --- DRIVER FOR SDIRK4 AT VAN DER POL'S EQUATION
-C * * * * * * * * * * * * * * * * * * * * * * * * *
-clink dr_sdirk4 sdirk4 decsol
-        IMPLICIT REAL*8 (A-H,O-Z)
-C --- PARAMETERS FOR SDIRK4 (FULL JACOBIAN)
-        PARAMETER (ND=2,LWORK=2*ND*ND+12*ND+7,LIWORK=2*ND+7)
-        PARAMETER (LRCONT=5*ND+2)
-        COMMON /CONT/ICONT(4),RCONT(LRCONT)
-C --- DECLARATIONS
-        DIMENSION Y(ND),WORK(LWORK),IWORK(LIWORK)
-        COMMON/STAT/NFCN,NJAC,NSTEP,NACCPT,NREJCT,NDEC,NSOL
-        EXTERNAL FVPOL,JVPOL,SOLOUT
-C --- DIMENSION OF THE SYSTEM
-        N=2
-C --- COMPUTE THE JACOBIAN ANALYTICALLY
-        IJAC=1
-C --- JACOBIAN IS A FULL MATRIX
-        MLJAC=N
-C --- DIFFERENTIAL EQUATION IS IN EXPLICIT FORM
-        IMAS=0
-C --- OUTPUT ROUTINE IS USED DURING INTEGRATION
-        IOUT=1
-C --- INITIAL VALUES
-        X=0.0D0
-        Y(1)=2.0D0
-        Y(2)=-0.66D0
-C --- ENDPOINT OF INTEGRATION
-        XEND=2.0D0
-C --- REQUIRED TOLERANCE
-        RTOL=1.0D-5
-        ATOL=1.0D0*RTOL
-        ITOL=0
-C --- INITIAL STEP SIZE
-        H=1.0D-6 
-C --- SET DEFAULT VALUES 
-        DO 10 I=1,7
-        IWORK(I)=0
-  10    WORK(I)=0.D0
-          IWORK(4)=0
-C --- CALL OF THE SUBROUTINE SDIRK4
-        CALL SDIRK4(N,FVPOL,X,Y,XEND,H,
-     &                  RTOL,ATOL,ITOL,
-     &                  JVPOL,IJAC,MLJAC,MUJAC,
[…]

### 2017-03-09 sdirk4.f

@@ -1,1175 +1,0 @@
-      SUBROUTINE SDIRK4(N,FCN,X,Y,XEND,H,
-     &                  RTOL,ATOL,ITOL,
-     &                  JAC ,IJAC,MLJAC,MUJAC,
-     &                  MAS ,IMAS,MLMAS,MUMAS,
-     &                  SOLOUT,IOUT,
-     &                  WORK,LWORK,IWORK,LIWORK,LRCONT,IDID)
-C ----------------------------------------------------------
-C     NUMERICAL SOLUTION OF A STIFF 
-C     SYSTEM OF FIRST 0RDER ORDINARY DIFFERENTIAL EQUATIONS  BY'=F(X,Y).
-C     TH METHOD USED IS A SINGLY DIAGONALLY IMPLICIT RUNGE-KUTTA METHOD
-C     OF ORDER 4 (WITH STEP SIZE CONTROL).
-C     C.F. SECTION IV.6
-C
-C     AUTHORS: E. HAIRER AND G. WANNER
-C              UNIVERSITE DE GENEVE, DEPT. DE MATHEMATIQUES
-C              CH-1211 GENEVE 24, SWITZERLAND 
-C              E-MAIL:  HAIRER@CGEUGE51.BITNET,  WANNER@CGEUGE51.BITNET
-C     
-C     THIS CODE IS PART OF THE BOOK:
-C         E. HAIRER AND G. WANNER, SOLVING ORDINARY DIFFERENTIAL
-C         EQUATIONS II. STIFF AND DIFFERENTIAL-ALGEBRAIC PROBLEMS.
-C         SPRINGER SERIES IN COMPUTATIONAL MATHEMATICS,
-C         SPRINGER-VERLAG (1990)               
-C      
-C     VERSION OF JUNE 22, 1995
-C
-C     INPUT PARAMETERS  
-C     ----------------  
-C     N           DIMENSION OF THE SYSTEM 
-C
-C     FCN         NAME (EXTERNAL) OF SUBROUTINE COMPUTING THE
-C                 VALUE OF F(X,Y):
-C                    SUBROUTINE FCN(N,X,Y,F)
-C                    REAL*8 X,Y(N),F(N)
-C                    F(1)=...   ETC.
-C
-C     X           INITIAL X-VALUE
-C
-C     Y(N)        INITIAL VALUES FOR Y
-C
[…]

### 2017-03-14 doc/Thesis/data/kinetics.tex
Thesis: Changes after thesis comments

@@
 
-\fig{fig:1D_power} shows the resulting baseline relative power profile of the one-dimensional problem.
+\fig{fig:1D_flux} shows the resulting baseline flux and relative power profile of the one-dimensional problem.
 The baseline was computed using MATLAB's ode15s function which is a embedded Runge-Kutta time adaptive method for stiff problems, the error tolerance was set very tightly ($10^{-12}$).
+The flux distribution shows that the core is relatively large, thus different regions are weakly coupled.
 This baseline computation is used to compute the error of the other time discretization methods.
@@
 \begin{figure}[!htbp]
-\begin{center}
+\centering
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/1D_flux.png}
+\caption{Flux profile at various times}
+\end{subfigure}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
 \includegraphics[width=\linewidth]{\FiguresDir/1D_power_base.png}
-\caption{Baseline total power level for the 1D slab test case}
-\label{fig:1D_power}
-\end{center}
+\caption{Power profile over transient}
+\end{subfigure}
+\caption{Baseline flux and power distribution}
+\label{fig:1D_flux}
 \end{figure}
@@
 \begin{figure}[!htbp]
-\begin{center}
-\includegraphics[width=\linewidth]{\FiguresDir/1D_shape.png}
-\caption{Flux profile of 1D slab at various times}
-\label{fig:1D_shape}
-\end{center}
+\centering
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/1D_shape2.png}
+\caption{Shape profile at various times}
+\label{1D_shape}
+\end{subfigure}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/1D_iqs_power.png}
+\caption{Power and reactivity profile over transient}
+\end{subfigure}
+\caption{IQS flux and power distribution}
+\label{fig:1D_res}
 \end{figure}
@@
 \end{center}
+\end{figure}
+
+\subsection{One-Dimensional Mini-Core Problem}
+
+This problem is exactly the same as the previous one-dimensional problem, except the the core was reduced to 80 cm in length.
+The purpose of testing this problem is to determine if a more tightly coupled core will yield better performance for IQS.
+\fig{fig:1Dmini_flux} shows the resulting baseline flux and relative power profile of the one-dimensional mini-core problem.
+This plot shows that the perturbed regions affect the domain more evidently.
+\fig{fig:1Dmini_shape} shows the shape profile at various times during the transient.
+This plot shows that the shape is much less time-dependent than the previous large core.
+\fig{fig:1Dmini_conv} shows that IQS performs significantly better with this example than the large core.
+
+\begin{figure}[!htbp]
+\centering
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/1Dmini_flux.png}
+\caption{Flux profile at various times}
+\end{subfigure}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{\FiguresDir/1Dmini_power_base.png}
+\caption{Power profile over transient}
+\end{subfigure}
+\caption{Mini-core baseline flux and power distribution}
+\label{fig:1Dmini_flux}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.7\linewidth]{\FiguresDir/1Dmini_shape.png}
+\caption{Shape profile at various times for one-dimensional mini-core}
+\label{fig:1Dmini_shape}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.7\linewidth]{\FiguresDir/1Dmini_conv_IE.png}
+\caption{Time step convergence for one-dimensional mini-core with implicit Euler discretization}
+\label{fig:1Dmini_conv}
 \end{figure}

### 2017-03-21 doc/Papers/IQS_convergence/IQS_convergence.tex
Convergence Paper: Initial draft

%----------------------------------------------------------------------------------------

\documentclass[twoside,twocolumn]{article}

\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage[text={7in,9in},centering,columnsep=18pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{enumitem} % Customized lists
\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\usepackage{titling} % Customizing the title section
\usepackage[hidelinks]{hyperref} % For hyperlinks in the PDF
\usepackage[numbers,sort&compress]{natbib}[2010/09/13]
\usepackage{bm}
\usepackage{graphicx} 
\usepackage{tikz} 
\usepackage{changepage}
\usepackage{subcaption}

\setlength{\abovecaptionskip}{0.25\baselineskip}
\setlength{\belowcaptionskip}{0.25\baselineskip}
\setlength{\textfloatsep}{0.5\baselineskip}
\setlength{\intextsep}{0.5\baselineskip}
\renewcommand{\bottomfraction}{0.9}
\renewcommand{\topfraction}{0.9}

\newcommand{\SN}{S$_N$}
\renewcommand{\vec}[1]{\bm{#1}} %vector is bold italic
\newcommand{\vd}{\bm{\cdot}} % slightly bold vector dot
\newcommand{\ud}{\mathop{}\!\mathrm{d}} % upright derivative symbol
%  new definitions
\newcommand{\bs}[1]{\mathbf{#1}}
\renewcommand{\div}{\bs{\nabla}\! \cdot \!}
\newcommand{\grad}{\bs{\nabla}}
% extra space
\newcommand{\qq}{\quad\quad}
% common reference commands
\newcommand{\eqt}[1]{Eq.~(\ref{#1})}                     % equation
\newcommand{\eqts}[2]{Eqs.~(\ref{#1})~and~(\ref{#2})}    % equation
\newcommand{\eqtss}[2]{Eqs.~(\ref{#1})-(\ref{#2})}       % equation
\newcommand{\fig}[1]{Fig.~\ref{#1}}                      % figure
\newcommand{\figs}[2]{Figs.~\ref{#1}~and~\ref{#2}}		% two figures
\newcommand{\figss}[2]{Figs.~\ref{#1}~-~\ref{#2}}			% multiple figures
\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table
\newcommand{\sct}[1]{Section~\ref{#1}}                   % section
\newcommand{\app}[1]{Appendix~\ref{#1}}                  % appendix

\newcommand{\keff}{k_\textit{eff}}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\vn}{\vec{n}}
\newcommand{\vel}{\vec{\mathrm{v}}}
\newcommand{\adj}{\Phi^\dagger_0}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert_{L^2}}

\definecolor{britishracinggreen}{rgb}{0.0, 0.26, 0.15}
\newcommand{\tcr}[1]{\textcolor{red}{#1}}
\newcommand{\tcb}[1]{\textcolor{blue}{#1}}
\newcommand{\tcm}[1]{\textcolor{magenta}{#1}}
\newcommand{\tcp}[1]{\textcolor{violet}{#1}}
\newcommand{\tcg}[1]{\textcolor{britishracinggreen}{#1}}

% tikz stuff
\usetikzlibrary{shapes,arrows,positioning}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\tikzstyle{redblock}=[rectangle, draw, align=center, top color=red!25, bottom color=red!75,  minimum width=10mm, minimum height=10mm,]
\tikzstyle{blueblock}=[rectangle,draw, align=center, top color=blue!25, bottom color=blue!75,  minimum width=10mm, minimum height=10mm]
\tikzstyle{purpleblock}=[rectangle,  draw, align=center, top color=purple!25, bottom color=purple!75, minimum width=10mm, minimum height=10mm]
\tikzstyle{orangeblock}=[rectangle, draw, align=center, top color=orange!25, bottom color=orange!75,  minimum width=10mm, minimum height=10mm]
\tikzstyle{greendiamond}=[diamond, draw, align=center, top color=green!25, bottom color=green!75,  minimum width=8mm, aspect=2]
\tikzstyle{greenblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=green!20, ultra thick, minimum width=60mm, minimum height=15mm,]
\tikzstyle{blueblock2}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=blue!20, ultra thick, minimum width=60mm, minimum height=15mm]
\tikzstyle{reddiamond}=[diamond, draw, align=center, top color=white, bottom color=red!20, ultra thick, minimum width=60mm, aspect=2]
\newcommand{\tikzback}[5]{
 \begin{pgfonlayer}{background}
  \path (#1.west |- #2.north)+(-0.5,0.75) node (a1) {};
  \path (#3.east |- #4.south)+(+0.5,-0.25) node (a2) {};
  \path[fill=yellow!10,rounded corners, draw=black!100, dashed] (a1) rectangle (a2);
   \path (#3.east |- #2.north)+(0,1.25)--(#1.west |- #2.north) node[midway] (#5-n) {};
   \path (#3.east |- #2.south)+(0,-0.35)--(#1.west |- #2.south) node[midway] (#5-s) {};
   \path (#1.west |- #2.north)+(-0.75,0)--(#1.west |- #4.south) node[midway] (#5-w) {};
 \end{pgfonlayer}}

%----------------------------------------------------------------------------------------

\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Large\bfseries} % Article title formatting
\posttitle{\end{center}} % Article title closing formatting
\title{High-Order Time Discretization and Nonlinear Iteration Analysis of the Improved Quasi-Static Method} % Article title
\author{%
Zachary M. Prince\thanks{Corresponding author, \newline\indent\indent Phone: +1-360-362-6622 \newline\indent\indent email: zachmprince@tamu.edu} \qquad Jean C. Ragusa  \\
\normalsize Texas A\& M University, College Station, TX \\
}
\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent The improved quasi-static method (IQS) is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent amplitude and a time-space-energy dependent shape. The objective of the IQS factorization is to evaluate amplitude and shape on different time scales in order to reduce computational burden associated with solving the multi-dimensional flux equations, while maintaining solution accuracy. IQS factorization leads to a nonlinear system of equations that requires iteration of shape and amplitude. IQS iteration techniques involve fixed-point (Picard) iteration with various convergence criteria and Newton iteration, namely preconditioned Jacobian-free Newton Krylov (PJFNK) method. Nonlinear convergence of each of these techniques is investigated. Validation of IQS with analysis of time step convergence is also investigated. Proper time step convergence is vital for implementation of time adaptive methods and error prediction. The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicolson, backward difference formulae (BDF), and singly-diagonally-implicit Runge-Kutta (SDIRK) methods.
\end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

IQS is a spatial kinetics method that involves factorizing the flux solution into space- and time-dependent components \cite{Ott_1966,Dulla2008,Devooght_1984,Monier_diss,Sissaoui_1995}.  These components are the flux's amplitude and its shape. Amplitude is only time-dependent, while the shape is both space- and time-dependent (as opposed to the PRKE approach where the shape was constant in time).  However, the impetus of the method is the assumption that the shape is only weakly dependent on time. Therefore, the variable for the multi-dimensional diffusion evaluation is expected to be less stiff than the flux itself, without making the assumption that spatial variance is time-independent.  As a result, the shape may not require an update at the same frequency as the amplitude function, but only on larger macro-time steps.

Due to the factorization of the flux into a shape and an amplitude, the latter two variables are coupled. Ott in \cite{Ott_1966} first investigated the coupling of shape and amplitude in a quasi-static nature, but did not include the time derivative in the shape equation. Later, in \cite{Ott_1969}, Ott incorporated the time derivative of shape in the equation, yielding better results; this also led to the technique's name: the Improved Quasi-Static method.  The resulting system of equations is  nonlinear \cite{Dulla2008}. Nonlinear problems require an iterative process to evaluate, either fixed-point (Picard) or Newton iteration. Sissaoui et al. \cite{Sissaoui_1995}, Koclas et al. \cite{Koclas_1996}, Devooght et al. \cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their IQS simulations, the main difference among them is their criteria for convergence.  Devooght et al. in \cite{Devooght_1984} also utilizes a Newton iteration technique. This paper discusses the different nonlinear iteration techniques and tests the rigor of their implementation. IQS can also be linearized by using the IQS Predictor-Corrector method (IQS P-C) \cite{dulla2006}. IQS P-C entails evaluating the flux equation then correcting its amplitude using an amplitude evaluation. This method has proven to be effective for problems requiring a significant amount of shape updates when IQS is implemented \cite{Dulla2008}.

IQS has yet to be rigorously tested with high-order discretization of the shape equation. All publications that test IQS do not go beyond first-order time discretization. Testing IQS with higher-order schemes for proper time step error convergence is essential for implementation of time adaptive methods and quantification of model uncertainty. This paper discusses implementation of up though fourth-order time discretization for the shape equation of IQS and the flux equation of IQS P-C. Step doubling time adaptation is also implemented to test IQS and IQS P-C  performance with adaptation of shape/flux evaluation.

To test IQS and IQS P-C nonlinear iteration, time step convergence, and time adaptation performance (five?) problems were applied. (Three?) problems are purely neutronics and involve varying magnitudes of reactor size and complexity. Two problems involve temperature feedback and test an intermediate time scale for temperature evaluation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this Section, we recall the equation for the IQS method, starting from the standard multi-group diffusion equations with delayed neutron precursors in operator form:

\be
\frac{1}{v^g}\frac{\partial \phi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \phi^{g'} - L^g\phi^g + S_{d}^g
\label{eq:flux}
\ee 
\be
\frac{dC_i}{dt} = \sum_{g=1}^G P_{d,i}^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:precursor}
\ee
%
where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the diffusion operator, 
$S_{d}^g$ is the delay source, and $P_{d,i}^g$ is the delayed-neutron fission operator.

The flux factorization approach leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent 
multigroup shape ($\varphi$):
\be
\phi^g(\vec{r},t)=p(t)\varphi^g(\vec{r},t)
\ee
After implementing the factorization, the shape diffusion equations result:
\be
\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \varphi^{g'} - \left(L^g - \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right)\varphi^g + \boxed{\frac{1}{p}}S_{d}^g
\label{eq:shape}
\ee 
\be
\frac{dC_i}{dt} = \boxed{p}\sum_{g=1}^G P_{d,i}^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:prec}
\ee

Note that the time-dependent shape equation is similar to the time-dependent flux equation. 
However, the shape equations are now non-linearly coupled (boxed terms) to the amplitude equations.

To obtain the amplitude equations, the multigroup shape equations are multiplied by a weighting function, typically the initial adjoint flux ($\phi^{*g}$), and then integrated over the domain.  For brevity, the inner product over space will be represented with parenthetical notation ($\left(\phi^{*g},f^g\right) = \int_D \phi^{*g}(\vec{r})f^g(\vec{r})d^3r
$). In order to impose uniqueness of the factorization, one requires $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)$ to be constant.  After some manipulation, the point reactor kinetics equations (PRKE) for the amplitude solution are obtained:
\be
\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
\ee
\be
\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}p-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
\ee
where the functional coefficients are calculated using the space-/time-dependent shape function as follows:
\be
\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\sum_{g'}(H^{g' \to g}+P_p^{g' \to g}-L^{g'}\delta_{g'g})\varphi^{g'}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:rmb}
\ee
\be
\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{\sum_{g=1}^G(\phi^{*g}, P_{d,i}^g \varphi^g)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\ee
\be
\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
\label{eq:l}
\ee

Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three dimensions, and even more so when using the transport equations in lieu of the diffusion equations.  Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself, thus allowing for larger time step sizes in updating the shape. The PRKE equations form a small ODE system and can be solved using a much smaller time step size. In transients where the shape varies much less than the flux, IQS can be very computationally effective. The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{figures/IQS_visualization.jpg}
\caption{IQS method visualization}
\label{fig:iqsviz}
\end{figure}

It is important to note that the PRKE parameters are evaluated at each macro step and interpolated for the PRKE evaluation. In order to preserve the error convergence of high order discretization schemes for shape, higher order interpolation of the parameters is required.

%---------------------------------------------------------------------------------------%
\subsection{Iteration Methods}
\label{sect:iter}
%---------------------------------------------------------------------------------------%

As we noted in the previous section, shape-PRKE equations are a nonlinear system and thus may be solved in a iterative manner.  Each macro time step can be iterated so the best shape is used to compute power at the micro time steps. Sissaoui et al. from \cite{Sissaoui_1995}, Koclas et al. from \cite{Koclas_1996}, Devooght et al. from \cite{Devooght_1984}, and Monier from \cite{Monier_diss} all use iterative techniques for their quasi-static simulations.  They all undergo a similar process:
\begin{adjustwidth*}{0cm}{4mm}
\begin{itemize}
\item[\textit{Step 1:}] Compute the PRKE parameters at the end of the macro step using the last computed shape
\item[\textit{Step 2:}] Linearly interpolate the computed PRKE parameters over the macro step
\item[\textit{Step 3:}] Solve the PRKE on micro steps over the entire macro step
\item[\textit{Step 4:}] Solve the shape equation on the macro step using the computed values of $p$ and $dp/dt$.
\item[\textit{Step 5:}] Check if the shape solution has converged:
	\begin{itemize}
	\item \textit{No:} Repeat the same macro time step
	\item \textit{Yes:} Move on to the next macro time step
	\end{itemize}
\end{itemize}
\end{adjustwidth*}

The major difference between the methods of these authors is the convergence criteria used.  Sissaoui and Koclas \cite{Sissaoui_1995, Koclas_1996} use fixed point iteration where the criteria is the simply the normalized difference between the last two computed shapes.  Monier in \cite{Monier_diss} also does fixed point iterations with the same criteria, except the solution is scaled by $\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n})\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n+1})\right)}$ after each iteration.  Devooght in \cite{Devooght_1984} does a Newton-SOR iteration where the residual of the shape function evaluation is the convergence criteria and next iteration's solution is computed using Newton-Raphson method.

These techniques are by no means an exhaustive list of the possible iteration techniques for IQS. Dulla et al. in \cite{Dulla2008} does an in depth analysis of the fixed point iteration technique most similar to Sissaoui and Koclas, involving convergence rates and solution results.  However, no comprehensive analysis of iteration techniques exists, comparing both Newton and fixed-point convergence rates.  The following describes each iteration technique investigated by this research:

\begin{enumerate}
\item $L^\infty$ norm of shape: $\text{max}\left|\varphi_{n}^{(k+1)} - \varphi_{n}^{(k)}\right|/\text{max}\left|\varphi_{n}^{(k+1)}\right| < \epsilon_{\varphi}$
\item $L^2$ norm of shape: $\norm{\varphi_{n}^{(k+1)} - \varphi_{n}^{(k)}}/\norm{\varphi_{n}^{(k+1)}} < \epsilon_{\varphi}$
\item Reactivity convergence: $\left(\rho/\Lambda\right)^{(k+1)} - \left(\rho/\Lambda\right)^{(k)} < \epsilon_{\rho}$
\item Amplitude convergence: $p_n^{k+1} - p_n^{k} < \epsilon_p$
\item Uniqueness consistency: $(K_n^{(k+1)} - K_0)/K_0 < \epsilon_K$ where,
\be
K_n=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)
\ee
\end{enumerate}

In order to preserve the uniqueness criteria, it is beneficial to scale the shape such that the $K_n$ is constant, shown in \eqt{eq:shape_scale}.  This scaling can also be done after each iteration, to insure that the uniqueness criteria is satisfied whenever the shape is evaluated.
\be 
\varphi^{g}_n = \varphi^{g,(\text{last})}_n \frac{K_0}{K^{(\text{last})}_{n}}
\label{eq:shape_scale}
\ee

%---------------------------------------------------------------------------------------%
\subsection{IQS Predictor-Corrector}
%---------------------------------------------------------------------------------------%

The Predictor-Corrector (P-C) version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the solution of the coupled system of equations is different.  In the IQS P-C version, the flux equations (not the shape equations) are solved (represented by \eqts{eq:flux}{eq:precursor}) in order to obtain a predicted flux solution. This predicted flux is then converted to a shape by rescaling it as follows:
\be
\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
\label{eq:rescale}
\ee
where the scaling factors are given by
\be
K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
\ee
\be
K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
\ee

The PRKE parameters are then computed with this shape using \eqtss{eq:rmb}{eq:l} and interpolated over the macro step, then the PRKE ODE system is solved on the micro time scale.  With the newly computed amplitude, the shape is rescaled into a flux and the final corrected flux is given by:
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1} \,.
\ee

The advantage to the predictor-corrector method is there is no iteration necessary for this method and, in turn, is much simpler and faster than the standard IQS.  Ikeda et al. in \cite{Ikeda_2001} and Goluoglu et al. in \cite{Goluoglu_2001} both use IQS P-C for complex, three-dimensional problems.  Their results prove IQS P-C to be impressively effective, despite the de-coupling of the system.  Dulla et al. in \cite{Dulla2008} also describes an in depth comparison of IQS P-C with traditional IQS.

%---------------------------------------------------------------------------------------%
\subsection{Temperature Feedback Treatment}
%---------------------------------------------------------------------------------------%

Two problems that are tested with IQS and IQS P-C involve adiabatic temperature feedback. The governing temperature equation is described by \eqt{eq:temp} with cross-section feedback described by \eqt{eq:dopp}.

\be
\rho c_p \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
\label{eq:temp}
\ee

\be
\Sigma_a^{thermal}(\vec{r},t) = \Sigma_a^{thermal}(\vec{r},0)\left[1+\gamma\left(\sqrt{T}-\sqrt{T_0}\right)\right]
\label{eq:dopp}
\ee

A typical implicit solver would simply use the interpolated flux at end of the temperature time step for the right hand side of the equation, a theta-method discretization.  However, IQS has much more information about the profile of the flux along the time step because of the micro-step amplitude evaluation.  Therefore, it is possible to solve for temperature using a semi-analytical approach, shown by \eqt{eq:temp_an}.

\be
T^{n+1} = T^n + \frac{\kappa_f}{\rho c_p} \left(a_2 \varphi^{n+1} + a_1 \varphi^{n}\right)
\label{eq:temp_an}
\ee

Where $n$ corresponds to the beginning of the temperature step.  $a_1$ and $a_2$ are integration coefficients defined by \eqt{eq:a1} and \eqt{eq:a2}.  Any interpolation of the amplitude along the micro steps is possible for the integration, this application uses piece-wise linear.

\be
a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)p(t')dt'
\label{eq:a1}
\ee
\be
a_2 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)p(t')dt'
\label{eq:a2}
\ee

Temperature feedback affects both the shape equation and the reactivity of the PRKE; thus, it is an additional nonlinear component to the already coupled shape-amplitude equations. In foresight to the application of this component, temperature is much more time dependent than the shape, but less so than the amplitude.  Therefore, the evaluation of temperature will have its own time scale.  A possible solution process for a problem with temperature feedback will have time three time scales portrayed in \fig{fig:time}.  The first time scale is the shape solve, the second is the temperature evaluation as well as the computation of PRKE parameters, and the third is the PRKE scale.  It is important to note that the number of time steps in each scale is arbitrary and meant only for visual purposes.

\begin{figure}[htpb!]
\centering
\resizebox{\linewidth}{!}{
\begin{tikzpicture}[scale=2]
%Shape
\draw[] (0,2) -- (6,2) ;
\foreach \x in  {0,6}
\draw[shift={(\x,2)},color=black] (0pt,6pt) -- (0pt,-6pt);
\draw[shift={(0,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(shape) at (-.25,2) {$\varphi:$};

%Temp/Params
\draw[] (0,1) -- (6,1) ;
\foreach \x in  {0,6}
\draw[shift={(\x,1)},color=black] (0pt,6pt) -- (0pt,-6pt);
\foreach \x in  {2,4}
\draw[shift={(\x,1)},color=black] (0pt,4pt) -- (0pt,-4pt);
\draw[shift={(0,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(temp) at (-.75,1) {$T \rightarrow \rho, \bar{\beta}, \bar{\lambda}, \Lambda$:};

% PRKE
\draw[] (0,0) -- (6,0) ;
\foreach \x in  {0,6}
\draw[shift={(\x,0)},color=black] (0pt,6pt) -- (0pt,-6pt);
\foreach \x in  {0,2,4,6}
\draw[shift={(\x,0)},color=black] (0pt,4pt) -- (0pt,-4pt);
\foreach \x in  {0,0.2,0.4,0.6,0.8,1,1.2,1.4,1.6,1.8,2,2.2,2.4,2.6,2.8,3,3.2,3.4,3.6,3.8,4,4.2,4.4,4.6,4.8,5,5.2,5.4,5.6,5.8,6}
\draw[shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt);
\draw[shift={(0,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(prke) at (-.5,0) {$p, \frac{dp}{dt}:$};

\draw (0,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 2} (2,0);
\draw (2,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 4} (4,0);
\draw (4,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 6} (6,0);
\draw (0,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 1} (2,1);
\draw (2,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 3} (4,1);
\draw (4,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 5} (6,1);
\draw (0,2) edge[out=15,in=165,->,thick] node[above,sloped] {Step 7} (6,2);

\draw (0,2) edge[out=-135,in=135,->,thick] node[below,sloped] {$\varphi$} (0,1);
\draw (2,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (2,0);
\draw (2,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (2,1);
\draw (4,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (4,0);
\draw (4,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (4,1);
\draw (6,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (6,0);
\draw (6,0) edge[out=45,in=0,->,thick] node[below,sloped] {$p, \frac{dp}{dt}$} (6,2);

\end{tikzpicture}
}
\caption{Time scales and process of IQS with temperature feedback}
\label{fig:time}
\end{figure}

Iteration processes are needed for each time scale.  The amplitude and temperature need to be iterated on the middle time scale until convergence on each temperature step.  Then another iterative process needs to occur in the shape time scale on all three variables. \fig{fig:proc} shows the programming diagram implement to execute this process. The time increment of $\Delta t/3$ for the temperature solve is arbitrary and is meant only to match \fig{fig:time} where three temperature updates have been used as illustration.

\begin{figure}[!htpb]
\centering
\resizebox{\linewidth}{!}{
\begin{tikzpicture}[every node/.style = {font=\normalsize}]

\node[greenblock](p1) at (0,0) {Save Old Solution};
\node[greenblock](p2) at (0,-2.5) {Save Current Solution};
\node[greenblock](p3) at (0,-5) {Interpolate Solution};
\node[blueblock2](p4) at (0,-7.5) {Solve Temperature};
\node[greenblock] (p5) at (0,-10) {Update PRKE Params};
\node[greenblock] (p6) at (0,-12.5) {Solve PRKE};
\node[reddiamond] (check1) at (0,-17.5) {Check for \\ convergence ($p$)};
\node[reddiamond] (check2) at (7,-12) {If $t=t_{n+1}$};
\node[greenblock] (p7) at (8,-6) {Shape Solve};
\node[reddiamond] (check3) at (8,-2.5) {Check for \\ convergence ($\varphi$)};
%\node [above =0.1mm of p1] {\large{IQS Solve:}};
%\node [above =0.1mm of p5] {\large{Multi-physics:}};

%\tikzback{p1}{p1}{p4}{p4}{bk1}
%\tikzback{p5}{p5}{p5}{p5}{bk2}

\draw[->,ultra thick](p1.south) -- node[right] {$t=t_{n+1}$}(p2.north);
\draw[->,ultra thick](p2.south) -- node[right] {$t=t_{n}+\Delta t/3$}(p3.north);
\draw[->,ultra thick](p3.south) -- (p4.north);
\draw [->,ultra thick] (p4.south)-- (p5.north);
\draw[->,ultra thick](p5.south) -- (p6.north);
\draw[->,ultra thick](p6.south) -- (check1.north);
\draw[->,ultra thick](check1.west) -- node[above,sloped] {\large{no}} (-5,-17.5) |-  (p4.west);
\draw[->,ultra thick](check1.east) -| node[right] {\large{yes}} (check2.south);
\draw[->,ultra thick](check2.west) -- node[above,sloped] {\large{no}\small \qq $t=t+\Delta t/3$} (4,-5) -- (p3.east);
\draw[->,ultra thick](check2.north) --  (8,-10) -- node[right] {\large{yes}} (p7.south);
\draw[->,ultra thick](p7.north) -- (check3.south);
\draw[->,ultra thick](check3.west) -- node[above] {no} (p2.east);
\draw[->,ultra thick](check3.north) -- node[right] {yes} (8,0) -- node[above] {$t_n=t_{n+1}$} (p1.east);

\end{tikzpicture}
}
\caption{Visualization of fixed-point iteration and temperature update process for IQS}
\label{fig:proc}
\end{figure}

%---------------------------------------------------------------------------------------%
\subsection{Delayed Neutron Precursor Treatment}
%---------------------------------------------------------------------------------------%

This section presents the time-integration method used to solve coupled flux/shape and precursor equations, represented by Equations~\eqref{eq:flux}/\eqref{eq:shape}~and~\eqref{eq:precursor}/\eqref{eq:prec}. First, we note that we could keep this system of two time-dependent equations and solve it as a coupled system. However, this is unnecessary and a memory expensive endeavor because the precursor equation is only an ODE and not a PDE. Instead, one may discretize in time the shape equation, which typically requires the knowledge of the precursor concentrations at the end of the time step. This precursor value is taken from the solution, numerical or analytical, of the precursors ODE.

A common precursor evaluation technique is using the theta method; applying this to \eqt{eq:prec}:

\be
C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta}{1+\theta\Delta t\lambda}S_f^n p^n +  \frac{\theta\Delta t \beta}{1+\theta\Delta t\lambda}S_f^{n+1} p^{n+1}
\label{eq:dnp_theta}
\ee

Where $S_f$ is the fission source equivalent for shape ($S_f^n=(\nu\Sigma_f)^n\varphi^n$). Another technique to evaluating the precursor equation is to use a exponential operator and integrate the time derivative analytically.  Applying this operation to \eqt{eq:prec} yields:

\be
C^{n+1} =  C^n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
\label{eq:prec_an}
\ee

Again, this can be applied to implicit discretization by altering the definition of the fission source and eliminating $p$.  Because $S_f$ is not known continuously over the time step, the integration can be done using any scheme (Riemann, trapezoid, Simpson's, etc.). However, there is a very accurate representation of $p(t)$ over the macro step from the PRKE solve. In order to utilize this information, another possibility is to interpolate $S_f$ linearly over the macro step.  Such that:

\be
S_f(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}S_f^n  + \frac{t-t_n}{t_{n+1}-t_n}S_f^{n+1}  \quad t_n \le t \le t_{n+1}
\ee
Applying this to \eqt{eq:prec_an} yields:
\be
C^{n+1} = C^n e^{-\lambda \Delta t} + \left(\hat{a}_2 S_f^{n+1}+\hat{a}_1 S_f^n\right)\beta
\label{eq:dnp_an}
\ee

With integration coefficients defined as:

\begin{align}
&\hat{a}_1= \int_{t_n}^{t_{n+1}}\frac{t_{n+1}-t'}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
&\hat{a}_2 = \int_{t_n}^{t_{n+1}}\frac{t'-t_n}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt'
\end{align}

%---------------------------------------------------------------------------------------%
\subsection{Step Doubling Time Adaptation}
%---------------------------------------------------------------------------------------%

Further enhancements to IQS performance can be gained by using time adaptation (or time step control) in order to increase or reduce the macro time step size for the shape evaluation, depending on error estimates. A step-doubling technique is chosen as the time adaptation technique \cite{NumC}. The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps ($\varphi^g_{\Delta t/2}$). Note: $\varphi$ is changed to $\phi$ for implicit discretization and IQS P-C.

The relative error is computed as follows:
\be
e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2} - \sum_{g=1}^G\varphi^g_{\Delta t}}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2}},\norm{\sum_{g=1}^G\varphi^g_{\Delta t}}\right)}
\label{eq:edt2}
\ee
If the error is smaller than the user-specified tolerance, $e_{tol}$, the time step is accepted. In addition, a new time step size is estimated as follows:
\be
\Delta t_{new} = S \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}}
\label{eq:dt2}
\ee

Where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor. If the error is larger than the user-specified tolerance, the time step is rejected. A new time step size is estimated using  \eqt{eq:edt2} as well. This process can be visualized by \figs{fig:dt2_1}{fig:dt2_2}.  Where a step involves a full convergence of shape, amplitude, and any multiphysics on the respective time step.

To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to implicit discretization method, traditional IQS, and IQS P-C.  Each of these methods will be applied to several diffusion problems; the number of time steps taken and the resulting error will be used to compare the methods.

\begin{figure}[htpb!]
\centering
\resizebox{\linewidth}{!}{
\begin{tikzpicture}[scale=1.5]
\draw[] (0,2) -- (6,2) ;
\foreach \x in  {0,3,6}
\draw[shift={(\x,2)},color=black] (0pt,6pt) -- (0pt,-6pt);
\draw[shift={(0,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(3,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1/2}$};
\draw[shift={(6,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\draw (0,2) edge[out=45,in=135,->,thick,blue] node[above,sloped] {\tcb{Step 1}} (6,2);
\draw (0,2) edge[out=25,in=155,->,thick,red] node[above,sloped] {\tcr{Step 2}} (3,2);
\draw (3,2) edge[out=25,in=155,->,thick,red] node[above,sloped] {\tcr{Step 3}} (6,2);
\node[anchor=west](shape) at (6.1,2.5) {$\tcb{\phi_{\Delta t}^g}$};
\node[anchor=west](shape) at (6.1,2) {$\tcr{\phi_{\Delta t/2}^g}$};

\end{tikzpicture}
}
\caption{Visualization of step doubling process on time-line}
\label{fig:dt2_1}
\end{figure}

\begin{figure}[!htpb]
\centering
\resizebox{\linewidth}{!}{
\begin{tikzpicture}[every node/.style = {font=\large},scale=0.75]

\node[blueblock](p1) at (-6,3) {Step 1} ;
\node[redblock](p2) at (-8,0) {Step 2} ;
\node[redblock](p3) at (-4,0) {Step 3} ;
\node[purpleblock](p4) at (0,0) {Compute \\ $e_n$};
\node[purpleblock](p5) at (4,0) {Compute  \\ $\Delta t_{new}$};
\node[greendiamond](p6) at (8,0) {$e_n < e_{tol}$};
\node[orangeblock](p7) at (0,-3) {Restore \\ Solutions};

\draw[->,thick](p1.east) -| (p4.north);
\draw[->,thick](p2.east) -- (p3.west);
\draw[->,thick](p3.east) -- (p4.west);
\draw[->,thick](p4.east) -- (p5.west);
\draw[->,thick](p5.east) -- (p6.west);
\draw[->,thick](p6.south) |- node[right] {no} (p7.east);
\draw[->,thick](p6.east) -- node[above] {yes} (11,0);
\draw[->,thick](p7.west) -| (-10,0) -- (p2.west);
\draw[->,thick](p7.west) -| (-10,0) |- (p1.west);

\end{tikzpicture}
}
\caption{Visualization of step doubling process with coding logic}
\label{fig:dt2_2}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Kinetics Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section describes kinetics examples that IQS is tested with and analysis of its performance.  The examples range in complexity and application.  The first is a one-dimensional problem, designed for the prototype code in MATLAB.  The next one is from the Argonne National Lab (ANL) Benchmark Problem Book (BPB), and is a common problem for testing codes and developing methods \cite{ANL_BPB}.

%---------------------------------------------------------------------------------------%
\subsection{One-Dimensional Problem}
%---------------------------------------------------------------------------------------%

This example is very simple and computes quickly; it entails a one dimensional, homogeneous 400 cm slab with a heterogenous perturbation in absorption cross section.  \fig{fig:slab} shows how the regions of the slab are divided and \tbl{tab:1Dmat} shows the initial material properties.  Regions 2, 3, and 4 have slope perturbations at different points in time, \tbl{tab:1Dslope} shows the values of the absorption cross-section in each region at the times of interest.  The values of $\Sigma_a$ between these times of interest are linear interpolations between the given values.

\begin{figure}[!htbp]
\begin{center}
\resizebox{\linewidth}{!}{
\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l |}
\hline \hline \hline
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 4 & 4 & 1 & 1 & 1 & 1 \\
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
\hline \hline \hline
\end{tabular}
}
\caption{1-D slab region identification}
\label{fig:slab}
\end{center}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{1-D slab material properties and problem parameters}
\label{tab:1Dmat}
\resizebox{\linewidth}{!}{
\begin{tabular}{llllll}
\hline
$D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu \Sigma_f (cm^{-1})$ & $v (cm/s)$ & $\beta$ & $\lambda (s^{-1})$ \\
\hline
1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\

\hline
\end{tabular}
}
\end{center}
\end{table}

\begin{table}[!htbp]
\begin{center}
\caption{1-D slab absorption cross-section at times of interest}
\label{tab:1Dslope}
\resizebox{\linewidth}{!}{
\begin{tabular}{llllllll}
\hline
Region & Material Property & 0.0 s & 0.1 s & 0.6 s & 1.0 s & 1.7 s \\
\hline
2 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.095 & 1.095 & 1.095 \\
3 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.09 & 1.09 & 1.1 \\
4 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.105 & 1.105 & 1.105 \\
\hline
\end{tabular}
}
\end{center}
\end{table}

\fig{fig:1D_flux} shows the resulting baseline flux and relative power profile of the one-dimensional problem.  The baseline was computed using MATLAB's ode15s function which is a embedded Runge-Kutta time adaptive method for stiff problems, the error tolerance was set very tightly ($10^{-12}$).  The flux distribution shows that the core is relatively large, thus different regions are weakly coupled. This baseline computation is used to compute the error of the other time discretization methods.  \fig{fig:1D_shape} shows the shape profile at various times during the transient.  It is apparent that the shape is very time-dependent, so it is expected that IQS has marginal accuracy gain for a given time step size.

\begin{figure}[!htbp]
\centering
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=\linewidth]{figures/1D_flux.png}
\caption{Flux profile at various times}
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=\linewidth]{figures/1D_power_base.png}
\caption{Power profile over transient}
\end{subfigure}
\caption{Baseline flux and power distribution}
\label{fig:1D_flux}
\end{figure}

\begin{figure}[!htbp]
\centering
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=\linewidth]{figures/1D_shape2.png}
\caption{Shape profile at various times}
\label{fig:1D_shape}
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=\linewidth]{figures/1D_iqs_power.png}
\caption{Power and reactivity profile over transient}
\end{subfigure}
\caption{IQS flux and power distribution}
\label{fig:1D_res}
\end{figure}

%----------------------------------------------------------------------------------------
\subsubsection{Iteration Analysis}

IQS, as previously stated, is a system of nonlinear equations between shape and amplitude. These equations needed to be iterated to numerically converge to an accurate solution.  \sct{sect:iter} lists various iteration techniques for fixed-point and Newton schemes. \fig{fig:iter} shows the number of fixed-point iterations required for a $10^{-11}$ tolerance over the transient. The criteria listed in the legend correspond to the list as such:
\begin{enumerate}
\item $L^{\infty} \rightarrow$ 1.
\item $L^{2} \rightarrow$ 2.
\item Reactivity $\rightarrow$ 3.
\item Amplitude $\rightarrow$ 4.
\item K criteria $\rightarrow$ 5.
\item All properties $\rightarrow$ 3.-5.
\end{enumerate}
This plot shows that 1-4 have approximately the same convergence behavior, but the K criteria converges to a certain error. \fig{fig:iter_err} shows the resulting error the K criteria converges to for different points of rescaling the shape. The rescaling is described by \eqt{eq:shape_scale}. Rescaling shape more frequently helps the error.  However, rescaling every iteration is somewhat artificial because it does not consider changes spatially. Regardless, an error of $~10^{-5}$ is quite large and it is expected that the magnitude is due to the explicit treatment of precursors (\eqt{eq:dnp_theta}). Switching to an analytical elimination (\eqt{eq:dnp_an}) does not converge as well, but the error is much smaller, seen in \fig{fig:iter_err_an}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{figures/iter_renorm.png}
\caption{\# of iterations for various convergence criteria, tolerance$=10^{-11}$, max iterations$=20$}
\label{fig:iter}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{figures/iter_error.png}
\caption{Final iteration error for K convergence criteria}
\label{fig:iter_err}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{figures/iter_error_an.png}
\caption{Final iteration error for K convergence criteria with analytical precursor elimination}
\label{fig:iter_err_an}
\end{figure}

%----------------------------------------------------------------------------------------
\subsubsection{One-Dimensional Problem Time Step Convergence Analysis}

Time step convergence analysis involves evaluating a problem with various refinements in step size and comparing the resulting errors with the time step size. Plotting error versus $\Delta t$ on a log-scale should produce a relatively strait line with a slope equal to the order of the time discretization method. In order to evaluate the performance and error convergence of IQS, the slab was simulated with varying time discretization methods and time step sizes.  \fig{fig:1D_conv} shows these convergence plots of five different discretization methods for implicit discretization, IQS, and IQS P-C.  These plots were generated from the results using the MATLAB prototype program.  The plots show that IQS and IQS P-C are convergent through fourth order BDF.  Third order SDIRK did not show third order convergence, but, through extensive testing, SDIRK shows non-convergent behavior for too stiff of problems.

There are higher discretization order that can be tested, but most practical application do not go beyond second order.  This paper shows an analysis of the first publicized application of IQS with higher than second order discretization, which exposed unforeseen properties of IQS.  When using higher order techniques, the interpolation of PRKE parameters and shape for precursor integration become important to consider.  Every other application that was investigated linearly interpolates parameters for the PRKE evaluation.  Similarly, the shape used for the integration of the ODE for the precursors needs to have higher order interpolation to preserve high order error convergence.  This interpolation was done using Lagragian and Hermite methods, both leading to successful convergence.

\begin{figure*}[!htbp]
\centering
\begin{subfigure}[b]{0.49\textwidth}
\centering
\includegraphics[width=\linewidth]{figures/1D_conv_IE.png}
\caption{Implicit Euler}
\end{subfigure}
\begin{subfigure}[b]{0.49\textwidth}
\centering
\includegraphics[width=\linewidth]{figures/1D_conv_BDF2.png}
\caption{BDF2}
\end{subfigure}
\begin{subfigure}[b]{0.49\textwidth}
\centering
\includegraphics[width=\linewidth]{figures/1D_conv_BDF3.png}
\caption{BDF3}
\end{subfigure}
\begin{subfigure}[b]{0.49\textwidth}
\centering
\includegraphics[width=\linewidth]{figures/1D_conv_BDF4.png}
\caption{BDF4}
\end{subfigure}
\begin{subfigure}[b]{0.49\textwidth}
\centering
\includegraphics[width=\linewidth]{figures/1D_conv_SDIRK33.png}
\caption{SDIRK33}
\end{subfigure}
\caption{Error convergence plots of implicit discretization, IQS, and IQS P-C with various time discretization schemes}
\label{fig:1D_conv}
\end{figure*}

%----------------------------------------------------------------------------------------
\subsubsection{One-Dimensional Mini-Core Problem}

This problem is exactly the same as the previous one-dimensional problem, except the the core was reduced to 80 cm in length. The purpose of testing this problem is to determine if a more tightly coupled core will yield better performance for IQS. \fig{fig:1Dmini_flux} shows the resulting baseline flux and relative power profile of the one-dimensional mini-core problem. This plot shows that the perturbed regions affect the domain more evidently. \fig{fig:1Dmini_shape} shows the shape profile at various times during the transient. This plot shows that the shape is much less time-dependent than the previous large core. \fig{fig:1Dmini_conv} shows that IQS performs significantly better with this example than the large core.

\begin{figure}[!htbp]
\centering
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=\linewidth]{figures/1Dmini_flux.png}
\caption{Flux profile at various times}
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=\linewidth]{figures/1Dmini_power_base.png}
\caption{Power profile over transient}
\end{subfigure}
\caption{Mini-core baseline flux and power distribution}
\label{fig:1Dmini_flux}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{figures/1Dmini_shape.png}
\caption{Shape profile at various times for one-dimensional mini-core}
\label{fig:1Dmini_shape}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{figures/1Dmini_conv_IE.png}
\caption{Time step convergence for one-dimensional mini-core with implicit Euler discretization}
\label{fig:1Dmini_conv}
\end{figure}

%---------------------------------------------------------------------------------------%
\subsection{TWIGL Benchmark}
%---------------------------------------------------------------------------------------%

This benchmark problem originates from the Argonne National Lab Benchmark Problem Book \cite{ANL_BPB}.  It is a 2D, 2-group reactor core model with no reflector region \cite{TWIGL_benchmark}.  This example is meant to be of progressive complexity from the previous example.  The transient of this reactor is very geometrically symmetrical with very little temporal shape change.  Therefore, IQS is expected to perform significantly better than the implicit discretization method. Figs. \ref{fig:TWIGL_power} show the IQS  solution as compared with the implicit discretization solution. This shows that IQS is consistent in more complex, higher dimensional problems. These plots also serve to illustrate that IQS has a much more accurate solution, even at a significantly larger time step than the implicit discretization.

\begin{figure}[!htbp]
\centering
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{figures/TWIGL_power_plot.png}
\caption{Power profile for entire transient}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{figures/TWIGL_power_plot2.png}
\caption{Power at cusp of profile}
\end{subfigure}
\caption{Power level comparison of TWIGL Benchmark}
\label{fig:TWIGL_power}
\end{figure}

%----------------------------------------------------------------------------------------
\subsubsection{TWIGL Time Step Convergence Analysis}

In order to demonstrate asymptotic convergence of IQS, implicit Euler (IE) and second order BDF (BDF2) were applied to the TWIGL simulation. \fig{fig:TWIGL_conv} plots the error convergence of IQS and the implicit discretization methods.  The curves show the impressive convergence of IQS for the highly transient TWIGL example. The slope indicated in the legend are the linear slope of curves on the log plot, these slopes should be similar to the order of the method (1 for IE and 2 for BDF2).  IQS shows a increased order because the PRKE is performing much of accuracy convergence and it is computed using SDIRK33, a third order method.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{figures/TWIGL_convergence.jpg}
\caption{Error convergence comparison of TWIGL Benchmark}
\label{fig:TWIGL_conv}
\end{figure}

%----------------------------------------------------------------------------------------
\subsubsection{TWIGL Time Adaptation}

\tbl{tab:TWIGLdt2} and \fig{fig:TWIGL_power_dt2} show the results for TWIGL with time adaptation.  The results show that both IQS methods perform exceptionally well compared to implicit discretization.  It also shows that traditional IQS performed better with large $e_{tol}$, while IQS P-C was better with smaller $e_{tol}$.

\begin{table*}[!htbp]
\begin{center}
\caption{TWIGL step doubling results}
\label{tab:TWIGLdt2}
%\resizebox{\linewidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
\hline
  &  & \multicolumn{3}{|c|}{Implicit Discretization} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
\hline
1 &	0.05    &	0.00012677 &	9   &	29  &	0.03380433 &	4   &	20   &	0.00323100 &	4  &	9   \\
2 &	0.01    &	3.5555e-05 &	11  &	35  &	0.00166991 &	5   &	40   &	0.00263068 &	5  &	12  \\
3 &	0.005   &	4.0364e-05 &	11  &	31  &	0.00886584 &	5   &	40   &	0.00160486 &	6  &	21  \\
4 &	0.001   &	0.00294822 &	33  &	122 &	0.02976305 &	5   &	36   &	1.7527e-05 &	10 &	35  \\
5 &	0.0005  &	0.00099778 &	39  &	131 &	0.00143781 &	6   &	55   &	1.4185e-05 &	16 &	74  \\
6 &	0.0001  &	0.00019510 &	78  &	236 &	0.00016175 &	8   &	65   &	6.2903e-06 &	19 &	78  \\
7 &	5.0e-05 &	0.00018372 &	112 &	342 &	6.0328e-05 &	12  &	163  &	1.5247e-06 &	24 &	92  \\
8 &	1.0e-05 &	8.0564e-05 &	263 &	794 &	7.7103e-05 &	379 &	5729 &	9.8321e-07 &	48 &	210 \\
\hline

\end{tabular}%}
\end{center}
\end{table*}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{figures/TWIGL_power_plot_dt2.png}
\caption{Power level comparison of TWIGL Benchmark with time adaptation}
\label{fig:TWIGL_power_dt2}
\end{figure}

%---------------------------------------------------------------------------------------%
\subsection{LMW Benchmark?}
%---------------------------------------------------------------------------------------%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dynamics Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section describes two dynamics examples, including the LRA benchmark and a TREAT experiment.  These examples are of increased complexity from the previous kinetics examples.  This section also analyzes IQS's performance with these, which is vital for verification of IQS in real-world problems.

%---------------------------------------------------------------------------------------%
\subsection{LRA Benchmark}
%---------------------------------------------------------------------------------------%

The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor.  It is a super prompt-critical transient. The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.  The spacial discretization was performed using continuous finite element method with first order Lagrangian basis functions. The mesh consisted of blocks $11\times 11$ with five uniform refinements, totaling $165,165$ elements and $124,609$ nodes. Three different temporal techniques were applied: implicit discretization of the flux equation, IQS, and IQS-PC. Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.  Third order Runge-Kutta discretization with step doubling adaptation was used for the PRKE evaluation.  The performance of IQS and the temperature updates were measured by its improvement in accuracy at peak power over the implicit discretization method.

\fig{fig:lra_profile} shows the baseline power and temperature transient profile for the LRA benchmark. The baseline results are compared to the results achieved by Sutton and Aviles in \cite{Sutton_1996} and presented in \tbl{tab:base}.  The relative difference in the magnitude of the peak power ($t\approx1.44 s$) from the baseline was used for error comparison.

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{figures/lra_profile.png}
\caption{LRA baseline temperature and power profile}
\label{fig:lra_profile}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{LRA baseline verification}
\label{tab:base}
\resizebox{\linewidth}{!}{
\begin{tabular}{|l|cc|}
\hline
Calculation  &  Baseline & Sutton (Spandex 1936) \\
\hline
No. of Spatial Nodes	& 3872 		& 1936 \\
Eigenvalue 				& 0.99637	& 0.99637 \\
No. of Time Steps 		& 6000 		& 23,890 \\
Time to Peak Power (s) 	& 1.441 	& 1.441 \\
Peak Power (W/cm$^3$) 	& 5456 		& 5461 \\
\hline
\end{tabular}
}
\end{center}
\end{table}

%----------------------------------------------------------------------------------------
\subsubsection{LRA Results}

This section shows the time step error convergence of IQS for the LRA benchmark, as well as the effect of the intermediate temperature time scale. \fig{fig:lra_bad} is an error convergence plot comparing the three techniques where temperature is evaluated only on the macro step (1 temperature update).  \fig{fig:lra_mpconv} is an error convergence plot comparing the three techniques where temperature is evaluated 5 times within a macro step (5 temperature updates).  Finally, \fig{fig:mp} shows the effect of various temperature updates. The dashed lines correspond to implicit discretization at different flux step sizes, while the IQS macro step size is kept constant.

\begin{figure}[!htbp]
\centering
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{figures/lra_bad.png}
\caption{Only one temperature update per macro step}
\label{fig:lra_bad}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{figures/lra_mp_convergence.png}
\caption{Five temperature updates per macro step}
\label{fig:lra_mpconv}
\end{subfigure}
\caption{LRA error convergence plots}
\end{figure}

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{figures/lra_mp.png}
\caption{Error plot with various temperature updates per macro step}
\label{fig:mp}
\end{figure}

The convergence plots show that updating temperature and the PRKE parameters within a macro step has a significant effect on the performance of IQS.  With only one update, IQS was only slightly better than implicit discretization, implicit discretization required about 150\% more time steps than IQS for the same error.  While 5 temperature updates showed a much more significant IQS performance, implicit discretization required about 400\% more time steps than IQS for the same error.  \fig{fig:mp} shows that error has a convergent behavior for the number of temperature updates.  This convergence makes sense because temperature can only be so accurate before the error in shape is dominating. Table \ref{tab:ndiff_lra} shows the run time results for the implicit discretization calculations. The number of GMRES linear iterations is included because it is proportional measure of the computational effort. Tables \ref{tab:iqs_lra} and \ref{tab:iqspc_lra} present the IQS run-times with various numbers of temperature updates.  These run-times are based on total alive time of the execution where the diffusion evaluation is distributed over 24 processors. These run-times show a marginal performance for IQS and impressive performance for IQS P-C.  Some of the execution times were able to decrease from implicit discretization with the same number of macro steps because IQS is better equipped to resolve the nonlinearity between temperature and amplitude. Furthermore, there does seem to be an ideal number of temperature updates to optimize execution time: IQS only needs one and IQS P-C seems to be ideal at 4 updates. This discrepancy in the number of updates shows that a adaptive type implementation of the updates would be ideal, and could enforce a constant error over the transient. It is also important to compare the error of implicit discretization with IQS at one update and IQS P-C at 4 updates.  IQS shows an error comparable to implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by -34.1\%.  IQS P-C shows an error less than implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by <-34.9\%.

\begin{table}[!htbp]
\begin{center}
\caption{Implicit discretization run time results}
\label{tab:ndiff_lra}
\begin{tabular}{|l|l|ccc|}
\hline
Run  &  $\Delta t$ & Error & Runtime (hr) & Linear Iter.\\
\hline
1	& 4.0e-3	& 1.407e-2 	& 4.11	& 7.13e4	\\
2	& 2.0e-3	& 3.174e-3 	& 6.01	& 9.49e4 	\\
3 	& 1.0e-3 	& 7.690e-4 	& 10.38	& 1.45e5	\\
4 	& 5.0e-4 	& 1.892e-4 	& 21.91	& 2.08e5	\\
5 	& 2.5e-4	& 4.590e-5 	& 25.23	& 3.16e5	\\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[!htbp]
\begin{center}
\caption{IQS run time results with $\Delta t = 0.004$}
\label{tab:iqs_lra}
\begin{tabular}{|l|l|ccc|}
\hline
	&  Temperature 	&  		& Runtime 	& \% Increase	\\
Run	&  Updates 	& Error & (hr)		& in Runtime$^*$\\
\hline
1	& 1		& 2.612e-3 	& 3.96 	& -3.18\%	\\
2	& 2		& 9.893e-4 	& 6.02	&  47.1\%	\\
3 	& 4 	& 5.796e-4 	& 7.87	&  92.3\%	\\
4 	& 8 	& 4.772e-4 	& 12.61	& 207.9\% 	\\
5 	& 16	& 4.516e-4 	& 22.14	& 440.7\%	\\
\hline
\end{tabular}
\\
$^*$ difference in runtime from $\Delta t = 0.004$ implicit discretization 
\end{center}
\end{table}

\begin{table}[!htbp]
\begin{center}
\caption{IQS PC run time results with $\Delta t = 0.004$}
\label{tab:iqspc_lra}
\begin{tabular}{|l|l|ccc|}
\hline
	&  Temperature 	&  		& Runtime 	& \% Increase	\\
Run	&  Updates 	& Error & (hr)		& in Runtime$^*$\\
\hline
1	& 1		& 3.488e-3 	& 2.91 	& -28.9\%	\\
2	& 2		& 1.349e-3 	& 3.73	& -9.00\%	\\
3 	& 4 	& 9.161e-4 	& 3.97	& -3.04\%	\\
4 	& 8 	& 8.052e-4 	& 5.39	&  31.7\%	\\
5 	& 16	& 7.905e-4 	& 8.19	&  100\%	\\
\hline
\end{tabular}
\\
$^*$ difference in runtime from $\Delta t = 0.004$ implicit discretization 
\end{center}
\end{table}

%----------------------------------------------------------------------------------------
\subsubsection{LRA Time Adaptation}

\fig{fig:lra_dt2} shows the power profile of the LRA with time adaptation of implicit discretization and IQS P-C, and \tbl{tab:lra_dt2} compiles the results. These time adaptation results show the significant decrease in macro time steps required for IQS P-C. These profiles were obtaining with only one temperature update per macro step; so based on previous results, the IQS P-C performance would improve even more with more updates.

\begin{figure}[!htbp]
\centering
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{figures/LRA_DT2.png}
\caption{Full power profile}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{figures/LRA_DT2_peak.png}
\caption{Peak power profile at peak}
\end{subfigure}
\caption{LRA power profile with time adaptation of implicit discretization and IQS P-C}
\label{fig:lra_dt2}
\end{figure}

\begin{table}
\begin{center}
\caption{LRA step doubling adaptation results with implicit discretization and IQS P-C}
\label{tab:lra_dt2}
\resizebox{\linewidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
 & \multicolumn{3}{|c|}{Implicit Dis.} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Event & Power & Error & Steps & Power  & Error & Steps \\
 & (W/cm$^3$) &  &  &  (W/cm$^3$) &  &  \\
\hline
Max Power & 5567.3 & 0.019454 & 423 & 5568.3 & 0.019274 & 47 \\
End (3 s) & 109.66 & 2.3650e-4 & 603 & 109.65 & 3.0622e-4 & 97 \\
\hline
\end{tabular}}
\end{center}
\end{table}

%---------------------------------------------------------------------------------------%
\subsection{TREAT Transient-15 Problem}
%---------------------------------------------------------------------------------------%

%----------------------------------------------------------------------------------------
\subsubsection{Transient-15 Multiphysics Time Scale Results}

%----------------------------------------------------------------------------------------
\subsubsection{Transient-15 Time Adaptation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrvnat}
\bibliography{references_IQS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------

\end{document}

### 2017-03-21 doc/Thesis/data/dynamics.tex
Thesis: Some final changes before submission

@@
 \includegraphics[width=0.8\linewidth]{\FiguresDir/lra.png}
-\caption{LRA benchmark geometry with region assignment.}
+\caption{LRA benchmark geometry with region assignment \cite{ANL_BPB}}
 \label{fig:lra-geometry}
@@
     \centering
-    \caption{LRA benchmark initial two-group constants.\label{tab:lra-xs}}
+    \caption{LRA benchmark initial two-group constants \cite{ANL_BPB} \label{tab:lra-xs}}
     \resizebox{\textwidth}{!}{
@@
 \includegraphics[width=\linewidth]{\FiguresDir/lra_profile.png}
-\caption{LRA baseline temperature and power profile}
+\caption{LRA baseline temperature and power profile \cite{PrinceMC2017}}
 \label{fig:lra_profile}
@@
 \includegraphics[width=\linewidth]{\FiguresDir/lra_profile2D.png}
-\caption{LRA baseline spacial power profile at $t=1.44 s$}
+\caption{LRA baseline spacial power profile at $t=1.44 s$ \cite{PrinceMC2017}}
 \label{fig:lra_profile2D}
@@
 \begin{center}
+\caption{LRA baseline verification}
+\label{tab:base}
 %\resizebox{\linewidth}{!}{
@@
 \end{center}
-\caption{LRA baseline verification}
-\label{tab:base}
 \end{table}
@@
 \end{subfigure}
-\caption{LRA error convergence plots}
+\caption{LRA error convergence plots \cite{PrinceMC2017}}
 \end{figure}
@@
 \includegraphics[width=\linewidth]{\FiguresDir/lra_mp.png}
-\caption{Error plot with various temperature updates per macro step}
+\caption{Error plot with various temperature updates per macro step \cite{PrinceMC2017}}
 \label{fig:mp}
@@
 \begin{center}
+\caption{Implicit discretization run time results}
+\label{tab:ndiff_lra}
 %\resizebox{\linewidth}{!}{
@@
 \end{center}
-\caption{Implicit discretization run time results}
-\label{tab:ndiff_lra}
 \end{table}
@@
 \begin{center}
+\caption{IQS run time results with $\Delta t = 0.004$}
+\label{tab:iqs_lra}
 %\resizebox{\linewidth}{!}{
@@
 $^*$ difference in runtime from $\Delta t = 0.004$ implicit discretization 
-\caption{IQS run time results with $\Delta t = 0.004$}
-\label{tab:iqs_lra}
 \end{center}
@@
 \begin{center}
+\caption{IQS PC run time results with $\Delta t = 0.004$}
+\label{tab:iqspc_lra}
 %\resizebox{\linewidth}{!}{
@@
 $^*$ difference in runtime from $\Delta t = 0.004$ implicit discretization 
-\caption{IQS PC run time results with $\Delta t = 0.004$}
-\label{tab:iqspc_lra}
 \end{center}
@@
 \includegraphics[width=\linewidth]{\FiguresDir/time_constant_lra.png}
-\caption{Dynamical time scale for LRA benchmark}
+\caption{Dynamical time scale for LRA benchmark \cite{PrinceMC2017}}
 \label{fig:LRAtc}
@@
 \begin{center}
+\caption{LRA step doubling adaptation results with implicit discretization and IQS P-C}
+\label{tab:lra_dt2}
 \resizebox{\textwidth}{!}{
@@
 \end{center}
-\caption{LRA step doubling adaptation results with implicit discretization and IQS P-C}
-\label{tab:lra_dt2}
 \end{table}
@@
 \includegraphics[width=0.7\linewidth]{\FiguresDir/Tran15_config.png}
-\caption{Transient-15 159-element small core configuration}
+\caption{Transient-15 159-element small core configuration \cite{mammoth}}
 \label{fig:Tran15_config}
@@
 \end{subfigure}
-\caption{Top quarter view of Transient-15 mesh}
+\caption{Top quarter view of Transient-15 mesh \cite{mammoth}}
 \label{fig:Tran15_mesh}
@@
 \end{subfigure}
-\caption{Top single block view of Transient-15 mesh}
+\caption{Top single block view of Transient-15 mesh \cite{mammoth}}
 \label{fig:Tran15_mesh_zoom}
@@
 \includegraphics[width=0.6\linewidth]{\FiguresDir/Tran15_core2.png}
-\caption{Transient-15 core power profile at peak power}
+\caption{Transient-15 core power profile at peak power \cite{PrinceMC2017}}
 \label{fig:Tran15}
@@
 \includegraphics[width=\linewidth]{\FiguresDir/Tran15_profile.png}
-\caption{Transient-15 total power and average temperature profile during transient}
+\caption{Transient-15 total power and average temperature profile during transient \cite{PrinceMC2017}}
 \label{fig:Tran15_profile}
@@
 \begin{center}
+\caption{Transient-15 Error and Runtime Results}
+\label{tab:tran15}
 \resizebox{\linewidth}{!}{
@@
 $^*$ difference in runtime from implicit discretization 
-\caption{Transient-15 Error and Runtime Results}
-\label{tab:tran15}
 \end{table}
@@
 \includegraphics[width=0.7\linewidth]{\FiguresDir/time_constant_tran15.png}
-\caption{Dynamical time scale for the Transient-15 example}
+\caption{Dynamical time scale for the Transient-15 example \cite{PrinceMC2017}}
 \label{fig:tran15tc}

### 2017-03-21 doc/Thesis/data/kinetics.tex
Thesis: Some final changes before submission

@@
 
-Neutron dynamics is the study of the time-dependent nature of neutrons in a reactor.
+Neutron kinetics is the study of the time-dependent nature of neutrons in a reactor.
 More specifically, there is no coupling of neutron behavior with other physics, like thermal hydraulic feedback.
@@
 \end{tabular}
-\caption{1-D slab region identification}
+\caption{1-D slab region identification \cite{PrincePHYSOR2016}}
 \label{fig:slab}
@@
 \begin{center}
-\caption{1-D slab material properties and problem parameters}
+\caption{1-D slab material properties and problem parameters \cite{PrincePHYSOR2016}}
 \label{tab:1Dmat}
@@
 \begin{center}
-\caption{1-D slab absorption cross-section at times of interest}
+\caption{1-D slab absorption cross-section at times of interest \cite{PrincePHYSOR2016}}
 \label{tab:1Dslope}
@@
 \caption{Shape profile at various times}
-\label{1D_shape}
+\label{fig:1D_shape}
 \end{subfigure}
@@
 \includegraphics[width=0.75\textwidth]{\FiguresDir/TWIGL_regions.jpg}
-\caption{TWIGL benchmark problem description}
+\caption{TWIGL benchmark problem description \cite{Ban_2012}}
 \label{fig:TWIGL_reg}
@@
 \begin{center}
-\caption{TWIGL benchmark material properties and slope perturbation}
+\caption{TWIGL benchmark material properties and slope perturbation \cite{Ban_2012}}
 \label{tab:TWIGL_mat}
@@
 \ref{fig:TWIGL_power} and \ref{fig:TWIGL_plots} show the IQS  solution as compared with the implicit discretization solution.
-It is important to note the IQS shape plot is scaled differently than the Brute Force flux plot (\fig{fig:TWIGL_plots}) because the amplitude term is not included, but the gradients of colors is comparable.
+It is important to note the IQS shape plot is scaled differently than the implicit discretization flux plot (\fig{fig:TWIGL_plots}) because the amplitude term is not included, but the gradients of colors is comparable.
 These plots show that IQS is consistent in more complex, higher dimensional problems in Rattlesnake.
@@
 \end{subfigure}
-\caption{TWIGL Benchmark flux/shape comparison at $t=0.2$}
+\caption{TWIGL Benchmark flux/shape comparison at $t=0.2$ \cite{PrincePHYSOR2016}}
 \label{fig:TWIGL_plots}

### 2017-03-21 doc/Thesis/data/theory.tex
Thesis: Some final changes before submission

@@
 \includegraphics[width=\linewidth]{\FiguresDir/IQS_visualization.jpg}
-\caption{IQS method visualization}
+\caption{IQS method visualization \cite{PrincePHYSOR2016}}
 \label{fig:iqsviz}
@@
 By far, the most common nonlinear system iteration method in MOOSE is Preconditioned Jacobian-Free Newton-Krylov with Generalized Minimal RESidual method (GMRES) as the linear system solver.
-This section only descirbes the methods' application to IQS; a very detailed description of PJFNK is transcribed by Knoll in \cite{PJFNK_Knoll}.
+This section only describes the methods' application to IQS; a very detailed description of PJFNK is transcribed by Knoll in \cite{PJFNK_Knoll}.
 Essentially, the IQS system of equations can be described by \eqt{eq:IQSsys}.
@@
 \be
-T^{n+1} = T^n + \frac{\kappa_f}{\rho c_p} \left(a_2 \varphi^{n+1} + a_1 \varphi^{n}\right)
+T^{n+1} = T^n + \frac{\kappa_f}{\rho c_p} \sum_{g=1}^G\left(a_2(\Sigma_f^g\varphi^g)^{n+1} + a_1(\Sigma_f^g\varphi^g)^n\right)
 \label{eq:temp_an}
@@
 \draw[shift={(6,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
-\node(temp) at (-.75,1) {$T \rightarrow \rho, \bar{\beta}, \bar{\lambda}, \Lambda$:};
+\node(temp) at (-.4,1) {$T, \rho$:};
 
@@
 \draw[shift={(6,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
-\node(prke) at (-.5,0) {$p, \frac{dp}{dt}:$};
+\node(prke) at (-.2,0) {$p:$};
 
@@
 \draw (6,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (6,0);
-\draw (6,0) edge[out=45,in=0,->,thick] node[below,sloped] {$p, \frac{dp}{dt}$} (6,2);
+\draw (6,1) edge[out=45,in=-45,->,thick] node[below,sloped] {$T$} (6,2);
+\draw (6,0) edge[out=30,in=0,->,thick] node[below,sloped] {$p, \frac{dp}{dt}$} (6,2);
 
@@
 \fig{fig:proc} shows the programming diagram implement to execute this process.
-The time increment of $\Delta t/3$ for the temperature solve is arbitrary and is meant only to match \fig{fig:time} where three temperature updates have been used as illustration.
+The time increment of $\Delta t^T$ corresponds to the time step that the temperature is evaluated.
 
@@
 \node[greenblock] (p6) at (0,-12.5) {Solve PRKE};
-\node[reddiamond] (check1) at (0,-17.5) {Check for \\ convergence ($p$)};
+\node[reddiamond] (check1) at (0,-17.5) {Check for \\ convergence ($p,T$)};
 \node[reddiamond] (check2) at (7,-12) {If $t=t_{n+1}$};
 \node[greenblock] (p7) at (8,-6) {Shape Solve};
-\node[reddiamond] (check3) at (8,-2.5) {Check for \\ convergence ($\varphi$)};
+\node[reddiamond] (check3) at (8,-2.5) {Check for \\ convergence ($\varphi,K_{n+1}$)};
 %\node [above =0.1mm of p1] {\large{IQS Solve:}};
@@
 \draw[->,ultra thick](p1.south) -- node[right] {$t=t_{n+1}$}(p2.north);
-\draw[->,ultra thick](p2.south) -- node[right] {$t=t_{n}+\Delta t/3$}(p3.north);
+\draw[->,ultra thick](p2.south) -- node[right] {$t=t_{n}+\Delta t^T$}(p3.north);
 \draw[->,ultra thick](p3.south) -- (p4.north);
@@
 \draw[->,ultra thick](check1.east) -| node[right] {\large{yes}} (check2.south);
-\draw[->,ultra thick](check2.west) -- node[above,sloped] {\large{no}\small \qq $t=t+\Delta t/3$} (4,-5) -- (p3.east);
+\draw[->,ultra thick](check2.west) -- node[above,sloped] {\large{no}\small \qq $t=t+\Delta t^T$} (4,-5) -- (p3.east);
 \draw[->,ultra thick](check2.north) --  (8,-10) -- node[right] {\large{yes}} (p7.south);
@@
 A vital part of the verification and validation for IQS is analyzing error convergence.
-Since any time discretization scheme is capable of being applied to the shape equation of IQS, it is important to investigate IQS's performance to a variety of these schemes.
+Since any time discretization scheme can be applied to the shape equation of IQS, it is important to investigate IQS's performance to a variety of these schemes.
 There is lack of literature that applies IQS to schemes other than implicit Euler; higher order schemes are never rigorously tested.
@@
 \tbl{tab:bdfa} shows the values of $\alpha$ for every order implemented in this thesis.
-The benefit of using BDF is that any order of convergence can be applied and are unconditionally stable up through fourth order, so the IQS approximation can easily be validated for high order discretization.
+The benefit of using BDF is that any order of convergence can be applied and are practically stable up through fourth order, so the IQS approximation can easily be validated for high order discretization.
 
@@
 \be
-\frac{C^{n+1}-C^n}{\Delta t}=(1-\theta)\beta S_f^np^n-(1-\theta)\lambda C^n + \theta\beta S_f^{n+1}p^{n+1}-\theta\lambda C^{n+1}
+\frac{C^{n+1}_i-C^n_i}{\Delta t}=(1-\theta)\beta_i S_f^np^n-(1-\theta)\lambda_i C^n_i + \theta\beta_i S_f^{n+1}p^{n+1}-\theta\lambda_i C^{n+1}_i
 \ee
-Where $S_f$ is the fission source equivalent for shape ($S_f^n=(\nu\Sigma_f)^n\varphi^n$). Rearranging to solve for the precursor at the end of the time step yields
+Where $S_f$ is the fission source equivalent for shape ($S_f^n=\sum_{g=1}^G(\nu\Sigma_f^g\varphi^g)^n$). Rearranging to solve for the precursor at the end of the time step yields
 \be
-C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta}{1+\theta\Delta t\lambda}S_f^n p^n +  \frac{\theta\Delta t \beta}{1+\theta\Delta t\lambda}S_f^{n+1} p^{n+1}
+C^{n+1}_i = \frac{1-(1-\theta)\Delta t\lambda_i}{1+\theta\Delta t\lambda_i}C^n_i + \frac{(1-\theta)\Delta t \beta_i}{1+\theta\Delta t\lambda_i}S_f^n p^n +  \frac{\theta\Delta t \beta_i}{1+\theta\Delta t\lambda_i}S_f^{n+1} p^{n+1}
 \label{eq:dnp_theta}
 \ee
-Reporting this value of $C^{n+1}$, one can solve for the shape $\varphi^{n+1}$ as a function of $\varphi^n$ and $C^n$ (and $p^n$, $p^{n+1}$, $dp/dt|_n$ and  $dp/dt|_{n+1}$).
-Once $\varphi^{n+1}$ has been determined, $C^{n+1}$ is updated.
-Applying this technique to implicit discretization is done by changing the definition the fission source ($S_f^n=(\nu\Sigma_f)^n\phi^n$) and eliminating all $p$ terms.
+Reporting this value of $C^{n+1}_i$, one can solve for the shape $\varphi^{n+1,g}$ as a function of $\varphi^{n,g}$ and $C^n_i$ (and $p^n$, $p^{n+1}$, $dp/dt|_n$ and  $dp/dt|_{n+1}$).
+Once $\varphi^{n+1,g}$ has been determined, $C^{n+1}_i$ is updated.
+Applying this technique to implicit discretization is done by changing the definition the fission source ($S_f^n=\sum_{g=1}^G(\nu\Sigma_f^g\phi^g)^n$) and eliminating all $p$ terms.
 
@@
 \be
-C^{n+1} =  C^n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
+C^{n+1}_i =  C^n_i e^{-\lambda_i (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta_i(t') S_f(t') p(t')e^{-\lambda_i (t_{n+1}-t')}dt'
 \label{eq:prec_an}
@@
 \be
-C^{n+1} = C^n e^{-\lambda \Delta t} + \left(\hat{a}_2 S_f^{n+1}+\hat{a}_1 S_f^n\right)\beta
+C^{n+1}_i = C^n_i e^{-\lambda_i \Delta t} + \left(\hat{a}_{2,i} S_f^{n+1}+\hat{a}_{1,i} S_f^n\right)\beta_i
 \label{eq:dnp_an}
@@
 \begin{align}
-&\hat{a}_1= \int_{t_n}^{t_{n+1}}\frac{t_{n+1}-t'}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
-&\hat{a}_2 = \int_{t_n}^{t_{n+1}}\frac{t'-t_n}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt'
+&\hat{a}_{1,i}= \int_{t_n}^{t_{n+1}}\frac{t_{n+1}-t'}{\Delta t}p(t')e^{-\lambda_i(t_{n+1}-t')}dt' \\
+&\hat{a}_{2,i} = \int_{t_n}^{t_{n+1}}\frac{t'-t_n}{\Delta t}p(t')e^{-\lambda_i(t_{n+1}-t')}dt'
 \end{align}

### 2017-04-05 doc/Thesis/data/abstract.tex
Thesis: Changes after OGAPS round 1 feedback

@@
 Furthermore, reactor simulations often require coupling to additional multiphysics components for temperature feedback.
-These additional physics may evolve on different time scales and the applicability of IQS in multiphysic simulations needs to be evaluated from this perspective as well.
+These additional physics may evolve on different time scales, and the applicability of IQS in multiphysic simulations needs to be evaluated from this perspective as well.
 
-The objectives of this research is to establish IQS performance with various iteration techniques, validate time step convergence of IQS, and apply IQS to multiphysics simulation.
+The objectives of this research are to establish IQS performance with various iteration techniques, validate time step convergence of IQS, and apply IQS to multiphysics simulation.
 IQS iteration techniques involve fixed-point (Picard) iteration with various convergence criteria and Newton iteration, namely preconditioned Jacobian-free Newton Krylov (PJFNK) method.
@@
 IQS application to multiphysics simulations involves its implementation into the Rattlesnake/MOOSE framework.
-These simulations allow insight to the performance of IQS for full transient reactor simulations.
+These simulations allow insight into the performance of IQS for full transient reactor simulations.
 

### 2017-04-05 doc/Thesis/data/theory.tex
Thesis: Changes after OGAPS round 1 feedback

@@
 
-\chapter{Improved Quasi-Static Method}
+\chapter{\uppercase{Improved Quasi-Static Method}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \frac{\rho-\bar{\beta}}{\Lambda} = 
-\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}
+\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}\sum_{g'=1}^G \nu^{g'}_p \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} +\left( \div D^g \grad  - \Sigma_r^g \right)\varphi^g\right)}
 {\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
@@
 
-Where $n$ is the time step, $k$ is the iteration number, and $\epsilon_{\varphi}$ is the numerical criteria provided by a user.
+where $n$ is the time step, $k$ is the iteration number, and $\epsilon_{\varphi}$ is the numerical criteria provided by a user.
 Through testing, iterations do not always converge; in this circumstance, the maximum number of iterations is reached.
@@
 
-Where $K$ is the IQS uniqueness expression:
+where $K$ is the IQS uniqueness expression:
 \be 
@@
 \ee 
-Where $J$ is the Jacobian matrix defined as $J_{ij}=\partial R_i/\partial\varphi_j$, $\delta\varphi$ is the error in the iterating solution, and $R$ is the residual vector defined as $A\varphi-F$ (most methods simply define the residual instead of evaluating this expression).
+where $J$ is the Jacobian matrix defined as $J_{ij}=\partial R_i/\partial\varphi_j$, $\delta\varphi$ is the error in the iterating solution, and $R$ is the residual vector defined as $A\varphi-F$ (most methods simply define the residual instead of evaluating this expression).
 For IQS, a residual evaluation entails a PRKE parameter evaluation and the PRKE evaluation over the entire macro-step.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Predictor-Corrector version of IQS (IQS P-C)}
+\section{Predictor-Corrector IQS (IQS P-C)}
 \label{sect:pc}
@@
 \ee
-Where $n$ corresponds to the beginning of the temperature step.
+where $n$ corresponds to the beginning of the temperature step.
 $a_1$ and $a_2$ are integration coefficients defined by \eqt{eq:a1} and \eqt{eq:a2}.
@@
 \centering
+\resizebox{\linewidth}{!}{
 \begin{tikzpicture}[scale=2]
@@
 
-\end{tikzpicture}
+\end{tikzpicture}}
 \caption{Time scales and process of IQS with temperature feedback}
@@
 \centering
+\resizebox{\linewidth}{!}{
 \begin{tikzpicture}[every node/.style = {font=\normalsize}]
@@
 
-\end{tikzpicture}
+\end{tikzpicture}}
 \caption{Visualization of fixed-point iteration and temperature update process for IQS}
@@
 \ee
-Where $IV$ is the inverse velocity operator, $A$ contains all the operations on $\varphi$ from the left-hand-side of \eqt{eq:shape}, and $b$ is the source from the precursors.
+where $IV$ is the inverse velocity operator, $A$ contains all the operations on $\varphi$ from the left-hand-side of \eqt{eq:shape}, and $b$ is the source from the precursors.
 
@@
 \ee
-Where $n$ is the previous time step and $n+1$ is the time step being evaluated. Applying this to \eqt{eq:shape_mat} yields:
+where $n$ is the previous time step and $n+1$ is the time step being evaluated. Applying this to \eqt{eq:shape_mat} yields:
 \be
@@
 \ee
-Where,
+where,
 \be 
@@
 \end{center}
-Where $\lambda\approx0.4358665215$ satisfies $1-9\lambda+18\lambda^2-6\lambda^3=0$.
+where $\lambda\approx0.4358665215$ satisfies $1-9\lambda+18\lambda^2-6\lambda^3=0$.
 
@@
 \ee
-Where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor.
+where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor.
 If the error is larger than the user-specified tolerance, the time step is rejected.
@@
 \centering
+\resizebox{\linewidth}{!}{
 \begin{tikzpicture}[every node/.style = {font=\large},scale=0.75]
@@
 
-\end{tikzpicture}
+\end{tikzpicture}}
 \caption{Visualization of step doubling process with coding logic}
@@
 \ee
-Where $S_f$ is the fission source equivalent for shape ($S_f^n=\sum_{g=1}^G(\nu\Sigma_f^g\varphi^g)^n$). Rearranging to solve for the precursor at the end of the time step yields
+where $S_f$ is the fission source equivalent for shape ($S_f^n=\sum_{g=1}^G(\nu\Sigma_f^g\varphi^g)^n$). Rearranging to solve for the precursor at the end of the time step yields
 \be
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{PRKE coefficients}
+\subsection{PRKE Coefficients}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-05-12 doc/Papers/IQS_convergence/IQS_convergence.tex
title ideas

@@
 \posttitle{\end{center}} % Article title closing formatting
-\title{High-Order Time Discretization and Nonlinear Iteration Analysis of the Improved Quasi-Static Method} % Article title
+\title{High-Order Time Discretization and Nonlinear Iteration Analysis of the Improved Quasi-Static Method\\
+Time-adaptive Multiphysics Reactor Simulations based on the Improved Quasi-Static Method with high-order temporal discretization\\
+Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method\\
+} % Article title
 \author{%

### 2017-08-29 doc/Papers/IQS_convergence/IQS_convergence.tex
chaning to elsevier format

@@
 
-\documentclass[twoside,twocolumn]{article}
-
-\usepackage[english]{babel} % Language hyphenation and typographical rules
+\documentclass{elsarticle}
+
 \usepackage{amsmath}
-\usepackage[T1]{fontenc}
-\usepackage{txfonts}
-\usepackage[text={7in,9in},centering,columnsep=18pt]{geometry} % Document margins
-\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
+\usepackage{bm}
+\usepackage{subcaption}
 \usepackage{booktabs} % Horizontal rules in tables
 \usepackage{enumitem} % Customized lists
-\usepackage{abstract} % Allows abstract customization
-\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
-\usepackage{titling} % Customizing the title section
-\usepackage[hidelinks]{hyperref} % For hyperlinks in the PDF
-\usepackage[numbers,sort&compress]{natbib}[2010/09/13]
-\usepackage{bm}
 \usepackage{graphicx} 
+\usepackage{natbib}%[2010/09/13]
+\usepackage{changepage}
 \usepackage{tikz}
-\usepackage{changepage}
-\usepackage{subcaption}
-
-\setlength{\abovecaptionskip}{0.25\baselineskip}
-\setlength{\belowcaptionskip}{0.25\baselineskip}
-\setlength{\textfloatsep}{0.5\baselineskip}
-\setlength{\intextsep}{0.5\baselineskip}
-\renewcommand{\bottomfraction}{0.9}
-\renewcommand{\topfraction}{0.9}
+
+%%%%%%%%%%%% new commands
 
@@
 
-%----------------------------------------------------------------------------------------
-
-\setlength{\droptitle}{-4\baselineskip} % Move the title up
-
-\pretitle{\begin{center}\Large\bfseries} % Article title formatting
-\posttitle{\end{center}} % Article title closing formatting
+%------------------------------------------------------------------------------
+
+\begin{document}
+
+%------------------------------------------------------------------------------
+\begin{frontmatter}
+
+\journal{Annals of Nuclear Energy}
+
 \title{High-Order Time Discretization and Nonlinear Iteration Analysis of the Improved Quasi-Static Method\\
@@
 } % Article title
-\author{%
-Zachary M. Prince\thanks{Corresponding author, \newline\indent\indent Phone: +1-360-362-6622 \newline\indent\indent email: zachmprince@tamu.edu} \qquad Jean C. Ragusa  \\
-\normalsize Texas A\& M University, College Station, TX \\
-}
-\renewcommand{\maketitlehookd}{%
+
+\author[tamu]{Zachary M. Prince}
+\ead{zachmprince@tamu.edu}
+
+\author[tamu]{Jean C. Ragusa}
+\ead{jean.ragusa@tamu.edu}
+
+\address[tamu]{Texas A\&M University,
+  Department of Nuclear Engineering,
+  College Station, TX 77840, USA}
+
 \begin{abstract}
-\noindent The improved quasi-static method (IQS) is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent amplitude and a time-space-energy dependent shape.
+The improved quasi-static method (IQS) is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent amplitude and a time-space-energy dependent shape.
 The objective of the IQS factorization is to evaluate amplitude and shape on different time scales in order to reduce computational burden associated with solving the multi-dimensional flux equations, while maintaining solution accuracy.
@@
 \end{abstract}
-}
+
+\begin{keyword}
+Reactor dynamics \sep reactor kinetics \sep Improved quasi-static method \sep temporal convergence 
+\end{keyword}
+
+\end{frontmatter}
+%------------------------------------------------------------------------------
 
 %----------------------------------------------------------------------------------------
-
-\begin{document}
-
-\maketitle
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{abbrvnat}
-\bibliography{references_IQS}
+\bibliographystyle{elsarticle-num}
+\bibliography{references_IQS.bib}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-08-29 doc/Papers/IQS_convergence/IQS_convergence.tex
working on theory part of the paper

@@
 %%%%%%%%%%%% new commands
+ \def\todo#1#2{{\color{green}
+     \begin{tabular}{|l|}
+       \hline #1: todo!\\\hline #2 \\\hline
+     \end{tabular}}}
+ \def\question#1#2#3{{\color{red}
+     \begin{tabular}{|l|}
+       \hline Question #1$\to$#2 \\\hline #3 \\\hline
+     \end{tabular}}}
+ \def\answer#1#2{{\color{blue}
+     \begin{tabular}{|l|}
+       \hline{ #1}\\\hline{ #2}\\\hline
+     \end{tabular}}}
+ \def\comm#1{{\color{magenta}
+     \begin{tabular}{|l|}
+       \hline #1 \\\hline
+     \end{tabular}}}
 
@@
 Time-adaptive Multiphysics Reactor Simulations based on the Improved Quasi-Static Method with high-order temporal discretization\\
-Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method\\
+\tcb{Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method}\\
 } % Article title
@@
 \begin{abstract}
-The improved quasi-static method (IQS) is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent amplitude and a time-space-energy dependent shape.
-The objective of the IQS factorization is to evaluate amplitude and shape on different time scales in order to reduce computational burden associated with solving the multi-dimensional flux equations, while maintaining solution accuracy.
-IQS factorization leads to a nonlinear system of equations that requires iteration of shape and amplitude.
-IQS iteration techniques involve fixed-point (Picard) iteration with various convergence criteria and Newton iteration, namely preconditioned Jacobian-free Newton Krylov (PJFNK) method.
-Nonlinear convergence of each of these techniques is investigated.
-Validation of IQS with analysis of time step convergence is also investigated.
-Proper time step convergence is vital for implementation of time adaptive methods and error prediction.
-The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicolson, backward difference formulae (BDF), and singly-diagonally-implicit Runge-Kutta (SDIRK) methods.
+The improved quasi-static method (IQS) is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent amplitude and 
+a time-space-energy dependent shape. The objective of the IQS factorization is to evaluate amplitude and shape on different time scales in order to reduce 
+computational burden associated with solving the multi-dimensional flux equations, while maintaining solution accuracy. IQS factorization leads to a nonlinear 
+system of equations that requires iteration of shape and amplitude. IQS iteration techniques involve fixed-point (Picard) iteration with various convergence criteria 
+and Newton iteration, namely preconditioned Jacobian-free Newton Krylov (PJFNK) method. Nonlinear convergence of each of these techniques is investigated. 
+\tcb{Verification of IQS with analysis of time step convergence} is also investigated. Proper time step convergence is vital for implementation of time adaptive 
+methods and error prediction. The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicholson, backward 
+difference formulae (BDF), and singly-diagonally-implicit Runge-Kutta (SDIRK) methods.
 \end{abstract}
@@
 
+\todo{Zach}{I change the template to that of an elsevier journal but I get a ton of bibtex errors. can you look into that?}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-IQS is a spatial kinetics method that involves factorizing the flux solution into space- and time-dependent components \cite{Ott_1966,Dulla2008,Devooght_1984,Monier_diss,Sissaoui_1995}.
-These components are the flux's amplitude and its shape.
-Amplitude is only time-dependent, while the shape is both space- and time-dependent (as opposed to the PRKE approach where the shape was constant in time).
-However, the impetus of the method is the assumption that the shape is only weakly dependent on time.
-Therefore, the variable for the multi-dimensional diffusion evaluation is expected to be less stiff than the flux itself, without making the assumption that spatial variance is time-independent.
-As a result, the shape may not require an update at the same frequency as the amplitude function, but only on larger macro-time steps.
-
-Due to the factorization of the flux into a shape and an amplitude, the latter two variables are coupled.
-Ott in \cite{Ott_1966} first investigated the coupling of shape and amplitude in a quasi-static nature, but did not include the time derivative in the shape equation.
-Later, in \cite{Ott_1969}, Ott incorporated the time derivative of shape in the equation, yielding better results; this also led to the technique's name: the Improved Quasi-Static method.
-The resulting system of equations is  nonlinear \cite{Dulla2008}.
-Nonlinear problems require an iterative process to evaluate, either fixed-point (Picard) or Newton iteration.
+The improved quasi-static method (IQS) is a numerical technique devised for nuclear reactor transient analysis. It involves factorizing the neutron flux 
+solution into a time-only-dependent component, the amplitude, and a space- and time-dependent component, the shape \cite{Ott_1966,Devooght_1984,Monier_diss,Sissaoui_1995,Dulla2008}. 
+The amplitude solution satisfies the point reactor kinetic equations (PRKE) where the shape solution have been used to generate the PRKE coefficients 
+(reactivity, effective fraction of delayed neutrons, mean generation time). The shape solution satisfies a modified time-dependent neutron balance equation. 
+The rationale for the IQS method lies in the assumption that the shape function is weakly dependent on time. Therefore, it is expected that the modified time-dependent neutron 
+balance equations be less stiff than the original time-dependent neutron balance equations for the flux. As a result, the shape may not require to be solved for at the same frequency 
+as the amplitude, but only on larger macro-time steps, which is expected to yield wall clock savings, especially in multi-dimensional geometries. The PRKE form a small system of ordinary
+differential equations and solving them on a fine temporal grid is not a computational burden. As opposed to the standard PRKE approach, whereby a shape function is selected, typically once for
+an entire transient without updates, the IQS technique is obtained in a rigorously consistent manner from the time-dependent neutron balance equation.
+
+Due to the factorization of the flux into a shape and an amplitude, the latter two variables are nonlinearly coupled. Ott in \cite{Ott_1966} first investigated the coupling of 
+shape and amplitude in a quasi-static nature, but did not include the time derivative in the shape equation. Later, in \cite{Ott_1969}, Ott incorporated the time derivative of 
+shape in the equation, yielding better results but requiring a fixed-point approach to resolve the nonlinearities. This also led to the technique's name: the Improved Quasi-Static method
+although one may argue that such a denomination does not make it immediately clear that the technique fully solves a time-dependent problem.
+
+Nonlinear problems are solved in an iterative manner, typically using either a fixed-point (Picard) approach or Newton's iterations.
 Sissaoui et al.
@@
 \cite{Koclas_1996}, Devooght et al.
-\cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their IQS simulations, the main difference among them is their criteria for convergence.
+\cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their IQS simulations, the main difference among them being their criteria for convergence.
 Devooght et al.
-in \cite{Devooght_1984} also utilizes a Newton iteration technique.
-This paper discusses the different nonlinear iteration techniques and tests the rigor of their implementation.
+in \cite{Devooght_1984} proposed a Newton-based iteration technique.
+This \comm{This?
+Ours?
+his?
+I suppose you mean ours.
+If so, you need to complete the lit review\\ by talking about IQS-PC from Dulla and then their time-adaptive IQS work.\\ After that, start a new paragraph focusing on your paper} paper discusses the different nonlinear iteration techniques and tests the rigor of their implementation.
 IQS can also be linearized by using the IQS Predictor-Corrector method (IQS P-C) \cite{dulla2006}.
@@
 
-IQS has yet to be rigorously tested with high-order discretization of the shape equation.
-All publications that test IQS do not go beyond first-order time discretization.
-Testing IQS with higher-order schemes for proper time step error convergence is essential for implementation of time adaptive methods and quantification of model uncertainty.
+IQS has yet to be rigorously tested with high-order discretization \question{jcr}{zep}{spatial discretization?} of the shape equation.
+All \comm{all?}publications that test IQS do not go beyond first-order time discretization \comm{I thought the Canadians used theta scheme, hence crank nicholson, hence 2nd order ....}.
+Testing IQS with higher-order schemes for proper time step error convergence is essential for implementation of time adaptive methods and quantification of model uncertainty \comm{explain a bit or rephase}.
 This paper discusses implementation of up though fourth-order time discretization for the shape equation of IQS and the flux equation of IQS P-C.
@@
 
-In this Section, we recall the equation for the IQS method, starting from the standard multi-group diffusion equations with delayed neutron precursors in operator form:
-
-\be
-\frac{1}{v^g}\frac{\partial \phi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \phi^{g'} - L^g\phi^g + S_{d}^g
+In this Section, we recall the equations for the IQS method, starting from multi-group neutron conservation statements in operator form:
+
+\begin{subequations}
+\be
+\frac{1}{v^g}\frac{\partial \phi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \phi^{g'} - L^g\phi^g + S_{d}^g \,,
 \label{eq:flux}
@@
 \be
-\frac{dC_i}{dt} = \sum_{g=1}^G P_{d,i}^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
+\frac{dC_i}{dt} = \sum_{g=1}^G P_{d,i}^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I  \,.
 \label{eq:precursor}
 \ee
+\end{subequations}
 %
-where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the diffusion operator, 
-$S_{d}^g$ is the delay source, and $P_{d,i}^g$ is the delayed-neutron fission operator.
-
-The flux factorization approach leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent 
+where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the leakage and interaction operator, 
+$S_{d}^g$ is the delayed neutron source, and $P_{d,i}^g$ is the delayed-neutron fission operator. We assumed $G$ neutron groups and $I$ precursors groups.
+The operator $L^g$ is block diagonal in energy groups: for neutron transport, $L^g = \vec \Omega \cdot \grad + \Sigma_t^g$ whereas for neutron diffusion, 
+$L^g = -\div D^g \grad + \Sigma_r^g$, with  $\Sigma_t^g$ and $\Sigma_r^g$ the total and removal macroscopic cross sections. In the following, we specialize the 
+IQS derivation for the diffusion approximation without loss of generality. Obtaining the IQS formulation for neutron transport is similarly straightforward.
+
+The flux factorization in IQS leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent 
 multigroup shape ($\varphi$):
 \be
-\phi^g(\vec{r},t)=p(t)\varphi^g(\vec{r},t)
-\ee
-After implementing the factorization, the shape diffusion equations result:
-\be
-\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \varphi^{g'} - \left(L^g - \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right)\varphi^g + \boxed{\frac{1}{p}}S_{d}^g
+\phi^g(\vec{r},t)=p(t)\varphi^g(\vec{r},t) \,.
+\ee
+After reporting the above factorization in the balance equations, the shape diffusion equations result:
+\begin{subequations}
+\be
+\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \varphi^{g'} - \left(L^g \tcr{+} \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right)\varphi^g + \boxed{\frac{1}{p}}S_{d}^g
 \label{eq:shape}
 \ee 
+\comm{see the red + above. I think you had a typo}
 \be
@@
 \ee
-
-Note that the time-dependent shape equation is similar to the time-dependent flux equation. 
-However, the shape equations are now non-linearly coupled (boxed terms) to the amplitude equations.
-
-To obtain the amplitude equations, the multigroup shape equations are multiplied by a weighting function, typically the initial adjoint flux ($\phi^{*g}$), and then integrated over the domain.
+\end{subequations}
+
+Note that the time-dependent shape equations are similar to the time-dependent flux equations. One may introduce
+a new block diagonal operator $\tilde{L}^g = L^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ where the reaction term (total/removal cross section)
+simply needs to be augmented by $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$.
+Note that the shape equations are now nonlinearly coupled (boxed terms) to the amplitude equations.
+
+To obtain the amplitude equations, the multigroup shape equations are multiplied by a time-independent weighting function, typically 
+the initial adjoint multigroup flux ($\phi^{*g}$), and then integrated over the phase-space domain.  
 For brevity, the inner product over space will be represented with parenthetical notation ($\left(\phi^{*g},f^g\right) = \int_D \phi^{*g}(\vec{r})f^g(\vec{r})d^3r
-$).
-In order to impose uniqueness of the factorization, one requires $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)$ to be constant.
-After some manipulation, the point reactor kinetics equations (PRKE) for the amplitude solution are obtained:
+$). In order to impose uniqueness of the factorization, one requires that 
+\be
+K(t) = \sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t) \right)
+\ee
+to be constant (hence $K(t)=K_0$).  
+After some manipulation, the point reactor kinetics equations (PRKE) for the amplitude solution are finally obtained:
+\begin{subequations}
 \be
@@
 \ee
+\end{subequations}
 where the functional coefficients are calculated using the space-/time-dependent shape function as follows:
+\begin{subequations}
 \be
@@
 \ee
-
-Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three dimensions, and even more so when using the transport equations in lieu of the diffusion equations.
-Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself, thus allowing for larger time step sizes in updating the shape.
-The PRKE equations form a small ODE system and can be solved using a much smaller time step size.
-In transients where the shape varies much less than the flux, IQS can be very computationally effective.
-The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.
+\end{subequations}
+
+Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three spatial dimensions, and even more so when 
+using the transport equations in lieu of the diffusion equations.  Using IQS, one expects the time dependence of the shape to be 
+weaker than that of the flux itself, thus allowing for larger time step sizes in updating the shape. The PRKE equations form a 
+small ODE system and can be solved using a much smaller time step size. In transients where the shape varies much less than the 
+flux, IQS can be very computationally effective. The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, 
+is illustrated in \fig{fig:iqsviz}.  
+\comm{technically there is an intermediate step for reactivity updates. a reviewer will criticize you easily for not mentioning it}
 
@@
 
-It is important to note that the PRKE parameters are evaluated at each macro step and interpolated for the PRKE evaluation.
-In order to preserve the error convergence of high order discretization schemes for shape, higher order interpolation of the parameters is required.
+It is important to note that the PRKE parameters are evaluated at each macro step and interpolated for the PRKE evaluation. In order to preserve the 
+error convergence rate of high order temporal discretization schemes for shape, higher order interpolation of the parameters will be required.
 
@@
 
-As we noted in the previous section, shape-PRKE equations are a nonlinear system and thus may be solved in a iterative manner.
-Each macro time step can be iterated so the best shape is used to compute power at the micro time steps.
-Sissaoui et al.
-from \cite{Sissaoui_1995}, Koclas et al.
-from \cite{Koclas_1996}, Devooght et al.
-from \cite{Devooght_1984}, and Monier from \cite{Monier_diss} all use iterative techniques for their quasi-static simulations.
-They all undergo a similar process:
+As noted in the previous section, shape-PRKE equations form a nonlinear system and must be solved in a iterative manner.  Over each macro time step,  one can 
+use the latest end-time shape iterate to compute/interpolate the PRKE coefficents over the micro time step intervals. 
+Sissaoui et al. from \cite{Sissaoui_1995}, Koclas et al. from \cite{Koclas_1996}, Devooght et al. from \cite{Devooght_1984}, and Monier from \cite{Monier_diss} all use 
+iterative techniques for their IQS implementations.  They all undergo a similar process:
 \begin{adjustwidth*}{0cm}{4mm}
@@
 
-The major difference between the methods of these authors is the convergence criteria used.
-Sissaoui and Koclas \cite{Sissaoui_1995, Koclas_1996} use fixed point iteration where the criteria is the simply the normalized difference between the last two computed shapes.
-Monier in \cite{Monier_diss} also does fixed point iterations with the same criteria, except the solution is scaled by $\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n})\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n+1})\right)}$ after each iteration.
-Devooght in \cite{Devooght_1984} does a Newton-SOR iteration where the residual of the shape function evaluation is the convergence criteria and next iteration's solution is computed using Newton-Raphson method.
+The major difference between the methods of these authors is the convergence criteria used.  Sissaoui and Koclas \cite{Sissaoui_1995, Koclas_1996} use 
+fixed point iteration where the criteria is the simply the normalized difference between the last two computed shapes.  Monier in \cite{Monier_diss} also 
+employs fixed point iterations with the same criteria, except that the solution is rescaled by 
+$\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n})\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n+1})\right)}$ after each iteration. 
+\comm{can you give a small numerical insight into the scaling in addition to mentioning it; it would make this flow better}
+\comm{also, isn't the scaling simply $K_n/K_{n+1}$?}
+\comm{finally, wasn't there another option where the shape is renormalized after so many iterations and the procedure moves on to the next macro interval?}
+Devooght in \cite{Devooght_1984} uses a Newton-SOR iteration where the residual of the shape function evaluation is the convergence criteria and next iteration's 
+solution is computed using Newton-Raphson method.
 
@@
 Dulla et al.
-in \cite{Dulla2008} does an in depth analysis of the fixed point iteration technique most similar to Sissaoui and Koclas, involving convergence rates and solution results.
+in \cite{Dulla2008} provide an in depth analysis of the fixed point iteration technique most similar to Sissaoui and Koclas, involving convergence rates and solution results.
 However, no comprehensive analysis of iteration techniques exists, comparing both Newton and fixed-point convergence rates.
-The following describes each iteration technique investigated by this research:
-
-\begin{enumerate}
-\item $L^\infty$ norm of shape: $\text{max}\left|\varphi_{n}^{(k+1)} - \varphi_{n}^{(k)}\right|/\text{max}\left|\varphi_{n}^{(k+1)}\right| < \epsilon_{\varphi}$
-\item $L^2$ norm of shape: $\norm{\varphi_{n}^{(k+1)} - \varphi_{n}^{(k)}}/\norm{\varphi_{n}^{(k+1)}} < \epsilon_{\varphi}$
-\item Reactivity convergence: $\left(\rho/\Lambda\right)^{(k+1)} - \left(\rho/\Lambda\right)^{(k)} < \epsilon_{\rho}$
-\item Amplitude convergence: $p_n^{k+1} - p_n^{k} < \epsilon_p$
-\item Uniqueness consistency: $(K_n^{(k+1)} - K_0)/K_0 < \epsilon_K$ where,
-\be
-K_n=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)
-\ee
-\end{enumerate}
+The following describes each iteration convergence criterion investigated in this paper:
+\comm{you never show Newton results, do you?}
+\begin{itemize}
+\item $L^\infty$ norm of shape: 
+\[
+\frac{\text{max}\left|\varphi_{n}^{(k+1)} - \varphi_{n}^{(k)}\right|}{\text{max}\left|\varphi_{n}^{(k+1)}\right|} < \epsilon_{\varphi}
+\]
+\item $L^2$ norm of shape: 
+\[
+\frac{\norm{\varphi_{n}^{(k+1)} - \varphi_{n}^{(k)}}}{\norm{\varphi_{n}^{(k+1)}}} < \epsilon_{\varphi}
+\]
+\item Reactivity convergence: 
+\[
+\left(\rho/\Lambda\right)^{(k+1)} - \left(\rho/\Lambda\right)^{(k)} < \epsilon_{\rho}
+\]
+\item Amplitude convergence: 
+\[
+p_n^{k+1} - p_n^{k} < \epsilon_p
+\]
+\item Uniqueness consistency: 
+\[
+(K_n^{(k+1)} - K_0)/K_0 < \epsilon_K
+\]
+where $k$ denotes the nonlinear iteration index within a given macro time step interval.
+%\be
+%K_n=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)
+%\ee
+\end{itemize}
+\comm{why not $n+1$ in the subscripts? that would make more sense since you use n+1 as the end-time afterwards}
 

### 2017-08-30 doc/Papers/IQS_convergence/IQS_convergence.tex
some update to IQS-PC text

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Theory}
+\section{Background Theory on IQS}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %---------------------------------------------------------------------------------------%
-\subsection{Iteration Methods}
+\subsection{IQS Iterative Schemes}
 \label{sect:iter}
@@
 \]
+\comm{why no time subscripts?}
 \item Amplitude convergence: 
 \[
-p_n^{k+1} - p_n^{k} < \epsilon_p
+p_n^{(k+1)} - p_n^{(k)} < \epsilon_p
 \]
+\comm{why is this not a relative error criterion like the others? what's the rationale?}
 \item Uniqueness consistency: 
 \[
-(K_n^{(k+1)} - K_0)/K_0 < \epsilon_K
+\frac{K_n^{(k+1)} - K_0}{K_0} < \epsilon_K
 \]
@@
 
-In order to preserve the uniqueness criteria, it is beneficial to scale the shape such that the $K_n$ is constant, shown in \eqt{eq:shape_scale}.  
-This scaling can also be done after each iteration, to insure that the uniqueness criteria is satisfied whenever the shape is evaluated.
-\be 
-\varphi^{g}_n = \varphi^{g,(\text{last})}_n \frac{K_0}{K^{(\text{last})}_{n}}
+In order to ensure that uniqueness criteria is preserved, some authors (\comm{which ones?})
+explicitly scale the shape solution, either at each nonlinear iteration or upon exiting the nonlinear loop
+on the macro time time step. This re-scaling is simply
+\be 
+\varphi^{g}_n \leftarrow \varphi^{g,(\text{last})}_n \frac{K_0}{K^{(\text{last})}_{n}} \,.
 \label{eq:shape_scale}
 \ee
-
-%---------------------------------------------------------------------------------------%
-\subsection{IQS Predictor-Corrector}
-%---------------------------------------------------------------------------------------%
-
-The Predictor-Corrector (P-C) version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the solution of the coupled system of equations is different.
-In the IQS P-C version, the flux equations (not the shape equations) are solved (represented by \eqts{eq:flux}{eq:precursor}) in order to obtain a predicted flux solution.
-This predicted flux is then converted to a shape by rescaling it as follows:
-\be
-\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
+\comm{so, how does this jive with situations when the criterion on $K$ is employed?}
+
+%---------------------------------------------------------------------------------------%
+\subsection{IQS Predictor-Corrector Scheme}
+%---------------------------------------------------------------------------------------%
+\comm{make sure the intro has a decent overview -ie no equation, just words- on the IQS-PC. Dulla and Ravetto will likely review this paper}
+
+The Predictor-Corrector version of the IQS method (IQS-PC) also relies on a shape and amplitude factorization of the flux.
+The PRKE derivation is identical to that of the standard (nonlinear) version of the IQS scheme, in the sense that shape solutions
+at the beginning and end times of the macro time step are used (and interpolated in between).
+However, the manner in which the shape function is obtained is different.  In the IQS-PC version, the flux equations (not the shape equations) 
+are first solved (represented by \eqts{eq:flux}{eq:precursor}) in order to obtain a {\it predicted} flux solution at the end of the 
+macro time step. This predicted flux is then converted to a shape by normalizing it as follows:
+\be
+\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}} \,,
 \label{eq:rescale}
@@
 where the scaling factors are given by
-\be
-K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
-\ee
-\be
-K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
-\ee
-
-The PRKE parameters are then computed with this shape using \eqtss{eq:rmb}{eq:l} and interpolated over the macro step, then the PRKE ODE system is solved on the micro time scale.
-With the newly computed amplitude, the shape is rescaled into a flux and the final corrected flux is given by:
+\begin{subequations}
+\be
+K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right) \,,
+\ee
+\be
+K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right) \,.
+\ee
+\end{subequations}
+
+The PRKE parameters are then computed with this shape using \eqtss{eq:rmb}{eq:l} and interpolated over the macro step, then the 
+PRKE ODE system is solved on the micro time scale.  With the newly computed amplitude, the shape is rescaled into a flux and 
+the final {\it corrected} flux is given by:
 \be
@@
 
-The advantage to the predictor-corrector method is there is no iteration necessary for this method and, in turn, is much simpler and faster than the standard IQS.
-Ikeda et al.
-in \cite{Ikeda_2001} and Goluoglu et al.
-in \cite{Goluoglu_2001} both use IQS P-C for complex, three-dimensional problems.
+The advantage to the predictor-corrector method is there is no iteration necessary for this method and, in turn, 
+is much simpler and can be faster than the standard IQS.  Ikeda et al. in \cite{Ikeda_2001} and Goluoglu et al. 
+in \cite{Goluoglu_2001} both use IQS-PC for complex, three-dimensional problems.
 Their results prove IQS P-C to be impressively effective, despite the de-coupling of the system.

### 2017-09-01 doc/Papers/IQS_convergence/IQS_convergence.tex
commented up to the results section

@@
 IQS has yet to be rigorously tested with high-order discretization \question{jcr}{zep}{spatial discretization?} of the shape equation.
-All \comm{all?}publications that test IQS do not go beyond first-order time discretization \comm{I thought the Canadians used theta scheme, hence crank nicholson, hence 2nd order ....}.
+All \comm{all?}publications that test IQS do not go beyond first-order time discretization \comm{I thought the Canadians used theta scheme, hence crank nicholson, hence 2nd order ....; furthermore, I think Hebert also used a generalized RK method, no?}.
 Testing IQS with higher-order schemes for proper time step error convergence is essential for implementation of time adaptive methods and quantification of model uncertainty \comm{explain a bit or rephase}.
@@
 \be
+\label{eq:normalization}
 K(t) = \sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t) \right) 
 \ee
-to be constant (hence $K(t)=K_0$).  
-After some manipulation, the point reactor kinetics equations (PRKE) for the amplitude solution are finally obtained:
+to be constant (hence $K(t)=K(t_0)=K_0$).  
+After some manipulations, the point reactor kinetics equations (PRKE) for the amplitude solution are obtained:
 \begin{subequations}
@@
 %---------------------------------------------------------------------------------------%
-\comm{make sure the intro has a decent overview -ie no equation, just words- on the IQS-PC. Dulla and Ravetto will likely review this paper}
+\comm{make sure the intro has a decent overview -i.e., no equation, just words- on the IQS-PC. Dulla and Ravetto will likely review this paper}
 
@@
 The PRKE derivation is identical to that of the standard (nonlinear) version of the IQS scheme, in the sense that shape solutions
-at the beginning and end times of the macro time step are used (and interpolated in between).
+at the beginning and end times of the macro time interval are used (with interpolation in between).
 However, the manner in which the shape function is obtained is different.  In the IQS-PC version, the flux equations (not the shape equations) 
@@
 \ee
-where the scaling factors are given by
-\begin{subequations}
+where the flux scaling factor is given by
 \be
@@
 \ee
-\be
-K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right) \,.
-\ee
-\end{subequations}
+and $K_0$ is the shape normalization constant (see \eqt{eq:normalization}).
 
@@
 
-The advantage to the predictor-corrector method is there is no iteration necessary for this method and, in turn, 
-is much simpler and can be faster than the standard IQS.  Ikeda et al. in \cite{Ikeda_2001} and Goluoglu et al. 
+The advantage to the predictor-corrector method is that no nonlinear iterations are necessary in this method; it is 
+simpler to implement and can be faster than the standard IQS.  Ikeda et al. in \cite{Ikeda_2001} and Goluoglu et al. 
 in \cite{Goluoglu_2001} both use IQS-PC for complex, three-dimensional problems.
-Their results prove IQS P-C to be impressively effective, despite the de-coupling of the system.
+Their results prove IQS-PC to be effective.
 Dulla et al.
-in \cite{Dulla2008} also describes an in depth comparison of IQS P-C with traditional IQS.
+in \cite{Dulla2008} also describes an in depth comparison of IQS-PC with traditional IQS.
+
+\comm{make sure you have an IQS-PC paragraph in the lit review section as well, not just here}
 
@@
 %---------------------------------------------------------------------------------------%
-
-Two problems that are tested with IQS and IQS P-C involve adiabatic temperature feedback.
-The governing temperature equation is described by \eqt{eq:temp} with cross-section feedback described by \eqt{eq:dopp}.
-
+\comm{have you seen multiphysics simulations with IQS in the literature before? 
+Some of Ott's paper or his old code manual have temperature. so this is something else
+that means to be discussed in the lit review section}
+
+Here, we consider multiphysics reactor physics simulations.  Fuel temperature feedback mechanisms are included 
+using an adiabatic heat conservation model given in \eqt{eq:temp}.  
 \be
@@
 \ee
-
+The thermal-range cross section's dependence on temperature is described by \eqt{eq:dopp}.
 \be
@@
 
-A typical implicit solver would simply use the interpolated flux at end of the temperature time step for the right hand side of the equation, a theta-method discretization.
-However, IQS has much more information about the profile of the flux along the time step because of the micro-step amplitude evaluation.
-Therefore, it is possible to solve for temperature using a semi-analytical approach, shown by \eqt{eq:temp_an}.
-
+A standard time-implicit solver for the heat equation would simply employ the flux values at 
+the extremities of a time step interval.  However, in an IQS neutronic solve, much more information about 
+space/time  distribution of the flux is known: the shape distributions are typically known at the beginning and end
+of a macro time step interval but the amplitude is known on the micro-step time scale, providing
+a richer space/time information for the neutron flux over the macro-step interval.  
+Thus, it is possible to solve for temperature in \eqt{eq:temp} using a semi-analytical approach, 
+shown in \eqt{eq:temp_an}.
+\comm{you keep mixing superscript and subscript for the temporal index. Fix this throughout.}
 \be
@@
 \ee
-
-Where $n$ corresponds to the beginning of the temperature step.
-$a_1$ and $a_2$ are integration coefficients defined by \eqt{eq:a1} and \eqt{eq:a2}.
-Any interpolation of the amplitude along the micro steps is possible for the integration, this application uses piece-wise linear.
-
+where $n$ corresponds to the beginning of the temperature step.  $a_1$ and $a_2$ are integration coefficients 
+given by \eqt{eq:a1} and \eqt{eq:a2}.  The shape function was assumed to vary linearly in time:
+\be
+\varphi(\vec{r},t) = \frac{t_{n}-t}{\Delta t}\varphi^{n+1}(\vec{r}) + \frac{t-t_n}{\Delta t}\varphi^{n+1}(\vec{r})  \,,
+\ee
+which leads to the coefficients' definitions:
+\begin{subequations}
 \be
@@
 \ee
-
-Temperature feedback affects both the shape equation and the reactivity of the PRKE; thus, it is an additional nonlinear component to the already coupled shape-amplitude equations.
-In foresight to the application of this component, temperature is much more time dependent than the shape, but less so than the amplitude.
-Therefore, the evaluation of temperature will have its own time scale.
-A possible solution process for a problem with temperature feedback will have time three time scales portrayed in \fig{fig:time}.
-The first time scale is the shape solve, the second is the temperature evaluation as well as the computation of PRKE parameters, and the third is the PRKE scale.
-It is important to note that the number of time steps in each scale is arbitrary and meant only for visual purposes.
+\end{subequations}
+Because the amplitude $p$ is known on a fine time scale, the integrals \eqt{eq:a1} and \eqt{eq:a2}
+are carried out along the micro steps, using a linear interpolant for the amplitude.
+
+Temperature feedback affects both the shape equation and the reactivity coefficients of the PRKE; thus, 
+it is an additional nonlinear component to the already coupled shape-amplitude equations. In foresight to 
+the application of this component, we expect temperature to be more rapidly varying than the shape, but 
+less so than the amplitude.  Therefore, the evaluation of temperature will have its own time scale, intermediary
+between the amplitude's fine time scale and the shape's coarse time scale.  
+Hence, a possible solution process for a problem with temperature feedback will have three time scales, as
+portrayed in \fig{fig:time}.  The first time scale is the shape solve, the second is the temperature evaluation 
+as well as the computation of PRKE parameters, and the third is the PRKE scale.
+It is important to note that the number of time steps in each scale is arbitrary and the number chosen in \fig{fig:time} are only meant for
+illustrative purposes.
 
@@
 }
-\caption{Time scales and process of IQS with temperature feedback}
+\caption{Time scales and process of IQS solve with temperature feedback}
 \label{fig:time}
@@
 
-Iteration processes are needed for each time scale.
+Iteration processes are needed in between each time scale.
 The amplitude and temperature need to be iterated on the middle time scale until convergence on each temperature step.
@@
 %---------------------------------------------------------------------------------------%
-
-This section presents the time-integration method used to solve coupled flux/shape and precursor equations, represented by Equations~\eqref{eq:flux}/\eqref{eq:shape}~and~\eqref{eq:precursor}/\eqref{eq:prec}.
+\comm{in the results, will you comment on the differences between using the theta scheme versus
+the semi-analytical integration for the precursors?}
+
+\comm{was this section supposed to be moved up?}
+
+%This section presents the time-integration method used to solve coupled flux/shape and precursor equations, represented by Equations~\eqref{eq:flux}/\eqref{eq:shape}~and~\eqref{eq:precursor}/\eqref{eq:prec}.
 First, we note that we could keep this system of two time-dependent equations and solve it as a coupled system.
@@
 
-A common precursor evaluation technique is using the theta method; applying this to \eqt{eq:prec}:
-
-\be
-C^{n+1} = \frac{1-(1-\theta)\Delta t\lambda}{1+\theta\Delta t\lambda}C^n + \frac{(1-\theta)\Delta t \beta}{1+\theta\Delta t\lambda}S_f^n p^n +  \frac{\theta\Delta t \beta}{1+\theta\Delta t\lambda}S_f^{n+1} p^{n+1}
+The precursors' equations, \eqt{eq:prec}, form a system of ODES. % ordinary differential equations.
+A theta-scheme is often employed to evaluate the precursor concentrations:
+\comm{temporal indices: superscripts or subscripts?}
+\be
+C^{n+1} = \frac{1-(1-\theta) \lambda\Delta t}{1+\theta \lambda\Delta t}C^n + \frac{(1-\theta) \beta\Delta t}{1+\theta \lambda\Delta t}S_f^n p^n +  \frac{\theta \beta\Delta t}{1+\theta \lambda\Delta t}S_f^{n+1} p^{n+1}
 \label{eq:dnp_theta}
 \ee
-
-Where $S_f$ is the fission source equivalent for shape ($S_f^n=(\nu\Sigma_f)^n\varphi^n$).
-Another technique to evaluating the precursor equation is to use a exponential operator and integrate the time derivative analytically.
-Applying this operation to \eqt{eq:prec} yields:
-
+where $S_f$ is the fission source computed using the shape solution ($S_f^n=(\nu\Sigma_f)^n\varphi^n$). With
+$\theta=1$, this yields the implicit Euler method, while with $\theta=\tfrac 1 2$, we obtain the Crank 
+Nicolson technique. In doing so, however, no micro-scale temporal information for the amplitude is used.
+One can also solve the precursors' equations analytically, yielding:
 \be
@@
 \ee
-
-Again, this can be applied to implicit discretization by altering the definition of the fission source and eliminating $p$.
-Because $S_f$ is not known continuously over the time step, the integration can be done using any scheme (Riemann, trapezoid, Simpson's, etc.).
-However, there is a very accurate representation of $p(t)$ over the macro step from the PRKE solve.
-In order to utilize this information, another possibility is to interpolate $S_f$ linearly over the macro step.
-Such that:
-
-\be
-S_f(t) = \frac{t_{n+1}-t}{t_{n+1}-t_n}S_f^n  + \frac{t-t_n}{t_{n+1}-t_n}S_f^{n+1}  \quad t_n \le t \le t_{n+1}
-\ee
-Applying this to \eqt{eq:prec_an} yields:
-\be
-C^{n+1} = C^n e^{-\lambda \Delta t} + \left(\hat{a}_2 S_f^{n+1}+\hat{a}_1 S_f^n\right)\beta
+The shape $S_f$ is not known continuously over the time step but we use the beginning/end times values
+and make a linear variation over the macro time step,
+\be
+S_f(t) = \frac{t_{n+1}-t}{\Delta t}S_f^n  + \frac{t-t_n}{\Delta t}S_f^{n+1}  \quad t_n \le t \le t_{n+1} \,.
+\ee
+However, we have a very accurate representation of $p(t)$ over the macro step from the PRKE solve (micro time scale). 
+Using this fine scale information, the analytical solve for the precursor values yields
+\be
+C^{n+1} = C^n e^{-\lambda \Delta t} + \left(\hat{a}_2 S_f^{n+1}+\hat{a}_1 S_f^n\right)\beta \,,
 \label{eq:dnp_an}
 \ee
-
-With integration coefficients defined as:
-
-\begin{align}
-&\hat{a}_1= \int_{t_n}^{t_{n+1}}\frac{t_{n+1}-t'}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt' \\
-&\hat{a}_2 = \int_{t_n}^{t_{n+1}}\frac{t'-t_n}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt'
-\end{align}
+with integration coefficients defined as:
+\begin{subequations}
+\be
+\hat{a}_1= \int_{t_n}^{t_{n+1}}\frac{t_{n+1}-t'}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt' \,,
+\ee
+\be
+\hat{a}_2 = \int_{t_n}^{t_{n+1}}\frac{t'-t_n}{\Delta t}p(t')e^{-\lambda(t_{n+1}-t')}dt'  \,.
+\ee
+\end{subequations}
 
@@
 
-Further enhancements to IQS performance can be gained by using time adaptation (or time step control) in order to increase or reduce the macro time step size for the shape evaluation, depending on error estimates.
-A step-doubling technique is chosen as the time adaptation technique \cite{NumC}.
-The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps ($\varphi^g_{\Delta t/2}$).
-Note: $\varphi$ is changed to $\phi$ for implicit discretization and IQS P-C.
-
+Further enhancements to the performance of the IQS methods can be gained by using time adaptation 
+(or time step control) in order to increase or reduce the macro time step size for the shape evaluation, 
+depending on error estimates. A step-doubling technique is chosen as the time adaptation technique \cite{NumC}. 
+The step doubling technique involves estimating the local error for a certain time step by taking the 
+difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps 
+($\varphi^g_{\Delta t/2}$). Note: $\varphi$ is changed to $\phi$ in the case of the IQS-PC tehcnique.
+%
 The relative error is computed as follows:
 \be
-e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2} - \sum_{g=1}^G\varphi^g_{\Delta t}}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2}},\norm{\sum_{g=1}^G\varphi^g_{\Delta t}}\right)}
+e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2} - \sum_{g=1}^G\varphi^g_{\Delta t}}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2}},\norm{\sum_{g=1}^G\varphi^g_{\Delta t}}\right)} \,.
 \label{eq:edt2}
@@
 \be
-\Delta t_{new} = S \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}}
+\Delta t_{new} = S \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}} \,,
 \label{eq:dt2}
 \ee
-
-Where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor.
-If the error is larger than the user-specified tolerance, the time step is rejected.
-A new time step size is estimated using  \eqt{eq:edt2} as well.
-This process can be visualized by \figs{fig:dt2_1}{fig:dt2_2}.
-Where a step involves a full convergence of shape, amplitude, and any multiphysics on the respective time step.
-
-To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to implicit discretization method, traditional IQS, and IQS P-C.
-Each of these methods will be applied to several diffusion problems; the number of time steps taken and the resulting error will be used to compare the methods.
+where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor. 
+If the error is larger than the user-specified tolerance, the time step is rejected and new (smaller) macro time 
+step size is estimated using \eqt{eq:edt2} as well. This process can be visualized by \figs{fig:dt2_1}{fig:dt2_2},
+where a step involves a full convergence of shape, amplitude, and any multiphysics on the respective time step.
+
+To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to 
+direct temporal discrezation of the flux equation (no IQS), the standard IQS method, and the IQS-PC version. 
+Each of these methods will be applied to several diffusion problems; the number of time steps taken and 
+the resulting error will be used to compare the methods.
 

### 2017-09-04 doc/Papers/IQS_convergence/IQS_convergence.tex
Addressed comments up to results section

@@
 \usepackage{tikz} 
+\usepackage[english]{babel}
 
@@
 
-\title{High-Order Time Discretization and Nonlinear Iteration Analysis of the Improved Quasi-Static Method\\
-Time-adaptive Multiphysics Reactor Simulations based on the Improved Quasi-Static Method with high-order temporal discretization\\
-\tcb{Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method}\\
-} % Article title
+\title{Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method}
 
@@
 \begin{abstract}
-The improved quasi-static method (IQS) is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent amplitude and 
-a time-space-energy dependent shape. The objective of the IQS factorization is to evaluate amplitude and shape on different time scales in order to reduce 
-computational burden associated with solving the multi-dimensional flux equations, while maintaining solution accuracy. IQS factorization leads to a nonlinear 
-system of equations that requires iteration of shape and amplitude. IQS iteration techniques involve fixed-point (Picard) iteration with various convergence criteria 
-and Newton iteration, namely preconditioned Jacobian-free Newton Krylov (PJFNK) method. Nonlinear convergence of each of these techniques is investigated. 
-\tcb{Verification of IQS with analysis of time step convergence} is also investigated. Proper time step convergence is vital for implementation of time adaptive 
-methods and error prediction. The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicholson, backward 
-difference formulae (BDF), and singly-diagonally-implicit Runge-Kutta (SDIRK) methods.
+The improved quasi-static method (IQS) is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent amplitude and a time-space-energy dependent shape.
+The objective of the IQS factorization is to evaluate amplitude and shape on different time scales in order to reduce computational burden associated with solving the multi-dimensional flux equations, while maintaining solution accuracy.
+IQS factorization leads to a nonlinear system of equations that requires iteration of shape and amplitude.
+IQS iteration techniques involve fixed-point (Picard) iteration with various convergence criteria and Newton iteration, namely preconditioned Jacobian-free Newton Krylov (PJFNK) method.
+Nonlinear convergence of each of these techniques is investigated.
+Verification of IQS with analysis of time step convergence is also investigated.
+Proper time step convergence is vital for implementation of time adaptive methods and error prediction.
+The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicholson, backward difference formulae (BDF), and singly-diagonally-implicit Runge-Kutta (SDIRK) methods.
 \end{abstract}
@@
 %----------------------------------------------------------------------------------------
-
-\todo{Zach}{I change the template to that of an elsevier journal but I get a ton of bibtex errors. can you look into that?}
 
@@
 in \cite{Devooght_1984} proposed a Newton-based iteration technique.
-This \comm{This?
-Ours?
-his?
-I suppose you mean ours.
-If so, you need to complete the lit review\\ by talking about IQS-PC from Dulla and then their time-adaptive IQS work.\\ After that, start a new paragraph focusing on your paper} paper discusses the different nonlinear iteration techniques and tests the rigor of their implementation.
-IQS can also be linearized by using the IQS Predictor-Corrector method (IQS P-C) \cite{dulla2006}.
+Another approach is to linearize the equations, so that no nonlinear iteration is necessary.
+IQS can be linearized using the IQS Predictor-Corrector method (IQS P-C) \cite{dulla2006}.
 IQS P-C entails evaluating the flux equation then correcting its amplitude using an amplitude evaluation.
 This method has proven to be effective for problems requiring a significant amount of shape updates when IQS is implemented \cite{Dulla2008}.
-
-IQS has yet to be rigorously tested with high-order discretization \question{jcr}{zep}{spatial discretization?} of the shape equation.
-All \comm{all?}publications that test IQS do not go beyond first-order time discretization \comm{I thought the Canadians used theta scheme, hence crank nicholson, hence 2nd order ....; furthermore, I think Hebert also used a generalized RK method, no?}.
-Testing IQS with higher-order schemes for proper time step error convergence is essential for implementation of time adaptive methods and quantification of model uncertainty \comm{explain a bit or rephase}.
+Dulla also investigates shape time adaptation with IQS with Caron in \cite{caron2017}.
+
+In order for IQS to be a truly useful reactor simulation method, it needs to be effective with multiphysics simulations.
+Meneley and Ott first implemented IQS into the one-dimensional fast-reactor code QX1 \cite{Meneley_1971}.
+This implementation does simple adiabatic heat up with tabular cross-section feedback.
+Later, Kereszt\'{u}ri et al.
+describe IQS implementation into KIKO3D, a three-dimensional pressurized water reactor code, in \cite{KIKO3D_2003}.
+This implementation computes fuel heat transfer and thermal hydraulic feedback.
+Ikeda and Takeda developed in the nodal expansion method code EPISODE which uses IQS P-C with adiabatic fuel heat and Doppler feedback \cite{Ikeda_2001}.
+
+This paper discusses the different nonlinear iteration techniques and tests the rigor of their implementation.
+Additionally, IQS has yet to be rigorously tested with high-order temporal discretization of the shape equation.
+All publications that test IQS do not go beyond first-order time discretization \comm{I thought the Canadians used theta scheme, hence crank nicholson, hence 2nd order ....; furthermore, I think Hebert also used a generalized RK method, no?} \answer{zmp}{the Canadians use the theta scheme for the precursors, but only first order for shape}.
+Testing IQS with higher-order schemes would show the method's effectiveness and stability with more rigorous temporal discretization.
 This paper discusses implementation of up though fourth-order time discretization for the shape equation of IQS and the flux equation of IQS P-C.
-Step doubling time adaptation is also implemented to test IQS and IQS P-C  performance with adaptation of shape/flux evaluation.
-
-To test IQS and IQS P-C nonlinear iteration, time step convergence, and time adaptation performance (five?) problems were applied.
-(Three?) problems are purely neutronics and involve varying magnitudes of reactor size and complexity.
+Step doubling time adaptation is also implemented to test IQS and IQS P-C  performance with adaptation of shape/flux evaluation; although, Caron et al.
+in \cite{caron2017} performs a more comprehensive shape time adaptation analysis.
+Step doubling is also applied to amplitude evaluation and amplitude-based adaptation was applied to reactivity updates.
+In regards to multiphysics simulation, this paper continues the analysis of IQS with adiabatic heat up and cross-section feedback.
+We develop a semi-analytical approach to evaluating fuel temperature and implement an intermediate time scale for temperature evaluation.
+
+To test IQS and IQS P-C nonlinear iteration, time step convergence, and time adaptation performance four problems were applied.
+Two problems are purely neutronics and involve varying magnitudes of reactor size and complexity.
 Two problems involve temperature feedback and test an intermediate time scale for temperature evaluation.
@@
 \be
-\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \varphi^{g'} - \left(L^g \tcr{+} \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right)\varphi^g + \boxed{\frac{1}{p}}S_{d}^g
+\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \varphi^{g'} - \left(L^g + \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right)\varphi^g + \boxed{\frac{1}{p}}S_{d}^g
 \label{eq:shape}
 \ee
-\comm{see the red + above. I think you had a typo}
+
 \be
@@
 
-Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three spatial dimensions, and even more so when 
-using the transport equations in lieu of the diffusion equations.  Using IQS, one expects the time dependence of the shape to be 
-weaker than that of the flux itself, thus allowing for larger time step sizes in updating the shape. The PRKE equations form a 
-small ODE system and can be solved using a much smaller time step size. In transients where the shape varies much less than the 
-flux, IQS can be very computationally effective. The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, 
-is illustrated in \fig{fig:iqsviz}.  
+Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three spatial dimensions, and even more so when using the transport equations in lieu of the diffusion equations.
+Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself, thus allowing for larger time step sizes in updating the shape.
+The PRKE equations form a small ODE system and can be solved using a much smaller time step size.
+In transients where the shape varies much less than the flux, IQS can be very computationally effective.
+%The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.  
 \comm{technically there is an intermediate step for reactivity updates. a reviewer will criticize you easily for not mentioning it}
 
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{figures/IQS_visualization.jpg}
-\caption{IQS method visualization}
-\label{fig:iqsviz}
-\end{figure}
-
-It is important to note that the PRKE parameters are evaluated at each macro step and interpolated for the PRKE evaluation. In order to preserve the 
-error convergence rate of high order temporal discretization schemes for shape, higher order interpolation of the parameters will be required.
+%\begin{figure}[!htbp]
+%\centering
+%\includegraphics[width=\linewidth]{figures/IQS_visualization.jpg}
+%\caption{IQS method visualization}
+%\label{fig:iqsviz}
+%\end{figure}
+
+It is important to note that the PRKE parameters are evaluated at each macro step and interpolated for the PRKE evaluation.
+In order to preserve the error convergence rate of high order temporal discretization schemes for shape, higher order interpolation of the parameters will be required.
 
@@
 
-The major difference between the methods of these authors is the convergence criteria used.  Sissaoui and Koclas \cite{Sissaoui_1995, Koclas_1996} use 
-fixed point iteration where the criteria is the simply the normalized difference between the last two computed shapes.  Monier in \cite{Monier_diss} also 
-employs fixed point iterations with the same criteria, except that the solution is rescaled by 
-$\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n})\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g(t_{n+1})\right)}$ after each iteration. 
-\comm{can you give a small numerical insight into the scaling in addition to mentioning it; it would make this flow better}
-\comm{also, isn't the scaling simply $K_n/K_{n+1}$?}
-\comm{finally, wasn't there another option where the shape is renormalized after so many iterations and the procedure moves on to the next macro interval?}
+The major difference between the methods of these authors is the convergence criteria used.
+Sissaoui and Koclas \cite{Sissaoui_1995, Koclas_1996} use fixed point iteration where the criteria is the simply the normalized difference between the last two computed shapes.
+Monier in \cite{Monier_diss} also employs fixed point iterations with the same criteria, except that the solution is rescaled by $K_n/K_{n+1}$ after each iteration.
+Dulla in \cite{Dulla2008} does the same fixed-point iteration, but the shape is scaled by $K_0/K_{n+1}$ at the end of each macro time step.
+Renormalizing the shape is essential for preserving the uniqueness condition ($K_{n+1} = K_0$); this condition is not inherently conserved, even when the solution has converged from shape iterations.
 Devooght in \cite{Devooght_1984} uses a Newton-SOR iteration where the residual of the shape function evaluation is the convergence criteria and next iteration's
@@
 \comm{you never show Newton results, do you?}
+\answer{zmp}{no, never figured out how to do it, plus Devooght actually uses GQS for his results.}
 \begin{itemize}
-\item $L^\infty$ norm of shape: 
+\item $L^\infty$ norm of shape \cite{Monier_diss}: 
 \[
-\frac{\text{max}\left|\varphi_{n}^{(k+1)} - \varphi_{n}^{(k)}\right|}{\text{max}\left|\varphi_{n}^{(k+1)}\right|} < \epsilon_{\varphi}
+\frac{\text{max}\left|\varphi_{n+1}^{(k+1)} - \varphi_{n+1}^{(k)}\right|}{\text{max}\left|\varphi_{n+1}^{(k+1)}\right|} < \epsilon_{\varphi}
 \]
@@
 \[
-\frac{\norm{\varphi_{n}^{(k+1)} - \varphi_{n}^{(k)}}}{\norm{\varphi_{n}^{(k+1)}}} < \epsilon_{\varphi}
+\frac{\norm{\varphi_{n+1}^{(k+1)} - \varphi_{n+1}^{(k)}}}{\norm{\varphi_{n+1}^{(k+1)}}} < \epsilon_{\varphi}
 \]
-\item Reactivity convergence: 
+\item Reactivity convergence \cite{Monier_diss}: 
 \[
-\left(\rho/\Lambda\right)^{(k+1)} - \left(\rho/\Lambda\right)^{(k)} < \epsilon_{\rho}
+\left(\rho/\Lambda\right)^{(k+1)}_{n+1} - \left(\rho/\Lambda\right)^{(k)}_{n+1} < \epsilon_{\rho}
 \]
-\comm{why no time subscripts?}
-\item Amplitude convergence: 
+\item Amplitude convergence \cite{Monier_diss}: 
 \[
-p_n^{(k+1)} - p_n^{(k)} < \epsilon_p
+p_{n+1}^{(k+1)} - p_{n+1}^{(k)} < \epsilon_p
 \]
 \comm{why is this not a relative error criterion like the others? what's the rationale?}
-\item Uniqueness consistency: 
+\answer{zmp}{I copied all of the above from \cite{Monier_diss}, not sure the rationale though}
+\item Uniqueness consistency \cite{Monier_diss}: 
 \[
-\frac{K_n^{(k+1)} - K_0}{K_0} < \epsilon_K
+\frac{K_{n+1}^{(k+1)} - K_0}{K_0} < \epsilon_K
 \]
 where $k$ denotes the nonlinear iteration index within a given macro time step interval.
-%\be
-%K_n=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)
-%\ee
 \end{itemize}
-\comm{why not $n+1$ in the subscripts? that would make more sense since you use n+1 as the end-time afterwards}
-
-In order to ensure that uniqueness criteria is preserved, some authors (\comm{which ones?})
-explicitly scale the shape solution, either at each nonlinear iteration or upon exiting the nonlinear loop
-on the macro time time step. This re-scaling is simply
+
+In order to ensure that uniqueness criteria is preserved, some authors (\cite{Monier_diss,Dulla2008}) explicitly scale the shape solution, either at each nonlinear iteration or upon exiting the nonlinear loop on the macro time time step.
+This re-scaling is simply
 \be 
@@
 \comm{so, how does this jive with situations when the criterion on $K$ is employed?}
+\answer{zmp}{if the criteria is satisfied, it doesn't rescale; if it isn't, the solution is rescaled and another iteration is performed, the resulting solution isn't guaranteed to satisfy the criteria}
 
@@
 
-The Predictor-Corrector version of the IQS method (IQS-PC) also relies on a shape and amplitude factorization of the flux.
-The PRKE derivation is identical to that of the standard (nonlinear) version of the IQS scheme, in the sense that shape solutions
-at the beginning and end times of the macro time interval are used (with interpolation in between).
-However, the manner in which the shape function is obtained is different.  In the IQS-PC version, the flux equations (not the shape equations) 
-are first solved (represented by \eqts{eq:flux}{eq:precursor}) in order to obtain a {\it predicted} flux solution at the end of the 
-macro time step. This predicted flux is then converted to a shape by normalizing it as follows:
+The Predictor-Corrector version of the IQS method (IQS-PC) is a linearized version of IQS, where no nonlinear iterations are required.
+The method relies on the shape and amplitude factorization of the flux and uniqueness condition.
+The PRKE derivation is identical to that of the standard (nonlinear) version of the IQS scheme, in the sense that shape solutions at the beginning and end times of the macro time interval are used (with interpolation in between).
+However, the manner in which the shape function is obtained is different.
+In the IQS-PC version, the flux equations (not the shape equations) are first solved (represented by \eqts{eq:flux}{eq:precursor}) in order to obtain a {\it predicted} flux solution at the end of the macro time step.
+This predicted flux is then converted to a shape by normalizing it as follows:
 \be
@@
 The PRKE parameters are then computed with this shape using \eqtss{eq:rmb}{eq:l} and interpolated over the macro step, then the 
-PRKE ODE system is solved on the micro time scale.  With the newly computed amplitude, the shape is rescaled into a flux and 
-the final {\it corrected} flux is given by:
+PRKE ODE system is solved on the micro time scale.  With the newly computed amplitude, the shape is rescaled into a flux and the final {\it corrected} flux is given by:
 \be
@@
 
-The advantage to the predictor-corrector method is that no nonlinear iterations are necessary in this method; it is 
-simpler to implement and can be faster than the standard IQS.  Ikeda et al. in \cite{Ikeda_2001} and Goluoglu et al. 
+The advantage to the predictor-corrector method is that no nonlinear iterations are necessary in this method; it is simpler to implement and can be faster than the standard IQS.
+Ikeda et al.
+in \cite{Ikeda_2001} and Goluoglu et al.
 in \cite{Goluoglu_2001} both use IQS-PC for complex, three-dimensional problems.
@@
 Dulla et al.
-in \cite{Dulla2008} also describes an in depth comparison of IQS-PC with traditional IQS.
-
-\comm{make sure you have an IQS-PC paragraph in the lit review section as well, not just here}
+in \cite{Dulla2008} also describes an in depth comparison of IQS-PC with traditional IQS, which originally brought the method to light.
 
@@
 %---------------------------------------------------------------------------------------%
-\comm{have you seen multiphysics simulations with IQS in the literature before? 
-Some of Ott's paper or his old code manual have temperature. so this is something else
-that means to be discussed in the lit review section}
 
@@
 shown in \eqt{eq:temp_an}.
-\comm{you keep mixing superscript and subscript for the temporal index. Fix this throughout.}
-\be
-T^{n+1} = T^n + \frac{\kappa_f}{\rho c_p} \left(a_2 \varphi^{n+1} + a_1 \varphi^{n}\right)
+
+\be
+T_{n+1} = T_n + \frac{\kappa_f}{\rho c_p} \left(a_2 \varphi_{n+1} + a_1 \varphi_{n}\right)
 \label{eq:temp_an}
@@
 \be
-\varphi(\vec{r},t) = \frac{t_{n}-t}{\Delta t}\varphi^{n+1}(\vec{r}) + \frac{t-t_n}{\Delta t}\varphi^{n+1}(\vec{r})  \,,
+\varphi(\vec{r},t) = \frac{t_{n+1}-t}{\Delta t}\varphi_{n}(\vec{r}) + \frac{t-t_n}{\Delta t}\varphi_{n+1}(\vec{r})  \,,
 \ee
@@
 the semi-analytical integration for the precursors?}
+\answer{zmp}{yes I do}
 
 \comm{was this section supposed to be moved up?}
+\answer{zmp}{I think we decided to put it after temperature, to make multiphysics seem more important}
 
@@
 A theta-scheme is often employed to evaluate the precursor concentrations:
-\comm{temporal indices: superscripts or subscripts?}
-\be
-C^{n+1} = \frac{1-(1-\theta) \lambda\Delta t}{1+\theta \lambda\Delta t}C^n + \frac{(1-\theta) \beta\Delta t}{1+\theta \lambda\Delta t}S_f^n p^n +  \frac{\theta \beta\Delta t}{1+\theta \lambda\Delta t}S_f^{n+1} p^{n+1}
+
+\be
+C_{n+1} = \frac{1-(1-\theta) \lambda\Delta t}{1+\theta \lambda\Delta t}C_n + \frac{(1-\theta) \beta\Delta t}{1+\theta \lambda\Delta t}S_{f,n} p_n +  \frac{\theta \beta\Delta t}{1+\theta \lambda\Delta t}S_{f,n+1} p_{n+1}
 \label{eq:dnp_theta}
 \ee
-where $S_f$ is the fission source computed using the shape solution ($S_f^n=(\nu\Sigma_f)^n\varphi^n$). With
+where $S_f$ is the fission source computed using the shape solution ($S_{f,n}=(\nu\Sigma_f)_n\varphi_n$). With
 $\theta=1$, this yields the implicit Euler method, while with $\theta=\tfrac 1 2$, we obtain the Crank 
@@
 \be
-C^{n+1} =  C^n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
+C_{n+1} =  C_n e^{-\lambda (t_{n+1} - t_n) }  + \int_{t_n}^{t_{n+1}} \beta(t') S_f(t') p(t')e^{-\lambda (t_{n+1}-t')}dt'
 \label{eq:prec_an}
@@
 \be
-S_f(t) = \frac{t_{n+1}-t}{\Delta t}S_f^n  + \frac{t-t_n}{\Delta t}S_f^{n+1}  \quad t_n \le t \le t_{n+1} \,.
+S_f(t) = \frac{t_{n+1}-t}{\Delta t}S_{f,n}  + \frac{t-t_n}{\Delta t}S_{f,n+1}  \quad t_n \le t \le t_{n+1} \,.
 \ee
@@
 \be
-C^{n+1} = C^n e^{-\lambda \Delta t} + \left(\hat{a}_2 S_f^{n+1}+\hat{a}_1 S_f^n\right)\beta \,,
+C_{n+1} = C_n e^{-\lambda \Delta t} + \left(\hat{a}_2 S_{f,n+1}+\hat{a}_1 S_{f,n}\right)\beta \,,
 \label{eq:dnp_an}
@@
 \label{tab:TWIGLdt2}
-%\resizebox{\linewidth}{!}{
+\resizebox{\linewidth}{!}{
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
@@
 
-\end{tabular}%}
+\end{tabular}}
 \end{center}
@@
 \bibliographystyle{elsarticle-num}
-\bibliography{references_IQS.bib}
+\bibliography{references_IQS}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-09-04 doc/Papers/IQS_convergence/IQS_convergence.tex
Adding more results

@@
 \centering
-\includegraphics[width=\linewidth]{figures/1D_flux.png}
+\includegraphics[height=3in]{figures/1D_flux.png}
 \caption{Flux profile at various times}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/1D_power_base.png}
+\includegraphics[height=3in]{figures/1D_power_base.png}
 \caption{Power profile over transient}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/1D_shape2.png}
+\includegraphics[height=3in]{figures/1D_shape2.png}
 \caption{Shape profile at various times}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/1D_iqs_power.png}
+\includegraphics[height=3in]{figures/1D_iqs_power.png}
 \caption{Power and reactivity profile over transient}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/iter_renorm.png}
+\includegraphics[height=3in]{figures/iter_renorm.png}
 \caption{\# of iterations for various convergence criteria, tolerance$=10^{-11}$, max iterations$=20$}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/iter_error.png}
+\includegraphics[height=3in]{figures/iter_error.png}
 \caption{Final iteration error for K convergence criteria}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/iter_error_an.png}
+\includegraphics[height=3in]{figures/iter_error_an.png}
 \caption{Final iteration error for K convergence criteria with analytical precursor elimination}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/1Dmini_flux.png}
+\includegraphics[height=3in]{figures/1Dmini_flux.png}
 \caption{Flux profile at various times}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/1Dmini_power_base.png}
+\includegraphics[height=3in]{figures/1Dmini_power_base.png}
 \caption{Power profile over transient}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/1Dmini_shape.png}
+\includegraphics[height=3in]{figures/1Dmini_shape.png}
 \caption{Shape profile at various times for one-dimensional mini-core}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/1Dmini_conv_IE.png}
+\includegraphics[height=3in]{figures/1Dmini_conv_IE.png}
 \caption{Time step convergence for one-dimensional mini-core with implicit Euler discretization}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/TWIGL_convergence.jpg}
+\includegraphics[height=3in]{figures/TWIGL_convergence.jpg}
 \caption{Error convergence comparison of TWIGL Benchmark}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/TWIGL_power_plot_dt2.png}
+\includegraphics[height=3in]{figures/TWIGL_power_plot_dt2.png}
 \caption{Power level comparison of TWIGL Benchmark with time adaptation}
@@
 \end{figure}
-
-%---------------------------------------------------------------------------------------%
-\subsection{LMW Benchmark?}
-%---------------------------------------------------------------------------------------%
 
@@
 \label{tab:base}
-\resizebox{\linewidth}{!}{
+%\resizebox{\linewidth}{!}{
 \begin{tabular}{|l|cc|}
@@
 \end{tabular}
-}
+%}
 \end{center}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/lra_mp.png}
+\includegraphics[height=3in]{figures/lra_mp.png}
 \caption{Error plot with various temperature updates per macro step}
@@
 IQS shows an error comparable to implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by -34.1\%.
-IQS P-C shows an error less than implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by <-34.9\%.
+IQS P-C shows an error less than implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by $<-34.9\%$.
 
@@
 
+Transient 15 is a test case based on the TREAT core.
+The purpose of the original creation of this simulation in Rattlesnake is to test the model's fidelity with the thermal feedback of TREAT, but it is not meant to exactly match any previous experiments.
+Nevertheless, the goal of the following simulations is to test IQS and its time scale based treatment of temperature with a more complex model.
+Transient 15 involves an 11-energy group diffusion approximation and is discretized into $355,712$ hexahedral continuous finite elements totaling $4,109,523$ degrees of freedom.
+The three-second transient involves a linear ramp decrease in the absorption cross section throughout the control rod region.
+\fig{fig:Tran15} shows a visualization of the flux profile within the core, hidden is the massive amount of graphite surrounding the core.
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[height=3in]{figures/Tran15_core2.png}
+\caption{Transient 15 core power profile at peak power}
+\label{fig:Tran15}
+\end{figure}
+
+The Transient-15 model uses a adiabatic temperature feedback mechanism, similar to the one explored by the LRA.
+\eqt{eq:temp2} describes the heat up of the fuel.
+It is very similar, except the specific heat is now dependent on temperature is described by \eqt{eq:cp}.
+The temperature evaluation is identical to the one described in LRA section, except a Newton iteration process is employed to resolve the nonlinearity from the specific heat term.
+The feedback to the cross-sections are applied using linear interpolation of tabular data provided by INL.
+
+\be
+\rho c_p(T) \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
+\label{eq:temp2}
+\ee
+
+\be
+c_p = -5.8219e-10T^3 - 4.3694e-7T^2 + 2.8369e-3T -1.009e-2
+\label{eq:cp}
+\ee
+
 %----------------------------------------------------------------------------------------
@@
 
-%----------------------------------------------------------------------------------------
-\subsubsection{Transient-15 Time Adaptation}
+In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time step, implicit discretization, IQS with one and 5 temperature updates per macro step, and IQS P-C with one and 5 updates.
+\fig{fig:Tran15_profile} shows the baseline power and temperature profile for the Transient-15 example.
+\tbl{tab:tran15} shows the error and runtime results.
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[height=3in]{figures/Tran15_profile.png}
+\caption{Transient-15 total power and average temperature profile during transient}
+\label{fig:Tran15_profile}
+\end{figure}
+
+\begin{table}[htb]
+\begin{center}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{|l|ccccccc|}
+\hline
+ & No.	& Max &	Time at Max &	Max Average &	\% Increase &	Max Power & Linear	\\
+Method & of Steps &	Power (W) &	Power (s) &	Temperature (K) & Runtime$^*$ & Error & Iterations \\
+\hline
+Baseline 			& 3000 	& 3.5039e+08 & 1.901 & 371 &	---		&	---		 & ---		\\
+Implicit Dis. 		& 300 	& 3.5011e+08 & 1.90  & 371 &	---		&	7.875e-4 & 41020	\\
+IQS 				& 300	& 3.5036e+08 & 1.90  & 371 & -11.9\%	&	8.385e-5 & 23949	\\
+IQS (5 updates) 	& 300 	& 3.5040e+08 & 1.90  & 371 &  49.7\%	&	3.687e-5 & 24035	\\
+IQS P-C 			& 300 	& 3.5065e+08 & 1.90  & 371 &  -2.1\%	&	7.527e-4 & 39020	\\
+IQS P-C (5 updates) & 300 	& 3.5043e+08 & 1.90  & 371 &  26.5\%	&	1.227e-4 & 37866	\\
+\hline
+\end{tabular}
+}
+\end{center}
+\vspace{-3mm}
+$^*$ difference in runtime from implicit discretization 
+\caption{Transient-15 Error and Runtime Results}
+\label{tab:tran15}
+\end{table}
+
+The results from \tbl{tab:tran15} show similar performance of IQS with the temperature updates as the LRA.
+Again, the number of linear GMRES iterations is shown as a measure of computational expense.
+However, these iterations do not consider the temperature updates, so the iterations of the simulations with multiple updates should be taken with a grain of salt.
+IQS with 1 temperature update shows a performance that reduces the error to approximately a tenth of the implicit discretization error, and reduces the execution time by about 12\%.
+This shows that IQS was able to resolve the nonlinearity between flux and temperature with significantly fewer diffusion evaluations.
+Having IQS with 5 updates significantly increased the execution time for the same time step, but the error was reduced.
+Comparing this error to a similar implicit discretization error at a smaller time step could show that the runtime was reduced.
+IQS P-C performed not nearly as well as it did with the LRA benchmark, but still proved to be effective.
+Having 5 updates for IQS P-C increased the runtime marginally, but decreased the error significantly.
 

### 2017-09-07 doc/Papers/IQS_convergence/IQS_convergence.tex
second review to up results

@@
 \usepackage[english]{babel}
+\usepackage{xspace}
 
@@
 \newcommand{\keff}{k_\textit{eff}}
+\newcommand{\iqspc}{IQS-PC\xspace}
 
@@
 \newcommand{\tcg}[1]{\textcolor{britishracinggreen}{#1}}
+\newcommand{\jcr}[1]{\textcolor{red}{jcr: #1}}
 
@@
 \begin{abstract}
-The improved quasi-static method (IQS) is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent amplitude and a time-space-energy dependent shape.
-The objective of the IQS factorization is to evaluate amplitude and shape on different time scales in order to reduce computational burden associated with solving the multi-dimensional flux equations, while maintaining solution accuracy.
-IQS factorization leads to a nonlinear system of equations that requires iteration of shape and amplitude.
-IQS iteration techniques involve fixed-point (Picard) iteration with various convergence criteria and Newton iteration, namely preconditioned Jacobian-free Newton Krylov (PJFNK) method.
-Nonlinear convergence of each of these techniques is investigated.
-Verification of IQS with analysis of time step convergence is also investigated.
-Proper time step convergence is vital for implementation of time adaptive methods and error prediction.
-The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicholson, backward difference formulae (BDF), and singly-diagonally-implicit Runge-Kutta (SDIRK) methods.
+The improved quasi-static method (IQS) is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent 
+amplitude and a time-space-energy dependent shape. The objective of the IQS factorization is to evaluate amplitude and shape on different time scales 
+in order to reduce the computational burden associated with solving the multi-dimensional flux equations, while maintaining solution accuracy. 
+The IQS decomposition leads to a nonlinear system of equations that requires iteration of shape and amplitude. IQS iteration techniques involve 
+fixed-point (Picard) iteration with various convergence criteria and Newton iteration, namely preconditioned Jacobian-free Newton Krylov (PJFNK) method. 
+\jcr{space is at a premium in the abstract. why talk about jfnk if no such results are presented? i must be missing sthg}
+Nonlinear convergence of each \jcr{each?} of these techniques is investigated. Verification of IQS with analysis of time step convergence is also investigated. 
+Proper time step convergence is vital for implementation of time adaptive methods and error prediction. The time derivative of the shape function is 
+discretized through fourth order using implicit-Euler, Crank-Nicholson, backward difference formulae (BDF), and singly-diagonally-implicit 
+Runge-Kutta (SDIRK) methods.
 \end{abstract}
@@
 as the amplitude, but only on larger macro-time steps, which is expected to yield wall clock savings, especially in multi-dimensional geometries. The PRKE form a small system of ordinary
-differential equations and solving them on a fine temporal grid is not a computational burden. As opposed to the standard PRKE approach, whereby a shape function is selected, typically once for
-an entire transient without updates, the IQS technique is obtained in a rigorously consistent manner from the time-dependent neutron balance equation.
+differential equations (ODE) and solving them on a fine temporal grid is not a computational burden. As opposed to the standard PRKE approach, whereby a 
+shape function is selected, typically once for an entire transient without updates, the IQS technique is obtained in a rigorously consistent manner from the 
+time-dependent neutron balance equation.
 
@@
 shape and amplitude in a quasi-static nature, but did not include the time derivative in the shape equation. Later, in \cite{Ott_1969}, Ott incorporated the time derivative of 
-shape in the equation, yielding better results but requiring a fixed-point approach to resolve the nonlinearities. This also led to the technique's name: the Improved Quasi-Static method
-although one may argue that such a denomination does not make it immediately clear that the technique fully solves a time-dependent problem.
-
-Nonlinear problems are solved in an iterative manner, typically using either a fixed-point (Picard) approach or Newton's iterations.
-Sissaoui et al.
-\cite{Sissaoui_1995}, Koclas et al.
-\cite{Koclas_1996}, Devooght et al.
-\cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their IQS simulations, the main difference among them being their criteria for convergence.
-Devooght et al.
-in \cite{Devooght_1984} proposed a Newton-based iteration technique.
-Another approach is to linearize the equations, so that no nonlinear iteration is necessary.
-IQS can be linearized using the IQS Predictor-Corrector method (IQS P-C) \cite{dulla2006}.
-IQS P-C entails evaluating the flux equation then correcting its amplitude using an amplitude evaluation.
-This method has proven to be effective for problems requiring a significant amount of shape updates when IQS is implemented \cite{Dulla2008}.
-Dulla also investigates shape time adaptation with IQS with Caron in \cite{caron2017}.
-
-In order for IQS to be a truly useful reactor simulation method, it needs to be effective with multiphysics simulations.
-Meneley and Ott first implemented IQS into the one-dimensional fast-reactor code QX1 \cite{Meneley_1971}.
-This implementation does simple adiabatic heat up with tabular cross-section feedback.
-Later, Kereszt\'{u}ri et al.
-describe IQS implementation into KIKO3D, a three-dimensional pressurized water reactor code, in \cite{KIKO3D_2003}.
-This implementation computes fuel heat transfer and thermal hydraulic feedback.
-Ikeda and Takeda developed in the nodal expansion method code EPISODE which uses IQS P-C with adiabatic fuel heat and Doppler feedback \cite{Ikeda_2001}.
-
-This paper discusses the different nonlinear iteration techniques and tests the rigor of their implementation.
-Additionally, IQS has yet to be rigorously tested with high-order temporal discretization of the shape equation.
-All publications that test IQS do not go beyond first-order time discretization \comm{I thought the Canadians used theta scheme, hence crank nicholson, hence 2nd order ....; furthermore, I think Hebert also used a generalized RK method, no?} \answer{zmp}{the Canadians use the theta scheme for the precursors, but only first order for shape}.
-Testing IQS with higher-order schemes would show the method's effectiveness and stability with more rigorous temporal discretization.
-This paper discusses implementation of up though fourth-order time discretization for the shape equation of IQS and the flux equation of IQS P-C.
-Step doubling time adaptation is also implemented to test IQS and IQS P-C  performance with adaptation of shape/flux evaluation; although, Caron et al.
+shape in the equation, yielding better results but requiring a fixed-point approach to resolve the nonlinear coupling between the amplitude and shape equations. This also 
+led to the technique's name: the Improved Quasi-Static method although one may argue that such a denomination does not make it immediately clear that the technique rigorously 
+solves a time-dependent problem.
+
+Nonlinear problems are solved in an iterative manner, typically using either a fixed-point (Picard) approach or Newton's iterations. Sissaoui et al. \cite{Sissaoui_1995}, 
+Koclas et al. \cite{Koclas_1996}, Devooght et al. \cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their IQS simulations, 
+the main difference among them being their criteria for convergence.  Devooght et al. in \cite{Devooght_1984} proposed a Newton-based iteration technique. Another approach 
+is to linearize the equations, so that no nonlinear iteration is necessary. IQS can be linearized using the IQS Predictor-Corrector method (\iqspc) \cite{dulla2006}. 
+\iqspc entails evaluating the flux equation then correcting its amplitude using an amplitude evaluation. This method has proven to be effective for problems 
+requiring a significant amount of shape updates when IQS is implemented \cite{Dulla2008}. Dulla also investigates shape time adaptation with IQS with Caron in \cite{caron2017}.
+
+For use in multiphysics reactor physics applications, the interplay of the IQS solution technique (neutronics) and other physics components (e.g., fuel temperature) needs to be 
+resolved effectively. Meneley and Ott first implemented IQS into the one-dimensional fast-reactor code QX1 \cite{Meneley_1971}. This implementation does simple adiabatic heat up 
+with tabular cross-section feedback. Later, Kereszt\'{u}ri et al. describe IQS implementation into KIKO3D, a three-dimensional pressurized water reactor code, in \cite{KIKO3D_2003}. 
+This implementation computes fuel heat transfer and thermal hydraulic feedback. Ikeda and Takeda developed in the nodal expansion method code EPISODE which uses \iqspc with adiabatic 
+fuel heat and Doppler feedback \cite{Ikeda_2001}. \jcr{here you just say who did IQS multiphysics before. but how did they do it? are there differences between these implementations,
+is there something lacking that your paper will provide insight to?}
+
+This paper (i) discusses the different nonlinear iteration techniques for IQS and tests the rigor of their implementation, (ii) investigates high-order temporal discretization of the shape equation,
+and (iii) analyzes IQS coupled with adiabatic heat up and cross-section feedback. 
+%
+Prior IQS publications do not go beyond first-order time discretization for the shape equation. Here, we test IQS with higher-order schemes (up to fourth order) and investigate the method's 
+effectiveness with such temporal discretizations. 
+Step doubling time adaptation is also implemented to test IQS and \iqspc  performance with adaptation of shape/flux evaluation; although, Caron et al.
 in \cite{caron2017} performs a more comprehensive shape time adaptation analysis.
 Step doubling is also applied to amplitude evaluation and amplitude-based adaptation was applied to reactivity updates.
-In regards to multiphysics simulation, this paper continues the analysis of IQS with adiabatic heat up and cross-section feedback.
-We develop a semi-analytical approach to evaluating fuel temperature and implement an intermediate time scale for temperature evaluation.
-
-To test IQS and IQS P-C nonlinear iteration, time step convergence, and time adaptation performance four problems were applied.
-Two problems are purely neutronics and involve varying magnitudes of reactor size and complexity.
-Two problems involve temperature feedback and test an intermediate time scale for temperature evaluation.
+In regards to multiphysics simulation, we develop a semi-analytical approach to evaluate adiabtic fuel temperature and introduce an intermediate time scale for temperature evaluation.
+%
+To test IQS and \iqspc nonlinear iteration, time step convergence, and time adaptation performance, four different test cases are employed: two problems are purely neutronics and involve varying magnitudes of reactor size and complexity; two problems involve temperature feedback and test an intermediate time scale for temperature evaluation.
 
@@
 \ee
-After reporting the above factorization in the balance equations, the shape diffusion equations result:
+After reporting the above factorization in the balance equations, the shape diffusion equations result (the main differences are highlighted using boxes around modified/new terms):
 \begin{subequations}
 \be
-\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \varphi^{g'} - \left(L^g + \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right)\varphi^g + \boxed{\frac{1}{p}}S_{d}^g
+\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \varphi^{g'} 
+    - \left(L^g + \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right)\varphi^g + \boxed{\frac{1}{p}}S_{d}^g
 \label{eq:shape}
@@
 the initial adjoint multigroup flux ($\phi^{*g}$), and then integrated over the phase-space domain.  
-For brevity, the inner product over space will be represented with parenthetical notation ($\left(\phi^{*g},f^g\right) = \int_D \phi^{*g}(\vec{r})f^g(\vec{r})d^3r
-$). In order to impose uniqueness of the factorization, one requires that 
+For brevity, the inner product over space will be represented with parenthetical notation ($\left(\phi^{*g},f^g\right) = \int_D \phi^{*g}(\vec{r})f^g(\vec{r})d^3r$). 
+In order to impose uniqueness of the factorization, one requires that 
 \be
@@
 \ee
-to be constant (hence $K(t)=K(t_0)=K_0$).  
+be constant (hence $K(t)=K(t_0)=K_0$).  
 After some manipulations, the point reactor kinetics equations (PRKE) for the amplitude solution are obtained:
@@
 
-Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three spatial dimensions, and even more so when using the transport equations in lieu of the diffusion equations.
-Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself, thus allowing for larger time step sizes in updating the shape.
-The PRKE equations form a small ODE system and can be solved using a much smaller time step size.
-In transients where the shape varies much less than the flux, IQS can be very computationally effective.
+Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three spatial dimensions, and even more so when using 
+the transport equations in lieu of the diffusion equations.  Using IQS, one expects the time dependence of the shape to be weaker than that 
+of the flux itself, thus allowing for larger time step sizes in updating the shape. The PRKE equations form a small ODE system and can be 
+solved using a much smaller time step size. In transients where the shape varies much less than the flux, IQS can be very computationally effective. 
 %The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.  
-\comm{technically there is an intermediate step for reactivity updates. a reviewer will criticize you easily for not mentioning it}
-
 %\begin{figure}[!htbp]
@@
 %\end{figure}
-
-It is important to note that the PRKE parameters are evaluated at each macro step and interpolated for the PRKE evaluation.
-In order to preserve the error convergence rate of high order temporal discretization schemes for shape, higher order interpolation of the parameters will be required.
+%
+Note that the PRKE parameters are evaluated at each macro step and interpolated for the PRKE evaluation. In order to preserve 
+the error convergence rate of high order temporal discretization schemes for shape, higher order interpolation of the parameters will be required.
 
@@
 
-The major difference between the methods of these authors is the convergence criteria used.
-Sissaoui and Koclas \cite{Sissaoui_1995, Koclas_1996} use fixed point iteration where the criteria is the simply the normalized difference between the last two computed shapes.
-Monier in \cite{Monier_diss} also employs fixed point iterations with the same criteria, except that the solution is rescaled by $K_n/K_{n+1}$ after each iteration.
-Dulla in \cite{Dulla2008} does the same fixed-point iteration, but the shape is scaled by $K_0/K_{n+1}$ at the end of each macro time step.
-Renormalizing the shape is essential for preserving the uniqueness condition ($K_{n+1} = K_0$); this condition is not inherently conserved, even when the solution has converged from shape iterations.
-Devooght in \cite{Devooght_1984} uses a Newton-SOR iteration where the residual of the shape function evaluation is the convergence criteria and next iteration's
+The major difference between the methods of these authors is the convergence criteria used.  Sissaoui and Koclas \cite{Sissaoui_1995, Koclas_1996} 
+use fixed point iteration where the criteria is the simply the normalized difference between the last two computed shapes.  Monier in \cite{Monier_diss} 
+also employs fixed point iterations with the same criteria, except that the solution is rescaled by $K_n/K_{n+1}$ after each iteration. Dulla in 
+\cite{Dulla2008} does the same fixed-point iteration, but the shape is scaled by $K_0/K_{n+1}$ at the end of each macro time step. Renormalizing 
+the shape is essential for preserving the uniqueness condition ($K_{n+1} = K_0$); this condition is not inherently conserved, even when the 
+solution has converged from shape iterations. Devooght in \cite{Devooght_1984} uses a Newton-SOR iteration where the residual of the shape function 
+evaluation is the convergence criteria and next iteration's 
 solution is computed using Newton-Raphson method.
 
-These techniques are by no means an exhaustive list of the possible iteration techniques for IQS.
-Dulla et al.
-in \cite{Dulla2008} provide an in depth analysis of the fixed point iteration technique most similar to Sissaoui and Koclas, involving convergence rates and solution results.
-However, no comprehensive analysis of iteration techniques exists, comparing both Newton and fixed-point convergence rates.
-The following describes each iteration convergence criterion investigated in this paper:
+These techniques are by no means an exhaustive list of the possible iteration techniques for IQS. Dulla et al. in \cite{Dulla2008} provide an 
+in-depth analysis of the fixed point iteration technique most similar to Sissaoui and Koclas, involving convergence rates and solution results.  
+However, no comprehensive analysis of iteration techniques exists, comparing both Newton and fixed-point convergence rates.  The following describes 
+each iteration convergence criterion investigated in this paper:
 \comm{you never show Newton results, do you?}
+
 \answer{zmp}{no, never figured out how to do it, plus Devooght actually uses GQS for his results.}
+
+\comm{there are parts of the paper where it feels like the reader is going to expect Newton-based results\\
+(I feel like you said so in the abstract). You should amend these statements}
+
 \begin{itemize}
@@
 \]
-\comm{why is this not a relative error criterion like the others? what's the rationale?}
-\answer{zmp}{I copied all of the above from \cite{Monier_diss}, not sure the rationale though}
+%\comm{why is this not a relative error criterion like the others? what's the rationale?}
+%\answer{zmp}{I copied all of the above from \cite{Monier_diss}, not sure the rationale though}
 \item Uniqueness consistency \cite{Monier_diss}: 
@@
 
-In order to ensure that uniqueness criteria is preserved, some authors (\cite{Monier_diss,Dulla2008}) explicitly scale the shape solution, either at each nonlinear iteration or upon exiting the nonlinear loop on the macro time time step.
-This re-scaling is simply
+In order to ensure that uniqueness criteria is preserved, some authors \cite{Monier_diss,Dulla2008} explicitly scale the shape solution, 
+either at each nonlinear iteration or upon exiting the nonlinear loop on the macro time time step. This re-scaling is simply
 \be 
@@
 \ee
-\comm{so, how does this jive with situations when the criterion on $K$ is employed?}
-\answer{zmp}{if the criteria is satisfied, it doesn't rescale; if it isn't, the solution is rescaled and another iteration is performed, the resulting solution isn't guaranteed to satisfy the criteria}
+%\comm{so, how does this jive with situations when the criterion on $K$ is employed?}
+%\answer{zmp}{if the criteria is satisfied, it doesn't rescale; if it isn't, the solution is rescaled and another iteration is performed, the resulting solution isn't guaranteed to satisfy the criteria}
 
@@
 %---------------------------------------------------------------------------------------%
-\comm{make sure the intro has a decent overview -i.e., no equation, just words- on the IQS-PC. Dulla and Ravetto will likely review this paper}
-
-The Predictor-Corrector version of the IQS method (IQS-PC) is a linearized version of IQS, where no nonlinear iterations are required.
-The method relies on the shape and amplitude factorization of the flux and uniqueness condition.
-The PRKE derivation is identical to that of the standard (nonlinear) version of the IQS scheme, in the sense that shape solutions at the beginning and end times of the macro time interval are used (with interpolation in between).
-However, the manner in which the shape function is obtained is different.
-In the IQS-PC version, the flux equations (not the shape equations) are first solved (represented by \eqts{eq:flux}{eq:precursor}) in order to obtain a {\it predicted} flux solution at the end of the macro time step.
-This predicted flux is then converted to a shape by normalizing it as follows:
-\be
-\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}} \,,
+
+The Predictor-Corrector version of the IQS method (\iqspc) is a linearized version of IQS, where no nonlinear iterations are required. 
+The method relies on the shape and amplitude factorization of the flux and uniqueness condition. The PRKE derivation is 
+identical to that of the standard (nonlinear) version of the IQS scheme, in the sense that shape solutions at the beginning 
+and end times of the macro time interval are used (with interpolation in between).
+However, the manner in which the shape function is obtained is different.  In the \iqspc version, the flux equations 
+(not the shape equations) are first solved (represented by \eqts{eq:flux}{eq:precursor}) in order to obtain a 
+{\it predicted} flux solution at the end of the macro time step. This predicted flux is then converted to a shape by 
+normalizing it as follows:
+\be
+\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{\mathcal{K}_{n+1}} \,,
 \label{eq:rescale}
@@
 \be
-K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right) \,,
+\mathcal{K}_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right) \,,
 \ee
@@
 The PRKE parameters are then computed with this shape using \eqtss{eq:rmb}{eq:l} and interpolated over the macro step, then the 
-PRKE ODE system is solved on the micro time scale.  With the newly computed amplitude, the shape is rescaled into a flux and the final {\it corrected} flux is given by:
+PRKE ODE system is solved on the micro time scale.  With the newly computed amplitude, the shape is rescaled into a flux and the final 
+{\it corrected} flux is given by:
 \be
@@
 
-The advantage to the predictor-corrector method is that no nonlinear iterations are necessary in this method; it is simpler to implement and can be faster than the standard IQS.
-Ikeda et al.
-in \cite{Ikeda_2001} and Goluoglu et al.
-in \cite{Goluoglu_2001} both use IQS-PC for complex, three-dimensional problems.
-Their results prove IQS-PC to be effective.
-Dulla et al.
-in \cite{Dulla2008} also describes an in depth comparison of IQS-PC with traditional IQS, which originally brought the method to light.
+The advantage to the predictor-corrector method is that no nonlinear iterations are necessary in this method; it is simpler to 
+implement and can be faster than the standard IQS.  Ikeda et al. in \cite{Ikeda_2001} and Goluoglu et al. in \cite{Goluoglu_2001} 
+both use \iqspc for complex, three-dimensional problems.  Their results prove \iqspc to be effective.  Dulla et al. in \cite{Dulla2008} 
+also describes an in depth comparison of \iqspc with traditional IQS, which originally brought the method to light.
 
@@
 portrayed in \fig{fig:time}.  The first time scale is the shape solve, the second is the temperature evaluation 
-as well as the computation of PRKE parameters, and the third is the PRKE scale.
-It is important to note that the number of time steps in each scale is arbitrary and the number chosen in \fig{fig:time} are only meant for
+as well as the computation of PRKE parameters, and the third is the PRKE scale.  It is important to note that the 
+number of time steps in each scale is arbitrary and the number chosen in \fig{fig:time} are only meant for 
 illustrative purposes.
@@
 
-Iteration processes are needed in between each time scale.
-The amplitude and temperature need to be iterated on the middle time scale until convergence on each temperature step.
-Then another iterative process needs to occur in the shape time scale on all three variables.
-\fig{fig:proc} shows the programming diagram implement to execute this process.
-The time increment of $\Delta t/3$ for the temperature solve is arbitrary and is meant only to match \fig{fig:time} where three temperature updates have been used as illustration.
+Iteration processes are needed in between each time scale.  The amplitude and temperature need to be iterated on the middle 
+time scale until convergence on each temperature step.  Then another iterative process needs to occur in the shape time scale on 
+all three variables. \fig{fig:proc} shows the programming diagram implement to execute this process. The time increment of 
+$\Delta t/3$ for the temperature solve is arbitrary and is meant only to match \fig{fig:time} where three temperature updates 
+have been used as illustration.
 
@@
 %---------------------------------------------------------------------------------------%
-\comm{in the results, will you comment on the differences between using the theta scheme versus
-the semi-analytical integration for the precursors?}
-\answer{zmp}{yes I do}
-
-\comm{was this section supposed to be moved up?}
-\answer{zmp}{I think we decided to put it after temperature, to make multiphysics seem more important}
 
@@
 \ee
-The shape $S_f$ is not known continuously over the time step but we use the beginning/end times values
-and make a linear variation over the macro time step,
+The shape function (using in $S_f$) is not known continuously over the time step but one may use the beginning/end times values
+to linearly interpolate the fission source over the macro time step,
 \be
@@
 \ee
-However, we have a very accurate representation of $p(t)$ over the macro step from the PRKE solve (micro time scale). 
+Furthermore, a very accurate representation of $p(t)$ over the macro step is available from the PRKE solve (micro time scale). 
 Using this fine scale information, the analytical solve for the precursor values yields
@@
 \ee
+and
 \be
@@
 difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps 
-($\varphi^g_{\Delta t/2}$). Note: $\varphi$ is changed to $\phi$ in the case of the IQS-PC tehcnique.
+($\varphi^g_{\Delta t/2}$). Note: $\varphi$ is changed to $\phi$ in the case of the \iqspc tehcnique.
 %
@@
 \be
-\Delta t_{new} = S \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}} \,,
+\Delta t_{new} = s \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}} \,,
 \label{eq:dt2}
 \ee
-where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor. 
+where $q$ is the convergence order of the time integration scheme being used and $s\simeq 0.8$ is a safety factor. 
 If the error is larger than the user-specified tolerance, the time step is rejected and new (smaller) macro time 
@@
 To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to 
-direct temporal discrezation of the flux equation (no IQS), the standard IQS method, and the IQS-PC version. 
+direct temporal discrezation of the flux equation (no IQS), the standard IQS method, and the \iqspc version. 
 Each of these methods will be applied to several diffusion problems; the number of time steps taken and 
@@
 In order to evaluate the performance and error convergence of IQS, the slab was simulated with varying time discretization methods and time step sizes.
-\fig{fig:1D_conv} shows these convergence plots of five different discretization methods for implicit discretization, IQS, and IQS P-C.
+\fig{fig:1D_conv} shows these convergence plots of five different discretization methods for implicit discretization, IQS, and \iqspc.
 These plots were generated from the results using the MATLAB prototype program.
-The plots show that IQS and IQS P-C are convergent through fourth order BDF.
+The plots show that IQS and \iqspc are convergent through fourth order BDF.
 Third order SDIRK did not show third order convergence, but, through extensive testing, SDIRK shows non-convergent behavior for too stiff of problems.
@@
 \end{subfigure}
-\caption{Error convergence plots of implicit discretization, IQS, and IQS P-C with various time discretization schemes}
+\caption{Error convergence plots of implicit discretization, IQS, and \iqspc with various time discretization schemes}
 \label{fig:1D_conv}
@@
 The results show that both IQS methods perform exceptionally well compared to implicit discretization.
-It also shows that traditional IQS performed better with large $e_{tol}$, while IQS P-C was better with smaller $e_{tol}$.
+It also shows that traditional IQS performed better with large $e_{tol}$, while \iqspc was better with smaller $e_{tol}$.
 
@@
 \hline
-  &  & \multicolumn{3}{|c|}{Implicit Discretization} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
+  &  & \multicolumn{3}{|c|}{Implicit Discretization} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{\iqspc} \\
 \hline
@@
 The mesh consisted of blocks $11\times 11$ with five uniform refinements, totaling $165,165$ elements and $124,609$ nodes.
-Three different temporal techniques were applied: implicit discretization of the flux equation, IQS, and IQS-PC.
+Three different temporal techniques were applied: implicit discretization of the flux equation, IQS, and \iqspc.
 Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.
@@
 These run-times are based on total alive time of the execution where the diffusion evaluation is distributed over 24 processors.
-These run-times show a marginal performance for IQS and impressive performance for IQS P-C.
+These run-times show a marginal performance for IQS and impressive performance for \iqspc.
 Some of the execution times were able to decrease from implicit discretization with the same number of macro steps because IQS is better equipped to resolve the nonlinearity between temperature and amplitude.
-Furthermore, there does seem to be an ideal number of temperature updates to optimize execution time: IQS only needs one and IQS P-C seems to be ideal at 4 updates.
+Furthermore, there does seem to be an ideal number of temperature updates to optimize execution time: IQS only needs one and \iqspc seems to be ideal at 4 updates.
 This discrepancy in the number of updates shows that a adaptive type implementation of the updates would be ideal, and could enforce a constant error over the transient.
-It is also important to compare the error of implicit discretization with IQS at one update and IQS P-C at 4 updates.
+It is also important to compare the error of implicit discretization with IQS at one update and \iqspc at 4 updates.
 IQS shows an error comparable to implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by -34.1\%.
-IQS P-C shows an error less than implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by $<-34.9\%$.
+\iqspc shows an error less than implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by $<-34.9\%$.
 
@@
 
-\fig{fig:lra_dt2} shows the power profile of the LRA with time adaptation of implicit discretization and IQS P-C, and \tbl{tab:lra_dt2} compiles the results.
-These time adaptation results show the significant decrease in macro time steps required for IQS P-C.
-These profiles were obtaining with only one temperature update per macro step; so based on previous results, the IQS P-C performance would improve even more with more updates.
+\fig{fig:lra_dt2} shows the power profile of the LRA with time adaptation of implicit discretization and \iqspc, and \tbl{tab:lra_dt2} compiles the results.
+These time adaptation results show the significant decrease in macro time steps required for \iqspc.
+These profiles were obtaining with only one temperature update per macro step; so based on previous results, the \iqspc performance would improve even more with more updates.
 
@@
 \end{subfigure}
-\caption{LRA power profile with time adaptation of implicit discretization and IQS P-C}
+\caption{LRA power profile with time adaptation of implicit discretization and \iqspc}
 \label{fig:lra_dt2}
@@
 \begin{center}
-\caption{LRA step doubling adaptation results with implicit discretization and IQS P-C}
+\caption{LRA step doubling adaptation results with implicit discretization and \iqspc}
 \label{tab:lra_dt2}
@@
 \hline
- & \multicolumn{3}{|c|}{Implicit Dis.} & \multicolumn{3}{|c|}{IQS P-C} \\
+ & \multicolumn{3}{|c|}{Implicit Dis.} & \multicolumn{3}{|c|}{\iqspc} \\
 \hline
@@
 
-In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time step, implicit discretization, IQS with one and 5 temperature updates per macro step, and IQS P-C with one and 5 updates.
+In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time step, implicit discretization, IQS with one and 5 temperature updates per macro step, and \iqspc with one and 5 updates.
 \fig{fig:Tran15_profile} shows the baseline power and temperature profile for the Transient-15 example.
@@
 IQS (5 updates) 	& 300 	& 3.5040e+08 & 1.90  & 371 &  49.7\%	&	3.687e-5 & 24035	\\
-IQS P-C 			& 300 	& 3.5065e+08 & 1.90  & 371 &  -2.1\%	&	7.527e-4 & 39020	\\
-IQS P-C (5 updates) & 300 	& 3.5043e+08 & 1.90  & 371 &  26.5\%	&	1.227e-4 & 37866	\\
+\iqspc 			& 300 	& 3.5065e+08 & 1.90  & 371 &  -2.1\%	&	7.527e-4 & 39020	\\
+\iqspc (5 updates) & 300 	& 3.5043e+08 & 1.90  & 371 &  26.5\%	&	1.227e-4 & 37866	\\
 \hline
@@
 Comparing this error to a similar implicit discretization error at a smaller time step could show that the runtime was reduced.
-IQS P-C performed not nearly as well as it did with the LRA benchmark, but still proved to be effective.
-Having 5 updates for IQS P-C increased the runtime marginally, but decreased the error significantly.
+\iqspc performed not nearly as well as it did with the LRA benchmark, but still proved to be effective.
+Having 5 updates for \iqspc increased the runtime marginally, but decreased the error significantly.
 

### 2017-09-09 doc/Papers/IQS_convergence/IQS_convergence.tex
reviewed kinetic results

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Kinetics Results}
+\section{Neutronics-only Transient Results}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-This section describes kinetics examples that IQS is tested with and analysis of its performance.
-The examples range in complexity and application.
-The first is a one-dimensional problem, designed for the prototype code in MATLAB.
-The next one is from the Argonne National Lab (ANL) Benchmark Problem Book (BPB), and is a common problem for testing codes and developing methods \cite{ANL_BPB}.
+In this section, we analyze various convergence criteria proposed in the nonlinear solution tehcnique for the IQDS method.
+We also investigate the use of higher-temporal discretization for the shape equations.
+Both series of tests are carried out using neutronics-only problems.  The first test case uses a one-dimensional problem.  
+The second test is from the ANL Benchmark Problem Book (BPB) \cite{ANL_BPB}.
 
@@
 
-This example is very simple and computes quickly; it entails a one dimensional, homogeneous 400 cm slab with a heterogenous perturbation in absorption cross section.
-\fig{fig:slab} shows how the regions of the slab are divided and \tbl{tab:1Dmat} shows the initial material properties.
-Regions 2, 3, and 4 have slope perturbations at different points in time, \tbl{tab:1Dslope} shows the values of the absorption cross-section in each region at the times of interest.
-The values of $\Sigma_a$ between these times of interest are linear interpolations between the given values.
+This one-dimensional problem uses a 400-cm slab. This initial configuration is homogeneous and the transient is initiated by a perturbation in absorption cross section. 
+\fig{fig:slab} shows problem layout and \tbl{tab:1Dmat} contains the material properties; one-group values are used for this problem. 
+Regions 2, 3, and 4 have linear ramp perturbations at different moments in time, \tbl{tab:1Dslope} shows the values of the absorption cross-section in each region at the 
+times of interest.  The values of $\Sigma_a$ between these times of interest vary linearly between the given values.
 
@@
 }
-\caption{1-D slab region identification}
+\caption{1-D slab: Region identification number}
 \label{fig:slab}
@@
 
-\fig{fig:1D_flux} shows the resulting baseline flux and relative power profile of the one-dimensional problem.
-The baseline was computed using MATLAB's ode15s function which is a embedded Runge-Kutta time adaptive method for stiff problems, the error tolerance was set very tightly ($10^{-12}$).
-The flux distribution shows that the core is relatively large, thus different regions are weakly coupled.
-This baseline computation is used to compute the error of the other time discretization methods.
+\fig{fig:1D_flux} shows the reference flux solution at different instants as well as total power as a function of time.  The baseline reference solution 
+was computed by solving the flux equation (not the IQS formulation) using time-step control with a tight relative error tolerance of $10^{-12}$.  
+The problem being relatively large, the various zones are weakly coupled
+and the flux perturbation is not global. This baseline computation is used to compute the error of the other time discretization methods.  
 \fig{fig:1D_shape} shows the shape profile at various times during the transient.  
-It is apparent that the shape is very time-dependent, so it is expected that IQS has marginal accuracy gain for a given time step size.
+% It is apparent that the shape is time-dependent, so it is expected that IQS has marginal accuracy gain for a given time step size.
 
@@
 \centering
-\begin{subfigure}[b]{\linewidth}
-\centering
-\includegraphics[height=3in]{figures/1D_flux.png}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=0.99\textwidth]{figures/1D_flux.png}
 \caption{Flux profile at various times}
 \end{subfigure}
-\begin{subfigure}[b]{\linewidth}
-\centering
-\includegraphics[height=3in]{figures/1D_power_base.png}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=0.99\textwidth]{figures/1D_power_base.png}
 \caption{Power profile over transient}
@@
 \end{figure}
+\jcr{the editor will like complain that the figure do not have the same height.
+you must have save them in matlab after playing with them.
+i recommend you save them witht he size and aspect ratio produced by matlab.}
 
@@
 \centering
-\begin{subfigure}[b]{\linewidth}
-\centering
-\includegraphics[height=3in]{figures/1D_shape2.png}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=0.99\textwidth]{figures/1D_shape2.png}
 \caption{Shape profile at various times}
@@
 \end{subfigure}
-\begin{subfigure}[b]{\linewidth}
-\centering
-\includegraphics[height=3in]{figures/1D_iqs_power.png}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=0.99\textwidth]{figures/1D_iqs_power.png}
 \caption{Power and reactivity profile over transient}
@@
 \end{figure}
-
-%----------------------------------------------------------------------------------------
-\subsubsection{Iteration Analysis}
-
-IQS, as previously stated, is a system of nonlinear equations between shape and amplitude.
-These equations needed to be iterated to numerically converge to an accurate solution.
-\sct{sect:iter} lists various iteration techniques for fixed-point and Newton schemes.
-\fig{fig:iter} shows the number of fixed-point iterations required for a $10^{-11}$ tolerance over the transient.
-The criteria listed in the legend correspond to the list as such:
-\begin{enumerate}
-\item $L^{\infty} \rightarrow$ 1.
-\item $L^{2} \rightarrow$ 2.
-\item Reactivity $\rightarrow$ 3.
-\item Amplitude $\rightarrow$ 4.
-\item K criteria $\rightarrow$ 5.
-\item All properties $\rightarrow$ 3.-5.
-\end{enumerate}
-This plot shows that 1-4 have approximately the same convergence behavior, but the K criteria converges to a certain error.
-\fig{fig:iter_err} shows the resulting error the K criteria converges to for different points of rescaling the shape.
-The rescaling is described by \eqt{eq:shape_scale}.
-Rescaling shape more frequently helps the error.
+\jcr{not sure that the Fig.7b is useful}
+
+In the standard IQS technique, a nonlinear system of equations is to be solved for the shape and the amplitude solutions.
+\sct{sect:iter} lists various iteration stopping criteria typically employed with IQS.
+\fig{fig:iter} gives the number of fixed-point iterations required to reach tolerance of $10^{-11}$ as a function of time during the transient.
+A theta-scheme is employed for the precurosrs equations (\jcr{recall the eq}).
+%The criteria listed in the legend correspond to the list as such:
+%\begin{enumerate}
+%\item $L^{\infty} \rightarrow$ 1.
+%\item $L^{2} \rightarrow$ 2.
+%\item Reactivity $\rightarrow$ 3.
+%\item Amplitude $\rightarrow$ 4.
+%\item $K$ criteria $\rightarrow$ 5.
+%\item The last 3 criteria combined $\rightarrow$ 3.-5.
+%\end{enumerate}
+This figures shows that criteria based on the  $L^{\infty}$ norm of the shape, the $L^2$ norm of the shape, the reactivity change, and the amplitude have approximately the same convergence behavior, but the criteria
+based on the uniqueness constant $K$ never reaches the tight tolerance and runs up to the maximum number of iterations allowed (20 in this case). 
+\fig{fig:iter_err} shows the error in the factor $K$ saturates to different levels, whether the shape is rescaled or not. Rescaling (described by \eqt{eq:shape_scale}) 
+%Rescaling shape more frequently helps the error.
 However, rescaling every iteration is somewhat artificial because it does not consider changes spatially.
 Regardless, an error of $~10^{-5}$ is quite large and it is expected that the magnitude is due to the explicit treatment of precursors (\eqt{eq:dnp_theta}).
-Switching to an analytical elimination (\eqt{eq:dnp_an}) does not converge as well, but the error is much smaller, seen in \fig{fig:iter_err_an}.
+However, switching the precursors solve to be performed semi-analytically (\eqt{eq:dnp_an}) significantly improves the convergence in the uniqueness factor $K$, see \fig{fig:iter_err_an}.
 
@@
 \end{figure}
-
 \begin{figure}[!htbp]
@@
 \includegraphics[height=3in]{figures/iter_error.png}
-\caption{Final iteration error for K convergence criteria}
+\caption{Final iteration error for $K$ factor convergence criteria}
 \label{fig:iter_err}
 \end{figure}
-
 \begin{figure}[!htbp]
@@
 \includegraphics[height=3in]{figures/iter_error_an.png}
-\caption{Final iteration error for K convergence criteria with analytical precursor elimination}
+\caption{Final iteration error for $K$ convergence criteria with analytical precursor elimination}
 \label{fig:iter_err_an}
 \end{figure}
+\jcr{can you make subfigures with the last 2 figures to group them. otherwise, they will be scattered around by the editorial staff}
+
+\clearpage
+\newpage
 
 %----------------------------------------------------------------------------------------
-\subsubsection{One-Dimensional Problem Time Step Convergence Analysis}
-
-Time step convergence analysis involves evaluating a problem with various refinements in step size and comparing the resulting errors with the time step size.
+%\subsubsection{One-Dimensional Problem Time Step Convergence Analysis}
+
+Next, we investigate the use of higher-order temporal methods to solve the shape equations (IQS and \iqspc). We compare these results with the same high-order discretization for the flux equations.
+%Time step convergence analysis involves evaluating a problem with various refinements in step size and comparing the resulting errors with the time step size.
 Plotting error versus $\Delta t$ on a log-scale should produce a relatively strait line with a slope equal to the order of the time discretization method.
 In order to evaluate the performance and error convergence of IQS, the slab was simulated with varying time discretization methods and time step sizes.
-\fig{fig:1D_conv} shows these convergence plots of five different discretization methods for implicit discretization, IQS, and \iqspc.
-These plots were generated from the results using the MATLAB prototype program.
-The plots show that IQS and \iqspc are convergent through fourth order BDF.
-Third order SDIRK did not show third order convergence, but, through extensive testing, SDIRK shows non-convergent behavior for too stiff of problems.
-
-There are higher discretization order that can be tested, but most practical application do not go beyond second order.
-This paper shows an analysis of the first publicized application of IQS with higher than second order discretization, which exposed unforeseen properties of IQS.
+\fig{fig:1D_conv} shows the convergence results for four different time-implicit discretization: backward difference formulae (for order 1, which is the same as the implicit Euler, to order 4) 
+\jcr{give refs for BDF methods of any order, do not pick wikipedia!}. % as well as a 3rd-order implicit Runge-Kutta method, SDIRK33. \jcr{also give a ref}
+The plots show that IQS and \iqspc converge with the expected order (1 through 4) using BDF1 through BDF4. Recall that in the amplitude equations, the PRKE parameters
+are interpolated using the shape solutions. We used higher-order interpolants in time for the shape solution when a higher-order method was used
+(the convergence rates were degraded for higher-order BDF methods when the PRKE parameters were only linearly interpolated over the macro step). Lagrange
+and Hermite interpolants yielded the same results. Likewise, we used semi-analytical integration of the precursors equations, along with the same higher-order
+interpolation of the shape values when a higher-order method was used. Without using the same order interpolants as the BDF formula, the error in the
+PRKE parameters and the precursors led to loss of convergence rate.
+
+% The Third-order SDIRK method did not show third order convergence, but, through extensive testing, SDIRK shows non-convergent behavior for too stiff of problems.  
+\jcr{I am removing sdirk, the point we want to make is that high-order methods do not buy anything for iqs. the sdirk issue is unrelated so why confuse the audience}
+
+The convergence rates for the straightforward discretization of the flux equations are also shown in \fig{fig:1D_conv} and they followed the same expected convergence rates.
+The flux results are indicated by the legend ``Implicit Dis.''. It is worthy to note that, for lower order discretization, the error in the flux equations is always much higher 
+than that in the IQS results. However, as the temporal order is increased, the gap between the discretization error in the flux equations and the IQS equations vanishes. 
+In previous publications, she shape solve int he IQS method always employed a lower-order temporal discretization. Our results seem to indicate that a high-order discretization 
+of the flux equations reduces the usefulness of the IQS approach.
+%This paper shows an analysis of the first publicized application of IQS with higher than second order discretization, which exposed unforeseen properties of IQS.
 When using higher order techniques, the interpolation of PRKE parameters and shape for precursor integration become important to consider.
@@
 \includegraphics[width=\linewidth]{figures/1D_conv_IE.png}
-\caption{Implicit Euler}
+\caption{Implicit Euler (BDF1)}
 \end{subfigure}
@@
 \end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{figures/1D_conv_SDIRK33.png}
-\caption{SDIRK33}
-\end{subfigure}
+%\begin{subfigure}[b]{0.49\textwidth}
+%\centering
+%\includegraphics[width=\linewidth]{figures/1D_conv_SDIRK33.png}
+%\caption{SDIRK33}
+%\end{subfigure}
 \caption{Error convergence plots of implicit discretization, IQS, and \iqspc with various time discretization schemes}
@@
 \subsubsection{One-Dimensional Mini-Core Problem}
-
+\jcr{i do not understand the purpose of the entire section.
+shouldn't the mini core be doen with bdf4 too, not just bdf1?
+and aren't the twigl results already such a demonstration so why not move on to twigl right away?}
 This problem is exactly the same as the previous one-dimensional problem, except the the core was reduced to 80 cm in length.
@@
 
-This benchmark problem originates from the Argonne National Lab Benchmark Problem Book \cite{ANL_BPB}.
+This benchmark problem originates from the ANL Benchmark Problem Book \cite{ANL_BPB}.
 It is a 2D, 2-group reactor core model with no reflector region \cite{TWIGL_benchmark}.
@@
 %----------------------------------------------------------------------------------------
-\subsubsection{TWIGL Time Step Convergence Analysis}
+%\subsubsection{TWIGL Time Step Convergence Analysis}
 
@@
 %----------------------------------------------------------------------------------------
-\subsubsection{TWIGL Time Adaptation}
+%\subsubsection{TWIGL Time Adaptation}
 
@@
 7 &	5.0e-05 &	0.00018372 &	112 &	342 &	6.0328e-05 &	12  &	163  &	1.5247e-06 &	24 &	92  \\
-8 &	1.0e-05 &	8.0564e-05 &	263 &	794 &	7.7103e-05 &	379 &	5729 &	9.8321e-07 &	48 &	210 \\
+% 8 &	1.0e-05 &	8.0564e-05 &	263 &	794 &	7.7103e-05 &	379 &	5729 &	9.8321e-07 &	48 &	210 \\
 \hline
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Dynamics Results}
+\section{Multiphysics Results}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-This section describes two dynamics examples, including the LRA benchmark and a TREAT experiment.
-These examples are of increased complexity from the previous kinetics examples.
+This section describes two multiphysics examples with adiabatic heat up: the LRA benchmark from the  ANL Benchmark Problem Book \cite{ANL_BPB} 
+and a modelisation of an experiment carried out of the Transient Reactor Experimental Facility, TREAT \jcr{provide a reference for this model; it could be an INL report or one of your conference paper}.
+% These examples are of increased complexity from the previous kinetics examples.
 This section also analyzes IQS's performance with these, which is vital for verification of IQS in real-world problems.
@@
 
-The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor.
-It is a super prompt-critical transient.
-The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.
-The spacial discretization was performed using continuous finite element method with first order Lagrangian basis functions.
-The mesh consisted of blocks $11\times 11$ with five uniform refinements, totaling $165,165$ elements and $124,609$ nodes.
-Three different temporal techniques were applied: implicit discretization of the flux equation, IQS, and \iqspc.
-Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.
+The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor.  It is a super prompt-critical transient. 
+The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.  The spatial discretization uses continuous finite elements
+with first-order Lagrangian basis functions.
+The mesh consists of blocks $11\times 11$ with five uniform refinements, totaling $165,165$ elements and $124,609$ nodes.
+Three different temporal techniques are
+applied: implicit discretization of the flux equation, IQS, and \iqspc.
+Crank-Nicolson \jcr{check spelling for CN emthod throughout} time discretization scheme is used for the diffusion evaluation of each technique.
 Third order Runge-Kutta discretization with step doubling adaptation was used for the PRKE evaluation.
+\jcr{actually, you never say what is used for the PRKE solve.
+this shoild be moved way up because it applies to everything!}
 The performance of IQS and the temperature updates were measured by its improvement in accuracy at peak power over the implicit discretization method.
@@
 %\resizebox{\linewidth}{!}{
-\begin{tabular}{|l|cc|}
-\hline
-Calculation  &  Baseline & Sutton (Spandex 1936) \\
-\hline
+\begin{tabular}{|l||c|c|}
+\hline
+Calculation   &  Baseline Rattlesnake & Sutton (Spandex 1936) \\
+\hline\hline
 No. of Spatial Nodes	& 3872 		& 1936 \\

### 2017-09-10 doc/Papers/IQS_convergence/IQS_convergence.tex
second pass on the kinetic results

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Implementation in a multiphysics simulation environment}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Both the standard IQS and \iqspc methods have been implemented in the Multiphysics Object-Oriented
+Simulation Envrionement, MOOSE \cite{}, as part of its radiation transport solver, Rattlesnake \cite{}.
+All multi-dimensional results presented here have been computed using the diffusion solver of Rattlesnake.
+MOOSE calculation flows naturally allows for a relatively straightforward implementation of the IQS methodologies.
+MOOSE's nonlinear solvers (fixed-point iterations and Jacobian-free Newton-Krylov) are naturally well
+suited to handle the nonlinear solve of the tradiationa IQS method during a macro time step. \iqspc's implementation
+is MOOSE is even simpler. We briefly describe the their implementation next.
+
+First, it is important to note that MOOSE provides entry points in its calculation flow at various moments:
+\begin{itemize}
+\item during a time step, with a pre-step and post-step call;
+\item at each nonlinear iteration;
+\item at each linear iteration.
+\end{itemize}
+At each of these instants, an auxiliary calculations may be initiated (e.g., a convergence check, a full PRKE
+solve, areactivity update, ...).
+
+\jcr{in plain English, can you describe how both iqs methods were implemented so that we give some recognition
+to moose a bit. thanks}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Neutronics-only Transient Results}
@@
 
-In this section, we analyze various convergence criteria proposed in the nonlinear solution tehcnique for the IQDS method.
+In this section, we analyze various convergence criteria proposed in the nonlinear solution technique for the IQDS method.
 We also investigate the use of higher-temporal discretization for the shape equations.
@@
 
-\fig{fig:1D_flux} shows the reference flux solution at different instants as well as total power as a function of time.  The baseline reference solution 
-was computed by solving the flux equation (not the IQS formulation) using time-step control with a tight relative error tolerance of $10^{-12}$.  
+\fig{fig:1D} shows the reference flux solution at different instants as well as total power as a function of time.  The baseline reference solution 
+was computed by solving the flux equations (not the IQS formulation) using time-step control with a tight relative error tolerance of $10^{-12}$.  
 The problem being relatively large, the various zones are weakly coupled
 and the flux perturbation is not global. This baseline computation is used to compute the error of the other time discretization methods.  
-\fig{fig:1D_shape} shows the shape profile at various times during the transient.  
+\fig{fig:1D_shape} shows the shape profile at the same instants during the transient as in \fig{fig:1D_flux};
+that solution was also obtained using time-step control with tight tolerance. Both flux solve and IQS solve yield
+the same answer, as expected.  
+% Even though the magnitude of the shape varies less during the transient, its spatial distribution shows more variation than that 
 % It is apparent that the shape is time-dependent, so it is expected that IQS has marginal accuracy gain for a given time step size.
@@
 \caption{Flux profile at various times}
+\label{fig:1D_flux}
 \end{subfigure}
@@
 \caption{Power profile over transient}
+\label{fig:1D_tot_power}
 \end{subfigure}
 \caption{Baseline flux and power distribution}
-\label{fig:1D_flux}
-\end{figure}
-\jcr{the editor will like complain that the figure do not have the same height.
-you must have save them in matlab after playing with them.
-i recommend you save them witht he size and aspect ratio produced by matlab.}
+\label{fig:1D}
+\end{figure}
+\jcr{the editor will like complain that the figure does not have the same height.
+you must have saved them in matlab after playing with them.
+i recommend you save them with the size and aspect ratio as produced by matlab.}
 
 \begin{figure}[!htbp]
-\centering
-\begin{subfigure}[b]{0.49\textwidth}
-\centering
-\includegraphics[width=0.99\textwidth]{figures/1D_shape2.png}
-\caption{Shape profile at various times}
+%\centering
+%\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=0.49\textwidth]{figures/1D_shape2.png}
+\caption{IQS Shape profile at various times}
 \label{fig:1D_shape}
-\end{subfigure}
-\begin{subfigure}[b]{0.49\textwidth}
-\centering
-\includegraphics[width=0.99\textwidth]{figures/1D_iqs_power.png}
-\caption{Power and reactivity profile over transient}
-\end{subfigure}
-\caption{IQS flux and power distribution}
-\label{fig:1D_res}
-\end{figure}
-\jcr{not sure that the Fig.7b is useful}
+%\end{subfigure}
+%\begin{subfigure}[b]{0.49\textwidth}
+%\centering
+%\includegraphics[width=0.99\textwidth]{figures/1D_iqs_power.png}
+%\caption{Power and reactivity profile over transient}
+%\end{subfigure}
+%\caption{IQS flux and power distribution}
+%\label{fig:1D_res}
+\end{figure}
+%\jcr{not sure that the Fig.7b is useful}
 
@@
 \centering
-\includegraphics[height=3in]{figures/iter_renorm.png}
+\includegraphics[width=0.49\textwidth]{figures/iter_renorm.png}
 \caption{\# of iterations for various convergence criteria, tolerance$=10^{-11}$, max iterations$=20$}
@@
 \end{figure}
+
 \begin{figure}[!htbp]
 \centering
-\includegraphics[height=3in]{figures/iter_error.png}
-\caption{Final iteration error for $K$ factor convergence criteria}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=0.99\textwidth]{figures/iter_error.png}
+\caption{Precursors solved using a theta-scheme}
 \label{fig:iter_err}
-\end{figure}
-\begin{figure}[!htbp]
-\centering
-\includegraphics[height=3in]{figures/iter_error_an.png}
-\caption{Final iteration error for $K$ convergence criteria with analytical precursor elimination}
+\end{subfigure}
+\begin{subfigure}[b]{0.49\textwidth}
+\centering
+\includegraphics[width=0.99\textwidth]{figures/iter_error_an.png}
+\caption{Analytical precursor solve}
 \label{fig:iter_err_an}
-\end{figure}
-\jcr{can you make subfigures with the last 2 figures to group them. otherwise, they will be scattered around by the editorial staff}
+\end{subfigure}
+\caption{Final iteration error for $K$ factor convergence criterion}
+\end{figure}
 
@@
 
-Next, we investigate the use of higher-order temporal methods to solve the shape equations (IQS and \iqspc). We compare these results with the same high-order discretization for the flux equations.
+Next, we investigate the use of higher-order temporal methods to solve the shape equations (standard IQS and \iqspc methods).
+We compare these results with the same high-order discretization for the flux equations.
 %Time step convergence analysis involves evaluating a problem with various refinements in step size and comparing the resulting errors with the time step size.
@@
 In order to evaluate the performance and error convergence of IQS, the slab was simulated with varying time discretization methods and time step sizes.
-\fig{fig:1D_conv} shows the convergence results for four different time-implicit discretization: backward difference formulae (for order 1, which is the same as the implicit Euler, to order 4) 
+\fig{fig:1D_conv} shows the convergence results for four different time-implicit discretization: backward difference formulae, from order 1, which is the same as the implicit Euler, to order 4.
 \jcr{give refs for BDF methods of any order, do not pick wikipedia!}. % as well as a 3rd-order implicit Runge-Kutta method, SDIRK33. \jcr{also give a ref}
@@
 (the convergence rates were degraded for higher-order BDF methods when the PRKE parameters were only linearly interpolated over the macro step). Lagrange
-and Hermite interpolants yielded the same results. Likewise, we used semi-analytical integration of the precursors equations, along with the same higher-order
+and Hermite interpolants yielded the same results. In addition, we used semi-analytical integration of the precursors equations, along with the same higher-order
 interpolation of the shape values when a higher-order method was used. Without using the same order interpolants as the BDF formula, the error in the
-PRKE parameters and the precursors led to loss of convergence rate.
+PRKE parameters and the precursors led to reduction in the observed convergence rate.
 
@@
 The convergence rates for the straightforward discretization of the flux equations are also shown in \fig{fig:1D_conv} and they followed the same expected convergence rates.
-The flux results are indicated by the legend ``Implicit Dis.''. It is worthy to note that, for lower order discretization, the error in the flux equations is always much higher 
-than that in the IQS results. However, as the temporal order is increased, the gap between the discretization error in the flux equations and the IQS equations vanishes. 
-In previous publications, she shape solve int he IQS method always employed a lower-order temporal discretization. Our results seem to indicate that a high-order discretization 
-of the flux equations reduces the usefulness of the IQS approach.
+These flux-solve results are indicated by the legend ``Implicit Dis.''. It is worthy to note that, for lower-order discretizations, the error in the flux equations is always significanrtly higher 
+than that of the IQS results. However, as the order of the temporal scheme is increased, the gap between the discretization error in the flux equations and the IQS equations vanishes. 
+In previous publications, the shape solve in the IQS method always employed a lower-order temporal discretization
+(often implicit Euler). Our results seem to indicate that a high-order discretization 
+of the flux equations may reduce the usefulness of the IQS approach (at least for large domains with weakly coupled spatial regions).
 %This paper shows an analysis of the first publicized application of IQS with higher than second order discretization, which exposed unforeseen properties of IQS.
@@
 \centering
-\includegraphics[height=3in]{figures/1Dmini_flux.png}
+\includegraphics[height=1.5in]{figures/1Dmini_flux.png}
 \caption{Flux profile at various times}
@@
 \centering
-\includegraphics[height=3in]{figures/1Dmini_power_base.png}
+\includegraphics[height=1.5in]{figures/1Dmini_power_base.png}
 \caption{Power profile over transient}
@@
 \centering
-\includegraphics[height=3in]{figures/1Dmini_shape.png}
+\includegraphics[height=1.5in]{figures/1Dmini_shape.png}
 \caption{Shape profile at various times for one-dimensional mini-core}
@@
 \centering
-\includegraphics[height=3in]{figures/1Dmini_conv_IE.png}
+\includegraphics[height=1.5in]{figures/1Dmini_conv_IE.png}
 \caption{Time step convergence for one-dimensional mini-core with implicit Euler discretization}
@@
 
-This benchmark problem originates from the ANL Benchmark Problem Book \cite{ANL_BPB}.
-It is a 2D, 2-group reactor core model with no reflector region \cite{TWIGL_benchmark}.
-This example is meant to be of progressive complexity from the previous example.
-The transient of this reactor is very geometrically symmetrical with very little temporal shape change.
-Therefore, IQS is expected to perform significantly better than the implicit discretization method.
-Figs.
-\ref{fig:TWIGL_power} show the IQS  solution as compared with the implicit discretization solution.
-This shows that IQS is consistent in more complex, higher dimensional problems.
-These plots also serve to illustrate that IQS has a much more accurate solution, even at a significantly larger time step than the implicit discretization.
+This benchmark problem originates from the ANL Benchmark Problem Book \cite{ANL_BPB}.  It is a 2D, 2-group reactor core model with no reflector region \cite{TWIGL_benchmark}.  
+% This example is meant to be of progressive complexity from the previous example.  
+The core is smaller (hence the spatial regions are more tightly coupled) and the spatial variation of the flux
+varies moderately over the transient. 
+Therefore, IQS is expected to perform significantly better than an implicit flux solve.
+\fig{fig:TWIGL_power} shows the IQS  solution as compared with the implicit flux solution.
+This illustrates that IQS can yield a much more accurate solution, even at a significantly larger time step than the implicit flux discretization.
 
@@
 
-In order to demonstrate asymptotic convergence of IQS, implicit Euler (IE) and second order BDF (BDF2) were applied to the TWIGL simulation.
+In order to demonstrate asymptotic convergence of IQS, implicit Euler (IE) and second order BDF (BDF2) were applied to solve the TWIGL problem.
 \fig{fig:TWIGL_conv} plots the error convergence of IQS and the implicit discretization methods.
-The curves show the impressive convergence of IQS for the highly transient TWIGL example.
-The slope indicated in the legend are the linear slope of curves on the log plot, these slopes should be similar to the order of the method (1 for IE and 2 for BDF2).
-IQS shows a increased order because the PRKE is performing much of accuracy convergence and it is computed using SDIRK33, a third order method.
+The curves show the superior convergence of IQS.
+The observed convergence rates are indicated in the legend.
+They show orders 1 for IE and 2 for BDF2 for the flux solves. However, the convergence rates for IQS show higher values,
+order 2 for IE solves for the shape, and above 2.5 for BDF2 solve of the shape.
+When there is moderate variation in the shape over the transient, the PRKE parameters are accurately estimated, and the fine-scale PRKE solve provides a well resolved amplitude solution.
 
@@
 
-\tbl{tab:TWIGLdt2} and \fig{fig:TWIGL_power_dt2} show the results for TWIGL with time adaptation.
-The results show that both IQS methods perform exceptionally well compared to implicit discretization.
+\tbl{tab:TWIGLdt2} and \fig{fig:TWIGL_power_dt2} show the results for the TWIGL benchmark with time adaptation.
+The results show that both IQS methods perform exceptionally well compared to implicit flux discretization.
 It also shows that traditional IQS performed better with large $e_{tol}$, while \iqspc was better with smaller $e_{tol}$.
@@
 \centering
-\includegraphics[height=3in]{figures/TWIGL_power_plot_dt2.png}
-\caption{Power level comparison of TWIGL Benchmark with time adaptation}
+\includegraphics[height=2in]{figures/TWIGL_power_plot_dt2.png}
+\caption{Power level comparison of TWIGL Benchmark with time step control}
 \label{fig:TWIGL_power_dt2}
@@
 This section describes two multiphysics examples with adiabatic heat up: the LRA benchmark from the  ANL Benchmark Problem Book \cite{ANL_BPB} 
-and a modelisation of an experiment carried out of the Transient Reactor Experimental Facility, TREAT \jcr{provide a reference for this model; it could be an INL report or one of your conference paper}.
+and a modelisation of an experiment carried out at the Transient Reactor Test Facility, TREAT \jcr{provide a reference for this model; it could be an INL report or one of your conference paper}. 
 % These examples are of increased complexity from the previous kinetics examples.
@@
 
-The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor.  It is a super prompt-critical transient. 
+The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up 
+and Doppler feedback in thermal reactor.  It is a super prompt-critical transient. 
 The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.  The spatial discretization uses continuous finite elements
@@
 \jcr{actually, you never say what is used for the PRKE solve.
-this shoild be moved way up because it applies to everything!}
+this should be moved way up because it applies to everything!}
 The performance of IQS and the temperature updates were measured by its improvement in accuracy at peak power over the implicit discretization method.

### 2017-09-13 doc/Papers/IQS_convergence/IQS_convergence.tex
Addressing kinetics part comments

@@
 \begin{abstract}
-The improved quasi-static method (IQS) is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent 
-amplitude and a time-space-energy dependent shape. The objective of the IQS factorization is to evaluate amplitude and shape on different time scales 
-in order to reduce the computational burden associated with solving the multi-dimensional flux equations, while maintaining solution accuracy. 
-The IQS decomposition leads to a nonlinear system of equations that requires iteration of shape and amplitude. IQS iteration techniques involve 
-fixed-point (Picard) iteration with various convergence criteria and Newton iteration, namely preconditioned Jacobian-free Newton Krylov (PJFNK) method. 
-\jcr{space is at a premium in the abstract. why talk about jfnk if no such results are presented? i must be missing sthg}
-Nonlinear convergence of each \jcr{each?} of these techniques is investigated. Verification of IQS with analysis of time step convergence is also investigated. 
-Proper time step convergence is vital for implementation of time adaptive methods and error prediction. The time derivative of the shape function is 
-discretized through fourth order using implicit-Euler, Crank-Nicholson, backward difference formulae (BDF), and singly-diagonally-implicit 
-Runge-Kutta (SDIRK) methods.
+The improved quasi-static method (IQS) is a rigorous space/time multiscale approach whereby the neutron flux is represented by a time-dependent amplitude and a time-, space-, and energy-dependent shape.
+The objective of the IQS factorization is to evaluate amplitude and shape on different time scales in order to reduce the computational burden associated with solving the multi-dimensional flux equations, while maintaining solution accuracy.
+The IQS decomposition leads to a nonlinear system of equations that requires iteration of shape and amplitude.
+IQS iteration techniques involve fixed-point (Picard) iteration with various convergence criteria and shape rescaling.
+Nonlinear convergence of each of these techniques is investigated.
+Verification of IQS with analysis of time step convergence is also investigated in order to investigate the method's effectiveness with high-order schemes.
+The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicolson, and backward difference formulae (BDF).
 \end{abstract}
@@
 
-The improved quasi-static method (IQS) is a numerical technique devised for nuclear reactor transient analysis. It involves factorizing the neutron flux 
-solution into a time-only-dependent component, the amplitude, and a space- and time-dependent component, the shape \cite{Ott_1966,Devooght_1984,Monier_diss,Sissaoui_1995,Dulla2008}. 
-The amplitude solution satisfies the point reactor kinetic equations (PRKE) where the shape solution have been used to generate the PRKE coefficients 
-(reactivity, effective fraction of delayed neutrons, mean generation time). The shape solution satisfies a modified time-dependent neutron balance equation. 
-The rationale for the IQS method lies in the assumption that the shape function is weakly dependent on time. Therefore, it is expected that the modified time-dependent neutron 
-balance equations be less stiff than the original time-dependent neutron balance equations for the flux. As a result, the shape may not require to be solved for at the same frequency 
-as the amplitude, but only on larger macro-time steps, which is expected to yield wall clock savings, especially in multi-dimensional geometries. The PRKE form a small system of ordinary
-differential equations (ODE) and solving them on a fine temporal grid is not a computational burden. As opposed to the standard PRKE approach, whereby a 
-shape function is selected, typically once for an entire transient without updates, the IQS technique is obtained in a rigorously consistent manner from the 
-time-dependent neutron balance equation.
-
-Due to the factorization of the flux into a shape and an amplitude, the latter two variables are nonlinearly coupled. Ott in \cite{Ott_1966} first investigated the coupling of 
-shape and amplitude in a quasi-static nature, but did not include the time derivative in the shape equation. Later, in \cite{Ott_1969}, Ott incorporated the time derivative of 
-shape in the equation, yielding better results but requiring a fixed-point approach to resolve the nonlinear coupling between the amplitude and shape equations. This also 
-led to the technique's name: the Improved Quasi-Static method although one may argue that such a denomination does not make it immediately clear that the technique rigorously 
-solves a time-dependent problem.
-
-Nonlinear problems are solved in an iterative manner, typically using either a fixed-point (Picard) approach or Newton's iterations. Sissaoui et al. \cite{Sissaoui_1995}, 
-Koclas et al. \cite{Koclas_1996}, Devooght et al. \cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their IQS simulations, 
-the main difference among them being their criteria for convergence.  Devooght et al. in \cite{Devooght_1984} proposed a Newton-based iteration technique. Another approach 
-is to linearize the equations, so that no nonlinear iteration is necessary. IQS can be linearized using the IQS Predictor-Corrector method (\iqspc) \cite{dulla2006}. 
-\iqspc entails evaluating the flux equation then correcting its amplitude using an amplitude evaluation. This method has proven to be effective for problems 
-requiring a significant amount of shape updates when IQS is implemented \cite{Dulla2008}. Dulla also investigates shape time adaptation with IQS with Caron in \cite{caron2017}.
-
-For use in multiphysics reactor physics applications, the interplay of the IQS solution technique (neutronics) and other physics components (e.g., fuel temperature) needs to be 
-resolved effectively. Meneley and Ott first implemented IQS into the one-dimensional fast-reactor code QX1 \cite{Meneley_1971}. This implementation does simple adiabatic heat up 
-with tabular cross-section feedback. Later, Kereszt\'{u}ri et al. describe IQS implementation into KIKO3D, a three-dimensional pressurized water reactor code, in \cite{KIKO3D_2003}. 
-This implementation computes fuel heat transfer and thermal hydraulic feedback. Ikeda and Takeda developed in the nodal expansion method code EPISODE which uses \iqspc with adiabatic 
-fuel heat and Doppler feedback \cite{Ikeda_2001}. \jcr{here you just say who did IQS multiphysics before. but how did they do it? are there differences between these implementations,
-is there something lacking that your paper will provide insight to?}
-
-This paper (i) discusses the different nonlinear iteration techniques for IQS and tests the rigor of their implementation, (ii) investigates high-order temporal discretization of the shape equation,
-and (iii) analyzes IQS coupled with adiabatic heat up and cross-section feedback. 
-%
-Prior IQS publications do not go beyond first-order time discretization for the shape equation. Here, we test IQS with higher-order schemes (up to fourth order) and investigate the method's 
-effectiveness with such temporal discretizations. 
+The improved quasi-static method (IQS) is a numerical technique devised for nuclear reactor transient analysis.
+It involves factorizing the neutron flux solution into a time-only-dependent component, the amplitude, and a space- and time-dependent component, the shape \cite{Ott_1966,Devooght_1984,Monier_diss,Sissaoui_1995,Dulla2008}.
+The amplitude solution satisfies the point reactor kinetic equations (PRKE) where the shape solution has been used to generate the PRKE coefficients (reactivity, effective fraction of delayed neutrons, mean generation time).
+The shape solution satisfies a modified time-dependent neutron balance equation.
+The rationale for the IQS method lies in the assumption that the shape function is weakly dependent on time.
+Therefore, it is expected that the modified time-dependent neutron balance equations be less stiff than the original time-dependent neutron balance equations for the flux.
+As a result, the shape may not require to be solved for at the same frequency as the amplitude.
+Instead, shape is evaluated only on larger macro-time steps, which is expected to yield wall clock savings, especially in multi-dimensional geometries.
+The PRKE form a small system of ordinary differential equations (ODE) and solving them on a fine temporal grid is not a computational burden.
+As opposed to the standard PRKE approach, whereby a shape function is selected, typically once for an entire transient without updates, the IQS technique is obtained in a rigorously consistent manner from the time-dependent neutron balance equation.
+
+Due to the factorization of the flux into a shape and an amplitude, the latter two variables are nonlinearly coupled.
+Ott in \cite{Ott_1966} first investigated the coupling of shape and amplitude in a quasi-static nature, but did not include the time derivative in the shape equation.
+Later, in \cite{Ott_1969}, Ott incorporated the time derivative of shape in the equation, yielding better results but requiring a fixed-point approach to resolve the nonlinear coupling between the amplitude and shape equations.
+This also led to the technique's name: the Improved Quasi-Static method although one may argue that such a denomination does not make it immediately clear that the technique rigorously solves a time-dependent problem.
+
+Nonlinear problems are solved in an iterative manner, typically using either a fixed-point (Picard) approach or Newton's iterations.
+Sissaoui et al.
+\cite{Sissaoui_1995}, Koclas et al.
+\cite{Koclas_1996}, Devooght et al.
+\cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their IQS simulations, the main difference among them being their criteria for convergence.
+Devooght et al.
+in \cite{Devooght_1984} proposed a Newton-based iteration technique, but this type of iteration in IQS is not investigated in this paper.
+Another approach is to linearize the equations, so that no nonlinear iteration is necessary.
+IQS can be linearized using the IQS Predictor-Corrector method (\iqspc) \cite{dulla2006}.
+\iqspc entails evaluating the flux equation then correcting its amplitude using an amplitude evaluation.
+This method has proven to be effective for problems requiring a significant amount of shape updates when IQS is implemented \cite{Dulla2008}.
+Dulla also investigates time adaptation of the shape evaluations with IQS with Caron in \cite{caron2017}.
+
+For use in multiphysics reactor physics applications, the interplay of the IQS solution technique (neutronics) and other physics components (e.g., fuel temperature) needs to be resolved effectively.
+Meneley and Ott first implemented IQS into the one-dimensional fast-reactor code QX1 \cite{Meneley_1971}.
+This implementation does simple adiabatic heat up with tabular cross-section feedback.
+Later, Kereszt\'{u}ri et al.
+describe IQS implementation into KIKO3D, a three-dimensional pressurized water reactor code, in \cite{KIKO3D_2003}.
+This implementation computes fuel heat transfer and thermal hydraulic feedback.
+Ikeda and Takeda developed in the nodal expansion method code EPISODE which uses \iqspc with adiabatic fuel heat and Doppler feedback, as well as full neutronics and thermal-hydraulics coupling \cite{Ikeda_2001}.
+These multiphysics applications range in complexity and problem type, but seem to neglect the advantage of including the non-neutronics physics into the quasi-static process.
+Although Meneley and Ikeda hint at using an intermediate time-scale for fuel temperature evaluation, their numerical experiments show that the evaluation is on the same scale as shape.
+Additionally, these applications only use heat source values at the shape time-scale from the neutronics calculation, while much more information about the transient heat source can be extracted from the amplitude scale.
+
+This paper (i) discusses the different nonlinear iteration techniques for IQS and tests the rigor of their implementation, (ii) investigates high-order temporal discretization of the shape equation, and (iii) analyzes IQS coupled with adiabatic heat up and cross-section feedback.
+Prior IQS publications do not go beyond first-order time discretization for the shape equation.
+Here, we test IQS with higher-order schemes (up to fourth order) and investigate the method's effectiveness with such temporal discretizations.
 Step doubling time adaptation is also implemented to test IQS and \iqspc performance with adaptation of shape/flux evaluation; although, Caron et al.
 in \cite{caron2017} performs a more comprehensive shape time adaptation analysis.
-Step doubling is also applied to amplitude evaluation and amplitude-based adaptation was applied to reactivity updates.
-In regards to multiphysics simulation, we develop a semi-analytical approach to evaluate adiabtic fuel temperature and introduce an intermediate time scale for temperature evaluation.
-%
+Step doubling is also applied to amplitude evaluation.
+Regarding multiphysics simulation, we develop a semi-analytical approach to evaluate adiabatic fuel temperature and introduce an intermediate time scale for temperature evaluation.
 To test IQS and \iqspc nonlinear iteration, time step convergence, and time adaptation performance, four different test cases are employed: two problems are purely neutronics and involve varying magnitudes of reactor size and complexity; two problems involve temperature feedback and test an intermediate time scale for temperature evaluation.
@@
 the error convergence rate of high order temporal discretization schemes for shape, higher order interpolation of the parameters will be required.
+Third order Runge-Kutta discretization with step doubling adaptation was used for the PRKE evaluation, to insure a negligibly small error in the amplitude for a given set of parameters.
 
@@
 As noted in the previous section, shape-PRKE equations form a nonlinear system and must be solved in a iterative manner.  Over each macro time step,  one can 
-use the latest end-time shape iterate to compute/interpolate the PRKE coefficents over the micro time step intervals. 
+use the latest end-time shape iterate to compute/interpolate the PRKE coefficients over the micro time step intervals. 
 Sissaoui et al. from \cite{Sissaoui_1995}, Koclas et al. from \cite{Koclas_1996}, Devooght et al. from \cite{Devooght_1984}, and Monier from \cite{Monier_diss} all use 
@@
 the shape is essential for preserving the uniqueness condition ($K_{n+1} = K_0$); this condition is not inherently conserved, even when the 
-solution has converged from shape iterations. Devooght in \cite{Devooght_1984} uses a Newton-SOR iteration where the residual of the shape function 
-evaluation is the convergence criteria and next iteration's 
-solution is computed using Newton-Raphson method.
+solution has converged from shape iterations.
 
 These techniques are by no means an exhaustive list of the possible iteration techniques for IQS. Dulla et al. in \cite{Dulla2008} provide an 
-in-depth analysis of the fixed point iteration technique most similar to Sissaoui and Koclas, involving convergence rates and solution results.  
-However, no comprehensive analysis of iteration techniques exists, comparing both Newton and fixed-point convergence rates.  The following describes 
+in-depth analysis of the fixed-point iteration technique most similar to Sissaoui and Koclas, involving convergence rates and solution results.  
+However, no comprehensive analysis of iteration techniques exists. The following describes 
 each iteration convergence criterion investigated in this paper:
-\comm{you never show Newton results, do you?}
-
-\answer{zmp}{no, never figured out how to do it, plus Devooght actually uses GQS for his results.}
-
-\comm{there are parts of the paper where it feels like the reader is going to expect Newton-based results\\
-(I feel like you said so in the abstract). You should amend these statements}
 
@@
 The Predictor-Corrector version of the IQS method (\iqspc) is a linearized version of IQS, where no nonlinear iterations are required.
-The method relies on the shape and amplitude factorization of the flux and uniqueness condition. The PRKE derivation is 
-identical to that of the standard (nonlinear) version of the IQS scheme, in the sense that shape solutions at the beginning 
-and end times of the macro time interval are used (with interpolation in between).
-However, the manner in which the shape function is obtained is different.  In the \iqspc version, the flux equations 
-(not the shape equations) are first solved (represented by \eqts{eq:flux}{eq:precursor}) in order to obtain a 
-{\it predicted} flux solution at the end of the macro time step. This predicted flux is then converted to a shape by 
-normalizing it as follows:
+The method relies on the shape and amplitude factorization of the flux and uniqueness condition.
+The PRKE derivation is identical to that of the standard (nonlinear) version of the IQS scheme, in the sense that shape solutions at the beginning and end times of the macro time interval are used (with interpolation in between).
+However, the manner in which the shape function is obtained is different.
+In the \iqspc version, the flux equations (not the shape equations) are first solved (represented by \eqts{eq:flux}{eq:precursor}) in order to obtain a {\it predicted} flux solution at the end of the macro time step.
+This {\it predicted} flux is then converted to a shape by normalizing it as follows:
+
 \be
@@
 
-The PRKE parameters are then computed with this shape using \eqtss{eq:rmb}{eq:l} and interpolated over the macro step, then the 
-PRKE ODE system is solved on the micro time scale.  With the newly computed amplitude, the shape is rescaled into a flux and the final 
-{\it corrected} flux is given by:
+The PRKE parameters are then computed with this shape using \eqtss{eq:rmb}{eq:l} and interpolated over the macro step, then the PRKE ODE system is solved on the micro time scale.
+With the newly computed amplitude, the shape is rescaled into a flux and the final {\it corrected} flux is given by:
+
 \be
@@
 
-The advantage to the predictor-corrector method is that no nonlinear iterations are necessary in this method; it is simpler to 
-implement and can be faster than the standard IQS.  Ikeda et al. in \cite{Ikeda_2001} and Goluoglu et al. in \cite{Goluoglu_2001} 
-both use \iqspc for complex, three-dimensional problems.  Their results prove \iqspc to be effective.  Dulla et al. in \cite{Dulla2008} 
-also describes an in depth comparison of \iqspc with traditional IQS, which originally brought the method to light.
+The advantage to the predictor-corrector method is that no nonlinear iterations are necessary in this method; it is simpler to implement and can be faster than the standard IQS.
+Ikeda et al.
+in \cite{Ikeda_2001} and Goluoglu et al.
+in \cite{Goluoglu_2001}
+both use \iqspc for complex, three-dimensional problems.
+Their results prove \iqspc to be effective.
+Dulla et al.
+in \cite{Dulla2008} also describes an in-depth comparison of \iqspc with traditional IQS, which originally brought the method to light.
 
@@
 
-Here, we consider multiphysics reactor physics simulations.  Fuel temperature feedback mechanisms are included 
-using an adiabatic heat conservation model given in \eqt{eq:temp}.  
+Here, we consider multiphysics reactor physics simulations.  Fuel temperature feedback mechanisms are included using an adiabatic heat conservation model given in \eqt{eq:temp}.  
 \be
@@
 
-A standard time-implicit solver for the heat equation would simply employ the flux values at 
-the extremities of a time step interval.  However, in an IQS neutronic solve, much more information about 
-space/time  distribution of the flux is known: the shape distributions are typically known at the beginning and end
-of a macro time step interval but the amplitude is known on the micro-step time scale, providing
-a richer space/time information for the neutron flux over the macro-step interval.  
-Thus, it is possible to solve for temperature in \eqt{eq:temp} using a semi-analytical approach, 
-shown in \eqt{eq:temp_an}.
-
+A standard time-implicit solver for the heat equation would simply employ the flux values at the extremities of a time step interval.
+However, in an IQS neutronic solve, much more information about space/time distribution of the flux is known: the shape distributions are typically known at the beginning and end of a macro time step interval but the amplitude is known on the micro-step time scale, providing a richer space/time information for the neutron flux over the macro-step interval.
+Thus, it is possible to solve for temperature in \eqt{eq:temp} using a semi-analytical approach, shown in \eqt{eq:temp_an}.
 \be
@@
 where $S_f$ is the fission source computed using the shape solution ($S_{f,n}=(\nu\Sigma_f)_n\varphi_n$). With
-$\theta=1$, this yields the implicit Euler method, while with $\theta=\tfrac 1 2$, we obtain the Crank 
-Nicolson technique. In doing so, however, no micro-scale temporal information for the amplitude is used.
+$\theta=1$, this yields the implicit Euler method, while with $\theta=\tfrac 1 2$, we obtain the Crank-Nicolson technique.
+In doing so, however, no micro-scale temporal information for the amplitude is used.
 One can also solve the precursors' equations analytically, yielding:
@@
 
-Further enhancements to the performance of the IQS methods can be gained by using time adaptation 
-(or time step control) in order to increase or reduce the macro time step size for the shape evaluation, 
-depending on error estimates. A step-doubling technique is chosen as the time adaptation technique \cite{NumC}. 
-The step doubling technique involves estimating the local error for a certain time step by taking the 
-difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps 
-($\varphi^g_{\Delta t/2}$). Note: $\varphi$ is changed to $\phi$ in the case of the \iqspc tehcnique.
-%
+Further enhancements to the performance of the IQS methods can be gained by using time adaptation (or time step control) in order to increase or reduce the macro time step size for the shape evaluation, depending on error estimates.
+A step-doubling technique is chosen as the time adaptation technique \cite{NumC}.
+The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps ($\varphi^g_{\Delta t/2}$).
+Note: $\varphi$ is changed to $\phi$ in the case of the \iqspc technique.
 The relative error is computed as follows:
@@
 where $q$ is the convergence order of the time integration scheme being used and $s\simeq 0.8$ is a safety factor.
-If the error is larger than the user-specified tolerance, the time step is rejected and new (smaller) macro time 
-step size is estimated using \eqt{eq:edt2} as well. This process can be visualized by \figs{fig:dt2_1}{fig:dt2_2},
-where a step involves a full convergence of shape, amplitude, and any multiphysics on the respective time step.
-
-To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to 
-direct temporal discrezation of the flux equation (no IQS), the standard IQS method, and the \iqspc version. 
-Each of these methods will be applied to several diffusion problems; the number of time steps taken and 
-the resulting error will be used to compare the methods.
+If the error is larger than the user-specified tolerance, the time step is rejected and new (smaller) macro time step size is estimated using \eqt{eq:edt2} as well.
+This process can be visualized by \figs{fig:dt2_1}{fig:dt2_2}, where a step involves a full convergence of shape, amplitude, and any multiphysics on the respective time step.
+
+To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to direct temporal discretization of the flux equation (no IQS), the standard IQS method, and the \iqspc version.
+Each of these methods will be applied to several diffusion problems; the number of time steps taken and the resulting error will be used to compare the methods.
 
@@
 
-Both the standard IQS and \iqspc methods have been implemented in the Multiphysics Object-Oriented
-Simulation Envrionement, MOOSE \cite{}, as part of its radiation transport solver, Rattlesnake \cite{}.
+Both the standard IQS and \iqspc methods have been implemented in the Multiphysics Object-Oriented Simulation Environment, MOOSE \cite{moose}, as part of its radiation transport solver, Rattlesnake \cite{wang2013}.
 All multi-dimensional results presented here have been computed using the diffusion solver of Rattlesnake.
-MOOSE calculation flows naturally allows for a relatively straightforward implementation of the IQS methodologies.
-MOOSE's nonlinear solvers (fixed-point iterations and Jacobian-free Newton-Krylov) are naturally well
-suited to handle the nonlinear solve of the tradiationa IQS method during a macro time step. \iqspc's implementation
-is MOOSE is even simpler. We briefly describe the their implementation next.
-
-First, it is important to note that MOOSE provides entry points in its calculation flow at various moments:
-\begin{itemize}
-\item during a time step, with a pre-step and post-step call;
-\item at each nonlinear iteration;
-\item at each linear iteration.
-\end{itemize}
-At each of these instants, an auxiliary calculations may be initiated (e.g., a convergence check, a full PRKE
-solve, areactivity update, ...).
-
-\jcr{in plain English, can you describe how both iqs methods were implemented so that we give some recognition
-to moose a bit. thanks}.
+MOOSE calculation flows naturally, which allows for a relatively straightforward implementation of the IQS methodologies.
+MOOSE's nonlinear solvers (fixed-point iterations and Jacobian-free Newton-Krylov) are well suited to handle the nonlinear solve of the traditional IQS method during a macro time step.
+\iqspc's implementation is MOOSE is even simpler.
+We briefly describe the implementation next.
+
+First, the IQS shape equation was represented by the gathering of appropriate kernels.
+Most of the shape equation kernels are the same as the flux kernels already implemented in Rattlesnake, so only two kernels were modified.
+\eqt{eq:kernels} shows which kernels needed to be written to make sure MOOSE is solving the shape equation.
+The PRKE ODE was written as a sub-function in the Transient executioner, which progresses the system in time and tells when to evaluate the main shape/flux solve.
+This executioner has breakpoints that allows specified calls to the PRKE, PRKE parameter evaluation, and other multiphysics.
+The PRKE parameters are written as user-objects that loop over elements to execute the proper integration when the executioner calls them.
+Similarly, the semi-analytic treatment of fuel temperature and precursors are auxiliary kernels that do a node-by-node evaluation of the variable when the executioner calls it.
+For a more detailed description of the MOOSE application of IQS, see \cite{PrinceTR2016}.
+
+\begin{align}
+\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
+& - \underbrace{\frac{1}{v^g} \boxed{\overbrace{\frac{1}{p}\frac{dp}{dt}}^{From PRKE}}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
+\label{eq:kernels}
+\end{align}
+
+%First, it is important to note that MOOSE provides entry points in its calculation flow at various moments:
+%\begin{itemize}
+%\item during a time step, with a pre-step and post-step call;
+%\item at each nonlinear iteration;
+%\item at each linear iteration.
+%\end{itemize}
+%At each of these instants, an auxiliary calculations may be initiated (e.g., a convergence check, a full PRKE
+%solve, a reactivity update, ...).
 
@@
 We also investigate the use of higher-temporal discretization for the shape equations.
-Both series of tests are carried out using neutronics-only problems.  The first test case uses a one-dimensional problem.  
+Both series of tests are carried out using neutronics-only problems.
+The first test case uses a one-dimensional problem.
 The second test is from the ANL Benchmark Problem Book (BPB) \cite{ANL_BPB}.
@@
 
-This one-dimensional problem uses a 400-cm slab. This initial configuration is homogeneous and the transient is initiated by a perturbation in absorption cross section. 
+This one-dimensional problem uses a 400-cm slab.
+This initial configuration is homogeneous and the transient is initiated by a perturbation in absorption cross section.
 \fig{fig:slab} shows problem layout and \tbl{tab:1Dmat} contains the material properties; one-group values are used for this problem.
-Regions 2, 3, and 4 have linear ramp perturbations at different moments in time, \tbl{tab:1Dslope} shows the values of the absorption cross-section in each region at the 
-times of interest.  The values of $\Sigma_a$ between these times of interest vary linearly between the given values.
+Regions 2, 3, and 4 have linear ramp perturbations at different moments in time, \tbl{tab:1Dslope} shows the values of the absorption cross-section in each region at the times of interest.
+The values of $\Sigma_a$ between these times of interest vary linearly between the given values.
 
@@
 
-\fig{fig:1D} shows the reference flux solution at different instants as well as total power as a function of time.  The baseline reference solution 
-was computed by solving the flux equations (not the IQS formulation) using time-step control with a tight relative error tolerance of $10^{-12}$.  
-The problem being relatively large, the various zones are weakly coupled
-and the flux perturbation is not global. This baseline computation is used to compute the error of the other time discretization methods.  
-\fig{fig:1D_shape} shows the shape profile at the same instants during the transient as in \fig{fig:1D_flux};
-that solution was also obtained using time-step control with tight tolerance. Both flux solve and IQS solve yield
-the same answer, as expected.  
+\fig{fig:1D} shows the reference flux solution at different instants as well as total power as a function of time.
+The baseline reference solution was computed by solving the flux equations (not the IQS formulation) using time-step control with a tight relative error tolerance of $10^{-12}$.
+The problem being relatively large, the various zones are weakly coupled and the flux perturbation is not global.
+This baseline computation is used to compute the error of the other time discretization methods.
+\fig{fig:1D_shape} shows the shape profile at the same instants during the transient as in \fig{fig:1D_flux}; that solution was also obtained using time-step control with tight tolerance.
+Both flux solve and IQS solve yield the same answer, as expected.
 % Even though the magnitude of the shape varies less during the transient, its spatial distribution shows more variation than that 
@@
 \end{figure}
-\jcr{the editor will like complain that the figure does not have the same height.
-you must have saved them in matlab after playing with them.
-i recommend you save them with the size and aspect ratio as produced by matlab.}
 
@@
 \fig{fig:iter} gives the number of fixed-point iterations required to reach tolerance of $10^{-11}$ as a function of time during the transient.
-A theta-scheme is employed for the precurosrs equations (\jcr{recall the eq}).
+A theta-scheme is employed for the precursors equations (\eqt{eq:dnp_theta}).
 %The criteria listed in the legend correspond to the list as such:
@@
 %\end{enumerate}
-This figures shows that criteria based on the  $L^{\infty}$ norm of the shape, the $L^2$ norm of the shape, the reactivity change, and the amplitude have approximately the same convergence behavior, but the criteria
-based on the uniqueness constant $K$ never reaches the tight tolerance and runs up to the maximum number of iterations allowed (20 in this case). 
-\fig{fig:iter_err} shows the error in the factor $K$ saturates to different levels, whether the shape is rescaled or not. Rescaling (described by \eqt{eq:shape_scale}) 
+This figures shows that criteria based on the  $L^{\infty}$ norm of the shape, the $L^2$ norm of the shape, the reactivity change, and the amplitude have approximately the same convergence behavior, but the criteria based on the uniqueness constant $K$ never reaches the tight tolerance and runs up to the maximum number of iterations allowed (20 in this case).
+\fig{fig:iter_err} shows the error in the factor $K$ saturates to different levels, whether the shape is rescaled or not.
 %Rescaling shape more frequently helps the error.
@@
 \fig{fig:1D_conv} shows the convergence results for four different time-implicit discretization: backward difference formulae, from order 1, which is the same as the implicit Euler, to order 4.
-\jcr{give refs for BDF methods of any order, do not pick wikipedia!}. % as well as a 3rd-order implicit Runge-Kutta method, SDIRK33. \jcr{also give a ref}
-The plots show that IQS and \iqspc converge with the expected order (1 through 4) using BDF1 through BDF4. Recall that in the amplitude equations, the PRKE parameters
-are interpolated using the shape solutions. We used higher-order interpolants in time for the shape solution when a higher-order method was used
-(the convergence rates were degraded for higher-order BDF methods when the PRKE parameters were only linearly interpolated over the macro step). Lagrange
-and Hermite interpolants yielded the same results. In addition, we used semi-analytical integration of the precursors equations, along with the same higher-order
-interpolation of the shape values when a higher-order method was used. Without using the same order interpolants as the BDF formula, the error in the
-PRKE parameters and the precursors led to reduction in the observed convergence rate.
+For a detailed description on BDF methods, see \cite{Gear:2007}.
+The plots show that IQS and \iqspc converge with the expected order (1 through 4) using BDF1 through BDF4.
+Recall that in the amplitude equations, the PRKE parameters are interpolated using the shape solutions.
+We used higher-order interpolants in time for the shape solution when a higher-order method was used (the convergence rates were degraded for higher-order BDF methods when the PRKE parameters were only linearly interpolated over the macro step).
+Lagrange and Hermite interpolants yielded the same results.
+In addition, we used semi-analytical integration of the precursors equations, along with the same higher-order interpolation of the shape values when a higher-order method was used.
+Without using the same order interpolants as the BDF formula, the error in the PRKE parameters and the precursors led to reduction in the observed convergence rate.
 
 % The Third-order SDIRK method did not show third order convergence, but, through extensive testing, SDIRK shows non-convergent behavior for too stiff of problems.
-\jcr{I am removing sdirk, the point we want to make is that high-order methods do not buy anything for iqs. the sdirk issue is unrelated so why confuse the audience}
 
 The convergence rates for the straightforward discretization of the flux equations are also shown in \fig{fig:1D_conv} and they followed the same expected convergence rates.
-These flux-solve results are indicated by the legend ``Implicit Dis.''. It is worthy to note that, for lower-order discretizations, the error in the flux equations is always significanrtly higher 
-than that of the IQS results. However, as the order of the temporal scheme is increased, the gap between the discretization error in the flux equations and the IQS equations vanishes. 
-In previous publications, the shape solve in the IQS method always employed a lower-order temporal discretization
-(often implicit Euler). Our results seem to indicate that a high-order discretization 
-of the flux equations may reduce the usefulness of the IQS approach (at least for large domains with weakly coupled spatial regions).
+These flux-solve results are indicated by the legend ``Implicit Dis.''.
+It is worthy to note that, for lower-order discretizations, the error in the flux equations is always significantly higher than that of the IQS results.
+However, as the order of the temporal scheme is increased, the gap between the discretization error in the flux equations and the IQS equations vanishes.
+In previous publications, the shape solves with the IQS method always employed a lower-order temporal discretization (often implicit Euler).
+Our results seem to indicate that a high-order discretization of the flux equations may reduce the usefulness of the IQS approach (at least for large domains with weakly coupled spatial regions).
 %This paper shows an analysis of the first publicized application of IQS with higher than second order discretization, which exposed unforeseen properties of IQS.
@@
 
-%----------------------------------------------------------------------------------------
-\subsubsection{One-Dimensional Mini-Core Problem}
-\jcr{i do not understand the purpose of the entire section.
+%%----------------------------------------------------------------------------------------
+%\subsubsection{One-Dimensional Mini-Core Problem}
+%\jcr{i do not understand the purpose of the entire section.
 shouldn't the mini core be doen with bdf4 too, not just bdf1?
 and aren't the twigl results already such a demonstration so why not move on to twigl right away?}
-This problem is exactly the same as the previous one-dimensional problem, except the the core was reduced to 80 cm in length.
+%This problem is exactly the same as the previous one-dimensional problem, except the the core was reduced to 80 cm in length.
 The purpose of testing this problem is to determine if a more tightly coupled core will yield better performance for IQS.
@@
 \fig{fig:1Dmini_conv} shows that IQS performs significantly better with this example than the large core.
-
-\begin{figure}[!htbp]
-\centering
-\begin{subfigure}[b]{\linewidth}
-\centering
-\includegraphics[height=1.5in]{figures/1Dmini_flux.png}
-\caption{Flux profile at various times}
-\end{subfigure}
-\begin{subfigure}[b]{\linewidth}
-\centering
-\includegraphics[height=1.5in]{figures/1Dmini_power_base.png}
-\caption{Power profile over transient}
-\end{subfigure}
-\caption{Mini-core baseline flux and power distribution}
-\label{fig:1Dmini_flux}
-\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[height=1.5in]{figures/1Dmini_shape.png}
-\caption{Shape profile at various times for one-dimensional mini-core}
-\label{fig:1Dmini_shape}
-\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[height=1.5in]{figures/1Dmini_conv_IE.png}
-\caption{Time step convergence for one-dimensional mini-core with implicit Euler discretization}
-\label{fig:1Dmini_conv}
-\end{figure}
+%
+%\begin{figure}[!htbp]
+%\centering
+%\begin{subfigure}[b]{\linewidth}
+%\centering
+%\includegraphics[height=1.5in]{figures/1Dmini_flux.png}
+%\caption{Flux profile at various times}
+%\end{subfigure}
+%\begin{subfigure}[b]{\linewidth}
+%\centering
+%\includegraphics[height=1.5in]{figures/1Dmini_power_base.png}
+%\caption{Power profile over transient}
+%\end{subfigure}
+%\caption{Mini-core baseline flux and power distribution}
+%\label{fig:1Dmini_flux}
+%\end{figure}
+%
+%\begin{figure}[!htbp]
+%\centering
+%\includegraphics[height=1.5in]{figures/1Dmini_shape.png}
+%\caption{Shape profile at various times for one-dimensional mini-core}
+%\label{fig:1Dmini_shape}
+%\end{figure}
+%
+%\begin{figure}[!htbp]
+%\centering
+%\includegraphics[height=1.5in]{figures/1Dmini_conv_IE.png}
+%\caption{Time step convergence for one-dimensional mini-core with implicit Euler discretization}
+%\label{fig:1Dmini_conv}
+%\end{figure}
 
@@
 
-This benchmark problem originates from the ANL Benchmark Problem Book \cite{ANL_BPB}.  It is a 2D, 2-group reactor core model with no reflector region \cite{TWIGL_benchmark}.  
-% This example is meant to be of progressive complexity from the previous example.  
-The core is smaller (hence the spatial regions are more tightly coupled) and the spatial variation of the flux
-varies moderately over the transient. 
+This benchmark problem originates from the ANL Benchmark Problem Book \cite{ANL_BPB}.
+It is a 2D, 2-group reactor core model with no reflector region \cite{TWIGL_benchmark}.
+The core is smaller (hence the spatial regions are more tightly coupled) and the spatial variation of the flux varies moderately over the transient.
 Therefore, IQS is expected to perform significantly better than an implicit flux solve.
@@
 The observed convergence rates are indicated in the legend.
-They show orders 1 for IE and 2 for BDF2 for the flux solves. However, the convergence rates for IQS show higher values,
-order 2 for IE solves for the shape, and above 2.5 for BDF2 solve of the shape.
+They show orders 1 for IE and 2 for BDF2 for the flux solves.
+However, the convergence rates for IQS show higher values, order 2 for IE solves for the shape, and above 2.5 for BDF2 solve of the shape.
 When there is moderate variation in the shape over the transient, the PRKE parameters are accurately estimated, and the fine-scale PRKE solve provides a well resolved amplitude solution.
@@
 This section describes two multiphysics examples with adiabatic heat up: the LRA benchmark from the  ANL Benchmark Problem Book \cite{ANL_BPB} 
-and a modelisation of an experiment carried out at the Transient Reactor Test Facility, TREAT \jcr{provide a reference for this model; it could be an INL report or one of your conference paper}. 
+and a modelisation of an experiment carried out at the Transient Reactor Test Facility, TREAT \cite{mammoth, Tran15}. 
 % These examples are of increased complexity from the previous kinetics examples.
@@
 
-The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up 
-and Doppler feedback in thermal reactor.  It is a super prompt-critical transient. 
-The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.  The spatial discretization uses continuous finite elements
-with first-order Lagrangian basis functions.
+The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor.
+It is a super prompt-critical transient.
+The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.
+The spatial discretization uses continuous finite elements with first-order Lagrangian basis functions.
 The mesh consists of blocks $11\times 11$ with five uniform refinements, totaling $165,165$ elements and $124,609$ nodes.
-Three different temporal techniques are
-applied: implicit discretization of the flux equation, IQS, and \iqspc.
-Crank-Nicolson \jcr{check spelling for CN emthod throughout} time discretization scheme is used for the diffusion evaluation of each technique.
-Third order Runge-Kutta discretization with step doubling adaptation was used for the PRKE evaluation.
-\jcr{actually, you never say what is used for the PRKE solve.
-this should be moved way up because it applies to everything!}
+Three different temporal techniques are applied: implicit discretization of the flux equation, IQS, and \iqspc.
+Crank-Nicolson time discretization scheme is used for the diffusion evaluation of each technique.
 The performance of IQS and the temperature updates were measured by its improvement in accuracy at peak power over the implicit discretization method.
@@
 Furthermore, there does seem to be an ideal number of temperature updates to optimize execution time: IQS only needs one and \iqspc seems to be ideal at 4 updates.
-This discrepancy in the number of updates shows that a adaptive type implementation of the updates would be ideal, and could enforce a constant error over the transient.
+This discrepancy in the number of updates shows that an adaptive type implementation of the updates would be ideal, and could enforce a constant error over the transient.
 It is also important to compare the error of implicit discretization with IQS at one update and \iqspc at 4 updates.

### 2017-09-13 doc/Papers/IQS_convergence/IQS_convergence.tex
Adding conclusion

@@
 
-The Transient-15 model uses a adiabatic temperature feedback mechanism, similar to the one explored by the LRA.
+The Transient-15 model uses an adiabatic temperature feedback mechanism, similar to the one explored by the LRA.
 \eqt{eq:temp2} describes the heat up of the fuel.
@@
 
-In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time step, implicit discretization, IQS with one and 5 temperature updates per macro step, and \iqspc with one and 5 updates.
+In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time-step, implicit discretization, IQS with one and 5 temperature updates per macro step, and \iqspc with one and 5 updates.
 \fig{fig:Tran15_profile} shows the baseline power and temperature profile for the Transient-15 example.
@@
 
+The goal of this paper was to continue the investigation and development of the improved quasi-static method for the optimization of transient reactor simulations.
+In pursuit of this goal, this paper (i) discussed the different nonlinear iteration techniques for IQS and tests the rigor of their implementation, (ii) investigates high-order temporal discretization of the shape equation, and (iii) analyzes IQS coupled with adiabatic heat up and cross-section feedback.
+
+IQS is a nonlinear system of equations; therefore, fixed-point iteration was used to evaluate shape and amplitude.
+Investigating these iteration techniques is important for understanding the behavior of IQS, as well as determining the most appropriate technique for optimal performance.
+The iteration techniques were applied to a one-dimensional prototype problem for testing.
+For fixed-point iteration, five different criteria were tested:
+\begin{enumerate}
+\item $L^\infty$ norm of the change in shape between iterations
+\item $L^2$ norm of the change in shape between iterations
+\item Difference in reactivity between iterations
+\item Difference in amplitude between iterations
+\item IQS uniqueness consistency criteria
+\end{enumerate}
+The results showed that criteria 1-4 had relatively equivalent convergence behavior.
+However, iteration with the criteria 5 could not converge without a highly-accurate analytical treatment of the precursor equation.
+This criteria proved to be the most rigorous and thorough convergence criteria and is recommended for any application of IQS.
+
+IQS effectiveness with high-order temporal discretization schemes was also investigated.
+Implicit Euler and BDF discretization schemes were applied to the one-dimensional prototype problem.
+IQS showed proper error convergence up through fourth order BDF schemes.
+Implicit Euler and second order BDF schemes were applied to the TWIGL benchmark, where IQS showed normal error convergence.
+Step doubling adaptation was also applied to TWIGL, where IQS performed impressively, reducing the number of diffusion evaluations considerably.
+The Crank-Nicholson scheme was applied to the LRA benchmark with IQS.
+IQS again showed expected second order convergence and was able to reduce the number of diffusion evaluations with time adaptation by more than a factor of 6.
+The results of the TWIGL and LRA benchmarks proved that IQS is capable of proper convergence and time adaptation behavior for more complex problems.
+However, with high-order schemes, IQS becomes less and less effective, demonstrating almost the same performance as implicit discretization at fourth order discretization.
+
+Full transient reactor simulation requires the implementation of multiphysics solution methods.
+In order to test IQS with these types of simulations, it was implemented into the MOOSE/Rattlesnake framework.
+The LRA benchmark and a full core TREAT model were evaluated as test cases for multiphysics treatment.
+These examples involve a adiabatic heat up of the fuel with Doppler broadening feedback of cross-sections.
+The evaluation of temperature was integrated in the quasi-static process by introducing an intermediate time scale for temperature and PRKE parameter evaluation.
+The results of the LRA benchmark showed that quasi-static approach to temperature greatly improved the accuracy for a given time step size.
+However, the increase in computation time due to the extra temperature evaluations was significant.
+Therefore, for this benchmark, IQS with one update and IQS P-C with four updates is optimal.
+The results of the TREAT example showed similar behavior, except the updates only produced a marginal increase in accuracy.
+These examples' results show that performing a variant number of updates during the transient could further improve the performance of this quasi-static process.
+In conclusion, IQS and the quasi-static treatment of temperature can significantly improve the computational efficiency for burdensome full transient reactor simulations.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-09-25 doc/Papers/IQS_convergence/IQS_convergence.tex
finished reviewing the whole paper. almost ready to submit

--- 2017-09-13 doc/Papers/IQS_convergence/IQS_convergence.tex
@@
 the error convergence rate of high order temporal discretization schemes for shape, higher order interpolation of the parameters will be required.
-Third order Runge-Kutta discretization with step doubling adaptation was used for the PRKE evaluation, to insure a negligibly small error in the amplitude for a given set of parameters.
+Third order implicit Runge-Kutta discretization with time step adaptation was used for the PRKE evaluation, to insure a negligibly small error in the amplitude solution.
 
@@
 All multi-dimensional results presented here have been computed using the diffusion solver of Rattlesnake.
-MOOSE calculation flows naturally, which allows for a relatively straightforward implementation of the IQS methodologies.
+The sequence of
+calculations in MOOSE (temporal loops, nonlinear loops, linear solves) with appropriate entry points at any level in the simulation allows for a relatively straightforward implementation of the IQS methodologies.
 MOOSE's nonlinear solvers (fixed-point iterations and Jacobian-free Newton-Krylov) are well suited to handle the nonlinear solve of the traditional IQS method during a macro time step.
-\iqspc's implementation is MOOSE is even simpler.
+\iqspc's implementation in MOOSE is even simpler.
 We briefly describe the implementation next.
 
-First, the IQS shape equation was represented by the gathering of appropriate kernels.
-Most of the shape equation kernels are the same as the flux kernels already implemented in Rattlesnake, so only two kernels were modified.
-\eqt{eq:kernels} shows which kernels needed to be written to make sure MOOSE is solving the shape equation.
-The PRKE ODE was written as a sub-function in the Transient executioner, which progresses the system in time and tells when to evaluate the main shape/flux solve.
+First, the IQS shape equation was represented by the gathering of appropriate finite ELement kernels.
+Most of the shape equation kernels are the same as the flux kernels already implemented in Rattlesnake for its standard diffusion (or transport) solution, and only two kernels had to be modified.
+\eqt{eq:kernels} shows which kernels needed to be written or amended (denoted by an $\star$) to solve the shape equation in MOOSE.
+The PRKE ODE was written as a sub-function in the Transient executioner, which advances the simulation in time and determines when to perform a shape (or flux for \iqspc) solve.
 This executioner has breakpoints that allows specified calls to the PRKE, PRKE parameter evaluation, and other multiphysics.
-The PRKE parameters are written as user-objects that loop over elements to execute the proper integration when the executioner calls them.
-Similarly, the semi-analytic treatment of fuel temperature and precursors are auxiliary kernels that do a node-by-node evaluation of the variable when the executioner calls it.
-For a more detailed description of the MOOSE application of IQS, see \cite{PrinceTR2016}.
+The PRKE parameters are written as user-objects, looping over cells to evaluate them.
+Similarly, the semi-analytic treatment of fuel temperature and precursors are auxiliary kernels that do a node-by-node evaluation of the variable upon requests of the transient executioner.
+Additional information pertaining to these MOOSE objects can be found in \cite{PrinceTR2016}.
 
 \begin{align}
-\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
-& - \underbrace{\frac{1}{v^g} \boxed{\overbrace{\frac{1}{p}\frac{dp}{dt}}^{From PRKE}}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
+\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Prompt Production Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Scattering Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Diffusion Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Reaction Kernel} \nonumber \\
+& - \underbrace{\frac{1}{v^g} \boxed{\overbrace{\frac{1}{p}\frac{dp}{dt}}^{From PRKE}}\varphi^g}_{IQS Reaction Kernel\ ^\star}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Precursor Kernel\ ^\star}
 \label{eq:kernels}
@@
 %---------------------------------------------------------------------------------------%
-\subsection{One-Dimensional Problem}
+\subsection{A One-Dimensional Problem}
 %---------------------------------------------------------------------------------------%
@@
 
-This section describes two multiphysics examples with adiabatic heat up: the LRA benchmark from the  ANL Benchmark Problem Book \cite{ANL_BPB} 
+This section describes two multiphysics examples with adiabatic heat-up: the LRA benchmark from the  ANL Benchmark Problem Book \cite{ANL_BPB} 
 and a modelisation of an experiment carried out at the Transient Reactor Test Facility, TREAT \cite{mammoth, Tran15}. 
@@
 The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor.  
+\jcr{already said before}
 It is a super prompt-critical transient.
 The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.
-The spatial discretization uses continuous finite elements with first-order Lagrangian basis functions.
+The spatial discretization uses continuous finite elements with first-order Lagrangian basis functions \jcr{move up and say only one.
+I believe all rattlesnake results use FEM order 1}.
 The mesh consists of blocks $11\times 11$ with five uniform refinements, totaling $165,165$ elements and $124,609$ nodes.
-Three different temporal techniques are applied: implicit discretization of the flux equation, IQS, and \iqspc.
-Crank-Nicolson time discretization scheme is used for the diffusion evaluation of each technique.
-The performance of IQS and the temperature updates were measured by its improvement in accuracy at peak power over the implicit discretization method.
+Three different temporal solution techniques are used: implicit discretization of the flux equations, IQS formulation, and \iqspc.
+Crank-Nicolson time discretization scheme is used for the space-time equations.
+\jcr{when you recall all common numerical options, make sure you recall that the PRKE ODE are done with high-order implicit RK.
+Put all info that is common to all results in 1 location at the beginning}.
+The performance of IQS with temperature feedback is assessed by comparing the solution accuracy at peak power.
 
@@
 %----------------------------------------------------------------------------------------
-\subsubsection{LRA Results}
+\subsubsection{IQS Results for the LRA Benchmark without Time Adaptivity}
 
 This section shows the time step error convergence of IQS for the LRA benchmark, as well as the effect of the intermediate temperature time scale.
-\fig{fig:lra_bad} is an error convergence plot comparing the three techniques where temperature is evaluated only on the macro step (1 temperature update).
-\fig{fig:lra_mpconv} is an error convergence plot comparing the three techniques where temperature is evaluated 5 times within a macro step (5 temperature updates).
+\fig{fig:lra_bad} is an error convergence plot comparing the three tempoeral solution techniques where temperature is evaluated only once per macro step (1 temperature update).
+\fig{fig:lra_mpconv} is an error convergence plot comparing the same three techniques when temperature is evaluated 5 times within a macro step (5 temperature updates).
 Finally, \fig{fig:mp} shows the effect of various temperature updates.
-The dashed lines correspond to implicit discretization at different flux step sizes, while the IQS macro step size is kept constant.
+The dashed lines correspond to implicit flux discretization at different step sizes, while the IQS macro step size is kept constant.
 
@@
 The convergence plots show that updating temperature and the PRKE parameters within a macro step has a significant effect on the performance of IQS.
-With only one update, IQS was only slightly better than implicit discretization, implicit discretization required about 150\% more time steps than IQS for the same error.
-While 5 temperature updates showed a much more significant IQS performance, implicit discretization required about 400\% more time steps than IQS for the same error.
-\fig{fig:mp} shows that error has a convergent behavior for the number of temperature updates.
-This convergence makes sense because temperature can only be so accurate before the error in shape is dominating.
+With only one update, IQS was only slightly better than the implicit flux discretization which required about 150\% \jcr{2.5 more then?
+I think you mean +50\%}more time steps than IQS for the same error.
+While 5 temperature updates showed a much improved performance for IQS, with the implicit flux discretization required about 400\% more time steps than IQS for the same error level.
+\fig{fig:mp} shows that the solution converges with an increase in the number of temperature updates.
+%
 Table \ref{tab:ndiff_lra} shows the run time results for the implicit discretization calculations.
-The number of GMRES linear iterations is included because it is proportional measure of the computational effort.
+The number of GMRES linear iterations is included because it is a proportional measure of the computational effort.
 Tables \ref{tab:iqs_lra} and \ref{tab:iqspc_lra} present the IQS run-times with various numbers of temperature updates.
 These run-times are based on total alive time of the execution where the diffusion evaluation is distributed over 24 processors.
-These run-times show a marginal performance for IQS and impressive performance for \iqspc.
+These run-times show a marginal performance for IQS and superior performance for \iqspc.
 Some of the execution times were able to decrease from implicit discretization with the same number of macro steps because IQS is better equipped to resolve the nonlinearity between temperature and amplitude.
-Furthermore, there does seem to be an ideal number of temperature updates to optimize execution time: IQS only needs one and \iqspc seems to be ideal at 4 updates.
+Furthermore, there seems to be an optimal number of temperature updates to minimize execution time for a given accuracy.
+%: IQS only needs one and \iqspc seems to be ideal at 4 updates.
 This discrepancy in the number of updates shows that an adaptive type implementation of the updates would be ideal, and could enforce a constant error over the transient.
 It is also important to compare the error of implicit discretization with IQS at one update and \iqspc at 4 updates.
-IQS shows an error comparable to implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by -34.1\%.
+IQS shows an error comparable to implicit discretization at $\Delta t = 0.002$ \jcr{where's that table with that time step?
+or do you mean it's not there and you just talk about results that are not in the paper?
+which would be a strange way of conveying information}, signifying an actual increase in runtime by -34.1\%.
 \iqspc shows an error less than implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by $<-34.9\%$.
+\jcr{where do I need to look for the -34.1 and -34.9 values?}
 
@@
 \begin{center}
-\caption{Implicit discretization run time results}
+\caption{Implicit flux discretization run-time results}
 \label{tab:ndiff_lra}
@@
 \begin{center}
-\caption{IQS run time results with $\Delta t = 0.004$}
+\caption{IQS run-time results with $\Delta t = 0.004$}
 \label{tab:iqs_lra}
@@
 \\
-$^*$ difference in runtime from $\Delta t = 0.004$ implicit discretization 
+$^*$ difference in run-time from $\Delta t = 0.004$ implicit discretization 
 \end{center}
@@
 \begin{center}
-\caption{IQS PC run time results with $\Delta t = 0.004$}
+\caption{\iqspc run time results with $\Delta t = 0.004$}
 \label{tab:iqspc_lra}
@@
 %----------------------------------------------------------------------------------------
-\subsubsection{LRA Time Adaptation}
-
-\fig{fig:lra_dt2} shows the power profile of the LRA with time adaptation of implicit discretization and \iqspc, and \tbl{tab:lra_dt2} compiles the results.
-These time adaptation results show the significant decrease in macro time steps required for \iqspc.
-These profiles were obtaining with only one temperature update per macro step; so based on previous results, the \iqspc performance would improve even more with more updates.
+\subsubsection{IQS Results for the LRA Benchmark with Time Adaptivity}
+
+\fig{fig:lra_dt2} shows the power profile of the LRA with time a of implicit discretization and \iqspc, and \tbl{tab:lra_dt2} compiles the results.
+These adaptivity results show the significant decrease in the number of macro time steps required for \iqspc.
+The power profiles were obtaining with only one temperature update per macro step; so based on previous results, the \iqspc performance would improve even more with more updates.
+\jcr{feels strange to suppose results behavior.
+if i were a reviewer, I'd ask for them to be included}
 
@@
 
-Transient 15 is a test case based on the TREAT core.
-The purpose of the original creation of this simulation in Rattlesnake is to test the model's fidelity with the thermal feedback of TREAT, but it is not meant to exactly match any previous experiments.
-Nevertheless, the goal of the following simulations is to test IQS and its time scale based treatment of temperature with a more complex model.
-Transient 15 involves an 11-energy group diffusion approximation and is discretized into $355,712$ hexahedral continuous finite elements totaling $4,109,523$ degrees of freedom.
-The three-second transient involves a linear ramp decrease in the absorption cross section throughout the control rod region.
-\fig{fig:Tran15} shows a visualization of the flux profile within the core, hidden is the massive amount of graphite surrounding the core.
+Transient-15 is a test case based on the TREAT core at the INL \jcr{some ref for TREAT? The batman report maybe?}. 
+This test case is not meant to model an actual TREAT experiment but was created to evaluate Rattlesnake’s physics 
+models for TREAT transient with feedback.  However, the model was complex enough and representative enough to test 
+the performance of IQS and its time scale based treatment of temperature.
+Transient-15 involves an 11-energy group diffusion approximation and the geometry is discretized into $355,712$ hexahedral continuous finite elements totaling $4,109,523$ degrees of freedom.
+The three-second transient involves a linear ramp decrease in the absorption cross section throughout the control rod region (to simulate a rod movement out of the core).
+\fig{fig:Tran15} shows a visualization of the flux profile within the core (the graphite reflector has been removed in the picture to show only the active core).
 
@@
 \centering
-\includegraphics[height=3in]{figures/Tran15_core2.png}
-\caption{Transient 15 core power profile at peak power}
+\includegraphics[height=2.5in]{figures/Tran15_core2.png}
+\caption{Transient-15 core power profile at peak power ($t=???$ \jcr{fill in the blanks})}
 \label{fig:Tran15}
@@
 
-The Transient-15 model uses a adiabatic temperature feedback mechanism, similar to the one explored by the LRA.
-\eqt{eq:temp2} describes the heat up of the fuel.
-It is very similar, except the specific heat is now dependent on temperature is described by \eqt{eq:cp}.
-The temperature evaluation is identical to the one described in LRA section, except a Newton iteration process is employed to resolve the nonlinearity from the specific heat term.
-The feedback to the cross-sections are applied using linear interpolation of tabular data provided by INL.
+The Transient-15 model uses an adiabatic temperature feedback mechanism, similar to the one used in the LRA benchmark.
+\eqt{eq:temp2} describes the heat up of the fuel, where, this time, the fuel specific heat capacity is itself also temperature-dependent.
+Hence, the temperature evaluation is identical to the one described in LRA section, except a Newton iteration process is employed to resolve the nonlinearity from the specific heat term.
+The 11-g neutron cross sections are linearly interpolated at tabulated temperatures.
 
@@
 
-\be
-c_p = -5.8219e-10T^3 - 4.3694e-7T^2 + 2.8369e-3T -1.009e-2
-\label{eq:cp}
-\ee
+%\be
+%c_p = -5.8219e-10T^3 - 4.3694e-7T^2 + 2.8369e-3T -1.009e-2
+%\label{eq:cp}
+%\ee
 
 %----------------------------------------------------------------------------------------
-\subsubsection{Transient-15 Multiphysics Time Scale Results}
-
-In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time step, implicit discretization, IQS with one and 5 temperature updates per macro step, and \iqspc with one and 5 updates.
+%\subsubsection{Transient-15 Multiphysics Time Scale Results}
+
+In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time-step, implicit flux discretization, IQS with one and 5 temperature updates per macro step, and \iqspc with one and 5 temperature updates.
 \fig{fig:Tran15_profile} shows the baseline power and temperature profile for the Transient-15 example.
-\tbl{tab:tran15} shows the error and runtime results.
+\tbl{tab:tran15} shows the error and run-time results.
 
@@
 Baseline 			& 3000 	& 3.5039e+08 & 1.901 & 371 &	---		&	---		 & ---		\\
-Implicit Dis. 		& 300 	& 3.5011e+08 & 1.90  & 371 &	---		&	7.875e-4 & 41020	\\
+Implicit Disc. 		& 300 	& 3.5011e+08 & 1.90  & 371 &	---		&	7.875e-4 & 41020	\\
 IQS 				& 300	& 3.5036e+08 & 1.90  & 371 & -11.9\%	&	8.385e-5 & 23949	\\
@@
 \vspace{-3mm}
-$^*$ difference in runtime from implicit discretization 
+$^*$ difference in run-time from implicit discretization 
 \caption{Transient-15 Error and Runtime Results}
@@
 
-The results from \tbl{tab:tran15} show similar performance of IQS with the temperature updates as the LRA.
-Again, the number of linear GMRES iterations is shown as a measure of computational expense.
-However, these iterations do not consider the temperature updates, so the iterations of the simulations with multiple updates should be taken with a grain of salt.
-IQS with 1 temperature update shows a performance that reduces the error to approximately a tenth of the implicit discretization error, and reduces the execution time by about 12\%.
+The results from \tbl{tab:tran15} show similar performance of IQS with the temperature updates as in the LRA
+test case.
+Again, the number of linear GMRES iterations is shown as a measure of computational cost.
+However, these iterations do not consider the temperature updates (auxiliary kernels solved outside of the GMRES solver for temperature updates).
+IQS with 1 temperature update shows a performance that reduces the error to approximately a tenth of the implicit flux discretization error, and reduces the execution time by about 12\%.
 This shows that IQS was able to resolve the nonlinearity between flux and temperature with significantly fewer diffusion evaluations.
-Having IQS with 5 updates significantly increased the execution time for the same time step, but the error was reduced.
-Comparing this error to a similar implicit discretization error at a smaller time step could show that the runtime was reduced.
+Having IQS with 5 updates increased the execution time for the same time step, but the error was further reduced.
+Comparing this error to a similar implicit flux discretization error at a smaller time step could show that the run-time was reduced.
 \iqspc performed not nearly as well as it did with the LRA benchmark, but still proved to be effective.
-Having 5 updates for \iqspc increased the runtime marginally, but decreased the error significantly.
+Having 5 updates for \iqspc increased the run-time marginally, but decreased the error significantly.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The goal of this paper was to investigate the convergence property of IQS schemes and to evaluate
+additional time-scales (updates) for temperature feedback in the IQS schemes.
+Specifically, this paper (i) discussed the different nonlinear iteration techniques for IQS and tests the rigor of their implementation (convergence criterion), (ii) investigates high-order temporal discretization of the shape equation, and (iii) analyzes IQS coupled with adiabatic heat-up and cross-section feedback.
+
+IQS is a nonlinear system of equations; therefore, a fixed-point iteration technique was used to evaluate shape and amplitude.
+Investigating these iterative techniques is important for understanding the behavior of IQS, as well as determining the most appropriate technique for optimal performance.
+The iteration techniques were applied to a one-dimensional prototype problem for testing.
+For fixed-point iteration, five different criteria were tested:
+\begin{enumerate}
+\item $L^\infty$ norm of the change in shape between iterations
+\item $L^2$ norm of the change in shape between iterations
+\item Difference in reactivity between iterations
+\item Difference in amplitude between iterations
+\item IQS uniqueness consistency criteria
+\end{enumerate}
+The results showed that criteria 1-4 had relatively equivalent convergence behavior.
+However, iteration with the criteria 5 could not converge without a highly-accurate analytical treatment of the precursor equation.
+This criteria proved to be the most rigorous and thorough convergence criteria and is recommended for any application of IQS.
+
+IQS effectiveness with high-order temporal discretization schemes was also investigated.
+BDF schemes up to order 4 were applied to the one-dimensional prototype problem.
+IQS showed expected error convergence up.
+However,
+a reduction in effectiveness was noted when comparing IQS (and \iqspc) with a straight implicit discretization of the flux equations: for high-order schemes, it appeared that the solving the flux equations over any IQS technique yielded
+almost the same performance.
+
+Several multi-dimensional test cases were investigated.
+IQS showed the expected error convergence on all cases.
+Step adaptation was also applied to the TWIGL benchmark, where IQS's performance is significant, reducing the number of diffusion evaluations considerably.
+Multi-dimensional test cases with temperature feedback were performed as well (the LRA benchmark and a representative case of the TREAT core). 
+These examples involve a adiabatic heat up of the fuel with Doppler broadening feedback of the cross sections.
+The evaluation of temperature was integrated in the quasi-static process by introducing an intermediate time scale for temperature and PRKE parameter evaluation.
+The results of the LRA benchmark showed that quasi-static approach to temperature greatly improved the accuracy for a given time step size.
+However, the increase in computation time due to the extra temperature evaluations was significant.
+The results of the TREAT example showed similar behavior, except the temperature updates only produced a smaller increase in accuracy.
+These examples show that performing a variant number of updates during the transient could further improve the performance of this quasi-static process.
+
+\jcr{you need an outlook. what should be next? and then you're done}
 

### 2017-09-26 doc/Papers/IQS_convergence/IQS_convergence.tex
Addressing comments and adding future work

@@
 \usepackage{xspace} 
+\usepackage{color, colortbl}
 
@@
 \newcommand{\jcr}[1]{\textcolor{red}{jcr: #1}}
+\newcommand{\zmp}[1]{\textcolor{blue}{zmp: #1}}
 
@@
 
-First, the IQS shape equation was represented by the gathering of appropriate finite ELement kernels.
+First, the IQS shape equation was represented by the gathering of appropriate finite element kernels.
 Most of the shape equation kernels are the same as the flux kernels already implemented in Rattlesnake for its standard diffusion (or transport) solution, and only two kernels had to be modified.
@@
 
-In this section, we analyze various convergence criteria proposed in the nonlinear solution technique for the IQDS method.
+In this section, we analyze various convergence criteria proposed in the nonlinear solution technique for the IQS method.
 We also investigate the use of higher-temporal discretization for the shape equations.
@@
 
-This section describes two multiphysics examples with adiabatic heat-up: the LRA benchmark from the  ANL Benchmark Problem Book \cite{ANL_BPB} 
+This section describes two transient multiphysics examples with adiabatic heat-up and Doppler feedback in a thermal reactor: the LRA benchmark from the  ANL Benchmark Problem Book \cite{ANL_BPB} 
 and a modelisation of an experiment carried out at the Transient Reactor Test Facility, TREAT \cite{mammoth, Tran15}. 
-% These examples are of increased complexity from the previous kinetics examples.
-This section also analyzes IQS's performance with these, which is vital for verification of IQS in real-world problems.
+Both examples were performed using the Rattlesnake/MOOSE framework at INL. The spatial discretization uses continuous finite elements with first-order Lagrangian basis functions.
 
@@
 
-The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor.  
-\jcr{already said before}
+The LRA benchmark is a two-dimensional, two-group neutron diffusion problem.  
 It is a super prompt-critical transient.
 The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.
-The spatial discretization uses continuous finite elements with first-order Lagrangian basis functions \jcr{move up and say only one.
-I believe all rattlesnake results use FEM order 1}.
+The spatial discretization uses continuous finite elements with first-order Lagrangian basis functions.
 The mesh consists of blocks $11\times 11$ with five uniform refinements, totaling $165,165$ elements and $124,609$ nodes.
@@
 Put all info that is common to all results in 1 location at the beginning}.
+\zmp{I put the PRKE evaluation detail in the theory section. Should I put it again here?}
 The performance of IQS with temperature feedback is assessed by comparing the solution accuracy at peak power.
@@
 The convergence plots show that updating temperature and the PRKE parameters within a macro step has a significant effect on the performance of IQS.
-With only one update, IQS was only slightly better than the implicit flux discretization which required about 150\% \jcr{2.5 more then?
-I think you mean +50\%}more time steps than IQS for the same error.
+With only one update, IQS was only slightly better than the implicit flux discretization which required about 50\% more time steps than IQS for the same error.
 While 5 temperature updates showed a much improved performance for IQS, with the implicit flux discretization required about 400\% more time steps than IQS for the same error level.
@@
 It is also important to compare the error of implicit discretization with IQS at one update and \iqspc at 4 updates.
-IQS shows an error comparable to implicit discretization at $\Delta t = 0.002$ \jcr{where's that table with that time step?
-or do you mean it's not there and you just talk about results that are not in the paper?
-which would be a strange way of conveying information}, signifying an actual increase in runtime by -34.1\%.
+IQS shows an error comparable to implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by -34.1\%.
 \iqspc shows an error less than implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by $<-34.9\%$.
-\jcr{where do I need to look for the -34.1 and -34.9 values?}
+These specific runs are highlighted in Tables \ref{tab:ndiff_lra} - \ref{tab:iqspc_lra}.
 
@@
 1	& 4.0e-3	& 1.407e-2 	& 4.11	& 7.13e4	\\
-2	& 2.0e-3	& 3.174e-3 	& 6.01	& 9.49e4 	\\
+\rowcolor{yellow} 2	& 2.0e-3	& 3.174e-3 	& 6.01	& 9.49e4 	\\
 3 	& 1.0e-3 	& 7.690e-4 	& 10.38	& 1.45e5	\\
@@
 \hline
-1	& 1		& 2.612e-3 	& 3.96 	& -3.18\%	\\
+\rowcolor{yellow} 1	& 1		& 2.612e-3 	& 3.96 	& -3.18\%	\\
 2	& 2		& 9.893e-4 	& 6.02	&  47.1\%	\\
@@
 2	& 2		& 1.349e-3 	& 3.73	& -9.00\%	\\
-3 	& 4 	& 9.161e-4 	& 3.97	& -3.04\%	\\
+\rowcolor{yellow} 3 	& 4 	& 9.161e-4 	& 3.97	& -3.04\%	\\
 4 	& 8 	& 8.052e-4 	& 5.39	&  31.7\%	\\
@@
 These adaptivity results show the significant decrease in the number of macro time steps required for \iqspc.
-The power profiles were obtaining with only one temperature update per macro step; so based on previous results, the \iqspc performance would improve even more with more updates.
-\jcr{feels strange to suppose results behavior.
-if i were a reviewer, I'd ask for them to be included}
+The power profiles were obtaining with only one temperature update per macro step.
 
@@
 
-Transient-15 is a test case based on the TREAT core at the INL \jcr{some ref for TREAT? The batman report maybe?}. 
+Transient-15 is a test case based on the TREAT core at the INL \cite{mammoth, Tran15}. 
 This test case is not meant to model an actual TREAT experiment but was created to evaluate Rattlesnake’s physics 
@@
 \includegraphics[height=2.5in]{figures/Tran15_core2.png}
-\caption{Transient-15 core power profile at peak power ($t=???$ \jcr{fill in the blanks})}
+\caption{Transient-15 core power profile at peak power ($t=1.90 s$)}
 \label{fig:Tran15}
@@
 
-\jcr{you need an outlook. what should be next? and then you're done}
+This work shows how effective IQS can be with different problem and time-discretization types.
+Possible future work would be to further optimize IQS for multi-physics problems in a more general sense.
+This proposed work would include an adaptation-like approach to the quasi-static evaluation of the non-neutronics physics.
+For instance, instead of a fixed number of temperature evaluations per macro step, adapting the number of these evaluations through the transient could improve results.
+This adaption could also be applied to other physics like thermal hydraulic and structural feedback.
 

### 2017-09-26 doc/Papers/IQS_convergence/IQS_convergence.tex
Addressing comments and adding future work

@@
 
-This work shows how effective IQS can be with different problem and time-discretization types.
+This work shows, to a degree, how effective IQS can be with different problem and time-discretization types.
 Possible future work would be to further optimize IQS for multi-physics problems in a more general sense.
@@
 For instance, instead of a fixed number of temperature evaluations per macro step, adapting the number of these evaluations through the transient could improve results.
-This adaption could also be applied to other physics like thermal hydraulic and structural feedback.
+This adaption could be applied to other physics like thermal hydraulic and structural feedback.
 

### 2018-04-04 doc/Papers/IQS_convergence/IQS_convergence.tex
Adding acknowlegments section

--- 2017-09-26 doc/Papers/IQS_convergence/IQS_convergence.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Acknowledgments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+This work was supported by the Department of Energy, Idaho National Laboratory, and the Integrated University Program Fellowship.
+We thank Mark Dehart, Yaqi Wang, the NEAMS program, and INL's MOOSE/Rattlesnake team for their support.
+This research did not receive any specific grant from funding agencies in the public, commercial, or
+not-for-profit sectors.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliographystyle{elsarticle-num}

### 2018-04-05 doc/Papers/IQS_convergence/IQS_convergence.tex
final review and sdirk33 ref

--- 2017-09-26 doc/Papers/IQS_convergence/IQS_convergence.tex
@@
 Note that the PRKE parameters are evaluated at each macro step and interpolated for the PRKE evaluation. In order to preserve 
-the error convergence rate of high order temporal discretization schemes for shape, higher order interpolation of the parameters will be required.
-Third order implicit Runge-Kutta discretization with time step adaptation was used for the PRKE evaluation, to insure a negligibly small error in the amplitude solution.
+the error convergence rate of high order temporal discretization schemes for shape, higher order interpolation of the parameters is required.
+A third-order implicit Runge-Kutta \cite{sdirk33} discretization with time-step adaptation is used for all PRKE evaluations in this paper in
+order to ensure a negligibly small error in the amplitude solution.
 
@@
 Crank-Nicolson time discretization scheme is used for the space-time equations.
-\jcr{when you recall all common numerical options, make sure you recall that the PRKE ODE are done with high-order implicit RK.
-Put all info that is common to all results in 1 location at the beginning}.
-\zmp{I put the PRKE evaluation detail in the theory section. Should I put it again here?}
 The performance of IQS with temperature feedback is assessed by comparing the solution accuracy at peak power.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Acknowledgments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+This work was supported by the Department of Energy through grant DE-AC07-05ID14517 (Idaho National Laboratory) and the Integrated University Program Fellowship (Cooperative Agreement DE-NE0000112).
+We thank Mark D.
+Dehart, Yaqi Wang, and INL's MOOSE/Rattlesnake team for their support.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliographystyle{elsarticle-num}

### 2018-04-06 doc/Papers/IQS_convergence/cover-letter-iqs.tex
Adding cover letter for submission

\documentstyle[11pt]{letter}

%%%%%% Letter Size Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%        \addtolength{\textwidth}{2.5cm}     %% For longer or shorter text width
        \addtolength{\topmargin}{-3.5cm}    %% For more or less top margin
       \addtolength{\textheight}{7cm}    %% For longer or shorter textheight
%        \addtolength{\oddsidemargin}{-1.25cm} %% For odd side margin (twoside)
                                            %% or margin (oneside)

\address{Zachary M. Prince\\ 
Department of Nuclear Engineering \\
Texas A\&M University\\
College Station, TX 77843-3133, USA\\
phone: (360) 362 6622\\
e-mail: zachmprince@tamu.edu \vspace{0.5cm}}

%%%%%% The Signature  and Date %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\signature{\vspace{-1.25cm}Zachary M. Prince, Jean C. Ragusa}

\begin{document}

\begin{letter}{Professor Sara A. Pozzi\\
    Editor,\\
    Elsevier Annals of Nuclear Energy}
\date{\today}
%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%         \vspace{3.0cm}

\opening{Dear Professor Pozzi,}
         \vspace{0.25cm}
%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Please find attached a copy of our manuscript titled ``Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method'' for submission to the {\em Annals of Nuclear Energy}.

In this paper, we revisit the improved quasi-static method (IQS) for its application to high order time discretizations and temperature feedback dynamics. We derive the IQS and IQS predictor-corrector method (IQS-PC) with the semi-analytical treatment of delayed neutron precursors and adiabatic heat-up. Temperature computation and feedback are treated on a separate time scale from shape and amplitude to increase the efficiency of nonlinear quasi-static process.

We show numerical results from four different test cases where we compare error from IQS computation and typical implicit discretization of the flux equation. A one-dimensional test was constructed to analyze error and nonlinear iteration convergence for IQS with up to fourth-order backward-difference-formulae discretization. IQS was applied to the TWIGL benchmark to show its performance with step doubling time adaptation. Analysis of the quasi-static treatment of temperature feedback are shown with results from the LRA benchmark and a test case from a model of the Transient Reactor Testing Facility.

IQS and IQS-PC were implemented into the MOOSE-Rattlesnake framework at Idaho National Laboratory in order to compute the results from the multi-dimensional examples.

Thank you for considering this manuscript for publication in the {\it Annals of Nuclear Energy}.

\vspace{0.25cm}

%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%% The Closing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\closing{Best regards, }

\end{letter}
\end{document}

### 2018-04-06 doc/Papers/IQS_convergence/IQS_convergence.tex

@@
 
-This work shows how effective IQS can be with different problem and time-discretization types.
-Possible future work would be to further optimize IQS for multi-physics problems in a more general sense.
-This proposed work would include an adaptation-like approach to the quasi-static evaluation of the non-neutronics physics.
-For instance, instead of a fixed number of temperature evaluations per macro step, adapting the number of these evaluations through the transient could improve results.
-This adaption could also be applied to other physics like thermal hydraulic and structural feedback.
+This work investigated the effectiveness of for several time discretizations.
+Extension to this work could include a further optimization of the IQS algorithm for multi-physics problems by extending, for instance,  the use of adaptivity in the evaluation of the other physic components.
+For example, instead of a fixed number of temperature evaluations per macro time-step, adapting the number of these evaluations during the transient could further reduce simulation times.
+This adaption could be applied to coupled physics such as thermal-hydraulic and structural feedback.
 

### 2018-04-06 doc/Papers/IQS_convergence/cover-letter-iqs.tex
cover

@@
 
-In this paper, we revisit the improved quasi-static method (IQS) for its application to high order time discretizations and temperature feedback dynamics.
+In this paper, we revisit the improved quasi-static method (IQS) for its application to high-order time discretizations and temperature feedback dynamics.
 We derive the IQS and IQS predictor-corrector method (IQS-PC) with the semi-analytical treatment of delayed neutron precursors and adiabatic heat-up.
-Temperature computation and feedback are treated on a separate time scale from shape and amplitude to increase the efficiency of nonlinear quasi-static process.
+Temperature computation and feedback are treated on a separate time scale from the neutronic shape and amplitude functions to increase the efficiency of nonlinear quasi-static process.
 
-We show numerical results from four different test cases where we compare error from IQS computation and typical implicit discretization of the flux equation.
+We show numerical results from four different test cases where we compare error norms from IQS computation and standard implicit discretization of the flux equations.
 A one-dimensional test was constructed to analyze error and nonlinear iteration convergence for IQS with up to fourth-order backward-difference-formulae discretization.
 IQS was applied to the TWIGL benchmark to show its performance with step doubling time adaptation.
-Analysis of the quasi-static treatment of temperature feedback are shown with results from the LRA benchmark and a test case from a model of the Transient Reactor Testing Facility.
+Analysis of the quasi-static treatment of temperature feedback are shown with results from the LRA benchmark and a test case from a model of the Transient Reactor Testing Facility (TREAT).
 
-IQS and IQS-PC were implemented into the MOOSE-Rattlesnake framework at Idaho National Laboratory in order to compute the results from the multi-dimensional examples.
+IQS and IQS-PC were implemented into the MOOSE-Rattlesnake framework of Idaho National Laboratory.
 

### 2018-05-29 doc/Papers/IQS_convergence/highlights-iqs.tex
Adding highlights for submission

\documentclass{elsarticle}

\begin{document}

\section*{Highlights}

\begin{itemize}
\item Improved Quasi-Static Method (IQS) shows proper error convergence through fourth-order discretization, but performance diminishes compared to standard discretization.
\item IQS uniqueness criteria can only be preserved through a highly-accurate analytical treatment of delayed neutron precursors.
\item The effectiveness of IQS is maximized with time adaptation of shape updates.
\item Quasi-static treatment of other kinetic variables, such as temperature, significantly improves IQS performance for multi-physics problems.
\end{itemize}

\end{document}

### 2018-08-28 doc/Papers/IQS_convergence/reviewer_notes.tex
Itemizing and commenting on reviewer's notes

\documentclass{elsarticle}

\usepackage[letterpaper, margin=1in]{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{graphicx,amsmath, amssymb, amsthm, amsfonts,subcaption}
\usepackage[usenames, dvipsnames]{xcolor}

\newcommand{\fix}{\textsquare}
\newcommand{\working}{$\boxdot$}
\newcommand{\nofix}{?}
\newcommand{\done}{\checkmark}

\newcommand{\easy}[1]{\textcolor{ForestGreen}{#1}}
\newcommand{\medm}[1]{\textcolor{BurntOrange}{#1}}
\newcommand{\hard}[1]{\textcolor{Red}{#1}}

\begin{document}

\section*{Major Issues}

\begin{itemize}

\item[\fix] \medm{ Justify page 8: "we expect temperature to be more rapidly varying than the shape but less so than the amplitude" }
\begin{itemize}
\item Should we produce some kind of mathematical observation, or can we allude to the dynamical time constant data that we can include for the LRA and TREAT?
\end{itemize}

\item[\fix] \easy{ Justify page 5: "Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself, this allowing for larger time step sizes in updating the shape". }
\begin{itemize}
\item Reference first IQS paper
\end{itemize}

\item[\fix] \easy{  Results are often just stated and not discussed, e.g. page 19 "It also shows that traditional IQS performed better with large etol, while IQS-PC was better with smaller etol". The statement simply bloats the manuscripts; discuss the reasons and implications or remove it. Another example is page 15, "However, switching the precursors solve to be performed semi-analytically…".  In this case, the authors should take the time to discuss the implications of this finding. }
\begin{itemize}
\item I will remove the first referenced quote.
\item I will add that if one expecting a certain tolerance to be reached in the iterative process, the semi-analytical integration of the precursors is necessary. Since, in the majority of cases, this evaluation is negligible compared to the evaluation of the shape, it is fortuitous to apply this technique in all cases.
\end{itemize}

\item[\fix] \medm{ For the comparison of LRA and TREAT results, the efficiency [how accurate can you be given a fixed execution time or given an accuracy how much execution time do you need] is a very important quantity; the authors discuss it for LRA and TREAT. However, the authors first present results for the errors and then runtime results. Then they try to synthesize efficiency by comparing different cases that have different runtimes and different errors, e.g. page 23 right before Table 5 "IQS-PC shows and error less than implicit discretization at delta t = 0.002, signifying …". I suggest the authors create error vs. execution time plots for different cases and based on these curves discuss efficiency. The manner efficiency is discussed is very confusing and not very useful for the reader. }
\begin{itemize}
\item Is just a plot of error vs. runtime sufficient?
\end{itemize}

\item[\fix] \easy{ In the results section, Error is sometimes used without introducing it, e.g. page 25, "Table 9 shows the error ….". Further "Error" is often used as y-label in plots., e.g. in the discussion of Figure 10 in which "convergence results" are presented; the sentence is not clear in this context because it was not stated if this is iterative convergence or convergence to the right solution as the time step is reduced. In some instances error is defined but it is tedious for the reader to find the corresponding passage in the text, figures, or tables. }
\begin{itemize}
\item Figure 9: Error is value of K error after 20 iterations
\item Figure 10,12,15,16,19 + Table 5,6,7,8: Error is relative difference in the peak power from highly refined solution
\end{itemize}

\item[\fix] \hard{ LRA and TREAT only compare errors computed with the peak powers. Peak powers are an integral quantity and may favor IQS over spatial kinetics. The authors should also compare power distributions or similar distributed quantities. }
\begin{itemize}
\item This would require a rerun of all simulations. There is a postprocessor in MOOSE that could do it, but the simulations would be ridiculously tedious.
\end{itemize}

\item[\fix] \medm{  Whenever errors are computed, baseline calculations are performed. How do the authors know that these baseline calculations are accurate enough to measure the error w.r.t. them? }
\begin{itemize}
\item Can we just say that a proper error convergence means that the baseline is good enough?
\end{itemize}

\end{itemize}

\section*{Minor Issues}

\begin{itemize}

\item[\fix] \easy{ Page 1: "… the amplitude, and a space- and time-dependent component, the shape [1,2,3,4,5]". What about energy? }
\begin{itemize}
\item space-, time-, energy-dependent component...
\end{itemize}

\item[\fix] \easy{ Is $\phi^g$ the scalar flux or scalar flux moments in Eq 1a? }
\begin{itemize}
\item Multigroup scalar flux
\end{itemize}

\item[\nofix] \easy{ All quantities in Eq. 1a and 1b need to be defined. }
\begin{itemize}
\item I thought we did in the paragraph below it?
\end{itemize}

\item[\nofix] \easy{ Eq. 3a: $S_d^g$ is not defined. }
\begin{itemize}
\item It is in the paragraph below Eq. (1)
\end{itemize}

\item[\fix] \easy{ For the uninitiated reader the term phase-space domain may not be familiar. The authors could introduce the term phase space. IQS may be interesting for people in other fields. }
\begin{itemize}
\item I guess we can include a definition in the first paragraph
\item Neutron flux lives in a 7 dimensional phase-space, dependent on time, space, energy, and direction. The neutron diffusion equation reduces this phase space by eliminating direction. IQS factorizes neutron flux into a time-only-dependent component, the amplitude, and a full phase-space dependent component, the shape...  
\end{itemize}

\item[\fix] \easy{ Algorithm in section 2.1 on page 5. Step 2: "linearly interpolate". It is later stated that interpolation of higher order is done for high order schemes. }
\begin{itemize}
\item Just remove linearly and state later that linear interpolation was used unless otherwise stated.
\end{itemize}

\item[\nofix] \easy{ For the Linf nor of the shape: why is the maximum difference over the maximum value used and not the maximum relative difference ($1 - \phi^{k+1}/\phi^k$). }
\begin{itemize}
\item I could include the reference for this error estimation, since I basically copied and there is no other reason why not to do a relative difference.
\end{itemize}

\item[\fix] \medm{ Eq (11) is restated in Eq (24). Why not just make Eq. 11 more general. Also, in Eq. (24) cp(T(t)) and originally the heat equation reads d(rho * cp(T) * T). Hence, the cp(T) cannot be pulled out so Eq. (24) is incorrect. }
\begin{itemize}
\item Eq. (11) is the linear case of Eq. (24) and makes the derivation of the semi-analytical integration much simpler.
\item Correcting Eq. (24) seems unnecessary, but it's easy so I'll fix it.
\end{itemize}

\item[\fix] \easy{ Figure 2. uses the delta t / 3 step that was claimed to be only for illustration purposes in Fig. 1. }
\begin{itemize}
\item I'll change the 3 to $N_{T}$ in Figure 2, which is the number of temperature solves within a macro step
\end{itemize}

\item[\fix] \medm{ Statement page 9: "Furthermore, a very accurate representation of p(t) over the macro step is available from the PRKE solve…". I think it should be precise or else one would have to show that the equations are solved accurately. Second, what does "very accurate" mean, "very accurate" is comparison to what? }
\begin{itemize}
\item We should say that it much more accurate to use the full point distribution than assuming an interpolation of the value between macro steps.
\end{itemize}

\item[\fix] \medm{ Implementation uses code specific jargon: kernels, auxkernels, Transient executioner, user-object. These need to be explained or cast in more general terms. }
\begin{itemize}
\item Two options I think: (1) Just reference a MOOSE paper that explains each term. (2) Itemize the definitions of all the terms used
\end{itemize}

\item[\fix] \easy{ Page 13 top of the page "looping over cells to evaluate", what cells, mesh cells? }
\begin{itemize}
\item Spatial mesh cells
\end{itemize}

\item[\fix] \easy{ LRA benchmark: the number of elements makes no sense 11x11 = 121 and when using five uniform refinement steps one would end up with 123,904 elements. }
\begin{itemize}
\item  Not sure how we got these numbers, but he's right
\end{itemize}

\item[\fix] \easy{ LRA benchmark description does not state what equations are solved, if that is provided in the ANL benchmark book it should be made clear. Equations could also be stated directly. }
\begin{itemize}
\item  Seems tedious to provide them, so will reference the book
\end{itemize}

\item[\fix] \easy{ In plots where delta t is plotted on the x-axis, it should be made clear that it is the macroscopic time step. }
\begin{itemize}
\item I would hate to alter all the plots, so I will just state in the intro of the results that $\Delta t$ is the macro step
\end{itemize}

\item[\fix] \easy{ Conclusions, page 27: incomplete sentence "IQS showed expected error convergence up". }
\begin{itemize}
\item ...through fourth-order discretization 
\end{itemize}

\item[\fix] \easy{ Timing estimates for LRA: the author should make clear how they execute the LRA problem, i.e. number of processors and processor model. Given the long runtimes, it appears that a single processor is used. Is this to avoid problems with asynchronous behavior in parallel? }
\begin{itemize}
\item I forgot how many were used. But my reasoning would be that longer runtimes give a more accurate representations of the actual runtime and that the parallel properties of MOOSE are somewhat of a black box. 
\item I'm not sure how relevant this is though. Can't we just say they are all on the same computer?
\end{itemize}

\end{itemize}

\end{document}

### 2018-08-29 doc/Papers/IQS_convergence/IQS_convergence.tex
Updating keywords
Adressing some of the straight-forward comments from the reviewer

@@
 \begin{keyword}
-Reactor dynamics \sep reactor kinetics \sep Improved quasi-static method \sep temporal convergence 
+Nuclear reactor dynamics \sep Quasi-static method \sep Temperature feedback \sep Time adaptivity 
 \end{keyword}
@@
 The improved quasi-static method (IQS) is a numerical technique devised for nuclear reactor transient analysis.
-It involves factorizing the neutron flux solution into a time-only-dependent component, the amplitude, and a space- and time-dependent component, the shape \cite{Ott_1966,Devooght_1984,Monier_diss,Sissaoui_1995,Dulla2008}.
+In neutron transport, the neutron flux solution lives in a seven dimensional phase-space, dependent on time, space, energy, and direction.
+The neutron diffusion equation reduces this phase-space by elimnating direction.
+IQS involves factorizing the neutron flux solution into a time-only-dependent component, the amplitude, and a full phase-space component, the shape \cite{Ott_1966,Devooght_1984,Monier_diss,Sissaoui_1995,Dulla2008}.
 The amplitude solution satisfies the point reactor kinetic equations (PRKE) where the shape solution has been used to generate the PRKE coefficients (reactivity, effective fraction of delayed neutrons, mean generation time).
@@
 \end{subequations}
-%
-where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the leakage and interaction operator, 
-$S_{d}^g$ is the delayed neutron source, and $P_{d,i}^g$ is the delayed-neutron fission operator. We assumed $G$ neutron groups and $I$ precursors groups.
-The operator $L^g$ is block diagonal in energy groups: for neutron transport, $L^g = \vec \Omega \cdot \grad + \Sigma_t^g$ whereas for neutron diffusion, 
-$L^g = -\div D^g \grad + \Sigma_r^g$, with  $\Sigma_t^g$ and $\Sigma_r^g$ the total and removal macroscopic cross sections. In the following, we specialize the 
+
+\begin{table}[htbp!]
+\centering
+\caption{Operator definition of multi-group neutron conservation equations}
+\label{tab:op_def}
+\begin{tabular}{|c|c|c|c|}
+\hline
+Operator & Description & Transport Def. & Diffusion Def. \\
+\hline
+$H^{g'\to g}$ & Scattering & $\frac{1}{4\pi}\int_{4\pi}\Sigma_s^{g'\rightarrow g}d\Omega '$ & $\Sigma_s^{g'\rightarrow g}$ \\
+\hline
+$P_p^{g'\to g}$ & Prompt fission & $\frac{1}{4\pi}\chi_p^{g'\rightarrow g}\frac{\nu_p\Sigma_f^{g'}}{k_{eff}}\int_{4\pi}d\Omega '$ & $\chi_p^{g'\rightarrow g}\frac{\nu_p\Sigma_f^{g'}}{k_{eff}}$ \\
+\hline
+$L^g$ & Streaming/Removal & $\vec \Omega \cdot \grad + \Sigma_t^g$ & $-\div D^g \grad + \Sigma_a^g$ \\
+\hline
+$S_d^g$ & Delayed neutron source & $\frac{1}{4\pi}\chi_d^g\sum_{i=1}^I\lambda_iC_i$ & $\chi_d^g\sum_{i=1}^I\lambda_iC_i$ \\
+\hline
+$P_{d,i}^g$ & Precursor production & $\beta_i\frac{\nu\Sigma_f^{g}}{k_{eff}}\int_{4\pi}d\Omega$ & $\beta_i\frac{\nu\Sigma_f^{g}}{k_{eff}}$ \\
+\hline
+\end{tabular}
+\end{table}
+
+where \tbl{tab:op_def} shows the definition of each operator for both the transport and diffusion case.
+Note that $\phi^g$ for neutron transport is the angular neutron flux ($\phi^g(\vec r, \vec\Omega, t)$), while for neutron diffusion it is scalar flux ($\phi^g(\vec r, t)$).
+%where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the leakage and interaction operator, 
+%$S_{d}^g$ is the delayed neutron source, and $P_{d,i}^g$ is the delayed-neutron fission operator. We assumed $G$ neutron groups and $I$ precursors groups.
+%The operator $L^g$ is block diagonal in energy groups: for neutron transport, $L^g = \vec \Omega \cdot \grad + \Sigma_t^g$ whereas for neutron diffusion, 
+%$L^g = -\div D^g \grad + \Sigma_r^g$, with  $\Sigma_t^g$ and $\Sigma_r^g$ the total and removal macroscopic cross sections. 
+In the following, we specialize the 
 IQS derivation for the diffusion approximation without loss of generality. Obtaining the IQS formulation for neutron transport is similarly straightforward.
@@
 the transport equations in lieu of the diffusion equations.  Using IQS, one expects the time dependence of the shape to be weaker than that 
-of the flux itself, thus allowing for larger time step sizes in updating the shape. The PRKE equations form a small ODE system and can be 
+of the flux itself, thus allowing for larger time step sizes in updating the shape \cite{Ott_1966}. The PRKE equations form a small ODE system and can be 
 solved using a much smaller time step size. In transients where the shape varies much less than the flux, IQS can be very computationally effective. 
@@
 \item[\textit{Step 1:}] Compute the PRKE parameters at the end of the macro step using the last computed shape
-\item[\textit{Step 2:}] Linearly interpolate the computed PRKE parameters over the macro step
+\item[\textit{Step 2:}] Interpolate the computed PRKE parameters over the macro step
 \item[\textit{Step 3:}] Solve the PRKE on micro steps over the entire macro step
@@
 solution has converged from shape iterations.
+It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order Hermite interpolation is necessary.
 
@@
 
-Here, we consider multiphysics reactor physics simulations.  Fuel temperature feedback mechanisms are included using an adiabatic heat conservation model given in \eqt{eq:temp}.  
+Here, we consider multiphysics reactor physics simulations.
+Fuel temperature feedback mechanisms are included using an adiabatic heat conservation model.
+For the purposes of this derivation, the adiabatic model, given in \eqt{eq:temp}, has linear heat capacity.
+However, a nonlinear heat capacity can easily be implemented with a iterative procedure.
 \be
@@
 between the amplitude's fine time scale and the shape's coarse time scale. 
+This idea stems from the fact that in a adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
+Because of the heat capacity of the fuel, the temperature amplitude is more slowly varying in time than the heat inflow from the flux amplitude, while still considerably faster than the time variance of the flux shape.
+This logic is empirically defended in the results of Section 5.
 Hence, a possible solution process for a problem with temperature feedback will have three time scales, as
@@
 time scale until convergence on each temperature step.  Then another iterative process needs to occur in the shape time scale on 
-all three variables. \fig{fig:proc} shows the programming diagram implement to execute this process. The time increment of 
-$\Delta t/3$ for the temperature solve is arbitrary and is meant only to match \fig{fig:time} where three temperature updates 
-have been used as illustration.
+all three variables. \fig{fig:proc} shows the programming diagram implement to execute this process. 
+Where $N_T$ is the number of temperature updates per macro step.
+%The time increment of $\Delta t/3$ for the temperature solve is arbitrary and is meant only to match \fig{fig:time} where three temperature updates have been used as illustration.
 
@@
 \draw[->,ultra thick](p1.south) -- node[right] {$t=t_{n+1}$}(p2.north);
-\draw[->,ultra thick](p2.south) -- node[right] {$t=t_{n}+\Delta t/3$}(p3.north);
+\draw[->,ultra thick](p2.south) -- node[right] {$t=t_{n}+\Delta t/N_{T}$}(p3.north);
 \draw[->,ultra thick](p3.south) -- (p4.north);
@@
 \draw[->,ultra thick](check1.east) -| node[right] {\large{yes}} (check2.south);
-\draw[->,ultra thick](check2.west) -- node[above,sloped] {\large{no}\small \qq $t=t+\Delta t/3$} (4,-5) -- (p3.east);
+\draw[->,ultra thick](check2.west) -- node[above,sloped] {\large{no}\small \qq $t=t+\Delta t/N_{T}$} (4,-5) -- (p3.east);
 \draw[->,ultra thick](check2.north) --  (8,-10) -- node[right] {\large{yes}} (p7.south);
@@
 \ee
-Furthermore, a very accurate representation of $p(t)$ over the macro step is available from the PRKE solve (micro time scale). 
-Using this fine scale information, the analytical solve for the precursor values yields
+Furthermore, a highly refined representation of $p(t)$ over the macro step is available from the PRKE solve (micro time scale). 
+Using this fine scale information yields a more precise integration of amplitude than assuming an interpolation of the amplitude between macro step points.
+Applying this integration procedure, the semi-analytical solve for the precursor values yields
 \be
@@
 
+A brief explanation of the MOOSE-specific objects for the IQS implementation can be seen below:
+\begin{itemize}
+\item Kernel - Evaluation of the weak-form residual for a particular piece of physics.
+\item Executioner - Establishes the type of simulation. The Transient type moves the simulation in time conducting spatial evaluations at each step.
+\item Auxiliary variable and kernel - "Optional" variables that live on the same mesh as the solution and are computed using auxiliary kernels. 
+\item Postprocessor - Computes scalar values over the entire spatial mesh, usually in an integral sense.
+\item User-object - A generic type of postprocessor that allows connectivity of relevant quantities between different MOOSE objects.
+\end{itemize}
+A more detailed and comprehensive description of MOOSE objects can be found in \cite{moose_training}.
+
 First, the IQS shape equation was represented by the gathering of appropriate finite element kernels.
@@
 This executioner has breakpoints that allows specified calls to the PRKE, PRKE parameter evaluation, and other multiphysics.
-The PRKE parameters are written as user-objects, looping over cells to evaluate them.
+The PRKE parameters are written as user-objects, looping over spatial mesh cells to evaluate them.
 Similarly, the semi-analytic treatment of fuel temperature and precursors are auxiliary kernels that do a node-by-node evaluation of the variable upon requests of the transient executioner.
@@
 The second test is from the ANL Benchmark Problem Book (BPB) \cite{ANL_BPB}.
+Note that any instance of $\Delta t$ in the following figures and commentary signifies the size of the macro time step.
 
@@
 \end{subfigure}
-\caption{Final iteration error for $K$ factor convergence criterion}
+\caption{Final iteration error $\left(K_{n+1}^{(k+1)}/K_0 - 1\right)$ for $K$ factor convergence criterion}
 \end{figure}
@@
 %\end{subfigure}
-\caption{Error convergence plots of implicit discretization, IQS, and \iqspc with various time discretization schemes}
+\caption{Error convergence plots of implicit discretization, IQS, and \iqspc with various time discretization schemes.
+Errors correspond to the relative difference in the peak power from the baseline solution.}
 \label{fig:1D_conv}
@@
 \includegraphics[height=3in]{figures/TWIGL_convergence.jpg}
-\caption{Error convergence comparison of TWIGL Benchmark}
+\caption{Error convergence comparison of TWIGL Benchmark. Errors correspond to the relative difference of the power at $t=0.2 s$ from the baseline solution.}
 \label{fig:TWIGL_conv}
@@
 \hline
-Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
+Test & $e_{tol}$ & Error$^*$ & Steps & Solves & Error$^*$ & Steps & Solves & Error$^*$ & Steps & Solves \\
 \hline
@@
 \hline
-
-\end{tabular}}
+\end{tabular}}\\
+\footnotesize{$^*$ relative difference of the power at $t=0.2 s$ from the baseline solution}
 \end{center}
@@
 Both examples were performed using the Rattlesnake/MOOSE framework at INL. The spatial discretization uses continuous finite elements with first-order Lagrangian basis functions.
+Note that any instance of $\Delta t$ in the following figures and commentary signifies the size of the macro time step.
 
@@
 It is a super prompt-critical transient.
+The equations being solved are explicitly defined in problem 14-A1 of the ANL Benchmark Problem Book \cite{ANL_BPB}.
 The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.
@@
 \hline
-Run  &  $\Delta t$ & Error & Runtime (hr) & Linear Iter.\\
+Run  &  $\Delta t$ & Error$^*$ & Runtime (hr) & Linear Iter.\\
 \hline
@@
 \hline
-\end{tabular}
+\end{tabular} \\
+\footnotesize{$^*$ relative difference of the peak power from the baseline solution}
 \end{center}
@@
 	&  Temperature 	&  		& Runtime 	& \% Increase	\\
-Run	&  Updates 	& Error & (hr)		& in Runtime$^*$\\
+Run	&  Updates 	& Error$^*$ & (hr)		& in Runtime$^{\dagger}$\\
 \hline
@@
 \\
-$^*$ difference in run-time from $\Delta t = 0.004$ implicit discretization 
+\footnotesize{$^*$ relative difference of the peak power from the baseline solution}\\
+\footnotesize{$^{\dagger}$ difference in run-time from $\Delta t = 0.004$ implicit discretization} 
 \end{center}
@@
 \be
-\rho c_p(T) \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
+\frac{\partial }{\partial t}\left[\rho c_p(T)T(\vec{r},t)\right] = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
 \label{eq:temp2}
@@
 BDF schemes up to order 4 were applied to the one-dimensional prototype problem.
-IQS showed expected error convergence up.
+IQS showed expected error convergence up through fourth-order time discretization of the shape.
 However,

### 2018-08-29 doc/Papers/IQS_convergence/reviewer_notes.tex
Adressing some of the straight-forward comments from the reviewer

@@
 
-\item[\fix] \medm{ Justify page 8: "we expect temperature to be more rapidly varying than the shape but less so than the amplitude" }
+\item[\done] \medm{ Justify page 8: "we expect temperature to be more rapidly varying than the shape but less so than the amplitude" }
 \begin{itemize}
-\item Should we produce some kind of mathematical observation, or can we allude to the dynamical time constant data that we can include for the LRA and TREAT?
+\item This idea stems from the fact that in a adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
+Because of the heat capacity of the fuel, the temperature amplitude is more slowly varying in time than the heat inflow from the flux amplitude, while still considerably faster than the time variance of the flux shape.
+This logic is empirically defended in the results of Section 5.
 \end{itemize}
 
-\item[\fix] \easy{ Justify page 5: "Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself, this allowing for larger time step sizes in updating the shape".
+\item[\done] \easy{ Justify page 5: "Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself, this allowing for larger time step sizes in updating the shape".
 }
@@
 
-\item[\fix] \easy{  Results are often just stated and not discussed, e.g.
+\item[\fix] \medm{  Results are often just stated and not discussed, e.g.
 page 19 "It also shows that traditional IQS performed better with large etol, while IQS-PC was better with smaller etol".
@@
 
-\item[\fix] \easy{ In the results section, Error is sometimes used without introducing it, e.g.
+\item[\done] \easy{ In the results section, Error is sometimes used without introducing it, e.g.
 page 25, "Table 9 shows the error ….".
@@
 \begin{itemize}
-\item Figure 9: Error is value of K error after 20 iterations
-\item Figure 10,12,15,16,19 + Table 5,6,7,8: Error is relative difference in the peak power from highly refined solution
+\item Included an explicit definition of error in each of the plots and tables
 \end{itemize}
@@
 
-\item[\fix] \easy{ Page 1: "… the amplitude, and a space- and time-dependent component, the shape [1,2,3,4,5]". What about energy? }
+\item[\done] \easy{ Page 1: "… the amplitude, and a space- and time-dependent component, the shape [1,2,3,4,5]". What about energy? }
 \begin{itemize}
-\item space-, time-, energy-dependent component...
+\item IQS involves factorizing the neutron flux solution into a time-only-dependent component, the amplitude, and a full phase-space component, the shape
 \end{itemize}
 
-\item[\fix] \easy{ Is $\phi^g$ the scalar flux or scalar flux moments in Eq 1a? }
+\item[\done] \easy{ Is $\phi^g$ the scalar flux or scalar flux moments in Eq 1a? }
 \begin{itemize}
-\item Multigroup scalar flux
+\item Note that $\phi^g$ for neutron transport is the angular neutron flux ($\phi^g(\vec r, \vec\Omega, t)$), while for neutron diffusion it is scalar flux ($\phi^g(\vec r, t)$).
 \end{itemize}
 
-\item[\nofix] \easy{ All quantities in Eq. 1a and 1b need to be defined. }
+\item[\working] \easy{ All quantities in Eq. 1a and 1b need to be defined. }
 \begin{itemize}
-\item I thought we did in the paragraph below it?
+\item Created a table of the operator definitions (Table 1)
+\item Do I need to include a definition of all the properties?
 \end{itemize}
 
-\item[\nofix] \easy{ Eq. 3a: $S_d^g$ is not defined. }
+\item[\done] \easy{ Eq. 3a: $S_d^g$ is not defined. }
 \begin{itemize}
-\item It is in the paragraph below Eq. (1)
+\item See Table 1
 \end{itemize}
 
-\item[\fix] \easy{ For the uninitiated reader the term phase-space domain may not be familiar. The authors could introduce the term phase space. IQS may be interesting for people in other fields. }
+\item[\done] \easy{ For the uninitiated reader the term phase-space domain may not be familiar. The authors could introduce the term phase space. IQS may be interesting for people in other fields. }
 \begin{itemize}
-\item I guess we can include a definition in the first paragraph
-\item Neutron flux lives in a 7 dimensional phase-space, dependent on time, space, energy, and direction.
-The neutron diffusion equation reduces this phase space by eliminating direction.
-IQS factorizes neutron flux into a time-only-dependent component, the amplitude, and a full phase-space dependent component, the shape...
+\item In neutron transport, the neutron flux solution lives in a seven dimensional phase-space, dependent on time, space, energy, and direction.
+The neutron diffusion equation reduces this phase-space by elimnating direction.
 \end{itemize}
 
-\item[\fix] \easy{ Algorithm in section 2.1 on page 5. Step 2: "linearly interpolate". It is later stated that interpolation of higher order is done for high order schemes. }
+\item[\done] \easy{ Algorithm in section 2.1 on page 5. Step 2: "linearly interpolate". It is later stated that interpolation of higher order is done for high order schemes. }
 \begin{itemize}
-\item Just remove linearly and state later that linear interpolation was used unless otherwise stated.
+\item Removed linearly from step 2
+\item It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order Hermite interpolation is necessary.
 \end{itemize}
@@
 \begin{itemize}
-\item I could include the reference for this error estimation, since I basically copied and there is no other reason why not to do a relative difference.
+\item I included the reference for this error estimation.
 \end{itemize}
 
-\item[\fix] \medm{ Eq (11) is restated in Eq (24).
+\item[\done] \medm{ Eq (11) is restated in Eq (24).
 Why not just make Eq.
@@
 \begin{itemize}
-\item Eq. (11) is the linear case of Eq. (24) and makes the derivation of the semi-analytical integration much simpler.
-\item Correcting Eq. (24) seems unnecessary, but it's easy so I'll fix it.
+\item For the purposes of this derivation, the adiabatic model, given in Eq. (11), has linear heat capacity. However, a nonlinear heat capacity can easily be implemented with a iterative procedure.
+\item Put $\rho C_p(T)$ inside the derivative 
 \end{itemize}
 
-\item[\fix] \easy{ Figure 2. uses the delta t / 3 step that was claimed to be only for illustration purposes in Fig. 1. }
+\item[\done] \easy{ Figure 2. uses the delta t / 3 step that was claimed to be only for illustration purposes in Fig. 1. }
 \begin{itemize}
@@
 
-\item[\fix] \medm{ Statement page 9: "Furthermore, a very accurate representation of p(t) over the macro step is available from the PRKE solve…".
+\item[\done] \medm{ Statement page 9: "Furthermore, a very accurate representation of p(t) over the macro step is available from the PRKE solve…".
 I think it should be precise or else one would have to show that the equations are solved accurately.
@@
 \begin{itemize}
-\item We should say that it much more accurate to use the full point distribution than assuming an interpolation of the value between macro steps.
+\item Furthermore, a highly refined representation of $p(t)$ over the macro step is available from the PRKE solve (micro time scale).
+Using this fine scale information yields a more precise integration of amplitude than assuming an interpolation of the amplitude between macro step points.
 \end{itemize}
 
-\item[\fix] \medm{ Implementation uses code specific jargon: kernels, auxkernels, Transient executioner, user-object. These need to be explained or cast in more general terms. }
+\item[\done] \medm{ Implementation uses code specific jargon: kernels, auxkernels, Transient executioner, user-object. These need to be explained or cast in more general terms. }
 \begin{itemize}
-\item Two options I think: (1) Just reference a MOOSE paper that explains each term. (2) Itemize the definitions of all the terms used
+\item Kernel - Evaluation of the weak-form residual for a particular piece of physics.
+\item Executioner - Establishes the type of simulation. The Transient type moves the simulation in time conducting spatial evaluations at each step.
+\item Auxiliary variable and kernel - "Optional" variables that live on the same mesh as the solution and are computed using auxiliary kernels. 
+\item Postprocessor - Computes scalar values over the entire spatial mesh, usually in an integral sense.
+\item User-object - A generic type of postprocessor that allows connectivity of relevant quantities between different MOOSE objects.
+\item A more detailed and comprehensive description of MOOSE objects can be found in [18]
 \end{itemize}
 
-\item[\fix] \easy{ Page 13 top of the page "looping over cells to evaluate", what cells, mesh cells? }
+\item[\done] \easy{ Page 13 top of the page "looping over cells to evaluate", what cells, mesh cells? }
 \begin{itemize}
@@
 
-\item[\fix] \easy{ LRA benchmark description does not state what equations are solved, if that is provided in the ANL benchmark book it should be made clear. Equations could also be stated directly. }
+\item[\done] \easy{ LRA benchmark description does not state what equations are solved, if that is provided in the ANL benchmark book it should be made clear.
+Equations could also be stated directly.
+}
 \begin{itemize}
@@
 
-\item[\fix] \easy{ In plots where delta t is plotted on the x-axis, it should be made clear that it is the macroscopic time step. }
+\item[\done] \easy{ In plots where delta t is plotted on the x-axis, it should be made clear that it is the macroscopic time step. }
 \begin{itemize}
-\item I would hate to alter all the plots, so I will just state in the intro of the results that $\Delta t$ is the macro step
+\item Put a note at the end of the intro of each results section
 \end{itemize}
 
-\item[\fix] \easy{ Conclusions, page 27: incomplete sentence "IQS showed expected error convergence up". }
+\item[\done] \easy{ Conclusions, page 27: incomplete sentence "IQS showed expected error convergence up". }
 \begin{itemize}
-\item ...through fourth-order discretization 
+\item ...through fourth-order discretization of the shape
 \end{itemize}
 
-\item[\fix] \easy{ Timing estimates for LRA: the author should make clear how they execute the LRA problem, i.e.
+\item[\nofix] \easy{ Timing estimates for LRA: the author should make clear how they execute the LRA problem, i.e.
 number of processors and processor model.
@@
 \begin{itemize}
-\item I forgot how many were used.
-But my reasoning would be that longer runtimes give a more accurate representations of the actual runtime and that the parallel properties of MOOSE are somewhat of a black box.
-\item I'm not sure how relevant this is though. Can't we just say they are all on the same computer?
+\item Already stated: These run-times are based on total alive time of the execution where the diffusion evaluation is distributed over 24 processors.
 \end{itemize}

### 2018-08-29 doc/Papers/IQS_convergence/IQS_convergence.tex
Including runtime vs error for LRA stuff

@@
 These specific runs are highlighted in Tables \ref{tab:ndiff_lra} - \ref{tab:iqspc_lra}.
+\fig{fig:lra_rt_vs_err} visualizes the dependence of runtime on error. This figure shows that IQS and IQS P-C generally has a lower runtime for certain error than implicit discretization.
 
@@
 \end{table}
+
+\begin{figure}[htbp!]
+\centering
+\includegraphics[height=3in]{figures/lra_rt_vs_err.png}
+\caption{Runtime vs. error for various number of temperature updates of IQS and IQS P-C, and various time step sizes of Implicit Dis. (lower is better)}
+\label{fig:lra_rt_vs_err}
+\end{figure}
 

### 2018-08-29 doc/Papers/IQS_convergence/reviewer_notes.tex
Including runtime vs error for LRA stuff

@@
 \usepackage[usenames, dvipsnames]{xcolor}
+\usepackage{float}
 
@@
 
-\item[\fix] \medm{ For the comparison of LRA and TREAT results, the efficiency [how accurate can you be given a fixed execution time or given an accuracy how much execution time do you need] is a very important quantity; the authors discuss it for LRA and TREAT.
+\item[\working] \medm{ For the comparison of LRA and TREAT results, the efficiency [how accurate can you be given a fixed execution time or given an accuracy how much execution time do you need] is a very important quantity; the authors discuss it for LRA and TREAT.
 However, the authors first present results for the errors and then runtime results.
@@
 \begin{itemize}
-\item Is just a plot of error vs. runtime sufficient?
+\item Fig. 17 visualizes the dependence of runtime on error. This figure shows that IQS and IQS P-C generally has a lower runtime for certain error than implicit discretization.
+\item I need help intepreting this result
 \end{itemize}
+\includegraphics[height=3in]{figures/lra_rt_vs_err.png}
 

### 2018-08-29 doc/Papers/IQS_convergence/figures/lra_rt_vs_err.m

@@ -1,0 +1,35 @@
+clear; clc; close all
+
+a = 0.5;
+
+bf.err = [1.407e-2 3.174e-3 7.690e-4 1.892e-4 4.590e-5];
+bf.rt = [4.11 6.01 10.38 21.91 25.23];
+bf.dt = [4e-3, 2e-3, 1e-3, 0.5e-3, 0.25e-3];
+bf.eff = bf.err.^a.*bf.rt;
+
+iqs.err = [2.612e-3 9.893e-4 5.796e-4 4.772e-4 4.516e-4];
+iqs.rt = [3.96 6.02 7.87 12.61 22.14];
+iqs.nt = [1, 2, 4, 8, 16];
+iqs.eff = iqs.err.^a.*iqs.rt;
+
+iqspc.err = [3.488e-3 1.349e-3 9.161e-4 8.052e-4 7.905e-4];
+iqspc.rt = [2.91 3.73 3.97 5.39 8.19];
+iqspc.nt = [1, 2, 4, 8, 16];
+iqspc.eff = iqspc.err.^a.*iqspc.rt;
+
+figure
+loglog(iqs.err,iqs.rt,'o-', iqspc.err,iqspc.rt,'o-', bf.err,bf.rt,'o-')
+xlabel('Error')
+ylabel('Runtime (hr)')
+legend('IQS, dt=0.004, N_T=1, 2, 4, 8, 16','IQS P-C, dt=0.004, N_T=1, 2, 4, 8, 16','Implicit Dis., dt=4e-3, 2e-3, 1e-3, 0.5e-3, 0.25e-3')
+grid on
+
+figure
+semilogy(iqs.nt,iqs.eff,'o-', iqspc.nt,iqspc.eff,'o-')
+hold on
+nn = linspace(min(iqs.nt),max(iqs.nt),100);
+for i=1:length(bf.eff)
+    eff = ones(size(nn))*bf.eff(i);
+    semilogy(nn,eff,'--');
+end
+

### 2018-09-20 doc/Papers/IQS_convergence/IQS_convergence.tex
Adding spatial profile of error for LRA

@@
 The spatial discretization uses continuous finite elements with first-order Lagrangian basis functions.
-The mesh consists of blocks $11\times 11$ with five uniform refinements, totaling $165,165$ elements and $124,609$ nodes.
+The mesh consists of blocks $11\times 11$ with five uniform refinements, totaling $123,904$ elements and $124,609$ nodes.
 Three different temporal solution techniques are used: implicit discretization of the flux equations, IQS formulation, and \iqspc.
@@
 
+All the values for error in this section are errors in the globally integrated flux.
+However, it is noteworthy to compare the spatial dependence of error.
+\fig{fig:lra_block} shows the difference in normalized power at each block of the LRA geometry for each method.
+The resulting spatially dependent error is in general an order of magnitude lower for IQS and IQS P-C than implicit discretization.
+This observation corresponds well with the globally integrated flux errors shown in Tables \ref{tab:ndiff_lra}-\ref{tab:iqspc_lra}.
+
+\begin{figure}[htbp!]
+\centering
+\resizebox{\textwidth}{!}{%
+\begin{tikzpicture}[every node/.style = {font=\normalsize}]
+
+\node[inner sep=0pt, anchor=south](fig) at (0,0) {\includegraphics[height=5in]{figures/lra_block.png}};
+\node[rectangle, draw, align=center,anchor=north east, style={font=\footnotesize}] (des) at (1,0) {\textbf{Normalized Power (Ref.)} \\ Difference (IQS) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (IQS P-C) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (Implicit Dis.) - $\Delta t = 0.004$};
+\node[anchor= north west] (dif) at (1,0) {Difference $= \frac{\text{Power}^\text{ref} - \text{Power}}{\text{Power}^\text{ref}}$};
+
+\end{tikzpicture}%
+}
+\caption{Difference in normalized power from reference solution for each fissile block of LRA geometry}
+\label{fig:lra_block}
+\end{figure}
+
 %----------------------------------------------------------------------------------------

### 2018-09-20 doc/Papers/IQS_convergence/reviewer_notes.tex
Adding spatial profile of error for LRA

@@
 \usepackage{float}
+\usepackage{tikz}
 
@@
 
-\item[\fix] \hard{ LRA and TREAT only compare errors computed with the peak powers.
+\item[\done] \hard{ LRA and TREAT only compare errors computed with the peak powers.
 Peak powers are an integral quantity and may favor IQS over spatial kinetics.
@@
 \begin{itemize}
-\item This would require a rerun of all simulations. There is a postprocessor in MOOSE that could do it, but the simulations would be ridiculously tedious.
+\item All the values for error in this section are errors in the globally integrated flux.
+However, it is noteworthy to compare the spatial dependence of error.
+Fig.
+18 shows the difference in normalized power at each block of the LRA geometry for each method.
+The resulting spatially dependent error is in general an order of magnitude lower for IQS and IQS P-C than implicit discretization.
+This observation corresponds well with the globally integrated flux errors shown in Tables 6-8.
 \end{itemize}
+\resizebox{\textwidth}{!}{%
+\begin{tikzpicture}[every node/.style = {font=\normalsize}]
+\node[inner sep=0pt, anchor=south](fig) at (0,0) {\includegraphics[height=5in]{figures/lra_block.png}};
+\node[rectangle, draw, align=center,anchor=north east, style={font=\footnotesize}] (des) at (1,0) {\textbf{Normalized Power (Ref.)} \\ Difference (IQS) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (IQS P-C) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (Implicit Dis.) - $\Delta t = 0.004$};
+\node[anchor= north west] (dif) at (1,0) {Difference $= \frac{\text{Power}^\text{ref} - \text{Power}}{\text{Power}^\text{ref}}$};
+
+\end{tikzpicture}%
+}
 
@@
 
-\item[\fix] \easy{ LRA benchmark: the number of elements makes no sense 11x11 = 121 and when using five uniform refinement steps one would end up with 123,904 elements. }
+\item[\done] \easy{ LRA benchmark: the number of elements makes no sense 11x11 = 121 and when using five uniform refinement steps one would end up with 123,904 elements. }
 \begin{itemize}
-\item  Not sure how we got these numbers, but he's right
+\item  Fixed it to 123,904 elements, number of nodes was originally correct
 \end{itemize}

### 2018-09-20 doc/Papers/IQS_convergence/figures/lra_rt_vs_err.m

@@ -1,35 +1,54 @@
 clear; clc; close all
 
-a = 0.5;
-
 bf.err = [1.407e-2 3.174e-3 7.690e-4 1.892e-4 4.590e-5];
 bf.rt = [4.11 6.01 10.38 21.91 25.23];
 bf.dt = [4e-3, 2e-3, 1e-3, 0.5e-3, 0.25e-3];
-bf.eff = bf.err.^a.*bf.rt;
+bf.eff = bf.err.*bf.rt;
 
 iqs.err = [2.612e-3 9.893e-4 5.796e-4 4.772e-4 4.516e-4];
 iqs.rt = [3.96 6.02 7.87 12.61 22.14];
 iqs.nt = [1, 2, 4, 8, 16];
-iqs.eff = iqs.err.^a.*iqs.rt;
+iqs.eff = iqs.err.*iqs.rt;
 
 iqspc.err = [3.488e-3 1.349e-3 9.161e-4 8.052e-4 7.905e-4];
 iqspc.rt = [2.91 3.73 3.97 5.39 8.19];
 iqspc.nt = [1, 2, 4, 8, 16];
-iqspc.eff = iqspc.err.^a.*iqspc.rt;
+iqspc.eff = iqspc.err.*iqspc.rt;
 
 figure
-loglog(iqs.err,iqs.rt,'o-', iqspc.err,iqspc.rt,'o-', bf.err,bf.rt,'o-')
-xlabel('Error')
-ylabel('Runtime (hr)')
-legend('IQS, dt=0.004, N_T=1, 2, 4, 8, 16','IQS P-C, dt=0.004, N_T=1, 2, 4, 8, 16','Implicit Dis., dt=4e-3, 2e-3, 1e-3, 0.5e-3, 0.25e-3')
+lw = 1.5;
+ms = 6;
+loglog(iqs.err,iqs.rt,'o-','LineWidth',lw,'MarkerSize',ms)
+hold on
+loglog(iqspc.err,iqspc.rt,'o-','LineWidth',lw,'MarkerSize',ms)
+loglog(bf.err,bf.rt,'o-','LineWidth',lw,'MarkerSize',ms)
+set(gca,'YLim',[2.91 40])
+xlabel('Error','Interpreter','latex','FontSize',14)
+ylabel('Runtime (hr)','Interpreter','latex','FontSize',14)
[…]

### 2018-09-20 doc/Papers/IQS_convergence/IQS_convergence.tex
Addressing all of reviewer comments

@@
 However, switching the precursors solve to be performed semi-analytically (\eqt{eq:dnp_an}) significantly improves the convergence in the uniqueness factor $K$, see \fig{fig:iter_err_an}.
+This result implies that uniqueness of the shape and amplitude can only be maintained with a consistent treatment of precursor term in the shape equation.
 
@@
 Our results seem to indicate that a high-order discretization of the flux equations may reduce the usefulness of the IQS approach (at least for large domains with weakly coupled spatial regions).
+It should be noted that the baseline for these error computations was performed using an adaptive method for the flux equations with a tight tolerance.
+The proper convergence rates seen by the methods with coarser time steps show that this baseline is accurate enough for computing error.
 %This paper shows an analysis of the first publicized application of IQS with higher than second order discretization, which exposed unforeseen properties of IQS.
@@
 \fig{fig:TWIGL_power} shows the IQS  solution as compared with the implicit flux solution.
-This illustrates that IQS can yield a much more accurate solution, even at a significantly larger time step than the implicit flux discretization.
+%This illustrates that IQS can yield a much more accurate solution, even at a significantly larger time step than the implicit flux discretization.
 
@@
 When there is moderate variation in the shape over the transient, the PRKE parameters are accurately estimated, and the fine-scale PRKE solve provides a well resolved amplitude solution.
+The baseline was computed using the BDF2 scheme on the flux equation with a time step of $10^{-5}$.
+The proper error convergences show that this solution is accurate enough to compare to for error computation for most of the convergence data points.
+Smallest time step scheme for IQS shows a break in the linearity of the convergence.
+These convergence results are consistent with the hypothesis that the performance of IQS is highly dependent on the spatial coupling of the flux.
 
@@
 It also shows that traditional IQS performed better with large $e_{tol}$, while \iqspc was better with smaller $e_{tol}$.
+This observation implies that the time adaptive options must be chosen differently when applied to a solver for shape versus a solver for flux.
 
@@
 Crank-Nicolson time discretization scheme is used for the space-time equations.
-The performance of IQS with temperature feedback is assessed by comparing the solution accuracy at peak power.
+%The performance of IQS with temperature feedback is assessed by comparing the solution accuracy at peak power.
 
@@
 These specific runs are highlighted in Tables \ref{tab:ndiff_lra} - \ref{tab:iqspc_lra}.
-\fig{fig:lra_rt_vs_err} visualizes the dependence of runtime on error. This figure shows that IQS and IQS P-C generally has a lower runtime for certain error than implicit discretization.
+\fig{fig:lra_rt_vs_err} visualizes the dependence of runtime on error.
+This figure shows that IQS and IQS P-C generally has a lower runtime for certain error than implicit discretization.
+Furthermore, note that the slopes on the right half of the \iqspc runs are shallower than the implicit discretization slope.
+This shows that by adding more temperature updates, within the 1-4 temperature updates region, \iqspc suffers a smaller increase in runtime for a decrease in error.
+However, the error in the shape function saturates the full error after adding more than four updates, leading to an inelastic dependence on runtime.
 

### 2018-09-20 doc/Papers/IQS_convergence/reviewer_notes.tex
Addressing all of reviewer comments

@@
 \usepackage{tikz}
+
+\newcommand{\iqspc}{IQS-PC}
 
@@
 
-\item[\fix] \medm{  Results are often just stated and not discussed, e.g.
+\item[\working] \medm{  Results are often just stated and not discussed, e.g.
 page 19 "It also shows that traditional IQS performed better with large etol, while IQS-PC was better with smaller etol".
@@
 \begin{itemize}
-\item I will remove the first referenced quote.
-\item I will add that if one expecting a certain tolerance to be reached in the iterative process, the semi-analytical integration of the precursors is necessary.
-Since, in the majority of cases, this evaluation is negligible compared to the evaluation of the shape, it is fortuitous to apply this technique in all cases.
+\item Page 16: Added "This result implies that uniqueness of the shape and amplitude can only be maintained with a consistent treatment of precursor term in the shape equation.
+\item Page 18: Removed "This illustrates that IQS can yield a much more accurate solution, even at a significantly larger time step than the implicit flux discretization."
+\item Page 20: Added "These convergence results are consistent with the hypothesis that the performance of IQS is highly dependent on the spatial coupling of the flux."
+\item Page 20: Added "This observation implies that the time adaptive options must be chosen differently when applied to a solver for shape versus a solver for flux."
 \end{itemize}
 
-\item[\working] \medm{ For the comparison of LRA and TREAT results, the efficiency [how accurate can you be given a fixed execution time or given an accuracy how much execution time do you need] is a very important quantity; the authors discuss it for LRA and TREAT.
+\item[\done] \medm{ For the comparison of LRA and TREAT results, the efficiency [how accurate can you be given a fixed execution time or given an accuracy how much execution time do you need] is a very important quantity; the authors discuss it for LRA and TREAT.
 However, the authors first present results for the errors and then runtime results.
@@
 \begin{itemize}
-\item Fig. 17 visualizes the dependence of runtime on error. This figure shows that IQS and IQS P-C generally has a lower runtime for certain error than implicit discretization.
-\item I need help intepreting this result
+\item Fig.
+17 visualizes the dependence of runtime on error.
+This figure shows that IQS and IQS P-C generally has a lower runtime for certain error than implicit discretization.
+Furthermore, note that the slopes on the right half of the \iqspc runs are shallower than the implicit discretization slope.
+This shows that by adding more temperature updates, within the 1-4 temperature updates region, \iqspc suffers a smaller increase in runtime for a decrease in error.
+However, the error in the shape function saturates the full error after adding more than four updates, leading to an inelastic dependence on runtime.
 \end{itemize}
@@
 
-\item[\fix] \medm{  Whenever errors are computed, baseline calculations are performed. How do the authors know that these baseline calculations are accurate enough to measure the error w.r.t. them? }
+\item[\done] \medm{  Whenever errors are computed, baseline calculations are performed. How do the authors know that these baseline calculations are accurate enough to measure the error w.r.t. them? }
 \begin{itemize}
-\item Can we just say that a proper error convergence means that the baseline is good enough?
+\item Page 18: Added "It should be noted that the baseline for these error computations was performed using an adaptive method for the flux equations with a tight tolerance.
+The proper convergence rates seen by the methods with coarser time steps show that this baseline is accurate enough for computing error."
+\item Page 20: Added "The baseline was computed using the BDF2 scheme on the flux equation with a time step of $10^{-5}$.
+The proper error convergences show that this solution is accurate enough to compare to for error computation for most of the convergence data points.
+Smallest time step scheme for IQS shows a break in the linearity of the convergence."
 \end{itemize}

### 2018-09-20 doc/Papers/IQS_convergence/IQS_convergence.tex
Making some small grammatical changes, ready for Ragusa review

@@
 \label{tab:op_def}
+\def\arraystretch{2}
 \begin{tabular}{|c|c|c|c|}
@@
 solution has converged from shape iterations.
-It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order Hermite interpolation is necessary.
+It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order interpolation is necessary.
 
@@
 between the amplitude's fine time scale and the shape's coarse time scale. 
-This idea stems from the fact that in a adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
+This idea stems from the fact that in an adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
 Because of the heat capacity of the fuel, the temperature amplitude is more slowly varying in time than the heat inflow from the flux amplitude, while still considerably faster than the time variance of the flux shape.
@@
 
-The precursors' equations, \eqt{eq:prec}, form a system of ODES. % ordinary differential equations.
+The precursors' equations, \eqt{eq:prec}, form a system of ODEs. % ordinary differential equations.
 A theta-scheme is often employed to evaluate the precursor concentrations:
@@
 \item Executioner - Establishes the type of simulation. The Transient type moves the simulation in time conducting spatial evaluations at each step.
-\item Auxiliary variable and kernel - "Optional" variables that live on the same mesh as the solution and are computed using auxiliary kernels. 
-\item Postprocessor - Computes scalar values over the entire spatial mesh, usually in an integral sense.
+\item Auxiliary variable and kernel - "Optional" variables that live on the same mesh as the solution and are computed algebraically using auxiliary kernels. 
+\item Postprocessor - Computes scalar values over the entire spatial mesh, usually involves integrated quantities.
 \item User-object - A generic type of postprocessor that allows connectivity of relevant quantities between different MOOSE objects.
@@
 The proper error convergences show that this solution is accurate enough to compare to for error computation for most of the convergence data points.
-Smallest time step scheme for IQS shows a break in the linearity of the convergence.
+The smallest time step scheme for IQS shows a break in the linearity of the convergence.
 These convergence results are consistent with the hypothesis that the performance of IQS is highly dependent on the spatial coupling of the flux.
@@
 This section shows the time step error convergence of IQS for the LRA benchmark, as well as the effect of the intermediate temperature time scale.
-\fig{fig:lra_bad} is an error convergence plot comparing the three tempoeral solution techniques where temperature is evaluated only once per macro step (1 temperature update).
+\fig{fig:lra_bad} is an error convergence plot comparing the three temporal solution techniques where temperature is evaluated only once per macro step (1 temperature update).
 \fig{fig:lra_mpconv} is an error convergence plot comparing the same three techniques when temperature is evaluated 5 times within a macro step (5 temperature updates).
@@
 It is also important to compare the error of implicit discretization with IQS at one update and \iqspc at 4 updates.
-IQS shows an error comparable to implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by -34.1\%.
-\iqspc shows an error less than implicit discretization at $\Delta t = 0.002$, signifying an actual increase in runtime by $<-34.9\%$.
+IQS shows an error comparable to implicit discretization at $\Delta t = 0.002$, signifying an actual increase in run-time by -34.1\%.
+\iqspc shows an error less than implicit discretization at $\Delta t = 0.002$, signifying an actual increase in run-time by $<-34.9\%$.
 These specific runs are highlighted in Tables \ref{tab:ndiff_lra} - \ref{tab:iqspc_lra}.
-\fig{fig:lra_rt_vs_err} visualizes the dependence of runtime on error.
-This figure shows that IQS and IQS P-C generally has a lower runtime for certain error than implicit discretization.
+\fig{fig:lra_rt_vs_err} visualizes the dependence of run-time on error.
+This figure shows that IQS and IQS P-C generally has a lower run-time for certain error than implicit discretization.
 Furthermore, note that the slopes on the right half of the \iqspc runs are shallower than the implicit discretization slope.
-This shows that by adding more temperature updates, within the 1-4 temperature updates region, \iqspc suffers a smaller increase in runtime for a decrease in error.
-However, the error in the shape function saturates the full error after adding more than four updates, leading to an inelastic dependence on runtime.
+This shows that by adding more temperature updates, within the 1-4 temperature updates region, \iqspc suffers a lesser increase in run-time for a decrease in error.
+However, the error in the shape function saturates the full error after adding more than four updates, leading to an inelastic dependence on run-time.
 
@@
 \hline
-Run  &  $\Delta t$ & Error$^*$ & Runtime (hr) & Linear Iter.\\
+Run  &  $\Delta t$ & Error$^*$ & Run-time (hr) & Linear Iter.\\
 \hline
@@
 \hline
-	&  Temperature 	&  		& Runtime 	& \% Increase	\\
-Run	&  Updates 	& Error$^*$ & (hr)		& in Runtime$^{\dagger}$\\
+	&  Temperature 	&  		& Run-time 	& \% Increase	\\
+Run	&  Updates 	& Error$^*$ & (hr)		& in Run-time$^{\dagger}$\\
 \hline
@@
 \includegraphics[height=3in]{figures/lra_rt_vs_err.png}
-\caption{Runtime vs. error for various number of temperature updates of IQS and IQS P-C, and various time step sizes of Implicit Dis. (lower is better)}
+\caption{Run-time vs. error for various number of temperature updates of IQS and IQS P-C, and various time step sizes of Implicit Dis. (lower is better)}
 \label{fig:lra_rt_vs_err}
@@
 
-\fig{fig:lra_dt2} shows the power profile of the LRA with time a of implicit discretization and \iqspc, and \tbl{tab:lra_dt2} compiles the results.
+\fig{fig:lra_dt2} shows the power profile of the LRA while using time adaptivity for implicit discretization and \iqspc; \tbl{tab:lra_dt2} compiles these results.
 These adaptivity results show the significant decrease in the number of macro time steps required for \iqspc.
@@
  & No.	& Max &	Time at Max &	Max Average &	\% Increase &	Max Power & Linear	\\
-Method & of Steps &	Power (W) &	Power (s) &	Temperature (K) & Runtime$^*$ & Error & Iterations \\
+Method & of Steps &	Power (W) &	Power (s) &	Temperature (K) & Run-time$^*$ & Error & Iterations \\
 \hline
@@
 $^*$ difference in run-time from implicit discretization 
-\caption{Transient-15 Error and Runtime Results}
+\caption{Transient-15 Error and Run-time Results}
 \label{tab:tran15}

### 2018-09-20 doc/Papers/IQS_convergence/figures/lra_rt_vs_err.m

@@ -24,7 +24,7 @@ loglog(iqspc.err,iqspc.rt,'o-','LineWidth',lw,'MarkerSize',ms)
 loglog(bf.err,bf.rt,'o-','LineWidth',lw,'MarkerSize',ms)
 set(gca,'YLim',[2.91 40])
 xlabel('Error','Interpreter','latex','FontSize',14)
-ylabel('Runtime (hr)','Interpreter','latex','FontSize',14)
+ylabel('Run-time (hr)','Interpreter','latex','FontSize',14)
 leg = {'IQS, $\Delta t=$0.004, $N_T=$1, 2, 4, 8, 16',...
        'IQS P-C, $\Delta t=$0.004, $N_T=$1, 2, 4, 8, 16',...
        'Implicit Dis., $\Delta t$=4e-3, 2e-3, 1e-3, 0.5e-3, 0.25e-3'};

### 2018-09-24 doc/Papers/IQS_convergence/IQS_convergence.tex
Some small fixes to the bibliography and reviewer notes

@@
 In neutron transport, the neutron flux solution lives in a seven dimensional phase-space, dependent on time, space, energy, and direction.
-The neutron diffusion equation reduces this phase-space by elimnating direction.
+The neutron diffusion equation reduces this phase-space by eliminating direction.
 IQS involves factorizing the neutron flux solution into a time-only-dependent component, the amplitude, and a full phase-space component, the shape \cite{Ott_1966,Devooght_1984,Monier_diss,Sissaoui_1995,Dulla2008}.
@@
 all three variables. \fig{fig:proc} shows the programming diagram implement to execute this process. 
-Where $N_T$ is the number of temperature updates per macro step.
+$N_T$ is the number of temperature updates per macro step.
 %The time increment of $\Delta t/3$ for the temperature solve is arbitrary and is meant only to match \fig{fig:time} where three temperature updates have been used as illustration.
@@
 
-Both the standard IQS and \iqspc methods have been implemented in the Multiphysics Object-Oriented Simulation Environment, MOOSE \cite{moose}, as part of its radiation transport solver, Rattlesnake \cite{wang2013}.
+Both the standard IQS and \iqspc methods have been implemented in the Multiphysics Object-Oriented Simulation Environment, MOOSE \cite{moose}, as part of its radiation transport solver, Rattlesnake \cite{Rattlesnake_theory}.
 All multi-dimensional results presented here have been computed using the diffusion solver of Rattlesnake.
@@
 The equations being solved are explicitly defined in problem 14-A1 of the ANL Benchmark Problem Book \cite{ANL_BPB}.
-The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.
+The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{Rattlesnake_user}.
 The spatial discretization uses continuous finite elements with first-order Lagrangian basis functions.
@@
 \fig{fig:lra_rt_vs_err} visualizes the dependence of run-time on error.
-This figure shows that IQS and IQS P-C generally has a lower run-time for certain error than implicit discretization.
+This figure shows that IQS and \iqspc generally has a lower run-time for certain error than implicit discretization.
 Furthermore, note that the slopes on the right half of the \iqspc runs are shallower than the implicit discretization slope.
@@
 \includegraphics[height=3in]{figures/lra_rt_vs_err.png}
-\caption{Run-time vs. error for various number of temperature updates of IQS and IQS P-C, and various time step sizes of Implicit Dis. (lower is better)}
+\caption{Run-time vs. error for various number of temperature updates of IQS and \iqspc, and various time step sizes of Implicit Dis. (lower is better)}
 \label{fig:lra_rt_vs_err}
@@
 \fig{fig:lra_block} shows the difference in normalized power at each block of the LRA geometry for each method.
-The resulting spatially dependent error is in general an order of magnitude lower for IQS and IQS P-C than implicit discretization.
+The resulting spatially dependent error is in general an order of magnitude lower for IQS and \iqspc than implicit discretization.
 This observation corresponds well with the globally integrated flux errors shown in Tables \ref{tab:ndiff_lra}-\ref{tab:iqspc_lra}.
@@
 \node[inner sep=0pt, anchor=south](fig) at (0,0) {\includegraphics[height=5in]{figures/lra_block.png}};
-\node[rectangle, draw, align=center,anchor=north east, style={font=\footnotesize}] (des) at (1,0) {\textbf{Normalized Power (Ref.)} \\ Difference (IQS) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (IQS P-C) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (Implicit Dis.) - $\Delta t = 0.004$};
+\node[rectangle, draw, align=center,anchor=north east, style={font=\footnotesize}] (des) at (1,0) {\textbf{Normalized Power (Ref.)} \\ Difference (IQS) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (\iqspc) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (Implicit Dis.) - $\Delta t = 0.004$};
 \node[anchor= north west] (dif) at (1,0) {Difference $= \frac{\text{Power}^\text{ref} - \text{Power}}{\text{Power}^\text{ref}}$};

### 2018-09-24 doc/Papers/IQS_convergence/reviewer_notes.tex
Some small fixes to the bibliography and reviewer notes

@@
 \usepackage{tikz}
+\usepackage{xspace}
 
-\newcommand{\iqspc}{IQS-PC}
+\newcommand{\iqspc}{IQS-PC\xspace}
 
@@
 
-\newcommand{\easy}[1]{\textcolor{ForestGreen}{#1}}
-\newcommand{\medm}[1]{\textcolor{BurntOrange}{#1}}
-\newcommand{\hard}[1]{\textcolor{Red}{#1}}
+%\newcommand{\easy}[1]{\textcolor{ForestGreen}{#1}}
+%\newcommand{\medm}[1]{\textcolor{BurntOrange}{#1}}
+%\newcommand{\hard}[1]{\textcolor{Red}{#1}}
+\newcommand{\easy}[1]{\textbf{\textit{#1}}}
+\newcommand{\medm}[1]{\textbf{\textit{#1}}}
+\newcommand{\hard}[1]{\textbf{\textit{#1}}}
 
@@
 
-\item[\done] \medm{ Justify page 8: "we expect temperature to be more rapidly varying than the shape but less so than the amplitude" }
+\item[\done] \medm{ Justify page 8: ``we expect temperature to be more rapidly varying than the shape but less so than the amplitude" }
 \begin{itemize}
-\item This idea stems from the fact that in a adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
+\item Page 9: ``This idea stems from the fact that in a adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
 Because of the heat capacity of the fuel, the temperature amplitude is more slowly varying in time than the heat inflow from the flux amplitude, while still considerably faster than the time variance of the flux shape.
-This logic is empirically defended in the results of Section 5.
+This logic is empirically defended in the results of Section 5."
 \end{itemize}
 
-\item[\done] \easy{ Justify page 5: "Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself, this allowing for larger time step sizes in updating the shape".
+\item[\done] \easy{ Justify page 5: ``Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself, this allowing for larger time step sizes in updating the shape".
 }
 \begin{itemize}
-\item Reference first IQS paper
+\item Page 5: References [1] as justification for this statement.
 \end{itemize}
@@
 \item[\working] \medm{  Results are often just stated and not discussed, e.g.
-page 19 "It also shows that traditional IQS performed better with large etol, while IQS-PC was better with smaller etol".
+page 19 ``It also shows that traditional IQS performed better with large etol, while IQS-PC was better with smaller etol".
 The statement simply bloats the manuscripts; discuss the reasons and implications or remove it.
-Another example is page 15, "However, switching the precursors solve to be performed semi-analytically…".
+Another example is page 15, ``However, switching the precursors solve to be performed semi-analytically…".
 In this case, the authors should take the time to discuss the implications of this finding.
@@
 \begin{itemize}
-\item Page 16: Added "This result implies that uniqueness of the shape and amplitude can only be maintained with a consistent treatment of precursor term in the shape equation.
-\item Page 18: Removed "This illustrates that IQS can yield a much more accurate solution, even at a significantly larger time step than the implicit flux discretization."
-\item Page 20: Added "These convergence results are consistent with the hypothesis that the performance of IQS is highly dependent on the spatial coupling of the flux."
-\item Page 20: Added "This observation implies that the time adaptive options must be chosen differently when applied to a solver for shape versus a solver for flux."
+\item Page 16: Added ``This result implies that uniqueness of the shape and amplitude can only be maintained with a consistent treatment of precursor term in the shape equation."
+\item Page 18: Removed ``This illustrates that IQS can yield a much more accurate solution, even at a significantly larger time step than the implicit flux discretization."
+\item Page 20: Added ``These convergence results are consistent with the hypothesis that the performance of IQS is highly dependent on the spatial coupling of the flux."
+\item Page 20: Added ``This observation implies that the time adaptive options must be chosen differently when applied to a solver for shape versus a solver for flux."
 \end{itemize}
@@
 Then they try to synthesize efficiency by comparing different cases that have different runtimes and different errors, e.g.
-page 23 right before Table 5 "IQS-PC shows and error less than implicit discretization at delta t = 0.002, signifying …".
+page 23 right before Table 5 ``IQS-PC shows and error less than implicit discretization at delta t = 0.002, signifying …".
 I suggest the authors create error vs.
@@
 \begin{itemize}
-\item Fig.
-17 visualizes the dependence of runtime on error.
-This figure shows that IQS and IQS P-C generally has a lower runtime for certain error than implicit discretization.
+\item Page 24: ``Fig.
+17 visualizes the dependence of run-time on error.
+This figure shows that IQS and IQS P-C generally has a lower run-time for certain error than implicit discretization.
 Furthermore, note that the slopes on the right half of the \iqspc runs are shallower than the implicit discretization slope.
-This shows that by adding more temperature updates, within the 1-4 temperature updates region, \iqspc suffers a smaller increase in runtime for a decrease in error.
-However, the error in the shape function saturates the full error after adding more than four updates, leading to an inelastic dependence on runtime.
+This shows that by adding more temperature updates, within the 1-4 temperature updates region, \iqspc suffers a smaller increase in run-time for a decrease in error.
+However, the error in the shape function saturates the full error after adding more than four updates, leading to an inelastic dependence on run-time.
+"
 \end{itemize}
-\includegraphics[height=3in]{figures/lra_rt_vs_err.png}
+%\includegraphics[height=3in]{figures/lra_rt_vs_err.png}
 
 \item[\done] \easy{ In the results section, Error is sometimes used without introducing it, e.g.
-page 25, "Table 9 shows the error ….".
-Further "Error" is often used as y-label in plots., e.g.
-in the discussion of Figure 10 in which "convergence results" are presented; the sentence is not clear in this context because it was not stated if this is iterative convergence or convergence to the right solution as the time step is reduced.
+page 25, ``Table 9 shows the error ….".
+Further ``Error" is often used as y-label in plots., e.g.
+in the discussion of Figure 10 in which ``convergence results" are presented; the sentence is not clear in this context because it was not stated if this is iterative convergence or convergence to the right solution as the time step is reduced.
 In some instances error is defined but it is tedious for the reader to find the corresponding passage in the text, figures, or tables.
@@
 \begin{itemize}
-\item Included an explicit definition of error in each of the plots and tables
+\item Page 17-25: Included an explicit definition of error in each of the plots and tables
 \end{itemize}
@@
 \begin{itemize}
-\item All the values for error in this section are errors in the globally integrated flux.
+\item Page 24-26: ``All the values for error in this section are errors in the globally integrated flux.
 However, it is noteworthy to compare the spatial dependence of error.
@@
 The resulting spatially dependent error is in general an order of magnitude lower for IQS and IQS P-C than implicit discretization.
-This observation corresponds well with the globally integrated flux errors shown in Tables 6-8.
+This observation corresponds well with the globally integrated flux errors shown in Tables 6-8."
 \end{itemize}
-\resizebox{\textwidth}{!}{%
-\begin{tikzpicture}[every node/.style = {font=\normalsize}]
-\node[inner sep=0pt, anchor=south](fig) at (0,0) {\includegraphics[height=5in]{figures/lra_block.png}};
-\node[rectangle, draw, align=center,anchor=north east, style={font=\footnotesize}] (des) at (1,0) {\textbf{Normalized Power (Ref.)} \\ Difference (IQS) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (IQS P-C) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (Implicit Dis.) - $\Delta t = 0.004$};
-\node[anchor= north west] (dif) at (1,0) {Difference $= \frac{\text{Power}^\text{ref} - \text{Power}}{\text{Power}^\text{ref}}$};
-
-\end{tikzpicture}%
-}
+%\resizebox{\textwidth}{!}{%
+%\begin{tikzpicture}[every node/.style = {font=\normalsize}]
+%\node[inner sep=0pt, anchor=south](fig) at (0,0) {\includegraphics[height=5in]{figures/lra_block.png}};
+%\node[rectangle, draw, align=center,anchor=north east, style={font=\footnotesize}] (des) at (1,0) {\textbf{Normalized Power (Ref.)} \\ Difference (IQS) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (IQS P-C) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (Implicit Dis.) - $\Delta t = 0.004$};
+%\node[anchor= north west] (dif) at (1,0) {Difference $= \frac{\text{Power}^\text{ref} - \text{Power}}{\text{Power}^\text{ref}}$};
+%
+%\end{tikzpicture}%
+%}
 
@@
 \begin{itemize}
-\item Page 18: Added "It should be noted that the baseline for these error computations was performed using an adaptive method for the flux equations with a tight tolerance.
+\item Page 18: Added ``It should be noted that the baseline for these error computations was performed using an adaptive method for the flux equations with a tight tolerance.
 The proper convergence rates seen by the methods with coarser time steps show that this baseline is accurate enough for computing error."
-\item Page 20: Added "The baseline was computed using the BDF2 scheme on the flux equation with a time step of $10^{-5}$.
+\item Page 20: Added ``The baseline was computed using the BDF2 scheme on the flux equation with a time step of $10^{-5}$.
 The proper error convergences show that this solution is accurate enough to compare to for error computation for most of the convergence data points.
@@
 
-\item[\done] \easy{ Page 1: "… the amplitude, and a space- and time-dependent component, the shape [1,2,3,4,5]". What about energy? }
+\item[\done] \easy{ Page 1: ``… the amplitude, and a space- and time-dependent component, the shape [1,2,3,4,5]". What about energy? }
 \begin{itemize}
-\item IQS involves factorizing the neutron flux solution into a time-only-dependent component, the amplitude, and a full phase-space component, the shape
+\item Page 1: ``IQS involves factorizing the neutron flux solution into a time-only-dependent component, the amplitude, and a full phase-space component, the shape."
 \end{itemize}
@@
 \begin{itemize}
-\item Note that $\phi^g$ for neutron transport is the angular neutron flux ($\phi^g(\vec r, \vec\Omega, t)$), while for neutron diffusion it is scalar flux ($\phi^g(\vec r, t)$).
+\item Page 3: ``Note that $\phi^g$ for neutron transport is the angular neutron flux ($\phi^g(\vec r, \vec\Omega, t)$), while for neutron diffusion it is scalar flux ($\phi^g(\vec r, t)$)."
 \end{itemize}
@@
 \begin{itemize}
-\item Created a table of the operator definitions (Table 1)
-\item Do I need to include a definition of all the properties?
+\item Page 4: Created a table of the operator definitions (Table 1)
 \end{itemize}
@@
 \begin{itemize}
-\item See Table 1
+\item Page 4: See Table 1
 \end{itemize}
@@
 \begin{itemize}
-\item In neutron transport, the neutron flux solution lives in a seven dimensional phase-space, dependent on time, space, energy, and direction.
-The neutron diffusion equation reduces this phase-space by elimnating direction.
+\item Page 1: ``In neutron transport, the neutron flux solution lives in a seven dimensional phase-space, dependent on time, space, energy, and direction.
+The neutron diffusion equation reduces this phase-space by eliminating direction."
 \end{itemize}
 
-\item[\done] \easy{ Algorithm in section 2.1 on page 5. Step 2: "linearly interpolate". It is later stated that interpolation of higher order is done for high order schemes. }
+\item[\done] \easy{ Algorithm in section 2.1 on page 5. Step 2: ``linearly interpolate". It is later stated that interpolation of higher order is done for high order schemes. }
 \begin{itemize}
-\item Removed linearly from step 2
-\item It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order Hermite interpolation is necessary.
+\item Page 5: Removed linearly from step 2
+\item Page 6: ``It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order interpolation is necessary."
 \end{itemize}
@@
 \begin{itemize}
-\item I included the reference for this error estimation.
+\item Included the reference for this error estimation.
 \end{itemize}
@@
 \begin{itemize}
-\item For the purposes of this derivation, the adiabatic model, given in Eq. (11), has linear heat capacity. However, a nonlinear heat capacity can easily be implemented with a iterative procedure.
-\item Put $\rho C_p(T)$ inside the derivative 
+\item Page 8: ``For the purposes of this derivation, the adiabatic model, given in Eq.
+(11), has linear heat capacity.
+However, a nonlinear heat capacity can easily be implemented with a iterative procedure."
+\item Page 28: Put $\rho C_p(T)$ inside the derivative 
 \end{itemize}
@@
 \begin{itemize}
-\item I'll change the 3 to $N_{T}$ in Figure 2, which is the number of temperature solves within a macro step
+\item Page 10: Included the variable $N_{T}$ in Figure 2
+\item Page 9: $N_T$ is the number of temperature updates per macro step.
 \end{itemize}
 
-\item[\done] \medm{ Statement page 9: "Furthermore, a very accurate representation of p(t) over the macro step is available from the PRKE solve…".
+\item[\done] \medm{ Statement page 9: ``Furthermore, a very accurate representation of p(t) over the macro step is available from the PRKE solve…".
 I think it should be precise or else one would have to show that the equations are solved accurately.
-Second, what does "very accurate" mean, "very accurate" is comparison to what?
+Second, what does ``very accurate" mean, ``very accurate" is comparison to what?
 }
 \begin{itemize}
-\item Furthermore, a highly refined representation of $p(t)$ over the macro step is available from the PRKE solve (micro time scale).
-Using this fine scale information yields a more precise integration of amplitude than assuming an interpolation of the amplitude between macro step points.
+\item Page 11: ``Furthermore, a highly refined representation of $p(t)$ over the macro step is available from the PRKE solve (micro time scale).
+Using this fine scale information yields a more precise integration of amplitude than assuming an interpolation of the amplitude between macro step points."
 \end{itemize}
@@
 \begin{itemize}
+\item Page 13:
+\begin{itemize}
 \item Kernel - Evaluation of the weak-form residual for a particular piece of physics.
 \item Executioner - Establishes the type of simulation. The Transient type moves the simulation in time conducting spatial evaluations at each step.
-\item Auxiliary variable and kernel - "Optional" variables that live on the same mesh as the solution and are computed using auxiliary kernels. 
-\item Postprocessor - Computes scalar values over the entire spatial mesh, usually in an integral sense.
+\item Auxiliary variable and kernel - ``Optional" variables that live on the same mesh as the solution and are computed algebraically using auxiliary kernels. 
+\item Postprocessor - Computes scalar values over the entire spatial mesh, usually involves integrated quantities.
 \item User-object - A generic type of postprocessor that allows connectivity of relevant quantities between different MOOSE objects.
-\item A more detailed and comprehensive description of MOOSE objects can be found in [18]
+A more detailed and comprehensive description of MOOSE objects can be found in [18]
+\end{itemize}
 \end{itemize}
 
-\item[\done] \easy{ Page 13 top of the page "looping over cells to evaluate", what cells, mesh cells? }
+\item[\done] \easy{ Page 13 top of the page ``looping over cells to evaluate", what cells, mesh cells? }
 \begin{itemize}
-\item Spatial mesh cells
+\item Page 14: ``The PRKE parameters are written as user-objects, looping over spatial mesh cells to evaluate them."
 \end{itemize}
@@
 \begin{itemize}
-\item  Fixed it to 123,904 elements, number of nodes was originally correct
+\item Page 22: Fixed it to 123,904 elements, number of nodes was originally correct
 \end{itemize}
@@
 \begin{itemize}
-\item  Seems tedious to provide them, so will reference the book
+\item  Page 22: ``The equations being solved are explicitly defined in problem 14-A1 of the ANL Benchmark Problem Book [20]"
 \end{itemize}
@@
 \begin{itemize}
-\item Put a note at the end of the intro of each results section
+\item Page 14 and 22: ``Note that any instance of $\Delta t$ in the following figures and commentary signifies the size of the macro time step."
 \end{itemize}
 
-\item[\done] \easy{ Conclusions, page 27: incomplete sentence "IQS showed expected error convergence up". }
+\item[\done] \easy{ Conclusions, page 27: incomplete sentence ``IQS showed expected error convergence up". }
 \begin{itemize}
-\item ...through fourth-order discretization of the shape
+\item Page 31: ``...through fourth-order time discretization of the shape"
 \end{itemize}
@@
 \begin{itemize}
-\item Already stated: These run-times are based on total alive time of the execution where the diffusion evaluation is distributed over 24 processors.
+\item Already stated on Page 24: ``These run-times are based on total alive time of the execution where the diffusion evaluation is distributed over 24 processors."
+\end{itemize}
+
+\end{itemize}
+
+\section*{Bibliography}
+
+\begin{itemize}
+
+\item[\done] \easy{ MOOSE reference is out of date. Recommend newer publication }
+\begin{itemize}
+\item See [16]
+\end{itemize}
+
+\item[\done] \easy{ Rattlesnake reference [17] is out of date. Recommend citing the theory manual INL/EXT-17-42103. }
+\begin{itemize}
+\item See [17]
+\end{itemize}
+
+\item[\done] \easy{ The LRA benchmark is extensively discussed in the Rattlesnake user manual, which might be a handy reference to use: INL/EXT-15-37337. }
+\begin{itemize}
+\item Page 22: ``The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) [25]"
 \end{itemize}

### 2018-09-24 doc/Papers/IQS_convergence/figures/lra_rt_vs_err.m

@@ -26,7 +26,7 @@ set(gca,'YLim',[2.91 40])
 xlabel('Error','Interpreter','latex','FontSize',14)
 ylabel('Run-time (hr)','Interpreter','latex','FontSize',14)
 leg = {'IQS, $\Delta t=$0.004, $N_T=$1, 2, 4, 8, 16',...
-       'IQS P-C, $\Delta t=$0.004, $N_T=$1, 2, 4, 8, 16',...
+       'IQS-PC, $\Delta t=$0.004, $N_T=$1, 2, 4, 8, 16',...
        'Implicit Dis., $\Delta t$=4e-3, 2e-3, 1e-3, 0.5e-3, 0.25e-3'};
 legend(leg(:),'Interpreter','latex','FontSize',12)
 grid on

### 2018-10-01 doc/Papers/IQS_convergence/IQS_convergence.tex
JCR review added to revision-1

@@
 Nonlinear convergence of each of these techniques is investigated.
-Verification of IQS with analysis of time step convergence is also investigated in order to investigate the method's effectiveness with high-order schemes.
+Verification of IQS with analysis of time step convergence is also \tcr{investigated in order to investigate} the method's effectiveness with high-order schemes.
 The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicolson, and backward difference formulae (BDF).
@@
 \begin{keyword}
-Nuclear reactor dynamics \sep Quasi-static method \sep Temperature feedback \sep Time adaptivity 
+Nuclear reactor dynamics \sep Improved quasi-static (IQS) method \sep Temperature feedback \sep Time adaptivity 
 \end{keyword}
@@
 In neutron transport, the neutron flux solution lives in a seven dimensional phase-space, dependent on time, space, energy, and direction.
-The neutron diffusion equation reduces this phase-space by eliminating direction.
+The neutron diffusion approximation reduces this phase-space by eliminating direction.
 IQS involves factorizing the neutron flux solution into a time-only-dependent component, the amplitude, and a full phase-space component, the shape \cite{Ott_1966,Devooght_1984,Monier_diss,Sissaoui_1995,Dulla2008}.
@@
 The shape solution satisfies a modified time-dependent neutron balance equation.
-The rationale for the IQS method lies in the assumption that the shape function is weakly dependent on time.
+The rationale for the IQS method lies in the observations that the shape function is less rapidly dependent on time than the amplitude function.
 Therefore, it is expected that the modified time-dependent neutron balance equations be less stiff than the original time-dependent neutron balance equations for the flux.
 As a result, the shape may not require to be solved for at the same frequency as the amplitude.
-Instead, shape is evaluated only on larger macro-time steps, which is expected to yield wall clock savings, especially in multi-dimensional geometries.
+Instead, the shape is evaluated only on larger macro-time steps, which is expected to yield wall-clock savings, especially in multi-dimensional geometries.
 The PRKE form a small system of ordinary differential equations (ODE) and solving them on a fine temporal grid is not a computational burden.
-As opposed to the standard PRKE approach, whereby a shape function is selected, typically once for an entire transient without updates, the IQS technique is obtained in a rigorously consistent manner from the time-dependent neutron balance equation.
+As opposed to the standard PRKE approach, whereby a shape function is selected, typically once for the entire transient without any updates, the PRKE used in the IQS technique are obtained in a rigorously consistent manner from the time-dependent neutron balance equation.
 
@@
 Later, in \cite{Ott_1969}, Ott incorporated the time derivative of shape in the equation, yielding better results but requiring a fixed-point approach to resolve the nonlinear coupling between the amplitude and shape equations.
-This also led to the technique's name: the Improved Quasi-Static method although one may argue that such a denomination does not make it immediately clear that the technique rigorously solves a time-dependent problem.
+This also led to the technique's name: the Improved Quasi-Static method although one may argue that such a denomination does not make it immediately clear that the technique solves rigorously a time-dependent problem.
 
@@
 Another approach is to linearize the equations, so that no nonlinear iteration is necessary.
-IQS can be linearized using the IQS Predictor-Corrector method (\iqspc) \cite{dulla2006}.
+For instance, IQS can be linearized using the IQS Predictor-Corrector method (\iqspc) \cite{dulla2006}.
 \iqspc entails evaluating the flux equation then correcting its amplitude using an amplitude evaluation.
@@
 This paper (i) discusses the different nonlinear iteration techniques for IQS and tests the rigor of their implementation, (ii) investigates high-order temporal discretization of the shape equation, and (iii) analyzes IQS coupled with adiabatic heat up and cross-section feedback.
-Prior IQS publications do not go beyond first-order time discretization for the shape equation.
+Prior IQS publications \tcr{do not go beyond first-order time discretization for the shape equation\footnote{are you sure?
+I thought the Canadians used a $\theta$ scheme, thus included 2nd order Crank-Nicolson}}.
 Here, we test IQS with higher-order schemes (up to fourth order) and investigate the method's effectiveness with such temporal discretizations.
@@
 \centering
-\caption{Operator definition of multi-group neutron conservation equations}
+\caption{Operator definition of multi-group neutron conservation equations\tcr{why do you have transfers in $\chi_p$ for $P$?}}
 \label{tab:op_def}
@@
 \hline
-$P_p^{g'\to g}$ & Prompt fission & $\frac{1}{4\pi}\chi_p^{g'\rightarrow g}\frac{\nu_p\Sigma_f^{g'}}{k_{eff}}\int_{4\pi}d\Omega '$ & $\chi_p^{g'\rightarrow g}\frac{\nu_p\Sigma_f^{g'}}{k_{eff}}$ \\
-\hline
-$L^g$ & Streaming/Removal & $\vec \Omega \cdot \grad + \Sigma_t^g$ & $-\div D^g \grad + \Sigma_a^g$ \\
+$P_p^{g'\to g}$ & Prompt fission & $\frac{1}{4\pi}\chi_p^{g'\rightarrow g}\frac{\nu_p\Sigma_f^{g'}}{\keff}\int_{4\pi}d\Omega '$ & $\chi_p^{g'\rightarrow g}\frac{\nu_p\Sigma_f^{g'}}{\keff}$ \\
+\hline
+$L^g$ & Streaming/Removal & $\vec \Omega \cdot \grad + \Sigma_t^g$ & $-\div D^g \grad + \Sigma_t^g$ \\
 \hline
@@
 \hline
-$P_{d,i}^g$ & Precursor production & $\beta_i\frac{\nu\Sigma_f^{g}}{k_{eff}}\int_{4\pi}d\Omega$ & $\beta_i\frac{\nu\Sigma_f^{g}}{k_{eff}}$ \\
+$P_{d,i}^g$ & Precursor production & $\beta_i\frac{\nu\Sigma_f^{g}}{\keff}\int_{4\pi}d\Omega$ & $\beta_i\frac{\nu\Sigma_f^{g}}{\keff}$ \\
 \hline
@@
 \end{table}
-
-where \tbl{tab:op_def} shows the definition of each operator for both the transport and diffusion case.
-Note that $\phi^g$ for neutron transport is the angular neutron flux ($\phi^g(\vec r, \vec\Omega, t)$), while for neutron diffusion it is scalar flux ($\phi^g(\vec r, t)$).
+\tcr{in transport, shouldn't $P^g_{d,i}$ also have a $1/4\pi$?}
+%
+\noindent
+where \tbl{tab:op_def} shows the definition of each operator for both the transport and diffusion cases.
+Note that $\phi^g$ for neutron transport is the angular neutron flux ($\phi^g(\vec r, \vec\Omega, t)$), while for neutron diffusion it is the scalar flux ($\phi^g(\vec r, t)$).
 %where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the leakage and interaction operator, 
@@
 the error convergence rate of high order temporal discretization schemes for shape, higher order interpolation of the parameters is required.
-A third-order implicit Runge-Kutta \cite{sdirk33} discretization with time-step adaptation is used for all PRKE evaluations in this paper in
+In this paper, a third-order implicit Runge-Kutta \cite{sdirk33} discretization with time-step adaptation is used for all PRKE solves in
 order to ensure a negligibly small error in the amplitude solution.
@@
 solution has converged from shape iterations.
-It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order interpolation is necessary.
+It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order interpolation was found to be necessary.
 
@@
 \end{itemize}
+\tcr{Why don't you have absolute values in the last 3 criteria?}
 
@@
 Dulla et al.
-in \cite{Dulla2008} also describes an in-depth comparison of \iqspc with traditional IQS, which originally brought the method to light.
+in \cite{Dulla2008} also describes an in-depth comparison of \iqspc with traditional IQS.
+%, which originally brought the method to light.
 
@@
 \be
-\rho c_p \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
+\rho C_p \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
 \label{eq:temp}
@@
 \be
-T_{n+1} = T_n + \frac{\kappa_f}{\rho c_p} \left(a_2 \varphi_{n+1} + a_1 \varphi_{n}\right)
+T_{n+1} = T_n + \frac{\kappa_f}{\rho C_p} \left(a_2 \varphi_{n+1} + a_1 \varphi_{n}\right)
 \label{eq:temp_an}
@@
 This idea stems from the fact that in an adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
-Because of the heat capacity of the fuel, the temperature amplitude is more slowly varying in time than the heat inflow from the flux amplitude, while still considerably faster than the time variance of the flux shape.
+Because of the heat capacity of the fuel, the temperature amplitude is more slowly varying in time than the heat \tcr{?inflow?} from the flux amplitude, while still considerably faster than the time \tcr{?variance?} of the flux shape.
 This logic is empirically defended in the results of Section 5.
@@
 time scale until convergence on each temperature step.  Then another iterative process needs to occur in the shape time scale on 
-all three variables. \fig{fig:proc} shows the programming diagram implement to execute this process. 
+all three variables. \fig{fig:proc} shows the solution workflow implemented to execute this process. 
 $N_T$ is the number of temperature updates per macro step.
@@
 The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps ($\varphi^g_{\Delta t/2}$).
-Note: $\varphi$ is changed to $\phi$ in the case of the \iqspc technique.
+Note: $\varphi$ (shape) is changed to $\phi$ (flux) in the case of the \iqspc technique.
 The relative error is computed as follows:
@@
 
-To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied to direct temporal discretization of the flux equation (no IQS), the standard IQS method, and the \iqspc version.
+To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied \tcr{?to direct temporal discretization?} of the flux equation (no IQS), the standard IQS method, and the \iqspc version.
 Each of these methods will be applied to several diffusion problems; the number of time steps taken and the resulting error will be used to compare the methods.
@@
 Most of the shape equation kernels are the same as the flux kernels already implemented in Rattlesnake for its standard diffusion (or transport) solution, and only two kernels had to be modified.
-\eqt{eq:kernels} shows which kernels needed to be written or amended (denoted by an $\star$) to solve the shape equation in MOOSE.
+\eqt{eq:kernels} shows which kernels needed to be written or amended (denoted by an $^\star$) to solve the shape equation in MOOSE.
 The PRKE ODE was written as a sub-function in the Transient executioner, which advances the simulation in time and determines when to perform a shape (or flux for \iqspc) solve.
@@
 \begin{align}
-\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Prompt Production Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Scattering Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Diffusion Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Reaction Kernel} \nonumber \\
-& - \underbrace{\frac{1}{v^g} \boxed{\overbrace{\frac{1}{p}\frac{dp}{dt}}^{From PRKE}}\varphi^g}_{IQS Reaction Kernel\ ^\star}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Precursor Kernel\ ^\star}
+\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{\textit{Prompt Production Kernel}} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{\textit{Scattering Kernel}} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{\textit{Diffusion Kernel}} - \underbrace{\Sigma_r^g\varphi^g}_{\textit{Reaction Kernel}} \nonumber \\
+& - \underbrace{\frac{1}{v^g} \boxed{\overbrace{\frac{1}{p}\frac{dp}{dt}}^{\textit{From PRKE}}}\varphi^g}_{\textit{IQS Reaction Kernel}^\star}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{\textit{Precursor Kernel}^\star}
 \label{eq:kernels}
@@
 The second test is from the ANL Benchmark Problem Book (BPB) \cite{ANL_BPB}.
-Note that any instance of $\Delta t$ in the following figures and commentary signifies the size of the macro time step.
+Note that any instance of $\Delta t$ in the following figures and commentary refers to the size of the macro time step.
 
@@
 \begin{center}
-\caption{1-D slab absorption cross-section at times of interest}
+\caption{1-D slab absorption cross sections at times of interest}
 \label{tab:1Dslope}
@@
 \includegraphics[width=0.99\textwidth]{figures/1D_power_base.png}
-\caption{Power profile over transient}
+\caption{Normalized total power as a function of time}
 \label{fig:1D_tot_power}
@@
 This figures shows that criteria based on the  $L^{\infty}$ norm of the shape, the $L^2$ norm of the shape, the reactivity change, and the amplitude have approximately the same convergence behavior, but the criteria based on the uniqueness constant $K$ never reaches the tight tolerance and runs up to the maximum number of iterations allowed (20 in this case).
-\fig{fig:iter_err} shows the error in the factor $K$ saturates to different levels, whether the shape is rescaled or not.
+\fig{fig:iter_err} shows that the error in the uniqueness factor $K$ saturates to different levels, whether the shape is rescaled or not.
 %Rescaling shape more frequently helps the error.
@@
 Next, we investigate the use of higher-order temporal methods to solve the shape equations (standard IQS and \iqspc methods).
-We compare these results with the same high-order discretization for the flux equations.
+We compare these results with the same high-order discretization of the flux equations.
 %Time step convergence analysis involves evaluating a problem with various refinements in step size and comparing the resulting errors with the time step size.
@@
 It should be noted that the baseline for these error computations was performed using an adaptive method for the flux equations with a tight tolerance.
-The proper convergence rates seen by the methods with coarser time steps show that this baseline is accurate enough for computing error.
+The proper convergence rates seen by the methods with coarser time steps show that this baseline is accurate enough for computing the error.
 %This paper shows an analysis of the first publicized application of IQS with higher than second order discretization, which exposed unforeseen properties of IQS.
@@
 \hline\hline
-No. of Spatial Nodes	& 3872 		& 1936 \\
+No. of Spatial Nodes	& 3,872 		& 1,936 \\
 Eigenvalue 				& 0.99637	& 0.99637 \\
-No. of Time Steps 		& 6000 		& 23,890 \\
+No. of Time Steps 		& 6,000 		& 23,890 \\
 Time to Peak Power (s) 	& 1.441 	& 1.441 \\
-Peak Power (W/cm$^3$) 	& 5456 		& 5461 \\
+Peak Power (W/cm$^3$) 	& 5,456 		& 5,461 \\
 \hline
@@
 Finally, \fig{fig:mp} shows the effect of various temperature updates.
-The dashed lines correspond to implicit flux discretization at different step sizes, while the IQS macro step size is kept constant.
+The dashed lines correspond to the implicit flux discretization solutions obtained using different time step sizes, while the IQS macro step size is kept constant.
 
@@
 \fig{fig:lra_rt_vs_err} visualizes the dependence of run-time on error.
-This figure shows that IQS and \iqspc generally has a lower run-time for certain error than implicit discretization.
+This figure shows that IQS and \iqspc generally have a lower run-time for a certain error than implicit discretization.
 Furthermore, note that the slopes on the right half of the \iqspc runs are shallower than the implicit discretization slope.
-This shows that by adding more temperature updates, within the 1-4 temperature updates region, \iqspc suffers a lesser increase in run-time for a decrease in error.
-However, the error in the shape function saturates the full error after adding more than four updates, leading to an inelastic dependence on run-time.
+This shows that by adding more temperature updates, in the 1-4 temperature-updates range, \iqspc suffers a lesser increase in run-time for a decrease in error.
+However, beyond 4 temperature updates, the error in the shape function saturates and no additional benefits can be gained by more frequent temperature updates.
 
@@
 
-All the values for error in this section are errors in the globally integrated flux.
-However, it is noteworthy to compare the spatial dependence of error.
-\fig{fig:lra_block} shows the difference in normalized power at each block of the LRA geometry for each method.
-The resulting spatially dependent error is in general an order of magnitude lower for IQS and \iqspc than implicit discretization.
-This observation corresponds well with the globally integrated flux errors shown in Tables \ref{tab:ndiff_lra}-\ref{tab:iqspc_lra}.
+Up to now, errors presented in this section were norms (i.e., space-integrated flux errors).
+However, for reactor physics problems, it is also noteworthy to compare the spatial dependence of error, for example, error in the power distribution.
+\fig{fig:lra_block} shows the difference in power for each assembly of the LRA geometry for each method.
+The resulting spatially dependent error is in general an order of magnitude lower for IQS and \iqspc than for the implicit flux discretization.
+This observation agrees well with the integrated flux errors shown in Tables \ref{tab:ndiff_lra}-\ref{tab:iqspc_lra}.
 
@@
 }
-\caption{Difference in normalized power from reference solution for each fissile block of LRA geometry}
+\caption{Difference in normalized power from reference solution for each fissile block of the LRA geometry}
 \label{fig:lra_block}
@@
 
-The Transient-15 model uses an adiabatic temperature feedback mechanism, similar to the one used in the LRA benchmark.
-\eqt{eq:temp2} describes the heat up of the fuel, where, this time, the fuel specific heat capacity is itself also temperature-dependent.
-Hence, the temperature evaluation is identical to the one described in LRA section, except a Newton iteration process is employed to resolve the nonlinearity from the specific heat term.
+The Transient-15 model uses an adiabatic temperature feedback mechanism, similar to the one used in the LRA benchmark (see, Eq.~\eqref{eq:temp}, except this time, the fuel specific heat capacity is itself also temperature-dependent).
+Hence, the temperature evaluation is identical to the one described in LRA section, except a Newton iteration process is employed to resolve the nonlinearity from the specific heat capacity term.
 The 11-g neutron cross sections are linearly interpolated at tabulated temperatures.
-
-\be
-\frac{\partial }{\partial t}\left[\rho c_p(T)T(\vec{r},t)\right] = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
-\label{eq:temp2}
-\ee
+%
+%\be
+%\frac{\partial }{\partial t}\left[\rho c_p(T)T(\vec{r},t)\right] = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
+%\label{eq:temp2}
+%\ee
 

### 2018-10-01 doc/Papers/IQS_convergence/reviewer_notes.tex
JCR review added to revision-1

@@
 
+\usepackage{color}
+\newcommand{\tcr}[1]{\textcolor{red}{#1}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
+%%%%%%%%%%%%%%%%%%%%%%%%
 
+\tcr{The different bullet sizes will be a cause for question.
+Make the reply to the reviewer uniform.
+Do NOT forgot to thank them as needed for every good suggestion/recommendation.
+I've done some of that.
+People review for free on their free time.
+We have to be grateful.}
 \section*{Major Issues}
@@
 \begin{itemize}
-\item Page 9: ``This idea stems from the fact that in a adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
-Because of the heat capacity of the fuel, the temperature amplitude is more slowly varying in time than the heat inflow from the flux amplitude, while still considerably faster than the time variance of the flux shape.
-This logic is empirically defended in the results of Section 5."
+\item We have added the following on Page 9: ``This idea stems from the fact that in a adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
+Because of the heat capacity/diffusivity of the fuel, the temperature amplitude is often more slowly varying in time than neutron amplitude function itself, while still considerably faster varying than the flux shape.'' This conjecture is empirically defended in the results of Section 5, for instance, the new Figure 17.
 \end{itemize}
@@
 \begin{itemize}
-\item Page 5: References [1] as justification for this statement.
+\item Page 5: References [1] as justification for this statement (and maybe its first empirical numerical verification.
+However, since K.
+Ott, many other researchers used and published IQS results.
+Our literature reviews includes several of these references, some recent, more older.
 \end{itemize}
@@
 }
+Thank you for the recommendations. The following sentences have been added:
 \begin{itemize}
@@
 \begin{itemize}
-\item Page 24: ``Fig.
+\item Thank you for the suggestion.
+We have added this on Page 24: ``Fig.
 17 visualizes the dependence of run-time on error.
@@
 Furthermore, note that the slopes on the right half of the \iqspc runs are shallower than the implicit discretization slope.
-This shows that by adding more temperature updates, within the 1-4 temperature updates region, \iqspc suffers a smaller increase in run-time for a decrease in error.
-However, the error in the shape function saturates the full error after adding more than four updates, leading to an inelastic dependence on run-time.
+This shows that by adding more temperature updates, in the 1-4 temperature-updates range, \iqspc suffers a lesser increase in run-time for a decrease in error.
+However, beyond 4 temperature updates, the error in the shape function saturates and no additional benefits can be gained by more frequent temperature updates.
 "
@@
 \begin{itemize}
-\item Page 17-25: Included an explicit definition of error in each of the plots and tables
+\item On  Pages 17-25: We included an explicit definition of error in each of the plots and tables. Thank you for the suggestion.
 \end{itemize}
@@
 \begin{itemize}
-\item Page 24-26: ``All the values for error in this section are errors in the globally integrated flux.
-However, it is noteworthy to compare the spatial dependence of error.
+\item Thank you for the suggestion.
+On Page 24-26, we added: ``Up to now, errors presented in this section were norms (i.e., space-integrated flux errors).
+However, for reactor physics problems, it is also noteworthy to compare the spatial dependence of error, for example, error in the power distribution.
 Fig.
-18 shows the difference in normalized power at each block of the LRA geometry for each method.
-The resulting spatially dependent error is in general an order of magnitude lower for IQS and IQS P-C than implicit discretization.
-This observation corresponds well with the globally integrated flux errors shown in Tables 6-8."
+18 shows the difference in power for each assembly of the LRA geometry for each method.
+The resulting spatially dependent error is in general an order of magnitude lower for IQS and IQS-PC than for the implicit flux discretization.
+This observation agrees well with the integrated flux errors shown in Tables 6-8.
+"
 \end{itemize}
@@
 \begin{itemize}
-\item Page 18: Added ``It should be noted that the baseline for these error computations was performed using an adaptive method for the flux equations with a tight tolerance.
-The proper convergence rates seen by the methods with coarser time steps show that this baseline is accurate enough for computing error."
-\item Page 20: Added ``The baseline was computed using the BDF2 scheme on the flux equation with a time step of $10^{-5}$.
+\item On Page 18: we added ``It should be noted that the baseline for these error computations was performed using an adaptive method for the flux equations with a tight tolerance.
+The proper convergence rates seen by the methods with coarser time steps show that this baseline is accurate enough for computing the error."
+\item On Page 20: We added ``The baseline was computed using the BDF2 scheme on the flux equation with a time step of $10^{-5}$.
 The proper error convergences show that this solution is accurate enough to compare to for error computation for most of the convergence data points.
@@
 \begin{itemize}
-\item Page 1: ``IQS involves factorizing the neutron flux solution into a time-only-dependent component, the amplitude, and a full phase-space component, the shape."
+\item We added this sentence on Page 1: ``IQS involves factorizing the neutron flux solution into a time-only-dependent component, the amplitude, and a full phase-space component, the shape."
 \end{itemize}
@@
 \begin{itemize}
-\item Page 4: Created a table of the operator definitions (Table 1)
+\item We agree. On Page 4: Created a table of the operator definitions (Table 1)
 \end{itemize}
@@
 \begin{itemize}
-\item Page 4: See Table 1
+\item We agree. On Page 4: Added in Table 1.
 \end{itemize}
@@
 \begin{itemize}
-\item Page 1: ``In neutron transport, the neutron flux solution lives in a seven dimensional phase-space, dependent on time, space, energy, and direction.
-The neutron diffusion equation reduces this phase-space by eliminating direction."
+\item We agree.
+On Page 1, we added: ``In neutron transport, the neutron flux solution lives in a seven dimensional phase-space, dependent on time, space, energy, and direction.
+The neutron diffusion approximation reduces this phase-space by eliminating direction."
 \end{itemize}
@@
 \begin{itemize}
-\item Page 5: Removed linearly from step 2
-\item Page 6: ``It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order interpolation is necessary."
+\item Page 5: Thank you for catching this. Indeed, we use high-order interpolation when a higher-order scheme is used. We removed ``linearly'' from step 2
+\item Page 6: We added: ``It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order interpolation is necessary."
 \end{itemize}
@@
 \begin{itemize}
-\item Included the reference for this error estimation.
+\item We included the reference for this error estimation.
 \end{itemize}
@@
 \begin{itemize}
-\item Page 8: ``For the purposes of this derivation, the adiabatic model, given in Eq.
-(11), has linear heat capacity.
-However, a nonlinear heat capacity can easily be implemented with a iterative procedure."
-\item Page 28: Put $\rho C_p(T)$ inside the derivative 
+\item Whether the heat capacity depends on temperature or not, the term  $\rho C_p(T)$ should always be outside of the temporal derivative.
+This form is given, for instance, in the Todreas \& Kazimi textbook ``Nuclear Systems I: Thermal-hydraulic fundamentals'', as Equation 4-123.
 \end{itemize}
@@
 \begin{itemize}
-\item Page 10: Included the variable $N_{T}$ in Figure 2
-\item Page 9: $N_T$ is the number of temperature updates per macro step.
+\item Page 10: We included the variable $N_{T}$ in Figure 2 for clarity.
+\item $N_T$ is defined on Page 9: ``$N_T$ is the number of temperature updates per macro step.''
 \end{itemize}
@@
 \begin{itemize}
-\item Page 11: ``Furthermore, a highly refined representation of $p(t)$ over the macro step is available from the PRKE solve (micro time scale).
+\item Thank you for noting this.
+We improve the phrasing on Page 11: ``Furthermore, a highly refined representation of $p(t)$ over the macro step is available from the PRKE solve (micro time scale).
 Using this fine scale information yields a more precise integration of amplitude than assuming an interpolation of the amplitude between macro step points."
@@
 \begin{itemize}
-\item Page 13:
+\item We are correct. This has been added on Page 13:
 \begin{itemize}
@@
 \begin{itemize}
-\item Page 22: Fixed it to 123,904 elements, number of nodes was originally correct
+\item Thank you. On Page 22: Fixed it to 123,904 elements, number of nodes was originally correct.
 \end{itemize}
@@
 \begin{itemize}
-\item Page 14 and 22: ``Note that any instance of $\Delta t$ in the following figures and commentary signifies the size of the macro time step."
+\item On Page 14 and 22, we added: ``Note that any instance of $\Delta t$ in the following figures and commentary signifies the size of the macro time step."
 \end{itemize}
@@
 \begin{itemize}
-\item Page 31: ``...through fourth-order time discretization of the shape"
+\item Thank you. Page 31: ``...through fourth-order time discretization of the shape"
 \end{itemize}
@@
 \begin{itemize}
-\item See [16]
+\item Thank you. We replaced it with [16].
 \end{itemize}
@@
 \begin{itemize}
-\item See [17]
+\item Thank you. We replaced it with [17].
 \end{itemize}

### 2018-10-02 doc/Papers/IQS_convergence/IQS_convergence.tex
Addressing JCR comments to revision

@@
 Nonlinear convergence of each of these techniques is investigated.
-Verification of IQS with analysis of time step convergence is also \tcr{investigated in order to investigate} the method's effectiveness with high-order schemes.
+Verification of IQS with analysis of time step convergence is also investigated in order to examine the method's effectiveness with high-order schemes.
 The time derivative of the shape function is discretized through fourth order using implicit-Euler, Crank-Nicolson, and backward difference formulae (BDF).
@@
 Prior IQS publications \tcr{do not go beyond first-order time discretization for the shape equation\footnote{are you sure?
-I thought the Canadians used a $\theta$ scheme, thus included 2nd order Crank-Nicolson}}.
+I thought the Canadians used a $\theta$ scheme, thus included 2nd order Crank-Nicolson}}\tcb{\footnote{I am pretty sure of this, the Canadians only do $\theta$ scheme on the precursor equation, they do implicit Euler on the shape}}.
 Here, we test IQS with higher-order schemes (up to fourth order) and investigate the method's effectiveness with such temporal discretizations.
@@
 \centering
-\caption{Operator definition of multi-group neutron conservation equations\tcr{why do you have transfers in $\chi_p$ for $P$?}}
+\caption{Operator definition of multi-group neutron conservation equations}
 \label{tab:op_def}
@@
 \hline
-$P_p^{g'\to g}$ & Prompt fission & $\frac{1}{4\pi}\chi_p^{g'\rightarrow g}\frac{\nu_p\Sigma_f^{g'}}{\keff}\int_{4\pi}d\Omega '$ & $\chi_p^{g'\rightarrow g}\frac{\nu_p\Sigma_f^{g'}}{\keff}$ \\
+$P_p^{g'\to g}$ & Prompt fission & $\frac{1}{4\pi}\chi_p^{g}\frac{\nu_p\Sigma_f^{g'}}{\keff}\int_{4\pi}d\Omega '$ & $\chi_p^{g}\frac{\nu_p\Sigma_f^{g'}}{\keff}$ \\
 \hline
@@
 \hline
-$P_{d,i}^g$ & Precursor production & $\beta_i\frac{\nu\Sigma_f^{g}}{\keff}\int_{4\pi}d\Omega$ & $\beta_i\frac{\nu\Sigma_f^{g}}{\keff}$ \\
+$P_{d,i}^g$ & Precursor production & $\frac{1}{4\pi}\beta_i\frac{\nu\Sigma_f^{g}}{\keff}\int_{4\pi}d\Omega$ & $\beta_i\frac{\nu\Sigma_f^{g}}{\keff}$ \\
 \hline
@@
 \end{table}
-\tcr{in transport, shouldn't $P^g_{d,i}$ also have a $1/4\pi$?}
 %
@@
 \[
-\left(\rho/\Lambda\right)^{(k+1)}_{n+1} - \left(\rho/\Lambda\right)^{(k)}_{n+1} < \epsilon_{\rho}
+\left|\left(\rho/\Lambda\right)^{(k+1)}_{n+1} - \left(\rho/\Lambda\right)^{(k)}_{n+1}\right| < \epsilon_{\rho}
 \]
@@
 \[
-p_{n+1}^{(k+1)} - p_{n+1}^{(k)} < \epsilon_p
+\left|p_{n+1}^{(k+1)} - p_{n+1}^{(k)}\right| < \epsilon_p
 \]
@@
 \[
-\frac{K_{n+1}^{(k+1)} - K_0}{K_0} < \epsilon_K
+\left|\frac{K_{n+1}^{(k+1)} - K_0}{K_0}\right| < \epsilon_K
 \]
@@
 \end{itemize}
-\tcr{Why don't you have absolute values in the last 3 criteria?}
 
@@
 This idea stems from the fact that in an adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
-Because of the heat capacity of the fuel, the temperature amplitude is more slowly varying in time than the heat \tcr{?inflow?} from the flux amplitude, while still considerably faster than the time \tcr{?variance?} of the flux shape.
-This logic is empirically defended in the results of Section 5.
+Because of the heat capacity of the fuel, the temperature amplitude is often more slowly varying in time than neutron amplitude function itself, while still considerably faster varying than the flux shape.
+This conjecture is empirically defended in the results of Section 5.
 Hence, a possible solution process for a problem with temperature feedback will have three time scales, as
@@
 
-To investigate IQS's performance with step-doubling time adaptation, the adaptation will be applied \tcr{?to direct temporal discretization?} of the flux equation (no IQS), the standard IQS method, and the \iqspc version.
+To investigate IQS's performance with step-doubling time adaptation, the adaptation is applied to implicit discretization of the flux equation (no IQS), the standard IQS method, and the \iqspc version.
 Each of these methods will be applied to several diffusion problems; the number of time steps taken and the resulting error will be used to compare the methods.

### 2018-10-02 doc/Papers/IQS_convergence/reviewer_notes.tex
Addressing JCR comments to revision

@@
-\documentclass{elsarticle}
+\documentclass[11pt]{letter}
 
-\usepackage[letterpaper, margin=1in]{geometry}
-\usepackage{titlesec}
-\usepackage{enumitem}
-\usepackage{graphicx,amsmath, amssymb, amsthm, amsfonts,subcaption}
-\usepackage[usenames, dvipsnames]{xcolor}
-\usepackage{float}
-\usepackage{tikz}
+%%%%%% Letter Size Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+        \setlength{\textwidth}{6in}     %% For longer or shorter text width
+        \setlength{\topmargin}{0in}    %% For more or less top margin
+       \setlength{\textheight}{8.6in}    %% For longer or shorter textheight
+        \setlength{\oddsidemargin}{0.25in} %% For odd side margin (twoside)
+                                            %% or margin (oneside)
+        \setlength{\paperwidth}{8.5in}
+        \setlength{\paperheight}{11in}
+        \setlength{\topmargin}{0pt}
+        \setlength{\headheight}{0pt}
+        \setlength{\headsep}{0pt}
+        \setlength{\footskip}{0.4in}
+        \setlength{\marginparwidth}{0in}
+        \setlength{\marginparsep}{0in}
+        \setlength{\hoffset}{0in}
+        \setlength{\voffset}{0in}
+
+\address{Zachary M. Prince\\ 
+Department of Nuclear Engineering \\
+Texas A\&M University\\
+College Station, TX 77843-3133, USA\\
+phone: (360) 362 6622\\
+e-mail: zachmprince@tamu.edu \vspace{0.5cm}}
+
+%%%%%% The Signature  and Date %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\signature{\vspace{-1.25cm}Zachary M. Prince, Jean C. Ragusa}
+
 \usepackage{xspace}
@@
 \newcommand{\iqspc}{IQS-PC\xspace}
+\newcommand{\section}[1]{\textbf{#1}}
 
-\newcommand{\fix}{\textsquare}
-\newcommand{\working}{$\boxdot$}
-\newcommand{\nofix}{?}
-\newcommand{\done}{\checkmark}
+%\newcommand{\fix}{\textsquare}
+%\newcommand{\working}{$\boxdot$}
+%\newcommand{\nofix}{?}
+%\newcommand{\done}{\checkmark}
+\newcommand{\fix}{$\bullet$}
+\newcommand{\working}{$\bullet$}
+\newcommand{\nofix}{$\bullet$}
+\newcommand{\done}{$\bullet$}
 
@@
 %\newcommand{\hard}[1]{\textcolor{Red}{#1}}
-\newcommand{\easy}[1]{\textbf{\textit{#1}}}
-\newcommand{\medm}[1]{\textbf{\textit{#1}}}
-\newcommand{\hard}[1]{\textbf{\textit{#1}}}
+\newcommand{\easy}[1]{{\textit{#1}}}
+\newcommand{\medm}[1]{{\textit{#1}}}
+\newcommand{\hard}[1]{{\textit{#1}}}
 
-\usepackage{color}
-\newcommand{\tcr}[1]{\textcolor{red}{#1}}
+%\usepackage{color}
+%\newcommand{\tcr}[1]{\textcolor{red}{#1}}
 
@@
 
-\tcr{The different bullet sizes will be a cause for question.
-Make the reply to the reviewer uniform.
-Do NOT forgot to thank them as needed for every good suggestion/recommendation.
-I've done some of that.
-People review for free on their free time.
-We have to be grateful.}
-\section*{Major Issues}
+\begin{letter}{Professor Sara A. Pozzi\\
+    Editor,\\
+    Elsevier Annals of Nuclear Energy\\
+    \bigskip
+    Re: Resubmission of manuscript ``Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method'', ANUCENE-D-18-00462}
+\date{\today}
+%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%         \vspace{3.0cm}
+
+\opening{Dear Professor Pozzi,}
+         \vspace{0.25cm}
+%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Thank you for the opportunity to revise our manuscript, ``Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method''.
+We appreciate the prompt review and constructive suggestions.
+It is our belief that the manuscript is substantially improved after making the suggested edits.
+
+Overall, the comments from the reviewer were positive and pivotal in the improved quality of the manuscript.
+Following this letter are the review comments in italics with our responses below, including how and where the text was modified.
+
+We hope you will find the revised manuscript suitable for publication. Thank you again for considering this manuscript for publication in the {\it Annals of Nuclear Energy}.
+
+\vspace{0.25cm}
+
+%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%% The Closing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\closing{Sincerely, }
+
+\end{letter}
+
+\large{\textbf{Reviewer Comments}}\hrule\small
+
+\section{Major Issues}
 
@@
 \begin{itemize}
-\item We have added the following on Page 9: ``This idea stems from the fact that in a adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
-Because of the heat capacity/diffusivity of the fuel, the temperature amplitude is often more slowly varying in time than neutron amplitude function itself, while still considerably faster varying than the flux shape.'' This conjecture is empirically defended in the results of Section 5, for instance, the new Figure 17.
+\item We have added the following on Page 9: ``This idea stems from the fact that in an adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude.
+Because of the heat capacity of the fuel, the temperature amplitude is often more slowly varying in time than neutron amplitude function itself, while still considerably faster varying than the flux shape.'' This conjecture is empirically defended in the results of Section 5, for instance, the new Figure 17.
 \end{itemize}
@@
 \begin{itemize}
-\item Page 5: References [1] as justification for this statement (and maybe its first empirical numerical verification.
+\item Page 5: References [1] as justification for this statement and maybe its first empirical numerical verification.
 However, since K.
@@
 \item[\working] \medm{  Results are often just stated and not discussed, e.g.
-page 19 ``It also shows that traditional IQS performed better with large etol, while IQS-PC was better with smaller etol".
+page 19 ``It also shows that traditional IQS performed better with large etol, while \iqspc was better with smaller etol".
 The statement simply bloats the manuscripts; discuss the reasons and implications or remove it.
@@
 }
+
 Thank you for the recommendations. The following sentences have been added:
@@
 Then they try to synthesize efficiency by comparing different cases that have different runtimes and different errors, e.g.
-page 23 right before Table 5 ``IQS-PC shows and error less than implicit discretization at delta t = 0.002, signifying …".
+page 23 right before Table 5 ``\iqspc shows and error less than implicit discretization at delta t = 0.002, signifying …".
 I suggest the authors create error vs.
@@
 \end{itemize}
-%\includegraphics[height=3in]{figures/lra_rt_vs_err.png}
 
@@
 18 shows the difference in power for each assembly of the LRA geometry for each method.
-The resulting spatially dependent error is in general an order of magnitude lower for IQS and IQS-PC than for the implicit flux discretization.
+The resulting spatially dependent error is in general an order of magnitude lower for IQS and \iqspc than for the implicit flux discretization.
 This observation agrees well with the integrated flux errors shown in Tables 6-8.
@@
 \end{itemize}
-%\resizebox{\textwidth}{!}{%
-%\begin{tikzpicture}[every node/.style = {font=\normalsize}]
-%\node[inner sep=0pt, anchor=south](fig) at (0,0) {\includegraphics[height=5in]{figures/lra_block.png}};
-%\node[rectangle, draw, align=center,anchor=north east, style={font=\footnotesize}] (des) at (1,0) {\textbf{Normalized Power (Ref.)} \\ Difference (IQS) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (IQS P-C) - $\Delta t = 0.004, \ N_T = 4$ \\ Difference (Implicit Dis.) - $\Delta t = 0.004$};
-%\node[anchor= north west] (dif) at (1,0) {Difference $= \frac{\text{Power}^\text{ref} - \text{Power}}{\text{Power}^\text{ref}}$};
-%
-%\end{tikzpicture}%
-%}
 
 \item[\done] \medm{  Whenever errors are computed, baseline calculations are performed. How do the authors know that these baseline calculations are accurate enough to measure the error w.r.t. them? }
+
+Thank you for suggesting this clarification. We've made the following modifications for justification:
 \begin{itemize}
@@
 The proper convergence rates seen by the methods with coarser time steps show that this baseline is accurate enough for computing the error."
-\item On Page 20: We added ``The baseline was computed using the BDF2 scheme on the flux equation with a time step of $10^{-5}$.
+\item On Page 20: we added ``The baseline was computed using the BDF2 scheme on the flux equation with a time step of $10^{-5}$.
 The proper error convergences show that this solution is accurate enough to compare to for error computation for most of the convergence data points.
@@
 
-\section*{Minor Issues}
+\newpage
+\section{Minor Issues}
 
@@
 \begin{itemize}
-\item Page 3: ``Note that $\phi^g$ for neutron transport is the angular neutron flux ($\phi^g(\vec r, \vec\Omega, t)$), while for neutron diffusion it is scalar flux ($\phi^g(\vec r, t)$)."
+\item Thank you for catching this discrepancy.
+On Page 3 we've added: ``Note that $\phi^g$ for neutron transport is the angular neutron flux ($\phi^g(\vec r, \vec\Omega, t)$), while for neutron diffusion it is scalar flux ($\phi^g(\vec r, t)$)."
 \end{itemize}
@@
 \begin{itemize}
-\item Page 5: Thank you for catching this. Indeed, we use high-order interpolation when a higher-order scheme is used. We removed ``linearly'' from step 2
-\item Page 6: We added: ``It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order interpolation is necessary."
+\item Thank you for catching this. Indeed, we use high-order interpolation when a higher-order scheme is used. We removed ``linearly'' from step 2 on Page 5.
+\item We've also add on Page 6: ``It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order interpolation is necessary."
 \end{itemize}
@@
 \item Thank you for noting this.
-We improve the phrasing on Page 11: ``Furthermore, a highly refined representation of $p(t)$ over the macro step is available from the PRKE solve (micro time scale).
+We improved the phrasing on Page 11: ``Furthermore, a highly refined representation of $p(t)$ over the macro step is available from the PRKE solve (micro time scale).
 Using this fine scale information yields a more precise integration of amplitude than assuming an interpolation of the amplitude between macro step points."
@@
 \begin{itemize}
-\item We are correct. This has been added on Page 13:
+\item Thank you for the valuable suggestion. This has been added on Page 13:
 \begin{itemize}
@@
 \begin{itemize}
-\item Page 14: ``The PRKE parameters are written as user-objects, looping over spatial mesh cells to evaluate them."
+\item Thank you for suggesting this clarification. Page 14: ``The PRKE parameters are written as user-objects, looping over spatial mesh cells to evaluate them."
 \end{itemize}
@@
 \begin{itemize}
-\item Thank you. On Page 22: Fixed it to 123,904 elements, number of nodes was originally correct.
+\item Thank you, that number is correct. On Page 22: Fixed it to 123,904 elements, number of nodes was originally correct. 
 \end{itemize}
@@
 
-\section*{Bibliography}
+\section{Bibliography}
 
@@
 \begin{itemize}
-\item Page 22: ``The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) [25]"
+\item Thank you for suggesting to include this reference.
+We've included it on Page 22: ``The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) [25]"
 \end{itemize}

### 2018-10-02 doc/Papers/IQS_convergence/IQS_convergence.tex
Changing wording for prior time discretization

@@
 This paper (i) discusses the different nonlinear iteration techniques for IQS and tests the rigor of their implementation, (ii) investigates high-order temporal discretization of the shape equation, and (iii) analyzes IQS coupled with adiabatic heat up and cross-section feedback.
-Prior IQS publications \tcr{do not go beyond first-order time discretization for the shape equation\footnote{are you sure?
-I thought the Canadians used a $\theta$ scheme, thus included 2nd order Crank-Nicolson}}\tcb{\footnote{I am pretty sure of this, the Canadians only do $\theta$ scheme on the precursor equation, they do implicit Euler on the shape}}.
+Prior IQS publications do not typically present results that include more than first-order time discretization of the shape equation.
 Here, we test IQS with higher-order schemes (up to fourth order) and investigate the method's effectiveness with such temporal discretizations.

### 2018-10-02 doc/Papers/IQS_convergence/reply_to_editor.tex
minor changes. splitting reply in 2 because I think you need to upload 2 letters

\documentclass[11pt]{letter}

%%%%%% Letter Size Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \setlength{\textwidth}{6in}     %% For longer or shorter text width
        \setlength{\topmargin}{0in}    %% For more or less top margin
       \setlength{\textheight}{8.6in}    %% For longer or shorter textheight
        \setlength{\oddsidemargin}{0.25in} %% For odd side margin (twoside)
                                            %% or margin (oneside)
        \setlength{\paperwidth}{8.5in}
        \setlength{\paperheight}{11in}
        \setlength{\topmargin}{0pt}
        \setlength{\headheight}{0pt}
        \setlength{\headsep}{0pt}
        \setlength{\footskip}{0.4in}
        \setlength{\marginparwidth}{0in}
        \setlength{\marginparsep}{0in}
        \setlength{\hoffset}{0in}
        \setlength{\voffset}{0in}

\address{Zachary M. Prince\\ 
Department of Nuclear Engineering \\
Texas A\&M University\\
College Station, TX 77843-3133, USA\\
phone: (360) 362 6622\\
e-mail: zachmprince@tamu.edu \vspace{0.5cm}}

%%%%%% The Signature  and Date %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\signature{\vspace{-1.25cm}Zachary M. Prince, Jean C. Ragusa}

\usepackage{xspace}

\newcommand{\iqspc}{IQS-PC\xspace}
\newcommand{\section}[1]{\textbf{#1}}

%\newcommand{\fix}{\textsquare}
%\newcommand{\working}{$\boxdot$}
%\newcommand{\nofix}{?}
%\newcommand{\done}{\checkmark}
\newcommand{\fix}{$\bullet$}
\newcommand{\working}{$\bullet$}
\newcommand{\nofix}{$\bullet$}
\newcommand{\done}{$\bullet$}

%\newcommand{\easy}[1]{\textcolor{ForestGreen}{#1}}
%\newcommand{\medm}[1]{\textcolor{BurntOrange}{#1}}
%\newcommand{\hard}[1]{\textcolor{Red}{#1}}
\newcommand{\easy}[1]{{\textit{#1}}}
\newcommand{\medm}[1]{{\textit{#1}}}
\newcommand{\hard}[1]{{\textit{#1}}}

%\usepackage{color}
%\newcommand{\tcr}[1]{\textcolor{red}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%

\begin{letter}{Professor Sara A. Pozzi\\
    Editor,\\
    Elsevier Annals of Nuclear Energy\\
    \bigskip
    Re: Revision of manuscript ``Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method'', ANUCENE-D-18-00462}
\date{\today}
%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%         \vspace{3.0cm}

\opening{Dear Professor Pozzi,}
         \vspace{0.25cm}
%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Thank you for the opportunity to revise our manuscript, ``Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method''. We appreciate the prompt review and constructive suggestions by the reviewer. We believe our manuscript is substantially improved after making the suggested edits.

Overall, the comments from the reviewer were positive and pivotal in the improved quality of the manuscript. We are following a reply to the reviewer where the review comments are in italics with our responses below, including how and where the manuscript was modified.

We hope you will find the revised manuscript suitable for publication. Thank you again for considering this manuscript for publication in the {\it Annals of Nuclear Energy}.

\vspace{0.25cm}

%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%% The Closing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\closing{Sincerely, }

\end{letter}

\end{document}

### 2018-10-02 doc/Papers/IQS_convergence/reply_to_reviewer.tex
minor changes. splitting reply in 2 because I think you need to upload 2 letters

\documentclass[11pt]{letter}

%%%%%% Letter Size Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \setlength{\textwidth}{6in}     %% For longer or shorter text width
        \setlength{\topmargin}{0in}    %% For more or less top margin
       \setlength{\textheight}{8.6in}    %% For longer or shorter textheight
        \setlength{\oddsidemargin}{0.25in} %% For odd side margin (twoside)
                                            %% or margin (oneside)
        \setlength{\paperwidth}{8.5in}
        \setlength{\paperheight}{11in}
        \setlength{\topmargin}{0pt}
        \setlength{\headheight}{0pt}
        \setlength{\headsep}{0pt}
        \setlength{\footskip}{0.4in}
        \setlength{\marginparwidth}{0in}
        \setlength{\marginparsep}{0in}
        \setlength{\hoffset}{0in}
        \setlength{\voffset}{0in}

\address{Zachary M. Prince\\ 
Department of Nuclear Engineering \\
Texas A\&M University\\
College Station, TX 77843-3133, USA\\
phone: (360) 362 6622\\
e-mail: zachmprince@tamu.edu \vspace{0.5cm}}

%%%%%% The Signature  and Date %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\signature{\vspace{-1.25cm}Zachary M. Prince, Jean C. Ragusa}

\usepackage{xspace}

\newcommand{\iqspc}{IQS-PC\xspace}
\newcommand{\section}[1]{\textbf{#1}}

%\newcommand{\fix}{\textsquare}
%\newcommand{\working}{$\boxdot$}
%\newcommand{\nofix}{?}
%\newcommand{\done}{\checkmark}
\newcommand{\fix}{$\bullet$}
\newcommand{\working}{$\bullet$}
\newcommand{\nofix}{$\bullet$}
\newcommand{\done}{$\bullet$}

%\newcommand{\easy}[1]{\textcolor{ForestGreen}{#1}}
%\newcommand{\medm}[1]{\textcolor{BurntOrange}{#1}}
%\newcommand{\hard}[1]{\textcolor{Red}{#1}}
\newcommand{\easy}[1]{{\textit{#1}}}
\newcommand{\medm}[1]{{\textit{#1}}}
\newcommand{\hard}[1]{{\textit{#1}}}

%\usepackage{color}
%\newcommand{\tcr}[1]{\textcolor{red}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%

\large{\textbf{Reply to Reviewer \#1}}\hrule\small

Thank you for the thorough review our our manuscript titled ``Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method''. Please find below our reply to the comments.

\section{Major Issues}

\begin{itemize}

\item[\done] \medm{ Justify page 8: ``we expect temperature to be more rapidly varying than the shape but less so than the amplitude" }
\begin{itemize}
\item We have added the following on Page 9: ``This idea stems from the fact that in an adiabatic temperature model, the temperature's spatial variation follows the neutron flux shape, while the temperature amplitude follows the flux amplitude. Because of the heat capacity of the fuel, the temperature amplitude is often more slowly varying in time than neutron amplitude function itself, while still considerably faster varying than the flux shape.'' This conjecture is empirically defended in the results of Section 5, for instance, the new Figure 17.
\end{itemize}

\item[\done] \easy{ Justify page 5: ``Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself, this allowing for larger time step sizes in updating the shape". }
\begin{itemize}
\item Page 5: References [1] as justification for this statement and maybe its first empirical numerical verification. However, since K. Ott, many other researchers used and published IQS results. Our literature reviews includes several of these references, some recent, more older.
\end{itemize}

\item[\working] \medm{  Results are often just stated and not discussed, e.g. page 19 ``It also shows that traditional IQS performed better with large etol, while \iqspc was better with smaller etol". The statement simply bloats the manuscripts; discuss the reasons and implications or remove it. Another example is page 15, ``However, switching the precursors solve to be performed semi-analytically…".  In this case, the authors should take the time to discuss the implications of this finding. }

Thank you for the recommendations. The following sentences have been added:
\begin{itemize}
\item Page 16: Added ``This result implies that uniqueness of the shape and amplitude can only be maintained with a consistent treatment of precursor term in the shape equation."
\item Page 18: Removed ``This illustrates that IQS can yield a much more accurate solution, even at a significantly larger time step than the implicit flux discretization."
\item Page 20: Added ``These convergence results are consistent with the hypothesis that the performance of IQS is highly dependent on the spatial coupling of the flux."
\item Page 20: Added ``This observation implies that the time adaptive options must be chosen differently when applied to a solver for shape versus a solver for flux."
\end{itemize}

\item[\done] \medm{ For the comparison of LRA and TREAT results, the efficiency [how accurate can you be given a fixed execution time or given an accuracy how much execution time do you need] is a very important quantity; the authors discuss it for LRA and TREAT. However, the authors first present results for the errors and then runtime results. Then they try to synthesize efficiency by comparing different cases that have different runtimes and different errors, e.g. page 23 right before Table 5 ``\iqspc shows and error less than implicit discretization at delta t = 0.002, signifying …". I suggest the authors create error vs. execution time plots for different cases and based on these curves discuss efficiency. The manner efficiency is discussed is very confusing and not very useful for the reader. }
\begin{itemize}
\item Thank you for the suggestion. We have added this on Page 24: ``Fig. 17 visualizes the dependence of run-time on error. This figure shows that IQS and IQS P-C generally has a lower run-time for certain error than implicit discretization. Furthermore, note that the slopes on the right half of the \iqspc runs are shallower than the implicit discretization slope. This shows that by adding more temperature updates, in the 1-4 temperature-updates range, \iqspc suffers a lesser increase in run-time for a decrease in error. However, beyond 4 temperature updates, the error in the shape function saturates and no additional benefits can be gained by more frequent temperature updates. "
\end{itemize}

\item[\done] \easy{ In the results section, Error is sometimes used without introducing it, e.g. page 25, ``Table 9 shows the error ….". Further ``Error" is often used as y-label in plots., e.g. in the discussion of Figure 10 in which ``convergence results" are presented; the sentence is not clear in this context because it was not stated if this is iterative convergence or convergence to the right solution as the time step is reduced. In some instances error is defined but it is tedious for the reader to find the corresponding passage in the text, figures, or tables. }
\begin{itemize}
\item On  Pages 17-25: We included an explicit definition of error in each of the plots and tables. Thank you for the suggestion.
\end{itemize}

\item[\done] \hard{ LRA and TREAT only compare errors computed with the peak powers. Peak powers are an integral quantity and may favor IQS over spatial kinetics. The authors should also compare power distributions or similar distributed quantities. }
\begin{itemize}
\item Thank you for the suggestion. On Page 24-26, we added: ``Up to now, errors presented in this section were norms (i.e., space-integrated flux errors). However, for reactor physics problems, it is also noteworthy to compare the spatial dependence of error, for example, error in the power distribution. Fig. 18 shows the difference in power for each assembly of the LRA geometry for each method. The resulting spatially dependent error is in general an order of magnitude lower for IQS and \iqspc than for the implicit flux discretization. This observation agrees well with the integrated flux errors shown in Tables 6-8. "
\end{itemize}

\item[\done] \medm{  Whenever errors are computed, baseline calculations are performed. How do the authors know that these baseline calculations are accurate enough to measure the error w.r.t. them? }

Thank you for suggesting this clarification. We've made the following modifications for justification:
\begin{itemize}
\item On Page 18: we added ``It should be noted that the baseline for these error computations was performed using an adaptive method for the flux equations with a tight tolerance. The proper convergence rates seen by the methods with coarser time steps show that this baseline is accurate enough for computing the error."
\item On Page 20: we added ``The baseline was computed using the BDF2 scheme on the flux equation with a time step of $10^{-5}$. The proper error convergences show that this solution is accurate enough to compare to for error computation for most of the convergence data points. Smallest time step scheme for IQS shows a break in the linearity of the convergence."
\end{itemize}

\end{itemize}

\newpage
\section{Minor Issues}

\begin{itemize}

\item[\done] \easy{ Page 1: ``… the amplitude, and a space- and time-dependent component, the shape [1,2,3,4,5]". What about energy? }
\begin{itemize}
\item We added this sentence on Page 1: ``IQS involves factorizing the neutron flux solution into a time-only-dependent component, the amplitude, and a full phase-space component, the shape."
\end{itemize}

\item[\done] \easy{ Is $\phi^g$ the scalar flux or scalar flux moments in Eq 1a? }
\begin{itemize}
\item Thank you for catching this discrepancy. On Page 3 we've added: ``Note that $\phi^g$ for neutron transport is the angular neutron flux ($\phi^g(\vec r, \vec\Omega, t)$), while for neutron diffusion it is scalar flux ($\phi^g(\vec r, t)$)."
\end{itemize}

\item[\working] \easy{ All quantities in Eq. 1a and 1b need to be defined. }
\begin{itemize}
\item We agree. On Page 4: Created a table of the operator definitions (Table 1)
\end{itemize}

\item[\done] \easy{ Eq. 3a: $S_d^g$ is not defined. }
\begin{itemize}
\item We agree. On Page 4: Added in Table 1.
\end{itemize}

\item[\done] \easy{ For the uninitiated reader the term phase-space domain may not be familiar. The authors could introduce the term phase space. IQS may be interesting for people in other fields. }
\begin{itemize}
\item We agree. On Page 1, we added: ``In neutron transport, the neutron flux solution lives in a seven dimensional phase-space, dependent on time, space, energy, and direction. The neutron diffusion approximation reduces this phase-space by eliminating direction."
\end{itemize}

\item[\done] \easy{ Algorithm in section 2.1 on page 5. Step 2: ``linearly interpolate". It is later stated that interpolation of higher order is done for high order schemes. }
\begin{itemize}
\item Thank you for catching this. Indeed, we use high-order interpolation when a higher-order scheme is used. We removed ``linearly'' from step 2 on Page 5.
\item We've also add on Page 6: ``It should be noted that Step 2 of the process generally involves a linear interpolation; however, for greater than second order shape discretizations, a higher-order interpolation is necessary."
\end{itemize}

\item[\nofix] \easy{ For the Linf nor of the shape: why is the maximum difference over the maximum value used and not the maximum relative difference ($1 - \phi^{k+1}/\phi^k$). }
\begin{itemize}
\item We included the reference for this error estimation.
\end{itemize}

\item[\done] \medm{ Eq (11) is restated in Eq (24). Why not just make Eq. 11 more general. Also, in Eq. (24) cp(T(t)) and originally the heat equation reads d(rho * cp(T) * T). Hence, the cp(T) cannot be pulled out so Eq. (24) is incorrect. }
\begin{itemize}
\item Whether the heat capacity depends on temperature or not, the term  $\rho C_p(T)$ should always be outside of the temporal derivative. This form is given, for instance, in the Todreas \& Kazimi textbook ``Nuclear Systems I: Thermal-hydraulic fundamentals'', as Equation 4-123.
\end{itemize}

\item[\done] \easy{ Figure 2. uses the delta t / 3 step that was claimed to be only for illustration purposes in Fig. 1. }
\begin{itemize}
\item Page 10: We included the variable $N_{T}$ in Figure 2 for clarity.
\item $N_T$ is defined on Page 9: ``$N_T$ is the number of temperature updates per macro step.''
\end{itemize}

\item[\done] \medm{ Statement page 9: ``Furthermore, a very accurate representation of p(t) over the macro step is available from the PRKE solve…". I think it should be precise or else one would have to show that the equations are solved accurately. Second, what does ``very accurate" mean, ``very accurate" is comparison to what? }
\begin{itemize}
\item Thank you for noting this. We improved the phrasing on Page 11: ``Furthermore, a highly refined representation of $p(t)$ over the macro step is available from the PRKE solve (micro time scale). Using this fine scale information yields a more precise integration of amplitude than assuming an interpolation of the amplitude between macro step points."
\end{itemize}

\item[\done] \medm{ Implementation uses code specific jargon: kernels, auxkernels, Transient executioner, user-object. These need to be explained or cast in more general terms. }
\begin{itemize}
\item Thank you for the valuable suggestion. This has been added on Page 13:
\begin{itemize}
\item Kernel - Evaluation of the weak-form residual for a particular piece of physics.
\item Executioner - Establishes the type of simulation. The Transient type moves the simulation in time conducting spatial evaluations at each step.
\item Auxiliary variable and kernel - ``Optional" variables that live on the same mesh as the solution and are computed algebraically using auxiliary kernels. 
\item Postprocessor - Computes scalar values over the entire spatial mesh, usually involves integrated quantities.
\item User-object - A generic type of postprocessor that allows connectivity of relevant quantities between different MOOSE objects.
A more detailed and comprehensive description of MOOSE objects can be found in [18]
\end{itemize}
\end{itemize}

\item[\done] \easy{ Page 13 top of the page ``looping over cells to evaluate", what cells, mesh cells? }
\begin{itemize}
\item Thank you for suggesting this clarification. Page 14: ``The PRKE parameters are written as user-objects, looping over spatial mesh cells to evaluate them."
\end{itemize}

\item[\done] \easy{ LRA benchmark: the number of elements makes no sense 11x11 = 121 and when using five uniform refinement steps one would end up with 123,904 elements. }
\begin{itemize}
\item Thank you, that number is correct. On Page 22: Fixed it to 123,904 elements, number of nodes was originally correct. 
\end{itemize}

\item[\done] \easy{ LRA benchmark description does not state what equations are solved, if that is provided in the ANL benchmark book it should be made clear. Equations could also be stated directly. }
\begin{itemize}
\item  Page 22: ``The equations being solved are explicitly defined in problem 14-A1 of the ANL Benchmark Problem Book [20]"
\end{itemize}

\item[\done] \easy{ In plots where delta t is plotted on the x-axis, it should be made clear that it is the macroscopic time step. }
\begin{itemize}
\item On Page 14 and 22, we added: ``Note that any instance of $\Delta t$ in the following figures and commentary signifies the size of the macro time step."
\end{itemize}

\item[\done] \easy{ Conclusions, page 27: incomplete sentence ``IQS showed expected error convergence up". }
\begin{itemize}
\item Thank you. Page 31: ``...through fourth-order time discretization of the shape"
\end{itemize}

\item[\nofix] \easy{ Timing estimates for LRA: the author should make clear how they execute the LRA problem, i.e. number of processors and processor model. Given the long runtimes, it appears that a single processor is used. Is this to avoid problems with asynchronous behavior in parallel? }
\begin{itemize}
\item Already stated on Page 24: ``These run-times are based on total alive time of the execution where the diffusion evaluation is distributed over 24 processors."
\end{itemize}

\end{itemize}

\section{Bibliography}

\begin{itemize}

\item[\done] \easy{ MOOSE reference is out of date. Recommend newer publication }
\begin{itemize}
\item Thank you. We replaced it with [16].
\end{itemize}

\item[\done] \easy{ Rattlesnake reference [17] is out of date. Recommend citing the theory manual INL/EXT-17-42103. }
\begin{itemize}
\item Thank you. We replaced it with [17].
\end{itemize}

\item[\done] \easy{ The LRA benchmark is extensively discussed in the Rattlesnake user manual, which might be a handy reference to use: INL/EXT-15-37337. }
\begin{itemize}
\item Thank you for suggesting to include this reference. We've included it on Page 22: ``The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) [25]"
\end{itemize}

\end{itemize}

\end{document}

### 2018-10-02 doc/Papers/IQS_convergence/IQS_convergence.tex
minor changes. splitting reply in 2 because I think you need to upload 2 letters

@@
 This paper (i) discusses the different nonlinear iteration techniques for IQS and tests the rigor of their implementation, (ii) investigates high-order temporal discretization of the shape equation, and (iii) analyzes IQS coupled with adiabatic heat up and cross-section feedback.
-Prior IQS publications do not typically present results that include more than first-order time discretization of the shape equation.
+Prior IQS publications do not typically present results that include more than first-order time discretizations of the shape equation.
 Here, we test IQS with higher-order schemes (up to fourth order) and investigate the method's effectiveness with such temporal discretizations.
@@
 
-To investigate IQS's performance with step-doubling time adaptation, the adaptation is applied to implicit discretization of the flux equation (no IQS), the standard IQS method, and the \iqspc version.
+To investigate IQS's performance with step-doubling time adaptation, the adaptation is applied to implicit discretization of the flux equation (no \iqs), the standard IQS method, and the \iqspc version.
 Each of these methods will be applied to several diffusion problems; the number of time steps taken and the resulting error will be used to compare the methods.
